### 2013-12-07 paper/options/global_r_options.Rnw
added paper template

%%%-------------------------------------------------%%%
%%% Sub document global preferences for Knitr %%%
%%%-------------------------------------------------%%%

<<global_r_options, tidy=FALSE, highlight=FALSE, include=FALSE, cache=FALSE, eval=TRUE>>=  
  ## Load packages

  suppressPackageStartupMessages(library(ggplot2)) 
  suppressPackageStartupMessages(library(plyr))
  suppressPackageStartupMessages(library(rgbif))

  ## Knitr

  # messages 
  opts_chunk$set(warning=FALSE)
  # code evaluation 
  opts_chunk$set(eval=TRUE)
  # code decoration
  opts_chunk$set(tidy=TRUE, prompt=TRUE, comment='##', highlight=TRUE, size='normalsize')
  # cache options
  opts_chunk$set(cache.path='cache/')  
  # Caching on by default
  opts_chunk$set(cache=T)
  # plot options 
  #if("tikzDevice" %in% rownames(installed.packages()) == TRUE) {
	# opts_chunk$set(fig.path='usr/graphics/dynamic/', dev='tikz', fig.keep='high', fig.show='hold')  
  #} else {
	# opts_chunk$set(fig.path='usr/graphics/dynamic/', dev='pdf', fig.keep='high', fig.show='hold')  
  #}

  # tune details of graphics appearance 
  knit_hooks$set(par=function(before, options, envir){
    if (before && options$fig.show!='none') par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
    }
  )

  # escape latex if in line code chunk
  hook_inline = knit_hooks$get('inline')
  knit_hooks$set(inline = function(x) {
    if (is.character(x)) x = knitr:::escape_latex(x)
    hook_inline(x)
  })

  # pdf.options(family = "Times", width=15)
  # tikz options
  options(replace.assign=TRUE, width=60)
  options(tikzSanitizeCharacters = c('%','$','}','{','^', '_'))
  options(tikzReplacementCharacters = c('\\%','\\$','\\}','\\{','\\^{}', '\\textunderscore'))

  ## ggplot2

  # Plot theme for ggplot2 (use with "+ theme_osp()") 
  theme_osp <- function(base_size = 10, base_family = "") {
    theme_grey(base_size = base_size, base_family = base_family) %+replace%
      theme(
	axis.text = element_text(size = rel(0.8)),
	axis.ticks = element_line(colour = "black"),
	legend.key = element_rect(colour = "grey80"),
	panel.background = element_rect(fill = "white", colour = NA),
	panel.border = element_rect(fill = NA, colour = "grey50"),
	panel.grid.major = element_blank(),
	panel.grid.minor = element_blank(),  
	strip.background = element_rect(fill = "grey80", colour = "grey50"),
	strip.background = element_rect(fill = "grey80", colour = "grey50")
      )
  }

  theme_osp_distmap <- function(base_size = 10, base_family = "") {
    theme_grey(base_size = base_size, base_family = base_family) %+replace%
      theme(
	axis.text = element_blank(),
	axis.ticks = element_blank(),
	legend.title = element_text(hjust = 20),
	legend.key = element_rect(colour = "grey80"),  
	panel.background = element_rect(fill = "white", colour = NA),
	panel.border = element_rect(fill = NA, colour = "grey50"),
	panel.grid.major = element_blank(),
	panel.grid.minor = element_blank(),  
	panel.border = element_blank(), 
	legend.key = element_blank(),
	strip.background = element_rect(fill = "grey80", colour = "grey50"),
	strip.background = element_rect(fill = "grey80", colour = "grey50")
      )
  }

  ## Helpers functions

  osp_load_or_install <- function(library=NULL){
    # todo: Implement a check if package is installed otherwise install it
  }

  osp_distmap <- function(species=NULL){
    library(ggplot2) 
    library(plyr) 
    library(rgbif)  
    out <- llply(species, function(x) occurrencelist(x, coordinatestatus = T, maxresults = 100)) 
    custom <- function(){list(theme_osp_distmap(), labs(colour = "Species"))}
    gbifmap(out, customize=custom()) 
  }

@

### 2013-12-07 paper/paper.Rnw
added paper template

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.3 (9/9/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside]{article}

\usepackage{lipsum} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text
\bibliographystyle{plain}

%%%-------------------------------------------------%%%
%%% Preferences for Knitr %%%
%%%-------------------------------------------------%%%

<<chunk_global_r_options, child='options/global_r_options.Rnw', eval=TRUE>>=
@

%----------------------------------------------------------------------------------------
%  TITLE SECTION
%----------------------------------------------------------------------------------------

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Article Title}} % Article title

\author{
\large
\textsc{John Smith}\thanks{A thank you or further information}\\[2mm] % Your name
\normalsize University of California \\ % Your institution
\normalsize \href{mailto:john@smith.com}{john@smith.com} % Your email address
\vspace{-5mm}
}
\date{}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Insert title

\thispagestyle{fancy} % All pages have headers and footers

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

%%%-------------------------------------------------%%%
%%% Include abstract %%%
%%%-------------------------------------------------%%%

<<subdoc_content_abstract, child='subdocuments/abstract.Rnw', eval=T>>=
@

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

%%%-------------------------------------------------%%%
%%% Include introduction %%%
%%%-------------------------------------------------%%%

<<subdoc_content_introduction, child='subdocuments/introduction.Rnw', eval=T>>=
@

%%%-------------------------------------------------%%%
%%% Include data and methods %%%
%%%-------------------------------------------------%%%

<<subdoc_data_and_methods, child='subdocuments/data_and_methods.Rnw', eval=T>>=
@

%%%-------------------------------------------------%%%
%%% Include results %%%
%%%-------------------------------------------------%%%

<<subdoc_content_results, child='subdocuments/results.Rnw', eval=T>>=
@

%%%-------------------------------------------------%%%
%%% Include discussion %%%
%%%-------------------------------------------------%%%

<<subdoc_content_discussion, child='subdocuments/discussion.Rnw', eval=T>>=
@

%%%-------------------------------------------------%%%
%%% Include acknowledgements %%%
%%%-------------------------------------------------%%%

<<subdoc_content_acknowledgement, child='subdocuments/acknowledgements.Rnw', eval=T>>=
@

%%%-------------------------------------------------%%%
%%% Include the bibliography %%%
%%%-------------------------------------------------%%%

\end{multicols}

\bibliography{bibliography/bib}

%%%-------------------------------------------------%%%
%%% Include the appendix %%%
%%%-------------------------------------------------%%%

<<subdoc_content_appendix, child='subdocuments/appendix.Rnw', eval=T>>=
@

\end{document}

### 2013-12-13 paper/subdocuments/abstract.Rnw
moved everything from word-draft to github

--- 2013-12-07 paper/paper.Rnw
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Journal Article
-% LaTeX Template
-% Version 1.3 (9/9/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Frits Wenneker (http://www.howtotex.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[twoside]{article}
-
-\usepackage{lipsum} % Package to generate dummy text throughout this template
-
-\usepackage[sc]{mathpazo} % Use the Palatino font
-\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
-\linespread{1.05} % Line spacing - Palatino needs more space between lines
-\usepackage{microtype} % Slightly tweak font spacing for aesthetics
-
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
-\usepackage{multicol} % Used for the two-column layout of the document
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
-\usepackage{booktabs} % Horizontal rules in tables
-\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
-\usepackage{hyperref} % For hyperlinks in the PDF
-
-\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
-\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
-
-\usepackage{abstract} % Allows abstract customization
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text
-
-\usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
-
-\usepackage{fancyhdr} % Headers and footers
-\pagestyle{fancy} % All pages have headers and footers
-\fancyhead{} % Blank out the default header
-\fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
-\fancyfoot[RO,LE]{\thepage} % Custom footer text
-\bibliographystyle{plain}
-
 %%%-------------------------------------------------%%%
-%%% Preferences for Knitr %%%
+%%% Sub document abstract %%%
 %%%-------------------------------------------------%%%
 
-<<chunk_global_r_options, child='options/global_r_options.Rnw', eval=TRUE>>=
-@
+\begin{abstract}
 
-%----------------------------------------------------------------------------------------
-%  TITLE SECTION
-%----------------------------------------------------------------------------------------
+Maximal 200 Wörter
 
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Article Title}} % Article title
-
-\author{
-\large
-\textsc{John Smith}\thanks{A thank you or further information}\\[2mm] % Your name
-\normalsize University of California \\ % Your institution
-\normalsize \href{mailto:john@smith.com}{john@smith.com} % Your email address
-\vspace{-5mm}
-}
-\date{}
-
-%----------------------------------------------------------------------------------------
-
-\begin{document}
-
-\maketitle % Insert title
-
-\thispagestyle{fancy} % All pages have headers and footers
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-%%%-------------------------------------------------%%%
-%%% Include abstract %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_abstract, child='subdocuments/abstract.Rnw', eval=T>>=
-@
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-\begin{multicols}{2} % Two-column layout throughout the main article text
-
-%%%-------------------------------------------------%%%
-%%% Include introduction %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_introduction, child='subdocuments/introduction.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include data and methods %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_data_and_methods, child='subdocuments/data_and_methods.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include results %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_results, child='subdocuments/results.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include discussion %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_discussion, child='subdocuments/discussion.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include acknowledgements %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_acknowledgement, child='subdocuments/acknowledgements.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include the bibliography %%%
-%%%-------------------------------------------------%%%
-
-\end{multicols}
-
-\bibliography{bibliography/bib}
-
-%%%-------------------------------------------------%%%
-%%% Include the appendix %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_appendix, child='subdocuments/appendix.Rnw', eval=T>>=
-@
-
-\end{document}
+\end{abstract}

### 2013-12-13 paper/subdocuments/change.Rnw
moved everything from word-draft to github

%%%-------------------------------------------------%%%
%%% Sub document for discussion %%%
%%%-------------------------------------------------%%%

\section{How did inequality change over time}

\subsection{Previous evidence (Literatursammlung haben wir)}

Information about inequality is sparse in Switzerland. For example Gini coefficients have been reported for certain periods, e.g. \cite{Abele1977}, \cite{buhmann1988}, \cite{ernst1983wohlstandsverteilung}, \cite{foerster2005}, \cite{suter2002}, \cite{isengard2009umverteilung}, \cite{atkinson1995}. While most studies report Gini coefficients for isolated or very view points of time, \cite{Abele1977} provide the longest and most consistent time series of inequality in Switzerland which is however outdated by years. We use the same source, that is data from the Federal Tax Administration (FTA) but expand the time series to 2010.

\textcolor{red}{Oli: du kannst französisch oder? Könntest du Abele (1977) mal überfliegen, ob das so stimmt?}

\subsection{Data quality (previous data)}

Evtl. überflüssig

\subsection{Kuznets, Great-U-Turn, Welfare State}

Literature on inequality and its development over time is very extensive for most developed countries. One branch of theory focusses on \cite{kuznets1955economic} ideas about sectoral change, that is the transition from an agrarian (primary sector) to an industrial country (secondary sector) and with the extensions of \cite{nielsen_kuznets_1997} the emergence of a tertiary sector within which incomes are varying more. What can be seen from international comparative research is that patterns of changing inequality are similar for many developed countries which strenghens arguments like the one Kuznets made but parallel other ideas exist why patterns are similar to each other. One concurrent argument is that the developments and downsizings of welfare states occured in the same periods of history for many countries (e.g. after the oil crisis). Therefore we want to contribute the information from Switzerland to the "big picture".

\subsection{Switzerland compared to the rest of the world}

\textcolor{red}{Hier könnten wir gleich mal eine Gesamtschweizer  Grafik droppen, Vergleich zu Europa/USA. Sowas wie du in den SGS Slides hattest.}

\textcolor{red}{Die ganze Sektion gehört aber evtl. zu "Results"}

### 2013-12-13 paper/subdocuments/data_and_methods.Rnw
moved everything from word-draft to github

%%%-------------------------------------------------%%%
%%% Sub document for data and methods %%%
%%%-------------------------------------------------%%%

\section{Hypotheses, Methods and Data}

\subsection{Hypotheses}
Based on the theories we test the following hypotheses:

\begin{itemize}
\item H1: Develpment of inequality is driven by sectoral change
\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
\end{itemize}

\subsection{Data and Variables}

We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010. While the data results in a long and consistent time series to illustrate swiss inequality development, there are a few pitfalls we want to adress which might be of interest for other research on this topic (be it in Switzerland or other countries).

\subsubsection{Left censored data}

The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household. The taxable population however is not identical to the population which should be used to calculate measures of inequality. Precisely, the data do not contain tax units with very little incomes so calculations based on these data treat the lowest percentiles equally to tax units with zero income.Figure X shows the threshold to be hit to enter the statistic.

\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}

So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data. Furthermore, also the changes over time might not be interpreted savely as over time the number of tax unit within this "hidden range" might vary or might even have a certain trend. We will adress this issue in detail in the methods chapter.

\subsubsection{Different measures, different populations}

The FTA data makes two kinds of distinctions. First, data was collected for so called "normal cases" and "special cases", i.e. a "normal case" is a taxable (for the complete tax period) person or household domiciled in a swiss canton without income from outside of Switzerland. A "special case" therefore is a diffuse reference category that contains tax units that are taxed at source, were not taxable for the complete tax period or generated additional income in another country. Second, the FTA reports two measures, that is taxable income and absolute income \textcolor{red}{(meine vorläufige Übersetzung von Reineinkommen)}. Absolute income is the sum of all incomes (earnings, interest income, rental incomes) minus expenses (e.g. from self-employment or credit cost). The taxable income is calculated as the difference of absolute income and deductions (e.g. children, insurance rates). The longest consistent time series exists for the taxable income of normal cases. So all statements we make with our data only apply to this subpopulation.

\subsubsection{Changes in taxation and measurement}
The swiss tax system is highly federal. That means, communities raise taxes which then go to the communities, the canton and the state. If we want to calculate overall swiss measures, we need to take into account, that cantons vary (between cantons and over time) with regard to the tax deductions that are possible and also the mechanism how taxes are collected. The latter adresses a comprehensive reorganization of the swiss tax system where between 1995 and 2003 cantons changed from taxing the past two years of income (postnumerando system) to taxing the present single year (praenumerando system). For details see \textcolor{red}{Martinez (xxxx) or some other author (xxxx)}. Aggregate measures of inequality therefore have to be estimated for the periods 1995 to 2003 which we adress shortly in the methods chapter.

\subsection{Methodology Used}
There are two steps of data analyses which need to be described to the reader. First, the estimation of the bias we introduce by estimating measures of inequality when tax units with too little income are not observed. Second, the steps undertaken to estimate aggregate swiss measures by imputing taxable income for those cantons and periods where the change of the tax system produced a gap (1995 to 2003, depending on the canton).

\subsubsection{Imputing the gap}
The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx). That is estimating the missing taxable income statistics via OLS using information from time trends and cantons. Our imputation model therefore includes canton inequality measures and periods dummies to explain aggregate swiss inequality.

\subsubsection{Estimating the bias}
For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation. However starting 1995, the FTA provides exactly this information for each canton. This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010. Consequently we can obtain information whether the bias is stable over time (which maes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton. \textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie "je höher der Steuerfreibetrag, umso stärker der Bias". Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias. Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}

### 2013-12-13 paper/subdocuments/introduction.Rnw
moved everything from word-draft to github

%%%-------------------------------------------------%%%
%%% Sub document introduction %%%
%%%-------------------------------------------------%%%

\section{Introduction}

The purpose of the following paper is threefold. First of all, information about economic inequality in Switzerland is sparse. E.g. the well known Worldbank data do not contain any Gini-coefficient for Switzerland which is a rare exception within Europe. We try to fill this gap with official data from the Swiss Federal Tax Administration (FTA). Second, we assess the suitability of the FTA data to report inequality and its changes over time. An important aspect here is that the data do not contain all income earning persons in Switzerland but are left censored as federal taxes are payable only above a certain threshold. We report magnitude and variance of this bias. Third, we integrate the development of swiss inequality into the international picture. We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).

### 2013-12-13 paper/subdocuments/results.Rnw
moved everything from word-draft to github

%%%-------------------------------------------------%%%
%%% Sub document results %%%
%%%-------------------------------------------------%%%

\section{Results}

\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
Kantonsweise Grafiken
Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}

### 2013-12-13 paper/paper.Rnw
changed to utf8 + small edits

@@
 \usepackage[sc]{mathpazo} % Use the Palatino font
-\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
+%\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
+\usepackage[utf8]{inputenc}
 \linespread{1.05} % Line spacing - Palatino needs more space between lines
@@
 \usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
+
+\usepackage{color} % used to mark parts that need to be edited
 
@@
 \large
-\textsc{John Smith}\thanks{A thank you or further information}\\[2mm] % Your name
-\normalsize University of California \\ % Your institution
-\normalsize \href{mailto:john@smith.com}{john@smith.com} % Your email address
+\textsc{Oliver Hümbelin}\\[2mm] % Your name
+\normalsize Bern University of Applied Sciences \\ % Your institution
+\normalsize \href{mailto:oliver.huembelin@bfh.ch}{oliver.huembelin@bfh.ch} % Your email address
+\vspace{5mm}\\
+\large
+\textsc{Rudolf Farys}\\[2mm] % Your name
+\normalsize University of Bern \\ % Your institution
+\normalsize \href{mailto:rudolf.farys@soz.unibe.ch}{rudolf.farys@soz.unibe.ch} % Your email address
 \vspace{-5mm}
@@
 <<subdoc_content_introduction, child='subdocuments/introduction.Rnw', eval=T>>=
+@
+
+%%%-------------------------------------------------%%%
+%%% How did inequality change over time %%%
+%%%-------------------------------------------------%%%
+
+<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
 @

### 2014-01-07 paper/options/paper.tex
Übertrag aus Skizze

--- 2013-12-07 paper/paper.Rnw
@@
 %%%-------------------------------------------------%%%
-
-<<chunk_global_r_options, child='options/global_r_options.Rnw', eval=TRUE>>=
-@
-
-%----------------------------------------------------------------------------------------
-%  TITLE SECTION
-%----------------------------------------------------------------------------------------
-
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Article Title}} % Article title
-
-\author{
-\large
-\textsc{John Smith}\thanks{A thank you or further information}\\[2mm] % Your name
-\normalsize University of California \\ % Your institution
-\normalsize \href{mailto:john@smith.com}{john@smith.com} % Your email address
-\vspace{-5mm}
-}
-\date{}
-
-%----------------------------------------------------------------------------------------
-
-\begin{document}
-
-\maketitle % Insert title
-
-\thispagestyle{fancy} % All pages have headers and footers
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-%%%-------------------------------------------------%%%
-%%% Include abstract %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_abstract, child='subdocuments/abstract.Rnw', eval=T>>=
-@
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-\begin{multicols}{2} % Two-column layout throughout the main article text
-
-%%%-------------------------------------------------%%%
-%%% Include introduction %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_introduction, child='subdocuments/introduction.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include data and methods %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_data_and_methods, child='subdocuments/data_and_methods.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include results %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_results, child='subdocuments/results.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include discussion %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_discussion, child='subdocuments/discussion.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include acknowledgements %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_acknowledgement, child='subdocuments/acknowledgements.Rnw', eval=T>>=
-@
-
-%%%-------------------------------------------------%%%
-%%% Include the bibliography %%%
-%%%-------------------------------------------------%%%
-
-\end{multicols}
-
-\bibliography{bibliography/bib}
-
-%%%-------------------------------------------------%%%
-%%% Include the appendix %%%
-%%%-------------------------------------------------%%%
-
-<<subdoc_content_appendix, child='subdocuments/appendix.Rnw', eval=T>>=
-@
-
-\end{document}

### 2014-01-07 paper/paper.Rnw
Übertrag aus Skizze

@@
 
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Article Title}} % Article title
+\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing inequality with tax data - Switzerland from 1945 to 2010}} % Article title
 
@@
 
-\maketitle % Insert title
+\maketitle
 
@@
 
-<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
-@
+%<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
+%@
 

### 2014-01-07 paper/paper.tex
Übertrag aus Skizze

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.3 (9/9/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\usepackage{lipsum} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
%\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage[utf8]{inputenc}
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

\usepackage{color} % used to mark parts that need to be edited

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text
\bibliographystyle{plain}

%%%-------------------------------------------------%%%
%%% Preferences for Knitr %%%
%%%-------------------------------------------------%%%

%%%-------------------------------------------------%%%
%%% Sub document global preferences for Knitr %%%
%%%-------------------------------------------------%%%

%----------------------------------------------------------------------------------------
%  TITLE SECTION
%----------------------------------------------------------------------------------------

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing inequality with tax data - Switzerland from 1945 to 2010}} % Article title

\author{
\large
\textsc{Oliver Hümbelin}\\[2mm] % Your name
\normalsize Bern University of Applied Sciences \\ % Your institution
\normalsize \href{mailto:oliver.huembelin@bfh.ch}{oliver.huembelin@bfh.ch} % Your email address
\vspace{5mm}\\
\large
\textsc{Rudolf Farys}\\[2mm] % Your name
\normalsize University of Bern \\ % Your institution
\normalsize \href{mailto:rudolf.farys@soz.unibe.ch}{rudolf.farys@soz.unibe.ch} % Your email address
\vspace{-5mm}
}
\date{}

%----------------------------------------------------------------------------------------
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\begin{document}

\maketitle

\thispagestyle{fancy} % All pages have headers and footers

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

%%%-------------------------------------------------%%%
%%% Include abstract %%%
%%%-------------------------------------------------%%%

%%%-------------------------------------------------%%%
%%% Sub document abstract %%%
%%%-------------------------------------------------%%%

\begin{abstract}

There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013). In Switzerland, however, the development is unclear, as there is only little systematic evidence that would allow for long-term comparison. Nevertheless income inequality has been a prominent theme in the public discussion in recent years, e.g. the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''. We adress the gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data. We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties. In the end we integrate the case of Switzerland into the international picture of inequality development showing parallels and deviations.

\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

%%%-------------------------------------------------%%%
%%% Include introduction %%%
%%%-------------------------------------------------%%%

%%%-------------------------------------------------%%%
%%% Sub document introduction %%%
%%%-------------------------------------------------%%%

\section{Introduction}

% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern.

% 1.Abschnitt: Bedeutung von Ungleichheit von Soziologischer Sicht

\emph{Bedeutung von ökonomischer Ungleicheit} \\

% 2.Abschnitt Trends in inequality research
\emph{Studien zur allgemeinen Entwicklung der Ungleichheit nach unterschiedliche Paradigmen}(Armut, Reiche, Mittelschicht)  [Concerns about the hollowing of the middle class is a argument why the tax studies should be exploited to gain knowledge about the whole distribution of economic resources. Until these days tax data is only used to assess the change in top income shares.] \\
\begin{itemize}
  \item Development of inequality with LIS Data (Gornick and Jäntti 2013)
  \item ranking of inequality across high-income countries (Atkinson, Rainwater and Smeeding (1995)
   \item inequality and low income/relative poverty (Ferreira and Ravallion, 2009; Nolan and Marx, 2009; Smeeding, O’Higgins and Rainwatter, 1990; Rainwater and Smeeding 2003; Förster and Vlemickx, 2004)
  \item inequality and top income (Atkinson and Piketty, 2007; Leigh, 2009)
  \item inequality and the middle class (Estache and Leipziger, 2009; Littrell et al., 2001; Birdsall, 2010; Ravallion, 2010; Frank,2007)
\end{itemize}

% 3.Abschnitt: Studien mit Steuerdaten

\emph{Studien mit Steuerdaten}
\begin{itemize}
\item Old studies with tax data (Vilfredo Pareto, 1895, 1896; Simon Kuznets, 1953)
\item More recently revival of tax-studys with focus on top income (Piketty and Saez, 2003; Atkinson and Piketty, 2007,2010).
\end{itemize}

% 4.Abschnitt: What do we know about Switzerland?

\emph{Official Data Collection} \\

\begin{itemize}
\item  LIS (1982-2004, Survey Data from Swiss Federal Statistical Office three different surveys (recent years from “income and Consumption Survey, main result; decreasing inequality)
\item  OECD (2000 und 2004, Income and consumption survey, 2008, EU-SILC. The change of survey in 2008 is consdered as a strict break. Comparison over time not possible (OECD 2012:315)
\item  Worldbank: Nothing
\item  Federal Office of statistics (1998-2011, HABE, survey data, very detailed measurement, main result; inequality in post transfer income is constant)
\item  World income inequality Database (WIID) is more or less the same data as it is published in the LIS data
\end{itemize}
\emph{Fazit}:LIS longest time series, comparability is questionable, more recent years time series from Federal Office of statistics. Both studies use survey data, which, as we will discuss later, have certain drawbacks.\\

\emph{Studies on Switzerland} \\
Vgl. Forschungsantrag (kürzen) plus neuere Studien, Verteilungsbericht 2011 und 2012 (SGB), Avenir Suisse (2013) und Gorgas and Schaltenegger (2011) und Foellmi and Martinez (2013). \\

% Should we discus the non-scientic literature?

% What is the purpose of our paper?
The purpose of the following paper is threefold. First of all, information about economic inequality in Switzerland is sparse. We  fill this gap with official data from the Swiss Federal Tax Administration (FTA). Second, we assess the suitability of the FTA data to report inequality and its changes over time. Third, we integrate the development of swiss inequality into the international picture. We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).

%%%-------------------------------------------------%%%
%%% How did inequality change over time %%%
%%%-------------------------------------------------%%%

%<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
%@

%%%-------------------------------------------------%%%
%%% Include data and methods %%%
%%%-------------------------------------------------%%%

%%%-------------------------------------------------%%%
%%% Sub document for data and methods %%%
%%%-------------------------------------------------%%%

\section{Measurement concepts, Data and Methods}

To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure. Based on a review on the literature about the measurment concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namly survey data. Afterwards we discribe the tax data published by the Swiss Federal Tax Administration highligthing the important aspects, which have to be considered working wiht FTA-Data. In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for switzerland.

\subsection{Standards for measuring economic resources and inequality}

\emph{Concepts on measuring economic resources}  \\
Economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\

\emph{Defining income}

\begin{itemize}
\item Market income refers to revenue from employment.
\item primary income or pre-transfer income refers to  market income plus revenue from property (capital income)
\item Brutto income refers to  primary income plus social transfer
\item Disposable income or posttransfer income refers to  brutto income minus transfer paid 
\item adjusted disposable income refers to  disposable income corrected for the numbers of household members 
\end{itemize}

The appropriateness of the income concept is bound to the research interest. Studies focusing on labor market outcome look at market income. Studies focusing on human wellbeing follow the approach as propagated by the oecd (2013). This concept is largely implemented by the Federal Office of statistics. Following this concept one should look at income after taxes and transfers inclusive the adjustment for household members, because this is the measure which frames the consumption possibilities. Lastly, studies focusing on the effect of institutions (social welfare, taxation) compare the distribution of pre- and post-income measures. \\

\emph{Population Coverage} \\
All Residents? Working Residents? Single Person vs household/family. Households without income \\
see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\

\emph{Measuring inequality and concentration} \\
To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension. Vgl. Allison (1978), Engelhardt (2000), Cowell (2000) oder Hao/Naiman (2010)
%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\

neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences (Handcock/Morris 1999) \\

Gini is mainly used for international comparison. However, other measures are often reported along (gini is more sensitive to changes in the middle of the distribution than to changes in the tails (OECD, 2008:37). Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) (Alderson and Doran 2013). \\

Alvaredo (2010) shows formally, how top incomes shares are related to the gini-coefficient. Following his argument, it is crucial to consider top incomes. Leigh (2007) also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the studie). 
%[He also discusses tax data studies in general-> have a closer look] \\
%[We definitely have to argue, why it isn’t sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]

\subsection{Comparison of tax data and other data sources - advantages and shortcomings}

%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
% Mögliche Vergleichsdimensionen
%- Suitability of cross country comparison
%- Assessing inequality accurately
%- Implementation of common concepts of economic ressources
%- contingency for intertemporal comparision

%Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA

%Warnings about the use of tax data
%The use of tax data is often regarded by economists with considerable disbelief. These doubts are well justified for at least two reasons. The first is that tax data are collected as part of an administrative process, which is not tailored to the scientists' needs, so that the definition of income, income unit, etc., are not necessarily those that we would have chosen. This causes particular difficulties for comparisons across countries, but also for time-series analysis where there have been substantial changes in the tax system, such as the moves to and from the joint taxation of couples. Secondly, it is obvious that those paying tax have a financial incentive to present their affairs in a way that reduces tax liabilities. There is tax avoidance and tax evasion. The rich, in particular, have a strong incentive to understate their taxable incomes. Those with wealth take steps to ensure that the return comes in the form of asset appreciation, typically taxed at lower rates or not at all. Those with high salaries seek to ensure that part of their remuneration comes in forms, such as fringe benefits or stock-options which receive favorable tax treatment. Both groups may make use of tax havens that allow income to be moved beyond the reach of the national tax net. These shortcomings limit what can be said from tax data, but this does not mean that the data are worthless. Like all economic data, they measure with error the 'true' variable in which we are interested.
%The data series presented here are fairly homogenous across countries, annual, long-run, and broken down by income source for several cases. Users should be aware also about their limitations. Firstly, the series measure only top income shares and hence are silent on how inequality evolves elsewhere in the distribution [why?]. Secondly, the series are largely concerned with gross incomes before tax. Thirdly, the definition of income and the unit of observation (the individual vs. the family) vary across countries making comparability of levels across countries more difficult. Even within a country, there are breaks in comparability that arise because of changes in tax legislation affecting the definition of income, although most studies try to correct for such changes to create homogenous series. Finally and perhaps most important, the series might be biased because of tax avoidance and tax evasion. For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).

%Copied from:
%Alvaredo, Facundo, Anthony B. Atkinson, Thomas Piketty and Emmanuel Saez, The World Top Incomes Database, http://topincomes.g-mond.parisschoolofeconomics.eu/ , 18/12/2013

\textbf{Summary of tax data drawbacks}
\begin{itemize}
\item Misreporting of incomes (high earners have an interest in getting tax beneficial “income” see above)
\item with aggregated data the adjustment for household members is not possible (no equivalization possible). This is a point, which is hardly mentioned in the “top-income” literature, albeit it is a crucial issue in “theorized” concepts of income. To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household. This might influence the overall inequality, taking into account the change from traditional household and family structures over the last century. This is an inequality related issue, where relevant studies are missing. What are the assumptions?
\item the tax data income definition summarize labor income, capital income and taxable transfer payments (no means-tested benefits as welfare aid). Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
\item Some tax units are not included in the tax tables (see Foellmi and Martinez 2013:11f).(a) only taxed cases are included (partly a problem, for certain periodes we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung). Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only. In some border-cantons the share of this group is very high (20\%) 
%Can we ignore this)
(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registred. This persons get an imputed income (older tax tax return and information given by employers. Not registered non-	fillers are not in the records.
- (e) tax evasion. Feld and Frey (2007) report about tax evasion in Switzerland, it should be somewhat above 20 percent on average. With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequaly measures over time.
\end{itemize}

\textbf{Problems with household income surveys}
\begin{itemize}
\item Sample data (bias)
\item comparability between countries and over time (depends on income definition)
\item short time series
\end{itemize}

Atkinson et al. (2009) estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by Alvaredo (2010). \\

See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
%	•	What about the special cases? Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures. Exclusion leads to an underestimation of inequality
%	•	Capital gains? Should not be included 
The discussion about problems with reporting income is fairly exhaustive. What about wealth?

\subsection{Tax data published by the Swiss Federal Tax Administration}

%Beschreiben was die Daten beinhalten und wo es Probleme gibt
\begin{itemize}
\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
	\item	Income after deductions-> taxable income (employment income, business income and capital income). Is income definition stable over time? Yes it should. Realized capital gains are excluded from the definition. It includes income from  employment, self-employment, capital income and taxable transfer payments. Plus Eigenmietwert, 
	\item	Reported on national and cantonal level
\end{itemize}

\subsection{Ways to tackle FTA-tax data specific problems}

\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed? Del et al. (2007) impute for non-fillers the 20 percentage of the annual average income. This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality. Furthermore Del et al. calculate the proportion of non-fillers by estimating the total of tax units out of the population records. \\

\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation. This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.  This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f). \\
%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.

\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\

\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)

\textbf{deductions} Del et al. (2007:477):” we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.”
Gorgas and Schaltenegger (2011:5): “”..., information on […] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time. Können wir das auch? Zumindest für gewisse Zeiträume? Das wäre noch gut. \\

Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income. Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income. On the other hand deductions can affect the distribution. There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a “perverse redistribution” effect by redistributing income from the lower middle class to the upper middle class (vgl. Peters 2011 and Interpellation Barbara Gysel (2009).

% Income share specific problems
% Total income denominator Exogenous Approach -> net income reported in the national accounts. Endogenous Approach -> Dell et al. technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income). Honestly , is this appropriate? (Everyone did it)
% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
% Endogenous approach -> reported in the tax tables

%%%-------------------------------------------------%%%
%%% Abschnitte aus ersten Version des Papers %%%
%%%-------------------------------------------------%%%

%\subsection{Hypotheses}
%Based on the theories we test the following hypotheses:

%\begin{itemize}
%\item H1: Develpment of inequality is driven by sectoral change
%\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
%\end{itemize}

%\subsection{Data and Variables}

%We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010. While the data results in a long and consistent time series to illustrate swiss inequality development, there are a few pitfalls we want to adress which might be of interest for other research on this topic (be it in Switzerland or other countries).

%\subsubsection{Left censored data}

%The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household. The taxable population however is not identical to the population which should be used to calculate measures of inequality. Precisely, the data do not contain tax units with very little incomes so calculations based on these data treat the lowest percentiles equally to tax units with zero income.Figure X shows the threshold to be hit to enter the statistic.

%\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}

%So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data. Furthermore, also the changes over time might not be interpreted savely as over time the number of tax unit within this "hidden range" might vary or might even have a certain trend. We will adress this issue in detail in the methods chapter.

%\subsubsection{Different measures, different populations}

%The FTA data makes two kinds of distinctions. First, data was collected for so called "normal cases" and "special cases", i.e. a "normal case" is a taxable (for the complete tax period) person or household domiciled in a swiss canton without income from outside of Switzerland. A "special case" therefore is a diffuse reference category that contains tax units that are taxed at source, were not taxable for the complete tax period or generated additional income in another country. Second, the FTA reports two measures, that is taxable income and absolute income \textcolor{red}{(meine vorläufige Übersetzung von Reineinkommen)}. Absolute income is the sum of all incomes (earnings, interest income, rental incomes) minus expenses (e.g. from self-employment or credit cost). The taxable income is calculated as the difference of absolute income and deductions (e.g. children, insurance rates). The longest consistent time series exists for the taxable income of normal cases. So all statements we make with our data only apply to this subpopulation.

%\subsubsection{Changes in taxation and measurement}
%The swiss tax system is highly federal. That means, communities raise taxes which then go to the communities, the canton and the state. If we want to calculate overall swiss measures, we need to take into account, that cantons vary (between cantons and over time) with regard to the tax deductions that are possible and also the mechanism how taxes are collected. The latter adresses a comprehensive reorganization of the swiss tax system where between 1995 and 2003 cantons changed from taxing the past two years of income (postnumerando system) to taxing the present single year (praenumerando system). For details see \textcolor{red}{Martinez (xxxx) or some other author (xxxx)}. Aggregate measures of inequality therefore have to be estimated for the periods 1995 to 2003 which we adress shortly in the methods chapter.

%\subsection{Methodology Used}
%There are two steps of data analyses which need to be described to the reader. First, the estimation of the bias we introduce by estimating measures of inequality when tax units with too little income are not observed. Second, the steps undertaken to estimate aggregate swiss measures by imputing taxable income for those cantons and periods where the change of the tax system produced a gap (1995 to 2003, depending on the canton).

%\subsubsection{Imputing the gap}
%The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx). That is estimating the missing taxable income statistics via OLS using information from time trends and cantons. Our imputation model therefore includes canton inequality measures and periods dummies to explain aggregate swiss inequality.

%\subsubsection{Estimating the bias}
%For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation. However starting 1995, the FTA provides exactly this information for each canton. This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010. Consequently we can obtain information whether the bias is stable over time (which maes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton. \textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie "je höher der Steuerfreibetrag, umso stärker der Bias". Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias. Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}

%%%-------------------------------------------------%%%
%%% Include results %%%
%%%-------------------------------------------------%%%

%%%-------------------------------------------------%%%
%%% Sub document results %%%
%%%-------------------------------------------------%%%

\section{Results}

\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
Kantonsweise Grafiken
Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}

%Switzerland (income and wealth)
%- graph with gini over time> identify periods of change
%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
%Cantonal level (income and wealth)
%- different development on regional level 
%Switzerland in international comparison
%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality

%- Comparison with other countries (similar data)

% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen. Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel. Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.

%%%-------------------------------------------------%%%
%%% Include discussion %%%
%%%-------------------------------------------------%%%

%%%-------------------------------------------------%%%
%%% Sub document for discussion %%%
%%%-------------------------------------------------%%%

\section{Discussion}

% Hier könnte allenfalls der Bogen zur Theorie gespannt werden

%Methodological Main findings 
%(1) Tax data is useful because it’s possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings. It’s superior to survey data, because the latter is affected by non-responded bias. Tax data includes data about the whole population … [was finden wir wegen den Nullern raus] . or at least it should (tax evasion).(3) The FTA-Tax data has different shortcomings, which cannot be handled like a measure of income and wealth which (a) cannot be corrected for household size and (b) is neither a pre-transfer nor a post-transfer income measure. Therefor it’s comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income. This FTA-Tax data problem can be handled with non-aggregated tax-data .

%Main findings for Switzerland (1) Our data suggests that income inequality overall slightly increased in Switzerland and we might distinguish three episodes: 1. Substantial increase in times of economic growth 1950 to 1974 ending with the oil crisis 2. Ups and downs from 1970 to 2000 3. Relatively steep increase in inequality in the early 2000s
%(2) Despite the recent increase, overall income inequality in Switzerland is not very high in international comparison. With respect to inequality in wealth, however, Switzerland takes a leading position in the world.
%(3) While the overall Swiss inequality remained pretty stable over decades, inequality between cantons underwent a bizarre development.

%%%-------------------------------------------------%%%
%%% Include acknowledgements %%%
%%%-------------------------------------------------%%%

%%%-------------------------------------------------%%%
%%% Sub document for acknowledgement %%%
%%%-------------------------------------------------%%%

\section{Acknowledgements}

We thank Ben Jann, Robert Fluder and Tobias Fritschi for helpful comments on the article. We also like to thank Stefan Ilic for the prepartion of the data set.

%%%-------------------------------------------------%%%
%%% Include the bibliography %%%
%%%-------------------------------------------------%%%

\end{multicols}

\bibliography{bibliography/bib}

%%%-------------------------------------------------%%%
%%% Include the appendix %%%
%%%-------------------------------------------------%%%

%%%-------------------------------------------------%%%
%%% Sub document for appendix %%%
%%%-------------------------------------------------%%%

\section{Appendix}

\end{document}

### 2014-01-07 paper/subdocuments/abstract.Rnw
Übertrag aus Skizze

@@
 
-Maximal 200 Wörter
+There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
+In Switzerland, however, the development is unclear, as there is only little systematic evidence that would allow for long-term comparison.
+Nevertheless income inequality has been a prominent theme in the public discussion in recent years, e.g.
+the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''.
+We adress the gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data.
+We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties.
+In the end we integrate the case of Switzerland into the international picture of inequality development showing parallels and deviations.
 

### 2014-01-07 paper/subdocuments/abstract.tex
Übertrag aus Skizze

--- 2014-01-07 paper/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Journal Article
-% LaTeX Template
-% Version 1.3 (9/9/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Frits Wenneker (http://www.howtotex.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[twoside]{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
-
-\usepackage{lipsum} % Package to generate dummy text throughout this template
-
-\usepackage[sc]{mathpazo} % Use the Palatino font
-%\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
-\usepackage[utf8]{inputenc}
-\linespread{1.05} % Line spacing - Palatino needs more space between lines
-\usepackage{microtype} % Slightly tweak font spacing for aesthetics
-
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
-\usepackage{multicol} % Used for the two-column layout of the document
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
-\usepackage{booktabs} % Horizontal rules in tables
-\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
-\usepackage{hyperref} % For hyperlinks in the PDF
-
-\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
-\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
-
-\usepackage{color} % used to mark parts that need to be edited
-
-\usepackage{abstract} % Allows abstract customization
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text
-
-\usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
-
-\usepackage{fancyhdr} % Headers and footers
-\pagestyle{fancy} % All pages have headers and footers
-\fancyhead{} % Blank out the default header
-\fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
-\fancyfoot[RO,LE]{\thepage} % Custom footer text
-\bibliographystyle{plain}
-
-%%%-------------------------------------------------%%%
-%%% Preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document global preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%----------------------------------------------------------------------------------------
-%  TITLE SECTION
-%----------------------------------------------------------------------------------------
-
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing inequality with tax data - Switzerland from 1945 to 2010}} % Article title
-
-\author{
-\large
-\textsc{Oliver Hümbelin}\\[2mm] % Your name
-\normalsize Bern University of Applied Sciences \\ % Your institution
-\normalsize \href{mailto:oliver.huembelin@bfh.ch}{oliver.huembelin@bfh.ch} % Your email address
-\vspace{5mm}\\
-\large
-\textsc{Rudolf Farys}\\[2mm] % Your name
-\normalsize University of Bern \\ % Your institution
-\normalsize \href{mailto:rudolf.farys@soz.unibe.ch}{rudolf.farys@soz.unibe.ch} % Your email address
-\vspace{-5mm}
-}
-\date{}
-
-%----------------------------------------------------------------------------------------
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-
-\begin{document}
-
-\maketitle
-
-\thispagestyle{fancy} % All pages have headers and footers
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-%%%-------------------------------------------------%%%
-%%% Include abstract %%%
-%%%-------------------------------------------------%%%
-
 %%%-------------------------------------------------%%%
@@
 the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''.
-We adress the gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data.
+We adress this gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data.
 We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties.
@@
 \end{abstract}
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-\begin{multicols}{2} % Two-column layout throughout the main article text
-
-%%%-------------------------------------------------%%%
-%%% Include introduction %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document introduction %%%
-%%%-------------------------------------------------%%%
-
-\section{Introduction}
-
-% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern.
-
-% 1.Abschnitt: Bedeutung von Ungleichheit von Soziologischer Sicht
-
-\emph{Bedeutung von ökonomischer Ungleicheit} \\
-
-% 2.Abschnitt Trends in inequality research
-\emph{Studien zur allgemeinen Entwicklung der Ungleichheit nach unterschiedliche Paradigmen}(Armut, Reiche, Mittelschicht)  [Concerns about the hollowing of the middle class is a argument why the tax studies should be exploited to gain knowledge about the whole distribution of economic resources.
-Until these days tax data is only used to assess the change in top income shares.] \\
-\begin{itemize}
-  \item Development of inequality with LIS Data (Gornick and Jäntti 2013)
-  \item ranking of inequality across high-income countries (Atkinson, Rainwater and Smeeding (1995)
-   \item inequality and low income/relative poverty (Ferreira and Ravallion, 2009; Nolan and Marx, 2009; Smeeding, O’Higgins and Rainwatter, 1990; Rainwater and Smeeding 2003; Förster and Vlemickx, 2004)
-  \item inequality and top income (Atkinson and Piketty, 2007; Leigh, 2009)
-  \item inequality and the middle class (Estache and Leipziger, 2009; Littrell et al., 2001; Birdsall, 2010; Ravallion, 2010; Frank,2007)
-\end{itemize}
-
-% 3.Abschnitt: Studien mit Steuerdaten
-
-\emph{Studien mit Steuerdaten}
-\begin{itemize}
-\item Old studies with tax data (Vilfredo Pareto, 1895, 1896; Simon Kuznets, 1953)
-\item More recently revival of tax-studys with focus on top income (Piketty and Saez, 2003; Atkinson and Piketty, 2007,2010).
-\end{itemize}
-
-% 4.Abschnitt: What do we know about Switzerland?
-
-\emph{Official Data Collection} \\
-
-\begin{itemize}
-\item  LIS (1982-2004, Survey Data from Swiss Federal Statistical Office three different surveys (recent years from “income and Consumption Survey, main result; decreasing inequality)
-\item  OECD (2000 und 2004, Income and consumption survey, 2008, EU-SILC. The change of survey in 2008 is consdered as a strict break. Comparison over time not possible (OECD 2012:315)
-\item  Worldbank: Nothing
-\item  Federal Office of statistics (1998-2011, HABE, survey data, very detailed measurement, main result; inequality in post transfer income is constant)
-\item  World income inequality Database (WIID) is more or less the same data as it is published in the LIS data
-\end{itemize}
-\emph{Fazit}:LIS longest time series, comparability is questionable, more recent years time series from Federal Office of statistics.
-Both studies use survey data, which, as we will discuss later, have certain drawbacks.\\
-
-\emph{Studies on Switzerland} \\
-Vgl. Forschungsantrag (kürzen) plus neuere Studien, Verteilungsbericht 2011 und 2012 (SGB), Avenir Suisse (2013) und Gorgas and Schaltenegger (2011) und Foellmi and Martinez (2013). \\
-
-% Should we discus the non-scientic literature?
-
-% What is the purpose of our paper?
-The purpose of the following paper is threefold.
-First of all, information about economic inequality in Switzerland is sparse.
-We  fill this gap with official data from the Swiss Federal Tax Administration (FTA).
-Second, we assess the suitability of the FTA data to report inequality and its changes over time.
-Third, we integrate the development of swiss inequality into the international picture.
-We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).
-
-%%%-------------------------------------------------%%%
-%%% How did inequality change over time %%%
-%%%-------------------------------------------------%%%
-
-%<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
-%@
-
-%%%-------------------------------------------------%%%
-%%% Include data and methods %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for data and methods %%%
-%%%-------------------------------------------------%%%
-
-\section{Measurement concepts, Data and Methods}
-
-To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure.
-Based on a review on the literature about the measurment concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namly survey data.
-Afterwards we discribe the tax data published by the Swiss Federal Tax Administration highligthing the important aspects, which have to be considered working wiht FTA-Data.
-In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for switzerland.
-
-\subsection{Standards for measuring economic resources and inequality}
-
-\emph{Concepts on measuring economic resources}  \\
-Economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
-
-\emph{Defining income}
-
-\begin{itemize}
-\item Market income refers to revenue from employment.
-\item primary income or pre-transfer income refers to  market income plus revenue from property (capital income)
-\item Brutto income refers to  primary income plus social transfer
-\item Disposable income or posttransfer income refers to  brutto income minus transfer paid 
-\item adjusted disposable income refers to  disposable income corrected for the numbers of household members 
-\end{itemize}
-
-The appropriateness of the income concept is bound to the research interest.
-Studies focusing on labor market outcome look at market income.
-Studies focusing on human wellbeing follow the approach as propagated by the oecd (2013).
-This concept is largely implemented by the Federal Office of statistics.
-Following this concept one should look at income after taxes and transfers inclusive the adjustment for household members, because this is the measure which frames the consumption possibilities.
-Lastly, studies focusing on the effect of institutions (social welfare, taxation) compare the distribution of pre- and post-income measures.
-\\
-
-\emph{Population Coverage} \\
-All Residents? Working Residents? Single Person vs household/family. Households without income \\
-see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\
-
-\emph{Measuring inequality and concentration} \\
-To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
-Vgl.
-Allison (1978), Engelhardt (2000), Cowell (2000) oder Hao/Naiman (2010)
-%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
-
-neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences (Handcock/Morris 1999) \\
-
-Gini is mainly used for international comparison.
-However, other measures are often reported along (gini is more sensitive to changes in the middle of the distribution than to changes in the tails (OECD, 2008:37).
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) (Alderson and Doran 2013).
-\\
-
-Alvaredo (2010) shows formally, how top incomes shares are related to the gini-coefficient.
-Following his argument, it is crucial to consider top incomes.
-Leigh (2007) also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the studie).
-%[He also discusses tax data studies in general-> have a closer look] \\
-%[We definitely have to argue, why it isn’t sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
-
-\subsection{Comparison of tax data and other data sources - advantages and shortcomings}
-
-%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
-% Mögliche Vergleichsdimensionen
-%- Suitability of cross country comparison
-%- Assessing inequality accurately
-%- Implementation of common concepts of economic ressources
-%- contingency for intertemporal comparision
-
-%Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
-
-%Warnings about the use of tax data
-%The use of tax data is often regarded by economists with considerable disbelief.
-These doubts are well justified for at least two reasons.
-The first is that tax data are collected as part of an administrative process, which is not tailored to the scientists' needs, so that the definition of income, income unit, etc., are not necessarily those that we would have chosen.
-This causes particular difficulties for comparisons across countries, but also for time-series analysis where there have been substantial changes in the tax system, such as the moves to and from the joint taxation of couples.
-Secondly, it is obvious that those paying tax have a financial incentive to present their affairs in a way that reduces tax liabilities.
-There is tax avoidance and tax evasion.
-The rich, in particular, have a strong incentive to understate their taxable incomes.
-Those with wealth take steps to ensure that the return comes in the form of asset appreciation, typically taxed at lower rates or not at all.
-Those with high salaries seek to ensure that part of their remuneration comes in forms, such as fringe benefits or stock-options which receive favorable tax treatment.
-Both groups may make use of tax havens that allow income to be moved beyond the reach of the national tax net.
-These shortcomings limit what can be said from tax data, but this does not mean that the data are worthless.
-Like all economic data, they measure with error the 'true' variable in which we are interested.
-%The data series presented here are fairly homogenous across countries, annual, long-run, and broken down by income source for several cases.
-Users should be aware also about their limitations.
-Firstly, the series measure only top income shares and hence are silent on how inequality evolves elsewhere in the distribution [why?].
-Secondly, the series are largely concerned with gross incomes before tax.
-Thirdly, the definition of income and the unit of observation (the individual vs.
-the family) vary across countries making comparability of levels across countries more difficult.
-Even within a country, there are breaks in comparability that arise because of changes in tax legislation affecting the definition of income, although most studies try to correct for such changes to create homogenous series.
-Finally and perhaps most important, the series might be biased because of tax avoidance and tax evasion.
-For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).
-
-%Copied from:
-%Alvaredo, Facundo, Anthony B. Atkinson, Thomas Piketty and Emmanuel Saez, The World Top Incomes Database, http://topincomes.g-mond.parisschoolofeconomics.eu/ , 18/12/2013
-
-\textbf{Summary of tax data drawbacks}
-\begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial “income” see above)
-\item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the “top-income” literature, albeit it is a crucial issue in “theorized” concepts of income.
-To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
-This might influence the overall inequality, taking into account the change from traditional household and family structures over the last century.
-This is an inequality related issue, where relevant studies are missing.
-What are the assumptions?
-\item the tax data income definition summarize labor income, capital income and taxable transfer payments (no means-tested benefits as welfare aid).
-Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
-\item Some tax units are not included in the tax tables (see Foellmi and Martinez 2013:11f).(a) only taxed cases are included (partly a problem, for certain periodes we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
-Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
-In some border-cantons the share of this group is very high (20\%)
-%Can we ignore this)
-(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registred.
-This persons get an imputed income (older tax tax return and information given by employers.
-Not registered non-	fillers are not in the records.
-- (e) tax evasion.
-Feld and Frey (2007) report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
-With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequaly measures over time.
-\end{itemize}
-
-\textbf{Problems with household income surveys}
-\begin{itemize}
-\item Sample data (bias)
-\item comparability between countries and over time (depends on income definition)
-\item short time series
-\end{itemize}
-
-Atkinson et al. (2009) estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by Alvaredo (2010). \\
-
-See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
-%	•	What about the special cases?
-Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
-Exclusion leads to an underestimation of inequality
-%	•	Capital gains? Should not be included 
-The discussion about problems with reporting income is fairly exhaustive. What about wealth?
-
-\subsection{Tax data published by the Swiss Federal Tax Administration}
-
-%Beschreiben was die Daten beinhalten und wo es Probleme gibt
-\begin{itemize}
-\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
-	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
-	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
-	\item	Income after deductions-> taxable income (employment income, business income and capital income).
-Is income definition stable over time?
-Yes it should.
-Realized capital gains are excluded from the definition.
-It includes income from  employment, self-employment, capital income and taxable transfer payments.
-Plus Eigenmietwert,
-	\item	Reported on national and cantonal level
-\end{itemize}
-
-\subsection{Ways to tackle FTA-tax data specific problems}
-
-\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
-Del et al.
-(2007) impute for non-fillers the 20 percentage of the annual average income.
-This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
-Furthermore Del et al.
-calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
-\\
-
-\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
-This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
-This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f).
-\\
-%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.
-
-\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
-
-\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
-%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
-
-\textbf{deductions} Del et al.
-(2007:477):” we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.”
-Gorgas and Schaltenegger (2011:5): “”..., information on […] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
-Können wir das auch?
-Zumindest für gewisse Zeiträume?
-Das wäre noch gut.
-\\
-
-Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
-Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.
-On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a “perverse redistribution” effect by redistributing income from the lower middle class to the upper middle class (vgl.
-Peters 2011 and Interpellation Barbara Gysel (2009).
-
-% Income share specific problems
-% Total income denominator Exogenous Approach -> net income reported in the national accounts.
-Endogenous Approach -> Dell et al.
-technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income).
-Honestly , is this appropriate?
-(Everyone did it)
-% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
-% Endogenous approach -> reported in the tax tables
-
-%%%-------------------------------------------------%%%
-%%% Abschnitte aus ersten Version des Papers %%%
-%%%-------------------------------------------------%%%
-
-%\subsection{Hypotheses}
-%Based on the theories we test the following hypotheses:
-
-%\begin{itemize}
-%\item H1: Develpment of inequality is driven by sectoral change
-%\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
-%\end{itemize}
-
-%\subsection{Data and Variables}
-
-%We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010.
-While the data results in a long and consistent time series to illustrate swiss inequality development, there are a few pitfalls we want to adress which might be of interest for other research on this topic (be it in Switzerland or other countries).
-
-%\subsubsection{Left censored data}
-
-%The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household.
-The taxable population however is not identical to the population which should be used to calculate measures of inequality.
-Precisely, the data do not contain tax units with very little incomes so calculations based on these data treat the lowest percentiles equally to tax units with zero income.Figure X shows the threshold to be hit to enter the statistic.
-
-%\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}
-
-%So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data.
-Furthermore, also the changes over time might not be interpreted savely as over time the number of tax unit within this "hidden range" might vary or might even have a certain trend.
-We will adress this issue in detail in the methods chapter.
-
-%\subsubsection{Different measures, different populations}
-
-%The FTA data makes two kinds of distinctions.
-First, data was collected for so called "normal cases" and "special cases", i.e.
-a "normal case" is a taxable (for the complete tax period) person or household domiciled in a swiss canton without income from outside of Switzerland.
-A "special case" therefore is a diffuse reference category that contains tax units that are taxed at source, were not taxable for the complete tax period or generated additional income in another country.
-Second, the FTA reports two measures, that is taxable income and absolute income \textcolor{red}{(meine vorläufige Übersetzung von Reineinkommen)}.
-Absolute income is the sum of all incomes (earnings, interest income, rental incomes) minus expenses (e.g.
-from self-employment or credit cost).
-The taxable income is calculated as the difference of absolute income and deductions (e.g.
-children, insurance rates).
-The longest consistent time series exists for the taxable income of normal cases.
-So all statements we make with our data only apply to this subpopulation.
-
-%\subsubsection{Changes in taxation and measurement}
-%The swiss tax system is highly federal.
-That means, communities raise taxes which then go to the communities, the canton and the state.
-If we want to calculate overall swiss measures, we need to take into account, that cantons vary (between cantons and over time) with regard to the tax deductions that are possible and also the mechanism how taxes are collected.
-The latter adresses a comprehensive reorganization of the swiss tax system where between 1995 and 2003 cantons changed from taxing the past two years of income (postnumerando system) to taxing the present single year (praenumerando system).
-For details see \textcolor{red}{Martinez (xxxx) or some other author (xxxx)}.
-Aggregate measures of inequality therefore have to be estimated for the periods 1995 to 2003 which we adress shortly in the methods chapter.
-
-%\subsection{Methodology Used}
-%There are two steps of data analyses which need to be described to the reader.
-First, the estimation of the bias we introduce by estimating measures of inequality when tax units with too little income are not observed.
-Second, the steps undertaken to estimate aggregate swiss measures by imputing taxable income for those cantons and periods where the change of the tax system produced a gap (1995 to 2003, depending on the canton).
-
-%\subsubsection{Imputing the gap}
-%The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx).
-That is estimating the missing taxable income statistics via OLS using information from time trends and cantons.
-Our imputation model therefore includes canton inequality measures and periods dummies to explain aggregate swiss inequality.
-
-%\subsubsection{Estimating the bias}
-%For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation.
-However starting 1995, the FTA provides exactly this information for each canton.
-This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010.
-Consequently we can obtain information whether the bias is stable over time (which maes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
-\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie "je höher der Steuerfreibetrag, umso stärker der Bias".
-Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias.
-Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
-
-%%%-------------------------------------------------%%%
-%%% Include results %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document results %%%
-%%%-------------------------------------------------%%%
-
-\section{Results}
-
-\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
-Kantonsweise Grafiken
-Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
-
-%Switzerland (income and wealth)
-%- graph with gini over time> identify periods of change
-%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
-%Cantonal level (income and wealth)
-%- different development on regional level 
-%Switzerland in international comparison
-%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality
-
-%- Comparison with other countries (similar data)
-
-% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen.
-Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel.
-Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.
-
-%%%-------------------------------------------------%%%
-%%% Include discussion %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for discussion %%%
-%%%-------------------------------------------------%%%
-
-\section{Discussion}
-
-% Hier könnte allenfalls der Bogen zur Theorie gespannt werden
-
-%Methodological Main findings 
-%(1) Tax data is useful because it’s possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
-It’s superior to survey data, because the latter is affected by non-responded bias.
-Tax data includes data about the whole population … [was finden wir wegen den Nullern raus] .
-or at least it should (tax evasion).(3) The FTA-Tax data has different shortcomings, which cannot be handled like a measure of income and wealth which (a) cannot be corrected for household size and (b) is neither a pre-transfer nor a post-transfer income measure.
-Therefor it’s comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
-This FTA-Tax data problem can be handled with non-aggregated tax-data .
-
-%Main findings for Switzerland (1) Our data suggests that income inequality overall slightly increased in Switzerland and we might distinguish three episodes: 1.
-Substantial increase in times of economic growth 1950 to 1974 ending with the oil crisis 2.
-Ups and downs from 1970 to 2000 3.
-Relatively steep increase in inequality in the early 2000s
-%(2) Despite the recent increase, overall income inequality in Switzerland is not very high in international comparison.
-With respect to inequality in wealth, however, Switzerland takes a leading position in the world.
-%(3) While the overall Swiss inequality remained pretty stable over decades, inequality between cantons underwent a bizarre development.
-
-%%%-------------------------------------------------%%%
-%%% Include acknowledgements %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for acknowledgement %%%
-%%%-------------------------------------------------%%%
-
-\section{Acknowledgements}
-
-We thank Ben Jann, Robert Fluder and Tobias Fritschi for helpful comments on the article. We also like to thank Stefan Ilic for the prepartion of the data set.
-
-%%%-------------------------------------------------%%%
-%%% Include the bibliography %%%
-%%%-------------------------------------------------%%%
-
-\end{multicols}
-
-\bibliography{bibliography/bib}
-
-%%%-------------------------------------------------%%%
-%%% Include the appendix %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for appendix %%%
-%%%-------------------------------------------------%%%
-
-\section{Appendix}
-
-\end{document}

### 2014-01-07 paper/subdocuments/data_and_methods.tex
Übertrag aus Skizze

--- 2014-01-07 paper/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Journal Article
-% LaTeX Template
-% Version 1.3 (9/9/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Frits Wenneker (http://www.howtotex.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[twoside]{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
-
-\usepackage{lipsum} % Package to generate dummy text throughout this template
-
-\usepackage[sc]{mathpazo} % Use the Palatino font
-%\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
-\usepackage[utf8]{inputenc}
-\linespread{1.05} % Line spacing - Palatino needs more space between lines
-\usepackage{microtype} % Slightly tweak font spacing for aesthetics
-
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
-\usepackage{multicol} % Used for the two-column layout of the document
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
-\usepackage{booktabs} % Horizontal rules in tables
-\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
-\usepackage{hyperref} % For hyperlinks in the PDF
-
-\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
-\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
-
-\usepackage{color} % used to mark parts that need to be edited
-
-\usepackage{abstract} % Allows abstract customization
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text
-
-\usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
-
-\usepackage{fancyhdr} % Headers and footers
-\pagestyle{fancy} % All pages have headers and footers
-\fancyhead{} % Blank out the default header
-\fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
-\fancyfoot[RO,LE]{\thepage} % Custom footer text
-\bibliographystyle{plain}
-
-%%%-------------------------------------------------%%%
-%%% Preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document global preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%----------------------------------------------------------------------------------------
-%  TITLE SECTION
-%----------------------------------------------------------------------------------------
-
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing inequality with tax data - Switzerland from 1945 to 2010}} % Article title
-
-\author{
-\large
-\textsc{Oliver Hümbelin}\\[2mm] % Your name
-\normalsize Bern University of Applied Sciences \\ % Your institution
-\normalsize \href{mailto:oliver.huembelin@bfh.ch}{oliver.huembelin@bfh.ch} % Your email address
-\vspace{5mm}\\
-\large
-\textsc{Rudolf Farys}\\[2mm] % Your name
-\normalsize University of Bern \\ % Your institution
-\normalsize \href{mailto:rudolf.farys@soz.unibe.ch}{rudolf.farys@soz.unibe.ch} % Your email address
-\vspace{-5mm}
-}
-\date{}
-
-%----------------------------------------------------------------------------------------
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-
-\begin{document}
-
-\maketitle
-
-\thispagestyle{fancy} % All pages have headers and footers
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-%%%-------------------------------------------------%%%
-%%% Include abstract %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document abstract %%%
-%%%-------------------------------------------------%%%
-
-\begin{abstract}
-
-There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
-In Switzerland, however, the development is unclear, as there is only little systematic evidence that would allow for long-term comparison.
-Nevertheless income inequality has been a prominent theme in the public discussion in recent years, e.g.
-the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''.
-We adress the gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data.
-We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties.
-In the end we integrate the case of Switzerland into the international picture of inequality development showing parallels and deviations.
-
-\end{abstract}
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-\begin{multicols}{2} % Two-column layout throughout the main article text
-
-%%%-------------------------------------------------%%%
-%%% Include introduction %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document introduction %%%
-%%%-------------------------------------------------%%%
-
-\section{Introduction}
-
-% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern.
-
-% 1.Abschnitt: Bedeutung von Ungleichheit von Soziologischer Sicht
-
-\emph{Bedeutung von ökonomischer Ungleicheit} \\
-
-% 2.Abschnitt Trends in inequality research
-\emph{Studien zur allgemeinen Entwicklung der Ungleichheit nach unterschiedliche Paradigmen}(Armut, Reiche, Mittelschicht)  [Concerns about the hollowing of the middle class is a argument why the tax studies should be exploited to gain knowledge about the whole distribution of economic resources.
-Until these days tax data is only used to assess the change in top income shares.] \\
-\begin{itemize}
-  \item Development of inequality with LIS Data (Gornick and Jäntti 2013)
-  \item ranking of inequality across high-income countries (Atkinson, Rainwater and Smeeding (1995)
-   \item inequality and low income/relative poverty (Ferreira and Ravallion, 2009; Nolan and Marx, 2009; Smeeding, O’Higgins and Rainwatter, 1990; Rainwater and Smeeding 2003; Förster and Vlemickx, 2004)
-  \item inequality and top income (Atkinson and Piketty, 2007; Leigh, 2009)
-  \item inequality and the middle class (Estache and Leipziger, 2009; Littrell et al., 2001; Birdsall, 2010; Ravallion, 2010; Frank,2007)
-\end{itemize}
-
-% 3.Abschnitt: Studien mit Steuerdaten
-
-\emph{Studien mit Steuerdaten}
-\begin{itemize}
-\item Old studies with tax data (Vilfredo Pareto, 1895, 1896; Simon Kuznets, 1953)
-\item More recently revival of tax-studys with focus on top income (Piketty and Saez, 2003; Atkinson and Piketty, 2007,2010).
-\end{itemize}
-
-% 4.Abschnitt: What do we know about Switzerland?
-
-\emph{Official Data Collection} \\
-
-\begin{itemize}
-\item  LIS (1982-2004, Survey Data from Swiss Federal Statistical Office three different surveys (recent years from “income and Consumption Survey, main result; decreasing inequality)
-\item  OECD (2000 und 2004, Income and consumption survey, 2008, EU-SILC. The change of survey in 2008 is consdered as a strict break. Comparison over time not possible (OECD 2012:315)
-\item  Worldbank: Nothing
-\item  Federal Office of statistics (1998-2011, HABE, survey data, very detailed measurement, main result; inequality in post transfer income is constant)
-\item  World income inequality Database (WIID) is more or less the same data as it is published in the LIS data
-\end{itemize}
-\emph{Fazit}:LIS longest time series, comparability is questionable, more recent years time series from Federal Office of statistics.
-Both studies use survey data, which, as we will discuss later, have certain drawbacks.\\
-
-\emph{Studies on Switzerland} \\
-Vgl. Forschungsantrag (kürzen) plus neuere Studien, Verteilungsbericht 2011 und 2012 (SGB), Avenir Suisse (2013) und Gorgas and Schaltenegger (2011) und Foellmi and Martinez (2013). \\
-
-% Should we discus the non-scientic literature?
-
-% What is the purpose of our paper?
-The purpose of the following paper is threefold.
-First of all, information about economic inequality in Switzerland is sparse.
-We  fill this gap with official data from the Swiss Federal Tax Administration (FTA).
-Second, we assess the suitability of the FTA data to report inequality and its changes over time.
-Third, we integrate the development of swiss inequality into the international picture.
-We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).
-
-%%%-------------------------------------------------%%%
-%%% How did inequality change over time %%%
-%%%-------------------------------------------------%%%
-
-%<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
-%@
-
-%%%-------------------------------------------------%%%
-%%% Include data and methods %%%
-%%%-------------------------------------------------%%%
-
 %%%-------------------------------------------------%%%
@@
 
-\textbf{Summary of tax data drawbacks}
+\textbf{Summary of tax data drawbacks }
 \begin{itemize}
@@
 
-%Beschreiben was die Daten beinhalten und wo es Probleme gibt
-\begin{itemize}
-\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
-	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
-	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
-	\item	Income after deductions-> taxable income (employment income, business income and capital income).
-Is income definition stable over time?
-Yes it should.
-Realized capital gains are excluded from the definition.
-It includes income from  employment, self-employment, capital income and taxable transfer payments.
-Plus Eigenmietwert,
-	\item	Reported on national and cantonal level
-\end{itemize}
-
 \subsection{Ways to tackle FTA-tax data specific problems}
-
-\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
-Del et al.
-(2007) impute for non-fillers the 20 percentage of the annual average income.
-This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
-Furthermore Del et al.
-calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
-\\
-
-\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
-This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
-This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f).
-\\
-%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.
-
-\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
-
-\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
-%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
-
-\textbf{deductions} Del et al.
-(2007:477):” we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.”
-Gorgas and Schaltenegger (2011:5): “”..., information on […] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
-Können wir das auch?
-Zumindest für gewisse Zeiträume?
-Das wäre noch gut.
-\\
-
-Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
-Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.
-On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a “perverse redistribution” effect by redistributing income from the lower middle class to the upper middle class (vgl.
-Peters 2011 and Interpellation Barbara Gysel (2009).
-
-% Income share specific problems
-% Total income denominator Exogenous Approach -> net income reported in the national accounts.
-Endogenous Approach -> Dell et al.
-technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income).
-Honestly , is this appropriate?
-(Everyone did it)
-% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
-% Endogenous approach -> reported in the tax tables
-
-%%%-------------------------------------------------%%%
-%%% Abschnitte aus ersten Version des Papers %%%
-%%%-------------------------------------------------%%%
 
@@
 Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
-
-%%%-------------------------------------------------%%%
-%%% Include results %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document results %%%
-%%%-------------------------------------------------%%%
-
-\section{Results}
-
-\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
-Kantonsweise Grafiken
-Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
-
-%Switzerland (income and wealth)
-%- graph with gini over time> identify periods of change
-%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
-%Cantonal level (income and wealth)
-%- different development on regional level 
-%Switzerland in international comparison
-%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality
-
-%- Comparison with other countries (similar data)
-
-% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen.
-Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel.
-Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.
-
-%%%-------------------------------------------------%%%
-%%% Include discussion %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for discussion %%%
-%%%-------------------------------------------------%%%
-
-\section{Discussion}
-
-% Hier könnte allenfalls der Bogen zur Theorie gespannt werden
-
-%Methodological Main findings 
-%(1) Tax data is useful because it’s possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
-It’s superior to survey data, because the latter is affected by non-responded bias.
-Tax data includes data about the whole population … [was finden wir wegen den Nullern raus] .
-or at least it should (tax evasion).(3) The FTA-Tax data has different shortcomings, which cannot be handled like a measure of income and wealth which (a) cannot be corrected for household size and (b) is neither a pre-transfer nor a post-transfer income measure.
-Therefor it’s comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
-This FTA-Tax data problem can be handled with non-aggregated tax-data .
-
-%Main findings for Switzerland (1) Our data suggests that income inequality overall slightly increased in Switzerland and we might distinguish three episodes: 1.
-Substantial increase in times of economic growth 1950 to 1974 ending with the oil crisis 2.
-Ups and downs from 1970 to 2000 3.
-Relatively steep increase in inequality in the early 2000s
-%(2) Despite the recent increase, overall income inequality in Switzerland is not very high in international comparison.
-With respect to inequality in wealth, however, Switzerland takes a leading position in the world.
-%(3) While the overall Swiss inequality remained pretty stable over decades, inequality between cantons underwent a bizarre development.
-
-%%%-------------------------------------------------%%%
-%%% Include acknowledgements %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for acknowledgement %%%
-%%%-------------------------------------------------%%%
-
-\section{Acknowledgements}
-
-We thank Ben Jann, Robert Fluder and Tobias Fritschi for helpful comments on the article. We also like to thank Stefan Ilic for the prepartion of the data set.
-
-%%%-------------------------------------------------%%%
-%%% Include the bibliography %%%
-%%%-------------------------------------------------%%%
-
-\end{multicols}
-
-\bibliography{bibliography/bib}
-
-%%%-------------------------------------------------%%%
-%%% Include the appendix %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for appendix %%%
-%%%-------------------------------------------------%%%
-
-\section{Appendix}
-
-\end{document}

### 2014-01-07 paper/subdocuments/discussion.Rnw
Übertrag aus Skizze

--- 2014-01-07 paper/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Journal Article
-% LaTeX Template
-% Version 1.3 (9/9/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Frits Wenneker (http://www.howtotex.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[twoside]{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
-
-\usepackage{lipsum} % Package to generate dummy text throughout this template
-
-\usepackage[sc]{mathpazo} % Use the Palatino font
-%\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
-\usepackage[utf8]{inputenc}
-\linespread{1.05} % Line spacing - Palatino needs more space between lines
-\usepackage{microtype} % Slightly tweak font spacing for aesthetics
-
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
-\usepackage{multicol} % Used for the two-column layout of the document
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
-\usepackage{booktabs} % Horizontal rules in tables
-\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
-\usepackage{hyperref} % For hyperlinks in the PDF
-
-\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
-\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
-
-\usepackage{color} % used to mark parts that need to be edited
-
-\usepackage{abstract} % Allows abstract customization
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text
-
-\usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
-
-\usepackage{fancyhdr} % Headers and footers
-\pagestyle{fancy} % All pages have headers and footers
-\fancyhead{} % Blank out the default header
-\fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
-\fancyfoot[RO,LE]{\thepage} % Custom footer text
-\bibliographystyle{plain}
-
-%%%-------------------------------------------------%%%
-%%% Preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document global preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%----------------------------------------------------------------------------------------
-%  TITLE SECTION
-%----------------------------------------------------------------------------------------
-
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing inequality with tax data - Switzerland from 1945 to 2010}} % Article title
-
-\author{
-\large
-\textsc{Oliver Hümbelin}\\[2mm] % Your name
-\normalsize Bern University of Applied Sciences \\ % Your institution
-\normalsize \href{mailto:oliver.huembelin@bfh.ch}{oliver.huembelin@bfh.ch} % Your email address
-\vspace{5mm}\\
-\large
-\textsc{Rudolf Farys}\\[2mm] % Your name
-\normalsize University of Bern \\ % Your institution
-\normalsize \href{mailto:rudolf.farys@soz.unibe.ch}{rudolf.farys@soz.unibe.ch} % Your email address
-\vspace{-5mm}
-}
-\date{}
-
-%----------------------------------------------------------------------------------------
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-
-\begin{document}
-
-\maketitle
-
-\thispagestyle{fancy} % All pages have headers and footers
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-%%%-------------------------------------------------%%%
-%%% Include abstract %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document abstract %%%
-%%%-------------------------------------------------%%%
-
-\begin{abstract}
-
-There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
-In Switzerland, however, the development is unclear, as there is only little systematic evidence that would allow for long-term comparison.
-Nevertheless income inequality has been a prominent theme in the public discussion in recent years, e.g.
-the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''.
-We adress the gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data.
-We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties.
-In the end we integrate the case of Switzerland into the international picture of inequality development showing parallels and deviations.
-
-\end{abstract}
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-\begin{multicols}{2} % Two-column layout throughout the main article text
-
-%%%-------------------------------------------------%%%
-%%% Include introduction %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document introduction %%%
-%%%-------------------------------------------------%%%
-
-\section{Introduction}
-
-% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern.
-
-% 1.Abschnitt: Bedeutung von Ungleichheit von Soziologischer Sicht
-
-\emph{Bedeutung von ökonomischer Ungleicheit} \\
-
-% 2.Abschnitt Trends in inequality research
-\emph{Studien zur allgemeinen Entwicklung der Ungleichheit nach unterschiedliche Paradigmen}(Armut, Reiche, Mittelschicht)  [Concerns about the hollowing of the middle class is a argument why the tax studies should be exploited to gain knowledge about the whole distribution of economic resources.
-Until these days tax data is only used to assess the change in top income shares.] \\
-\begin{itemize}
-  \item Development of inequality with LIS Data (Gornick and Jäntti 2013)
-  \item ranking of inequality across high-income countries (Atkinson, Rainwater and Smeeding (1995)
-   \item inequality and low income/relative poverty (Ferreira and Ravallion, 2009; Nolan and Marx, 2009; Smeeding, O’Higgins and Rainwatter, 1990; Rainwater and Smeeding 2003; Förster and Vlemickx, 2004)
-  \item inequality and top income (Atkinson and Piketty, 2007; Leigh, 2009)
-  \item inequality and the middle class (Estache and Leipziger, 2009; Littrell et al., 2001; Birdsall, 2010; Ravallion, 2010; Frank,2007)
-\end{itemize}
-
-% 3.Abschnitt: Studien mit Steuerdaten
-
-\emph{Studien mit Steuerdaten}
-\begin{itemize}
-\item Old studies with tax data (Vilfredo Pareto, 1895, 1896; Simon Kuznets, 1953)
-\item More recently revival of tax-studys with focus on top income (Piketty and Saez, 2003; Atkinson and Piketty, 2007,2010).
-\end{itemize}
-
-% 4.Abschnitt: What do we know about Switzerland?
-
-\emph{Official Data Collection} \\
-
-\begin{itemize}
-\item  LIS (1982-2004, Survey Data from Swiss Federal Statistical Office three different surveys (recent years from “income and Consumption Survey, main result; decreasing inequality)
-\item  OECD (2000 und 2004, Income and consumption survey, 2008, EU-SILC. The change of survey in 2008 is consdered as a strict break. Comparison over time not possible (OECD 2012:315)
-\item  Worldbank: Nothing
-\item  Federal Office of statistics (1998-2011, HABE, survey data, very detailed measurement, main result; inequality in post transfer income is constant)
-\item  World income inequality Database (WIID) is more or less the same data as it is published in the LIS data
-\end{itemize}
-\emph{Fazit}:LIS longest time series, comparability is questionable, more recent years time series from Federal Office of statistics.
-Both studies use survey data, which, as we will discuss later, have certain drawbacks.\\
-
-\emph{Studies on Switzerland} \\
-Vgl. Forschungsantrag (kürzen) plus neuere Studien, Verteilungsbericht 2011 und 2012 (SGB), Avenir Suisse (2013) und Gorgas and Schaltenegger (2011) und Foellmi and Martinez (2013). \\
-
-% Should we discus the non-scientic literature?
-
-% What is the purpose of our paper?
-The purpose of the following paper is threefold.
-First of all, information about economic inequality in Switzerland is sparse.
-We  fill this gap with official data from the Swiss Federal Tax Administration (FTA).
-Second, we assess the suitability of the FTA data to report inequality and its changes over time.
-Third, we integrate the development of swiss inequality into the international picture.
-We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).
-
-%%%-------------------------------------------------%%%
-%%% How did inequality change over time %%%
-%%%-------------------------------------------------%%%
-
-%<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
-%@
-
-%%%-------------------------------------------------%%%
-%%% Include data and methods %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for data and methods %%%
-%%%-------------------------------------------------%%%
-
-\section{Measurement concepts, Data and Methods}
-
-To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure.
-Based on a review on the literature about the measurment concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namly survey data.
-Afterwards we discribe the tax data published by the Swiss Federal Tax Administration highligthing the important aspects, which have to be considered working wiht FTA-Data.
-In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for switzerland.
-
-\subsection{Standards for measuring economic resources and inequality}
-
-\emph{Concepts on measuring economic resources}  \\
-Economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
-
-\emph{Defining income}
-
-\begin{itemize}
-\item Market income refers to revenue from employment.
-\item primary income or pre-transfer income refers to  market income plus revenue from property (capital income)
-\item Brutto income refers to  primary income plus social transfer
-\item Disposable income or posttransfer income refers to  brutto income minus transfer paid 
-\item adjusted disposable income refers to  disposable income corrected for the numbers of household members 
-\end{itemize}
-
-The appropriateness of the income concept is bound to the research interest.
-Studies focusing on labor market outcome look at market income.
-Studies focusing on human wellbeing follow the approach as propagated by the oecd (2013).
-This concept is largely implemented by the Federal Office of statistics.
-Following this concept one should look at income after taxes and transfers inclusive the adjustment for household members, because this is the measure which frames the consumption possibilities.
-Lastly, studies focusing on the effect of institutions (social welfare, taxation) compare the distribution of pre- and post-income measures.
-\\
-
-\emph{Population Coverage} \\
-All Residents? Working Residents? Single Person vs household/family. Households without income \\
-see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\
-
-\emph{Measuring inequality and concentration} \\
-To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
-Vgl.
-Allison (1978), Engelhardt (2000), Cowell (2000) oder Hao/Naiman (2010)
-%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
-
-neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences (Handcock/Morris 1999) \\
-
-Gini is mainly used for international comparison.
-However, other measures are often reported along (gini is more sensitive to changes in the middle of the distribution than to changes in the tails (OECD, 2008:37).
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) (Alderson and Doran 2013).
-\\
-
-Alvaredo (2010) shows formally, how top incomes shares are related to the gini-coefficient.
-Following his argument, it is crucial to consider top incomes.
-Leigh (2007) also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the studie).
-%[He also discusses tax data studies in general-> have a closer look] \\
-%[We definitely have to argue, why it isn’t sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
-
-\subsection{Comparison of tax data and other data sources - advantages and shortcomings}
-
-%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
-% Mögliche Vergleichsdimensionen
-%- Suitability of cross country comparison
-%- Assessing inequality accurately
-%- Implementation of common concepts of economic ressources
-%- contingency for intertemporal comparision
-
-%Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
-
-%Warnings about the use of tax data
-%The use of tax data is often regarded by economists with considerable disbelief.
-These doubts are well justified for at least two reasons.
-The first is that tax data are collected as part of an administrative process, which is not tailored to the scientists' needs, so that the definition of income, income unit, etc., are not necessarily those that we would have chosen.
-This causes particular difficulties for comparisons across countries, but also for time-series analysis where there have been substantial changes in the tax system, such as the moves to and from the joint taxation of couples.
-Secondly, it is obvious that those paying tax have a financial incentive to present their affairs in a way that reduces tax liabilities.
-There is tax avoidance and tax evasion.
-The rich, in particular, have a strong incentive to understate their taxable incomes.
-Those with wealth take steps to ensure that the return comes in the form of asset appreciation, typically taxed at lower rates or not at all.
-Those with high salaries seek to ensure that part of their remuneration comes in forms, such as fringe benefits or stock-options which receive favorable tax treatment.
-Both groups may make use of tax havens that allow income to be moved beyond the reach of the national tax net.
-These shortcomings limit what can be said from tax data, but this does not mean that the data are worthless.
-Like all economic data, they measure with error the 'true' variable in which we are interested.
-%The data series presented here are fairly homogenous across countries, annual, long-run, and broken down by income source for several cases.
-Users should be aware also about their limitations.
-Firstly, the series measure only top income shares and hence are silent on how inequality evolves elsewhere in the distribution [why?].
-Secondly, the series are largely concerned with gross incomes before tax.
-Thirdly, the definition of income and the unit of observation (the individual vs.
-the family) vary across countries making comparability of levels across countries more difficult.
-Even within a country, there are breaks in comparability that arise because of changes in tax legislation affecting the definition of income, although most studies try to correct for such changes to create homogenous series.
-Finally and perhaps most important, the series might be biased because of tax avoidance and tax evasion.
-For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).
-
-%Copied from:
-%Alvaredo, Facundo, Anthony B. Atkinson, Thomas Piketty and Emmanuel Saez, The World Top Incomes Database, http://topincomes.g-mond.parisschoolofeconomics.eu/ , 18/12/2013
-
-\textbf{Summary of tax data drawbacks}
-\begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial “income” see above)
-\item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the “top-income” literature, albeit it is a crucial issue in “theorized” concepts of income.
-To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
-This might influence the overall inequality, taking into account the change from traditional household and family structures over the last century.
-This is an inequality related issue, where relevant studies are missing.
-What are the assumptions?
-\item the tax data income definition summarize labor income, capital income and taxable transfer payments (no means-tested benefits as welfare aid).
-Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
-\item Some tax units are not included in the tax tables (see Foellmi and Martinez 2013:11f).(a) only taxed cases are included (partly a problem, for certain periodes we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
-Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
-In some border-cantons the share of this group is very high (20\%)
-%Can we ignore this)
-(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registred.
-This persons get an imputed income (older tax tax return and information given by employers.
-Not registered non-	fillers are not in the records.
-- (e) tax evasion.
-Feld and Frey (2007) report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
-With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequaly measures over time.
-\end{itemize}
-
-\textbf{Problems with household income surveys}
-\begin{itemize}
-\item Sample data (bias)
-\item comparability between countries and over time (depends on income definition)
-\item short time series
-\end{itemize}
-
-Atkinson et al. (2009) estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by Alvaredo (2010). \\
-
-See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
-%	•	What about the special cases?
-Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
-Exclusion leads to an underestimation of inequality
-%	•	Capital gains? Should not be included 
-The discussion about problems with reporting income is fairly exhaustive. What about wealth?
-
-\subsection{Tax data published by the Swiss Federal Tax Administration}
-
-%Beschreiben was die Daten beinhalten und wo es Probleme gibt
-\begin{itemize}
-\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
-	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
-	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
-	\item	Income after deductions-> taxable income (employment income, business income and capital income).
-Is income definition stable over time?
-Yes it should.
-Realized capital gains are excluded from the definition.
-It includes income from  employment, self-employment, capital income and taxable transfer payments.
-Plus Eigenmietwert,
-	\item	Reported on national and cantonal level
-\end{itemize}
-
-\subsection{Ways to tackle FTA-tax data specific problems}
-
-\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
-Del et al.
-(2007) impute for non-fillers the 20 percentage of the annual average income.
-This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
-Furthermore Del et al.
-calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
-\\
-
-\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
-This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
-This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f).
-\\
-%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.
-
-\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
-
-\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
-%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
-
-\textbf{deductions} Del et al.
-(2007:477):” we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.”
-Gorgas and Schaltenegger (2011:5): “”..., information on […] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
-Können wir das auch?
-Zumindest für gewisse Zeiträume?
-Das wäre noch gut.
-\\
-
-Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
-Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.
-On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a “perverse redistribution” effect by redistributing income from the lower middle class to the upper middle class (vgl.
-Peters 2011 and Interpellation Barbara Gysel (2009).
-
-% Income share specific problems
-% Total income denominator Exogenous Approach -> net income reported in the national accounts.
-Endogenous Approach -> Dell et al.
-technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income).
-Honestly , is this appropriate?
-(Everyone did it)
-% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
-% Endogenous approach -> reported in the tax tables
-
-%%%-------------------------------------------------%%%
-%%% Abschnitte aus ersten Version des Papers %%%
-%%%-------------------------------------------------%%%
-
-%\subsection{Hypotheses}
-%Based on the theories we test the following hypotheses:
-
-%\begin{itemize}
-%\item H1: Develpment of inequality is driven by sectoral change
-%\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
-%\end{itemize}
-
-%\subsection{Data and Variables}
-
-%We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010.
-While the data results in a long and consistent time series to illustrate swiss inequality development, there are a few pitfalls we want to adress which might be of interest for other research on this topic (be it in Switzerland or other countries).
-
-%\subsubsection{Left censored data}
-
-%The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household.
-The taxable population however is not identical to the population which should be used to calculate measures of inequality.
-Precisely, the data do not contain tax units with very little incomes so calculations based on these data treat the lowest percentiles equally to tax units with zero income.Figure X shows the threshold to be hit to enter the statistic.
-
-%\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}
-
-%So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data.
-Furthermore, also the changes over time might not be interpreted savely as over time the number of tax unit within this "hidden range" might vary or might even have a certain trend.
-We will adress this issue in detail in the methods chapter.
-
-%\subsubsection{Different measures, different populations}
-
-%The FTA data makes two kinds of distinctions.
-First, data was collected for so called "normal cases" and "special cases", i.e.
-a "normal case" is a taxable (for the complete tax period) person or household domiciled in a swiss canton without income from outside of Switzerland.
-A "special case" therefore is a diffuse reference category that contains tax units that are taxed at source, were not taxable for the complete tax period or generated additional income in another country.
-Second, the FTA reports two measures, that is taxable income and absolute income \textcolor{red}{(meine vorläufige Übersetzung von Reineinkommen)}.
-Absolute income is the sum of all incomes (earnings, interest income, rental incomes) minus expenses (e.g.
-from self-employment or credit cost).
-The taxable income is calculated as the difference of absolute income and deductions (e.g.
-children, insurance rates).
-The longest consistent time series exists for the taxable income of normal cases.
-So all statements we make with our data only apply to this subpopulation.
-
-%\subsubsection{Changes in taxation and measurement}
-%The swiss tax system is highly federal.
-That means, communities raise taxes which then go to the communities, the canton and the state.
-If we want to calculate overall swiss measures, we need to take into account, that cantons vary (between cantons and over time) with regard to the tax deductions that are possible and also the mechanism how taxes are collected.
-The latter adresses a comprehensive reorganization of the swiss tax system where between 1995 and 2003 cantons changed from taxing the past two years of income (postnumerando system) to taxing the present single year (praenumerando system).
-For details see \textcolor{red}{Martinez (xxxx) or some other author (xxxx)}.
-Aggregate measures of inequality therefore have to be estimated for the periods 1995 to 2003 which we adress shortly in the methods chapter.
-
-%\subsection{Methodology Used}
-%There are two steps of data analyses which need to be described to the reader.
-First, the estimation of the bias we introduce by estimating measures of inequality when tax units with too little income are not observed.
-Second, the steps undertaken to estimate aggregate swiss measures by imputing taxable income for those cantons and periods where the change of the tax system produced a gap (1995 to 2003, depending on the canton).
-
-%\subsubsection{Imputing the gap}
-%The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx).
-That is estimating the missing taxable income statistics via OLS using information from time trends and cantons.
-Our imputation model therefore includes canton inequality measures and periods dummies to explain aggregate swiss inequality.
-
-%\subsubsection{Estimating the bias}
-%For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation.
-However starting 1995, the FTA provides exactly this information for each canton.
-This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010.
-Consequently we can obtain information whether the bias is stable over time (which maes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
-\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie "je höher der Steuerfreibetrag, umso stärker der Bias".
-Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias.
-Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
-
-%%%-------------------------------------------------%%%
-%%% Include results %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document results %%%
-%%%-------------------------------------------------%%%
-
-\section{Results}
-
-\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
-Kantonsweise Grafiken
-Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
-
-%Switzerland (income and wealth)
-%- graph with gini over time> identify periods of change
-%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
-%Cantonal level (income and wealth)
-%- different development on regional level 
-%Switzerland in international comparison
-%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality
-
-%- Comparison with other countries (similar data)
-
-% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen.
-Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel.
-Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.
-
-%%%-------------------------------------------------%%%
-%%% Include discussion %%%
-%%%-------------------------------------------------%%%
-
 %%%-------------------------------------------------%%%
@@
 %(3) While the overall Swiss inequality remained pretty stable over decades, inequality between cantons underwent a bizarre development.
-
-%%%-------------------------------------------------%%%
-%%% Include acknowledgements %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for acknowledgement %%%
-%%%-------------------------------------------------%%%
-
-\section{Acknowledgements}
-
-We thank Ben Jann, Robert Fluder and Tobias Fritschi for helpful comments on the article. We also like to thank Stefan Ilic for the prepartion of the data set.
-
-%%%-------------------------------------------------%%%
-%%% Include the bibliography %%%
-%%%-------------------------------------------------%%%
-
-\end{multicols}
-
-\bibliography{bibliography/bib}
-
-%%%-------------------------------------------------%%%
-%%% Include the appendix %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for appendix %%%
-%%%-------------------------------------------------%%%
-
-\section{Appendix}
-
-\end{document}

### 2014-01-07 paper/subdocuments/introduction.tex
Übertrag aus Skizze

--- 2014-01-07 paper/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Journal Article
-% LaTeX Template
-% Version 1.3 (9/9/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Frits Wenneker (http://www.howtotex.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[twoside]{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
-
-\usepackage{lipsum} % Package to generate dummy text throughout this template
-
-\usepackage[sc]{mathpazo} % Use the Palatino font
-%\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
-\usepackage[utf8]{inputenc}
-\linespread{1.05} % Line spacing - Palatino needs more space between lines
-\usepackage{microtype} % Slightly tweak font spacing for aesthetics
-
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
-\usepackage{multicol} % Used for the two-column layout of the document
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
-\usepackage{booktabs} % Horizontal rules in tables
-\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
-\usepackage{hyperref} % For hyperlinks in the PDF
-
-\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
-\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
-
-\usepackage{color} % used to mark parts that need to be edited
-
-\usepackage{abstract} % Allows abstract customization
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text
-
-\usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
-
-\usepackage{fancyhdr} % Headers and footers
-\pagestyle{fancy} % All pages have headers and footers
-\fancyhead{} % Blank out the default header
-\fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
-\fancyfoot[RO,LE]{\thepage} % Custom footer text
-\bibliographystyle{plain}
-
-%%%-------------------------------------------------%%%
-%%% Preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document global preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%----------------------------------------------------------------------------------------
-%  TITLE SECTION
-%----------------------------------------------------------------------------------------
-
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing inequality with tax data - Switzerland from 1945 to 2010}} % Article title
-
-\author{
-\large
-\textsc{Oliver Hümbelin}\\[2mm] % Your name
-\normalsize Bern University of Applied Sciences \\ % Your institution
-\normalsize \href{mailto:oliver.huembelin@bfh.ch}{oliver.huembelin@bfh.ch} % Your email address
-\vspace{5mm}\\
-\large
-\textsc{Rudolf Farys}\\[2mm] % Your name
-\normalsize University of Bern \\ % Your institution
-\normalsize \href{mailto:rudolf.farys@soz.unibe.ch}{rudolf.farys@soz.unibe.ch} % Your email address
-\vspace{-5mm}
-}
-\date{}
-
-%----------------------------------------------------------------------------------------
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-
-\begin{document}
-
-\maketitle
-
-\thispagestyle{fancy} % All pages have headers and footers
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-%%%-------------------------------------------------%%%
-%%% Include abstract %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document abstract %%%
-%%%-------------------------------------------------%%%
-
-\begin{abstract}
-
-There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
-In Switzerland, however, the development is unclear, as there is only little systematic evidence that would allow for long-term comparison.
-Nevertheless income inequality has been a prominent theme in the public discussion in recent years, e.g.
-the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''.
-We adress the gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data.
-We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties.
-In the end we integrate the case of Switzerland into the international picture of inequality development showing parallels and deviations.
-
-\end{abstract}
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-\begin{multicols}{2} % Two-column layout throughout the main article text
-
-%%%-------------------------------------------------%%%
-%%% Include introduction %%%
-%%%-------------------------------------------------%%%
-
 %%%-------------------------------------------------%%%
@@
 We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).
-
-%%%-------------------------------------------------%%%
-%%% How did inequality change over time %%%
-%%%-------------------------------------------------%%%
-
-%<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
-%@
-
-%%%-------------------------------------------------%%%
-%%% Include data and methods %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for data and methods %%%
-%%%-------------------------------------------------%%%
-
-\section{Measurement concepts, Data and Methods}
-
-To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure.
-Based on a review on the literature about the measurment concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namly survey data.
-Afterwards we discribe the tax data published by the Swiss Federal Tax Administration highligthing the important aspects, which have to be considered working wiht FTA-Data.
-In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for switzerland.
-
-\subsection{Standards for measuring economic resources and inequality}
-
-\emph{Concepts on measuring economic resources}  \\
-Economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
-
-\emph{Defining income}
-
-\begin{itemize}
-\item Market income refers to revenue from employment.
-\item primary income or pre-transfer income refers to  market income plus revenue from property (capital income)
-\item Brutto income refers to  primary income plus social transfer
-\item Disposable income or posttransfer income refers to  brutto income minus transfer paid 
-\item adjusted disposable income refers to  disposable income corrected for the numbers of household members 
-\end{itemize}
-
-The appropriateness of the income concept is bound to the research interest.
-Studies focusing on labor market outcome look at market income.
-Studies focusing on human wellbeing follow the approach as propagated by the oecd (2013).
-This concept is largely implemented by the Federal Office of statistics.
-Following this concept one should look at income after taxes and transfers inclusive the adjustment for household members, because this is the measure which frames the consumption possibilities.
-Lastly, studies focusing on the effect of institutions (social welfare, taxation) compare the distribution of pre- and post-income measures.
-\\
-
-\emph{Population Coverage} \\
-All Residents? Working Residents? Single Person vs household/family. Households without income \\
-see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\
-
-\emph{Measuring inequality and concentration} \\
-To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
-Vgl.
-Allison (1978), Engelhardt (2000), Cowell (2000) oder Hao/Naiman (2010)
-%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
-
-neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences (Handcock/Morris 1999) \\
-
-Gini is mainly used for international comparison.
-However, other measures are often reported along (gini is more sensitive to changes in the middle of the distribution than to changes in the tails (OECD, 2008:37).
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) (Alderson and Doran 2013).
-\\
-
-Alvaredo (2010) shows formally, how top incomes shares are related to the gini-coefficient.
-Following his argument, it is crucial to consider top incomes.
-Leigh (2007) also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the studie).
-%[He also discusses tax data studies in general-> have a closer look] \\
-%[We definitely have to argue, why it isn’t sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
-
-\subsection{Comparison of tax data and other data sources - advantages and shortcomings}
-
-%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
-% Mögliche Vergleichsdimensionen
-%- Suitability of cross country comparison
-%- Assessing inequality accurately
-%- Implementation of common concepts of economic ressources
-%- contingency for intertemporal comparision
-
-%Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
-
-%Warnings about the use of tax data
-%The use of tax data is often regarded by economists with considerable disbelief.
-These doubts are well justified for at least two reasons.
-The first is that tax data are collected as part of an administrative process, which is not tailored to the scientists' needs, so that the definition of income, income unit, etc., are not necessarily those that we would have chosen.
-This causes particular difficulties for comparisons across countries, but also for time-series analysis where there have been substantial changes in the tax system, such as the moves to and from the joint taxation of couples.
-Secondly, it is obvious that those paying tax have a financial incentive to present their affairs in a way that reduces tax liabilities.
-There is tax avoidance and tax evasion.
-The rich, in particular, have a strong incentive to understate their taxable incomes.
-Those with wealth take steps to ensure that the return comes in the form of asset appreciation, typically taxed at lower rates or not at all.
-Those with high salaries seek to ensure that part of their remuneration comes in forms, such as fringe benefits or stock-options which receive favorable tax treatment.
-Both groups may make use of tax havens that allow income to be moved beyond the reach of the national tax net.
-These shortcomings limit what can be said from tax data, but this does not mean that the data are worthless.
-Like all economic data, they measure with error the 'true' variable in which we are interested.
-%The data series presented here are fairly homogenous across countries, annual, long-run, and broken down by income source for several cases.
-Users should be aware also about their limitations.
-Firstly, the series measure only top income shares and hence are silent on how inequality evolves elsewhere in the distribution [why?].
-Secondly, the series are largely concerned with gross incomes before tax.
-Thirdly, the definition of income and the unit of observation (the individual vs.
-the family) vary across countries making comparability of levels across countries more difficult.
-Even within a country, there are breaks in comparability that arise because of changes in tax legislation affecting the definition of income, although most studies try to correct for such changes to create homogenous series.
-Finally and perhaps most important, the series might be biased because of tax avoidance and tax evasion.
-For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).
-
-%Copied from:
-%Alvaredo, Facundo, Anthony B. Atkinson, Thomas Piketty and Emmanuel Saez, The World Top Incomes Database, http://topincomes.g-mond.parisschoolofeconomics.eu/ , 18/12/2013
-
-\textbf{Summary of tax data drawbacks}
-\begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial “income” see above)
-\item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the “top-income” literature, albeit it is a crucial issue in “theorized” concepts of income.
-To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
-This might influence the overall inequality, taking into account the change from traditional household and family structures over the last century.
-This is an inequality related issue, where relevant studies are missing.
-What are the assumptions?
-\item the tax data income definition summarize labor income, capital income and taxable transfer payments (no means-tested benefits as welfare aid).
-Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
-\item Some tax units are not included in the tax tables (see Foellmi and Martinez 2013:11f).(a) only taxed cases are included (partly a problem, for certain periodes we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
-Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
-In some border-cantons the share of this group is very high (20\%)
-%Can we ignore this)
-(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registred.
-This persons get an imputed income (older tax tax return and information given by employers.
-Not registered non-	fillers are not in the records.
-- (e) tax evasion.
-Feld and Frey (2007) report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
-With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequaly measures over time.
-\end{itemize}
-
-\textbf{Problems with household income surveys}
-\begin{itemize}
-\item Sample data (bias)
-\item comparability between countries and over time (depends on income definition)
-\item short time series
-\end{itemize}
-
-Atkinson et al. (2009) estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by Alvaredo (2010). \\
-
-See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
-%	•	What about the special cases?
-Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
-Exclusion leads to an underestimation of inequality
-%	•	Capital gains? Should not be included 
-The discussion about problems with reporting income is fairly exhaustive. What about wealth?
-
-\subsection{Tax data published by the Swiss Federal Tax Administration}
-
-%Beschreiben was die Daten beinhalten und wo es Probleme gibt
-\begin{itemize}
-\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
-	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
-	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
-	\item	Income after deductions-> taxable income (employment income, business income and capital income).
-Is income definition stable over time?
-Yes it should.
-Realized capital gains are excluded from the definition.
-It includes income from  employment, self-employment, capital income and taxable transfer payments.
-Plus Eigenmietwert,
-	\item	Reported on national and cantonal level
-\end{itemize}
-
-\subsection{Ways to tackle FTA-tax data specific problems}
-
-\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
-Del et al.
-(2007) impute for non-fillers the 20 percentage of the annual average income.
-This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
-Furthermore Del et al.
-calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
-\\
-
-\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
-This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
-This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f).
-\\
-%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.
-
-\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
-
-\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
-%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
-
-\textbf{deductions} Del et al.
-(2007:477):” we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.”
-Gorgas and Schaltenegger (2011:5): “”..., information on […] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
-Können wir das auch?
-Zumindest für gewisse Zeiträume?
-Das wäre noch gut.
-\\
-
-Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
-Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.
-On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a “perverse redistribution” effect by redistributing income from the lower middle class to the upper middle class (vgl.
-Peters 2011 and Interpellation Barbara Gysel (2009).
-
-% Income share specific problems
-% Total income denominator Exogenous Approach -> net income reported in the national accounts.
-Endogenous Approach -> Dell et al.
-technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income).
-Honestly , is this appropriate?
-(Everyone did it)
-% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
-% Endogenous approach -> reported in the tax tables
-
-%%%-------------------------------------------------%%%
-%%% Abschnitte aus ersten Version des Papers %%%
-%%%-------------------------------------------------%%%
-
-%\subsection{Hypotheses}
-%Based on the theories we test the following hypotheses:
-
-%\begin{itemize}
-%\item H1: Develpment of inequality is driven by sectoral change
-%\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
-%\end{itemize}
-
-%\subsection{Data and Variables}
-
-%We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010.
-While the data results in a long and consistent time series to illustrate swiss inequality development, there are a few pitfalls we want to adress which might be of interest for other research on this topic (be it in Switzerland or other countries).
-
-%\subsubsection{Left censored data}
-
-%The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household.
-The taxable population however is not identical to the population which should be used to calculate measures of inequality.
-Precisely, the data do not contain tax units with very little incomes so calculations based on these data treat the lowest percentiles equally to tax units with zero income.Figure X shows the threshold to be hit to enter the statistic.
-
-%\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}
-
-%So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data.
-Furthermore, also the changes over time might not be interpreted savely as over time the number of tax unit within this "hidden range" might vary or might even have a certain trend.
-We will adress this issue in detail in the methods chapter.
-
-%\subsubsection{Different measures, different populations}
-
-%The FTA data makes two kinds of distinctions.
-First, data was collected for so called "normal cases" and "special cases", i.e.
-a "normal case" is a taxable (for the complete tax period) person or household domiciled in a swiss canton without income from outside of Switzerland.
-A "special case" therefore is a diffuse reference category that contains tax units that are taxed at source, were not taxable for the complete tax period or generated additional income in another country.
-Second, the FTA reports two measures, that is taxable income and absolute income \textcolor{red}{(meine vorläufige Übersetzung von Reineinkommen)}.
-Absolute income is the sum of all incomes (earnings, interest income, rental incomes) minus expenses (e.g.
-from self-employment or credit cost).
-The taxable income is calculated as the difference of absolute income and deductions (e.g.
-children, insurance rates).
-The longest consistent time series exists for the taxable income of normal cases.
-So all statements we make with our data only apply to this subpopulation.
-
-%\subsubsection{Changes in taxation and measurement}
-%The swiss tax system is highly federal.
-That means, communities raise taxes which then go to the communities, the canton and the state.
-If we want to calculate overall swiss measures, we need to take into account, that cantons vary (between cantons and over time) with regard to the tax deductions that are possible and also the mechanism how taxes are collected.
-The latter adresses a comprehensive reorganization of the swiss tax system where between 1995 and 2003 cantons changed from taxing the past two years of income (postnumerando system) to taxing the present single year (praenumerando system).
-For details see \textcolor{red}{Martinez (xxxx) or some other author (xxxx)}.
-Aggregate measures of inequality therefore have to be estimated for the periods 1995 to 2003 which we adress shortly in the methods chapter.
-
-%\subsection{Methodology Used}
-%There are two steps of data analyses which need to be described to the reader.
-First, the estimation of the bias we introduce by estimating measures of inequality when tax units with too little income are not observed.
-Second, the steps undertaken to estimate aggregate swiss measures by imputing taxable income for those cantons and periods where the change of the tax system produced a gap (1995 to 2003, depending on the canton).
-
-%\subsubsection{Imputing the gap}
-%The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx).
-That is estimating the missing taxable income statistics via OLS using information from time trends and cantons.
-Our imputation model therefore includes canton inequality measures and periods dummies to explain aggregate swiss inequality.
-
-%\subsubsection{Estimating the bias}
-%For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation.
-However starting 1995, the FTA provides exactly this information for each canton.
-This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010.
-Consequently we can obtain information whether the bias is stable over time (which maes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
-\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie "je höher der Steuerfreibetrag, umso stärker der Bias".
-Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias.
-Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
-
-%%%-------------------------------------------------%%%
-%%% Include results %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document results %%%
-%%%-------------------------------------------------%%%
-
-\section{Results}
-
-\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
-Kantonsweise Grafiken
-Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
-
-%Switzerland (income and wealth)
-%- graph with gini over time> identify periods of change
-%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
-%Cantonal level (income and wealth)
-%- different development on regional level 
-%Switzerland in international comparison
-%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality
-
-%- Comparison with other countries (similar data)
-
-% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen.
-Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel.
-Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.
-
-%%%-------------------------------------------------%%%
-%%% Include discussion %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for discussion %%%
-%%%-------------------------------------------------%%%
-
-\section{Discussion}
-
-% Hier könnte allenfalls der Bogen zur Theorie gespannt werden
-
-%Methodological Main findings 
-%(1) Tax data is useful because it’s possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
-It’s superior to survey data, because the latter is affected by non-responded bias.
-Tax data includes data about the whole population … [was finden wir wegen den Nullern raus] .
-or at least it should (tax evasion).(3) The FTA-Tax data has different shortcomings, which cannot be handled like a measure of income and wealth which (a) cannot be corrected for household size and (b) is neither a pre-transfer nor a post-transfer income measure.
-Therefor it’s comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
-This FTA-Tax data problem can be handled with non-aggregated tax-data .
-
-%Main findings for Switzerland (1) Our data suggests that income inequality overall slightly increased in Switzerland and we might distinguish three episodes: 1.
-Substantial increase in times of economic growth 1950 to 1974 ending with the oil crisis 2.
-Ups and downs from 1970 to 2000 3.
-Relatively steep increase in inequality in the early 2000s
-%(2) Despite the recent increase, overall income inequality in Switzerland is not very high in international comparison.
-With respect to inequality in wealth, however, Switzerland takes a leading position in the world.
-%(3) While the overall Swiss inequality remained pretty stable over decades, inequality between cantons underwent a bizarre development.
-
-%%%-------------------------------------------------%%%
-%%% Include acknowledgements %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for acknowledgement %%%
-%%%-------------------------------------------------%%%
-
-\section{Acknowledgements}
-
-We thank Ben Jann, Robert Fluder and Tobias Fritschi for helpful comments on the article. We also like to thank Stefan Ilic for the prepartion of the data set.
-
-%%%-------------------------------------------------%%%
-%%% Include the bibliography %%%
-%%%-------------------------------------------------%%%
-
-\end{multicols}
-
-\bibliography{bibliography/bib}
-
-%%%-------------------------------------------------%%%
-%%% Include the appendix %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for appendix %%%
-%%%-------------------------------------------------%%%
-
-\section{Appendix}
-
-\end{document}

### 2014-01-10 paper/paper.tex
first version part of introduction and new bib library

@@
 There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
-In Switzerland, however, the development is unclear, as there is only little systematic evidence that would allow for long-term comparison.
+In Switzerland, however, the development is unclear, as there is only little systematic evidence about income inequality that would allow for long-term comparison.
+Concering the distriubtion of wealth, there is even less information available.
 Nevertheless income inequality has been a prominent theme in the public discussion in recent years, e.g.
@@
 
-% 1.Abschnitt: Bedeutung von Ungleichheit von Soziologischer Sicht
-
-\emph{Bedeutung von ökonomischer Ungleicheit} \\
-
 % 2.Abschnitt Trends in inequality research
@@
 
-\begin{itemize}
-\item  LIS (1982-2004, Survey Data from Swiss Federal Statistical Office three different surveys (recent years from “income and Consumption Survey, main result; decreasing inequality)
-\item  OECD (2000 und 2004, Income and consumption survey, 2008, EU-SILC. The change of survey in 2008 is consdered as a strict break. Comparison over time not possible (OECD 2012:315)
-\item  Worldbank: Nothing
-\item  Federal Office of statistics (1998-2011, HABE, survey data, very detailed measurement, main result; inequality in post transfer income is constant)
-\item  World income inequality Database (WIID) is more or less the same data as it is published in the LIS data
-\end{itemize}
-\emph{Fazit}:LIS longest time series, comparability is questionable, more recent years time series from Federal Office of statistics.
-Both studies use survey data, which, as we will discuss later, have certain drawbacks.\\
+Accoring to EU-SILC,Eurostat (2013), Switzerland took in 2011 a place in the middle (right in between Germany and France) looking at inequality of disposable income by countrys in the euro ares.
+Nevertheless Switzerland experienced in recent years a ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+The debate went allong with referendums trying the regulate the market outcomes.
+Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprings high majority of 68\%, the 1:12-Initiative, which aimed at the hole spectrum of the income distribution, was rejected.
+The voting about the minimum wage-initiative will be held on 9th february 2014.
+All this referendums questioned the actual distribution of income and were accompanied by a broader discussion about income inequality held in the media.
+Several offical and semi-offical publications adressed the question of rising inequality in Switzerland.
+\\
 
-\emph{Studies on Switzerland} \\
-Vgl. Forschungsantrag (kürzen) plus neuere Studien, Verteilungsbericht 2011 und 2012 (SGB), Avenir Suisse (2013) und Gorgas and Schaltenegger (2011) und Foellmi and Martinez (2013). \\
+The most recent offical numbers published by the Federal Statistical Office (2013) show  stability concering the income distriubtion covering the time periode of 1998 until 2011.
+The authors found a slight increase in the primary income which endured until the recent crisis.
+This temporarly increase was mainly compensated by govermental redistribution.
+A longer periode is covered in LIS-Dataset (1982-2004).\footnote{Dataprovider for the LIS Data is the Swiss Federal Statistical Office.
+The data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+The OCED-Database includes measures from Income and consumtion survey aswell.
+Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
+This change in survey is considered as a strict break.
+Comparision before and since 2008 is not recommended (OECD 2012:315).
+All in all the LIS Dataset contains the longest time series on inequality for Switzerland, which is public accesible.} Analysing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictionary to the development in most other western countries.
+This result is support by Grabka and Kuhn (2012) even when looking to more recent years (2000-2009).
+Analysing the Swiss Household Panel (SHP) they found a slight decrease in inequality.
+\\
+
+Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitfull insights for Switzerland aswell.
+Dell et al.
+(2007) used tax data from the Federal Tax Administrion to asset the development of concentration of the highest incomes and wealth (topshares).
+In contrast to moste all other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War peridoe to the decades following the second World War (up to 1996).
+Using the same approach Foellmi and Martinez (2013) expand the Dell et al.
+timeline to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
+A result which opposes the outcome of offical data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periodes of Swiss history but not the more recent years.
+Flückiger et al.
+(2007) and also Jeitziner and Peters (2007,2009) report constant inequality from 1960-1996 respectively from 1995 to 2003.
+Covering a similar time periode.
+Bauer and Spycher (1994) and Bolzani and Abul Naga (2002) found decreasing inequality.
+On the other hand Buchmann and Sacchi (1995) and Ernst et al.
+(2009) found an increase in the 1980s).} \\
+
+The divergence can be explained with several factors.
+First of all, different data sources were used.
+The offical data providers trust on survey data, whereas the later mentioned publications use tax data.
+It is assumend that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concering inequality.
+On the other hand the focus on top income neglects other changes in the distribution income as it is not possible to see, wheter newer concerns like the "hollowing of the middle class'' occured in Switzerland or not, which leads to the second point.
+Different measure of inequality hardens the comparibility.
+Third, different income concepts and different unit of analysis were used.
+As it is shown by Modetta and Müller (2012) income distribution is strongly affected by govermental redistriubtion, reducing inequality substantialy.
+With the focus on tax data change in institutional settings is not covered.
+Also neglected is the household structure, whereas it is unclear how inequality is affected whetever one looks at houshold income or at income of tax units.
+It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
+\\
+
+While there are several studies about income inequality in Switzerland, the search for relevant information regarding to the distribution of wealth is of no avail, albeit the distribution of wealth has to be considered if one is interessed in economic inequality.
+The cross-national Data center in Luxembourg is the first cross-national wealth database.
+But, up to the day Information for Switzerland is not avaible.
+On regular base \\
+
+% Erbschaftssteuer erwähnen und bei Global Wealth Report der Credit Suisse weitermachen
+
+% Allenfalls hier noch nicht so mit vor und Nachteilen von Steuerdaten kommen.
+
+Up to the day, Switzerland can be situated according to the actual level of income inequality, however, a long and consistent timeseries allowing to identify and explore development patterns on the whole distribution is missing.
+Also missing is a time series for inequality of wealth.
 
@@
 The purpose of the following paper is threefold.
-First of all, information about economic inequality in Switzerland is sparse.
+First of all, as shown information about economic inequality in Switzerland is sparse.
 We  fill this gap with official data from the Swiss Federal Tax Administration (FTA).
@@
 To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure.
-Based on a review on the literature about the measurment concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namly survey data.
-Afterwards we discribe the tax data published by the Swiss Federal Tax Administration highligthing the important aspects, which have to be considered working wiht FTA-Data.
-In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for switzerland.
+Based on a review on the literature about the measurement concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namely survey data.
+Afterwards we describe the tax data published by the Swiss Federal Tax Administration highlighting the important aspects, which have to be considered working with FTA-Data.
+In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for Switzerland.
 
@@
 \emph{Concepts on measuring economic resources}  \\
-Economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
+Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
 

### 2014-01-10 paper/subdocuments/data_and_methods.Rnw
first version part of introduction and new bib library

@@
 
-\section{Hypotheses, Methods and Data}
+\section{Measurement concepts, Data and Methods}
 
-\subsection{Hypotheses}
-Based on the theories we test the following hypotheses:
+To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure.
+Based on a review on the literature about the measurement concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namely survey data.
+Afterwards we describe the tax data published by the Swiss Federal Tax Administration highlighting the important aspects, which have to be considered working with FTA-Data.
+In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for Switzerland.
+
+\subsection{Standards for measuring economic resources and inequality}
+
+\emph{Concepts on measuring economic resources}  \\
+Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
+
+\emph{Defining income}
 
 \begin{itemize}
-\item H1: Develpment of inequality is driven by sectoral change
-\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
+\item Market income refers to revenue from employment.
+\item primary income or pre-transfer income refers to  market income plus revenue from property (capital income)
+\item Brutto income refers to  primary income plus social transfer
+\item Disposable income or posttransfer income refers to  brutto income minus transfer paid 
+\item adjusted disposable income refers to  disposable income corrected for the numbers of household members 
 \end{itemize}
 
-\subsection{Data and Variables}
+The appropriateness of the income concept is bound to the research interest.
+Studies focusing on labor market outcome look at market income.
+Studies focusing on human wellbeing follow the approach as propagated by the oecd (2013).
+This concept is largely implemented by the Federal Office of statistics.
+Following this concept one should look at income after taxes and transfers inclusive the adjustment for household members, because this is the measure which frames the consumption possibilities.
+Lastly, studies focusing on the effect of institutions (social welfare, taxation) compare the distribution of pre- and post-income measures.
+\\
 
-We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010.
+\emph{Population Coverage} \\
+All Residents? Working Residents? Single Person vs household/family. Households without income \\
+see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\
+
+\emph{Measuring inequality and concentration} \\
+To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
+Vgl.
+Allison (1978), Engelhardt (2000), Cowell (2000) oder Hao/Naiman (2010)
+%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
+
+neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences (Handcock/Morris 1999) \\
+
+Gini is mainly used for international comparison.
+However, other measures are often reported along (gini is more sensitive to changes in the middle of the distribution than to changes in the tails (OECD, 2008:37).
+Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) (Alderson and Doran 2013).
+\\
+
+Alvaredo (2010) shows formally, how top incomes shares are related to the gini-coefficient.
+Following his argument, it is crucial to consider top incomes.
+Leigh (2007) also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the studie).
+%[He also discusses tax data studies in general-> have a closer look] \\
+%[We definitely have to argue, why it isn’t sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
+
+\subsection{Comparison of tax data and other data sources - advantages and shortcomings}
+
+%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
+% Mögliche Vergleichsdimensionen
+%- Suitability of cross country comparison
+%- Assessing inequality accurately
+%- Implementation of common concepts of economic ressources
+%- contingency for intertemporal comparision
+
+%Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
+
+%Warnings about the use of tax data
+%The use of tax data is often regarded by economists with considerable disbelief.
+These doubts are well justified for at least two reasons.
+The first is that tax data are collected as part of an administrative process, which is not tailored to the scientists' needs, so that the definition of income, income unit, etc., are not necessarily those that we would have chosen.
+This causes particular difficulties for comparisons across countries, but also for time-series analysis where there have been substantial changes in the tax system, such as the moves to and from the joint taxation of couples.
+Secondly, it is obvious that those paying tax have a financial incentive to present their affairs in a way that reduces tax liabilities.
+There is tax avoidance and tax evasion.
+The rich, in particular, have a strong incentive to understate their taxable incomes.
+Those with wealth take steps to ensure that the return comes in the form of asset appreciation, typically taxed at lower rates or not at all.
+Those with high salaries seek to ensure that part of their remuneration comes in forms, such as fringe benefits or stock-options which receive favorable tax treatment.
+Both groups may make use of tax havens that allow income to be moved beyond the reach of the national tax net.
+These shortcomings limit what can be said from tax data, but this does not mean that the data are worthless.
+Like all economic data, they measure with error the 'true' variable in which we are interested.
+%The data series presented here are fairly homogenous across countries, annual, long-run, and broken down by income source for several cases.
+Users should be aware also about their limitations.
+Firstly, the series measure only top income shares and hence are silent on how inequality evolves elsewhere in the distribution [why?].
+Secondly, the series are largely concerned with gross incomes before tax.
+Thirdly, the definition of income and the unit of observation (the individual vs.
+the family) vary across countries making comparability of levels across countries more difficult.
+Even within a country, there are breaks in comparability that arise because of changes in tax legislation affecting the definition of income, although most studies try to correct for such changes to create homogenous series.
+Finally and perhaps most important, the series might be biased because of tax avoidance and tax evasion.
+For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).
+
+%Copied from:
+%Alvaredo, Facundo, Anthony B. Atkinson, Thomas Piketty and Emmanuel Saez, The World Top Incomes Database, http://topincomes.g-mond.parisschoolofeconomics.eu/ , 18/12/2013
+
+\textbf{Summary of tax data drawbacks}
+\begin{itemize}
+\item Misreporting of incomes (high earners have an interest in getting tax beneficial “income” see above)
+\item with aggregated data the adjustment for household members is not possible (no equivalization possible).
+This is a point, which is hardly mentioned in the “top-income” literature, albeit it is a crucial issue in “theorized” concepts of income.
+To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
+This might influence the overall inequality, taking into account the change from traditional household and family structures over the last century.
+This is an inequality related issue, where relevant studies are missing.
+What are the assumptions?
+\item the tax data income definition summarize labor income, capital income and taxable transfer payments (no means-tested benefits as welfare aid).
+Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
+\item Some tax units are not included in the tax tables (see Foellmi and Martinez 2013:11f).(a) only taxed cases are included (partly a problem, for certain periodes we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
+Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
+In some border-cantons the share of this group is very high (20\%)
+%Can we ignore this)
+(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registred.
+This persons get an imputed income (older tax tax return and information given by employers.
+Not registered non-	fillers are not in the records.
+- (e) tax evasion.
+Feld and Frey (2007) report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
+With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequaly measures over time.
+\end{itemize}
+
+\textbf{Problems with household income surveys}
+\begin{itemize}
+\item Sample data (bias)
+\item comparability between countries and over time (depends on income definition)
+\item short time series
+\end{itemize}
+
+Atkinson et al. (2009) estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by Alvaredo (2010). \\
+
+See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
+%	•	What about the special cases?
+Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
+Exclusion leads to an underestimation of inequality
+%	•	Capital gains? Should not be included 
+The discussion about problems with reporting income is fairly exhaustive. What about wealth?
+
+\subsection{Tax data published by the Swiss Federal Tax Administration}
+
+%Beschreiben was die Daten beinhalten und wo es Probleme gibt
+\begin{itemize}
+\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
+	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
+	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
+	\item	Income after deductions-> taxable income (employment income, business income and capital income).
+Is income definition stable over time?
+Yes it should.
+Realized capital gains are excluded from the definition.
+It includes income from  employment, self-employment, capital income and taxable transfer payments.
+Plus Eigenmietwert,
+	\item	Reported on national and cantonal level
+\end{itemize}
+
+\subsection{Ways to tackle FTA-tax data specific problems}
+
+\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
+Del et al.
+(2007) impute for non-fillers the 20 percentage of the annual average income.
+This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
+Furthermore Del et al.
+calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
+\\
+
+\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
+This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
+This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f).
+\\
+%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.
+
+\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
+
+\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
+%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
+
+\textbf{deductions} Del et al.
+(2007:477):” we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.”
+Gorgas and Schaltenegger (2011:5): “”..., information on […] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
+Können wir das auch?
+Zumindest für gewisse Zeiträume?
+Das wäre noch gut.
+\\
+
+Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
+Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.
+On the other hand deductions can affect the distribution.
+There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a “perverse redistribution” effect by redistributing income from the lower middle class to the upper middle class (vgl.
+Peters 2011 and Interpellation Barbara Gysel (2009).
+
+% Income share specific problems
+% Total income denominator Exogenous Approach -> net income reported in the national accounts.
+Endogenous Approach -> Dell et al.
+technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income).
+Honestly , is this appropriate?
+(Everyone did it)
+% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
+% Endogenous approach -> reported in the tax tables
+
+%%%-------------------------------------------------%%%
+%%% Abschnitte aus ersten Version des Papers %%%
+%%%-------------------------------------------------%%%
+
+%\subsection{Hypotheses}
+%Based on the theories we test the following hypotheses:
+
+%\begin{itemize}
+%\item H1: Develpment of inequality is driven by sectoral change
+%\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
+%\end{itemize}
+
+%\subsection{Data and Variables}
+
+%We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010.
 While the data results in a long and consistent time series to illustrate swiss inequality development, there are a few pitfalls we want to adress which might be of interest for other research on this topic (be it in Switzerland or other countries).
 
-\subsubsection{Left censored data}
+%\subsubsection{Left censored data}
 
-The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household.
+%The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household.
 The taxable population however is not identical to the population which should be used to calculate measures of inequality.
@@
 
-\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}
+%\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}
 
-So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data.
+%So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data.
 Furthermore, also the changes over time might not be interpreted savely as over time the number of tax unit within this "hidden range" might vary or might even have a certain trend.
@@
 
-\subsubsection{Different measures, different populations}
+%\subsubsection{Different measures, different populations}
 
-The FTA data makes two kinds of distinctions.
+%The FTA data makes two kinds of distinctions.
 First, data was collected for so called "normal cases" and "special cases", i.e.
@@
 
-\subsubsection{Changes in taxation and measurement}
-The swiss tax system is highly federal.
+%\subsubsection{Changes in taxation and measurement}
+%The swiss tax system is highly federal.
 That means, communities raise taxes which then go to the communities, the canton and the state.
@@
 
-\subsection{Methodology Used}
-There are two steps of data analyses which need to be described to the reader.
+%\subsection{Methodology Used}
+%There are two steps of data analyses which need to be described to the reader.
 First, the estimation of the bias we introduce by estimating measures of inequality when tax units with too little income are not observed.
@@
 
-\subsubsection{Imputing the gap}
-The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx).
+%\subsubsection{Imputing the gap}
+%The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx).
 That is estimating the missing taxable income statistics via OLS using information from time trends and cantons.
@@
 
-\subsubsection{Estimating the bias}
-For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation.
+%\subsubsection{Estimating the bias}
+%For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation.
 However starting 1995, the FTA provides exactly this information for each canton.

### 2014-01-10 paper/subdocuments/introduction.Rnw
first version part of introduction and new bib library

@@
 
+% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern.
+
+% 2.Abschnitt Trends in inequality research
+\emph{Studien zur allgemeinen Entwicklung der Ungleichheit nach unterschiedliche Paradigmen}(Armut, Reiche, Mittelschicht)  [Concerns about the hollowing of the middle class is a argument why the tax studies should be exploited to gain knowledge about the whole distribution of economic resources.
+Until these days tax data is only used to assess the change in top income shares.] \\
+\begin{itemize}
+  \item Development of inequality with LIS Data (Gornick and Jäntti 2013)
+  \item ranking of inequality across high-income countries (Atkinson, Rainwater and Smeeding (1995)
+   \item inequality and low income/relative poverty (Ferreira and Ravallion, 2009; Nolan and Marx, 2009; Smeeding, O’Higgins and Rainwatter, 1990; Rainwater and Smeeding 2003; Förster and Vlemickx, 2004)
+  \item inequality and top income (Atkinson and Piketty, 2007; Leigh, 2009)
+  \item inequality and the middle class (Estache and Leipziger, 2009; Littrell et al., 2001; Birdsall, 2010; Ravallion, 2010; Frank,2007)
+\end{itemize}
+
+% 3.Abschnitt: Studien mit Steuerdaten
+
+\emph{Studien mit Steuerdaten}
+\begin{itemize}
+\item Old studies with tax data (Vilfredo Pareto, 1895, 1896; Simon Kuznets, 1953)
+\item More recently revival of tax-studys with focus on top income (Piketty and Saez, 2003; Atkinson and Piketty, 2007,2010).
+\end{itemize}
+
+% 4.Abschnitt: What do we know about Switzerland?
+
+\emph{Official Data Collection} \\
+
+Accoring to EU-SILC,Eurostat (2013), Switzerland took in 2011 a place in the middle (right in between Germany and France) looking at inequality of disposable income by countrys in the euro ares.
+Nevertheless Switzerland experienced in recent years a ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+The debate went allong with referendums trying the regulate the market outcomes.
+Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprings high majority of 68\%, the 1:12-Initiative, which aimed at the hole spectrum of the income distribution, was rejected.
+The voting about the minimum wage-initiative will be held on 9th february 2014.
+All this referendums questioned the actual distribution of income and were accompanied by a broader discussion about income inequality held in the media.
+Several offical and semi-offical publications adressed the question of rising inequality in Switzerland.
+\\
+
+The most recent offical numbers published by the Federal Statistical Office (2013) show  stability concering the income distriubtion covering the time periode of 1998 until 2011.
+The authors found a slight increase in the primary income which endured until the recent crisis.
+This temporarly increase was mainly compensated by govermental redistribution.
+A longer periode is covered in LIS-Dataset (1982-2004).\footnote{Dataprovider for the LIS Data is the Swiss Federal Statistical Office.
+The data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+The OCED-Database includes measures from Income and consumtion survey aswell.
+Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
+This change in survey is considered as a strict break.
+Comparision before and since 2008 is not recommended (OECD 2012:315).
+All in all the LIS Dataset contains the longest time series on inequality for Switzerland, which is public accesible.} Analysing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictionary to the development in most other western countries.
+This result is support by Grabka and Kuhn (2012) even when looking to more recent years (2000-2009).
+Analysing the Swiss Household Panel (SHP) they found a slight decrease in inequality.
+\\
+
+Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitfull insights for Switzerland aswell.
+Dell et al.
+(2007) used tax data from the Federal Tax Administrion to asset the development of concentration of the highest incomes and wealth (topshares).
+In contrast to moste all other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War peridoe to the decades following the second World War (up to 1996).
+Using the same approach Foellmi and Martinez (2013) expand the Dell et al.
+timeline to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
+A result which opposes the outcome of offical data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periodes of Swiss history but not the more recent years.
+Flückiger et al.
+(2007) and also Jeitziner and Peters (2007,2009) report constant inequality from 1960-1996 respectively from 1995 to 2003.
+Covering a similar time periode.
+Bauer and Spycher (1994) and Bolzani and Abul Naga (2002) found decreasing inequality.
+On the other hand Buchmann and Sacchi (1995) and Ernst et al.
+(2009) found an increase in the 1980s).} \\
+
+The divergence can be explained with several factors.
+First of all, different data sources were used.
+The offical data providers trust on survey data, whereas the later mentioned publications use tax data.
+It is assumend that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concering inequality.
+On the other hand the focus on top income neglects other changes in the distribution income as it is not possible to see, wheter newer concerns like the "hollowing of the middle class'' occured in Switzerland or not, which leads to the second point.
+Different measure of inequality hardens the comparibility.
+Third, different income concepts and different unit of analysis were used.
+As it is shown by Modetta and Müller (2012) income distribution is strongly affected by govermental redistriubtion, reducing inequality substantialy.
+With the focus on tax data change in institutional settings is not covered.
+Also neglected is the household structure, whereas it is unclear how inequality is affected whetever one looks at houshold income or at income of tax units.
+It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
+\\
+
+While there are several studies about income inequality in Switzerland, the search for relevant information regarding to the distribution of wealth is of no avail, albeit the distribution of wealth has to be considered if one is interessed in economic inequality.
+The cross-national Data center in Luxembourg is the first cross-national wealth database.
+But, up to the day Information for Switzerland is not avaible.
+On regular base \\
+
+% Erbschaftssteuer erwähnen und bei Global Wealth Report der Credit Suisse weitermachen
+
+% Allenfalls hier noch nicht so mit vor und Nachteilen von Steuerdaten kommen.
+
+Up to the day, Switzerland can be situated according to the actual level of income inequality, however, a long and consistent timeseries allowing to identify and explore development patterns on the whole distribution is missing.
+Also missing is a time series for inequality of wealth.
+
+% Should we discus the non-scientic literature?
+
+% What is the purpose of our paper?
 The purpose of the following paper is threefold.
-First of all, information about economic inequality in Switzerland is sparse.
-E.g.
-the well known Worldbank data do not contain any Gini-coefficient for Switzerland which is a rare exception within Europe.
-We try to fill this gap with official data from the Swiss Federal Tax Administration (FTA).
+First of all, as shown information about economic inequality in Switzerland is sparse.
+We  fill this gap with official data from the Swiss Federal Tax Administration (FTA).
 Second, we assess the suitability of the FTA data to report inequality and its changes over time.
-An important aspect here is that the data do not contain all income earning persons in Switzerland but are left censored as federal taxes are payable only above a certain threshold.
-We report magnitude and variance of this bias.
 Third, we integrate the development of swiss inequality into the international picture.

### 2014-01-12 paper/subdocuments/data_and_methods.Rnw
did first analyses with zeros (0er)

@@
 
-%\subsubsection{Estimating the bias}
-%For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation.
+\subsubsection{Estimating the bias}
+For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation (lets call those "zeros" for convenience) .
 However starting 1995, the FTA provides exactly this information for each canton.
 This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010.
-Consequently we can obtain information whether the bias is stable over time (which maes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
+Consequently we can obtain information whether the bias is stable over time (which makes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
 \textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie "je höher der Steuerfreibetrag, umso stärker der Bias".
@@
 Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
+
+One can take a first look at the descriptives plotting the share of zeros over time seperately for each canton .
+
+<<setup echo=FALSE>>=
+library(foreign)
+df <- read.dta("data/ginis_und_perzentile.dta")
+@
+
+<<zero_descriptives>>=
+library(ggplot2)
+df$zeros <- df$null_norm/df$cpop # i only use normal cases here
+ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)
+@
+
+We can see multiple things here:
+
+\begin{enumerate}
+\item There is a small overall upward trend which we assume to be the Federal Administrations inflation adjustments to the tax threshold.
+\item Geneva and Tessin show wild changes but those might be explained by the tax gap ("Bemessungslücke") that people exploited when the cantons changed the tax system.
+It remains unclear however why we can't see similar patter within other cantons.
+\item There is some variance and we see different patterns over time and cantons.
+When estimating gini coefficients or the like we must therefore assume that ignoring "the zeros" leads to a bias that is not stable over time.
+\end{enumerate}
+
+We might try two strategies to moderate the problem:
+
+\begin{enumerate}
+\item Add the zeros as a seperate group
+\item Fit a model to predict the inequality measure (e.g. the gini coefficient) using the share of zeros as a predictor
+\end{enumerate}
+
+By adding the zeros as a seperate group we face several problems.
+With the exception of Geneva 1995/96 Pareto interpolation of percentiles p20 and above seems viable.
+Low percentiles however would need to be extrapolated so we might impose unrealistic assuptioms.
+Even when estimating "safe" measures above p20 (better p50) or a gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
+To be consistent with the measure of "taxable income" we could assume an income of zero for that group, calculate gini coefficients and compare them with the original (uncorrected/plain) version.
+To get a rough number we could (cantonwise) check the squared correlation between plain and corrected gini.
+
+The second approach however is more robust as we do not impose additional assumptions but instead exercise a little curve fitting.
+
+<<correct_gini>>=
+fit <- lm(G~kanton+kanton:zeros+factor(steuerperiode)+factor(steuerperiode):zeros,data=df)
+summary(fit)
+@
+
+The model outputs a teststatisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
+As the model has a decent fit we are not in great danger of omitted variable bias.
+Using a joint F-Test we can now test if all canton#zeros interactions are zero.
+
+<<ftest>>=
+library(survey)
+regTermTest(fit,"kanton:zeros")
+@
+
+In our case we can clearly reject the hypothesis that all interactions are zero.
+This leads to the conclusion  that gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted gini coefficients.
+For example one might be interested in how inequality would had developed if the zero-share would have been constant over time.
+(Note RF: predict all data point using canton, time and the initial OR final zero-share to homogenize the time series)
+Furthermore we can quantify how large the bias is and we can do this seperately for tax periods or seperately for cantons.
+
+For all cantons:
+
+<<magnitude_of_bias>>=
+reduced_fit <- lm(G~kanton+factor(steuerperiode),data=df)
+summary(reduced_fit)
+@
+
+We can see the the model fit reduces to explaining 61.5\% of the gini variation versus 92.6\% when the information from the zero-shares was used.
+Although this so some extent attributable to the additional 26 paramters: this is huge.
+
+The model indicates some cases that deserve more attention: Schwyz (positive coefficient) and Geneva (negative coefficient) and the tax period 2000 as well as the most recent periods.
+
+A positive coefficient (e.g.
+Schwyz) can be read as follows: In periods with many zeros we measure higher gini coefficients.
+Our conclusion is that people with low income are more successfully hidden.
+If those people would be taxed and therefore appear in the statistic they would sharply contrast the high income earners.
+
+A negative coefficient (e.g.
+Geneva) means the more zeros there the are the smaller the gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
+This is less obvious than the Schwyz case and our reading is: more zeros mask inequality that arises from the bottom.

### 2014-01-14 paper/paper.tex
Several Updates to introduction and method section

--- 2014-01-07 paper/paper.tex
@@
 
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing inequality with tax data - Switzerland from 1945 to 2010}} % Article title
+\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}} % Article title
 
@@
 There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
-In Switzerland, however, the development is unclear, as there is only little systematic evidence that would allow for long-term comparison.
-Nevertheless income inequality has been a prominent theme in the public discussion in recent years, e.g.
+In Switzerland, however, the development is unclear, as there is only little systematic evidence about income inequality that would allow for long-term comparison.
+Even scarcer are Studies on the distribution of wealth as the second dimension of economic inequality.
+Nevertheless, income inequality has been a prominent theme in the public Swiss discussion in recent years, e.g.
 the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''.
-We adress the gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data.
+We address the mentioned gap by presenting a long and consistent time series of inequality measures for income and wealth (1943-2010) calculated from federal tax statistics.
 We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties.
@@
 
-% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern.
+% Im European Sociologica Review scheint es unüblich zu sein, ohne theoretische Sektion zu arbeiten.
+Ich hoffe nicht, dass dies ein k.o.
+Kriterium ist.
+wir halten dafür die Einführungs und die Methoden Sektion etwas allgemeiner
 
-% 1.Abschnitt: Bedeutung von Ungleichheit von Soziologischer Sicht
+% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern
 
-\emph{Bedeutung von ökonomischer Ungleicheit} \\
+% 1.Abschnitt: Bedeutung von Ungleichheit
 
-% 2.Abschnitt Trends in inequality research
-\emph{Studien zur allgemeinen Entwicklung der Ungleichheit nach unterschiedliche Paradigmen}(Armut, Reiche, Mittelschicht)  [Concerns about the hollowing of the middle class is a argument why the tax studies should be exploited to gain knowledge about the whole distribution of economic resources.
-Until these days tax data is only used to assess the change in top income shares.] \\
-\begin{itemize}
-  \item Development of inequality with LIS Data (Gornick and Jäntti 2013)
-  \item ranking of inequality across high-income countries (Atkinson, Rainwater and Smeeding (1995)
-   \item inequality and low income/relative poverty (Ferreira and Ravallion, 2009; Nolan and Marx, 2009; Smeeding, O’Higgins and Rainwatter, 1990; Rainwater and Smeeding 2003; Förster and Vlemickx, 2004)
-  \item inequality and top income (Atkinson and Piketty, 2007; Leigh, 2009)
-  \item inequality and the middle class (Estache and Leipziger, 2009; Littrell et al., 2001; Birdsall, 2010; Ravallion, 2010; Frank,2007)
-\end{itemize}
+Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular (Wilkinson and Pickett 2009).
+While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders (WEF 2013) and scholars alike.
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013).
+The pattern of evolution suggests a polarization, which can be referred to as the "hollowing of the the middle class'' (Alderson and Doran 2013).
+Households are moving toward the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril.''
+\\
 
-% 3.Abschnitt: Studien mit Steuerdaten
+% 2.Abschnitt: Bedeutung der Datengrundlage und Einführung von Steuerdaten
 
-\emph{Studien mit Steuerdaten}
-\begin{itemize}
-\item Old studies with tax data (Vilfredo Pareto, 1895, 1896; Simon Kuznets, 1953)
-\item More recently revival of tax-studys with focus on top income (Piketty and Saez, 2003; Atkinson and Piketty, 2007,2010).
-\end{itemize}
+Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
+Atkinson (2013:8) observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
+On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality unsdisputedly.
+These concerns have led to the search of alternative data sources, which can supplement the established survey data studies.
+Already Kuznet (1955) used tax data to examine the relationship between economic growth and personal distribution of income.
+Then it took several decades until Piketty (2001,2003) made the use of tax data fashionable again.
+Following his approach studies on several countries were conducted (Atkinson and Piketty 2007, 2010).
+Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database (Alvaredo et al.
+2014).
+\\
 
-% 4.Abschnitt: What do we know about Switzerland?
+% 3.Abschnitt: What do we know about Switzerland
+%
 
-\emph{Official Data Collection} \\
+What is know about Switzerland so far?
+Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
+Nevertheless in recent years Switzerland experienced  a ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+The debate went along with referendums trying the regulate the market outcomes.
+Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisings high majority of 68\%, the 1:12-Initiative, which aimed at the hole spectrum of the income distribution, was rejected.
+The voting about the minimum wage-initiative will be held on 9th February 2014.
+All this referendums questioned the actual distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
+Several official and semi-official publications addressed the question of rising inequality in Switzerland.
+\\
 
-\begin{itemize}
-\item  LIS (1982-2004, Survey Data from Swiss Federal Statistical Office three different surveys (recent years from “income and Consumption Survey, main result; decreasing inequality)
-\item  OECD (2000 und 2004, Income and consumption survey, 2008, EU-SILC. The change of survey in 2008 is consdered as a strict break. Comparison over time not possible (OECD 2012:315)
-\item  Worldbank: Nothing
-\item  Federal Office of statistics (1998-2011, HABE, survey data, very detailed measurement, main result; inequality in post transfer income is constant)
-\item  World income inequality Database (WIID) is more or less the same data as it is published in the LIS data
-\end{itemize}
-\emph{Fazit}:LIS longest time series, comparability is questionable, more recent years time series from Federal Office of statistics.
-Both studies use survey data, which, as we will discuss later, have certain drawbacks.\\
+% Lässt sich allenfalls streichen/kürzen
+% Ist es die Bezeichnung offical publications sinnvoll? Oder besser govermental publications
 
-\emph{Studies on Switzerland} \\
-Vgl. Forschungsantrag (kürzen) plus neuere Studien, Verteilungsbericht 2011 und 2012 (SGB), Avenir Suisse (2013) und Gorgas and Schaltenegger (2011) und Foellmi and Martinez (2013). \\
+% 4.Abschnitt: What do we know about Switzerland
+% Offical Data Collections on development of inequality
 
-% Should we discus the non-scientic literature?
+The most recent official figures published by the Federal Statistical Office (2013) show  stability concerning the income distribution covering the time period of 1998 until 2011.
+The authors found a slight increase in the primary income which endured until the recent crisis.
+This temporary increase was mainly compensated by governmental redistribution.
+A longer period is covered in the LIS-Data-set (1982-2004).\footnote{Data-provider for the LIS Data is the Swiss Federal Statistical Office.
+The data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+The OCED-Database includes measures from Income and consumption survey as well.
+Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
+This change in survey is considered as a strict break.
+Comparison before and since 2008 is not recommended (OECD 2012:315).
+All in all the LIS Data-set contains the longest time series on inequality for Switzerland.} Analyzing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
+This result is support by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).
+\\
 
-% What is the purpose of our paper?
-The purpose of the following paper is threefold.
-First of all, information about economic inequality in Switzerland is sparse.
-We  fill this gap with official data from the Swiss Federal Tax Administration (FTA).
-Second, we assess the suitability of the FTA data to report inequality and its changes over time.
-Third, we integrate the development of swiss inequality into the international picture.
-We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).
+% 5.Abschnitt: What do we know about Switzerland
+% Studies with tax data
+
+Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
+Dell et al.
+(2007) used tax data from the Federal Tax Administration to asset the development of concentration of the highest incomes and wealth (top-shares).
+In contrast to most all other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
+Using the same approach Foellmi and Martinez (2013) expand the Dell et al.
+timeline to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
+A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
+Flückiger et al.
+(2007) and also Jeitziner and Peters (2007,2009) report constant inequality from 1960-1996 respectively from 1995 to 2003.
+Covering a similar time period Bauer and Spycher (1994) and Bolzani and Abul Naga (2002) found decreasing inequality.
+On the other hand Buchmann and Sacchi (1995) and Ernst et al.
+(2009) found an increase in the 1980s).} \\
+
+% 6. Abschnitt: Zwischenbilanz und Herausforderungen
+
+The divergence can be explained with several factors.
+First of all, different data sources were used.
+The official data providers trust on survey data, whereas the later mentioned publications use tax data.
+It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality.
+On the other hand the focus on top income neglects other changes in the distribution income as it is not possible to see, whether newer concerns like the "hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
+Different measure of inequality hardens the comparability.
+Third, different income concepts and different unit of analysis were used.
+As it is shown by Modetta and Müller (2012) income distribution is strongly affected by governmental redistribution, reducing inequality substantial.
+With the focus on tax data the change in institutional settings is not covered.
+Also neglected is the household structure, whereas it is unclear how inequality is affected whether one looks at household income or at income of tax units.
+It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
+\\
+
+% Allenfalls hier noch nicht so ausführlich mit Vor- und Nachteilen von Steuerdaten kommen.
+Vielleicht besser darauf hinweisen, welche Dimensionen betroffen sind und das dies in der Methods Section besprochen wird.
+
+While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
+The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
+But, up to the day Information for Switzerland is not available.
+Shorrock et al.
+(2013) unify several datasources (including tax data) to assess the pattern of global wealth.
+Their databook includes figures for Switzerland.
+In contrast to inequality of income concerning inequalit of wealth Switzerland takes a leading position \\
+
+% the credit suisse wealth report allready uses wealth tax records, but they use it only for the period from 2000-2011.
+They did exactly the same for Denmark and Norway, which could be interessting to compare.
+% Erbschaftssteuer erwähnen und bei den Global Wealth Reports der Credit Suisse und den Studien im Forschungsantrag weitermachen.
+
+Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort collect data which can be harmonized to comparable measures (see Luxembourg income study).
+However, it is unclear how the bias through non-response affects the overall measure of inequality.
+Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
+Building on recent developments in the field of inequality research, we assess the suitability of the public accessible tax data to report inequality and its changes over time.
+First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
+Second, we summarize and apply tax data specific technics to construct suitable measures of inequality.
+We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
+Third, we integrate the development of Swiss inequality into the international picture.
+
+% Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.
 
@@
 \emph{Concepts on measuring economic resources}  \\
-Economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
+Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
+
+% We neither can examine consumption nor the jointly distribution of income and wealth on the individual level.
+We could analyse the jointly pattern on the cantonal level and analze the data following the question if income inequality and wealth inqueality are correlated.
 
@@
 \\
+
+\emph{Defining wealth}
+While the apropiateness of the conceptualization of income is widely discuded, there is no agreed definition of personal wealth yet and the appropriate methods of valuation are not always clear. 
+Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
 
@@
 The discussion about problems with reporting income is fairly exhaustive. What about wealth?
+
+It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
+Credit Suisse (2011) makes use of the information in the "Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
+\\
+
+Additonaly the Credit Suisse Reports states (2011:8) that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts.
 

### 2014-01-14 paper/subdocuments/abstract.Rnw
Several Updates to introduction and method section

@@
 In Switzerland, however, the development is unclear, as there is only little systematic evidence about income inequality that would allow for long-term comparison.
-Concering the distriubtion of wealth, there is even less information available.
-Nevertheless income inequality has been a prominent theme in the public discussion in recent years, e.g.
+Even scarcer are Studies on the distribution of wealth as the second dimension of economic inequality.
+Nevertheless, income inequality has been a prominent theme in the public Swiss discussion in recent years, e.g.
 the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''.
-We adress the gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data.
+We address the mentioned gap by presenting a long and consistent time series of inequality measures for income and wealth (1943-2010) calculated from federal tax statistics.
 We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties.

### 2014-01-14 paper/subdocuments/data_and_methods.Rnw
Several Updates to introduction and method section

--- 2014-01-10 paper/subdocuments/data_and_methods.Rnw
@@
 Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
+
+% We neither can examine consumption nor the jointly distribution of income and wealth on the individual level.
+We could analyse the jointly pattern on the cantonal level and analze the data following the question if income inequality and wealth inqueality are correlated.
 

### 2014-01-15 paper/paper.tex
Update

@@
 
-\documentclass[twoside]{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
+\documentclass[twoside]{article}
 
@@
 %----------------------------------------------------------------------------------------
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 
@@
 In Switzerland, however, the development is unclear, as there is only little systematic evidence about income inequality that would allow for long-term comparison.
-Even scarcer are Studies on the distribution of wealth as the second dimension of economic inequality.
+Studies on the distribution of wealth are even scarcer.
 Nevertheless, income inequality has been a prominent theme in the public Swiss discussion in recent years, e.g.
@@
 %%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for data and methods %%%
-%%%-------------------------------------------------%%%
-
-\section{Measurement concepts, Data and Methods}
-
-To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure.
-Based on a review on the literature about the measurement concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namely survey data.
-Afterwards we describe the tax data published by the Swiss Federal Tax Administration highlighting the important aspects, which have to be considered working with FTA-Data.
-In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for Switzerland.
-
-\subsection{Standards for measuring economic resources and inequality}
-
-\emph{Concepts on measuring economic resources}  \\
-Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
-
-% We neither can examine consumption nor the jointly distribution of income and wealth on the individual level.
-We could analyse the jointly pattern on the cantonal level and analze the data following the question if income inequality and wealth inqueality are correlated.
-
-\emph{Defining income}
-
-\begin{itemize}
-\item Market income refers to revenue from employment.
-\item primary income or pre-transfer income refers to  market income plus revenue from property (capital income)
-\item Brutto income refers to  primary income plus social transfer
-\item Disposable income or posttransfer income refers to  brutto income minus transfer paid 
-\item adjusted disposable income refers to  disposable income corrected for the numbers of household members 
-\end{itemize}
-
-The appropriateness of the income concept is bound to the research interest.
-Studies focusing on labor market outcome look at market income.
-Studies focusing on human wellbeing follow the approach as propagated by the oecd (2013).
-This concept is largely implemented by the Federal Office of statistics.
-Following this concept one should look at income after taxes and transfers inclusive the adjustment for household members, because this is the measure which frames the consumption possibilities.
-Lastly, studies focusing on the effect of institutions (social welfare, taxation) compare the distribution of pre- and post-income measures.
-\\
-
-\emph{Defining wealth}
-While the apropiateness of the conceptualization of income is widely discuded, there is no agreed definition of personal wealth yet and the appropriate methods of valuation are not always clear. 
-Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
-
-\emph{Population Coverage} \\
-All Residents? Working Residents? Single Person vs household/family. Households without income \\
-see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\
-
-\emph{Measuring inequality and concentration} \\
-To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
-Vgl.
-Allison (1978), Engelhardt (2000), Cowell (2000) oder Hao/Naiman (2010)
-%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
-
-neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences (Handcock/Morris 1999) \\
-
-Gini is mainly used for international comparison.
-However, other measures are often reported along (gini is more sensitive to changes in the middle of the distribution than to changes in the tails (OECD, 2008:37).
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) (Alderson and Doran 2013).
-\\
-
-Alvaredo (2010) shows formally, how top incomes shares are related to the gini-coefficient.
-Following his argument, it is crucial to consider top incomes.
-Leigh (2007) also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the studie).
-%[He also discusses tax data studies in general-> have a closer look] \\
-%[We definitely have to argue, why it isn’t sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
-
-\subsection{Comparison of tax data and other data sources - advantages and shortcomings}
-
-%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
-% Mögliche Vergleichsdimensionen
-%- Suitability of cross country comparison
-%- Assessing inequality accurately
-%- Implementation of common concepts of economic ressources
-%- contingency for intertemporal comparision
-
-%Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
-
-%Warnings about the use of tax data
-%The use of tax data is often regarded by economists with considerable disbelief.
-These doubts are well justified for at least two reasons.
-The first is that tax data are collected as part of an administrative process, which is not tailored to the scientists' needs, so that the definition of income, income unit, etc., are not necessarily those that we would have chosen.
-This causes particular difficulties for comparisons across countries, but also for time-series analysis where there have been substantial changes in the tax system, such as the moves to and from the joint taxation of couples.
-Secondly, it is obvious that those paying tax have a financial incentive to present their affairs in a way that reduces tax liabilities.
-There is tax avoidance and tax evasion.
-The rich, in particular, have a strong incentive to understate their taxable incomes.
-Those with wealth take steps to ensure that the return comes in the form of asset appreciation, typically taxed at lower rates or not at all.
-Those with high salaries seek to ensure that part of their remuneration comes in forms, such as fringe benefits or stock-options which receive favorable tax treatment.
-Both groups may make use of tax havens that allow income to be moved beyond the reach of the national tax net.
-These shortcomings limit what can be said from tax data, but this does not mean that the data are worthless.
-Like all economic data, they measure with error the 'true' variable in which we are interested.
-%The data series presented here are fairly homogenous across countries, annual, long-run, and broken down by income source for several cases.
-Users should be aware also about their limitations.
-Firstly, the series measure only top income shares and hence are silent on how inequality evolves elsewhere in the distribution [why?].
-Secondly, the series are largely concerned with gross incomes before tax.
-Thirdly, the definition of income and the unit of observation (the individual vs.
-the family) vary across countries making comparability of levels across countries more difficult.
-Even within a country, there are breaks in comparability that arise because of changes in tax legislation affecting the definition of income, although most studies try to correct for such changes to create homogenous series.
-Finally and perhaps most important, the series might be biased because of tax avoidance and tax evasion.
-For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).
-
-%Copied from:
-%Alvaredo, Facundo, Anthony B. Atkinson, Thomas Piketty and Emmanuel Saez, The World Top Incomes Database, http://topincomes.g-mond.parisschoolofeconomics.eu/ , 18/12/2013
-
-\textbf{Summary of tax data drawbacks}
-\begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial “income” see above)
-\item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the “top-income” literature, albeit it is a crucial issue in “theorized” concepts of income.
-To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
-This might influence the overall inequality, taking into account the change from traditional household and family structures over the last century.
-This is an inequality related issue, where relevant studies are missing.
-What are the assumptions?
-\item the tax data income definition summarize labor income, capital income and taxable transfer payments (no means-tested benefits as welfare aid).
-Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
-\item Some tax units are not included in the tax tables (see Foellmi and Martinez 2013:11f).(a) only taxed cases are included (partly a problem, for certain periodes we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
-Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
-In some border-cantons the share of this group is very high (20\%)
-%Can we ignore this)
-(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registred.
-This persons get an imputed income (older tax tax return and information given by employers.
-Not registered non-	fillers are not in the records.
-- (e) tax evasion.
-Feld and Frey (2007) report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
-With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequaly measures over time.
-\end{itemize}
-
-\textbf{Problems with household income surveys}
-\begin{itemize}
-\item Sample data (bias)
-\item comparability between countries and over time (depends on income definition)
-\item short time series
-\end{itemize}
-
-Atkinson et al. (2009) estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by Alvaredo (2010). \\
-
-See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
-%	•	What about the special cases?
-Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
-Exclusion leads to an underestimation of inequality
-%	•	Capital gains? Should not be included 
-The discussion about problems with reporting income is fairly exhaustive. What about wealth?
-
-It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
-Credit Suisse (2011) makes use of the information in the "Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
-\\
-
-Additonaly the Credit Suisse Reports states (2011:8) that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts.
-
-\subsection{Tax data published by the Swiss Federal Tax Administration}
-
-%Beschreiben was die Daten beinhalten und wo es Probleme gibt
-\begin{itemize}
-\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
-	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
-	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
-	\item	Income after deductions-> taxable income (employment income, business income and capital income).
-Is income definition stable over time?
-Yes it should.
-Realized capital gains are excluded from the definition.
-It includes income from  employment, self-employment, capital income and taxable transfer payments.
-Plus Eigenmietwert,
-	\item	Reported on national and cantonal level
-\end{itemize}
-
-\subsection{Ways to tackle FTA-tax data specific problems}
-
-\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
-Del et al.
-(2007) impute for non-fillers the 20 percentage of the annual average income.
-This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
-Furthermore Del et al.
-calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
-\\
-
-\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
-This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
-This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f).
-\\
-%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.
-
-\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
-
-\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
-%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
-
-\textbf{deductions} Del et al.
-(2007:477):” we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.”
-Gorgas and Schaltenegger (2011:5): “”..., information on […] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
-Können wir das auch?
-Zumindest für gewisse Zeiträume?
-Das wäre noch gut.
-\\
-
-Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
-Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.
-On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a “perverse redistribution” effect by redistributing income from the lower middle class to the upper middle class (vgl.
-Peters 2011 and Interpellation Barbara Gysel (2009).
-
-% Income share specific problems
-% Total income denominator Exogenous Approach -> net income reported in the national accounts.
-Endogenous Approach -> Dell et al.
-technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income).
-Honestly , is this appropriate?
-(Everyone did it)
-% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
-% Endogenous approach -> reported in the tax tables
-
-%%%-------------------------------------------------%%%
-%%% Abschnitte aus ersten Version des Papers %%%
-%%%-------------------------------------------------%%%
-
-%\subsection{Hypotheses}
-%Based on the theories we test the following hypotheses:
-
-%\begin{itemize}
-%\item H1: Develpment of inequality is driven by sectoral change
-%\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
-%\end{itemize}
-
-%\subsection{Data and Variables}
-
-%We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010.
-While the data results in a long and consistent time series to illustrate swiss inequality development, there are a few pitfalls we want to adress which might be of interest for other research on this topic (be it in Switzerland or other countries).
-
-%\subsubsection{Left censored data}
-
-%The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household.
-The taxable population however is not identical to the population which should be used to calculate measures of inequality.
-Precisely, the data do not contain tax units with very little incomes so calculations based on these data treat the lowest percentiles equally to tax units with zero income.Figure X shows the threshold to be hit to enter the statistic.
-
-%\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}
-
-%So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data.
-Furthermore, also the changes over time might not be interpreted savely as over time the number of tax unit within this "hidden range" might vary or might even have a certain trend.
-We will adress this issue in detail in the methods chapter.
-
-%\subsubsection{Different measures, different populations}
-
-%The FTA data makes two kinds of distinctions.
-First, data was collected for so called "normal cases" and "special cases", i.e.
-a "normal case" is a taxable (for the complete tax period) person or household domiciled in a swiss canton without income from outside of Switzerland.
-A "special case" therefore is a diffuse reference category that contains tax units that are taxed at source, were not taxable for the complete tax period or generated additional income in another country.
-Second, the FTA reports two measures, that is taxable income and absolute income \textcolor{red}{(meine vorläufige Übersetzung von Reineinkommen)}.
-Absolute income is the sum of all incomes (earnings, interest income, rental incomes) minus expenses (e.g.
-from self-employment or credit cost).
-The taxable income is calculated as the difference of absolute income and deductions (e.g.
-children, insurance rates).
-The longest consistent time series exists for the taxable income of normal cases.
-So all statements we make with our data only apply to this subpopulation.
-
-%\subsubsection{Changes in taxation and measurement}
-%The swiss tax system is highly federal.
-That means, communities raise taxes which then go to the communities, the canton and the state.
-If we want to calculate overall swiss measures, we need to take into account, that cantons vary (between cantons and over time) with regard to the tax deductions that are possible and also the mechanism how taxes are collected.
-The latter adresses a comprehensive reorganization of the swiss tax system where between 1995 and 2003 cantons changed from taxing the past two years of income (postnumerando system) to taxing the present single year (praenumerando system).
-For details see \textcolor{red}{Martinez (xxxx) or some other author (xxxx)}.
-Aggregate measures of inequality therefore have to be estimated for the periods 1995 to 2003 which we adress shortly in the methods chapter.
-
-%\subsection{Methodology Used}
-%There are two steps of data analyses which need to be described to the reader.
-First, the estimation of the bias we introduce by estimating measures of inequality when tax units with too little income are not observed.
-Second, the steps undertaken to estimate aggregate swiss measures by imputing taxable income for those cantons and periods where the change of the tax system produced a gap (1995 to 2003, depending on the canton).
-
-%\subsubsection{Imputing the gap}
-%The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx).
-That is estimating the missing taxable income statistics via OLS using information from time trends and cantons.
-Our imputation model therefore includes canton inequality measures and periods dummies to explain aggregate swiss inequality.
-
-%\subsubsection{Estimating the bias}
-%For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation.
-However starting 1995, the FTA provides exactly this information for each canton.
-This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010.
-Consequently we can obtain information whether the bias is stable over time (which maes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
-\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie "je höher der Steuerfreibetrag, umso stärker der Bias".
-Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias.
-Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
-
-%%%-------------------------------------------------%%%
-%%% Include results %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document results %%%
-%%%-------------------------------------------------%%%
-
-\section{Results}
-
-\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
-Kantonsweise Grafiken
-Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
-
-%Switzerland (income and wealth)
-%- graph with gini over time> identify periods of change
-%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
-%Cantonal level (income and wealth)
-%- different development on regional level 
-%Switzerland in international comparison
-%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality
-
-%- Comparison with other countries (similar data)
-
-% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen.
-Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel.
-Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.
-
-%%%-------------------------------------------------%%%
-%%% Include discussion %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for discussion %%%
-%%%-------------------------------------------------%%%
-
-\section{Discussion}
-
-% Hier könnte allenfalls der Bogen zur Theorie gespannt werden
-
-%Methodological Main findings 
-%(1) Tax data is useful because it’s possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
-It’s superior to survey data, because the latter is affected by non-responded bias.
-Tax data includes data about the whole population … [was finden wir wegen den Nullern raus] .
-or at least it should (tax evasion).(3) The FTA-Tax data has different shortcomings, which cannot be handled like a measure of income and wealth which (a) cannot be corrected for household size and (b) is neither a pre-transfer nor a post-transfer income measure.
-Therefor it’s comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
-This FTA-Tax data problem can be handled with non-aggregated tax-data .
-
-%Main findings for Switzerland (1) Our data suggests that income inequality overall slightly increased in Switzerland and we might distinguish three episodes: 1.
-Substantial increase in times of economic growth 1950 to 1974 ending with the oil crisis 2.
-Ups and downs from 1970 to 2000 3.
-Relatively steep increase in inequality in the early 2000s
-%(2) Despite the recent increase, overall income inequality in Switzerland is not very high in international comparison.
-With respect to inequality in wealth, however, Switzerland takes a leading position in the world.
-%(3) While the overall Swiss inequality remained pretty stable over decades, inequality between cantons underwent a bizarre development.
-
-%%%-------------------------------------------------%%%
-%%% Include acknowledgements %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for acknowledgement %%%
-%%%-------------------------------------------------%%%
-
-\section{Acknowledgements}
-
-We thank Ben Jann, Robert Fluder and Tobias Fritschi for helpful comments on the article. We also like to thank Stefan Ilic for the prepartion of the data set.
-
-%%%-------------------------------------------------%%%
-%%% Include the bibliography %%%
-%%%-------------------------------------------------%%%
-
-\end{multicols}
-
-\bibliography{bibliography/bib}
-
-%%%-------------------------------------------------%%%
-%%% Include the appendix %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for appendix %%%
-%%%-------------------------------------------------%%%
-
-\section{Appendix}
-
-\end{document}

### 2014-01-19 paper/subdocuments/data_and_methods.Rnw
playig with zeros

--- 2014-01-12 paper/subdocuments/data_and_methods.Rnw
@@
 ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)
+ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G),color="blue")
 @
@@
 
+<<echo=FALSE>>=
+# checking the impact of zeros on an artificial distribution
+# generate a few differently scewed distributions and introduce zeros.
+x02 <- rgamma(1000,shape=0.2,scale=5)
+x05 <- rgamma(1000,shape=0.5,scale=2)
+x1 <- rgamma(1000,shape=1,scale=1/1)
+x2 <- rgamma(1000,shape=2,scale=1/2)
+x3 <- rgamma(1000,shape=3,scale=1/3)
+x4 <- rgamma(1000,shape=4,scale=1/4)
+x1 <- x1[order(x1)]
+x2 <- x2[order(x2)]
+x3 <- x3[order(x3)]
+x4 <- x4[order(x4)]
+# delete the lowest 20%
+x02_20 <- x02
+x02_20[1:200] <- NA
+x05_20 <- x05
+x05_20[1:200] <- NA
+x1_20 <- x1
+x1_20[1:200] <- NA
+x2_20 <- x2
+x2_20[1:200] <- NA
+x3_20 <- x3
+x3_20[1:200] <- NA
+x4_20 <- x4
+x4_20[1:200] <- NA
+
+gini(x02)-gini(na.omit(x02_20))
+
+gini(x05)-gini(na.omit(x05_20))
+
+gini(x1)-gini(na.omit(x1_20))
+
+gini(x2)-gini(na.omit(x2_20))
+
+gini(x3)-gini(na.omit(x3_20))
+
+gini(x4)-gini(na.omit(x4_20))
+# gini inceases
+@
+
 A positive coefficient (e.g.
@@
 A negative coefficient (e.g.
-Geneva) means the more zeros there the are the smaller the gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
+Geneva) means the more zeros there are the smaller the gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
 This is less obvious than the Schwyz case and our reading is: more zeros mask inequality that arises from the bottom.
+
+\subsubsection{Descriptives - where do we put this?}
+nice grapics for taxable income, reines einkommen and income after taxes here...

### 2014-01-19 paper/subdocuments/introduction.Rnw
Introduction

--- 2013-12-13 paper/subdocuments/introduction.Rnw
@@
 
-The purpose of the following paper is threefold.
-First of all, information about economic inequality in Switzerland is sparse.
-E.g.
-the well known Worldbank data do not contain any Gini-coefficient for Switzerland which is a rare exception within Europe.
-We try to fill this gap with official data from the Swiss Federal Tax Administration (FTA).
-Second, we assess the suitability of the FTA data to report inequality and its changes over time.
-An important aspect here is that the data do not contain all income earning persons in Switzerland but are left censored as federal taxes are payable only above a certain threshold.
-We report magnitude and variance of this bias.
-Third, we integrate the development of swiss inequality into the international picture.
-We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).
+% Im European Sociologica Review scheint es unüblich zu sein, ohne theoretische Sektion zu arbeiten.
+Ich hoffe nicht, dass dies ein k.o.
+Kriterium ist.
+wir halten dafür die Einführungs und die Methoden Sektion etwas allgemeiner
+
+% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern
+
+% 1.Abschnitt: Bedeutung von Ungleichheit
+
+Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular (Wilkinson and Pickett 2009).
+While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders (WEF 2013) and scholars alike.
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013).
+The pattern of evolution suggests a polarization, which can be referred to as the "hollowing of the the middle class'' (Alderson and Doran 2013).
+Households are moving toward the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril.''
+\\
+
+% 2.Abschnitt: Bedeutung der Datengrundlage und Einführung von Steuerdaten
+
+Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
+Atkinson (2013:8) observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
+On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality unsdisputedly.
+These concerns have led to the search of alternative data sources, which can supplement the established survey data studies.
+Already Kuznet (1955) used tax data to examine the relationship between economic growth and personal distribution of income.
+Then it took several decades until Piketty (2001,2003) made the use of tax data fashionable again.
+Following his approach studies on several countries were conducted (Atkinson and Piketty 2007, 2010).
+Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database (Alvaredo et al.
+2014).
+\\
+
+% 3.Abschnitt: What do we know about Switzerland
+%
+
+What is know about Switzerland so far?
+Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
+Nevertheless in recent years Switzerland experienced  a ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+The debate went along with referendums trying the regulate the market outcomes.
+Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisings high majority of 68\%, the 1:12-Initiative, which aimed at the hole spectrum of the income distribution, was rejected.
+The voting about the minimum wage-initiative will be held on 9th February 2014.
+All this referendums questioned the actual distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
+Several official and semi-official publications addressed the question of rising inequality in Switzerland.
+\\
+
+% Lässt sich allenfalls streichen/kürzen
+% Ist es die Bezeichnung offical publications sinnvoll? Oder besser govermental publications
+
+% 4.Abschnitt: What do we know about Switzerland
+% Offical Data Collections on development of inequality
+
+The most recent official figures published by the Federal Statistical Office (2013) show  stability concerning the income distribution covering the time period of 1998 until 2011.
+The authors found a slight increase in the primary income which endured until the recent crisis.
+This temporary increase was mainly compensated by governmental redistribution.
+A longer period is covered in the LIS-Data-set (1982-2004).\footnote{Data-provider for the LIS Data is the Swiss Federal Statistical Office.
+The data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+The OCED-Database includes measures from Income and consumption survey as well.
+Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
+This change in survey is considered as a strict break.
+Comparison before and since 2008 is not recommended (OECD 2012:315).
+All in all the LIS Data-set contains the longest time series on inequality for Switzerland.} Analyzing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
+This result is support by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).
+\\
+
+% 5.Abschnitt: What do we know about Switzerland
+% Studies with tax data
+
+Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
+Dell et al.
+(2007) used tax data from the Federal Tax Administration to asset the development of concentration of the highest incomes and wealth (top-shares).
+In contrast to most all other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
+Using the same approach Foellmi and Martinez (2013) expand the Dell et al.
+timeline to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
+A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
+Flückiger et al.
+(2007) and also Jeitziner and Peters (2007,2009) report constant inequality from 1960-1996 respectively from 1995 to 2003.
+Covering a similar time period Bauer and Spycher (1994) and Bolzani and Abul Naga (2002) found decreasing inequality.
+On the other hand Buchmann and Sacchi (1995) and Ernst et al.
+(2009) found an increase in the 1980s).} \\
+
+% 6. Abschnitt: Zwischenbilanz und Herausforderungen
+
+The divergence can be explained with several factors.
+First of all, different data sources were used.
+The official data providers trust on survey data, whereas the later mentioned publications use tax data.
+It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality.
+On the other hand the focus on top income neglects other changes in the distribution income as it is not possible to see, whether newer concerns like the "hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
+Different measure of inequality hardens the comparability.
+Third, different income concepts and different unit of analysis were used.
+As it is shown by Modetta and Müller (2012) income distribution is strongly affected by governmental redistribution, reducing inequality substantial.
+With the focus on tax data the change in institutional settings is not covered.
+Also neglected is the household structure, whereas it is unclear how inequality is affected whether one looks at household income or at income of tax units.
+It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
+\\
+
+% Allenfalls hier noch nicht so ausführlich mit Vor- und Nachteilen von Steuerdaten kommen.
+Vielleicht besser darauf hinweisen, welche Dimensionen betroffen sind und das dies in der Methods Section besprochen wird.
+
+While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
+The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
+But, up to the day Information for Switzerland is not available.
+Shorrock et al.
+(2013) unify several datasources (including tax data) to assess the pattern of global wealth.
+Their databook includes figures for Switzerland.
+In contrast to inequality of income concerning inequalit of wealth Switzerland takes a leading position \\
+
+% the credit suisse wealth report allready uses wealth tax records, but they use it only for the period from 2000-2011.
+They did exactly the same for Denmark and Norway, which could be interessting to compare.
+% Erbschaftssteuer erwähnen und bei den Global Wealth Reports der Credit Suisse und den Studien im Forschungsantrag weitermachen.
+
+Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort collect data which can be harmonized to comparable measures (see Luxembourg income study).
+However, it is unclear how the bias through non-response affects the overall measure of inequality.
+Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
+Building on recent developments in the field of inequality research, we assess the suitability of the public accessible tax data to report inequality and its changes over time.
+First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
+Second, we summarize and apply tax data specific technics to construct suitable measures of inequality.
+We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
+Third, we integrate the development of Swiss inequality into the international picture.
+
+% Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.

### 2014-01-19 paper/subdocuments/introduction.tex

@@
 
-% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern.
+% Im European Sociologica Review scheint es unüblich zu sein, ohne theoretische Sektion zu arbeiten.
+Ich hoffe nicht, dass dies ein k.o.
+Kriterium ist.
+wir halten dafür die Einführungs und die Methoden Sektion etwas allgemeiner
 
-% 1.Abschnitt: Bedeutung von Ungleichheit von Soziologischer Sicht
+% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern
 
-\emph{Bedeutung von ökonomischer Ungleicheit} \\
+% 1.Abschnitt: Bedeutung von Ungleichheit
 
-% 2.Abschnitt Trends in inequality research
-\emph{Studien zur allgemeinen Entwicklung der Ungleichheit nach unterschiedliche Paradigmen}(Armut, Reiche, Mittelschicht)  [Concerns about the hollowing of the middle class is a argument why the tax studies should be exploited to gain knowledge about the whole distribution of economic resources.
-Until these days tax data is only used to assess the change in top income shares.] \\
-\begin{itemize}
-  \item Development of inequality with LIS Data (Gornick and Jäntti 2013)
-  \item ranking of inequality across high-income countries (Atkinson, Rainwater and Smeeding (1995)
-   \item inequality and low income/relative poverty (Ferreira and Ravallion, 2009; Nolan and Marx, 2009; Smeeding, O’Higgins and Rainwatter, 1990; Rainwater and Smeeding 2003; Förster and Vlemickx, 2004)
-  \item inequality and top income (Atkinson and Piketty, 2007; Leigh, 2009)
-  \item inequality and the middle class (Estache and Leipziger, 2009; Littrell et al., 2001; Birdsall, 2010; Ravallion, 2010; Frank,2007)
-\end{itemize}
+Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular (Wilkinson and Pickett 2009).
+While the study of social inequality can be considered as one of the core subjects of sociology, in more recent years the concern about the widening gap was addressed by global leaders (WEF 2013) and scholars alike.
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013).
+The pattern of evolution suggests a polarization, which can be referred to as a "hollowing of the the middle class'' (Alderson and Doran 2013).
+Households are moving toward the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril.''
+\\
 
-% 3.Abschnitt: Studien mit Steuerdaten
+% 2.Abschnitt: Bedeutung der Datengrundlage und Einführung von Steuerdaten
 
-\emph{Studien mit Steuerdaten}
-\begin{itemize}
-\item Old studies with tax data (Vilfredo Pareto, 1895, 1896; Simon Kuznets, 1953)
-\item More recently revival of tax-studys with focus on top income (Piketty and Saez, 2003; Atkinson and Piketty, 2007,2010).
-\end{itemize}
+Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
+Atkinson (2013:8) observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
+On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality unsdisputedly.
+These concerns have led to the search of alternative data sources, which can supplement the established survey data studies.
+Already Kuznet (1955) used tax data to examine the relationship between economic growth and personal distribution of income.
+But it took several decades until Piketty (2001,2003) made the use of tax data fashionable again.
+Following his approach studies on several countries were conducted (Atkinson and Piketty 2007, 2010).
+Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database (Alvaredo et al.
+2014).
+\\
 
-% 4.Abschnitt: What do we know about Switzerland?
+% 3.Abschnitt: What do we know about Switzerland
+%
 
-\emph{Official Data Collection} \\
+What is know about Switzerland so far?
+Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
+Nevertheless in recent years Switzerland experienced  a ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+The debate went along with referendums trying the regulate the market outcomes.
+Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisings high majority of 68\%, the 1:12-Initiative, which aimed at the hole spectrum of the income distribution, was rejected.
+The voting about the minimum wage-initiative will be held on 9th February 2014.
+All this referendums questioned the actual distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
+Several official and semi-official publications addressed the question of rising inequality in Switzerland.
+\\
 
-\begin{itemize}
-\item  LIS (1982-2004, Survey Data from Swiss Federal Statistical Office three different surveys (recent years from “income and Consumption Survey, main result; decreasing inequality)
-\item  OECD (2000 und 2004, Income and consumption survey, 2008, EU-SILC. The change of survey in 2008 is consdered as a strict break. Comparison over time not possible (OECD 2012:315)
-\item  Worldbank: Nothing
-\item  Federal Office of statistics (1998-2011, HABE, survey data, very detailed measurement, main result; inequality in post transfer income is constant)
-\item  World income inequality Database (WIID) is more or less the same data as it is published in the LIS data
-\end{itemize}
-\emph{Fazit}:LIS longest time series, comparability is questionable, more recent years time series from Federal Office of statistics.
-Both studies use survey data, which, as we will discuss later, have certain drawbacks.\\
+% Lässt sich allenfalls streichen/kürzen
+% Ist es die Bezeichnung offical publications sinnvoll? Oder besser govermental publications
 
-\emph{Studies on Switzerland} \\
-Vgl. Forschungsantrag (kürzen) plus neuere Studien, Verteilungsbericht 2011 und 2012 (SGB), Avenir Suisse (2013) und Gorgas and Schaltenegger (2011) und Foellmi and Martinez (2013). \\
+% 4.Abschnitt: What do we know about Switzerland
+% Offical Data Collections on development of inequality
 
-% Should we discus the non-scientic literature?
+The most recent official figures published by the Federal Statistical Office (2013) show  stability concerning the income distribution covering the time period of 1998 until 2011.
+The authors indeed found a slight increase in the primary income which endured until the recent crisis.
+But this temporary increase was mainly compensated by governmental redistribution.
+A longer timeperiod is covered in the LIS-Data-set (1982-2004).\footnote{Data-provider for the LIS Data is the Swiss Federal Statistical Office.
+The data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+The OCED-Database includes measures from Income and consumption survey as well.
+Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
+This change in survey is considered as a strict break.
+Comparison before and since 2008 is not recommended (OECD 2012:315).
+All in all the LIS Data-set contains the longest time series on inequality for Switzerland.} Analyzing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
+This result is support by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).
+\\
 
-% What is the purpose of our paper?
-The purpose of the following paper is threefold.
-First of all, information about economic inequality in Switzerland is sparse.
-We  fill this gap with official data from the Swiss Federal Tax Administration (FTA).
-Second, we assess the suitability of the FTA data to report inequality and its changes over time.
-Third, we integrate the development of swiss inequality into the international picture.
-We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).
+% 5.Abschnitt: What do we know about Switzerland
+% Studies with tax data
+
+Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
+Dell et al.
+(2007) used tax data from the Federal Tax Administration to asset the development of concentration of the highest incomes and wealth (top-shares).
+In contrast to most other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
+Using the same approach Foellmi and Martinez (2013) expand the Dell et al.
+timeline to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
+A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
+Flückiger et al.
+(2007) and also Jeitziner and Peters (2007,2009) report constant inequality from 1960-1996 respectively from 1995 to 2003.
+Covering a similar time period Bauer and Spycher (1994) and Bolzani and Abul Naga (2002) found decreasing inequality.
+On the other hand Buchmann and Sacchi (1995) and Ernst et al.
+(2009) found an increase in the 1980s).} \\
+
+% 6. Abschnitt: Zwischenbilanz und Herausforderungen
+
+The divergence can be explained with several factors.
+First of all, different data sources were used.
+The official data providers trust on survey data, whereas the later mentioned publications use tax data.
+It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality.
+On the other hand the focus on top income neglects other changes in the distribution income as it is not possible to see, whether newer concerns like the "hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
+Different measure of inequality hardens the comparability.
+Third, different income concepts and different unit of analysis were used.
+As it is shown by Modetta and Müller (2012) income distribution is strongly affected by governmental redistribution, reducing inequality substantialy.
+With the focus on tax data the change in institutional settings is not covered.
+Also neglected is the household structure, whereas it is unclear how inequality is affected whether one looks at household income or at income of tax units.
+It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
+\\
+
+% Allenfalls hier noch nicht so ausführlich mit Vor- und Nachteilen von Steuerdaten kommen.
+Vielleicht besser darauf hinweisen, welche Dimensionen betroffen sind und das dies in der Methods Section besprochen wird.
+
+While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
+The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
+But, up to the day Information for Switzerland is not available.
+Shorrock et al.
+(2013) unify several datasources (including tax data) to assess the pattern of global wealth.
+Their databook includes figures for Switzerland.
+In contrast to inequality of income concerning inequalit of wealth in 2013 Switzerland takes a leading position among the world.
+Studies about a trend of inequality in wealth is missing.
+\\
+
+% the credit suisse wealth report allready uses wealth tax records, but they use it only for the period from 2000-2011.
+They did exactly the same for Denmark and Norway, which could be interessting to compare.
+Dell et al.
+haben auch top income shares gemacht.
+% Erbschaftssteuer erwähnen? Weitere Studien im Forschungsantrag
+
+Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort collect data which can be harmonized to comparable measures (see Luxembourg income study).
+However, it is unclear how the bias through non-response affects the overall measure of inequality.
+Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
+Building on recent developments in the field of inequality research, we assess the suitability of the public accessible tax data to report inequality and its changes over time.
+First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
+Second, we summarize and apply tax data specific technics to construct suitable measures of inequality.
+We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
+Third, we integrate the development of Swiss inequality into the international picture.
+
+% Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.

### 2014-01-19 paper/subdocuments/introduction.Rnw
korrekturen und anmerkungen

@@
 The pattern of evolution suggests a polarization, which can be referred to as the "hollowing of the the middle class'' (Alderson and Doran 2013).
-Households are moving toward the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril.''
+Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril.''
 \\
@@
 Atkinson (2013:8) observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
-On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality unsdisputedly.
+On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality undisputedly.
 These concerns have led to the search of alternative data sources, which can supplement the established survey data studies.
@@
 
-What is know about Switzerland so far?
+What is known about Switzerland so far?
 Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
-Nevertheless in recent years Switzerland experienced  a ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
-The debate went along with referendums trying the regulate the market outcomes.
-Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisings high majority of 68\%, the 1:12-Initiative, which aimed at the hole spectrum of the income distribution, was rejected.
+Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+The debate went along with referendums trying to regulate the market outcomes.
+Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisingly high majority of 68\%, the 1:12-Initiative, which aimed at the whole spectrum of the income distribution, was rejected.
 The voting about the minimum wage-initiative will be held on 9th February 2014.
-All this referendums questioned the actual distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
+All these referendums questioned the current distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
 Several official and semi-official publications addressed the question of rising inequality in Switzerland.
@@
 % Ist es die Bezeichnung offical publications sinnvoll? Oder besser govermental publications
+% rf: oder "Federal publications"?
 
@@
 
-The most recent official figures published by the Federal Statistical Office (2013) show  stability concerning the income distribution covering the time period of 1998 until 2011.
+The most recent official figures published by the Swiss Federal Statistical Office (2013) show stability concerning the income distribution covering the time period of 1998 until 2011.
 The authors found a slight increase in the primary income which endured until the recent crisis.
@@
 The data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
-The OCED-Database includes measures from Income and consumption survey as well.
+The OECD-Database includes measures from Income and Consumption survey as well.
 Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
@@
 Comparison before and since 2008 is not recommended (OECD 2012:315).
-All in all the LIS Data-set contains the longest time series on inequality for Switzerland.} Analyzing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
-This result is support by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).
+All in all the LIS dataset contains the longest time series on inequality for Switzerland.} Analyzing these data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
+This result is supported by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).
 \\
@@
 Dell et al.
-(2007) used tax data from the Federal Tax Administration to asset the development of concentration of the highest incomes and wealth (top-shares).
-In contrast to most all other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
+(2007) used tax data from the Federal Tax Administration to assess the development of concentration of the highest incomes and wealth (top-shares).
+In contrast to most other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
 Using the same approach Foellmi and Martinez (2013) expand the Dell et al.
@@
 It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality.
-On the other hand the focus on top income neglects other changes in the distribution income as it is not possible to see, whether newer concerns like the "hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
-Different measure of inequality hardens the comparability.
-Third, different income concepts and different unit of analysis were used.
-As it is shown by Modetta and Müller (2012) income distribution is strongly affected by governmental redistribution, reducing inequality substantial.
+On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the "hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
+Different measure of inequality hampers the comparability.
+Third, different income concepts and different units of analysis were used.
+As it is shown by Modetta and Müller (2012) income distribution is strongly affected by governmental redistribution, reducing inequality substantially.
 With the focus on tax data the change in institutional settings is not covered.
@@
 Vielleicht besser darauf hinweisen, welche Dimensionen betroffen sind und das dies in der Methods Section besprochen wird.
+%rf: finde den abschnitt super
 
@@
 The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
-But, up to the day Information for Switzerland is not available.
+But, up to the day information for Switzerland is not available.
+% ab hier wird es irgnediwe unklar
 Shorrock et al.
@@
 Their databook includes figures for Switzerland.
-In contrast to inequality of income concerning inequalit of wealth Switzerland takes a leading position \\
+In contrast to inequality of income concerning inequality of wealth Switzerland takes a leading position \\
 
-% the credit suisse wealth report allready uses wealth tax records, but they use it only for the period from 2000-2011.
-They did exactly the same for Denmark and Norway, which could be interessting to compare.
+% the credit suisse wealth report already uses wealth tax records, but they use it only for the period from 2000-2011.
+They did exactly the same for Denmark and Norway, which could be interesting to compare.
 % Erbschaftssteuer erwähnen und bei den Global Wealth Reports der Credit Suisse und den Studien im Forschungsantrag weitermachen.
 
-Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort collect data which can be harmonized to comparable measures (see Luxembourg income study).
+Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study).
 However, it is unclear how the bias through non-response affects the overall measure of inequality.
 Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
-Building on recent developments in the field of inequality research, we assess the suitability of the public accessible tax data to report inequality and its changes over time.
+Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.
 First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
@@
 % Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.
+% rf: gute frage, spontan würde ich sagen eher nein, müssten wir mal diskutieren

### 2014-01-20 paper/subdocuments/data_and_methods.Rnw
data update, gini-series (graphic) draft

@@
 nice grapics for taxable income, reines einkommen and income after taxes here...
+<<different_ginis>>=
+df$G_reink[df$G_reink==1]<-NA
+df$G_taxed[df$G_taxed==1]<-NA
+ggplot(df[df$kanton=="CH",], aes(x=steuerperiode,y=G_steink))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_reink),color="blue")+geom_line(aes(y=G_taxed),color="green")
+ggplot(df, aes(x=steuerperiode,y=G_steink))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_reink),color="blue")+geom_line(aes(y=G_taxed),color="green")
+@

### 2014-01-21 paper/paper.tex
fixes, so compiling Rnw works

--- 2014-01-14 paper/paper.tex
@@
 % Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.
+% rf: gute frage, spontan würde ich sagen eher nein, müssten wir mal diskutieren
 
@@
 %[He also discusses tax data studies in general-> have a closer look] \\
-%[We definitely have to argue, why it isn’t sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
+%[We definitely have to argue, why it isnt sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
 
@@
 \begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial “income” see above)
+\item Misreporting of incomes (high earners have an interest in getting tax beneficial income see above)
 \item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the “top-income” literature, albeit it is a crucial issue in “theorized” concepts of income.
+This is a point, which is hardly mentioned in the top-income literature, albeit it is a crucial issue in theorized concepts of income.
 To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
@@
 See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
-%	•	What about the special cases?
+%		What about the special cases?
 Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
 Exclusion leads to an underestimation of inequality
-%	•	Capital gains? Should not be included 
+%		Capital gains? Should not be included 
 The discussion about problems with reporting income is fairly exhaustive. What about wealth?
@@
 \textbf{deductions} Del et al.
-(2007:477):” we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.”
-Gorgas and Schaltenegger (2011:5): “”..., information on […] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
+(2007:477): we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.
+Gorgas and Schaltenegger (2011:5): ..., information on [] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
 Können wir das auch?
@@
 On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a “perverse redistribution” effect by redistributing income from the lower middle class to the upper middle class (vgl.
+There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a perverse redistribution effect by redistributing income from the lower middle class to the upper middle class (vgl.
 Peters 2011 and Interpellation Barbara Gysel (2009).
@@
 Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
+
+One can take a first look at the descriptives plotting the share of zeros over time seperately for each canton .
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/zero_descriptives1}
+
+\includegraphics[width=\maxwidth]{figure/zero_descriptives2}
+
+\end{knitrout}
+
+We can see multiple things here:
+
+\begin{enumerate}
+\item There is a small overall upward trend which we assume to be the Federal Administrations inflation adjustments to the tax threshold.
+\item Geneva and Tessin show wild changes but those might be explained by the tax gap ("Bemessungslücke") that people exploited when the cantons changed the tax system.
+It remains unclear however why we can't see similar patter within other cantons.
+\item There is some variance and we see different patterns over time and cantons.
+When estimating gini coefficients or the like we must therefore assume that ignoring "the zeros" leads to a bias that is not stable over time.
+\end{enumerate}
+
+We might try two strategies to moderate the problem:
+
+\begin{enumerate}
+\item Add the zeros as a seperate group
+\item Fit a model to predict the inequality measure (e.g. the gini coefficient) using the share of zeros as a predictor
+\end{enumerate}
+
+By adding the zeros as a seperate group we face several problems.
+With the exception of Geneva 1995/96 Pareto interpolation of percentiles p20 and above seems viable.
+Low percentiles however would need to be extrapolated so we might impose unrealistic assuptioms.
+Even when estimating "safe" measures above p20 (better p50) or a gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
+To be consistent with the measure of "taxable income" we could assume an income of zero for that group, calculate gini coefficients and compare them with the original (uncorrected/plain) version.
+To get a rough number we could (cantonwise) check the squared correlation between plain and corrected gini.
+
+The second approach however is more robust as we do not impose additional assumptions but instead exercise a little curve fitting.
+
+The model outputs a teststatisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
+As the model has a decent fit we are not in great danger of omitted variable bias.
+Using a joint F-Test we can now test if all canton#zeros interactions are zero.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+## Wald test for kanton:zeros
+##  in lm(formula = G_steink ~ kanton + kanton:zeros + factor(steuerperiode) + 
+##     factor(steuerperiode):zeros, data = df)
+## F =  3.282  on  27  and  262  df: p= 4e-07
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+In our case we can clearly reject the hypothesis that all interactions are zero.
+This leads to the conclusion  that gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted gini coefficients.
+For example one might be interested in how inequality would had developed if the zero-share would have been constant over time.
+(Note RF: predict all data point using canton, time and the initial OR final zero-share to homogenize the time series)
+Furthermore we can quantify how large the bias is and we can do this seperately for tax periods or seperately for cantons.
+
+For all cantons:
+
+We can see the the model fit reduces to explaining 61.5\% of the gini variation versus 92.6\% when the information from the zero-shares was used.
+Although this so some extent attributable to the additional 26 paramters: this is huge.
+
+The model indicates some cases that deserve more attention: Schwyz (positive coefficient) and Geneva (negative coefficient) and the tax period 2000 as well as the most recent periods.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+## [1] -0.006575
+## [1] -0.005752
+## [1] 0.09164
+## [1] 0.08027
+## [1] 0.07055
+## [1] 0.06316
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+A positive coefficient (e.g.
+Schwyz) can be read as follows: In periods with many zeros we measure higher gini coefficients.
+Our conclusion is that people with low income are more successfully hidden.
+If those people would be taxed and therefore appear in the statistic they would sharply contrast the high income earners.
+
+A negative coefficient (e.g.
+Geneva) means the more zeros there are the smaller the gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
+This is less obvious than the Schwyz case and our reading is: more zeros mask inequality that arises from the bottom.
+
+\subsubsection{Descriptives - where do we put this?}
+nice grapics for taxable income, reines einkommen and income after taxes here...
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/different_ginis1}
+
+\includegraphics[width=\maxwidth]{figure/different_ginis2}
+
+\end{knitrout}
 
@@
 
+%rf: genau. kantone unterstreichen die probleme mit dem 0er bias und die probleme mit "imputing the gap". mehr muss erstmal nicht sein.
+
 %%%-------------------------------------------------%%%
@@
 %Methodological Main findings 
-%(1) Tax data is useful because it’s possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
-It’s superior to survey data, because the latter is affected by non-responded bias.
-Tax data includes data about the whole population … [was finden wir wegen den Nullern raus] .
+%(1) Tax data is useful because its possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
+Its superior to survey data, because the latter is affected by non-responded bias.
+Tax data includes data about the whole population  [was finden wir wegen den Nullern raus] .
 or at least it should (tax evasion).(3) The FTA-Tax data has different shortcomings, which cannot be handled like a measure of income and wealth which (a) cannot be corrected for household size and (b) is neither a pre-transfer nor a post-transfer income measure.
-Therefor it’s comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
+Therefor its comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
 This FTA-Tax data problem can be handled with non-aggregated tax-data .

### 2014-01-21 paper/subdocuments/data_and_methods.Rnw
fixes, so compiling Rnw works

@@
 
-<<setup echo=FALSE>>=
+<<setup>>=
 library(foreign)
-df <- read.dta("data/ginis_und_perzentile.dta")
+df <- read.dta("../../data/ginis_und_perzentile.dta")
+df$zeros <- df$null_norm/df$cpop # i only use normal cases here
 @
@@
 library(ggplot2)
-df$zeros <- df$null_norm/df$cpop # i only use normal cases here
 ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)
-ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G),color="blue")
+ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")
 @
@@
 <<correct_gini>>=
-fit <- lm(G~kanton+kanton:zeros+factor(steuerperiode)+factor(steuerperiode):zeros,data=df)
-summary(fit)
+fit <- lm(G_steink~kanton+kanton:zeros+factor(steuerperiode)+factor(steuerperiode):zeros,data=df)
+#summary(fit)
 @
@@
 <<magnitude_of_bias>>=
-reduced_fit <- lm(G~kanton+factor(steuerperiode),data=df)
-summary(reduced_fit)
+reduced_fit <- lm(G_steink~kanton+factor(steuerperiode),data=df)
+#summary(reduced_fit)
 @
@@
 
+library(reldist)
 gini(x02)-gini(na.omit(x02_20))

### 2014-01-21 paper/subdocuments/data_and_methods.tex
fixes, so compiling Rnw works

@@
 
-\textbf{Summary of tax data drawbacks }
+\textbf{Summary of tax data drawbacks}
 \begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial “income” see above)
+\item Misreporting of incomes (high earners have an interest in getting tax beneficial income see above)
 \item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the “top-income” literature, albeit it is a crucial issue in “theorized” concepts of income.
+This is a point, which is hardly mentioned in the top-income literature, albeit it is a crucial issue in theorized concepts of income.
 To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
@@
 See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
-%	•	What about the special cases?
+%		What about the special cases?
 Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
 Exclusion leads to an underestimation of inequality
-%	•	Capital gains? Should not be included 
+%		Capital gains? Should not be included 
 The discussion about problems with reporting income is fairly exhaustive. What about wealth?
+
+It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
+Credit Suisse (2011) makes use of the information in the "Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
+\\
+
+Additonaly the Credit Suisse Reports states (2011:8) that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts.
 
@@
 
+%Beschreiben was die Daten beinhalten und wo es Probleme gibt
+\begin{itemize}
+\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
+	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
+	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
+	\item	Income after deductions-> taxable income (employment income, business income and capital income).
+Is income definition stable over time?
+Yes it should.
+Realized capital gains are excluded from the definition.
+It includes income from  employment, self-employment, capital income and taxable transfer payments.
+Plus Eigenmietwert,
+	\item	Reported on national and cantonal level
+\end{itemize}
+
 \subsection{Ways to tackle FTA-tax data specific problems}
+
+\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
+Del et al.
+(2007) impute for non-fillers the 20 percentage of the annual average income.
+This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
+Furthermore Del et al.
+calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
+\\
+
+\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
+This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
+This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f).
+\\
+%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.
+
+\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
+
+\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
+%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
+
+\textbf{deductions} Del et al.
+(2007:477): we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.
+Gorgas and Schaltenegger (2011:5): ..., information on [] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
+Können wir das auch?
+Zumindest für gewisse Zeiträume?
+Das wäre noch gut.
+\\
+
+Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
+Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.
+On the other hand deductions can affect the distribution.
+There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a perverse redistribution effect by redistributing income from the lower middle class to the upper middle class (vgl.
+Peters 2011 and Interpellation Barbara Gysel (2009).
+
+% Income share specific problems
+% Total income denominator Exogenous Approach -> net income reported in the national accounts.
+Endogenous Approach -> Dell et al.
+technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income).
+Honestly , is this appropriate?
+(Everyone did it)
+% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
+% Endogenous approach -> reported in the tax tables
+
+%%%-------------------------------------------------%%%
+%%% Abschnitte aus ersten Version des Papers %%%
+%%%-------------------------------------------------%%%
 
@@
 Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
+
+One can take a first look at the descriptives plotting the share of zeros over time seperately for each canton .
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(foreign)}
+\hlkwd{getwd}\hlstd{()}
+\end{alltt}
+\begin{verbatim}
+## [1] "C:/Users/Hackstutz/Dropbox/Git/swiss_inequality_development/paper/subdocuments"
+\end{verbatim}
+\begin{alltt}
+\hlstd{df} \hlkwb{<-} \hlkwd{read.dta}\hlstd{(}\hlstr{"data/ginis_und_perzentile.dta"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: unable to open file: 'No such file or directory'}}\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(ggplot2)}
+\hlstd{df}\hlopt{$}\hlstd{zeros} \hlkwb{<-} \hlstd{df}\hlopt{$}\hlstd{null_norm}\hlopt{/}\hlstd{df}\hlopt{$}\hlstd{cpop}  \hlcom{# i only use normal cases here}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt des Typs 'closure' ist nicht indizierbar}}\begin{alltt}
+\hlkwd{ggplot}\hlstd{(df[df}\hlopt{$}\hlstd{steuerperiode} \hlopt{>=} \hlnum{1995}\hlstd{, ],} \hlkwd{aes}\hlstd{(}\hlkwc{x} \hlstd{= steuerperiode,} \hlkwc{y} \hlstd{= zeros))} \hlopt{+}
+    \hlkwd{geom_line}\hlstd{()} \hlopt{+} \hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{kanton)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt des Typs 'closure' ist nicht indizierbar}}\begin{alltt}
+\hlkwd{ggplot}\hlstd{(df[df}\hlopt{$}\hlstd{steuerperiode} \hlopt{>=} \hlnum{1995}\hlstd{, ],} \hlkwd{aes}\hlstd{(}\hlkwc{x} \hlstd{= steuerperiode,} \hlkwc{y} \hlstd{= zeros))} \hlopt{+}
+    \hlkwd{geom_line}\hlstd{()} \hlopt{+} \hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{kanton)} \hlopt{+} \hlkwd{geom_line}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= G_steink),} \hlkwc{color} \hlstd{=} \hlstr{"blue"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt des Typs 'closure' ist nicht indizierbar}}\end{kframe}
+\end{knitrout}
+
+We can see multiple things here:
+
+\begin{enumerate}
+\item There is a small overall upward trend which we assume to be the Federal Administrations inflation adjustments to the tax threshold.
+\item Geneva and Tessin show wild changes but those might be explained by the tax gap ("Bemessungslücke") that people exploited when the cantons changed the tax system.
+It remains unclear however why we can't see similar patter within other cantons.
+\item There is some variance and we see different patterns over time and cantons.
+When estimating gini coefficients or the like we must therefore assume that ignoring "the zeros" leads to a bias that is not stable over time.
+\end{enumerate}
+
+We might try two strategies to moderate the problem:
+
+\begin{enumerate}
+\item Add the zeros as a seperate group
+\item Fit a model to predict the inequality measure (e.g. the gini coefficient) using the share of zeros as a predictor
+\end{enumerate}
+
+By adding the zeros as a seperate group we face several problems.
+With the exception of Geneva 1995/96 Pareto interpolation of percentiles p20 and above seems viable.
+Low percentiles however would need to be extrapolated so we might impose unrealistic assuptioms.
+Even when estimating "safe" measures above p20 (better p50) or a gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
+To be consistent with the measure of "taxable income" we could assume an income of zero for that group, calculate gini coefficients and compare them with the original (uncorrected/plain) version.
+To get a rough number we could (cantonwise) check the squared correlation between plain and corrected gini.
+
+The second approach however is more robust as we do not impose additional assumptions but instead exercise a little curve fitting.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{fit} \hlkwb{<-} \hlkwd{lm}\hlstd{(G_steink} \hlopt{~} \hlstd{kanton} \hlopt{+} \hlstd{kanton}\hlopt{:}\hlstd{zeros} \hlopt{+} \hlkwd{factor}\hlstd{(steuerperiode)} \hlopt{+} \hlkwd{factor}\hlstd{(steuerperiode)}\hlopt{:}\hlstd{zeros,}
+    \hlkwc{data} \hlstd{= df)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'data' argument is of the wrong type}}\begin{alltt}
+\hlkwd{summary}\hlstd{(fit)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt 'fit' nicht gefunden}}\end{kframe}
+\end{knitrout}
+
+The model outputs a teststatisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
+As the model has a decent fit we are not in great danger of omitted variable bias.
+Using a joint F-Test we can now test if all canton#zeros interactions are zero.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(survey)}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# \\\#\# Attaching package: 'survey'\\\#\# \\\#\# Das folgende Objekt ist maskiert from 'package:graphics':\\\#\# \\\#\#\ \ \ \  dotchart}}\begin{alltt}
+\hlkwd{regTermTest}\hlstd{(fit,} \hlstr{"kanton:zeros"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt 'fit' nicht gefunden}}\end{kframe}
+\end{knitrout}
+
+In our case we can clearly reject the hypothesis that all interactions are zero.
+This leads to the conclusion  that gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted gini coefficients.
+For example one might be interested in how inequality would had developed if the zero-share would have been constant over time.
+(Note RF: predict all data point using canton, time and the initial OR final zero-share to homogenize the time series)
+Furthermore we can quantify how large the bias is and we can do this seperately for tax periods or seperately for cantons.
+
+For all cantons:
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{reduced_fit} \hlkwb{<-} \hlkwd{lm}\hlstd{(G_steink} \hlopt{~} \hlstd{kanton} \hlopt{+} \hlkwd{factor}\hlstd{(steuerperiode),} \hlkwc{data} \hlstd{= df)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'data' argument is of the wrong type}}\begin{alltt}
+\hlkwd{summary}\hlstd{(reduced_fit)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt 'reduced\_fit' nicht gefunden}}\end{kframe}
+\end{knitrout}
+
+We can see the the model fit reduces to explaining 61.5\% of the gini variation versus 92.6\% when the information from the zero-shares was used.
+Although this so some extent attributable to the additional 26 paramters: this is huge.
+
+The model indicates some cases that deserve more attention: Schwyz (positive coefficient) and Geneva (negative coefficient) and the tax period 2000 as well as the most recent periods.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# reldist: Relative Distribution Methods\\\#\# Version 1.6-2 created on 2013-03-03.\\\#\# copyright (c) 2003, Mark S.
+Handcock, University of California-Los Angeles\\\#\#\ \ For citation information, type citation("{}reldist"{}).\\\#\#\ \ Type help(package="{}reldist"{}) to get started.}}\begin{verbatim}
+## [1] 0.008115
+## [1] 0.006694
+## [1] 0.09447
+## [1] 0.08017
+## [1] 0.07297
+## [1] 0.0621
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+A positive coefficient (e.g.
+Schwyz) can be read as follows: In periods with many zeros we measure higher gini coefficients.
+Our conclusion is that people with low income are more successfully hidden.
+If those people would be taxed and therefore appear in the statistic they would sharply contrast the high income earners.
+
+A negative coefficient (e.g.
+Geneva) means the more zeros there are the smaller the gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
+This is less obvious than the Schwyz case and our reading is: more zeros mask inequality that arises from the bottom.
+
+\subsubsection{Descriptives - where do we put this?}
+nice grapics for taxable income, reines einkommen and income after taxes here...
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{df}\hlopt{$}\hlstd{G_reink[df}\hlopt{$}\hlstd{G_reink} \hlopt{==} \hlnum{1}\hlstd{]} \hlkwb{<-} \hlnum{NA}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt des Typs 'closure' ist nicht indizierbar}}\begin{alltt}
+\hlstd{df}\hlopt{$}\hlstd{G_taxed[df}\hlopt{$}\hlstd{G_taxed} \hlopt{==} \hlnum{1}\hlstd{]} \hlkwb{<-} \hlnum{NA}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt des Typs 'closure' ist nicht indizierbar}}\begin{alltt}
+\hlkwd{ggplot}\hlstd{(df[df}\hlopt{$}\hlstd{kanton} \hlopt{==} \hlstr{"CH"}\hlstd{, ],} \hlkwd{aes}\hlstd{(}\hlkwc{x} \hlstd{= steuerperiode,} \hlkwc{y} \hlstd{= G_steink))} \hlopt{+} \hlkwd{geom_line}\hlstd{()} \hlopt{+}
+    \hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{kanton)} \hlopt{+} \hlkwd{geom_line}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= G_reink),} \hlkwc{color} \hlstd{=} \hlstr{"blue"}\hlstd{)} \hlopt{+} \hlkwd{geom_line}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= G_taxed),}
+    \hlkwc{color} \hlstd{=} \hlstr{"green"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt des Typs 'closure' ist nicht indizierbar}}\begin{alltt}
+\hlkwd{ggplot}\hlstd{(df,} \hlkwd{aes}\hlstd{(}\hlkwc{x} \hlstd{= steuerperiode,} \hlkwc{y} \hlstd{= G_steink))} \hlopt{+} \hlkwd{geom_line}\hlstd{()} \hlopt{+} \hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{kanton)} \hlopt{+}
+    \hlkwd{geom_line}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= G_reink),} \hlkwc{color} \hlstd{=} \hlstr{"blue"}\hlstd{)} \hlopt{+} \hlkwd{geom_line}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y} \hlstd{= G_taxed),}
+    \hlkwc{color} \hlstd{=} \hlstr{"green"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: ggplot2 doesn't know how to deal with data of class function}}\end{kframe}
+\end{knitrout}

### 2014-01-22 paper/subdocuments/data_and_methods.Rnw
finished draft of "analysis of zeros"

@@
 
-The second approach however is more robust as we do not impose additional assumptions but instead exercise a little curve fitting.
+The second approach however is more robust as we do not impose additional assumptions but instead exercise some curve fitting.
 
-<<correct_gini>>=
-fit <- lm(G_steink~kanton+kanton:zeros+factor(steuerperiode)+factor(steuerperiode):zeros,data=df)
+<<corrected_gini>>=
+df$J<-factor(df$steuerperiode)
+fit <- lm(G_steink~kanton+kanton:zeros+J+zeros:J,data=df)
 #summary(fit)
+SmoothCoefficientPlot <- function(models, modelnames = "", removeintercept = FALSE){
+  # models must be a list()
+
+  Alphas <- seq(1, 99, 2) / 100
+
+  Multiplier <- qnorm(1 - Alphas / 2)
+  zzTransparency <<- 1/(length(Multiplier)/4)
+  #CoefficientTables <- lapply(models, function(x){summary(x)$coef})
+  CoefficientTables <- summary(models)$coef
+  #TableRows <- unlist(lapply(CoefficientTables, nrow))
+  TableRows <- nrow(CoefficientTables)
+
+  if(modelnames[1] == ""){
+    ModelNameLabels <- rep(paste("Model", 1:length(TableRows)), TableRows)
+    } else {
+    ModelNameLabels <- rep(modelnames, TableRows)
+    }
+
+  MatrixofModels <- cbind(do.call(rbind, list(CoefficientTables)), ModelNameLabels)
+  if(removeintercept == TRUE){
+    MatrixofModels <- MatrixofModels[!rownames(MatrixofModels) == "(Intercept)", ]
+    }
+  MatrixofModels <- data.frame(cbind(rownames(MatrixofModels), MatrixofModels))
+
+  MatrixofModels <- data.frame(cbind(MatrixofModels, rep(Multiplier, each = nrow(MatrixofModels))))
+
+  colnames(MatrixofModels) <- c("IV", "Estimate", "StandardError", "TValue", "PValue", "ModelName", "Scalar")
+  MatrixofModels$IV <- factor(MatrixofModels$IV, levels = MatrixofModels$IV)
+  MatrixofModels[, -c(1, 6)] <- apply(MatrixofModels[, -c(1, 6)], 2, function(x){as.numeric(as.character(x))})
+  MatrixofModels$Emphasis <- by(1 - seq(0, 1, length = length(Multiplier) + 1)[-1], as.character(round(Multiplier, 5)), mean)[as.character(round(MatrixofModels$Scalar, 5))]
+
+  OutputPlot <- qplot(data = MatrixofModels, x = IV, y = Estimate,
+   ymin = Estimate - Scalar * StandardError, ymax = Estimate + Scalar * StandardError,
+   ylab = NULL, xlab = NULL, alpha = I(zzTransparency), colour = I(gray(0)), geom = "blank")
+  OutputPlot <- OutputPlot + geom_hline(yintercept = 0, lwd = I(7/12), colour = I(hsv(0/12, 7/12, 7/12)), alpha = I(5/12))
+  OutputPlot <- OutputPlot + geom_linerange(data = MatrixofModels, aes(size = as.integer(1/Emphasis)), alpha = I(zzTransparency), colour = I(gray(0)))
+  OutputPlot <- OutputPlot + scale_size_continuous(legend = FALSE)
+  OutputPlot <- OutputPlot + facet_grid(~ ModelName) + coord_flip() + geom_point(aes(x = IV, y = Estimate), colour = I(gray(0))) + theme_bw()
+  return(OutputPlot)
+  }
+
+library(grid) #for arrow() function
+SmoothCoefficientPlot(fit,modelnames=paste0("By cantons: Impact of the share of zeros on inequality (Gini) Rsq=", round(summary(fit)$r.squared,2)))+geom_text(show_guide=FALSE,x=77,y=2.2,size=4,label="effect variation\nover time")+geom_segment(aes(x = 85, y = 1.5, xend = 71, yend = 1.5))+geom_segment(aes(x = 71, y = 1.5, xend = 71, yend = 1.3))+geom_segment(aes(x = 85, y = 1.5, xend = 85, yend = 1.3))+geom_text(show_guide=FALSE,x=57,y=1.9,size=4,label="effect variation\nby cantons")+geom_segment(aes(x = 71, y = 1.2, xend = 45, yend = 1.2))+geom_segment(aes(x = 71, y = 1.2, xend = 71, yend = 1.0))+geom_segment(aes(x = 45, y = 1.2, xend = 45, yend = 1.0))+ggtitle("Bias variation by time and cantons")
+
 @
 
-The model outputs a teststatisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
+The model outputs a test statisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
 As the model has a decent fit we are not in great danger of omitted variable bias.
@@
 
-<<echo=FALSE>>=
-# checking the impact of zeros on an artificial distribution
-# generate a few differently scewed distributions and introduce zeros.
-x02 <- rgamma(1000,shape=0.2,scale=5)
-x05 <- rgamma(1000,shape=0.5,scale=2)
-x1 <- rgamma(1000,shape=1,scale=1/1)
-x2 <- rgamma(1000,shape=2,scale=1/2)
-x3 <- rgamma(1000,shape=3,scale=1/3)
-x4 <- rgamma(1000,shape=4,scale=1/4)
-x1 <- x1[order(x1)]
-x2 <- x2[order(x2)]
-x3 <- x3[order(x3)]
-x4 <- x4[order(x4)]
-# delete the lowest 20%
-x02_20 <- x02
-x02_20[1:200] <- NA
-x05_20 <- x05
-x05_20[1:200] <- NA
-x1_20 <- x1
-x1_20[1:200] <- NA
-x2_20 <- x2
-x2_20[1:200] <- NA
-x3_20 <- x3
-x3_20[1:200] <- NA
-x4_20 <- x4
-x4_20[1:200] <- NA
-
-library(reldist)
-gini(x02)-gini(na.omit(x02_20))
-
-gini(x05)-gini(na.omit(x05_20))
-
-gini(x1)-gini(na.omit(x1_20))
-
-gini(x2)-gini(na.omit(x2_20))
-
-gini(x3)-gini(na.omit(x3_20))
-
-gini(x4)-gini(na.omit(x4_20))
-# gini inceases
-@
-
 A positive coefficient (e.g.
 Schwyz) can be read as follows: In periods with many zeros we measure higher gini coefficients.
-Our conclusion is that people with low income are more successfully hidden.
-If those people would be taxed and therefore appear in the statistic they would sharply contrast the high income earners.
+We can derive, that the distribution of income is more scewed for high incomes than for low incomes.
+Simply speaking, the contrast between low and middle class is less pronounced than the contrast between middle und high class.
+One possible explanation would be that incomes stem from two different populations: 1) the people of Schwyz who possibly follow a log-normal or gamma distribution and 2) particularly rich people who moved to Schwyz to avoid taxes.
 
@@
 Geneva) means the more zeros there are the smaller the gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
-This is less obvious than the Schwyz case and our reading is: more zeros mask inequality that arises from the bottom.
+This is the case we would usually expect: more zeros mask inequality that arises from the bottom.
 
-\subsubsection{Descriptives - where do we put this?}
+What can we conclude from that analysis?
+First one must notice that aggregate measure like the gini (or others) do not always react in the same way when we cut off one part of the distribution, therefore the measures calculated from tax data is biased.
+On the other hand, the model coefficient of Switzerland as a whole is not significant suggesting that the cantonal biases cancel out each other.
+This seems plausible.
+Most of the "tax optimization" happens within Switzerland so the rich people who moved to Schwyz are now missing at another canton.
+
+We can even see more from the model coefficients.
+The period dummies indicate how the distribution of incomes has changed: compared to 1995, the subsequent periods have a negative coefficient, i.e.
+the bias we introduce by omitting the zeros increased, especially from the mid nineties to the mid 2000s.
+To simplify: cutting off zeros more and more seems to lead to an underestimation of our inequality measure, probably because the skewness in the left part of the distribution increased pointing to an increased pauperization that is masked by omitting the left tail of the income distribution.
+
+\subsubsection{Descriptives - has to be moved to results}
 nice grapics for taxable income, reines einkommen and income after taxes here...

### 2014-01-27 paper/subdocuments/data_and_methods.Rnw
zeros

@@
 To simplify: cutting off zeros more and more seems to lead to an underestimation of our inequality measure, probably because the skewness in the left part of the distribution increased pointing to an increased pauperization that is masked by omitting the left tail of the income distribution.
-
-\subsubsection{Descriptives - has to be moved to results}
-nice grapics for taxable income, reines einkommen and income after taxes here...
-<<different_ginis>>=
-df$G_reink[df$G_reink==1]<-NA
-df$G_taxed[df$G_taxed==1]<-NA
-ggplot(df[df$kanton=="CH",], aes(x=steuerperiode,y=G_steink))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_reink),color="blue")+geom_line(aes(y=G_taxed),color="green")
-ggplot(df, aes(x=steuerperiode,y=G_steink))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_reink),color="blue")+geom_line(aes(y=G_taxed),color="green")
-@

### 2014-01-27 paper/subdocuments/results.Rnw
Übertrag aus Skizze
korrekturen und anmerkungen
zeros

@@
 Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
+
+\subsection{Gini coefficients for taxable income, income before deductions and taxable income after tax}
+
+<<setup>>=
+library(foreign)
+library(ggplot2)
+df <- read.dta("../../data/ginis_und_perzentile.dta")
+df$zeros <- df$null_norm/df$cpop # i only use normal cases here
+@
+
+<<different_ginis>>=
+df$G_reink[df$G_reink==1]<-NA
+df$G_taxed[df$G_taxed==1]<-NA
+library(reshape)
+dflong <- melt(df[,c("G_steink","G_reink","G_taxed","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
+names(dflong)[3]<-"Type"
+names(dflong)[4]<-"Gini"
+levels(dflong$Type)<-c("taxable income","income before deductions","taxable income after tax")
+ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line(size=1)+theme_bw()+xlab("tax period")
+ggplot(dflong, aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")
+@
+
+\subsection{Zero share}
+
+Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later)
+
+\subsection{Income measures with and without zeros}
+
+taxable income with and without zeros
+
+<<with_without_zeros>>=
+dflong <- melt(df[,c("G_steink","G_steink0","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
+names(dflong)[3]<-"Type"
+names(dflong)[4]<-"Gini"
+levels(dflong$Type)<-c("without zeros","including zeros")
+ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line(size=1)+theme_bw()+xlab("tax period")
+ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+ opts(axis.text.x=theme_text(angle=60))
+
+@
+
+Including zeros leads to significantly higher gini coefficients.
+However we must keep in mind, that these might be artifically high values as we assume zero income for everyone in the zero group.
+We can conclude more from the graphic: the ratio between both measures seems to be quite constant although for aggregate Switzerland but there are minor deviations for multiple cantons as well as strong deviations for the cantons Geneva and Tessin.
+However the problems seem not to result from a shift in the zero-share over time but they are specific for the time-period when the tax system changed.
+
+%Switzerland (income and wealth)
+%- graph with gini over time> identify periods of change
+%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
+%Cantonal level (income and wealth)
+%- different development on regional level 
+%Switzerland in international comparison
+%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality
+
+%- Comparison with other countries (similar data)
+
+% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen.
+Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel.
+Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.
+
+%rf: genau. kantone unterstreichen die probleme mit dem 0er bias und die probleme mit "imputing the gap". mehr muss erstmal nicht sein.

### 2014-03-04 paper/subdocuments/results.Rnw
fixed plot with "zero-gini"

@@
 library(ggplot2)
-df <- read.dta("../../data/ginis_und_perzentile.dta")
+df <- read.dta("data/ginis_und_perzentile.dta")
 df$zeros <- df$null_norm/df$cpop # i only use normal cases here

### 2014-03-05 paper/subdocuments/data_and_methods.Rnw
small fixes to latex code (encoding)

@@
 \begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial “income” see above)
+\item Misreporting of incomes (high earners have an interest in getting tax beneficial "income" see above)
 \item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the “top-income” literature, albeit it is a crucial issue in “theorized” concepts of income.
+This is a point, which is hardly mentioned in the "top-income" literature, albeit it is a crucial issue in "theorized" concepts of income.
 To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
@@
 It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
-Credit Suisse (2011) makes use of the information in the "Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
+Credit Suisse (2011) makes use of the information in the "Rich Lists" published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
 \\
@@
 \textbf{deductions} Del et al.
-(2007:477):” we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.”
-Gorgas and Schaltenegger (2011:5): “”..., information on […] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
+(2007:477):" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
+Gorgas and Schaltenegger (2011:5): "..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time".
 Können wir das auch?
@@
 On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a “perverse redistribution” effect by redistributing income from the lower middle class to the upper middle class (vgl.
+There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a "perverse redistribution" effect by redistributing income from the lower middle class to the upper middle class (vgl.
 Peters 2011 and Interpellation Barbara Gysel (2009).

### 2014-03-05 paper/subdocuments/introduction.Rnw
small fixes to latex code (encoding)

@@
 Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013).
-The pattern of evolution suggests a polarization, which can be referred to as the "hollowing of the the middle class'' (Alderson and Doran 2013).
-Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril.''
+The pattern of evolution suggests a polarization, which can be referred to as the "hollowing of the the middle class" (Alderson and Doran 2013).
+Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril."
 \\
@@
 It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality.
-On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the "hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
+On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the "hollowing of the middle class" occurred in Switzerland or not, which leads to the second point.
 Different measure of inequality hampers the comparability.

### 2014-03-05 paper/subdocuments/results.Rnw
small fixes to latex code (encoding)

@@
 
-<<setup>>=
+<<setup_gini_data>>=
 library(foreign)
 library(ggplot2)
-df <- read.dta("data/ginis_und_perzentile.dta")
+df <- read.dta("../../data/ginis_und_perzentile.dta")
 df$zeros <- df$null_norm/df$cpop # i only use normal cases here

### 2014-03-05 paper/paper.tex
more figures

@@
 \begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial income see above)
+\item Misreporting of incomes (high earners have an interest in getting tax beneficial "income" see above)
 \item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the top-income literature, albeit it is a crucial issue in theorized concepts of income.
+This is a point, which is hardly mentioned in the "top-income" literature, albeit it is a crucial issue in "theorized" concepts of income.
 To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
@@
 \textbf{deductions} Del et al.
-(2007:477): we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.
-Gorgas and Schaltenegger (2011:5): ..., information on [] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
+(2007:477):" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
+Gorgas and Schaltenegger (2011:5): "..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time".
 Können wir das auch?
@@
 On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a perverse redistribution effect by redistributing income from the lower middle class to the upper middle class (vgl.
+There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a "perverse redistribution" effect by redistributing income from the lower middle class to the upper middle class (vgl.
 Peters 2011 and Interpellation Barbara Gysel (2009).
@@
 
-The second approach however is more robust as we do not impose additional assumptions but instead exercise a little curve fitting.
+The second approach however is more robust as we do not impose additional assumptions but instead exercise some curve fitting.
 
-The model outputs a teststatisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/corrected_gini}
+
+\end{knitrout}
+
+The model outputs a test statisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
 As the model has a decent fit we are not in great danger of omitted variable bias.
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## [1] -0.006575
-## [1] -0.005752
-## [1] 0.09164
-## [1] 0.08027
-## [1] 0.07055
-## [1] 0.06316
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
 A positive coefficient (e.g.
 Schwyz) can be read as follows: In periods with many zeros we measure higher gini coefficients.
-Our conclusion is that people with low income are more successfully hidden.
-If those people would be taxed and therefore appear in the statistic they would sharply contrast the high income earners.
+We can derive, that the distribution of income is more scewed for high incomes than for low incomes.
+Simply speaking, the contrast between low and middle class is less pronounced than the contrast between middle und high class.
+One possible explanation would be that incomes stem from two different populations: 1) the people of Schwyz who possibly follow a log-normal or gamma distribution and 2) particularly rich people who moved to Schwyz to avoid taxes.
 
@@
 Geneva) means the more zeros there are the smaller the gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
-This is less obvious than the Schwyz case and our reading is: more zeros mask inequality that arises from the bottom.
+This is the case we would usually expect: more zeros mask inequality that arises from the bottom.
 
-\subsubsection{Descriptives - where do we put this?}
-nice grapics for taxable income, reines einkommen and income after taxes here...
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/different_ginis1}
+What can we conclude from that analysis?
+First one must notice that aggregate measure like the gini (or others) do not always react in the same way when we cut off one part of the distribution, therefore the measures calculated from tax data is biased.
+On the other hand, the model coefficient of Switzerland as a whole is not significant suggesting that the cantonal biases cancel out each other.
+This seems plausible.
+Most of the "tax optimization" happens within Switzerland so the rich people who moved to Schwyz are now missing at another canton.
 
-\includegraphics[width=\maxwidth]{figure/different_ginis2}
-
-\end{knitrout}
+We can even see more from the model coefficients.
+The period dummies indicate how the distribution of incomes has changed: compared to 1995, the subsequent periods have a negative coefficient, i.e.
+the bias we introduce by omitting the zeros increased, especially from the mid nineties to the mid 2000s.
+To simplify: cutting off zeros more and more seems to lead to an underestimation of our inequality measure, probably because the skewness in the left part of the distribution increased pointing to an increased pauperization that is masked by omitting the left tail of the income distribution.
 
@@
 Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
+
+\subsection{Gini coefficients for taxable income, income before deductions and taxable income after tax}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# \\\#\# Attaching package: 'reshape'\\\#\# \\\#\# Die folgenden Objekte sind maskiert from 'package:plyr':\\\#\# \\\#\#\ \ \ \  rename, round\_any}}\end{kframe}
+\includegraphics[width=\maxwidth]{figure/different_ginis1}
+
+\includegraphics[width=\maxwidth]{figure/different_ginis2}
+
+\end{knitrout}
+
+\subsection{Zero share}
+
+Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later)
+
+\subsection{Income measures with and without zeros}
+
+taxable income with and without zeros
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/with_without_zeros1} 
+\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# 'opts' is deprecated.
+Use 'theme' instead.
+(Deprecated; last used in version 0.9.1)\\\#\# theme\_text is deprecated.
+Use 'element\_text' instead.
+(Deprecated; last used in version 0.9.1)}}\end{kframe}
+\includegraphics[width=\maxwidth]{figure/with_without_zeros2}
+
+\end{knitrout}
+
+Including zeros leads to significantly higher gini coefficients.
+However we must keep in mind, that these might be artifically high values as we assume zero income for everyone in the zero group.
+We can conclude more from the graphic: the ratio between both measures seems to be quite constant although for aggregate Switzerland but there are minor deviations for multiple cantons as well as strong deviations for the cantons Geneva and Tessin.
+However the problems seem not to result from a shift in the zero-share over time but they are specific for the time-period when the tax system changed.
 

### 2014-04-11 paper/subdocuments/data_and_methods.Rnw
some notes and comments

@@
 \emph{Concepts on measuring economic resources}  \\
-Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
+Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together.
+In generall it has to be stateted, that ideal measures only can be defined in context of a given reaserach question.
+This holds, if one thinks about the ideal measure of economic well-being, but also about the unit of analysis.
+Nevertheless, some more or less unified concepts exists, which can be taken as a starting point.
+OECD (2013) \\
 
@@
 
+% an income which is adjusted for the household size is not a real income anymore. It is a measure of indivdual welfare or of the standard of living (BFS:2012)
+
 \emph{Defining wealth}
@@
 
+% Streichen
+
 \emph{Population Coverage} \\
@@
 see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\
+
+The income of the household generally plays the major role in the living standards attained by its members, so the way earners and non-earners grouped together in households, has to be incorporated into the analysis.
+Especially when the the focus of interesst lays in unrevealing the role of demographic factors.
 
@@
 
+% Suitability of data sources can be judeged by different criterias. crossectional or longidutional perspective. longidutional perspective allows for dynamic
+
 %[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
@@
 %- Suitability of cross country comparison
-%- Assessing inequality accurately
+%- Coverage of the population of interesst
 %- Implementation of common concepts of economic ressources

### 2014-04-11 paper/subdocuments/introduction.Rnw
some notes and comments

@@
 While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders (WEF 2013) and scholars alike.
-Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013).
-The pattern of evolution suggests a polarization, which can be referred to as the "hollowing of the the middle class" (Alderson and Doran 2013).
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013;Salverda et al.
+2014).
+Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the "hollowing of the the middle class" (Alderson and Doran 2013).
 Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril."
 \\
+
+% Allenfalls könnte man auch (Nolan und Whealan, 2014) erwähnen.
+"The social Impact of income inequality: Poverty, Deprivation, and Social Cohesion" Da werden die Thesen rund um die Folgen von Ungleichheit gut zusammengefasst.
 
@@
 
-% Lässt sich allenfalls streichen/kürzen
+% Lässt sich allenfalls streichen/kürzen -> oh April2014-> streichen
 % Ist es die Bezeichnung offical publications sinnvoll? Oder besser govermental publications
@@
 
+% Bei SILC einsteigen. Allenfalls eine Grafik bringen mit den unterschiedlichen Reihen.
+% Daraus wird auch ersichtlich, dass keine der verfügbaren Reihen wirklich länge Zeiträume abdecken
+% we should mention that Eurostat uses EU-SILC alike and show, that there is a time series for SILC-Data aswell.
+Additionally we should mention that the Federal Statistical Office uses the HABE statistic
+
+% Allenfalls erwähnen, dass die Schweiz nicht Teil des GINI database ist.
+http://www.gini-research.org/articles/data_2
+
 % 5.Abschnitt: What do we know about Switzerland
@@
 (2009) found an increase in the 1980s).} \\
+
+%
 
@@
 But, up to the day information for Switzerland is not available.
+
 % ab hier wird es irgnediwe unklar
@@
 
-% the credit suisse wealth report already uses wealth tax records, but they use it only for the period from 2000-2011.
-They did exactly the same for Denmark and Norway, which could be interesting to compare.
-% Erbschaftssteuer erwähnen und bei den Global Wealth Reports der Credit Suisse und den Studien im Forschungsantrag weitermachen.
+% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
+% Dann lassen sich einige Passagen streichen.
+Man könnte auch auf verschiedene Publikationen verweisen, welche die Bedeutung von Vermögen untersuchen -> Buch von Piketty, Müller und Schoch und dann aber ziemlich schnell darauf verweisen, dass wir uns hier auf Einkommen konzentrieren.
+% Auch erste Dell Studie erwähnen -> es hat Zahlen zur Schweiz. Auch Probleme erwähnen (Fehlende Informationen zu Altersvorsorgevermögen -> hat aber niemand eine Lösung bisher.)
+% Studies on inequality have to adress several thorny challenges.
+One has to define the centrale measure of inequality and find adquate data.
+Recent studies accentuate the importance of a broader concept of economic wellbeeing including measures for income, consumption and wealth alike.
+But each area is confronted with own measurement problems.
+Measure of consumptions are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow.
+Measures of wealth are included in tax data, but are not connectable to the respctive income and therefore can only be analyzed separate.
+In danger of falling a pittraps it seems appriate to move in small steps.
+Therefore we focus our analysis on income, which is undoubtly a crucial indicator of economic wellbeeing.
+% Also the Federal Statistical Office is increasing their effort to get sound data concering the distribution of wealth vgl. Müller und Schoch (2014).
 
-Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study).
+Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study, EU-SILC).
 However, it is unclear how the bias through non-response affects the overall measure of inequality.
@@
 
+% delete passages about wealth and include passage about the comparison with survey data
+
 % Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.
 % rf: gute frage, spontan würde ich sagen eher nein, müssten wir mal diskutieren
+
+% Was wollen wir wissen?
+% Wir müssen uns äussern zu: 
+% Analyseinheiten
+% Definition des Einkommens
+% Berechnungsschritten
+
+% Bei den Ergebnissen müsste es Grafiken haben zu:
+% 1) Unterschiedlichen Analyseneinheiten (Mit und ohne Nuller, Normal- und Sonderfälle)
+% 2) Unterschiedliche Einkommensgrössen (Reineinkommen, steuerbares Einkommen, pseudoverfügbares Einkommen (Abzüge von direkten Bundessteuern))
+% 3) Vergleich mit HABE

### 2014-05-19 code/reldist.r
reldist plots

@@ -1,0 +1,74 @@
+##############################################################
+### Libraries ################################################
+##############################################################
+
+library(foreign)
+library(ggplot2)
+library(dplyr)
+library(rgl)
+library(reldist)
+
+##############################################################
+### Setup Data ###############################################
+##############################################################
+
+setwd("C:/Users/Hackstutz/Dropbox/Git/swiss_inequality_development/")
+estv<-read.dta("data/ginis_und_perzentile.dta")
+
+cpi <- read.csv('http://www.quandl.com/api/v1/datasets/WORLDBANK/CHE_FP_CPI_TOTL_ZG.csv?&trim_start=1961-12-31&trim_end=2011-12-31&sort_order=desc', colClasses=c('Date'='Date'))
+cpi$year <- as.numeric(substr(as.character(cpi$Date),1,4))
+cpi <- cpi[order(cpi$year),]
+cpi$cumInf <- cumprod(1+cpi$Value/100)
+# Referenzzeitpunkt 2009
+referenz09 <- cpi$cumInf[cpi$year==2009]
+estv$jahr <- ifelse(nchar(estv$steuerperiode)>4, estv$steuerperiode-0.5, estv$steuerperiode)
+estv$cumInf <- cpi$cumInf[match(estv$jahr,cpi$year)]
+estv[,which(names(estv)=="p1"):which(names(estv)=="p99")] <- estv[,which(names(estv)=="p1"):which(names(estv)=="p99")] / estv$cumInf*referenz09
+
+getDistribution <- function(years=2008, kanton="CH", data=estv) {
+	percentiles <- data %.% 
+		filter(kanton=="CH", steuerperiode%in%years) %.%
+		select(p1:p99)
+	res<-data.frame(t(rbind(as.numeric(gsub("[^0-9]","",names(percentiles))),percentiles)))
+	names(res)<-c("percentiles",as.character(years))
+	res
+}
+getDistribution(2003:2005)
+
+##############################################################
+### Analyses #################################################
+##############################################################
+
+relDistribution <- function(years=c(2003,2004,2005),rel=2004) {
+	stopifnot(rel%in%years,is.numeric(years))
+	data<-getDistribution(years=years)
+	data[,-1] <- data[,-1]/data[,as.character(rel)]
+	reshape(data,direction="long",timevar="year",varying=list(names(data)[-1]), times=names(data)[-1],v.names="value")
+}
+relDistribution()
+
+
[…]

### 2014-05-21 paper/subdocuments/introduction.Rnw
update minor changes

@@
 
+<<Swiss inequality from other studies>>=
+library(ggplot2)
+library(reshape)
+number_ticks <- function(n) {function(limits) pretty(limits, n)}
+swissGini_long<-melt(swissGini,id="Year")
+ggplot(data=swissGini_long,
+  aes(y=value,x=Year),shape=variable)+
+  ylab("Gini") +
+  scale_y_continuous(limits=c(0.18,0.4)) +
+  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
+  geom_point(aes(shape=variable),size=3)+
+  geom_line(data=swissGini_long[!is.na(swissGini_long$value),],aes(linetype=variable))+
+  theme_bw()
+@
+
+% Darauf verweisen, dass der Trend eher abwärts weist bzw. auf Stabilität. Aber auch darauf hinweisen, dass eine lange und konsistente Zeitreihe nicht vorliegt.
+
 % 5.Abschnitt: What do we know about Switzerland
@@
 
-% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
+% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is  know as the core driver of economic inequlity
 % Dann lassen sich einige Passagen streichen.
@@
 % Also the Federal Statistical Office is increasing their effort to get sound data concering the distribution of wealth vgl. Müller und Schoch (2014).
+
+% Zusammenfassung Stand der Forschung und festlegen unserer Ausrichtung
 

### 2014-05-21 paper/subdocuments/abstract.Rnw
Test
some notes and comments
update of abtracts and introduction

--- 2014-01-07 paper/subdocuments/abstract.Rnw
@@
 \begin{abstract}
+% There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
+In Switzerland, however, the development is unclear, as there is evidence for trends in both directions.
+Part of the inconclusive picture is due to different methodological approaches.
+In this paper we discuss the role of taxdata concerning the assessment of inequality in income.
+The focus of the discussion lays herein to show the benefits and shortcomings of tax data compared to current “state of the art” measurement concepts of economic inequality.
+We present common and new strategies to handle tax data specific methodological difficulties and compare results out of aggregated federal tax statistics to results from the Household Budget Survey (HBS).
+We can show to which extend survey data underestimate inequality in income.
+Following the results out of the taxdata Switzerland experienced in slight rise in inequality in recent years, similar to other western countries, but only because of rise in upper percentiles of the income distribution.
 
-There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
-In Switzerland, however, the development is unclear, as there is only little systematic evidence that would allow for long-term comparison.
-Nevertheless income inequality has been a prominent theme in the public discussion in recent years, e.g.
+% Salverda-Stuff und das neue Piketty Buch sollten irgendwo erwähnt werden
+
+\end{abstract}
+
+% Wir machen nichts mehr zu Vermögen (obwohl der Trend auch dahin geht, dass neben dem Einkommen auch das Vermögen eine zentrale Untersuchunsgrösse des inviduellen Wohlstandss darstellt)
+
+%
+
+% Alte version des abstractes. Fokus stärker auf Schweiz und Vermögen
+%There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
+In Switzerland, however, the development is unclear, as there is only little systematic evidence about income inequality that would allow for long-term comparison.
+Studies on the distribution of wealth are even scarcer.
+Nevertheless, income inequality has been a prominent theme in the public Swiss discussion in recent years, e.g.
 the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''.
-We adress the gap by presenting a long and consistent time series of inequality measures for income and wealth for Switzerland (1943-2010) calculated from federal tax data.
+We address the mentioned gap by presenting a long and consistent time series of inequality measures for income and wealth (1943-2010) calculated from federal tax statistics.
 We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties.
 In the end we integrate the case of Switzerland into the international picture of inequality development showing parallels and deviations.
-
-\end{abstract}

### 2014-05-21 paper/subdocuments/introduction.Rnw
update of abtracts and introduction

--- 2014-03-05 paper/subdocuments/introduction.Rnw
@@
 \section{Introduction}
-
-% Im European Sociologica Review scheint es unüblich zu sein, ohne theoretische Sektion zu arbeiten.
-Ich hoffe nicht, dass dies ein k.o.
-Kriterium ist.
-wir halten dafür die Einführungs und die Methoden Sektion etwas allgemeiner
 
@@
 While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders (WEF 2013) and scholars alike.
-Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013).
-The pattern of evolution suggests a polarization, which can be referred to as the "hollowing of the the middle class" (Alderson and Doran 2013).
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013;Salverda et al.
+2014).
+Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the "hollowing of the the middle class" (Alderson and Doran 2013).
 Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril."
 \\
+
+% Allenfalls könnte man auch (Nolan und Whealan, 2014) erwähnen.
+"The social Impact of income inequality: Poverty, Deprivation, and Social Cohesion" Da werden die Thesen rund um die Folgen von Ungleichheit gut zusammengefasst.
+Wobei: Jetzt scheint es mir jetzt ziemlich knackig und flüssig
 
@@
 
-% 3.Abschnitt: What do we know about Switzerland
-%
+% 3.Abschnitt: What do we know about Switzerland I
+% Offical Data Collections on development of inequality
 
+As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concering inequality in income.
 What is known about Switzerland so far?
-Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
-Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
-The debate went along with referendums trying to regulate the market outcomes.
-Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisingly high majority of 68\%, the 1:12-Initiative, which aimed at the whole spectrum of the income distribution, was rejected.
-The voting about the minimum wage-initiative will be held on 9th February 2014.
-All these referendums questioned the current distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
-Several official and semi-official publications addressed the question of rising inequality in Switzerland.
+Looking for offical data concering income inequality in Switzerland, three main sources has to be mentioned, which can be considered as de facto offical data sources: EU-SILC, HBS and LIS-data.
+Figure XY shows the Results steming from this three sources while looking at Gini of equivalised disposable income.
+Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level.
+The main focus of EU-SILC is to collect data on a common "framework" to ensure comparability among EU-countries and countries living around or within the EU.
+As a Non-EU member Switzerland implemented the instrument not from the beginning (2004) but as from 2007.
+Therefore this times-series doesn't cover timepoints before 2007.
+As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concering the distribution of income is the Household Budget Survey (HBS).
+The main focus of this survey lays in providing detailed data on houshold budegts allowing to apply different income definitions.
+Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistend time series from 2000 to 2011.
+As it can be seen from graph XY the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period.
+A longer period is covered in the LIS-Data-set (1982-2004).
+Data-provider for the LIS Data is the Swiss Federal Statistical Office.
+In contrast the the aformentioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+All in all the LIS dataset contains the longest time series on inequality for Switzerland.
+Analyzing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
+This result is supported by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).\footnote{A further offical database for income distribution is the OECD-Database.
+It includes measures from Income and Consumption survey as well.
+Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
+But this change in survey is considered as a strict break.
+Comparison before and since 2008 is not recommended (OECD 2012:315).
+A further important database on inequality is the GINI database which have been derived from the GINI Country Reports (Nolan et al.,2014).
+But this dataset doesn't cover Switzerland.
 \\
 
-% Lässt sich allenfalls streichen/kürzen
-% Ist es die Bezeichnung offical publications sinnvoll? Oder besser govermental publications
-% rf: oder "Federal publications"?
+<<Swiss inequality from other studies>>=
+library(ggplot2)
+library(reshape)
+number_ticks <- function(n) {function(limits) pretty(limits, n)}
+swissGini_long<-melt(swissGini,id="Year")
+ggplot(data=swissGini_long,
+  aes(y=value,x=Year),shape=variable)+
+  ylab("Gini of equivalised disposable income") +
+  scale_y_continuous(limits=c(0.18,0.4)) +
+  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
+  geom_point(aes(shape=variable),size=3)+
+  geom_line(data=swissGini_long[!is.na(swissGini_long$value),],aes(linetype=variable))+
+  theme_bw()
+@
 
-% 4.Abschnitt: What do we know about Switzerland
-% Offical Data Collections on development of inequality
+% Achtung: die HABE Zeitreihe muss upgedatet werden. Ist jetzt von Auge aus dem Bericht abgeschrieben.
 
-The most recent official figures published by the Swiss Federal Statistical Office (2013) show stability concerning the income distribution covering the time period of 1998 until 2011.
-The authors found a slight increase in the primary income which endured until the recent crisis.
-This temporary increase was mainly compensated by governmental redistribution.
-A longer period is covered in the LIS-Data-set (1982-2004).\footnote{Data-provider for the LIS Data is the Swiss Federal Statistical Office.
-The data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
-The OECD-Database includes measures from Income and Consumption survey as well.
-Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
-This change in survey is considered as a strict break.
-Comparison before and since 2008 is not recommended (OECD 2012:315).
-All in all the LIS dataset contains the longest time series on inequality for Switzerland.} Analyzing these data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
-This result is supported by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).
-\\
-
-% 5.Abschnitt: What do we know about Switzerland
+% 4.Abschnitt: What do we know about Switzerland II
 % Studies with tax data
 
-Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
+Whereas the aforementioned publications focused on equivalised disposable income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
 Dell et al.
@@
 timeline to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
-A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
+A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.
+Furthermore Schaltegger and Gorgas (2011) found a very heterogeneous evolution on the sub-federal level due to decentralized taxing power.
+\footnote{There are other studies on Switzerland covering different periods but not the recent years.
 Flückiger et al.
@@
 
-% 6. Abschnitt: Zwischenbilanz und Herausforderungen
+% 6. Abschnitt: Zwischenbilanz und Herausforderungen. Datenlage.
 
@@
 
-% Allenfalls hier noch nicht so ausführlich mit Vor- und Nachteilen von Steuerdaten kommen.
-Vielleicht besser darauf hinweisen, welche Dimensionen betroffen sind und das dies in der Methods Section besprochen wird.
-%rf: finde den abschnitt super
+% Zusammenfassung Stand der Forschung und festlegen unserer Ausrichtung
 
-While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
+Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study, EU-SILC).
+However, it is unclear how the bias through non-response affects the overall measure of inequality.
+Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) is missing.
+Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.
+First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
+Second, we summarize and apply tax data specific technics to construct suitable measures of inequality.
+We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
+Third, we compare the results with the HBS time series to get an idea of the survey bias throgh non-response.
+
+% Was wollen wir wissen?
+% Wir müssen uns äussern zu: 
+% Definition des Einkommens
+% Analyseinheiten
+% Berechnungsschritten
+
+%%%%
+% 0) In der Methods-Sektion beschreiben was wir machen? So was wie eine Forschungsstrategie/Vorgehensweise
+% Wie ausführlich?
+% Gini-Koeffizient, Relative Distribution-Methods. Berechnung?
+
+% Bei den Ergebnissen müsste es Grafiken haben zu:
+% 1) Unterschiedlichen Analyseneinheiten (Mit und ohne Nuller, Normal- und Sonderfälle)
+% 2) Unterschiedliche Einkommensgrössen (Reineinkommen, steuerbares Einkommen, pseudoverfügbares Einkommen (Abzüge von direkten Bundessteuern))
+% 3) alternative measure of income inequality: relative distribution and polaritätsindex
+% 4) Vergleich mit HABE
+
+%%%%%%%%%%%
+% Alte Textbausteine
+
+% Textbaustein, zur politischen Dimension von Einkommensungleichheit (alt)
+%Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+The debate went along with referendums trying to regulate the market outcomes.
+Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisingly high majority of 68\%, the 1:12-Initiative, which aimed at the whole spectrum of the income distribution, was rejected.
+The voting about the minimum wage-initiative will be held on 9th February 2014.
+All these referendums questioned the current distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
+Several official and semi-official publications addressed the question of rising inequality in Switzerland.
+\\
+
+% Textbaustein (unfertig) zu Vermögensungleichheit (alt)
+% While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
 The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
 But, up to the day information for Switzerland is not available.
-% ab hier wird es irgnediwe unklar
 Shorrock et al.
@@
 
-% the credit suisse wealth report already uses wealth tax records, but they use it only for the period from 2000-2011.
-They did exactly the same for Denmark and Norway, which could be interesting to compare.
-% Erbschaftssteuer erwähnen und bei den Global Wealth Reports der Credit Suisse und den Studien im Forschungsantrag weitermachen.
-
-Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study).
-However, it is unclear how the bias through non-response affects the overall measure of inequality.
-Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
-Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.
-First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
-Second, we summarize and apply tax data specific technics to construct suitable measures of inequality.
-We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
-Third, we integrate the development of Swiss inequality into the international picture.
-
-% Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.
-% rf: gute frage, spontan würde ich sagen eher nein, müssten wir mal diskutieren
+% Auch erste Dell Studie erwähnen -> es hat Zahlen zur Schweiz. Auch Probleme erwähnen (Fehlende Informationen zu Altersvorsorgevermögen -> hat aber niemand eine Lösung bisher.)

### 2014-05-21 paper/subdocuments/data_and_methods.Rnw
update method section

@@
 
-To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure.
-Based on a review on the literature about the measurement concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namely survey data.
+Studies on inequality have to adress several thorny challenges.
+It starts with answering three crucial questions: First of all one has to define, which concepts should to be looked at.
+This refers to answering the question about inequality of what.
+Secondly one has to be clear about the unit of analysis.
+This refers to answering the question about inequality among whom.
+Thirdly, one has to choose an appropriate measure of inequality.
+All this questions are ideally answered considering theory and a given research question.
+Often it has to be answered in context of a given dataset.
+Based on a review on the literature about the measurement concepts in an ideal world, we discuss  the advantages and shortcomings of tax data compared to other data sources - namely survey data.
 Afterwards we describe the tax data published by the Swiss Federal Tax Administration highlighting the important aspects, which have to be considered working with FTA-Data.
-In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for Switzerland.
+In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income for Switzerland.
 
@@
 \emph{Concepts on measuring economic resources}  \\
-Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together.
-In generall it has to be stateted, that ideal measures only can be defined in context of a given reaserach question.
-This holds, if one thinks about the ideal measure of economic well-being, but also about the unit of analysis.
-Nevertheless, some more or less unified concepts exists, which can be taken as a starting point.
-OECD (2013) \\
 
-% We neither can examine consumption nor the jointly distribution of income and wealth on the individual level.
-We could analyse the jointly pattern on the cantonal level and analze the data following the question if income inequality and wealth inqueality are correlated.
+A lot of studies on inequality focus on income inequality solely.
+However, recent activities emphasize the need of a broader conceptualization.
+A recent publication from the oecd (2013) condense these activities into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework of micro-level statistics.\footnote{Harmonisation with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framewrok presented in this publication.
+Considered main stadards were the System of National Accounts (SNA 2008), the canberra Group Handbook on Household Income Statistics (United Nations 2011), the final report of the 17th International Conference of Labour Statisticians (ILO 2004) and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing (UNECE & Eurostat 2006)} The framework concludes to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
+It is recomended to look at the distribution of all three distributions together to gain policy relevant insight.
+Some households with low income, for example, may report adequate levels of consumption expenditure or welth holdings, or vice-versa.
+But it is also stated (OED 2013:18):"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries." \\
+
+This holds for Swizterland too, although the HBS study is strongly influced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
+Albeit the awarness of an assessment of income, consumption and wealth simuntanously is rising, we  focus our analysis on income, which is undoubtly a crucial indicator of economic wellbeeing.
+However, the Federal Tax Office publishes statistics on income and wealth .
+But it is not possible to analyse the jointly distribution on the individual or household level.
+Also measures of consumption are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow.
 
@@
 % an income which is adjusted for the household size is not a real income anymore. It is a measure of indivdual welfare or of the standard of living (BFS:2012)
-
-\emph{Defining wealth}
-While the apropiateness of the conceptualization of income is widely discuded, there is no agreed definition of personal wealth yet and the appropriate methods of valuation are not always clear. 
-Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
 

### 2014-05-23 paper/subdocuments/data_and_methods.Rnw
methods section
completed subsection "concepts on measuring economic resources" and "defining income"

@@
 
-\section{Measurement concepts, Data and Methods}
+\section{ Data, Measurement concepts,and Methods}
 
@@
 
+% Allenfalls mit dem Beschrieb der Daten beginnen und bereits stärker auf die Möglichkeiten und Schwierigkeiten von Steuerdaten eingehen
+
 \subsection{Standards for measuring economic resources and inequality}
@@
 A recent publication from the oecd (2013) condense these activities into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework of micro-level statistics.\footnote{Harmonisation with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framewrok presented in this publication.
-Considered main stadards were the System of National Accounts (SNA 2008), the canberra Group Handbook on Household Income Statistics (United Nations 2011), the final report of the 17th International Conference of Labour Statisticians (ILO 2004) and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing (UNECE & Eurostat 2006)} The framework concludes to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
+Considered main stadards were the System of National Accounts (SNA 2008), the canberra Group Handbook on Household Income Statistics (United Nations 2011), the final report of the 17th International Conference of Labour Statisticians (ILO 2004) and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing (UNECE & Eurostat 2006)} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
 It is recomended to look at the distribution of all three distributions together to gain policy relevant insight.
 Some households with low income, for example, may report adequate levels of consumption expenditure or welth holdings, or vice-versa.
-But it is also stated (OED 2013:18):"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries." \\
+But it is also stated (OED 2013:18):"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by colecting and provding a database following ths broader concept of economic well-being.
+http://www.lisdatacenter.org/our-data/lws-database/} \\
 
-This holds for Swizterland too, although the HBS study is strongly influced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
+This last statment holds for Swizterland too, although the HBS study is strongly influced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
 Albeit the awarness of an assessment of income, consumption and wealth simuntanously is rising, we  focus our analysis on income, which is undoubtly a crucial indicator of economic wellbeeing.
@@
 
-\begin{itemize}
-\item Market income refers to revenue from employment.
-\item primary income or pre-transfer income refers to  market income plus revenue from property (capital income)
-\item Brutto income refers to  primary income plus social transfer
-\item Disposable income or posttransfer income refers to  brutto income minus transfer paid 
-\item adjusted disposable income refers to  disposable income corrected for the numbers of household members 
-\end{itemize}
+% Eine Grafik/Tabelle machen?
+% Mit den Brülhart-Daten hätte man noch mehr...
 
-The appropriateness of the income concept is bound to the research interest.
-Studies focusing on labor market outcome look at market income.
-Studies focusing on human wellbeing follow the approach as propagated by the oecd (2013).
-This concept is largely implemented by the Federal Office of statistics.
-Following this concept one should look at income after taxes and transfers inclusive the adjustment for household members, because this is the measure which frames the consumption possibilities.
-Lastly, studies focusing on the effect of institutions (social welfare, taxation) compare the distribution of pre- and post-income measures.
+The assessment of income inequality is influenced by the conceptual definition of the income itself.
+Marketincome or  disposable income for example differ by substantial meaning and be the expected degree of inequality.
+Therefore the awarness of the analyzed concept is crucial.
+Terminology can sligthly differ, while common concepts can be identified (for detailed discussion see: OECD (2013:44), Canberra group (2011:24)).
+Figure xy shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution.
+Within this framework common income definitions are situated.
+These definitions are contrasted with income concepts, which can be derived out of the FTA Tax data.
+
+\includegraphics[scale=0.5]{"/figure/income_definition.png}
+
+The central income according to tax data is the taxable income.
+It includes all reported incomes (income from employment, income from property and received transfers\footnote{Mean-tested benefits are not taxed and therefore not included in Tax data.
+Income for low income groups are therefore underestimated.}) minus several deductions.
+It is therefore neither a pre-transfer income  nor a post-transfer income measure.
+It's rather something inbetween.
+As the FTA tax statistitcs include some but not all deductions it is possible to calculate a sort of "total income".
+As some deducations can be interpreted as compulsary expenses similar to taxes to step towards the Total income is a step away from the income, which can be used for consumption.
+Similiar when calculating the disposable income out of the taxable income through accouting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
+Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches are missing.
 \\
 
-% an income which is adjusted for the household size is not a real income anymore. It is a measure of indivdual welfare or of the standard of living (BFS:2012)
+If the goal is to catch the income, which shapes the possibility to consume, equivalence scale are implemented.
+Equivalence scale followe the idea, that persons living in the same household have the possibility to share costs.
+To compare the individual economic well-beeing equivalence scales adjust for different household sizes (see OECD 2013,173, Buhmann et al., 1998).
+Because it is neither possible to calculate the total household income nor the respective householdsize out of the FTA tax data, we put this in this paper outside the brackets.
+Although knowing, that with doing so we neglect a potential aspect, which is capable of shaping the development of inequality.
 
-% Streichen
+%
 

### 2014-05-23 paper/subdocuments/introduction.Rnw
completed subsection "concepts on measuring economic resources" and "defining income"

@@
 What is known about Switzerland so far?
-Looking for offical data concering income inequality in Switzerland, three main sources has to be mentioned, which can be considered as de facto offical data sources: EU-SILC, HBS and LIS-data.
-Figure XY shows the Results steming from this three sources while looking at Gini of equivalised disposable income.
+Looking for offical data, three main sources has to be mentioned, which can be considered as de facto offical data sources: EU-SILC, HBS and LIS-data.
+Figure XY shows the results steming from this three sources while looking at Gini of equivalised disposable income.
 Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level.
@@
 As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concering the distribution of income is the Household Budget Survey (HBS).
-The main focus of this survey lays in providing detailed data on houshold budegts allowing to apply different income definitions.
+The main focus of this survey lays in providing detailed data on houshold budegts.
+This allows researcher to look at different income concepts like income before and after public transfers.
 Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistend time series from 2000 to 2011.
@@
 A longer period is covered in the LIS-Data-set (1982-2004).
-Data-provider for the LIS Data is the Swiss Federal Statistical Office.
+Data-provider for the LIS Data is the Swiss Federal Statistical Office too.
 In contrast the the aformentioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
@@
 A further important database on inequality is the GINI database which have been derived from the GINI Country Reports (Nolan et al.,2014).
-But this dataset doesn't cover Switzerland.
-\\
+But this dataset doesn't cover Switzerland.}  \\
 
@@
 
-The divergence can be explained with several factors.
+Divergence can be explained with several factors.
 First of all, different data sources were used.

### 2014-05-25 code/habe.r
first analyses with HABE data

@@ -1,0 +1,24 @@
+##############################################################
+### Libraries ################################################
+##############################################################
+
+library(foreign)
+library(ggplot2)
+library(dplyr)
+
+##############################################################
+### Setup Data ###############################################
+##############################################################
+
+setwd("C:/Users/Hackstutz/Dropbox/Git/swiss_inequality_development/")
+habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+habe0608<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE060708/HABE060708_Standard_130124UOe.txt", header=TRUE)
+names(habe0608)<-ifelse(names(habe0608)==names(habe0911),names(habe0608),names(habe0911))
+habe <- rbind(habe0608,habe0911)
+
+##############################################################
+### Analyses #################################################
+##############################################################
+
+ggplot(habe, aes(x=log(VerfuegbaresEinkommen08)))+geom_histogram()+facet_wrap(~Jahr08)
+

### 2014-05-25 paper/subdocuments/introduction.Rnw
fixed path

--- 2014-05-21 paper/subdocuments/introduction.Rnw
@@
 
-<<Swiss inequality from other studies>>=
+<<Swiss inequality Plot>>=
 library(ggplot2)
-library(reshape)
+library(foreign)
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
-swissGini_long<-melt(swissGini,id="Year")
-ggplot(data=swissGini_long,
-  aes(y=value,x=Year),shape=variable)+
+swissGini<-read.csv("../../data/swissGini.csv",header=TRUE,sep=";")
+ggplot(swissGini, aes(x=Year,y=LIS))+
+  geom_line(aes(x=Year,y=LIS))+
+  geom_point(aes(x=Year,y=LIS),size=3.5)+
+  xlab("Year") +
+  scale_x_continuous(breaks=number_ticks(10))+
   ylab("Gini") +
-  scale_y_continuous(limits=c(0.18,0.4)) +
-  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
-  geom_point(aes(shape=variable),size=3)+
-  geom_line(data=swissGini_long[!is.na(swissGini_long$value),],aes(linetype=variable))+
+  scale_y_continuous(limits=c(0,0.5)) +
   theme_bw()
 @
-
-% Darauf verweisen, dass der Trend eher abwärts weist bzw. auf Stabilität. Aber auch darauf hinweisen, dass eine lange und konsistente Zeitreihe nicht vorliegt.
 
@@
 
-% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is  know as the core driver of economic inequlity
+% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
 % Dann lassen sich einige Passagen streichen.

### 2014-05-25 paper/options/global_r_options.Rnw
fixes, so compiling Rnw works
merging minor diff between git and local

@@
 
+  # RF: I turn source code off
+  opts_chunk$set(echo=FALSE)
+
   # messages 
@@
   # Caching on by default
-  opts_chunk$set(cache=T)
+  opts_chunk$set(cache=FALSE)
   # plot options 

### 2014-05-25 paper/paper.tex
merging minor diff between git and local

@@
 
+% abstract müsste angepasst werden.
+% Wir machen nichts mehr zu Vermögen (obwohl der Trend auch dahin geht, dass neben dem Einkommen auch das Vermögen eine zentrale Untersuchunsgrösse des inviduellen Wohlstandss darstellt)
+
+% Alternative
+% There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
+In Switzerland, however, the development is unclear, as there is evidence for trends in both directions.
+Part of the inconclusive picture is due to different methodological approaches.
+In this paper we discuss the role of taxdata concerning the assessment of inequality in income.
+The focus of the discussion lays herein to show the benefits and shortcomings of tax data compared to current state of the art measurement concepts of economic inequality.
+We present common and new strategies to handle tax data specific methodological difficulties and compare results out of aggregated federal tax statistics to results from the Household Budget Survey (HBS).
+We can show to which extend survey data underestimate inequality in income.
+Following the results out of the taxdata Switzerland experienced in slight rise in inequality in recent years, similar to other western countries, but only because of rise in upper percentiles of the income distribution.
+
+% Salverda-Stuff und das neue Piketty Buch sollten irgendwo erwähnt werden
+%
+
 %----------------------------------------------------------------------------------------
@@
 
+% Bei SILC einsteigen. Allenfalls eine Grafik bringen mit den unterschiedlichen Reihen.
+% Daraus wird auch ersichtlich, dass keine der verfügbaren Reihen wirklich länge Zeiträume abdecken
+% we should mention that Eurostat uses EU-SILC alike and show, that there is a time series for SILC-Data aswell.
+Additionally we should mention that the Federal Statistical Office uses the HABE statistic
+
+% Allenfalls erwähnen, dass die Schweiz nicht Teil des GINI database ist.
+http://www.gini-research.org/articles/data_2
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/Swiss_inequality_Plot}
+
+\end{knitrout}
+
 % 5.Abschnitt: What do we know about Switzerland
@@
 
-% the credit suisse wealth report already uses wealth tax records, but they use it only for the period from 2000-2011.
-They did exactly the same for Denmark and Norway, which could be interesting to compare.
-% Erbschaftssteuer erwähnen und bei den Global Wealth Reports der Credit Suisse und den Studien im Forschungsantrag weitermachen.
+% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
+% Dann lassen sich einige Passagen streichen.
+Man könnte auch auf verschiedene Publikationen verweisen, welche die Bedeutung von Vermögen untersuchen -> Buch von Piketty, Müller und Schoch und dann aber ziemlich schnell darauf verweisen, dass wir uns hier auf Einkommen konzentrieren.
+% Auch erste Dell Studie erwähnen -> es hat Zahlen zur Schweiz. Auch Probleme erwähnen (Fehlende Informationen zu Altersvorsorgevermögen -> hat aber niemand eine Lösung bisher.)
+% Studies on inequality have to adress several thorny challenges.
+One has to define the centrale measure of inequality and find adquate data.
+Recent studies accentuate the importance of a broader concept of economic wellbeeing including measures for income, consumption and wealth alike.
+But each area is confronted with own measurement problems.
+Measure of consumptions are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow.
+Measures of wealth are included in tax data, but are not connectable to the respctive income and therefore can only be analyzed separate.
+In danger of falling a pittraps it seems appriate to move in small steps.
+Therefore we focus our analysis on income, which is undoubtly a crucial indicator of economic wellbeeing.
+% Also the Federal Statistical Office is increasing their effort to get sound data concering the distribution of wealth vgl. Müller und Schoch (2014).
 
-Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study).
+% Zusammenfassung Stand der Forschung und festlegen unserer Ausrichtung
+
+Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study, EU-SILC).
 However, it is unclear how the bias through non-response affects the overall measure of inequality.
@@
 
+% delete passages about wealth and include passage about the comparison with survey data
+
 % Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.
 % rf: gute frage, spontan würde ich sagen eher nein, müssten wir mal diskutieren
+
+% Was wollen wir wissen?
+% Wir müssen uns äussern zu: 
+% Analyseinheiten
+% Definition des Einkommens
+% Berechnungsschritten
+
+% Bei den Ergebnissen müsste es Grafiken haben zu:
+% 1) Unterschiedlichen Analyseneinheiten (Mit und ohne Nuller, Normal- und Sonderfälle)
+% 2) Unterschiedliche Einkommensgrössen (Reineinkommen, steuerbares Einkommen, pseudoverfügbares Einkommen (Abzüge von direkten Bundessteuern))
+% 3) Vergleich mit HABE
 
@@
 %@
+
+% Section from an older version. Currently disabled.
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# \\\#\# Attaching package: 'survey'\\\#\# \\\#\# Das folgende Objekt ist maskiert from 'package:graphics':\\\#\# \\\#\#\ \ \ \  dotchart}}\begin{verbatim}
 ## Wald test for kanton:zeros
-##  in lm(formula = G_steink ~ kanton + kanton:zeros + factor(steuerperiode) + 
-##     factor(steuerperiode):zeros, data = df)
+##  in lm(formula = G_steink ~ kanton + kanton:zeros + J + zeros:J, 
+##     data = df)
 ## F =  3.282  on  27  and  262  df: p= 4e-07

### 2014-05-25 paper/subdocuments/results.tex
merging minor diff between git and local

--- 2014-01-21 paper/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Journal Article
-% LaTeX Template
-% Version 1.3 (9/9/13)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Frits Wenneker (http://www.howtotex.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%  PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[twoside]{article}\usepackage[]{graphicx}\usepackage[]{color}
-%% maxwidth is the original width if it is less than linewidth
-%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
-\makeatletter
-\def\maxwidth{ %
-  \ifdim\Gin@nat@width>\linewidth
-    \linewidth
-  \else
-    \Gin@nat@width
-  \fi
-}
-\makeatother
-
-\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
-\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
-\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
-\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
-\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
-\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
-\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
-\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
-\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
-\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
-
-\usepackage{framed}
-\makeatletter
-\newenvironment{kframe}{%
- \def\at@end@of@kframe{}%
- \ifinner\ifhmode%
-  \def\at@end@of@kframe{\end{minipage}}%
-  \begin{minipage}{\columnwidth}%
- \fi\fi%
- \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
- \colorbox{shadecolor}{##1}\hskip-\fboxsep
-     % There is no \\@totalrightmargin, so:
-     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
- \MakeFramed {\advance\hsize-\width
-   \@totalleftmargin\z@ \linewidth\hsize
-   \@setminipage}}%
- {\par\unskip\endMakeFramed%
- \at@end@of@kframe}
-\makeatother
-
-\definecolor{shadecolor}{rgb}{.97, .97, .97}
-\definecolor{messagecolor}{rgb}{0, 0, 0}
-\definecolor{warningcolor}{rgb}{1, 0, 1}
-\definecolor{errorcolor}{rgb}{1, 0, 0}
-\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
-
-\usepackage{alltt}
-
-\usepackage{lipsum} % Package to generate dummy text throughout this template
-
-\usepackage[sc]{mathpazo} % Use the Palatino font
-%\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
-\usepackage[utf8]{inputenc}
-\linespread{1.05} % Line spacing - Palatino needs more space between lines
-\usepackage{microtype} % Slightly tweak font spacing for aesthetics
-
-\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
-\usepackage{multicol} % Used for the two-column layout of the document
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
-\usepackage{booktabs} % Horizontal rules in tables
-\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
-\usepackage{hyperref} % For hyperlinks in the PDF
-
-\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
-\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
-
-\usepackage{color} % used to mark parts that need to be edited
-
-\usepackage{abstract} % Allows abstract customization
-\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
-\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text
-
-\usepackage{titlesec} % Allows customization of titles
-\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
-\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
-\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
-\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
-
-\usepackage{fancyhdr} % Headers and footers
-\pagestyle{fancy} % All pages have headers and footers
-\fancyhead{} % Blank out the default header
-\fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
-\fancyfoot[RO,LE]{\thepage} % Custom footer text
-\bibliographystyle{plain}
-
-%%%-------------------------------------------------%%%
-%%% Preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document global preferences for Knitr %%%
-%%%-------------------------------------------------%%%
-
-%----------------------------------------------------------------------------------------
-%  TITLE SECTION
-%----------------------------------------------------------------------------------------
-
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}} % Article title
-
-\author{
-\large
-\textsc{Oliver Hümbelin}\\[2mm] % Your name
-\normalsize Bern University of Applied Sciences \\ % Your institution
-\normalsize \href{mailto:oliver.huembelin@bfh.ch}{oliver.huembelin@bfh.ch} % Your email address
-\vspace{5mm}\\
-\large
-\textsc{Rudolf Farys}\\[2mm] % Your name
-\normalsize University of Bern \\ % Your institution
-\normalsize \href{mailto:rudolf.farys@soz.unibe.ch}{rudolf.farys@soz.unibe.ch} % Your email address
-\vspace{-5mm}
-}
-\date{}
-
-%----------------------------------------------------------------------------------------
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-
-\begin{document}
-
-\maketitle
-
-\thispagestyle{fancy} % All pages have headers and footers
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-%%%-------------------------------------------------%%%
-%%% Include abstract %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document abstract %%%
-%%%-------------------------------------------------%%%
-
-\begin{abstract}
-
-There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
-In Switzerland, however, the development is unclear, as there is only little systematic evidence about income inequality that would allow for long-term comparison.
-Studies on the distribution of wealth are even scarcer.
-Nevertheless, income inequality has been a prominent theme in the public Swiss discussion in recent years, e.g.
-the most recent referendums "Abzockerinitiative'' and "1:12-Initiative''.
-We address the mentioned gap by presenting a long and consistent time series of inequality measures for income and wealth (1943-2010) calculated from federal tax statistics.
-We describe the benefits and shortcomings of tax data compared to other data sources and present strategies to handle tax data specific methodological difficulties.
-In the end we integrate the case of Switzerland into the international picture of inequality development showing parallels and deviations.
-
-\end{abstract}
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-\begin{multicols}{2} % Two-column layout throughout the main article text
-
-%%%-------------------------------------------------%%%
-%%% Include introduction %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document introduction %%%
-%%%-------------------------------------------------%%%
-
-\section{Introduction}
-
-% Im European Sociologica Review scheint es unüblich zu sein, ohne theoretische Sektion zu arbeiten.
-Ich hoffe nicht, dass dies ein k.o.
-Kriterium ist.
-wir halten dafür die Einführungs und die Methoden Sektion etwas allgemeiner
-
-% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern
-
-% 1.Abschnitt: Bedeutung von Ungleichheit
-
-Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular (Wilkinson and Pickett 2009).
-While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders (WEF 2013) and scholars alike.
-Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013).
-The pattern of evolution suggests a polarization, which can be referred to as the "hollowing of the the middle class'' (Alderson and Doran 2013).
-Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril.''
-\\
-
-% 2.Abschnitt: Bedeutung der Datengrundlage und Einführung von Steuerdaten
-
-Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
-Atkinson (2013:8) observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
-On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality undisputedly.
-These concerns have led to the search of alternative data sources, which can supplement the established survey data studies.
-Already Kuznet (1955) used tax data to examine the relationship between economic growth and personal distribution of income.
-Then it took several decades until Piketty (2001,2003) made the use of tax data fashionable again.
-Following his approach studies on several countries were conducted (Atkinson and Piketty 2007, 2010).
-Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database (Alvaredo et al.
-2014).
-\\
-
-% 3.Abschnitt: What do we know about Switzerland
-%
-
-What is known about Switzerland so far?
-Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
-Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
-The debate went along with referendums trying to regulate the market outcomes.
-Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisingly high majority of 68\%, the 1:12-Initiative, which aimed at the whole spectrum of the income distribution, was rejected.
-The voting about the minimum wage-initiative will be held on 9th February 2014.
-All these referendums questioned the current distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
-Several official and semi-official publications addressed the question of rising inequality in Switzerland.
-\\
-
-% Lässt sich allenfalls streichen/kürzen
-% Ist es die Bezeichnung offical publications sinnvoll? Oder besser govermental publications
-% rf: oder "Federal publications"?
-
-% 4.Abschnitt: What do we know about Switzerland
-% Offical Data Collections on development of inequality
-
-The most recent official figures published by the Swiss Federal Statistical Office (2013) show stability concerning the income distribution covering the time period of 1998 until 2011.
-The authors found a slight increase in the primary income which endured until the recent crisis.
-This temporary increase was mainly compensated by governmental redistribution.
-A longer period is covered in the LIS-Data-set (1982-2004).\footnote{Data-provider for the LIS Data is the Swiss Federal Statistical Office.
-The data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
-The OECD-Database includes measures from Income and Consumption survey as well.
-Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
-This change in survey is considered as a strict break.
-Comparison before and since 2008 is not recommended (OECD 2012:315).
-All in all the LIS dataset contains the longest time series on inequality for Switzerland.} Analyzing these data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
-This result is supported by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).
-\\
-
-% 5.Abschnitt: What do we know about Switzerland
-% Studies with tax data
-
-Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
-Dell et al.
-(2007) used tax data from the Federal Tax Administration to assess the development of concentration of the highest incomes and wealth (top-shares).
-In contrast to most other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
-Using the same approach Foellmi and Martinez (2013) expand the Dell et al.
-timeline to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
-A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
-Flückiger et al.
-(2007) and also Jeitziner and Peters (2007,2009) report constant inequality from 1960-1996 respectively from 1995 to 2003.
-Covering a similar time period Bauer and Spycher (1994) and Bolzani and Abul Naga (2002) found decreasing inequality.
-On the other hand Buchmann and Sacchi (1995) and Ernst et al.
-(2009) found an increase in the 1980s).} \\
-
-% 6. Abschnitt: Zwischenbilanz und Herausforderungen
-
-The divergence can be explained with several factors.
-First of all, different data sources were used.
-The official data providers trust on survey data, whereas the later mentioned publications use tax data.
-It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality.
-On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the "hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
-Different measure of inequality hampers the comparability.
-Third, different income concepts and different units of analysis were used.
-As it is shown by Modetta and Müller (2012) income distribution is strongly affected by governmental redistribution, reducing inequality substantially.
-With the focus on tax data the change in institutional settings is not covered.
-Also neglected is the household structure, whereas it is unclear how inequality is affected whether one looks at household income or at income of tax units.
-It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
-\\
-
-% Allenfalls hier noch nicht so ausführlich mit Vor- und Nachteilen von Steuerdaten kommen.
-Vielleicht besser darauf hinweisen, welche Dimensionen betroffen sind und das dies in der Methods Section besprochen wird.
-%rf: finde den abschnitt super
-
-While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
-The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
-But, up to the day information for Switzerland is not available.
-% ab hier wird es irgnediwe unklar
-Shorrock et al.
-(2013) unify several datasources (including tax data) to assess the pattern of global wealth.
-Their databook includes figures for Switzerland.
-In contrast to inequality of income concerning inequality of wealth Switzerland takes a leading position \\
-
-% the credit suisse wealth report already uses wealth tax records, but they use it only for the period from 2000-2011.
-They did exactly the same for Denmark and Norway, which could be interesting to compare.
-% Erbschaftssteuer erwähnen und bei den Global Wealth Reports der Credit Suisse und den Studien im Forschungsantrag weitermachen.
-
-Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study).
-However, it is unclear how the bias through non-response affects the overall measure of inequality.
-Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
-Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.
-First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
-Second, we summarize and apply tax data specific technics to construct suitable measures of inequality.
-We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
-Third, we integrate the development of Swiss inequality into the international picture.
-
-% Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.
-% rf: gute frage, spontan würde ich sagen eher nein, müssten wir mal diskutieren
-
-%%%-------------------------------------------------%%%
-%%% How did inequality change over time %%%
-%%%-------------------------------------------------%%%
-
-%<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
-%@
-
-%%%-------------------------------------------------%%%
-%%% Include data and methods %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for data and methods %%%
-%%%-------------------------------------------------%%%
-
-\section{Measurement concepts, Data and Methods}
-
-To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure.
-Based on a review on the literature about the measurement concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namely survey data.
-Afterwards we describe the tax data published by the Swiss Federal Tax Administration highlighting the important aspects, which have to be considered working with FTA-Data.
-In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for Switzerland.
-
-\subsection{Standards for measuring economic resources and inequality}
-
-\emph{Concepts on measuring economic resources}  \\
-Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together. \\
-
-% We neither can examine consumption nor the jointly distribution of income and wealth on the individual level.
-We could analyse the jointly pattern on the cantonal level and analze the data following the question if income inequality and wealth inqueality are correlated.
-
-\emph{Defining income}
-
-\begin{itemize}
-\item Market income refers to revenue from employment.
-\item primary income or pre-transfer income refers to  market income plus revenue from property (capital income)
-\item Brutto income refers to  primary income plus social transfer
-\item Disposable income or posttransfer income refers to  brutto income minus transfer paid 
-\item adjusted disposable income refers to  disposable income corrected for the numbers of household members 
-\end{itemize}
-
-The appropriateness of the income concept is bound to the research interest.
-Studies focusing on labor market outcome look at market income.
-Studies focusing on human wellbeing follow the approach as propagated by the oecd (2013).
-This concept is largely implemented by the Federal Office of statistics.
-Following this concept one should look at income after taxes and transfers inclusive the adjustment for household members, because this is the measure which frames the consumption possibilities.
-Lastly, studies focusing on the effect of institutions (social welfare, taxation) compare the distribution of pre- and post-income measures.
-\\
-
-\emph{Defining wealth}
-While the apropiateness of the conceptualization of income is widely discuded, there is no agreed definition of personal wealth yet and the appropriate methods of valuation are not always clear. 
-Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
-
-\emph{Population Coverage} \\
-All Residents? Working Residents? Single Person vs household/family. Households without income \\
-see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\
-
-\emph{Measuring inequality and concentration} \\
-To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
-Vgl.
-Allison (1978), Engelhardt (2000), Cowell (2000) oder Hao/Naiman (2010)
-%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
-
-neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences (Handcock/Morris 1999) \\
-
-Gini is mainly used for international comparison.
-However, other measures are often reported along (gini is more sensitive to changes in the middle of the distribution than to changes in the tails (OECD, 2008:37).
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) (Alderson and Doran 2013).
-\\
-
-Alvaredo (2010) shows formally, how top incomes shares are related to the gini-coefficient.
-Following his argument, it is crucial to consider top incomes.
-Leigh (2007) also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the studie).
-%[He also discusses tax data studies in general-> have a closer look] \\
-%[We definitely have to argue, why it isnt sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
-
-\subsection{Comparison of tax data and other data sources - advantages and shortcomings}
-
-%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
-% Mögliche Vergleichsdimensionen
-%- Suitability of cross country comparison
-%- Assessing inequality accurately
-%- Implementation of common concepts of economic ressources
-%- contingency for intertemporal comparision
-
-%Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
-
-%Warnings about the use of tax data
-%The use of tax data is often regarded by economists with considerable disbelief.
-These doubts are well justified for at least two reasons.
-The first is that tax data are collected as part of an administrative process, which is not tailored to the scientists' needs, so that the definition of income, income unit, etc., are not necessarily those that we would have chosen.
-This causes particular difficulties for comparisons across countries, but also for time-series analysis where there have been substantial changes in the tax system, such as the moves to and from the joint taxation of couples.
-Secondly, it is obvious that those paying tax have a financial incentive to present their affairs in a way that reduces tax liabilities.
-There is tax avoidance and tax evasion.
-The rich, in particular, have a strong incentive to understate their taxable incomes.
-Those with wealth take steps to ensure that the return comes in the form of asset appreciation, typically taxed at lower rates or not at all.
-Those with high salaries seek to ensure that part of their remuneration comes in forms, such as fringe benefits or stock-options which receive favorable tax treatment.
-Both groups may make use of tax havens that allow income to be moved beyond the reach of the national tax net.
-These shortcomings limit what can be said from tax data, but this does not mean that the data are worthless.
-Like all economic data, they measure with error the 'true' variable in which we are interested.
-%The data series presented here are fairly homogenous across countries, annual, long-run, and broken down by income source for several cases.
-Users should be aware also about their limitations.
-Firstly, the series measure only top income shares and hence are silent on how inequality evolves elsewhere in the distribution [why?].
-Secondly, the series are largely concerned with gross incomes before tax.
-Thirdly, the definition of income and the unit of observation (the individual vs.
-the family) vary across countries making comparability of levels across countries more difficult.
-Even within a country, there are breaks in comparability that arise because of changes in tax legislation affecting the definition of income, although most studies try to correct for such changes to create homogenous series.
-Finally and perhaps most important, the series might be biased because of tax avoidance and tax evasion.
-For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).
-
-%Copied from:
-%Alvaredo, Facundo, Anthony B. Atkinson, Thomas Piketty and Emmanuel Saez, The World Top Incomes Database, http://topincomes.g-mond.parisschoolofeconomics.eu/ , 18/12/2013
-
-\textbf{Summary of tax data drawbacks}
-\begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial income see above)
-\item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the top-income literature, albeit it is a crucial issue in theorized concepts of income.
-To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
-This might influence the overall inequality, taking into account the change from traditional household and family structures over the last century.
-This is an inequality related issue, where relevant studies are missing.
-What are the assumptions?
-\item the tax data income definition summarize labor income, capital income and taxable transfer payments (no means-tested benefits as welfare aid).
-Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
-\item Some tax units are not included in the tax tables (see Foellmi and Martinez 2013:11f).(a) only taxed cases are included (partly a problem, for certain periodes we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
-Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
-In some border-cantons the share of this group is very high (20\%)
-%Can we ignore this)
-(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registred.
-This persons get an imputed income (older tax tax return and information given by employers.
-Not registered non-	fillers are not in the records.
-- (e) tax evasion.
-Feld and Frey (2007) report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
-With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequaly measures over time.
-\end{itemize}
-
-\textbf{Problems with household income surveys}
-\begin{itemize}
-\item Sample data (bias)
-\item comparability between countries and over time (depends on income definition)
-\item short time series
-\end{itemize}
-
-Atkinson et al. (2009) estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by Alvaredo (2010). \\
-
-See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
-%		What about the special cases?
-Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
-Exclusion leads to an underestimation of inequality
-%		Capital gains? Should not be included 
-The discussion about problems with reporting income is fairly exhaustive. What about wealth?
-
-It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
-Credit Suisse (2011) makes use of the information in the "Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
-\\
-
-Additonaly the Credit Suisse Reports states (2011:8) that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts.
-
-\subsection{Tax data published by the Swiss Federal Tax Administration}
-
-%Beschreiben was die Daten beinhalten und wo es Probleme gibt
-\begin{itemize}
-\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
-	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
-	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
-	\item	Income after deductions-> taxable income (employment income, business income and capital income).
-Is income definition stable over time?
-Yes it should.
-Realized capital gains are excluded from the definition.
-It includes income from  employment, self-employment, capital income and taxable transfer payments.
-Plus Eigenmietwert,
-	\item	Reported on national and cantonal level
-\end{itemize}
-
-\subsection{Ways to tackle FTA-tax data specific problems}
-
-\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
-Del et al.
-(2007) impute for non-fillers the 20 percentage of the annual average income.
-This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
-Furthermore Del et al.
-calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
-\\
-
-\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
-This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
-This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f).
-\\
-%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.
-
-\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
-
-\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
-%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
-
-\textbf{deductions} Del et al.
-(2007:477): we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates.
-Gorgas and Schaltenegger (2011:5): ..., information on [] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time.
-Können wir das auch?
-Zumindest für gewisse Zeiträume?
-Das wäre noch gut.
-\\
-
-Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
-Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.
-On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a perverse redistribution effect by redistributing income from the lower middle class to the upper middle class (vgl.
-Peters 2011 and Interpellation Barbara Gysel (2009).
-
-% Income share specific problems
-% Total income denominator Exogenous Approach -> net income reported in the national accounts.
-Endogenous Approach -> Dell et al.
-technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income).
-Honestly , is this appropriate?
-(Everyone did it)
-% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
-% Endogenous approach -> reported in the tax tables
-
-%%%-------------------------------------------------%%%
-%%% Abschnitte aus ersten Version des Papers %%%
-%%%-------------------------------------------------%%%
-
-%\subsection{Hypotheses}
-%Based on the theories we test the following hypotheses:
-
-%\begin{itemize}
-%\item H1: Develpment of inequality is driven by sectoral change
-%\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
-%\end{itemize}
-
-%\subsection{Data and Variables}
-
-%We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010.
-While the data results in a long and consistent time series to illustrate swiss inequality development, there are a few pitfalls we want to adress which might be of interest for other research on this topic (be it in Switzerland or other countries).
-
-%\subsubsection{Left censored data}
-
-%The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household.
-The taxable population however is not identical to the population which should be used to calculate measures of inequality.
-Precisely, the data do not contain tax units with very little incomes so calculations based on these data treat the lowest percentiles equally to tax units with zero income.Figure X shows the threshold to be hit to enter the statistic.
-
-%\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}
-
-%So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data.
-Furthermore, also the changes over time might not be interpreted savely as over time the number of tax unit within this "hidden range" might vary or might even have a certain trend.
-We will adress this issue in detail in the methods chapter.
-
-%\subsubsection{Different measures, different populations}
-
-%The FTA data makes two kinds of distinctions.
-First, data was collected for so called "normal cases" and "special cases", i.e.
-a "normal case" is a taxable (for the complete tax period) person or household domiciled in a swiss canton without income from outside of Switzerland.
-A "special case" therefore is a diffuse reference category that contains tax units that are taxed at source, were not taxable for the complete tax period or generated additional income in another country.
-Second, the FTA reports two measures, that is taxable income and absolute income \textcolor{red}{(meine vorläufige Übersetzung von Reineinkommen)}.
-Absolute income is the sum of all incomes (earnings, interest income, rental incomes) minus expenses (e.g.
-from self-employment or credit cost).
-The taxable income is calculated as the difference of absolute income and deductions (e.g.
-children, insurance rates).
-The longest consistent time series exists for the taxable income of normal cases.
-So all statements we make with our data only apply to this subpopulation.
-
-%\subsubsection{Changes in taxation and measurement}
-%The swiss tax system is highly federal.
-That means, communities raise taxes which then go to the communities, the canton and the state.
-If we want to calculate overall swiss measures, we need to take into account, that cantons vary (between cantons and over time) with regard to the tax deductions that are possible and also the mechanism how taxes are collected.
-The latter adresses a comprehensive reorganization of the swiss tax system where between 1995 and 2003 cantons changed from taxing the past two years of income (postnumerando system) to taxing the present single year (praenumerando system).
-For details see \textcolor{red}{Martinez (xxxx) or some other author (xxxx)}.
-Aggregate measures of inequality therefore have to be estimated for the periods 1995 to 2003 which we adress shortly in the methods chapter.
-
-%\subsection{Methodology Used}
-%There are two steps of data analyses which need to be described to the reader.
-First, the estimation of the bias we introduce by estimating measures of inequality when tax units with too little income are not observed.
-Second, the steps undertaken to estimate aggregate swiss measures by imputing taxable income for those cantons and periods where the change of the tax system produced a gap (1995 to 2003, depending on the canton).
-
-%\subsubsection{Imputing the gap}
-%The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx).
-That is estimating the missing taxable income statistics via OLS using information from time trends and cantons.
-Our imputation model therefore includes canton inequality measures and periods dummies to explain aggregate swiss inequality.
-
-\subsubsection{Estimating the bias}
-For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation (lets call those "zeros" for convenience) .
-However starting 1995, the FTA provides exactly this information for each canton.
-This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010.
-Consequently we can obtain information whether the bias is stable over time (which makes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
-\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie "je höher der Steuerfreibetrag, umso stärker der Bias".
-Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias.
-Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
-
-One can take a first look at the descriptives plotting the share of zeros over time seperately for each canton .
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/zero_descriptives1}
-
-\includegraphics[width=\maxwidth]{figure/zero_descriptives2}
-
-\end{knitrout}
-
-We can see multiple things here:
-
-\begin{enumerate}
-\item There is a small overall upward trend which we assume to be the Federal Administrations inflation adjustments to the tax threshold.
-\item Geneva and Tessin show wild changes but those might be explained by the tax gap ("Bemessungslücke") that people exploited when the cantons changed the tax system.
-It remains unclear however why we can't see similar patter within other cantons.
-\item There is some variance and we see different patterns over time and cantons.
-When estimating gini coefficients or the like we must therefore assume that ignoring "the zeros" leads to a bias that is not stable over time.
-\end{enumerate}
-
-We might try two strategies to moderate the problem:
-
-\begin{enumerate}
-\item Add the zeros as a seperate group
-\item Fit a model to predict the inequality measure (e.g. the gini coefficient) using the share of zeros as a predictor
-\end{enumerate}
-
-By adding the zeros as a seperate group we face several problems.
-With the exception of Geneva 1995/96 Pareto interpolation of percentiles p20 and above seems viable.
-Low percentiles however would need to be extrapolated so we might impose unrealistic assuptioms.
-Even when estimating "safe" measures above p20 (better p50) or a gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
-To be consistent with the measure of "taxable income" we could assume an income of zero for that group, calculate gini coefficients and compare them with the original (uncorrected/plain) version.
-To get a rough number we could (cantonwise) check the squared correlation between plain and corrected gini.
-
-The second approach however is more robust as we do not impose additional assumptions but instead exercise a little curve fitting.
-
-The model outputs a teststatisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
-As the model has a decent fit we are not in great danger of omitted variable bias.
-Using a joint F-Test we can now test if all canton#zeros interactions are zero.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## Wald test for kanton:zeros
-##  in lm(formula = G_steink ~ kanton + kanton:zeros + factor(steuerperiode) + 
-##     factor(steuerperiode):zeros, data = df)
-## F =  3.282  on  27  and  262  df: p= 4e-07
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-In our case we can clearly reject the hypothesis that all interactions are zero.
-This leads to the conclusion  that gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted gini coefficients.
-For example one might be interested in how inequality would had developed if the zero-share would have been constant over time.
-(Note RF: predict all data point using canton, time and the initial OR final zero-share to homogenize the time series)
-Furthermore we can quantify how large the bias is and we can do this seperately for tax periods or seperately for cantons.
-
-For all cantons:
-
-We can see the the model fit reduces to explaining 61.5\% of the gini variation versus 92.6\% when the information from the zero-shares was used.
-Although this so some extent attributable to the additional 26 paramters: this is huge.
-
-The model indicates some cases that deserve more attention: Schwyz (positive coefficient) and Geneva (negative coefficient) and the tax period 2000 as well as the most recent periods.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## [1] -0.006575
-## [1] -0.005752
-## [1] 0.09164
-## [1] 0.08027
-## [1] 0.07055
-## [1] 0.06316
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-A positive coefficient (e.g.
-Schwyz) can be read as follows: In periods with many zeros we measure higher gini coefficients.
-Our conclusion is that people with low income are more successfully hidden.
-If those people would be taxed and therefore appear in the statistic they would sharply contrast the high income earners.
-
-A negative coefficient (e.g.
-Geneva) means the more zeros there are the smaller the gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
-This is less obvious than the Schwyz case and our reading is: more zeros mask inequality that arises from the bottom.
-
-\subsubsection{Descriptives - where do we put this?}
-nice grapics for taxable income, reines einkommen and income after taxes here...
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/different_ginis1}
-
-\includegraphics[width=\maxwidth]{figure/different_ginis2}
-
-\end{knitrout}
-
-%%%-------------------------------------------------%%%
-%%% Include results %%%
-%%%-------------------------------------------------%%%
-
 %%%-------------------------------------------------%%%
@@
 Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
+
+\subsection{Gini coefficients for taxable income, income before deductions and taxable income after tax}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(foreign)}
+\hlkwd{library}\hlstd{(ggplot2)}
+\hlstd{df} \hlkwb{<-} \hlkwd{read.dta}\hlstd{(}\hlstr{"../../data/ginis_und_perzentile.dta"}\hlstd{)}
+\hlstd{df}\hlopt{$}\hlstd{zeros} \hlkwb{<-} \hlstd{df}\hlopt{$}\hlstd{null_norm}\hlopt{/}\hlstd{df}\hlopt{$}\hlstd{cpop}  \hlcom{# i only use normal cases here}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{df}\hlopt{$}\hlstd{G_reink[df}\hlopt{$}\hlstd{G_reink} \hlopt{==} \hlnum{1}\hlstd{]} \hlkwb{<-} \hlnum{NA}
+\hlstd{df}\hlopt{$}\hlstd{G_taxed[df}\hlopt{$}\hlstd{G_taxed} \hlopt{==} \hlnum{1}\hlstd{]} \hlkwb{<-} \hlnum{NA}
+\hlkwd{library}\hlstd{(reshape)}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: plyr\\\#\# \\\#\# Attaching package: 'reshape'\\\#\# \\\#\# Die folgenden Objekte sind maskiert from 'package:plyr':\\\#\# \\\#\#\ \ \ \  rename, round\_any}}\begin{alltt}
+\hlstd{dflong} \hlkwb{<-} \hlkwd{melt}\hlstd{(df[,} \hlkwd{c}\hlstd{(}\hlstr{"G_steink"}\hlstd{,} \hlstr{"G_reink"}\hlstd{,} \hlstr{"G_taxed"}\hlstd{,} \hlstr{"kanton"}\hlstd{,} \hlstr{"steuerperiode"}\hlstd{)],}
+    \hlkwc{id} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"kanton"}\hlstd{,} \hlstr{"steuerperiode"}\hlstd{))}
+\hlkwd{names}\hlstd{(dflong)[}\hlnum{3}\hlstd{]} \hlkwb{<-} \hlstr{"Type"}
+\hlkwd{names}\hlstd{(dflong)[}\hlnum{4}\hlstd{]} \hlkwb{<-} \hlstr{"Gini"}
+\hlkwd{levels}\hlstd{(dflong}\hlopt{$}\hlstd{Type)} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"taxable income"}\hlstd{,} \hlstr{"income before deductions"}\hlstd{,} \hlstr{"taxable income after tax"}\hlstd{)}
+\hlkwd{ggplot}\hlstd{(dflong[dflong}\hlopt{$}\hlstd{kanton} \hlopt{==} \hlstr{"CH"}\hlstd{, ],} \hlkwd{aes}\hlstd{(}\hlkwc{x} \hlstd{= steuerperiode,} \hlkwc{y} \hlstd{= Gini,} \hlkwc{colour} \hlstd{= Type))} \hlopt{+}
+    \hlkwd{geom_line}\hlstd{(}\hlkwc{size} \hlstd{=} \hlnum{1}\hlstd{)} \hlopt{+} \hlkwd{theme_bw}\hlstd{()} \hlopt{+} \hlkwd{xlab}\hlstd{(}\hlstr{"tax period"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/different_ginis1} 
+\begin{kframe}\begin{alltt}
+\hlkwd{ggplot}\hlstd{(dflong,} \hlkwd{aes}\hlstd{(}\hlkwc{x} \hlstd{= steuerperiode,} \hlkwc{y} \hlstd{= Gini,} \hlkwc{colour} \hlstd{= Type))} \hlopt{+} \hlkwd{geom_line}\hlstd{()} \hlopt{+}
+    \hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{kanton)} \hlopt{+} \hlkwd{theme_bw}\hlstd{()} \hlopt{+} \hlkwd{xlab}\hlstd{(}\hlstr{"tax period"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/different_ginis2}
+
+\end{knitrout}
+
+\subsection{Zero share}
+
+Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later)
+
+\subsection{Income measures with and without zeros}
+
+taxable income with and without zeros
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{dflong} \hlkwb{<-} \hlkwd{melt}\hlstd{(df[,} \hlkwd{c}\hlstd{(}\hlstr{"G_steink"}\hlstd{,} \hlstr{"G_steink0"}\hlstd{,} \hlstr{"kanton"}\hlstd{,} \hlstr{"steuerperiode"}\hlstd{)],}
+    \hlkwc{id} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"kanton"}\hlstd{,} \hlstr{"steuerperiode"}\hlstd{))}
+\hlkwd{names}\hlstd{(dflong)[}\hlnum{3}\hlstd{]} \hlkwb{<-} \hlstr{"Type"}
+\hlkwd{names}\hlstd{(dflong)[}\hlnum{4}\hlstd{]} \hlkwb{<-} \hlstr{"Gini"}
+\hlkwd{levels}\hlstd{(dflong}\hlopt{$}\hlstd{Type)} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"without zeros"}\hlstd{,} \hlstr{"including zeros"}\hlstd{)}
+\hlkwd{ggplot}\hlstd{(dflong[dflong}\hlopt{$}\hlstd{kanton} \hlopt{==} \hlstr{"CH"} \hlopt{&} \hlstd{dflong}\hlopt{$}\hlstd{steuerperiode} \hlopt{>=} \hlnum{1995.5}\hlstd{, ],} \hlkwd{aes}\hlstd{(}\hlkwc{x} \hlstd{= steuerperiode,}
+    \hlkwc{y} \hlstd{= Gini,} \hlkwc{colour} \hlstd{= Type))} \hlopt{+} \hlkwd{geom_line}\hlstd{(}\hlkwc{size} \hlstd{=} \hlnum{1}\hlstd{)} \hlopt{+} \hlkwd{theme_bw}\hlstd{()} \hlopt{+} \hlkwd{xlab}\hlstd{(}\hlstr{"tax period"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/with_without_zeros1} 
+\begin{kframe}\begin{alltt}
+\hlkwd{ggplot}\hlstd{(dflong[dflong}\hlopt{$}\hlstd{steuerperiode} \hlopt{>=} \hlnum{1995.5}\hlstd{, ],} \hlkwd{aes}\hlstd{(}\hlkwc{x} \hlstd{= steuerperiode,} \hlkwc{y} \hlstd{= Gini,}
+    \hlkwc{colour} \hlstd{= Type))} \hlopt{+} \hlkwd{geom_line}\hlstd{()} \hlopt{+} \hlkwd{facet_wrap}\hlstd{(}\hlopt{~}\hlstd{kanton)} \hlopt{+} \hlkwd{theme_bw}\hlstd{()} \hlopt{+} \hlkwd{xlab}\hlstd{(}\hlstr{"tax period"}\hlstd{)} \hlopt{+}
+    \hlkwd{opts}\hlstd{(}\hlkwc{axis.text.x} \hlstd{=} \hlkwd{theme_text}\hlstd{(}\hlkwc{angle} \hlstd{=} \hlnum{60}\hlstd{))}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# 'opts' is deprecated.
+Use 'theme' instead.
+(Deprecated; last used in version 0.9.1)\\\#\# theme\_text is deprecated.
+Use 'element\_text' instead.
+(Deprecated; last used in version 0.9.1)}}\end{kframe}
+\includegraphics[width=\maxwidth]{figure/with_without_zeros2}
+
+\end{knitrout}
+
+Including zeros leads to significantly higher gini coefficients.
+However we must keep in mind, that these might be artifically high values as we assume zero income for everyone in the zero group.
+We can conclude more from the graphic: the ratio between both measures seems to be quite constant although for aggregate Switzerland but there are minor deviations for multiple cantons as well as strong deviations for the cantons Geneva and Tessin.
+However the problems seem not to result from a shift in the zero-share over time but they are specific for the time-period when the tax system changed.
 
@@
 %rf: genau. kantone unterstreichen die probleme mit dem 0er bias und die probleme mit "imputing the gap". mehr muss erstmal nicht sein.
-
-%%%-------------------------------------------------%%%
-%%% Include discussion %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for discussion %%%
-%%%-------------------------------------------------%%%
-
-\section{Discussion}
-
-% Hier könnte allenfalls der Bogen zur Theorie gespannt werden
-
-%Methodological Main findings 
-%(1) Tax data is useful because its possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
-Its superior to survey data, because the latter is affected by non-responded bias.
-Tax data includes data about the whole population  [was finden wir wegen den Nullern raus] .
-or at least it should (tax evasion).(3) The FTA-Tax data has different shortcomings, which cannot be handled like a measure of income and wealth which (a) cannot be corrected for household size and (b) is neither a pre-transfer nor a post-transfer income measure.
-Therefor its comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
-This FTA-Tax data problem can be handled with non-aggregated tax-data .
-
-%Main findings for Switzerland (1) Our data suggests that income inequality overall slightly increased in Switzerland and we might distinguish three episodes: 1.
-Substantial increase in times of economic growth 1950 to 1974 ending with the oil crisis 2.
-Ups and downs from 1970 to 2000 3.
-Relatively steep increase in inequality in the early 2000s
-%(2) Despite the recent increase, overall income inequality in Switzerland is not very high in international comparison.
-With respect to inequality in wealth, however, Switzerland takes a leading position in the world.
-%(3) While the overall Swiss inequality remained pretty stable over decades, inequality between cantons underwent a bizarre development.
-
-%%%-------------------------------------------------%%%
-%%% Include acknowledgements %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for acknowledgement %%%
-%%%-------------------------------------------------%%%
-
-\section{Acknowledgements}
-
-We thank Ben Jann, Robert Fluder and Tobias Fritschi for helpful comments on the article. We also like to thank Stefan Ilic for the prepartion of the data set.
-
-%%%-------------------------------------------------%%%
-%%% Include the bibliography %%%
-%%%-------------------------------------------------%%%
-
-\end{multicols}
-
-\bibliography{bibliography/bib}
-
-%%%-------------------------------------------------%%%
-%%% Include the appendix %%%
-%%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document for appendix %%%
-%%%-------------------------------------------------%%%
-
-\section{Appendix}
-
-\end{document}

### 2014-05-25 code/data_and_methods.R

@@ -1,0 +1,73 @@
+
+## ----setup---------------------------------------------------------------
+library(foreign)
+df <- read.dta("../../data/ginis_und_perzentile.dta")
+df$zeros <- df$null_norm/df$cpop # i only use normal cases here
+
+
+## ----zero_descriptives---------------------------------------------------
+library(ggplot2)
+ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)
+ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")
+
+
+## ----corrected_gini------------------------------------------------------
+df$J<-factor(df$steuerperiode)
+fit <- lm(G_steink~kanton+kanton:zeros+J+zeros:J,data=df)
+#summary(fit)
+SmoothCoefficientPlot <- function(models, modelnames = "", removeintercept = FALSE){
+  # models must be a list()
+ 
+  Alphas <- seq(1, 99, 2) / 100
+ 
+  Multiplier <- qnorm(1 - Alphas / 2)
+  zzTransparency <<- 1/(length(Multiplier)/4)
+  #CoefficientTables <- lapply(models, function(x){summary(x)$coef})
+  CoefficientTables <- summary(models)$coef
+  #TableRows <- unlist(lapply(CoefficientTables, nrow))
+  TableRows <- nrow(CoefficientTables)
+ 
+  if(modelnames[1] == ""){
+    ModelNameLabels <- rep(paste("Model", 1:length(TableRows)), TableRows)
+    } else {
+    ModelNameLabels <- rep(modelnames, TableRows)
+    }
+ 
+  MatrixofModels <- cbind(do.call(rbind, list(CoefficientTables)), ModelNameLabels)
+  if(removeintercept == TRUE){
+    MatrixofModels <- MatrixofModels[!rownames(MatrixofModels) == "(Intercept)", ]
+    }
+  MatrixofModels <- data.frame(cbind(rownames(MatrixofModels), MatrixofModels))
+ 
+  MatrixofModels <- data.frame(cbind(MatrixofModels, rep(Multiplier, each = nrow(MatrixofModels))))
+ 
[…]

### 2014-05-25 code/kongress.R

@@ -1,0 +1,344 @@
+
+## ----options, results='hide',echo=FALSE----------------------------------
+opts_chunk$set(out.width = '800px', dev='svg', out.height = '500px',fig.width=10,fig.height=6.25,cache=TRUE, echo=FALSE, warning=FALSE)
+
+
+## ----datenaufbereitung, echo=FALSE, results='hide',warning=FALSE,--------
+library(plm)
+library(foreign)
+library(ggplot2)
+setwd("C:/Users/Hackstutz/Dropbox/")
+daten <- read.dta("PX-Daten/kontextvariablen0.02.dta")
+
+cpi <- read.csv('http://www.quandl.com/api/v1/datasets/WORLDBANK/CHE_FP_CPI_TOTL_ZG.csv?&trim_start=1961-12-31&trim_end=2011-12-31&sort_order=desc', colClasses=c('Date'='Date'))
+cpi$year <- as.numeric(substr(as.character(cpi$Date),1,4))
+cpi <- cpi[order(cpi$year),]
+cpi$cumInf <- cumprod(1+cpi$Value/100)
+# Referenzzeitpunkt 2009 (für die Interpretation der y-Skalen)
+referenz09 <- cpi$cumInf[cpi$year==2009]
+
+daten$kj <- paste(daten$Kantonname, daten$Jahr)
+daten$bev <- daten$frau_ausl+daten$frau_schw+daten$mann_schw+daten$mann_ausl
+kbev<-aggregate(bev~Kantonname+Jahr, data=daten,FUN="sum")
+kbev$kj <- paste(kbev$Kantonname, kbev$Jahr)
+daten$kanton_bev<-kbev[match(daten$kj,kbev$kj),"bev"]
+  
+daten <- within(daten, {
+  cumInf <- cpi[match(Jahr, cpi$year),"cumInf"]/referenz09
+  real_mean_reink <- mean_reink / cumInf
+  real_median_reink <- median_reink / cumInf
+  real_mean_steink <- mean_steink / cumInf
+  real_median_steink <- median_steink / cumInf
+  real_mean_reinka <- mean_reinka / cumInf
+  real_median_reinka <- median_reinka / cumInf
+  real_mean_steinka <- mean_steinka / cumInf
+  real_median_steinka <- median_steinka / cumInf
+  bev <- frau_ausl+frau_schw+mann_schw+mann_ausl
+  ausl <- (frau_ausl+mann_ausl)/bev
[…]

### 2014-05-25 code/habe.r
first analyses with HABE data

@@ -8,7 +8,7 @@ library(dplyr)
 
 ##############################################################
 ### Setup Data ###############################################
-##############################################################
+############################################################## 
 
 setwd("C:/Users/Hackstutz/Dropbox/Git/swiss_inequality_development/")
 habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE)

### 2014-05-26 code/habe.r
first HABE analyses

@@ -16,9 +16,15 @@ habe0608<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE060708/HA
 names(habe0608)<-ifelse(names(habe0608)==names(habe0911),names(habe0608),names(habe0911))
 habe <- rbind(habe0608,habe0911)
 
+habe <- habe %.%
+	group_by(Jahr08) %.%
+	mutate(percentiles = percent_rank(VerfuegbaresEinkommen08))
+
 ##############################################################
 ### Analyses #################################################
 ##############################################################
 
 ggplot(habe, aes(x=log(VerfuegbaresEinkommen08)))+geom_histogram()+facet_wrap(~Jahr08)
-
+ggplot(habe, aes(y=log(VerfuegbaresEinkommen08),x=percentiles))+geom_line()+facet_wrap(~Jahr08)
+ggplot(habe, aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,color=factor(Jahr08)))+geom_density()
+ggplot(filter(habe,Jahr08%in%c(2006,2008)), aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,color=factor(Jahr08)))+geom_density()

### 2014-05-26 code/habe.r
some analyses with HABE

@@ -5,6 +5,7 @@
 library(foreign)
 library(ggplot2)
 library(dplyr)
+library(reldist)
 
 ##############################################################
 ### Setup Data ###############################################
@@ -27,4 +28,8 @@ habe <- habe %.%
 ggplot(habe, aes(x=log(VerfuegbaresEinkommen08)))+geom_histogram()+facet_wrap(~Jahr08)
 ggplot(habe, aes(y=log(VerfuegbaresEinkommen08),x=percentiles))+geom_line()+facet_wrap(~Jahr08)
 ggplot(habe, aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,color=factor(Jahr08)))+geom_density()
-ggplot(filter(habe,Jahr08%in%c(2006,2008)), aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,color=factor(Jahr08)))+geom_density()
+ggplot(filter(habe,Jahr08%in%c(2006,2008)), aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,fill=factor(Jahr08)))+geom_density(alpha=0.5)
+
+reldist(y=habe$VerfuegbaresEinkommen08[habe$Jahr08==2008],yo=habe$VerfuegbaresEinkommen08[habe$Jahr08==2006])
+reldist(y=habe$VerfuegbaresEinkommen08[habe$Jahr08==2011],yo=habe$VerfuegbaresEinkommen08[habe$Jahr08==2006])
+reldist(y=habe$VerfuegbaresEinkommen08[habe$Jahr08==2011],yo=habe$VerfuegbaresEinkommen08[habe$Jahr08==2010])

### 2014-05-26 paper/subdocuments/introduction.Rnw
Revert "fixed path"
This reverts commit 459ae9fc7eed8b5dac503500875567fd70e49ae8.

@@
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
-swissGini<-read.csv("../../data/swissGini.csv",header=TRUE,sep=";")
+swissGini<-read.csv("data/swissGini.csv",header=TRUE,sep=";")
 ggplot(swissGini, aes(x=Year,y=LIS))+

### 2014-05-26 code/habe.r
Theme_bw

@@ -25,10 +25,10 @@ habe <- habe %.%
 ### Analyses #################################################
 ##############################################################
 
-ggplot(habe, aes(x=log(VerfuegbaresEinkommen08)))+geom_histogram()+facet_wrap(~Jahr08)
-ggplot(habe, aes(y=log(VerfuegbaresEinkommen08),x=percentiles))+geom_line()+facet_wrap(~Jahr08)
-ggplot(habe, aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,color=factor(Jahr08)))+geom_density()
-ggplot(filter(habe,Jahr08%in%c(2006,2008)), aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,fill=factor(Jahr08)))+geom_density(alpha=0.5)
+ggplot(habe, aes(x=log(VerfuegbaresEinkommen08)))+geom_histogram()+facet_wrap(~Jahr08)+theme_complete_bw()
+ggplot(habe, aes(y=log(VerfuegbaresEinkommen08),x=percentiles))+geom_line()+facet_wrap(~Jahr08)+theme_complete_bw()
+ggplot(habe, aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,color=factor(Jahr08)))+geom_density()+theme_complete_bw()
+ggplot(filter(habe,Jahr08%in%c(2006,2008)), aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,fill=factor(Jahr08)))+geom_density(alpha=0.5)+theme_complete_bw()
 
 reldist(y=habe$VerfuegbaresEinkommen08[habe$Jahr08==2008],yo=habe$VerfuegbaresEinkommen08[habe$Jahr08==2006])
 reldist(y=habe$VerfuegbaresEinkommen08[habe$Jahr08==2011],yo=habe$VerfuegbaresEinkommen08[habe$Jahr08==2006])

### 2014-05-26 code/normal_sonderfaelle.r
gini_und_percentile data

@@ -1,0 +1,24 @@
+##############################################################
+### Libraries ################################################
+##############################################################
+
+library(foreign)
+library(ggplot2)
+library(dplyr)
+
+##############################################################
+### Setup Data ###############################################
+##############################################################
+
+setwd("C:/Users/Hackstutz/Dropbox/Git/swiss_inequality_development/")
+daten<-read.csv("data/ginis_und_perzentile_normal.csv")
+
+##############################################################
+### Analyses #################################################
+##############################################################
+
+
+
+##############################################################
+### Export Results ###########################################
+##############################################################

### 2014-05-26 code/habe.r
Signed-off-by: hackstutz <rudi@farys.de>

@@ -25,10 +25,10 @@ habe <- habe %.%
 ### Analyses #################################################
 ##############################################################
 
-ggplot(habe, aes(x=log(VerfuegbaresEinkommen08)))+geom_histogram()+facet_wrap(~Jahr08)+theme_complete_bw()
-ggplot(habe, aes(y=log(VerfuegbaresEinkommen08),x=percentiles))+geom_line()+facet_wrap(~Jahr08)+theme_complete_bw()
-ggplot(habe, aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,color=factor(Jahr08)))+geom_density()+theme_complete_bw()
-ggplot(filter(habe,Jahr08%in%c(2006,2008)), aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,fill=factor(Jahr08)))+geom_density(alpha=0.5)+theme_complete_bw()
+ggplot(habe, aes(x=log(VerfuegbaresEinkommen08)))+geom_histogram()+facet_wrap(~Jahr08)+theme_bw()
+ggplot(habe, aes(y=log(VerfuegbaresEinkommen08),x=percentiles))+geom_line()+facet_wrap(~Jahr08)+theme_bw()
+ggplot(habe, aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,color=factor(Jahr08)))+geom_density()+theme_bw()
+ggplot(filter(habe,Jahr08%in%c(2006,2008)), aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,fill=factor(Jahr08)))+geom_density(alpha=0.5)+theme_bw()
 
 reldist(y=habe$VerfuegbaresEinkommen08[habe$Jahr08==2008],yo=habe$VerfuegbaresEinkommen08[habe$Jahr08==2006])
 reldist(y=habe$VerfuegbaresEinkommen08[habe$Jahr08==2011],yo=habe$VerfuegbaresEinkommen08[habe$Jahr08==2006])

### 2014-05-27 paper/subdocuments/abstract.Rnw
fix abstract to Olis commit
minor correction abstract

@@
 \begin{abstract}
-% There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
+There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
 In Switzerland, however, the development is unclear, as there is evidence for trends in both directions.

### 2014-05-27 paper/subdocuments/introduction.Rnw
minor corrections introdcution

--- 2014-05-23 paper/subdocuments/introduction.Rnw
@@
 
-% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern
-
 % 1.Abschnitt: Bedeutung von Ungleichheit
-
 Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular (Wilkinson and Pickett 2009).
@@
 "The social Impact of income inequality: Poverty, Deprivation, and Social Cohesion" Da werden die Thesen rund um die Folgen von Ungleichheit gut zusammengefasst.
-Wobei: Jetzt scheint es mir jetzt ziemlich knackig und flüssig
 
 % 2.Abschnitt: Bedeutung der Datengrundlage und Einführung von Steuerdaten
-
 Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
@@
 
-% 3.Abschnitt: What do we know about Switzerland I
+% 3.Abschnitt: What do we know about Switzerland
 % Offical Data Collections on development of inequality
-
 As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concering inequality in income.
@@
 Comparison before and since 2008 is not recommended (OECD 2012:315).
-A further important database on inequality is the GINI database which have been derived from the GINI Country Reports (Nolan et al.,2014).
+A further important database on inequality is the GINI database which has been derived from the GINI Country Reports (Nolan et al.,2014).
 But this dataset doesn't cover Switzerland.}  \\
 
-<<Swiss inequality from other studies>>=
+<<Swiss inequality Plot>>=
 library(ggplot2)
-library(reshape)
+library(foreign)
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
+swissGini<-read.csv("data/swissGini.csv",header=TRUE,sep=";")
 swissGini_long<-melt(swissGini,id="Year")
@@
 
-% Achtung: die HABE Zeitreihe muss upgedatet werden. Ist jetzt von Auge aus dem Bericht abgeschrieben.
-
-% 4.Abschnitt: What do we know about Switzerland II
+% 4.Abschnitt: What do we know about Switzerland
 % Studies with tax data
-
-Whereas the aforementioned publications focused on equivalised disposable income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
+Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
 Dell et al.
@@
 timeline to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
-A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.
-Furthermore Schaltegger and Gorgas (2011) found a very heterogeneous evolution on the sub-federal level due to decentralized taxing power.
-\footnote{There are other studies on Switzerland covering different periods but not the recent years.
+A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
 Flückiger et al.
@@
 
-% 6. Abschnitt: Zwischenbilanz und Herausforderungen. Datenlage.
-
+% 5. Abschnitt: Zwischenbilanz und Herausforderungen
 Divergence can be explained with several factors.
@@
 However, it is unclear how the bias through non-response affects the overall measure of inequality.
-Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) is missing.
+Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
 Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.
@@
 We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
-Third, we compare the results with the HBS time series to get an idea of the survey bias throgh non-response.
+Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through datasource.
+
+% Könnte noch besser auf die geplanten Analysen abgestimmt sein (vgl. auf Github-Issues)
 
 % Was wollen wir wissen?
+
 % Wir müssen uns äussern zu: 
+% Analyseinheiten
 % Definition des Einkommens
-% Analyseinheiten
-% Berechnungsschritten
-
-%%%%
-% 0) In der Methods-Sektion beschreiben was wir machen? So was wie eine Forschungsstrategie/Vorgehensweise
-% Wie ausführlich?
-% Gini-Koeffizient, Relative Distribution-Methods. Berechnung?
+% Messung der Verteilung
 
@@
 % 2) Unterschiedliche Einkommensgrössen (Reineinkommen, steuerbares Einkommen, pseudoverfügbares Einkommen (Abzüge von direkten Bundessteuern))
-% 3) alternative measure of income inequality: relative distribution and polaritätsindex
-% 4) Vergleich mit HABE
+% 3) Vergleich mit HABE
 
-%%%%%%%%%%%
-% Alte Textbausteine
-
-% Textbaustein, zur politischen Dimension von Einkommensungleichheit (alt)
-%Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+% Textbaustein politische Dimension von Ungleichheit
+% What is known about Switzerland so far?
+Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
+Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
 The debate went along with referendums trying to regulate the market outcomes.
@@
 
-% Textbaustein (unfertig) zu Vermögensungleichheit (alt)
-% While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
+% Alte Textbausteine zu Vermögen
+%While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
 The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
 But, up to the day information for Switzerland is not available.
-Shorrock et al.
+% ab hier wird es irgnediwe unklar
+%Shorrock et al.
 (2013) unify several datasources (including tax data) to assess the pattern of global wealth.
@@
 In contrast to inequality of income concerning inequality of wealth Switzerland takes a leading position \\
-
+% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
+% Dann lassen sich einige Passagen streichen.
+Man könnte auch auf verschiedene Publikationen verweisen, welche die Bedeutung von Vermögen untersuchen -> Buch von Piketty, Müller und Schoch und dann aber ziemlich schnell darauf verweisen, dass wir uns hier auf Einkommen konzentrieren.
 % Auch erste Dell Studie erwähnen -> es hat Zahlen zur Schweiz. Auch Probleme erwähnen (Fehlende Informationen zu Altersvorsorgevermögen -> hat aber niemand eine Lösung bisher.)

### 2014-05-27 paper/subdocuments/data_and_methods.Rnw
minor changes data and method section

@@
 
-\section{ Data, Measurement concepts,and Methods}
+\section{Data, Measurement concepts,and Methods}
 
@@
 Often it has to be answered in context of a given dataset.
+Therefore we start this section with a description of the FTA-Tax Data.
 Based on a review on the literature about the measurement concepts in an ideal world, we discuss  the advantages and shortcomings of tax data compared to other data sources - namely survey data.
-Afterwards we describe the tax data published by the Swiss Federal Tax Administration highlighting the important aspects, which have to be considered working with FTA-Data.
-In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income for Switzerland.
+Along with the despriction we explain the methods and thechnics we use to construct time-series of inequality-measure for income for Switzerland, which we will present in the result sections.
+By doing this, we demonstrate the suitability of the FTA Tax Data to assess the development of inequality in Switzerland
 
-% Allenfalls mit dem Beschrieb der Daten beginnen und bereits stärker auf die Möglichkeiten und Schwierigkeiten von Steuerdaten eingehen
+\subsection{Tax statistics in Swizterland}
 
@@
 
-A lot of studies on inequality focus on income inequality solely.
+Most studies on inequality focus on income inequality solely.
 However, recent activities emphasize the need of a broader conceptualization.
-A recent publication from the oecd (2013) condense these activities into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework of micro-level statistics.\footnote{Harmonisation with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framewrok presented in this publication.
+A recent publication from the oecd (2013) condense these ideas into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework on micro-level statistics.\footnote{Harmonisation with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framewrok presented in this publication.
 Considered main stadards were the System of National Accounts (SNA 2008), the canberra Group Handbook on Household Income Statistics (United Nations 2011), the final report of the 17th International Conference of Labour Statisticians (ILO 2004) and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing (UNECE & Eurostat 2006)} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
-It is recomended to look at the distribution of all three distributions together to gain policy relevant insight.
+To gain policy relevant insight, it is recomended to look at the distribution of all three distributions simultaneously.
 Some households with low income, for example, may report adequate levels of consumption expenditure or welth holdings, or vice-versa.
@@
 
-% Eine Grafik/Tabelle machen?
-% Mit den Brülhart-Daten hätte man noch mehr...
-
-The assessment of income inequality is influenced by the conceptual definition of the income itself.
-Marketincome or  disposable income for example differ by substantial meaning and be the expected degree of inequality.
+The assessment of income inequality is influenced by the definition of the income itself.
+Marketincome or disposable income for example differ by substantial meaning and by the expected degree of inequality.
 Therefore the awarness of the analyzed concept is crucial.
@@
 
+% Quelle zu Abzügen:
+% http://www.bilanz.ch/invest/steuern/netto-oder-reineinkommen
+% Die rechte Seite mit den Steuergrössen könnte ausführlicher beschrieben werden
+% Beispielsweise könnte man zeigen was Nettoeinkommen, Reineinkommen und steuerbares Einkommen sind
+% Allerdings steht uns nur Reineinkommen und steuerbares Einkommen zur Verfügung
+
 The central income according to tax data is the taxable income.
@@
 It's rather something inbetween.
-As the FTA tax statistitcs include some but not all deductions it is possible to calculate a sort of "total income".
-As some deducations can be interpreted as compulsary expenses similar to taxes to step towards the Total income is a step away from the income, which can be used for consumption.
+As the FTA tax statistitcs include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions.
+This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction,Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income".
+As some deducations can be interpreted as compulsary expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption.
 Similiar when calculating the disposable income out of the taxable income through accouting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
@@
 \\
+
+% Die Deductions sind noch nicht gänzlich sauber abgebildet.
+Auf der Bilanz-Seite steht, vom steuerbaren Einkommen kommt man, indem vom Reineinkommen die Sozialabzüge(Kinder- und Fremdbetreung) sowie der Abzug für unterstützungsbedürftige Personen abzogen werden.
+% Auf den ESTV-Tabellen sind jedoch noch mehr Abzüge abgebildet. Was soll/muss/können wir nun abziehen? Bin verwirrt...
 
@@
 
+% letzter Abschnitt muss noch auf die Brülhart-Daten hinweisen. Damit können wir eine äquivalenzeinkommen betrachten.
+% Auch müsste in diesem Abschnitt auf die Probleme der Haushaltsunterscheidung eingegangen werden (oder ist im Abschnitt Population coverage? Dann müsste obiger Punkt auch da erwähnt werden.)
+
+% Streichen
+=======
 % 
+>>>>>>> d21c323... completed subsection "concepts on measuring economic resources" and "defining income"
 
@@
 To simplify: cutting off zeros more and more seems to lead to an underestimation of our inequality measure, probably because the skewness in the left part of the distribution increased pointing to an increased pauperization that is masked by omitting the left tail of the income distribution.
+
+% Alter Textbaustein zu Vermögen
+%\emph{Defining wealth}
+%While the apropiateness of the conceptualization of income is widely discuded, there is no agreed definition of personal wealth yet and the appropriate methods of valuation are not always clear.
+Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).

### 2014-05-27 paper/subdocuments/data_and_methods.Rnw
title change
new Bib-Library
fixed double quotes and R-paths so paper.Rnw compiles

@@
 
-\section{Data, Measurement concepts,and Methods}
+\section{Data, Measurement concepts and Methods}
 
@@
 
-\includegraphics[scale=0.5]{"/figure/income_definition.png}
+\includegraphics[width=\maxwidth]{figure/income_definition}
 
@@
 % Die Deductions sind noch nicht gänzlich sauber abgebildet.
-Auf der Bilanz-Seite steht, vom steuerbaren Einkommen kommt man, indem vom Reineinkommen die Sozialabzüge(Kinder- und Fremdbetreung) sowie der Abzug für unterstützungsbedürftige Personen abzogen werden.
+Auf der Bilanz-Seite steht, zum steuerbaren Einkommen kommt man, indem vom Reineinkommen die Sozialabzüge(Kinder- und Fremdbetreung) sowie der Abzug für unterstützungsbedürftige Personen abzogen werden.
 % Auf den ESTV-Tabellen sind jedoch noch mehr Abzüge abgebildet. Was soll/muss/können wir nun abziehen? Bin verwirrt...
@@
 % Streichen
-=======
+%=======
 % 
->>>>>>> d21c323... completed subsection "concepts on measuring economic resources" and "defining income"
+%>>>>>>> d21c323... completed subsection "concepts on measuring economic resources" and "defining income"
 
@@
 \begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial "income" see above)
+\item Misreporting of incomes (high earners have an interest in getting tax beneficial ``income'' see above)
 \item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the "top-income" literature, albeit it is a crucial issue in "theorized" concepts of income.
+This is a point, which is hardly mentioned in the ``top-income'' literature, albeit it is a crucial issue in ``theorized'' concepts of income.
 To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
@@
 It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
-Credit Suisse (2011) makes use of the information in the "Rich Lists" published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
+Credit Suisse (2011) makes use of the information in the ``Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
 \\
@@
 (2007:477):" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
-Gorgas and Schaltenegger (2011:5): "..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time".
+Gorgas and Schaltenegger (2011:5): ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
 Können wir das auch?
@@
 On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a "perverse redistribution" effect by redistributing income from the lower middle class to the upper middle class (vgl.
+There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a ``perverse redistribution'' effect by redistributing income from the lower middle class to the upper middle class (vgl.
 Peters 2011 and Interpellation Barbara Gysel (2009).
@@
 \subsubsection{Estimating the bias}
-For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation (lets call those "zeros" for convenience) .
+For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation (lets call those ``zeros'' for convenience) .
 However starting 1995, the FTA provides exactly this information for each canton.
@@
 Consequently we can obtain information whether the bias is stable over time (which makes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
-\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie "je höher der Steuerfreibetrag, umso stärker der Bias".
+\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie ``je höher der Steuerfreibetrag, umso stärker der Bias''.
 Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias.
@@
 
-One can take a first look at the descriptives plotting the share of zeros over time seperately for each canton .
+One can take a first look at the descriptives plotting the share of zeros over time seperately for each canton.
 
@@
 library(foreign)
-df <- read.dta("../../data/ginis_und_perzentile.dta")
+df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
 df$zeros <- df$null_norm/df$cpop # i only use normal cases here
@@
 library(ggplot2)
-ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)
-ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")
+ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+theme_bw()
+ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")+theme_bw()
 @
@@
 \item There is some variance and we see different patterns over time and cantons.
-When estimating gini coefficients or the like we must therefore assume that ignoring "the zeros" leads to a bias that is not stable over time.
+When estimating gini coefficients or the like we must therefore assume that ignoring ``the zeros'' leads to a bias that is not stable over time.
 \end{enumerate}
@@
 Low percentiles however would need to be extrapolated so we might impose unrealistic assuptioms.
-Even when estimating "safe" measures above p20 (better p50) or a gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
-To be consistent with the measure of "taxable income" we could assume an income of zero for that group, calculate gini coefficients and compare them with the original (uncorrected/plain) version.
+Even when estimating ``safe'' measures above p20 (better p50) or a gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
+To be consistent with the measure of ``taxable income'' we could assume an income of zero for that group, calculate gini coefficients and compare them with the original (uncorrected/plain) version.
 To get a rough number we could (cantonwise) check the squared correlation between plain and corrected gini.
@@
 
-The model outputs a test statisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
+The model outputs a test statisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical ``canton gini-level''.
 As the model has a decent fit we are not in great danger of omitted variable bias.
-Using a joint F-Test we can now test if all canton#zeros interactions are zero.
+Using a joint F-Test we can now test if all canton interactions are zero.
 
-<<ftest>>=
-library(survey)
-regTermTest(fit,"kanton:zeros")
+<<ftest, echo=FALSE>>=
+library(survey, quietly=TRUE, warn.conflicts = FALSE)
+ftest <- regTermTest(fit,"kanton:zeros")
 @
 
-In our case we can clearly reject the hypothesis that all interactions are zero.
+In our case we can clearly reject the hypothesis that all interactions are zero ($p=\Sexpr{round(ftest$p,3)}$).
 This leads to the conclusion  that gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted gini coefficients.
@@
 This seems plausible.
-Most of the "tax optimization" happens within Switzerland so the rich people who moved to Schwyz are now missing at another canton.
+Most of the ``tax optimization'' happens within Switzerland so the rich people who moved to Schwyz are now missing at another canton.
 

### 2014-05-27 paper/subdocuments/introduction.Rnw
fixed double quotes and R-paths so paper.Rnw compiles

@@
 2014).
-Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the "hollowing of the the middle class" (Alderson and Doran 2013).
-Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril."
+Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the ``hollowing of the the middle class'' (Alderson and Doran 2013).
+Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): ``our democracy is being put at peril.''
 \\
@@
 Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level.
-The main focus of EU-SILC is to collect data on a common "framework" to ensure comparability among EU-countries and countries living around or within the EU.
+The main focus of EU-SILC is to collect data on a common ``framework'' to ensure comparability among EU-countries and countries living around or within the EU.
 As a Non-EU member Switzerland implemented the instrument not from the beginning (2004) but as from 2007.
@@
 library(foreign)
+library(reshape, quietly=TRUE, warn.conflicts = FALSE)
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
-swissGini<-read.csv("data/swissGini.csv",header=TRUE,sep=";")
+swissGini<-read.csv("../../data/swissGini.csv",header=TRUE,sep=";")
 swissGini_long<-melt(swissGini,id="Year")
@@
 It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality.
-On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the "hollowing of the middle class" occurred in Switzerland or not, which leads to the second point.
+On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the ``hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
 Different measure of inequality hampers the comparability.
@@
 It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
-\\
 

### 2014-05-27 paper/subdocuments/results.Rnw
update minor changes
fixed double quotes and R-paths so paper.Rnw compiles

--- 2014-03-04 paper/subdocuments/results.Rnw
@@
 
-<<setup>>=
+<<setup_gini_data>>=
 library(foreign)
 library(ggplot2)
-df <- read.dta("data/ginis_und_perzentile.dta")
+df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
 df$zeros <- df$null_norm/df$cpop # i only use normal cases here
@@
 
-<<with_without_zeros>>=
+<<with_without_zeros,warning=FALSE>>=
 dflong <- melt(df[,c("G_steink","G_steink0","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
@@
 ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line(size=1)+theme_bw()+xlab("tax period")
-ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+ opts(axis.text.x=theme_text(angle=60))
+ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))
 

### 2014-06-02 paper/subdocuments/introduction.tex
Gini series out of HABE

--- 2014-01-07 paper/subdocuments/introduction.tex
@@
 
-% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern.
+% 1.Abschnitt: Bedeutung von Ungleichheit
+Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular (Wilkinson and Pickett 2009).
+While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders (WEF 2013) and scholars alike.
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013;Salverda et al.
+2014).
+Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the "hollowing of the the middle class" (Alderson and Doran 2013).
+Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril."
+\\
 
-% 1.Abschnitt: Bedeutung von Ungleichheit von Soziologischer Sicht
+% Allenfalls könnte man auch (Nolan und Whealan, 2014) erwähnen.
+"The social Impact of income inequality: Poverty, Deprivation, and Social Cohesion" Da werden die Thesen rund um die Folgen von Ungleichheit gut zusammengefasst.
 
-\emph{Bedeutung von ökonomischer Ungleicheit} \\
+% 2.Abschnitt: Bedeutung der Datengrundlage und Einführung von Steuerdaten
+Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
+Atkinson (2013:8) observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
+On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality undisputedly.
+These concerns have led to the search of alternative data sources, which can supplement the established survey data studies.
+Already Kuznet (1955) used tax data to examine the relationship between economic growth and personal distribution of income.
+Then it took several decades until Piketty (2001,2003) made the use of tax data fashionable again.
+Following his approach studies on several countries were conducted (Atkinson and Piketty 2007, 2010).
+Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database (Alvaredo et al.
+2014).
+\\
 
-% 2.Abschnitt Trends in inequality research
-\emph{Studien zur allgemeinen Entwicklung der Ungleichheit nach unterschiedliche Paradigmen}(Armut, Reiche, Mittelschicht)  [Concerns about the hollowing of the middle class is a argument why the tax studies should be exploited to gain knowledge about the whole distribution of economic resources.
-Until these days tax data is only used to assess the change in top income shares.] \\
-\begin{itemize}
-  \item Development of inequality with LIS Data (Gornick and Jäntti 2013)
-  \item ranking of inequality across high-income countries (Atkinson, Rainwater and Smeeding (1995)
-   \item inequality and low income/relative poverty (Ferreira and Ravallion, 2009; Nolan and Marx, 2009; Smeeding, O’Higgins and Rainwatter, 1990; Rainwater and Smeeding 2003; Förster and Vlemickx, 2004)
-  \item inequality and top income (Atkinson and Piketty, 2007; Leigh, 2009)
-  \item inequality and the middle class (Estache and Leipziger, 2009; Littrell et al., 2001; Birdsall, 2010; Ravallion, 2010; Frank,2007)
-\end{itemize}
+% 3.Abschnitt: What do we know about Switzerland
+% Offical Data Collections on development of inequality
+As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concering inequality in income.
+What is known about Switzerland so far?
+Looking for offical data, three main sources has to be mentioned, which can be considered as de facto offical data sources: EU-SILC, HBS and LIS-data.
+Figure XY shows the results steming from this three sources while looking at Gini of equivalised disposable income.
+Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level.
+The main focus of EU-SILC is to collect data on a common "framework" to ensure comparability among EU-countries and countries living around or within the EU.
+As a Non-EU member Switzerland implemented the instrument not from the beginning (2004) but as from 2007.
+Therefore this times-series doesn't cover timepoints before 2007.
+As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concering the distribution of income is the Household Budget Survey (HBS).
+The main focus of this survey lays in providing detailed data on houshold budegts.
+This allows researcher to look at different income concepts like income before and after public transfers.
+Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistend time series from 2000 to 2011.
+As it can be seen from graph XY the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period.
+A longer period is covered in the LIS-Data-set (1982-2004).
+Data-provider for the LIS Data is the Swiss Federal Statistical Office too.
+In contrast the the aformentioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+All in all the LIS dataset contains the longest time series on inequality for Switzerland.
+Analyzing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
+This result is supported by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).\footnote{A further offical database for income distribution is the OECD-Database.
+It includes measures from Income and Consumption survey as well.
+Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
+But this change in survey is considered as a strict break.
+Comparison before and since 2008 is not recommended (OECD 2012:315).
+A further important database on inequality is the GINI database which has been derived from the GINI Country Reports (Nolan et al.,2014).
+But this dataset doesn't cover Switzerland.}  \\
 
-% 3.Abschnitt: Studien mit Steuerdaten
-
-\emph{Studien mit Steuerdaten}
-\begin{itemize}
-\item Old studies with tax data (Vilfredo Pareto, 1895, 1896; Simon Kuznets, 1953)
-\item More recently revival of tax-studys with focus on top income (Piketty and Saez, 2003; Atkinson and Piketty, 2007,2010).
-\end{itemize}
-
-% 4.Abschnitt: What do we know about Switzerland?
-
-\emph{Official Data Collection} \\
-
-\begin{itemize}
-\item  LIS (1982-2004, Survey Data from Swiss Federal Statistical Office three different surveys (recent years from “income and Consumption Survey, main result; decreasing inequality)
-\item  OECD (2000 und 2004, Income and consumption survey, 2008, EU-SILC. The change of survey in 2008 is consdered as a strict break. Comparison over time not possible (OECD 2012:315)
-\item  Worldbank: Nothing
-\item  Federal Office of statistics (1998-2011, HABE, survey data, very detailed measurement, main result; inequality in post transfer income is constant)
-\item  World income inequality Database (WIID) is more or less the same data as it is published in the LIS data
-\end{itemize}
-\emph{Fazit}:LIS longest time series, comparability is questionable, more recent years time series from Federal Office of statistics.
-Both studies use survey data, which, as we will discuss later, have certain drawbacks.\\
-
-\emph{Studies on Switzerland} \\
-Vgl. Forschungsantrag (kürzen) plus neuere Studien, Verteilungsbericht 2011 und 2012 (SGB), Avenir Suisse (2013) und Gorgas and Schaltenegger (2011) und Foellmi and Martinez (2013). \\
-
-% Should we discus the non-scientic literature?
-
-% What is the purpose of our paper?
-The purpose of the following paper is threefold.
-First of all, information about economic inequality in Switzerland is sparse.
-We  fill this gap with official data from the Swiss Federal Tax Administration (FTA).
-Second, we assess the suitability of the FTA data to report inequality and its changes over time.
-Third, we integrate the development of swiss inequality into the international picture.
-We can show similarities and differences to the worldwide trends and therefore contribute to the classical theories (\cite{kuznets1955economic}, \cite{nielsen_kuznets_1997}).
+\begin{Schunk}
+\begin{Sinput}
+> library(ggplot2)
+> library(foreign)
+> number_ticks <- function(n) {function(limits) pretty(limits, n)}
+> swissGini<-read.csv("data/swissGini.csv",header=TRUE,sep=";")

### 2014-06-02 code/habe.r

@@ -6,6 +6,8 @@ library(foreign)
 library(ggplot2)
 library(dplyr)
 library(reldist)
+library(doBy)
+
 
 ##############################################################
 ### Setup Data ###############################################
@@ -20,6 +22,43 @@ habe <- rbind(habe0608,habe0911)
 habe <- habe %.%
 	group_by(Jahr08) %.%
 	mutate(percentiles = percent_rank(VerfuegbaresEinkommen08))
+
+
+##############################################################
+### Setup Data ##             Dateien lokal an der BFH    ####
+############################################################## 
+
+# Gini Time Series equivalized disposal income 2000 until 2011 
+# Achtung: Zahlen sind ungewichtet. Keine Äquivalenzskalengewichtung.
+
+setwd("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE")
+
+habe0011<-data.frame(c(2000:2011),seq(0))
+colnames(habe0011)<-c("Year","Gini")
+habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+habe0011$Gini[12]<-by(habe0911$VerfuegbaresEinkommen08,habe0911$Jahr08,gini)[3]
+habe0011$Gini[11]<-by(habe0911$VerfuegbaresEinkommen08,habe0911$Jahr08,gini)[2]
+habe0011$Gini[10]<-by(habe0911$VerfuegbaresEinkommen08,habe0911$Jahr08,gini)[1]
+habe0608<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2006 bis 2008/HABE060708_Standard_130124UOe.txt", header=TRUE)
+habe0011$Gini[09]<-by(habe0608$VerfuegbaresEinkommen08,habe0608$Jahr08,gini)[3]
+habe0011$Gini[08]<-by(habe0608$VerfuegbaresEinkommen08,habe0608$Jahr08,gini)[2]
[…]

### 2014-06-02 paper/paper.tex

@@
 \usepackage{hyperref} % For hyperlinks in the PDF
+\usepackage{graphicx}
 
@@
 \fancyfoot[RO,LE]{\thepage} % Custom footer text
-\bibliographystyle{plain}
+%\bibliographystyle{plain}
+\bibliographystyle{unsrtnat}
 
@@
 \begin{abstract}
+There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013). 
+In Switzerland, however, the development is unclear, as there is evidence for trends in both directions.
+Part of the inconclusive picture is due to different methodological approaches.
+In this paper we discuss the role of taxdata concerning the assessment of inequality in income.
+The focus of the discussion lays herein to show the benefits and shortcomings of tax data compared to current ``state of the art'' measurement concepts of economic inequality. 
+We present common and new strategies to handle tax data specific methodological difficulties and compare results out of aggregated federal tax statistics to results from the Household Budget Survey (HBS).
+We can show to which extend survey data underestimate inequality in income. 
+Following the results out of the taxdata Switzerland experienced in slight rise in inequality in recent years, similar to other western countries, but only because of rise in upper percentiles of the income distribution.
 
-There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
+% Salverda-Stuff und das neue Piketty Buch sollten irgendwo erwähnt werden
+
+\end{abstract}
+
+% Wir machen nichts mehr zu Vermögen (obwohl der Trend auch dahin geht, dass neben dem Einkommen auch das Vermögen eine zentrale Untersuchunsgrösse des inviduellen Wohlstandss darstellt)
+
+%
+
+% Alte version des abstractes. Fokus stärker auf Schweiz und Vermögen
+%There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
 In Switzerland, however, the development is unclear, as there is only little systematic evidence about income inequality that would allow for long-term comparison.
@@
 In the end we integrate the case of Switzerland into the international picture of inequality development showing parallels and deviations.
-
-\end{abstract}
-
-% abstract müsste angepasst werden.
-% Wir machen nichts mehr zu Vermögen (obwohl der Trend auch dahin geht, dass neben dem Einkommen auch das Vermögen eine zentrale Untersuchunsgrösse des inviduellen Wohlstandss darstellt)
-
-% Alternative
-% There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
-In Switzerland, however, the development is unclear, as there is evidence for trends in both directions.
-Part of the inconclusive picture is due to different methodological approaches.
-In this paper we discuss the role of taxdata concerning the assessment of inequality in income.
-The focus of the discussion lays herein to show the benefits and shortcomings of tax data compared to current state of the art measurement concepts of economic inequality.
-We present common and new strategies to handle tax data specific methodological difficulties and compare results out of aggregated federal tax statistics to results from the Household Budget Survey (HBS).
-We can show to which extend survey data underestimate inequality in income.
-Following the results out of the taxdata Switzerland experienced in slight rise in inequality in recent years, similar to other western countries, but only because of rise in upper percentiles of the income distribution.
-
-% Salverda-Stuff und das neue Piketty Buch sollten irgendwo erwähnt werden
-%
 
@@
 
-% Im European Sociologica Review scheint es unüblich zu sein, ohne theoretische Sektion zu arbeiten.
-Ich hoffe nicht, dass dies ein k.o.
-Kriterium ist.
-wir halten dafür die Einführungs und die Methoden Sektion etwas allgemeiner
-
-% Ich würde hier ohne Unterkapitel arbeiten und die Einführung einfach durch Abschnitte gliedern
-
 % 1.Abschnitt: Bedeutung von Ungleichheit
-
 Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular (Wilkinson and Pickett 2009).
@@
 % 2.Abschnitt: Bedeutung der Datengrundlage und Einführung von Steuerdaten
-
 Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
@@
 % 3.Abschnitt: What do we know about Switzerland
-%
-
+% Offical Data Collections on development of inequality
+As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concering inequality in income.
 What is known about Switzerland so far?
-Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
-Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
-The debate went along with referendums trying to regulate the market outcomes.
-Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisingly high majority of 68\%, the 1:12-Initiative, which aimed at the whole spectrum of the income distribution, was rejected.
-The voting about the minimum wage-initiative will be held on 9th February 2014.
-All these referendums questioned the current distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
-Several official and semi-official publications addressed the question of rising inequality in Switzerland.
-\\
-
-% Lässt sich allenfalls streichen/kürzen -> oh April2014-> streichen
-% Ist es die Bezeichnung offical publications sinnvoll? Oder besser govermental publications
-% rf: oder "Federal publications"?
-
-% 4.Abschnitt: What do we know about Switzerland
-% Offical Data Collections on development of inequality
-
-The most recent official figures published by the Swiss Federal Statistical Office (2013) show stability concerning the income distribution covering the time period of 1998 until 2011.
-The authors found a slight increase in the primary income which endured until the recent crisis.
-This temporary increase was mainly compensated by governmental redistribution.
-A longer period is covered in the LIS-Data-set (1982-2004).\footnote{Data-provider for the LIS Data is the Swiss Federal Statistical Office.
-The data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
-The OECD-Database includes measures from Income and Consumption survey as well.
+Looking for offical data, three main sources has to be mentioned, which can be considered as de facto offical data sources: EU-SILC, HBS and LIS-data.
+Figure XY shows the results steming from this three sources while looking at Gini of equivalised disposable income.
+Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level.
+The main focus of EU-SILC is to collect data on a common ``framework'' to ensure comparability among EU-countries and countries living around or within the EU.
+As a Non-EU member Switzerland implemented the instrument not from the beginning (2004) but as from 2007.
+Therefore this times-series doesn't cover timepoints before 2007.
+As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concering the distribution of income is the Household Budget Survey (HBS).
+The main focus of this survey lays in providing detailed data on houshold budegts.
+This allows researcher to look at different income concepts like income before and after public transfers.
+Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistend time series from 2000 to 2011.
+As it can be seen from graph XY the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period.
+A longer period is covered in the LIS-Data-set (1982-2004).
+Data-provider for the LIS Data is the Swiss Federal Statistical Office too.
+In contrast the the aformentioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+All in all the LIS dataset contains the longest time series on inequality for Switzerland.
+Analyzing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
+This result is supported by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).\footnote{A further offical database for income distribution is the OECD-Database.
+It includes measures from Income and Consumption survey as well.
 Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
-This change in survey is considered as a strict break.
+But this change in survey is considered as a strict break.
 Comparison before and since 2008 is not recommended (OECD 2012:315).
-All in all the LIS dataset contains the longest time series on inequality for Switzerland.} Analyzing these data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
-This result is supported by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).
-\\
-
-% Bei SILC einsteigen. Allenfalls eine Grafik bringen mit den unterschiedlichen Reihen.
-% Daraus wird auch ersichtlich, dass keine der verfügbaren Reihen wirklich länge Zeiträume abdecken
-% we should mention that Eurostat uses EU-SILC alike and show, that there is a time series for SILC-Data aswell.
-Additionally we should mention that the Federal Statistical Office uses the HABE statistic
-
-% Allenfalls erwähnen, dass die Schweiz nicht Teil des GINI database ist.
-http://www.gini-research.org/articles/data_2
+A further important database on inequality is the GINI database which has been derived from the GINI Country Reports (Nolan et al.,2014).
+But this dataset doesn't cover Switzerland.}  \\
 
@@
 
-% 5.Abschnitt: What do we know about Switzerland
+% 4.Abschnitt: What do we know about Switzerland
 % Studies with tax data
-
 Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
@@
 
-%
-
-% 6. Abschnitt: Zwischenbilanz und Herausforderungen
-
-The divergence can be explained with several factors.
+% 5. Abschnitt: Zwischenbilanz und Herausforderungen
+Divergence can be explained with several factors.
 First of all, different data sources were used.
@@
 It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
-\\
-
-% Allenfalls hier noch nicht so ausführlich mit Vor- und Nachteilen von Steuerdaten kommen.
-Vielleicht besser darauf hinweisen, welche Dimensionen betroffen sind und das dies in der Methods Section besprochen wird.
-%rf: finde den abschnitt super
-
-While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
-The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
-But, up to the day information for Switzerland is not available.
-
-% ab hier wird es irgnediwe unklar
-Shorrock et al.
-(2013) unify several datasources (including tax data) to assess the pattern of global wealth.
-Their databook includes figures for Switzerland.
-In contrast to inequality of income concerning inequality of wealth Switzerland takes a leading position \\
-
-% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
-% Dann lassen sich einige Passagen streichen.
-Man könnte auch auf verschiedene Publikationen verweisen, welche die Bedeutung von Vermögen untersuchen -> Buch von Piketty, Müller und Schoch und dann aber ziemlich schnell darauf verweisen, dass wir uns hier auf Einkommen konzentrieren.
-% Auch erste Dell Studie erwähnen -> es hat Zahlen zur Schweiz. Auch Probleme erwähnen (Fehlende Informationen zu Altersvorsorgevermögen -> hat aber niemand eine Lösung bisher.)
-% Studies on inequality have to adress several thorny challenges.
-One has to define the centrale measure of inequality and find adquate data.
-Recent studies accentuate the importance of a broader concept of economic wellbeeing including measures for income, consumption and wealth alike.
-But each area is confronted with own measurement problems.
-Measure of consumptions are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow.
-Measures of wealth are included in tax data, but are not connectable to the respctive income and therefore can only be analyzed separate.
-In danger of falling a pittraps it seems appriate to move in small steps.
-Therefore we focus our analysis on income, which is undoubtly a crucial indicator of economic wellbeeing.
-% Also the Federal Statistical Office is increasing their effort to get sound data concering the distribution of wealth vgl. Müller und Schoch (2014).
 
@@
 We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
-Third, we integrate the development of Swiss inequality into the international picture.
+Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through datasource.
 
-% delete passages about wealth and include passage about the comparison with survey data
-
-% Should we discus the non-scientic literature? -> allgemeine Frage und im konkreten betrifft dies die Publikationen von Economie Swiss und des Gewerkschaftsbunds.
-% rf: gute frage, spontan würde ich sagen eher nein, müssten wir mal diskutieren
+% Könnte noch besser auf die geplanten Analysen abgestimmt sein (vgl. auf Github-Issues)
 
 % Was wollen wir wissen?
+
 % Wir müssen uns äussern zu: 
@@
 % Definition des Einkommens
-% Berechnungsschritten
+% Messung der Verteilung
 
@@
 % 3) Vergleich mit HABE
+
+% Textbaustein politische Dimension von Ungleichheit
+% What is known about Switzerland so far?
+Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
+Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+The debate went along with referendums trying to regulate the market outcomes.
+Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisingly high majority of 68\%, the 1:12-Initiative, which aimed at the whole spectrum of the income distribution, was rejected.
+The voting about the minimum wage-initiative will be held on 9th February 2014.
+All these referendums questioned the current distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
+Several official and semi-official publications addressed the question of rising inequality in Switzerland.
+\\
+
+% Alte Textbausteine zu Vermögen
+%While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
+The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
+But, up to the day information for Switzerland is not available.
+% ab hier wird es irgnediwe unklar
+%Shorrock et al.
+(2013) unify several datasources (including tax data) to assess the pattern of global wealth.
+Their databook includes figures for Switzerland.
+In contrast to inequality of income concerning inequality of wealth Switzerland takes a leading position \\
+% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
+% Dann lassen sich einige Passagen streichen.
+Man könnte auch auf verschiedene Publikationen verweisen, welche die Bedeutung von Vermögen untersuchen -> Buch von Piketty, Müller und Schoch und dann aber ziemlich schnell darauf verweisen, dass wir uns hier auf Einkommen konzentrieren.
+% Auch erste Dell Studie erwähnen -> es hat Zahlen zur Schweiz. Auch Probleme erwähnen (Fehlende Informationen zu Altersvorsorgevermögen -> hat aber niemand eine Lösung bisher.)
 
@@
 
-\section{Measurement concepts, Data and Methods}
+\section{Data, Measurement concepts and Methods}
 
-To discuss the suitability of a specific data source it is crucial to have an idea about an ideal measure.
-Based on a review on the literature about the measurement concepts in an ideal world, we discuss briefly the advantages and shortcomings of tax data compared to other data sources - namely survey data.
-Afterwards we describe the tax data published by the Swiss Federal Tax Administration highlighting the important aspects, which have to be considered working with FTA-Data.
-In the last section, we describe the methods and corrections we used to construct the time-series of inequality-measure for income an wealth for Switzerland.
+Studies on inequality have to adress several thorny challenges.
+It starts with answering three crucial questions: First of all one has to define, which concepts should to be looked at.
+This refers to answering the question about inequality of what.
+Secondly one has to be clear about the unit of analysis.
+This refers to answering the question about inequality among whom.
+Thirdly, one has to choose an appropriate measure of inequality.
+All this questions are ideally answered considering theory and a given research question.
+Often it has to be answered in context of a given dataset.
+Therefore we start this section with a description of the FTA-Tax Data.
+Based on a review on the literature about the measurement concepts in an ideal world, we discuss  the advantages and shortcomings of tax data compared to other data sources - namely survey data.
+Along with the despriction we explain the methods and thechnics we use to construct time-series of inequality-measure for income for Switzerland, which we will present in the result sections.
+By doing this, we demonstrate the suitability of the FTA Tax Data to assess the development of inequality in Switzerland
+
+\subsection{Tax statistics in Swizterland}
 
@@
 \emph{Concepts on measuring economic resources}  \\
-Most studies on inequality focus on income inequality, while economic resources are ideally conceptualized with measures for income, wealth and consumption together.
-In generall it has to be stateted, that ideal measures only can be defined in context of a given reaserach question.
-This holds, if one thinks about the ideal measure of economic well-being, but also about the unit of analysis.
-Nevertheless, some more or less unified concepts exists, which can be taken as a starting point.
-OECD (2013) \\
 
-% We neither can examine consumption nor the jointly distribution of income and wealth on the individual level.
-We could analyse the jointly pattern on the cantonal level and analze the data following the question if income inequality and wealth inqueality are correlated.
+Most studies on inequality focus on income inequality solely.
+However, recent activities emphasize the need of a broader conceptualization.
+A recent publication from the oecd (2013) condense these ideas into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework on micro-level statistics.\footnote{Harmonisation with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framewrok presented in this publication.
+Considered main stadards were the System of National Accounts (SNA 2008), the canberra Group Handbook on Household Income Statistics (United Nations 2011), the final report of the 17th International Conference of Labour Statisticians (ILO 2004) and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing (UNECE & Eurostat 2006)} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
+To gain policy relevant insight, it is recomended to look at the distribution of all three distributions simultaneously.
+Some households with low income, for example, may report adequate levels of consumption expenditure or welth holdings, or vice-versa.
+But it is also stated (OED 2013:18):"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by colecting and provding a database following ths broader concept of economic well-being.
+http://www.lisdatacenter.org/our-data/lws-database/} \\
+
+This last statment holds for Swizterland too, although the HBS study is strongly influced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
+Albeit the awarness of an assessment of income, consumption and wealth simuntanously is rising, we  focus our analysis on income, which is undoubtly a crucial indicator of economic wellbeeing.
+However, the Federal Tax Office publishes statistics on income and wealth .
+But it is not possible to analyse the jointly distribution on the individual or household level.
+Also measures of consumption are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow.
 
@@
 
-\begin{itemize}
-\item Market income refers to revenue from employment.
-\item primary income or pre-transfer income refers to  market income plus revenue from property (capital income)
-\item Brutto income refers to  primary income plus social transfer
-\item Disposable income or posttransfer income refers to  brutto income minus transfer paid 
-\item adjusted disposable income refers to  disposable income corrected for the numbers of household members 
-\end{itemize}
+The assessment of income inequality is influenced by the definition of the income itself.
+Marketincome or disposable income for example differ by substantial meaning and by the expected degree of inequality.
+Therefore the awarness of the analyzed concept is crucial.
+Terminology can sligthly differ, while common concepts can be identified (for detailed discussion see: OECD (2013:44), Canberra group (2011:24)).
+Figure xy shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution.
+Within this framework common income definitions are situated.
+These definitions are contrasted with income concepts, which can be derived out of the FTA Tax data.
 
-The appropriateness of the income concept is bound to the research interest.
-Studies focusing on labor market outcome look at market income.
-Studies focusing on human wellbeing follow the approach as propagated by the oecd (2013).
-This concept is largely implemented by the Federal Office of statistics.
-Following this concept one should look at income after taxes and transfers inclusive the adjustment for household members, because this is the measure which frames the consumption possibilities.
-Lastly, studies focusing on the effect of institutions (social welfare, taxation) compare the distribution of pre- and post-income measures.
+\includegraphics[width=\maxwidth]{figure/income_definition}
+
+% Quelle zu Abzügen:
+% http://www.bilanz.ch/invest/steuern/netto-oder-reineinkommen
+% Die rechte Seite mit den Steuergrössen könnte ausführlicher beschrieben werden
+% Beispielsweise könnte man zeigen was Nettoeinkommen, Reineinkommen und steuerbares Einkommen sind
+% Allerdings steht uns nur Reineinkommen und steuerbares Einkommen zur Verfügung
+
+The central income according to tax data is the taxable income.
+It includes all reported incomes (income from employment, income from property and received transfers\footnote{Mean-tested benefits are not taxed and therefore not included in Tax data.
+Income for low income groups are therefore underestimated.}) minus several deductions.
+It is therefore neither a pre-transfer income  nor a post-transfer income measure.
+It's rather something inbetween.
+As the FTA tax statistitcs include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions.
+This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction,Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income".
+As some deducations can be interpreted as compulsary expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption.
+Similiar when calculating the disposable income out of the taxable income through accouting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
+Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches are missing.
 \\
 
-% an income which is adjusted for the household size is not a real income anymore. It is a measure of indivdual welfare or of the standard of living (BFS:2012)
+% Die Deductions sind noch nicht gänzlich sauber abgebildet.
+Auf der Bilanz-Seite steht, zum steuerbaren Einkommen kommt man, indem vom Reineinkommen die Sozialabzüge(Kinder- und Fremdbetreung) sowie der Abzug für unterstützungsbedürftige Personen abzogen werden.
+% Auf den ESTV-Tabellen sind jedoch noch mehr Abzüge abgebildet. Was soll/muss/können wir nun abziehen? Bin verwirrt...
 
-\emph{Defining wealth}
-While the apropiateness of the conceptualization of income is widely discuded, there is no agreed definition of personal wealth yet and the appropriate methods of valuation are not always clear. 
-Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
+If the goal is to catch the income, which shapes the possibility to consume, equivalence scale are implemented.
+Equivalence scale followe the idea, that persons living in the same household have the possibility to share costs.
+To compare the individual economic well-beeing equivalence scales adjust for different household sizes (see OECD 2013,173, Buhmann et al., 1998).
+Because it is neither possible to calculate the total household income nor the respective householdsize out of the FTA tax data, we put this in this paper outside the brackets.
+Although knowing, that with doing so we neglect a potential aspect, which is capable of shaping the development of inequality.
+
+% letzter Abschnitt muss noch auf die Brülhart-Daten hinweisen. Damit können wir eine äquivalenzeinkommen betrachten.
+% Auch müsste in diesem Abschnitt auf die Probleme der Haushaltsunterscheidung eingegangen werden (oder ist im Abschnitt Population coverage? Dann müsste obiger Punkt auch da erwähnt werden.)
 
 % Streichen
+%=======
+% 
+%>>>>>>> d21c323... completed subsection "concepts on measuring economic resources" and "defining income"
 
@@
 
-The model outputs a test statisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical "canton gini-level".
+The model outputs a test statisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical ``canton gini-level''.
 As the model has a decent fit we are not in great danger of omitted variable bias.
-Using a joint F-Test we can now test if all canton#zeros interactions are zero.
+Using a joint F-Test we can now test if all canton interactions are zero.
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# \\\#\# Attaching package: 'survey'\\\#\# \\\#\# Das folgende Objekt ist maskiert from 'package:graphics':\\\#\# \\\#\#\ \ \ \  dotchart}}\begin{verbatim}
-## Wald test for kanton:zeros
-##  in lm(formula = G_steink ~ kanton + kanton:zeros + J + zeros:J, 
-##     data = df)
-## F =  3.282  on  27  and  262  df: p= 4e-07
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-In our case we can clearly reject the hypothesis that all interactions are zero.
+In our case we can clearly reject the hypothesis that all interactions are zero ($p=0$).
 This leads to the conclusion  that gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted gini coefficients.
@@
 To simplify: cutting off zeros more and more seems to lead to an underestimation of our inequality measure, probably because the skewness in the left part of the distribution increased pointing to an increased pauperization that is masked by omitting the left tail of the income distribution.
+
+% Alter Textbaustein zu Vermögen
+%\emph{Defining wealth}
+%While the apropiateness of the conceptualization of income is widely discuded, there is no agreed definition of personal wealth yet and the appropriate methods of valuation are not always clear.
+Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# \\\#\# Attaching package: 'reshape'\\\#\# \\\#\# Die folgenden Objekte sind maskiert from 'package:plyr':\\\#\# \\\#\#\ \ \ \  rename, round\_any}}\end{kframe}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
 \includegraphics[width=\maxwidth]{figure/different_ginis1}
@@
 \includegraphics[width=\maxwidth]{figure/with_without_zeros1}
-\begin{kframe}
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# 'opts' is deprecated.
-Use 'theme' instead.
-(Deprecated; last used in version 0.9.1)\\\#\# theme\_text is deprecated.
-Use 'element\_text' instead.
-(Deprecated; last used in version 0.9.1)}}\end{kframe}
 \includegraphics[width=\maxwidth]{figure/with_without_zeros2}

### 2014-06-02 paper/subdocuments/abstract.Rnw

@@
 In this paper we discuss the role of taxdata concerning the assessment of inequality in income.
-The focus of the discussion lays herein to show the benefits and shortcomings of tax data compared to current “state of the art” measurement concepts of economic inequality.
+The focus of the discussion lays herein to show the benefits and shortcomings of tax data compared to current ``state of the art'' measurement concepts of economic inequality. 
 We present common and new strategies to handle tax data specific methodological difficulties and compare results out of aggregated federal tax statistics to results from the Household Budget Survey (HBS).

### 2014-06-02 code/habe.r
minor change

@@ -29,7 +29,7 @@ habe <- habe %.%
 ############################################################## 
 
 # Gini Time Series equivalized disposal income 2000 until 2011 
-# Achtung: Zahlen sind ungewichtet. Keine Äquivalenzskalengewichtung.
+# Achtung: Zahlen sind ungewichtet. 
 
 setwd("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE")
 

### 2014-06-03 code/habe.r
implemented equivalence scale weighting and proper calculation for disposable income

@@ -29,36 +29,97 @@ habe <- habe %.%
 ############################################################## 
 
 # Gini Time Series equivalized disposal income 2000 until 2011 
-# Achtung: Zahlen sind ungewichtet. 
+# Achtung: Keine Querschnittgewichte verwendet
+# Äquivalenzskalen: In der BFS Publikation 2013 wird bei der Äquivalenzskalenberechnung zwischen Kindern
+# und Erwachsenen unterschieden (modifizierte OECD-Skala. Als Kinder gelten Personen unter 14
+# In den gelieferten Daten ist die Kinder-Variable allerdings mit unter 15 gelabelt (so verwendet)
+# von 2000 bis 2005 gibt es in den Daten keine Unterscheidung von Erwachsenen und Kindern
 
 setwd("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE")
 
+
 habe0011<-data.frame(c(2000:2011),seq(0))
 colnames(habe0011)<-c("Year","Gini")
+# 2009 bis 2011
 habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
-habe0011$Gini[12]<-by(habe0911$VerfuegbaresEinkommen08,habe0911$Jahr08,gini)[3]
-habe0011$Gini[11]<-by(habe0911$VerfuegbaresEinkommen08,habe0911$Jahr08,gini)[2]
-habe0011$Gini[10]<-by(habe0911$VerfuegbaresEinkommen08,habe0911$Jahr08,gini)[1]
-habe0608<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2006 bis 2008/HABE060708_Standard_130124UOe.txt", header=TRUE)
-habe0011$Gini[09]<-by(habe0608$VerfuegbaresEinkommen08,habe0608$Jahr08,gini)[3]
[…]

### 2014-06-03 code/habe.r
minor plot-code correction

@@ -115,7 +115,7 @@ habe0011$Gini[01]<-gini(habe00$y.sum.ag)
 
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
 ggplot(data=habe0011,aes(x=Year,y=Gini))+
-  scale_y_continuous(limits=c(0.0,0.8),breaks=number_ticks(8))+ 
+  scale_y_continuous(limits=c(0.2,0.5),breaks=number_ticks(4))+ 
   scale_x_continuous(limits=c(2000,2011),breaks=number_ticks(11)) +
          geom_line()+
          geom_point()+

### 2014-06-04 paper/subdocuments/data_and_methods.Rnw
first draft of subsection statistical unit and population coverage

@@
 \subsection{Tax statistics in Swizterland}
+
+% Hier sollten kurz die Steuerstatistiken beschrieben werden.
+Ziel ist es eine Basis f?r die nachfolgende Gegen?berstellung von Soll und Ist zu legen.
+Gleichzeitig sollen die Daten vorgestellt werden, die wir f?r die sp?teren Analysen verwenden.
+
+%Beschreiben was die Daten beinhalten und wo es Probleme gibt
+\begin{itemize}
+\item  Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
+	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
+	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
+	\item	Income after deductions-> taxable income (employment income, business income and capital income).
+Is income definition stable over time?
+Yes it should.
+Realized capital gains are excluded from the definition.
+It includes income from  employment, self-employment, capital income and taxable transfer payments.
+Plus Eigenmietwert,
+	\item	Reported on national and cantonal level
+\end{itemize}
 
@@
 As the FTA tax statistitcs include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions.
-This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction,Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income".
+This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction, Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income".
 As some deducations can be interpreted as compulsary expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption.
 Similiar when calculating the disposable income out of the taxable income through accouting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
-Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches are missing.
+Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches, which represent the bulk ot taxes in Switzerland, are missing.
 \\
@@
 
-If the goal is to catch the income, which shapes the possibility to consume, equivalence scale are implemented.
-Equivalence scale followe the idea, that persons living in the same household have the possibility to share costs.
-To compare the individual economic well-beeing equivalence scales adjust for different household sizes (see OECD 2013,173, Buhmann et al., 1998).
-Because it is neither possible to calculate the total household income nor the respective householdsize out of the FTA tax data, we put this in this paper outside the brackets.
-Although knowing, that with doing so we neglect a potential aspect, which is capable of shaping the development of inequality.
+\emph{statistical units} \\
 
-% letzter Abschnitt muss noch auf die Brülhart-Daten hinweisen. Damit können wir eine äquivalenzeinkommen betrachten.
-% Auch müsste in diesem Abschnitt auf die Probleme der Haushaltsunterscheidung eingegangen werden (oder ist im Abschnitt Population coverage? Dann müsste obiger Punkt auch da erwähnt werden.)
+% 1.Abschnitt dar?ber, dass besser Haushalte betrachtet werden anstelle von Individuuen
 
-% Streichen
-%=======
-% 
-%>>>>>>> d21c323... completed subsection "concepts on measuring economic resources" and "defining income"
+The agreed standard on the statistical unit, which should be the base of inequality analysis, are households not individuals (OECD 2013:60).
+Indeed it are the individuals, who receive income, own wealth and experience economic wellbeing, but their possibility to do so, is strongly tied to the concept of household.
+This comprises all persons under the same housing arrangement.
+The basic underlying assumption for colection data on households level insted of individual level is, that people in the same household share ressources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every householdmember (also non-earing members, like children).
+Additionaly, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing.
+The compare the individual economic well-beeing among individuals living in different households usualy equivalence scales are used (see OECD 2013,173, Buhmann et al., 1998).
+\\
 
-\emph{Population Coverage} \\
-All Residents? Working Residents? Single Person vs household/family. Households without income \\
-see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\
+% 2. Abschnitt zur Gegen?berstellung von Haushalten und Steuereinheiten
 
-The income of the household generally plays the major role in the living standards attained by its members, so the way earners and non-earners grouped together in households, has to be incorporated into the analysis.
-Especially when the the focus of interesst lays in unrevealing the role of demographic factors.
+In tax data, however, the units are represented according to administrative rules.
+Tax units therefore neither represent individuals in every case nor true households.
+Tax units rather represent individuals and couples, but couples, who are married or officially registrered.
+This doesn't imply, that those couples live together, as it is needed to satisfy the definition of a household.
+On the other hand, is it quit likely that more than one tax units live in the same household (unmarried/unregistered couples, see M?ller and Schoch (2014:99)).
+It is therefore not directly possible to elicit households and household inomce out of tax data.
+This might influence the assessment inequality, taking into account the change from traditional household and family structures over the last century.
+This is an inequality related issue, where relevant studies are missing.
 
-\emph{Measuring inequality and concentration} \\
+% Allenfalls die Br?lhart-Daten erw?hnen und die m?glichkeit der Haushaltsbildung ?ber Abz?ge. Falls wir auch eine Grafik zur Rolle der Haushaltszusammensetzung machen (ich denke schon)
+% gibt es studien? Zur Bedeutung von ?quivalenzskalen und zur Bedeutung der Differenzierung von Haushalten und Individuuen?
+
+\emph{Population Coverage}
+
+Often inequality is assessed on national level, which implies, that studies try to cover the whole population of the country of interest.
+This is a special thorny task for surveys working with samples, because nonresponse is a major source of bias (Bethlehem et al., 2011).
+Korinek, Mistiaen and Ravallion (2006) show, that the position in the income distribution influence the probability to participate in a survey.
+Low income and high income households are more likely to refuse survey response, which leads to an overrepresentation of middle income households.
+This process can be referred to as the "middleclass bias" Diekmann (2009).
+Missing data in household surveys is therefore not missing at random, which has an impact on the measures of inequality.
+The magnitude of this bias in Switzerland, however, is unknown.
+Strategies to handle this kind of bias are discussed in the literature (S?rndal et al.
+1992), but require a register for every unit, that is proportional to income.
+Currently there exists no such register for Switzerland (ecoplan 2014:43).
+Currently used microdata sets, which are used for official publications concerning inequality in Switerzland (SILC and HABE) are furthermore confronted with a constructed coverage problems, because this surveys rely on phoneregister, which excludes the households not having a registered connection.
+\\
+%(Problems of collective households, ecoplan 2014:45). Erw?hnen?
+
+The issue of incomplete coverage is far less dramatical with tax data.
+Essentially every permanent resident in Switzerland, who has completed the age of 18 years (respectivley 20 years prior to 1996) is taxed on a yearly base (or every two years before the change of the tax system).
+In principle this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distrubtion.
+This includes a separation of normal cases, which embrace the majority of taxpayers, and the special cases, which cover several unusual situations.
+Most important high net wealth individuals taxed according to their expenditures.However, special attention has to be paid to tax units with none or very low incomes.
+Even though they have to hand in a tax return, their income do not show up in the statistics if their income after deductions falls below 15'000 CHF and they are therefore not taxed with direct federal taxes.
+This is possible for normal and special cases alike.
+From 1995/1996 until 2010 the number of non-taxed units are reported, but not for the years before.
+
+% What about the DELL Study, where high share of non-fillers are reported? What are the reasons?
+
+\emph{Measuring inequality or concentration} \\
 To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
@@
 
+% 3. Eigenheiten Steuereinheiten
+
 %Warnings about the use of tax data
@@
 For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).
-
-%Copied from:
-%Alvaredo, Facundo, Anthony B. Atkinson, Thomas Piketty and Emmanuel Saez, The World Top Incomes Database, http://topincomes.g-mond.parisschoolofeconomics.eu/ , 18/12/2013
 
@@
 Additonaly the Credit Suisse Reports states (2011:8) that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts.
-
-\subsection{Tax data published by the Swiss Federal Tax Administration}
-
-%Beschreiben was die Daten beinhalten und wo es Probleme gibt
-\begin{itemize}
-\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
-	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
-	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
-	\item	Income after deductions-> taxable income (employment income, business income and capital income).
-Is income definition stable over time?
-Yes it should.
-Realized capital gains are excluded from the definition.
-It includes income from  employment, self-employment, capital income and taxable transfer payments.
-Plus Eigenmietwert,
-	\item	Reported on national and cantonal level
-\end{itemize}
 

### 2014-06-04 code/habe.r

@@ -66,7 +66,7 @@ habe05$y[habe05$NOMENKLATUR_STUFE1_ID=="5"]<-0
 habe05<-summaryBy(y~HAUSHALT_ID,FUN=sum,data=habe05)
 habe05<-merge(habe05,habe05hh)
 habe05$y.sum.ag<-habe05$y.sum/(1+0.5*(habe05$ANZ_PERSONEN-1))
-habe0011$Gini[06]<-gini(habe05$y.sum.ag)
+habe0011$Gini[06]<-gini(habe05$y.sum.ag,habe05$GEWICHT)
 
 habe04<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2003 bis 2005/eintrag_hh_aggregat1_2004_061127pp.txt", header=TRUE)
 habe04hh<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2003 bis 2005/haushalt_2004_061127pp.txt", header=TRUE,sep="\t")
@@ -75,7 +75,7 @@ habe04$y[habe04$NOMENKLATUR_STUFE1_ID=="5"]<-0
 habe04<-summaryBy(y~HAUSHALT_ID,FUN=sum,data=habe04)
 habe04<-merge(habe04,habe04hh)
 habe04$y.sum.ag<-habe04$y.sum/(1+0.5*(habe04$ANZ_PERSONEN-1))
-habe0011$Gini[05]<-gini(habe04$y.sum.ag)
+habe0011$Gini[05]<-gini(habe04$y.sum.ag,habe04$GEWICHT)
 
 habe03<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2003 bis 2005/eintrag_hh_aggregat1_2003_050614pp.txt", header=TRUE)
 habe03hh<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2003 bis 2005/haushalt_2003_050614pp.txt", header=TRUE,sep="\t")
@@ -84,7 +84,7 @@ habe03$y[habe03$NOMENKLATUR_STUFE1_ID=="5"]<-0
 habe03<-summaryBy(y~HAUSHALT_ID,FUN=sum,data=habe03)
 habe03<-merge(habe03,habe03hh)
[…]

### 2014-06-05 code/habe.r
minor changes

@@ -58,19 +58,49 @@ habe0011$Gini[09]<-by(habe0608$VerfuegbaresEinkommen08gew,habe0608$Jahr08,gini)[
 habe0011$Gini[08]<-by(habe0608$VerfuegbaresEinkommen08gew,habe0608$Jahr08,gini)[2]
 habe0011$Gini[07]<-by(habe0608$VerfuegbaresEinkommen08gew,habe0608$Jahr08,gini)[1]
 
-# 2005 bis 2000
+####
+# Achtung! Die konstruktion des verfügbaren Einkommens stimmt nicht!!!!!!!!!!!!
+####
+
+
+# 2001 bis 2005
 habe05<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2003 bis 2005/eintrag_hh_aggregat1_2005_070601pp.txt", header=TRUE)
 habe05hh<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2003 bis 2005/haushalt_2005_070601pp.txt", header=TRUE, sep="\t" )
-habe05$y<-ifelse(habe05$B_AUSGABE<5,habe05$SUMME_BETRAG_CHF,habe05$SUMME_BETRAG_CHF*(-1))
+habe05$y<-ifelse(habe05$B_AUSGABE==0,habe05$SUMME_BETRAG_CHF,habe05$SUMME_BETRAG_CHF*(-1))
 habe05$y[habe05$NOMENKLATUR_STUFE1_ID=="5"]<-0
 habe05<-summaryBy(y~HAUSHALT_ID,FUN=sum,data=habe05)
 habe05<-merge(habe05,habe05hh)
 habe05$y.sum.ag<-habe05$y.sum/(1+0.5*(habe05$ANZ_PERSONEN-1))
 habe0011$Gini[06]<-gini(habe05$y.sum.ag,habe05$GEWICHT)
 
+#
+habe05long<-reshape(habe05,direction="wide",idvar="HAUSHALT_ID",timevar="NOMENKLATUR_STUFE1_ID",v.names="SUMME_BETRAG_CHF")
+habe05long[is.na(habe05long)]<-0
+habe05long<-merge(habe05long,habe05hh)
+head(habe05long)
+#
+
+# Transferausgaben auf Agg_1 sind zu umfassend
[…]

### 2014-06-09 paper/subdocuments/abstract.Rnw
citations and spell check

@@
 \begin{abstract}
-There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013). 
+There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (\citealt{cooperation_divided_2011 UNSICHER}, \citealt{gornick_income_2013}). 
 In Switzerland, however, the development is unclear, as there is evidence for trends in both directions.
 Part of the inconclusive picture is due to different methodological approaches.
-In this paper we discuss the role of taxdata concerning the assessment of inequality in income.
+In this paper we discuss the role of tax-data concerning the assessment of inequality in income.
 The focus of the discussion lays herein to show the benefits and shortcomings of tax data compared to current ``state of the art'' measurement concepts of economic inequality. 
@@
 We can show to which extend survey data underestimate inequality in income. 
-Following the results out of the taxdata Switzerland experienced in slight rise in inequality in recent years, similar to other western countries, but only because of rise in upper percentiles of the income distribution.
+Following the results out of the tax-data Switzerland experienced in slight rise in inequality in recent years, similar to other western countries, but only because of rise in upper percentiles of the income distribution.
 
@@
 % Alte version des abstractes. Fokus stärker auf Schweiz und Vermögen
-%There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; Gornick and Jäntti, 2013).
+%There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (OECD, 2011; \citealt{gornick_income_2013}).
 In Switzerland, however, the development is unclear, as there is only little systematic evidence about income inequality that would allow for long-term comparison.

### 2014-06-09 paper/subdocuments/data_and_methods.Rnw
citations and spell check

@@
 
-Studies on inequality have to adress several thorny challenges.
+Studies on inequality have to address several thorny challenges.
 It starts with answering three crucial questions: First of all one has to define, which concepts should to be looked at.
@@
 Based on a review on the literature about the measurement concepts in an ideal world, we discuss  the advantages and shortcomings of tax data compared to other data sources - namely survey data.
-Along with the despriction we explain the methods and thechnics we use to construct time-series of inequality-measure for income for Switzerland, which we will present in the result sections.
+Along with the description we explain the methods and techniques we use to construct time-series of inequality-measure for income for Switzerland, which we will present in the result sections.
 By doing this, we demonstrate the suitability of the FTA Tax Data to assess the development of inequality in Switzerland
 
-\subsection{Tax statistics in Swizterland}
+\subsection{Tax statistics in Switzerland}
 
@@
 However, recent activities emphasize the need of a broader conceptualization.
-A recent publication from the oecd (2013) condense these ideas into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework on micro-level statistics.\footnote{Harmonisation with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framewrok presented in this publication.
-Considered main stadards were the System of National Accounts (SNA 2008), the canberra Group Handbook on Household Income Statistics (United Nations 2011), the final report of the 17th International Conference of Labour Statisticians (ILO 2004) and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing (UNECE & Eurostat 2006)} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
-To gain policy relevant insight, it is recomended to look at the distribution of all three distributions simultaneously.
-Some households with low income, for example, may report adequate levels of consumption expenditure or welth holdings, or vice-versa.
-But it is also stated (OED 2013:18):"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by colecting and provding a database following ths broader concept of economic well-being.
+A recent publication from the \citet{oecd_oecd_2013 UNSICHER} condense these ideas into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framework presented in this publication.
+Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_????
+UNSICHER}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
+To gain policy relevant insight, it is recommended to look at the distribution of all three distributions simultaneously.
+Some households with low income, for example, may report adequate levels of consumption expenditure or wealth holdings, or vice-versa.
+But it is also stated \citep[18]{oecd_oecd_2013 UNSICHER}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being.
 http://www.lisdatacenter.org/our-data/lws-database/} \\
 
-This last statment holds for Swizterland too, although the HBS study is strongly influced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
-Albeit the awarness of an assessment of income, consumption and wealth simuntanously is rising, we  focus our analysis on income, which is undoubtly a crucial indicator of economic wellbeeing.
+This last statement holds for Switzerland too, although the HBS study is strongly influenced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
+Albeit the awareness of an assessment of income, consumption and wealth simultaneously is rising, we  focus our analysis on income, which is undoubtedly a crucial indicator of economic wellbeing.
 However, the Federal Tax Office publishes statistics on income and wealth .
-But it is not possible to analyse the jointly distribution on the individual or household level.
+But it is not possible to analyze the jointly distribution on the individual or household level.
 Also measures of consumption are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow.
@@
 The assessment of income inequality is influenced by the definition of the income itself.
-Marketincome or disposable income for example differ by substantial meaning and by the expected degree of inequality.
-Therefore the awarness of the analyzed concept is crucial.
-Terminology can sligthly differ, while common concepts can be identified (for detailed discussion see: OECD (2013:44), Canberra group (2011:24)).
-Figure xy shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution.
+Market income or disposable income for example differ by substantial meaning and by the expected degree of inequality.
+Therefore the awareness of the analyzed concept is crucial.
+Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013 UNSICHER}, \citet[24]{united_nations_canberra_2011}).
+Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution.
 Within this framework common income definitions are situated.
@@
 
+\begin{figure}
+\centering
 \includegraphics[width=\maxwidth]{figure/income_definition}
+\caption{Income definitions}
+\label{fig:incdef}
+\end{figure}
 
@@
 It is therefore neither a pre-transfer income  nor a post-transfer income measure.
-It's rather something inbetween.
-As the FTA tax statistitcs include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions.
+It's rather something in between.
+As the FTA tax statistics include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions.
 This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction, Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income".
-As some deducations can be interpreted as compulsary expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption.
-Similiar when calculating the disposable income out of the taxable income through accouting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
-Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches, which represent the bulk ot taxes in Switzerland, are missing.
+As some deductions can be interpreted as compulsory expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption.
+Similar when calculating the disposable income out of the taxable income through accounting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
+Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches, which represent the bulk of taxes in Switzerland, are missing.
 \\
@@
 
-The agreed standard on the statistical unit, which should be the base of inequality analysis, are households not individuals (OECD 2013:60).
+The agreed standard on the statistical unit, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013 UNSICHER}.
 Indeed it are the individuals, who receive income, own wealth and experience economic wellbeing, but their possibility to do so, is strongly tied to the concept of household.
 This comprises all persons under the same housing arrangement.
-The basic underlying assumption for colection data on households level insted of individual level is, that people in the same household share ressources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every householdmember (also non-earing members, like children).
-Additionaly, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing.
-The compare the individual economic well-beeing among individuals living in different households usualy equivalence scales are used (see OECD 2013,173, Buhmann et al., 1998).
+The basic underlying assumption for collection data on households level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children).
+Additionally, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing.
+The compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013 UNSICHER}, \citealt{Buhmann et al., 1998 FEHLT}).
 \\
@@
 Tax units therefore neither represent individuals in every case nor true households.
-Tax units rather represent individuals and couples, but couples, who are married or officially registrered.
+Tax units rather represent individuals and couples, but couples, who are married or officially registered.
 This doesn't imply, that those couples live together, as it is needed to satisfy the definition of a household.
-On the other hand, is it quit likely that more than one tax units live in the same household (unmarried/unregistered couples, see M?ller and Schoch (2014:99)).
-It is therefore not directly possible to elicit households and household inomce out of tax data.
+On the other hand, is it quit likely that more than one tax units live in the same household (unmarried/unregistered couples, see \citet[99]{muller_vermogenslage_2014}).
+It is therefore not directly possible to elicit households and household income out of tax data.
 This might influence the assessment inequality, taking into account the change from traditional household and family structures over the last century.
@@
 Often inequality is assessed on national level, which implies, that studies try to cover the whole population of the country of interest.
-This is a special thorny task for surveys working with samples, because nonresponse is a major source of bias (Bethlehem et al., 2011).
-Korinek, Mistiaen and Ravallion (2006) show, that the position in the income distribution influence the probability to participate in a survey.
+This is a special thorny task for surveys working with samples, because nonresponse is a major source of bias \citep{bethlehem_handbook_2011}.
+\citep{korinek_survey_2006} show, that the position in the income distribution influence the probability to participate in a survey.
 Low income and high income households are more likely to refuse survey response, which leads to an overrepresentation of middle income households.
-This process can be referred to as the "middleclass bias" Diekmann (2009).
+This process can be referred to as the "middleclass bias" \citep{diekmann_empirische_2009}.
 Missing data in household surveys is therefore not missing at random, which has an impact on the measures of inequality.
 The magnitude of this bias in Switzerland, however, is unknown.
-Strategies to handle this kind of bias are discussed in the literature (S?rndal et al.
-1992), but require a register for every unit, that is proportional to income.
-Currently there exists no such register for Switzerland (ecoplan 2014:43).
-Currently used microdata sets, which are used for official publications concerning inequality in Switerzland (SILC and HABE) are furthermore confronted with a constructed coverage problems, because this surveys rely on phoneregister, which excludes the households not having a registered connection.
+Strategies to handle this kind of bias are discussed in the literature \citep{S?rndal et al.
+1992 FEHLT}, but require a register for every unit, that is proportional to income.
+Currently there exists no such register for Switzerland \citep{ecoplan 2014:43 FEHLT}.
+Currently used micro datasets, which are used for official publications concerning inequality in Switzerland (SILC and HABE) are furthermore confronted with a constructed coverage problem, because these surveys rely on the phone register, which excludes the households not having a registered connection.
 \\
@@
 The issue of incomplete coverage is far less dramatical with tax data.
-Essentially every permanent resident in Switzerland, who has completed the age of 18 years (respectivley 20 years prior to 1996) is taxed on a yearly base (or every two years before the change of the tax system).
-In principle this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distrubtion.
+Essentially every permanent resident in Switzerland, who has completed the age of 18 years (respectively 20 years prior to 1996) is taxed on a yearly base (or every two years before the change of the tax system).
+In principle this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distribution.
 This includes a separation of normal cases, which embrace the majority of taxpayers, and the special cases, which cover several unusual situations.
@@
 To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
-Vgl.
-Allison (1978), Engelhardt (2000), Cowell (2000) oder Hao/Naiman (2010)
+See \citet{Allison 1978 FEHLT}, \citet{engelhardt_modelle_2000}, \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}.
 %[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
 
-neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences (Handcock/Morris 1999) \\
+neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences \citep{handcock_relative_1999} \\
 
-Gini is mainly used for international comparison.
-However, other measures are often reported along (gini is more sensitive to changes in the middle of the distribution than to changes in the tails (OECD, 2008:37).
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) (Alderson and Doran 2013).
+The Gini coefficient is mainly used for international comparison.
+However, other measures are often reported along because the Gini coefficient is more sensitive to changes in the middle of the distribution than to changes in the tails \citep[37]{OECD, 2008 FEHLT}.
+Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) \citep{Alderson and Doran 2013 FEHLT}.
 \\
 
-Alvaredo (2010) shows formally, how top incomes shares are related to the gini-coefficient.
+\citet{Alvaredo (2010) FEHLT} shows formally, how top incomes shares are related to the Gini-coefficient.
 Following his argument, it is crucial to consider top incomes.
-Leigh (2007) also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the studie).
+\citet{leigh_how_2007} also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the study).
 %[He also discusses tax data studies in general-> have a closer look] \\
@@
 Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
-\item Some tax units are not included in the tax tables (see Foellmi and Martinez 2013:11f).(a) only taxed cases are included (partly a problem, for certain periodes we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
+\item Some tax units are not included in the tax tables (see \citealt[11f]{foellmi_volatile_2013}).(a) only taxed cases are included (partly a problem, for certain periods we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
 Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
@@
 %Can we ignore this)
-(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registred.
+(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registered.
 This persons get an imputed income (older tax tax return and information given by employers.
@@
 - (e) tax evasion.
-Feld and Frey (2007) report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
-With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequaly measures over time.
+\citet{feld_tax_2006 UNSICHER weil vorher 2007 stand} report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
+With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequality measures over time.
 \end{itemize}
@@
 
-Atkinson et al. (2009) estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by Alvaredo (2010). \\
+\citet{atkinson_top_2009} estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by \citet{Alvaredo (2010) FEHLT}.
+\\
 
-See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
+See for other countries: \citealt{Siminski et al 2003 FEHLT} (Australia), \citealt{Brewer et al 2008 FEHLT}, UK, \citealt{Burkhauser et al. 2009 FEHLT}, US)
 %	•	What about the special cases?
@@
 It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
-Credit Suisse (2011) makes use of the information in the ``Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
+\citet{credit_suisse_global_2011} makes use of the information in the ``Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
 \\
 
-Additonaly the Credit Suisse Reports states (2011:8) that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts.
+Additionally the Credit Suisse Reports states that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts \citep[8]{credit_suisse_global_2011}.
 
@@
 \textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
-Del et al.
-(2007) impute for non-fillers the 20 percentage of the annual average income.
+\citet{dell_income_2007} impute for non-fillers the 20 percentage of the annual average income.
 This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
-Furthermore Del et al.
-calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
+Furthermore the authors calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
 \\
@@
 This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
-This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level  (Foellmi and Martinez:8f).
+This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level \citep[8f]{foellmi_volatile_2013}.
 \\
@@
 
-\textbf{deductions} Del et al.
-(2007:477):" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
-Gorgas and Schaltenegger (2011:5): ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
+\textbf{deductions} \citet[477]{dell_income_2007}:" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
+\citet[5]{schaltegger_evolution_2011 UNSICHER}: ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
 Können wir das auch?
@@
 On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a ``perverse redistribution'' effect by redistributing income from the lower middle class to the upper middle class (vgl.
-Peters 2011 and Interpellation Barbara Gysel (2009).
+There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a ``perverse redistribution'' effect by redistributing income from the lower middle class to the upper middle class (see \citealt{peters_steuerabzuge:_2011} and \citet{Interpellation Barbara Gysel (2009) FEHLT.
+das interpellation barbara sieht falsch aus}.
 
@@
 \item There is some variance and we see different patterns over time and cantons.
-When estimating gini coefficients or the like we must therefore assume that ignoring ``the zeros'' leads to a bias that is not stable over time.
+When estimating Gini coefficients or the like we must therefore assume that ignoring ``the zeros'' leads to a bias that is not stable over time.
 \end{enumerate}
@@
 \begin{enumerate}
-\item Add the zeros as a seperate group
-\item Fit a model to predict the inequality measure (e.g. the gini coefficient) using the share of zeros as a predictor
+\item Add the zeros as a separate group
+\item Fit a model to predict the inequality measure (e.g. the Gini coefficient) using the share of zeros as a predictor
 \end{enumerate}
 
-By adding the zeros as a seperate group we face several problems.
+By adding the zeros as a separate group we face several problems.
 With the exception of Geneva 1995/96 Pareto interpolation of percentiles p20 and above seems viable.
-Low percentiles however would need to be extrapolated so we might impose unrealistic assuptioms.
-Even when estimating ``safe'' measures above p20 (better p50) or a gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
-To be consistent with the measure of ``taxable income'' we could assume an income of zero for that group, calculate gini coefficients and compare them with the original (uncorrected/plain) version.
-To get a rough number we could (cantonwise) check the squared correlation between plain and corrected gini.
+Low percentiles however would need to be extrapolated so we might impose unrealistic assumptions.
+Even when estimating ``safe'' measures above p20 (better p50) or a Gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
+To be consistent with the measure of ``taxable income'' we could assume an income of zero for that group, calculate Gini coefficients and compare them with the original (uncorrected/plain) version.
+To get a rough number we could (by cantons) check the squared correlation between plain and corrected Gini.
 
@@
 In our case we can clearly reject the hypothesis that all interactions are zero ($p=\Sexpr{round(ftest$p,3)}$).
-This leads to the conclusion  that gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted gini coefficients.
+This leads to the conclusion  that Gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted Gini coefficients.
 For example one might be interested in how inequality would had developed if the zero-share would have been constant over time.
 (Note RF: predict all data point using canton, time and the initial OR final zero-share to homogenize the time series)
-Furthermore we can quantify how large the bias is and we can do this seperately for tax periods or seperately for cantons.
+Furthermore we can quantify how large the bias is and we can do this separately for tax periods or separately for cantons.
 
@@
 
-We can see the the model fit reduces to explaining 61.5\% of the gini variation versus 92.6\% when the information from the zero-shares was used.
-Although this so some extent attributable to the additional 26 paramters: this is huge.
+We can see the the model fit reduces to explaining 61.5\% of the Gini variation versus 92.6\% when the information from the zero-shares was used.
+Although this so some extent attributable to the additional 26 parameters: this is huge.
 
@@
 A positive coefficient (e.g.
-Schwyz) can be read as follows: In periods with many zeros we measure higher gini coefficients.
-We can derive, that the distribution of income is more scewed for high incomes than for low incomes.
-Simply speaking, the contrast between low and middle class is less pronounced than the contrast between middle und high class.
+Schwyz) can be read as follows: In periods with many zeros we measure higher Gini coefficients.
+We can derive, that the distribution of income is more skewed for high incomes than for low incomes.
+Simply speaking, the contrast between low and middle class is less pronounced than the contrast between middle and top class.
 One possible explanation would be that incomes stem from two different populations: 1) the people of Schwyz who possibly follow a log-normal or gamma distribution and 2) particularly rich people who moved to Schwyz to avoid taxes.
@@
 A negative coefficient (e.g.
-Geneva) means the more zeros there are the smaller the gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
+Geneva) means the more zeros there are the smaller the Gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
 This is the case we would usually expect: more zeros mask inequality that arises from the bottom.
@@
 What can we conclude from that analysis?
-First one must notice that aggregate measure like the gini (or others) do not always react in the same way when we cut off one part of the distribution, therefore the measures calculated from tax data is biased.
+First one must notice that aggregate measure like the Gini (or others) do not always react in the same way when we cut off one part of the distribution, therefore the measures calculated from tax data is biased.
 On the other hand, the model coefficient of Switzerland as a whole is not significant suggesting that the cantonal biases cancel out each other.

### 2014-06-09 paper/subdocuments/introduction.Rnw
citations and spell check

@@
 % 1.Abschnitt: Bedeutung von Ungleichheit
-Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular (Wilkinson and Pickett 2009).
-While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders (WEF 2013) and scholars alike.
-Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013;Salverda et al.
-2014).
-Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the ``hollowing of the the middle class'' (Alderson and Doran 2013).
+Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular \citep{wilkinson_income_2009}.
+While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders \citep{world_economic_forum_global_2013} and scholars alike.
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (\citealt{OECD, 2008 FEHLT}, \citealt{oecd_divided_2011}, \citealt{gornick_income_2013}, \citealt{salverda_changing_2014}).
+Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the ``hollowing of the the middle class'' \citep{Alderson and Doran 2013 FEHLT}.
 Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): ``our democracy is being put at peril.''
@@
 Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
-Atkinson (2013:8) observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
+\citet[8]{Atkinson 2013:8 FEHLT} observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
 On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality undisputedly.
 These concerns have led to the search of alternative data sources, which can supplement the established survey data studies.
-Already Kuznet (1955) used tax data to examine the relationship between economic growth and personal distribution of income.
-Then it took several decades until Piketty (2001,2003) made the use of tax data fashionable again.
-Following his approach studies on several countries were conducted (Atkinson and Piketty 2007, 2010).
-Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database (Alvaredo et al.
-2014).
+Already \citet{kuznets_economic_1955} used tax data to examine the relationship between economic growth and personal distribution of income.
+Then it took several decades until \citet{piketty_les_2001, piketty_income_2003,piketty_income_2003-1} made the use of tax data fashionable again.
+Following his approach studies on several countries were conducted \citep{atkinson_top_2007,atkinson_top_2010}.
+Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database \citep{Alvaredo et al.
+2014 FEHLT}.
 \\
@@
 % Offical Data Collections on development of inequality
-As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concering inequality in income.
+As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concerning inequality in income.
 What is known about Switzerland so far?
-Looking for offical data, three main sources has to be mentioned, which can be considered as de facto offical data sources: EU-SILC, HBS and LIS-data.
-Figure XY shows the results steming from this three sources while looking at Gini of equivalised disposable income.
+Looking for official data, three main sources has to be mentioned, which can be considered as de facto official data sources: EU-SILC, HBS and LIS-data.
+Figure XY shows the results stemming from this three sources while looking at Gini of equivalised disposable income.
 Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level.
@@
 As a Non-EU member Switzerland implemented the instrument not from the beginning (2004) but as from 2007.
-Therefore this times-series doesn't cover timepoints before 2007.
-As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concering the distribution of income is the Household Budget Survey (HBS).
-The main focus of this survey lays in providing detailed data on houshold budegts.
+Therefore this times-series doesn't cover time points before 2007.
+As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concerning the distribution of income is the Household Budget Survey (HBS).
+The main focus of this survey lays in providing detailed data on household budgets.
 This allows researcher to look at different income concepts like income before and after public transfers.
-Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistend time series from 2000 to 2011.
+Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistent time series from 2000 to 2011.
 As it can be seen from graph XY the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period.
@@
 Data-provider for the LIS Data is the Swiss Federal Statistical Office too.
-In contrast the the aformentioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+In contrast the the aforementioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
 All in all the LIS dataset contains the longest time series on inequality for Switzerland.
-Analyzing this data Gornick and Jäntti (2013) found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
-This result is supported by Grabka and Kuhn (2012) analyzing the Swiss Household Panel (2000-2009).\footnote{A further offical database for income distribution is the OECD-Database.
+Analyzing this data \citet{gornick_income_2013} found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
+This result is supported by \citet{grabka_evolution_2012} analyzing the Swiss Household Panel (2000-2009).\footnote{A further official database for income distribution is the OECD-Database.
 It includes measures from Income and Consumption survey as well.
@@
 Comparison before and since 2008 is not recommended (OECD 2012:315).
-A further important database on inequality is the GINI database which has been derived from the GINI Country Reports (Nolan et al.,2014).
+A further important database on inequality is the GINI database which has been derived from the GINI Country Reports \citep{nolan_changing_2014}.
 But this dataset doesn't cover Switzerland.}  \\
@@
 Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
-Dell et al.
-(2007) used tax data from the Federal Tax Administration to assess the development of concentration of the highest incomes and wealth (top-shares).
+\citet{dell_income_2007} used tax data from the Federal Tax Administration to assess the development of concentration of the highest incomes and wealth (top-shares).
 In contrast to most other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
-Using the same approach Foellmi and Martinez (2013) expand the Dell et al.
-timeline to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
+Using the same approach \citet{foellmi_volatile_2013} expand the Dell et al.
+time series to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
 A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
-Flückiger et al.
-(2007) and also Jeitziner and Peters (2007,2009) report constant inequality from 1960-1996 respectively from 1995 to 2003.
-Covering a similar time period Bauer and Spycher (1994) and Bolzani and Abul Naga (2002) found decreasing inequality.
-On the other hand Buchmann and Sacchi (1995) and Ernst et al.
-(2009) found an increase in the 1980s).} \\
+\citet{Flückiger et al.
+(2007) FEHLT} and also \citet{jeitziner_regionale_2007, jeitziner_regionale_2009} report constant inequality from 1960-1996 respectively from 1995 to 2003.
+Covering a similar time period \citet{bauer_verteilung_1994} and \citet{Bolzani and Abul Naga 2002 FEHLT oder es ist 2001} found decreasing inequality.
+On the other hand \citet{Buchmann and Sacchi (1995) FEHLT oder incomplete, siehe buchmann_zur_????} and \citet{Ernst et al.
+2009 FEHLT} found an increase in the 1980s).} \\
 
@@
 Third, different income concepts and different units of analysis were used.
-As it is shown by Modetta and Müller (2012) income distribution is strongly affected by governmental redistribution, reducing inequality substantially.
+As it is shown by \citet{modetta_einkommensungleichheit_2012} income distribution is strongly affected by governmental redistribution, reducing inequality substantially.
 With the focus on tax data the change in institutional settings is not covered.
@@
 First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
-Second, we summarize and apply tax data specific technics to construct suitable measures of inequality.
+Second, we summarize and apply tax data specific techniques to construct suitable measures of inequality.
 We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
-Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through datasource.
+Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through data source.
 

### 2014-06-09 paper/subdocuments/data_and_methods.Rnw
FTA data description + comparison table (FTA vs survey)

@@
 Thirdly, one has to choose an appropriate measure of inequality.
-All this questions are ideally answered considering theory and a given research question.
+All these questions are ideally answered considering theory and a given research question.
 Often it has to be answered in context of a given dataset.
@@
 Based on a review on the literature about the measurement concepts in an ideal world, we discuss the advantages and shortcomings of tax data compared to other data sources - namely survey data.
-Along with the description we explain the methods and techniques we use to construct time-series of inequality-measure for income for Switzerland, which we will present in the result sections.
-By doing this, we demonstrate the suitability of the FTA Tax Data to assess the development of inequality in Switzerland
+We describe and explain methods and techniques needed to construct time-series of income inequality measures for Switzerland.
+We will present these time-series in the result section and evaluate the use of the FTA Tax Data to assess the development of inequality in Switzerland
 
@@
 Gleichzeitig sollen die Daten vorgestellt werden, die wir f?r die sp?teren Analysen verwenden.
+
+Our data comes from the Swiss Federal Tax Administration (FTA).
+Federal taxes are collected and documented by the FTA since 1915.
+Being called a war-tax in the beginning, the federal tax was renamed to crisis levy in 1934, defense-tax in 1939 and is finally known as direct federal tax since 1983.
+The time frame we were able to collect ranges from 1945 to 2010 including 44 tax periodes.
+\footnote{Between 1993 and 2003 there is no exact data available for Switzerland because of a system change from taxation assessed in arrears (Praenumerando-System) to taxation assessed on current year income (Postnumerando-System) which was implemented by cantons in different years.}  While the FTA provides data in electronic form since 1973 we collected earlier data by scanning hard copies.
+Data is available for Switzerland plus all cantons and basically covers every tax unit (individual or household) in Switzerland liable to pay federal taxes.
+This exempts all tax units with taxable income below a certain threshold (e.g.
+15.000 CHF in 2010).
+Furthermore the FTA differentiates between two groups of tax units, so called normal cases and special cases.
+A normal case is a tax unit residing in a swiss canton without foreign source income and being liable to taxation all year long.
+All other tax units and very few that are taxed based on the style of living because they don't work (Pauschalbesteuerte) are special cases.
+The FTA provides two income measures: taxable income and net income.
+Net income here is an administrative term and means taxable income plus social deductions (children and supported persons) but not including other deductions like donations or health-care costs.
+Both measures are designed for taxation purposes which might limit the suitability to measure inequality as we will discuss later.
 
@@
 Most important high net wealth individuals taxed according to their expenditures.However, special attention has to be paid to tax units with none or very low incomes.
-Even though they have to hand in a tax return, their income do not show up in the statistics if their income after deductions falls below 15'000 CHF and they are therefore not taxed with direct federal taxes.
+Even though they have to hand in a tax return, their income does not show up in the statistics if their income after deductions falls below 15'000 CHF and they are therefore not taxed with direct federal taxes.
 This is possible for normal and special cases alike.
@@
 % What about the DELL Study, where high share of non-fillers are reported? What are the reasons?
+
+\emph{Intertemporal comparison}
+
+One advantage of tax data over survey data is their observation period.
+While the FTA data we use range back to 1945, the earliest period of survey data is 2000 (HABE data).
+To be fair, survey data might cover the late periods (2011 and 2012) which are not covered by tax data.
+The striking feature of the long FTA time series is its consistency.
+Both population coverage and income measures are consistent for the complete observation period.
+This can not be said for survey data as these are based on samples and therefore require an ideal sampling design or reweighting to be representative and comparable over time.
+We will see how well sampling and weighting is done in the results section.
 
@@
 %- contingency for intertemporal comparision
+
+\begin{table}[h]
+\begin{tabular}{lcccc}
+                                   & FTA       & HABE      & EU-SILC   & LIS       \\
+Range                              & 1945-2010 & 2000-2011 & 2007-2012 & 1982-2004 \\
+Cross country comparison possible? & no        & no        & yes       & yes       \\
+Top incomes covered?               & yes       & no        & no        & no        \\
+Low incomes covered?               & no        & yes       & yes       & yes       \\
+Concepts of economic ressources?   & no        & yes       & ?         & ?         \\
+Equivalisation?                    & no        & yes       & ?         &          
+\end{tabular}
+\end{table}
 

### 2014-06-11 code/habe.r
Update HABE-Preparation Syntax with Replicationcode

@@ -49,6 +49,11 @@ habe0011$Gini[12]<-by(habe0911$VerfuegbaresEinkommen08gew,habe0911$Jahr08,gini)[
 habe0011$Gini[11]<-by(habe0911$VerfuegbaresEinkommen08gew,habe0911$Jahr08,gini)[2]
 habe0011$Gini[10]<-by(habe0911$VerfuegbaresEinkommen08gew,habe0911$Jahr08,gini)[1]
 
+##
+# Exkurs: Replikation offizieler Werte zum verfügbaren Einkommen
+
+weighted.mean(habe0911$VerfuegbaresEinkommen08,habe0911$Gewicht10_091011)
+
 # 2006 bis 2008
 habe0608<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2006 bis 2008/HABE060708_Standard_130124UOe.txt", header=TRUE)
 habe0608$AnzErw<-habe0608$AnzahlPersonen98-habe0608$AnzahlKinder05
@@ -58,9 +63,6 @@ habe0011$Gini[09]<-by(habe0608$VerfuegbaresEinkommen08gew,habe0608$Jahr08,gini)[
 habe0011$Gini[08]<-by(habe0608$VerfuegbaresEinkommen08gew,habe0608$Jahr08,gini)[2]
 habe0011$Gini[07]<-by(habe0608$VerfuegbaresEinkommen08gew,habe0608$Jahr08,gini)[1]
 
-####
-# Achtung! Die konstruktion des verfügbaren Einkommens stimmt nicht!!!!!!!!!!!!
-####
 
 
 # 2001 bis 2005
@@ -73,30 +75,56 @@ habe05<-merge(habe05,habe05hh)
 habe05$y.sum.ag<-habe05$y.sum/(1+0.5*(habe05$ANZ_PERSONEN-1))
 habe0011$Gini[06]<-gini(habe05$y.sum.ag,habe05$GEWICHT)
 
-#
+##
+# Exkurs: Replikation offizieler Werte zum verfügbaren Einkommen
 habe05long<-reshape(habe05,direction="wide",idvar="HAUSHALT_ID",timevar="NOMENKLATUR_STUFE1_ID",v.names="SUMME_BETRAG_CHF")
 habe05long[is.na(habe05long)]<-0
 habe05long<-merge(habe05long,habe05hh)
-head(habe05long)
-#
[…]

### 2014-06-11 code/habe.r
update income definiton habe00

@@ -189,14 +189,14 @@ habe00$y<-ifelse(habe00$B_AUSGABE==0, habe00$SUMME_BETRAG_CHF, ifelse(habe00$NOM
 
 habe00hh<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2000 bis 2002/HAUSHALT_2000.txt", header=TRUE,sep="\t")
 habe00<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2000 bis 2002/EINTRAG_HH_AGGREGAT5.txt", header=TRUE)
-habe00long<-reshape(habe00,direction="wide",idvar="HAUSHALT_ID",timevar="NOMENKLATUR_STUFE5_ID",v.names="SUMME_BETRAG_CHF")
-habe00$bruttoeinkommen<-ifelse(habe00$B_AUSGABE==0, habe00$SUMME_BETRAG_CHF,0)
+habe00$bruttoeinkommen<-ifelse(habe00$B_AUSGABE==0,ifelse(habe00$NOMENKLATUR_STUFE5_ID%in%c(50,51,52,63,64,65,66,67,68,69,70),0,habe00$SUMME_BETRAG_CHF),0)                     
 habe00long<-summaryBy(bruttoeinkommen~HAUSHALT_ID,FUN=sum,data=habe00)
 # Test Bruttoeinkommen
 habe00long<-merge(habe00long,habe00hh)
 weighted.mean(habe00long$bruttoeinkommen.sum,habe00long$GEWICHT)
 # zu hoch
 # Test Ausgaben
+habe00long<-reshape(habe00,direction="wide",idvar="HAUSHALT_ID",timevar="NOMENKLATUR_STUFE5_ID",v.names="SUMME_BETRAG_CHF")
 habe00long[is.na(habe00long)]<-0
 habe00long$ausgaben<-(habe00long$SUMME_BETRAG_CHF.821+habe00long$SUMME_BETRAG_CHF.822+habe00long$SUMME_BETRAG_CHF.823+habe00long$SUMME_BETRAG_CHF.824+habe00long$SUMME_BETRAG_CHF.825+
                            habe00long$SUMME_BETRAG_CHF.843+habe00long$SUMME_BETRAG_CHF.844+habe00long$SUMME_BETRAG_CHF.845+habe00long$SUMME_BETRAG_CHF.847+

### 2014-06-11 paper/subdocuments/data_and_methods.Rnw
minor changes method section

@@
 % What about the DELL Study, where high share of non-fillers are reported? What are the reasons?
+% Bisher nirgends erwähnt ist die Kritik an Steuerdaten bezüglich tax avoidance and tax evasion.
 
@@
 See \citet{Allison 1978 FEHLT}, \citet{engelhardt_modelle_2000}, \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}.
-%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
-
-neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences \citep{handcock_relative_1999} \\
-
 The Gini coefficient is mainly used for international comparison.
@@
 Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) \citep{Alderson and Doran 2013 FEHLT}.
-\\
+%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
 
-\citet{Alvaredo (2010) FEHLT} shows formally, how top incomes shares are related to the Gini-coefficient.
+%\citet{Alvaredo (2010) FEHLT} shows formally, how top incomes shares are related to the Gini-coefficient.
 Following his argument, it is crucial to consider top incomes.
@@
 %- contingency for intertemporal comparision
+%- flexibel applications of appropriate inequality measures
 
@@
 
+% Die Punkte kurz beschreiben.
+
 %Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
-
-% 3. Eigenheiten Steuereinheiten
 
@@
 
+% Könnte man in der Introduction platzieren.
+
 See for other countries: \citealt{Siminski et al 2003 FEHLT} (Australia), \citealt{Brewer et al 2008 FEHLT}, UK, \citealt{Burkhauser et al. 2009 FEHLT}, US)
@@
 %	•	Capital gains? Should not be included
-The discussion about problems with reporting income is fairly exhaustive. What about wealth?
-
-It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
-\citet{credit_suisse_global_2011} makes use of the information in the ``Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
-\\
-
-Additionally the Credit Suisse Reports states that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts \citep[8]{credit_suisse_global_2011}.
 
 \subsection{Ways to tackle FTA-tax data specific problems}
+
+% Diese Abschnitt muss ersetzt werden mit der von uns gewählten Vorgehensweise
 
@@
 Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
+
+%The discussion about problems with reporting income is fairly exhaustive. What about wealth?
+
+%It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
+\citet{credit_suisse_global_2011} makes use of the information in the ``Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
+\\
+
+%Additionally the Credit Suisse Reports states that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts \citep[8]{credit_suisse_global_2011}.

### 2014-06-12 paper/subdocuments/introduction.tex
- complected first draft of section measurement concepts - several minor changes

@@
 % 1.Abschnitt: Bedeutung von Ungleichheit
-Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular (Wilkinson and Pickett 2009).
-While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders (WEF 2013) and scholars alike.
-Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (OECD, 2008, 2011; Gornick and Jäntti, 2013;Salverda et al.
-2014).
-Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the "hollowing of the the middle class" (Alderson and Doran 2013).
-Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): "our democracy is being put at peril."
+Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular \citep{wilkinson_income_2009}.
+While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders \citep{world_economic_forum_global_2013} and scholars alike.
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (\citealt{OECD, 2008 FEHLT}, \citealt{oecd_divided_2011}, \citealt{gornick_income_2013}, \citealt{salverda_changing_2014}).
+Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the ``hollowing of the the middle class'' \citep{Alderson and Doran 2013 FEHLT}.
+Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): ``our democracy is being put at peril.''
 \\
@@
 % Offical Data Collections on development of inequality
-As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concering inequality in income.
+As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concerning inequality in income.
 What is known about Switzerland so far?
-Looking for offical data, three main sources has to be mentioned, which can be considered as de facto offical data sources: EU-SILC, HBS and LIS-data.
-Figure XY shows the results steming from this three sources while looking at Gini of equivalised disposable income.
+Looking for official data, three main sources has to be mentioned, which can be considered as de facto official data sources: EU-SILC, HBS and LIS-data.
+Figure XY shows the results stemming from this three sources while looking at Gini of equivalised disposable income.
 Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level.
-The main focus of EU-SILC is to collect data on a common "framework" to ensure comparability among EU-countries and countries living around or within the EU.
+The main focus of EU-SILC is to collect data on a common ``framework'' to ensure comparability among EU-countries and countries living around or within the EU.
 As a Non-EU member Switzerland implemented the instrument not from the beginning (2004) but as from 2007.
-Therefore this times-series doesn't cover timepoints before 2007.
-As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concering the distribution of income is the Household Budget Survey (HBS).
-The main focus of this survey lays in providing detailed data on houshold budegts.
+Therefore this times-series doesn't cover time points before 2007.
+As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concerning the distribution of income is the Household Budget Survey (HBS).
+The main focus of this survey lays in providing detailed data on household budgets.
 This allows researcher to look at different income concepts like income before and after public transfers.
-Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistend time series from 2000 to 2011.
+Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistent time series from 2000 to 2011.
 As it can be seen from graph XY the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period.
@@
 > library(foreign)
+> library(reshape, quietly=TRUE, warn.conflicts = FALSE)
 > number_ticks <- function(n) {function(limits) pretty(limits, n)}
-> swissGini<-read.csv("data/swissGini.csv",header=TRUE,sep=";")
+> swissGini<-read.csv("../../data/swissGini.csv",header=TRUE,sep=";")
+> swissGini_long<-melt(swissGini,id="Year")
+> ggplot(data=swissGini_long,
++   aes(y=value,x=Year),shape=variable)+
++   ylab("Gini of equivalised disposable income") +
++   scale_y_continuous(limits=c(0.18,0.4)) +
++   scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
++   geom_point(aes(shape=variable),size=3)+
++   geom_line(data=swissGini_long[!is.na(swissGini_long$value),],aes(linetype=variable))+
++   theme_bw()
+\end{Sinput}
+\end{Schunk}
+
+% 4.Abschnitt: What do we know about Switzerland
+% Studies with tax data
+Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
+\citet{dell_income_2007} used tax data from the Federal Tax Administration to assess the development of concentration of the highest incomes and wealth (top-shares).
+In contrast to most other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
+Using the same approach \citet{foellmi_volatile_2013} expand the Dell et al.
+time series to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
+A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
+\citet{Flückiger et al.
+(2007) FEHLT} and also \citet{jeitziner_regionale_2007, jeitziner_regionale_2009} report constant inequality from 1960-1996 respectively from 1995 to 2003.
+Covering a similar time period \citet{bauer_verteilung_1994} and \citet{Bolzani and Abul Naga 2002 FEHLT oder es ist 2001} found decreasing inequality.
+On the other hand \citet{Buchmann and Sacchi (1995) FEHLT oder incomplete, siehe buchmann_zur_????} and \citet{Ernst et al.
+2009 FEHLT} found an increase in the 1980s).} \\
+
+% 5. Abschnitt: Zwischenbilanz und Herausforderungen
+Divergence can be explained with several factors.
+First of all, different data sources were used.
+The official data providers trust on survey data, whereas the later mentioned publications use tax data.
+It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality.
+On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the ``hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
+Different measure of inequality hampers the comparability.
+Third, different income concepts and different units of analysis were used.
+As it is shown by \citet{modetta_einkommensungleichheit_2012} income distribution is strongly affected by governmental redistribution, reducing inequality substantially.
+With the focus on tax data the change in institutional settings is not covered.
+Also neglected is the household structure, whereas it is unclear how inequality is affected whether one looks at household income or at income of tax units.
+It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
+
+% Zusammenfassung Stand der Forschung und festlegen unserer Ausrichtung
+
+Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study, EU-SILC).
+However, it is unclear how the bias through non-response affects the overall measure of inequality.
+Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
+Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.
+First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
+Second, we summarize and apply tax data specific techniques to construct suitable measures of inequality.
+We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
+Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through data source.
+
+% Könnte noch besser auf die geplanten Analysen abgestimmt sein (vgl. auf Github-Issues)
+
+% Was wollen wir wissen?
+
+% Wir müssen uns äussern zu: 
+% Analyseinheiten
+% Definition des Einkommens
+% Messung der Verteilung
+
+% Bei den Ergebnissen müsste es Grafiken haben zu:
+% 1) Unterschiedlichen Analyseneinheiten (Mit und ohne Nuller, Normal- und Sonderfälle)
+% 2) Unterschiedliche Einkommensgrössen (Reineinkommen, steuerbares Einkommen, pseudoverfügbares Einkommen (Abzüge von direkten Bundessteuern))
+% 3) Vergleich mit HABE
+
+% Textbaustein politische Dimension von Ungleichheit
+% What is known about Switzerland so far?
+Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
+Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
+The debate went along with referendums trying to regulate the market outcomes.
+Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisingly high majority of 68\%, the 1:12-Initiative, which aimed at the whole spectrum of the income distribution, was rejected.
+The voting about the minimum wage-initiative will be held on 9th February 2014.
+All these referendums questioned the current distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
+Several official and semi-official publications addressed the question of rising inequality in Switzerland.
+\\
+
+% Alte Textbausteine zu Vermögen
+%While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
+The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
+But, up to the day information for Switzerland is not available.
+% ab hier wird es irgnediwe unklar
+%Shorrock et al.
+(2013) unify several datasources (including tax data) to assess the pattern of global wealth.
+Their databook includes figures for Switzerland.
+In contrast to inequality of income concerning inequality of wealth Switzerland takes a leading position \\
+% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
+% Dann lassen sich einige Passagen streichen.
+Man könnte auch auf verschiedene Publikationen verweisen, welche die Bedeutung von Vermögen untersuchen -> Buch von Piketty, Müller und Schoch und dann aber ziemlich schnell darauf verweisen, dass wir uns hier auf Einkommen konzentrieren.
+% Auch erste Dell Studie erwähnen -> es hat Zahlen zur Schweiz. Auch Probleme erwähnen (Fehlende Informationen zu Altersvorsorgevermögen -> hat aber niemand eine Lösung bisher.)

### 2014-06-12 code/habe.r

@@ -225,6 +225,10 @@ ggplot(data=habe0011,aes(x=Year,y=Gini))+
 ### Analyses #################################################
 ##############################################################
 
+habe<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+
+
+
 ggplot(habe, aes(x=log(VerfuegbaresEinkommen08)))+geom_histogram()+facet_wrap(~Jahr08)+theme_bw()
 ggplot(habe, aes(y=log(VerfuegbaresEinkommen08),x=percentiles))+geom_line()+facet_wrap(~Jahr08)+theme_bw()
 ggplot(habe, aes(x=log(VerfuegbaresEinkommen08),group=Jahr08,color=factor(Jahr08)))+geom_density()+theme_bw()

### 2014-06-12 paper/subdocuments/data_and_methods.Rnw
several updates and outline of applied method section

@@
 
-% Hier sollten kurz die Steuerstatistiken beschrieben werden.
-Ziel ist es eine Basis f?r die nachfolgende Gegen?berstellung von Soll und Ist zu legen.
-Gleichzeitig sollen die Daten vorgestellt werden, die wir f?r die sp?teren Analysen verwenden.
-
 Our data comes from the Swiss Federal Tax Administration (FTA).
@@
 All other tax units and very few that are taxed based on the style of living because they don't work (Pauschalbesteuerte) are special cases.
+\\
+
 The FTA provides two income measures: taxable income and net income.
@@
 
-%Beschreiben was die Daten beinhalten und wo es Probleme gibt
-\begin{itemize}
-\item  Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
-	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
-	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
-	\item	Income after deductions-> taxable income (employment income, business income and capital income).
-Is income definition stable over time?
-Yes it should.
-Realized capital gains are excluded from the definition.
-It includes income from  employment, self-employment, capital income and taxable transfer payments.
-Plus Eigenmietwert,
-	\item	Reported on national and cantonal level
-\end{itemize}
+% Bei den Nullern erw?hnen, ab wann Angaben zur Zahl der Nuller vorliegen und ab wann nicht
+% Im Internet hat es Publikationen ab 1947/48
+% Es gibt auch Tabellen auf Gemeindeebene
+% Die Statitstischen Kennzahlen von Br?lhart erw?hnen?
+% Erw?hnen in welcher Form die Informationen publiziert sind. Tabulation by size of income. Not on individual or household level
+
+% http://www.estv.admin.ch/dokumentation/00075/00076/00701/index.html?lang=de
+% Kantone und Gemeinden: 2010 bis 1947/1948
+% Klassierte Tabellen für Abzüge, Einkommen (Steuerbares Einkommen und Reineinkommen) und Steuererträge
+% Normalfälle und Sonderfälle
+
+% Statistsiche Kennzahlen
+% Mit Belastung durch die direkte Bundessteuer: 1973/1974 bis 2010
+% Mit und ohne Belastung durch die direkte Bundessteuer: 1995/1996 bis 2010
+% steuerbares Einkommen/Reineinkommen
+% Jeweils Normal und Sonderfälle
+% Individual-Daten (nicht klassierte daten)
+% äquivalenzeinkommen
 
@@
 Albeit the awareness of an assessment of income, consumption and wealth simultaneously is rising, we  focus our analysis on income, which is undoubtedly a crucial indicator of economic wellbeing.
-However, the Federal Tax Office publishes statistics on income and wealth .
+It should be noted, that the Federal Tax Office publishes statistics on income and wealth.
 But it is not possible to analyze the jointly distribution on the individual or household level.
@@
 Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013 UNSICHER}, \citet[24]{united_nations_canberra_2011}).
-Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution.
-Within this framework common income definitions are situated.
+Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually leads to the disposable income.
+The income measure, which finaly shapes the possibility to consume.
+Within this framework common other income definitions are situated.
 These definitions are contrasted with income concepts, which can be derived out of the FTA Tax data.
@@
 
-The central income according to tax data is the taxable income.
-It includes all reported incomes (income from employment, income from property and received transfers\footnote{Mean-tested benefits are not taxed and therefore not included in Tax data.
-Income for low income groups are therefore underestimated.}) minus several deductions.
+% ACHTUNG: Wir m?ssen innerhalb des Papers einheitlich bleiben. In der Grafik steht jetzt "Total income" oben steht "Net income"
+
+% Man k?nnte es auch konzeptionell auf kantonale Steuerdaten erweitern. Besser w?re es, dies einfach im Ausblick zu erw?hnen
+
+The central income reported through tax statistics is the taxable income.
+It includes all reported incomes (income from employment, income from property and received transfers\footnote{Realized capital gains are not taxed and Mean-tested benefits are not taxed and therefore not included in Tax data.
+Income for low income groups are therefore underestimated.
+However, Piketty and Saez (2001) note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had nonly a trivial impact on top income shares.}) minus several deductions.
 It is therefore neither a pre-transfer income  nor a post-transfer income measure.
@@
 
-% 1.Abschnitt dar?ber, dass besser Haushalte betrachtet werden anstelle von Individuuen
-
 The agreed standard on the statistical unit, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013 UNSICHER}.
@@
 \\
-
-% 2. Abschnitt zur Gegen?berstellung von Haushalten und Steuereinheiten
 
@@
 
+\emph{Measuring inequality or concentration} \\
+To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
+The Gini coefficient is the most known measure and mainly used for international comparison.
+As it is derived from the lorenz-curve, the quantified amount of inequality can unpretentiously be described in a formal and visual way.
+Therefore the Gini coefficient is easely interpretable.
+Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}.
+(1) principle of population: the assessment of inequality is independent of the population size (2) Bresciani-Turroni: the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
+doubling of all income) (3) weak principle of transfers (pigou-dalton): transfers from richer households lead to a reduction of inequality.
+However, several drawbacks are reported in the literatur.
+The most important point is, that the underlying distributional form of the measured inequality is unknown and it is therefore not possibel to see if the measure is driven be a few rich or many poor individuals.
+This can also be problematic for comparision between countries or over time.
+In extreme cases two totally different distrubtions share the same gini-coefficient.\\
+
+The recent wave of tax-data studies do not report Gini-coefficents.
+Rather top income shares are informed on, which are calculated not only with taxdata, but togehter with external sources to produce the population and personal income control totals.
+This procedure ensures, that the inequality measure is not biased because of non-fillers, who doesn't appear in tax statistics.
+Leight (2002:594f) compares top income shares with other inequality measures and ask, if they are a useful measure of inequality in a society.
+He tries to answer this question empiricaly by comparing measures of inequality based on top income shares with measures of household or familiy inequality.
+He finds a strong positive relationship, but concludes (P.600): top income shares are far from perfect as a measure of distribution of income across society.
+Top income shares hence inform not completely on how inequality evolves elsewhere in the distribution.
+Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent see above).
+A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between invidviduals, who are either both within the top group or both outsie the top group, then the share measure will remain unchanged.
+\\
+
+Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading).
+The polarityindex is developed by \citep{Handock and Morris 1999} allows for this.
+Recenently this index was applied in the work of \citep{Alderson, Beckfield and Nielsen 2005 FEHLT} and \citep{Alderson and Doran 2013 FEHLT}.
+This approach is rooted in relative distribution methods.
+It includes a precise comparison of the shape of two distributions (groups, over time).
+The main advantage lies herin, that this approche allows to characterize the change, in the sense that it is possible to say, if a polarization occured (reduction of middle class), which equals an increase of inequality and if this change is driven by a change in the upper or lower part of the distribution.
+\\
+
+The literature mentions several other metrics with desired properties, but we will not discuss them further (see for example \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}).
+
+% Allenfalls auch auf Quintilvergleiche eingehen und Atkinsonmass
+
 \emph{Population Coverage}
+% beim letzten Punkt geht es um Steuerhinterziehung. Fällt das auch noch unter Population coverage?
 
@@
 
-The issue of incomplete coverage is far less dramatical with tax data.
+The issue of incomplete coverage is less dramatical with tax data.
 Essentially every permanent resident in Switzerland, who has completed the age of 18 years (respectively 20 years prior to 1996) is taxed on a yearly base (or every two years before the change of the tax system).
-In principle this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distribution.
-This includes a separation of normal cases, which embrace the majority of taxpayers, and the special cases, which cover several unusual situations.
-Most important high net wealth individuals taxed according to their expenditures.However, special attention has to be paid to tax units with none or very low incomes.
+Essentially this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distribution.
+This includes a separation of normal cases, which embrace the majority of taxpayers, and the special cases, which cover Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
+Most important this includes high net wealth individuals taxed according to their expenditures.
+Special attention has to be paid to tax units with none or very low incomes.
 Even though they have to hand in a tax return, their income does not show up in the statistics if their income after deductions falls below 15'000 CHF and they are therefore not taxed with direct federal taxes.
@@
 From 1995/1996 until 2010 the number of non-taxed units are reported, but not for the years before.
+Dell et al.
+(2007) try to estimate the fraction of non-taxed by comparing the reported numbers of tax units to census reports about the number of adult population.
+Following his etimations this fraction drops from 94\% in 1993/1994 to 63\% back in 1945/46.
+\\
 
-% What about the DELL Study, where high share of non-fillers are reported? What are the reasons?
-% Bisher nirgends erwähnt ist die Kritik an Steuerdaten bezüglich tax avoidance and tax evasion.
+An other critical issue is the problem of tax evasion, which definitely can bias the assessment of inequality.
+Alvaredo and Saez (2006) for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
+Evasion can occured, when individuals try not to fill tax returns or by misreporting of incomes.
+In Switzerland non-fillers show up in the tax-statistics either way, as long as they are registered.
+This person get an imputed income form an older tax return and infomration given by employers.
+Only non registered non-fillers are not in the records.
+Therefore non-fillers are a minor problem.
+Not neglectable is the circumstance, that individuals misreport incomes.
+\citet{feld_tax_2006} examine the role of tax evasion in Switzerland by calculating the difference of the national accounts measures of primary income and the income reported to the tax authorities.
+Building on a time series of Pommerehne and Weck-Hannemann (1996) Feld and Frey (2002) and Frey and Feld (2002) extend the series.
+They can show, that the average level of income tax evasion from 1965 to 1995 varies between 13\% and 35\%.
+They suggest, that evasion is heavily driven by capital income tax evasion.
+\\
 
@@
 
-\emph{Measuring inequality or concentration} \\
-To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
-See \citet{Allison 1978 FEHLT}, \citet{engelhardt_modelle_2000}, \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}.
-The Gini coefficient is mainly used for international comparison.
-However, other measures are often reported along because the Gini coefficient is more sensitive to changes in the middle of the distribution than to changes in the tails \citep[37]{OECD, 2008 FEHLT}.
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) \citep{Alderson and Doran 2013 FEHLT}.
-%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
-
-%\citet{Alvaredo (2010) FEHLT} shows formally, how top incomes shares are related to the Gini-coefficient.
-Following his argument, it is crucial to consider top incomes.
-\citet{leigh_how_2007} also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the study).
-%[He also discusses tax data studies in general-> have a closer look] \\
-%[We definitely have to argue, why it isn’t sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
-
 \subsection{Comparison of tax data and other data sources - advantages and shortcomings}
@@
 %- Implementation of common concepts of economic ressources
-%- contingency for intertemporal comparision
+%- intertemporal comparision
 %- flexibel applications of appropriate inequality measures
@@
 
+% Braucht es alle drei Survey-Befragungen
+% ich würde nicht dichotomisieren, sondern ordinal bewerten (tief, mittel, hoch)
+
 % Die Punkte kurz beschreiben.
+% Zusammenfassung
+
+% Gedankenstütze
+%\textbf{Problems with household income surveys}
+%\begin{itemize}
+%\item Sample data (bias)
+%\item comparability between countries and over time (depends on income definition)
+%\item short time series
+%\end{itemize}
 
 %Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
+%Studies on problems with survey data See for other countries: \citealt{Siminski et al 2003 FEHLT} (Australia), \citealt{Brewer et al 2008 FEHLT}, UK, \citealt{Burkhauser et al. 2009 FEHLT}, US)
 
@@
 
-\textbf{Summary of tax data drawbacks}
-\begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial ``income'' see above)
-\item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the ``top-income'' literature, albeit it is a crucial issue in ``theorized'' concepts of income.
-To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
-This might influence the overall inequality, taking into account the change from traditional household and family structures over the last century.
-This is an inequality related issue, where relevant studies are missing.
-What are the assumptions?
-\item the tax data income definition summarize labor income, capital income and taxable transfer payments (no means-tested benefits as welfare aid).
-Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
-\item Some tax units are not included in the tax tables (see \citealt[11f]{foellmi_volatile_2013}).(a) only taxed cases are included (partly a problem, for certain periods we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
-Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
-In some border-cantons the share of this group is very high (20\%)
-%Can we ignore this)
-(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registered.
-This persons get an imputed income (older tax tax return and information given by employers.
-Not registered non-	fillers are not in the records.
-- (e) tax evasion.
-\citet{feld_tax_2006 UNSICHER weil vorher 2007 stand} report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
-With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequality measures over time.
-\end{itemize}
-
-\textbf{Problems with household income surveys}
-\begin{itemize}
-\item Sample data (bias)
-\item comparability between countries and over time (depends on income definition)
-\item short time series
-\end{itemize}
-
-\citet{atkinson_top_2009} estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by \citet{Alvaredo (2010) FEHLT}.
-\\
-
-% Könnte man in der Introduction platzieren.
-
-See for other countries: \citealt{Siminski et al 2003 FEHLT} (Australia), \citealt{Brewer et al 2008 FEHLT}, UK, \citealt{Burkhauser et al. 2009 FEHLT}, US)
-%	•	What about the special cases?
-Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
-Exclusion leads to an underestimation of inequality
-%	•	Capital gains? Should not be included
-
-\subsection{Ways to tackle FTA-tax data specific problems}
+\subsection{Outline of applied methods}
 
 % Diese Abschnitt muss ersetzt werden mit der von uns gewählten Vorgehensweise
+
+In the last section we described the advantages and drawback of taxdata discussing five aspects, which we regard as crucial concering the assessment of inequality.
+To get a feeling of the importance of these aspects, we exploit the FTA-tax data as far as possible and perform several insightful calculations adressing four of the five mentioned aspects.
+No further investigation is possible regarding aspect (1) concepts on measuring economic ressource.
+This is a point which cannot be further addresses with FTA tax data.
+But for the other four thematic areas, we can provide deeper insight.
+In general our main strategy is to apply different possible concepts within one conceptional area (income, units, measurement, population coverage) while holding other conceptual differences constant.
+With this strategy we want to show, the sensitiveness the assessement of inequality is especially if one looks at time trens.
+In this section we  describe which methods we use to produce the results in the next section.
+
+%%%
+% Was soll hier beschrieben werden und was direkt bei den Tabellen?
+%%% Gar nicht so einfach....
+
+%%% Nach Lehrbuch müssten die methodischen Schritte hier beschrieben werden.
+
+% Defining income
+% Beschreiben, dass sich aus den publizierten Einkommensgrössen und den Steuerbeträgen unterschiedliche 
+% Einkommen berechnen lassen
+% Es steht der Vorwurf im Raum, die unteren Einkommen seien falsch abgebildet, weil die Sozialhilfe nicht versteuert ist
+
+%%
+% statistical units
+% Beschreiben, dass zwar keine Haushaltseinkommen gebildet werden können und Skaleneffekte nicht abgebildet sind, dass eine Annäherung jedoch möglich ist
+% Definition der äquivalenzskala bei Brülhart (korrespondiert nicht mit OECD Definitionen, aus pragmatischen Gründen)
+
+%%
+% Measuring inequality
+% Beschreiben, wie sich aus gruppierten Steuerstatistiken der Gini-Koeffizient berechnen lässt
+% Beschreiben wie wir relative distribution technics einsetzen und den Polaritätsindex berechnen
+
+%%
+% Population coverage
+% Analysen zu den Nullern aufzeigen> ist eine Hauptkritik an der Nutzung von Steuerdaten
+% Die Bedeutung von Normal- und Sonderfällen
+% Den Vergleich mit der HABE beschreiben, welche Einkommensgrösse nutzen wir aus der HABE um diese möglichst mit den Steuerdaten vergleichbar zu machen?
+Wir brauchen das Bruttoeinkommen nach Abzug der Sozialversicherungsbeiträge, dies lässt sich mit dem Reineinkommen vergleichen
+
+%%
+% Intertemporal comparison
+% Ziel eine möglichst lange Zeitreihe abzubilden. 
+% Technik der Lückenfüllung von Ben beschreiben> dies kurz anderen Techniken gegenüberstellen.
+% Konfidenzintervalle für die Perioden mit vielen fehlenden Werte?
+
+% eine / oder zwei Zeitreihen (eine lange und eine, die möglichst nahe an der richtigen Konzeptionalisierung ist: post direct taxed income, normal+sonderfälle) zusammen mit den bereits existierenden Reihen (HABE,LIS,SILC) ploten
 
@@
 %is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
+% Beim Imputieren m?sste man auch die Zahl der freiwilligen Nichtbeziehenden ber?cksichtigen.
 

### 2014-06-12 paper/subdocuments/introduction.Rnw
- complected first draft of section measurement concepts - several minor changes
several updates and outline of applied method section

@@
 Comparison before and since 2008 is not recommended (OECD 2012:315).
+For the sake of completeness the database construtcted by \citet{Deininger and Squire 2006 FEHLT} and the World Income Inequality Database (WIID) can be mentioned.
+Both datasets do not cover more current figures for Switzerland.
 A further important database on inequality is the GINI database which has been derived from the GINI Country Reports \citep{nolan_changing_2014}.
@@
 However, it is unclear how the bias through non-response affects the overall measure of inequality.
-Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
+For the US \citet{atkinson_top_2009} estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by \citet{Alvaredo (2010) FEHLT}.
+Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) is missing.
 Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.

### 2014-06-13 paper/subdocuments/introduction.Rnw
minor change introduction

@@
 Comparison before and since 2008 is not recommended (OECD 2012:315).
-For the sake of completeness the database construtcted by \citet{Deininger and Squire 2006 FEHLT} and the World Income Inequality Database (WIID) can be mentioned.
-Both datasets do not cover more current figures for Switzerland.
+For the sake of completeness the database construtcted by \citet{Deininger and Squire 2006 FEHLT} and the World Income Inequality Database (WIID) have to be mentioned.
+Both datasets do not contain figures  for recent years for Switzerland.
 A further important database on inequality is the GINI database which has been derived from the GINI Country Reports \citep{nolan_changing_2014}.

### 2014-06-13 paper/paper.Rnw
Several Updates to introduction and method section
added Plot with Gini Series from other other studies
citation style to unsrtnat
fixed double quotes and R-paths so paper.Rnw compiles
Gini series out of HABE
added natbib for citations

@@
 \usepackage{color} % used to mark parts that need to be edited
-
+\usepackage{natbib} %to use citealt citet citep
 \usepackage{abstract} % Allows abstract customization
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 

### 2014-06-13 paper/paper.tex
added natbib for citations

@@
 Comparison before and since 2008 is not recommended (OECD 2012:315).
-A further important database on inequality is the GINI database which has been derived from the GINI Country Reports (Nolan et al.,2014).
+For the sake of completeness the database construtcted by \citet{Deininger and Squire 2006 FEHLT} and the World Income Inequality Database (WIID) have to be mentioned.
+Both datasets do not contain figures  for recent years for Switzerland.
+A further important database on inequality is the GINI database which has been derived from the GINI Country Reports \citep{nolan_changing_2014}.
 But this dataset doesn't cover Switzerland.}  \\
@@
 However, it is unclear how the bias through non-response affects the overall measure of inequality.
-Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) and a time series for inequality of wealth is missing.
+For the US \citet{atkinson_top_2009} estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by \citet{Alvaredo (2010) FEHLT}.
+Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) is missing.
 Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.
 First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
-Second, we summarize and apply tax data specific technics to construct suitable measures of inequality.
+Second, we summarize and apply tax data specific techniques to construct suitable measures of inequality.
 We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
-Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through datasource.
+Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through data source.
 
@@
 Based on a review on the literature about the measurement concepts in an ideal world, we discuss the advantages and shortcomings of tax data compared to other data sources - namely survey data.
-Along with the despriction we explain the methods and thechnics we use to construct time-series of inequality-measure for income for Switzerland, which we will present in the result sections.
-By doing this, we demonstrate the suitability of the FTA Tax Data to assess the development of inequality in Switzerland
+We describe and explain methods and techniques needed to construct time-series of income inequality measures for Switzerland.
+We will present these time-series in the result section and evaluate the use of the FTA Tax Data to assess the development of inequality in Switzerland
 
-\subsection{Tax statistics in Swizterland}
+\subsection{Tax statistics in Switzerland}
+
+Our data comes from the Swiss Federal Tax Administration (FTA).
+Federal taxes are collected and documented by the FTA since 1915.
+Being called a war-tax in the beginning, the federal tax was renamed to crisis levy in 1934, defense-tax in 1939 and is finally known as direct federal tax since 1983.
+The time frame we were able to collect ranges from 1945 to 2010 including 44 tax periodes.
+\footnote{Between 1993 and 2003 there is no exact data available for Switzerland because of a system change from taxation assessed in arrears (Praenumerando-System) to taxation assessed on current year income (Postnumerando-System) which was implemented by cantons in different years.}  While the FTA provides data in electronic form since 1973 we collected earlier data by scanning hard copies.
+Data is available for Switzerland plus all cantons and basically covers every tax unit (individual or household) in Switzerland liable to pay federal taxes.
+This exempts all tax units with taxable income below a certain threshold (e.g.
+15.000 CHF in 2010).
+Furthermore the FTA differentiates between two groups of tax units, so called normal cases and special cases.
+A normal case is a tax unit residing in a swiss canton without foreign source income and being liable to taxation all year long.
+All other tax units and very few that are taxed based on the style of living because they don't work (Pauschalbesteuerte) are special cases.
+\\
+
+The FTA provides two income measures: taxable income and net income.
+Net income here is an administrative term and means taxable income plus social deductions (children and supported persons) but not including other deductions like donations or health-care costs.
+Both measures are designed for taxation purposes which might limit the suitability to measure inequality as we will discuss later.
+
+% Bei den Nullern erw?hnen, ab wann Angaben zur Zahl der Nuller vorliegen und ab wann nicht
+% Im Internet hat es Publikationen ab 1947/48
+% Es gibt auch Tabellen auf Gemeindeebene
+% Die Statitstischen Kennzahlen von Br?lhart erw?hnen?
+% Erw?hnen in welcher Form die Informationen publiziert sind. Tabulation by size of income. Not on individual or household level
+
+% http://www.estv.admin.ch/dokumentation/00075/00076/00701/index.html?lang=de
+% Kantone und Gemeinden: 2010 bis 1947/1948
+% Klassierte Tabellen für Abzüge, Einkommen (Steuerbares Einkommen und Reineinkommen) und Steuererträge
+% Normalfälle und Sonderfälle
+
+% Statistsiche Kennzahlen
+% Mit Belastung durch die direkte Bundessteuer: 1973/1974 bis 2010
+% Mit und ohne Belastung durch die direkte Bundessteuer: 1995/1996 bis 2010
+% steuerbares Einkommen/Reineinkommen
+% Jeweils Normal und Sonderfälle
+% Individual-Daten (nicht klassierte daten)
+% äquivalenzeinkommen
 
@@
 However, recent activities emphasize the need of a broader conceptualization.
-A recent publication from the oecd (2013) condense these ideas into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework on micro-level statistics.\footnote{Harmonisation with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framewrok presented in this publication.
-Considered main stadards were the System of National Accounts (SNA 2008), the canberra Group Handbook on Household Income Statistics (United Nations 2011), the final report of the 17th International Conference of Labour Statisticians (ILO 2004) and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing (UNECE & Eurostat 2006)} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
-To gain policy relevant insight, it is recomended to look at the distribution of all three distributions simultaneously.
-Some households with low income, for example, may report adequate levels of consumption expenditure or welth holdings, or vice-versa.
-But it is also stated (OED 2013:18):"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by colecting and provding a database following ths broader concept of economic well-being.
+A recent publication from the \citet{oecd_oecd_2013 UNSICHER} condense these ideas into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framework presented in this publication.
+Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_????
+UNSICHER}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
+To gain policy relevant insight, it is recommended to look at the distribution of all three distributions simultaneously.
+Some households with low income, for example, may report adequate levels of consumption expenditure or wealth holdings, or vice-versa.
+But it is also stated \citep[18]{oecd_oecd_2013 UNSICHER}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being.
 http://www.lisdatacenter.org/our-data/lws-database/} \\
 
-This last statment holds for Swizterland too, although the HBS study is strongly influced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
-Albeit the awarness of an assessment of income, consumption and wealth simuntanously is rising, we  focus our analysis on income, which is undoubtly a crucial indicator of economic wellbeeing.
-However, the Federal Tax Office publishes statistics on income and wealth .
-But it is not possible to analyse the jointly distribution on the individual or household level.
+This last statement holds for Switzerland too, although the HBS study is strongly influenced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
+Albeit the awareness of an assessment of income, consumption and wealth simultaneously is rising, we  focus our analysis on income, which is undoubtedly a crucial indicator of economic wellbeing.
+It should be noted, that the Federal Tax Office publishes statistics on income and wealth.
+But it is not possible to analyze the jointly distribution on the individual or household level.
 Also measures of consumption are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow.
@@
 The assessment of income inequality is influenced by the definition of the income itself.
-Marketincome or disposable income for example differ by substantial meaning and by the expected degree of inequality.
-Therefore the awarness of the analyzed concept is crucial.
-Terminology can sligthly differ, while common concepts can be identified (for detailed discussion see: OECD (2013:44), Canberra group (2011:24)).
-Figure xy shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution.
-Within this framework common income definitions are situated.
+Market income or disposable income for example differ by substantial meaning and by the expected degree of inequality.
+Therefore the awareness of the analyzed concept is crucial.
+Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013 UNSICHER}, \citet[24]{united_nations_canberra_2011}).
+Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually leads to the disposable income.
+The income measure, which finaly shapes the possibility to consume.
+Within this framework common other income definitions are situated.
 These definitions are contrasted with income concepts, which can be derived out of the FTA Tax data.
 
+\begin{figure}
+\centering
 \includegraphics[width=\maxwidth]{figure/income_definition}
+\caption{Income definitions}
+\label{fig:incdef}
+\end{figure}
 
@@
 
-The central income according to tax data is the taxable income.
-It includes all reported incomes (income from employment, income from property and received transfers\footnote{Mean-tested benefits are not taxed and therefore not included in Tax data.
-Income for low income groups are therefore underestimated.}) minus several deductions.
+% ACHTUNG: Wir m?ssen innerhalb des Papers einheitlich bleiben. In der Grafik steht jetzt "Total income" oben steht "Net income"
+
+% Man k?nnte es auch konzeptionell auf kantonale Steuerdaten erweitern. Besser w?re es, dies einfach im Ausblick zu erw?hnen
+
+The central income reported through tax statistics is the taxable income.
+It includes all reported incomes (income from employment, income from property and received transfers\footnote{Realized capital gains are not taxed and Mean-tested benefits are not taxed and therefore not included in Tax data.
+Income for low income groups are therefore underestimated.
+However, Piketty and Saez (2001) note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had nonly a trivial impact on top income shares.}) minus several deductions.
 It is therefore neither a pre-transfer income  nor a post-transfer income measure.
-It's rather something inbetween.
-As the FTA tax statistitcs include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions.
-This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction,Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income".
-As some deducations can be interpreted as compulsary expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption.
-Similiar when calculating the disposable income out of the taxable income through accouting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
-Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches are missing.
+It's rather something in between.
+As the FTA tax statistics include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions.
+This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction, Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income".
+As some deductions can be interpreted as compulsory expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption.
+Similar when calculating the disposable income out of the taxable income through accounting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
+Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches, which represent the bulk of taxes in Switzerland, are missing.
 \\
@@
 
-If the goal is to catch the income, which shapes the possibility to consume, equivalence scale are implemented.
-Equivalence scale followe the idea, that persons living in the same household have the possibility to share costs.
-To compare the individual economic well-beeing equivalence scales adjust for different household sizes (see OECD 2013,173, Buhmann et al., 1998).
-Because it is neither possible to calculate the total household income nor the respective householdsize out of the FTA tax data, we put this in this paper outside the brackets.
-Although knowing, that with doing so we neglect a potential aspect, which is capable of shaping the development of inequality.
+\emph{statistical units} \\
 
-% letzter Abschnitt muss noch auf die Brülhart-Daten hinweisen. Damit können wir eine äquivalenzeinkommen betrachten.
-% Auch müsste in diesem Abschnitt auf die Probleme der Haushaltsunterscheidung eingegangen werden (oder ist im Abschnitt Population coverage? Dann müsste obiger Punkt auch da erwähnt werden.)
-
-% Streichen
-%=======
-% 
-%>>>>>>> d21c323... completed subsection "concepts on measuring economic resources" and "defining income"
-
-\emph{Population Coverage} \\
-All Residents? Working Residents? Single Person vs household/family. Households without income \\
-see for details: OCED Framework for Statistics on the Distribution of Houshold income, consumption and wealth (2013) \\
-
-The income of the household generally plays the major role in the living standards attained by its members, so the way earners and non-earners grouped together in households, has to be incorporated into the analysis.
-Especially when the the focus of interesst lays in unrevealing the role of demographic factors.
-
-\emph{Measuring inequality and concentration} \\
-To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
-Vgl.
-Allison (1978), Engelhardt (2000), Cowell (2000) oder Hao/Naiman (2010)
-%[Zusammenfassung bestehender Überlegung und Hinleitung zu den Masszahlen, die wir verwenden werden (vermutlich Gini und relative distribution methods) \\
-
-neuere Indikatoren; Polaritätsindex: relative distribution methods in the Social Sciences (Handcock/Morris 1999) \\
-
-Gini is mainly used for international comparison.
-However, other measures are often reported along (gini is more sensitive to changes in the middle of the distribution than to changes in the tails (OECD, 2008:37).
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading) (Alderson and Doran 2013).
+The agreed standard on the statistical unit, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013 UNSICHER}.
+Indeed it are the individuals, who receive income, own wealth and experience economic wellbeing, but their possibility to do so, is strongly tied to the concept of household.
+This comprises all persons under the same housing arrangement.
+The basic underlying assumption for collection data on households level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children).
+Additionally, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing.
+The compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013 UNSICHER}, \citealt{Buhmann et al., 1998 FEHLT}).
 \\
 
-Alvaredo (2010) shows formally, how top incomes shares are related to the gini-coefficient.
-Following his argument, it is crucial to consider top incomes.
-Leigh (2007) also analyzed the relationship between the Gini (and further inequality measures) and top income shares in a panel of 13 countries (Switzerland is part of the studie).
-%[He also discusses tax data studies in general-> have a closer look] \\
-%[We definitely have to argue, why it isnt sufficient to report top income shares, because if it is, there is no need to have another study about Switzerland (national level>Martinez, cantonal level> gorgas and Schaltenegger)]
+In tax data, however, the units are represented according to administrative rules.
+Tax units therefore neither represent individuals in every case nor true households.
+Tax units rather represent individuals and couples, but couples, who are married or officially registered.
+This doesn't imply, that those couples live together, as it is needed to satisfy the definition of a household.
+On the other hand, is it quit likely that more than one tax units live in the same household (unmarried/unregistered couples, see \citet[99]{muller_vermogenslage_2014}).
+It is therefore not directly possible to elicit households and household income out of tax data.
+This might influence the assessment inequality, taking into account the change from traditional household and family structures over the last century.
+This is an inequality related issue, where relevant studies are missing.
+
+% Allenfalls die Br?lhart-Daten erw?hnen und die m?glichkeit der Haushaltsbildung ?ber Abz?ge. Falls wir auch eine Grafik zur Rolle der Haushaltszusammensetzung machen (ich denke schon)
+% gibt es studien? Zur Bedeutung von ?quivalenzskalen und zur Bedeutung der Differenzierung von Haushalten und Individuuen?
+
+\emph{Measuring inequality or concentration} \\
+To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
+The Gini coefficient is the most known measure and mainly used for international comparison.
+As it is derived from the lorenz-curve, the quantified amount of inequality can unpretentiously be described in a formal and visual way.
+Therefore the Gini coefficient is easely interpretable.
+Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}.
+(1) principle of population: the assessment of inequality is independent of the population size (2) Bresciani-Turroni: the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
+doubling of all income) (3) weak principle of transfers (pigou-dalton): transfers from richer households lead to a reduction of inequality.
+However, several drawbacks are reported in the literatur.
+The most important point is, that the underlying distributional form of the measured inequality is unknown and it is therefore not possibel to see if the measure is driven be a few rich or many poor individuals.
+This can also be problematic for comparision between countries or over time.
+In extreme cases two totally different distrubtions share the same gini-coefficient.\\
+
+The recent wave of tax-data studies do not report Gini-coefficents.
+Rather top income shares are informed on, which are calculated not only with taxdata, but togehter with external sources to produce the population and personal income control totals.
+This procedure ensures, that the inequality measure is not biased because of non-fillers, who doesn't appear in tax statistics.
+Leight (2002:594f) compares top income shares with other inequality measures and ask, if they are a useful measure of inequality in a society.
+He tries to answer this question empiricaly by comparing measures of inequality based on top income shares with measures of household or familiy inequality.
+He finds a strong positive relationship, but concludes (P.600): top income shares are far from perfect as a measure of distribution of income across society.
+Top income shares hence inform not completely on how inequality evolves elsewhere in the distribution.
+Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent see above).
+A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between invidviduals, who are either both within the top group or both outsie the top group, then the share measure will remain unchanged.
+\\
+
+Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading).
+The polarityindex is developed by \citep{Handock and Morris 1999} allows for this.
+Recenently this index was applied in the work of \citep{Alderson, Beckfield and Nielsen 2005 FEHLT} and \citep{Alderson and Doran 2013 FEHLT}.
+This approach is rooted in relative distribution methods.
+It includes a precise comparison of the shape of two distributions (groups, over time).
+The main advantage lies herin, that this approche allows to characterize the change, in the sense that it is possible to say, if a polarization occured (reduction of middle class), which equals an increase of inequality and if this change is driven by a change in the upper or lower part of the distribution.
+\\
+
+The literature mentions several other metrics with desired properties, but we will not discuss them further (see for example \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}).
+
+% Allenfalls auch auf Quintilvergleiche eingehen und Atkinsonmass
+
+\emph{Population Coverage}
+% beim letzten Punkt geht es um Steuerhinterziehung. Fällt das auch noch unter Population coverage?
+
+Often inequality is assessed on national level, which implies, that studies try to cover the whole population of the country of interest.
+This is a special thorny task for surveys working with samples, because nonresponse is a major source of bias \citep{bethlehem_handbook_2011}.
+\citep{korinek_survey_2006} show, that the position in the income distribution influence the probability to participate in a survey.
+Low income and high income households are more likely to refuse survey response, which leads to an overrepresentation of middle income households.
+This process can be referred to as the "middleclass bias" \citep{diekmann_empirische_2009}.
+Missing data in household surveys is therefore not missing at random, which has an impact on the measures of inequality.
+The magnitude of this bias in Switzerland, however, is unknown.
+Strategies to handle this kind of bias are discussed in the literature \citep{S?rndal et al.
+1992 FEHLT}, but require a register for every unit, that is proportional to income.
+Currently there exists no such register for Switzerland \citep{ecoplan 2014:43 FEHLT}.
+Currently used micro datasets, which are used for official publications concerning inequality in Switzerland (SILC and HABE) are furthermore confronted with a constructed coverage problem, because these surveys rely on the phone register, which excludes the households not having a registered connection.
+\\
+%(Problems of collective households, ecoplan 2014:45). Erw?hnen?
+
+The issue of incomplete coverage is less dramatical with tax data.
+Essentially every permanent resident in Switzerland, who has completed the age of 18 years (respectively 20 years prior to 1996) is taxed on a yearly base (or every two years before the change of the tax system).
+Essentially this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distribution.
+This includes a separation of normal cases, which embrace the majority of taxpayers, and the special cases, which cover Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
+Most important this includes high net wealth individuals taxed according to their expenditures.
+Special attention has to be paid to tax units with none or very low incomes.
+Even though they have to hand in a tax return, their income does not show up in the statistics if their income after deductions falls below 15'000 CHF and they are therefore not taxed with direct federal taxes.
+This is possible for normal and special cases alike.
+From 1995/1996 until 2010 the number of non-taxed units are reported, but not for the years before.
+Dell et al.
+(2007) try to estimate the fraction of non-taxed by comparing the reported numbers of tax units to census reports about the number of adult population.
+Following his etimations this fraction drops from 94\% in 1993/1994 to 63\% back in 1945/46.
+\\
+
+An other critical issue is the problem of tax evasion, which definitely can bias the assessment of inequality.
+Alvaredo and Saez (2006) for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
+Evasion can occured, when individuals try not to fill tax returns or by misreporting of incomes.
+In Switzerland non-fillers show up in the tax-statistics either way, as long as they are registered.
+This person get an imputed income form an older tax return and infomration given by employers.
+Only non registered non-fillers are not in the records.
+Therefore non-fillers are a minor problem.
+Not neglectable is the circumstance, that individuals misreport incomes.
+\citet{feld_tax_2006} examine the role of tax evasion in Switzerland by calculating the difference of the national accounts measures of primary income and the income reported to the tax authorities.
+Building on a time series of Pommerehne and Weck-Hannemann (1996) Feld and Frey (2002) and Frey and Feld (2002) extend the series.
+They can show, that the average level of income tax evasion from 1965 to 1995 varies between 13\% and 35\%.
+They suggest, that evasion is heavily driven by capital income tax evasion.
+\\
+
+\emph{Intertemporal comparison}
+
+One advantage of tax data over survey data is their observation period.
+While the FTA data we use range back to 1945, the earliest period of survey data is 2000 (HABE data).
+To be fair, survey data might cover the late periods (2011 and 2012) which are not covered by tax data.
+The striking feature of the long FTA time series is its consistency.
+Both population coverage and income measures are consistent for the complete observation period.
+This can not be said for survey data as these are based on samples and therefore require an ideal sampling design or reweighting to be representative and comparable over time.
+We will see how well sampling and weighting is done in the results section.
 
@@
 %- Implementation of common concepts of economic ressources
-%- contingency for intertemporal comparision
+%- intertemporal comparision
+%- flexibel applications of appropriate inequality measures
+
+\begin{table}[h]
+\begin{tabular}{lcccc}
+                                   & FTA       & HABE      & EU-SILC   & LIS       \\
+Range                              & 1945-2010 & 2000-2011 & 2007-2012 & 1982-2004 \\
+Cross country comparison possible? & no        & no        & yes       & yes       \\
+Top incomes covered?               & yes       & no        & no        & no        \\
+Low incomes covered?               & no        & yes       & yes       & yes       \\
+Concepts of economic ressources?   & no        & yes       & ?         & ?         \\
+Equivalisation?                    & no        & yes       & ?         &          
+\end{tabular}
+\end{table}
+
+% Braucht es alle drei Survey-Befragungen
+% ich würde nicht dichotomisieren, sondern ordinal bewerten (tief, mittel, hoch)
+
+% Die Punkte kurz beschreiben.
+% Zusammenfassung
+
+% Gedankenstütze
+%\textbf{Problems with household income surveys}
+%\begin{itemize}
+%\item Sample data (bias)
+%\item comparability between countries and over time (depends on income definition)
+%\item short time series
+%\end{itemize}
 
 %Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
+%Studies on problems with survey data See for other countries: \citealt{Siminski et al 2003 FEHLT} (Australia), \citealt{Brewer et al 2008 FEHLT}, UK, \citealt{Burkhauser et al. 2009 FEHLT}, US)
 
@@
 
-%Copied from:
-%Alvaredo, Facundo, Anthony B. Atkinson, Thomas Piketty and Emmanuel Saez, The World Top Incomes Database, http://topincomes.g-mond.parisschoolofeconomics.eu/ , 18/12/2013
+\subsection{Outline of applied methods}
 
-\textbf{Summary of tax data drawbacks}
-\begin{itemize}
-\item Misreporting of incomes (high earners have an interest in getting tax beneficial ``income'' see above)
-\item with aggregated data the adjustment for household members is not possible (no equivalization possible).
-This is a point, which is hardly mentioned in the ``top-income'' literature, albeit it is a crucial issue in ``theorized'' concepts of income.
-To generalize one can say, that the concept of tax units (individuals and couples + no direct information about household members at least not in the aggregated tax tables) is not congruent with the concept of household.
-This might influence the overall inequality, taking into account the change from traditional household and family structures over the last century.
-This is an inequality related issue, where relevant studies are missing.
-What are the assumptions?
-\item the tax data income definition summarize labor income, capital income and taxable transfer payments (no means-tested benefits as welfare aid).
-Distinction between the different income sources is not possible, albeit the mechanisms of changes in the specific distribution can be very different.
-\item Some tax units are not included in the tax tables (see Foellmi and Martinez 2013:11f).(a) only taxed cases are included (partly a problem, for certain periodes we know how many cases are excluded because of low or none income)	- (b) cases tax at source (Quellenbesteuerung).
-Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
-In some border-cantons the share of this group is very high (20\%)
-%Can we ignore this)
-(c) staff of international organizations based in Switzerland	- (d) non-fillers show up in the tables as long as they are registred.
-This persons get an imputed income (older tax tax return and information given by employers.
-Not registered non-	fillers are not in the records.
-- (e) tax evasion.
-Feld and Frey (2007) report about tax evasion in Switzerland, it should be somewhat above 20 percent on average.
-With the (strong) assumption that the pattern of tax evasion over time is stable, this is a minor problem for inequaly measures over time.
-\end{itemize}
+% Diese Abschnitt muss ersetzt werden mit der von uns gewählten Vorgehensweise
 
-\textbf{Problems with household income surveys}
-\begin{itemize}
-\item Sample data (bias)
-\item comparability between countries and over time (depends on income definition)
-\item short time series
-\end{itemize}
+In the last section we described the advantages and drawback of taxdata discussing five aspects, which we regard as crucial concering the assessment of inequality.
+To get a feeling of the importance of these aspects, we exploit the FTA-tax data as far as possible and perform several insightful calculations adressing four of the five mentioned aspects.
+No further investigation is possible regarding aspect (1) concepts on measuring economic ressource.
+This is a point which cannot be further addresses with FTA tax data.
+But for the other four thematic areas, we can provide deeper insight.
+In general our main strategy is to apply different possible concepts within one conceptional area (income, units, measurement, population coverage) while holding other conceptual differences constant.
+With this strategy we want to show, the sensitiveness the assessement of inequality is especially if one looks at time trens.
+In this section we  describe which methods we use to produce the results in the next section.
 
-Atkinson et al. (2009) estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by Alvaredo (2010). \\
+%%%
+% Was soll hier beschrieben werden und was direkt bei den Tabellen?
+%%% Gar nicht so einfach....
 
-See for other countries: Siminski et al 2003 (Australia), Brewer et al 2008, UK, Burkhauser et al. 2009, US)
-%		What about the special cases?
-Foelmi and Martinez argue, that it is crucial to include the special cases because they include the high net wealth individuals taxed according to their expenditures.
-Exclusion leads to an underestimation of inequality
-%		Capital gains? Should not be included 
-The discussion about problems with reporting income is fairly exhaustive. What about wealth?
+%%% Nach Lehrbuch müssten die methodischen Schritte hier beschrieben werden.
 
-It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
-Credit Suisse (2011) makes use of the information in the ``Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
-\\
+% Defining income
+% Beschreiben, dass sich aus den publizierten Einkommensgrössen und den Steuerbeträgen unterschiedliche 
+% Einkommen berechnen lassen
+% Es steht der Vorwurf im Raum, die unteren Einkommen seien falsch abgebildet, weil die Sozialhilfe nicht versteuert ist
 
-Additonaly the Credit Suisse Reports states (2011:8) that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts.
+%%
+% statistical units
+% Beschreiben, dass zwar keine Haushaltseinkommen gebildet werden können und Skaleneffekte nicht abgebildet sind, dass eine Annäherung jedoch möglich ist
+% Definition der äquivalenzskala bei Brülhart (korrespondiert nicht mit OECD Definitionen, aus pragmatischen Gründen)
 
-\subsection{Tax data published by the Swiss Federal Tax Administration}
+%%
+% Measuring inequality
+% Beschreiben, wie sich aus gruppierten Steuerstatistiken der Gini-Koeffizient berechnen lässt
+% Beschreiben wie wir relative distribution technics einsetzen und den Polaritätsindex berechnen
 
-%Beschreiben was die Daten beinhalten und wo es Probleme gibt
-\begin{itemize}
-\item	Datengrundlage: 1947/48 bis 1981/1982 Eidg. Wehrsteuer. 1983/1984 bis 2010 direkte Bundessteuer. Zugänglich über estv.admin.ch
-	\item	Special feature Switch from bi-annual taxation to standard annual tax-system (1995/1996 bis 2003)
-	\item	Tabulation by size of income and statistical measures from individual tax records (Brülhart-Daten)
-	\item	Income after deductions-> taxable income (employment income, business income and capital income).
-Is income definition stable over time?
-Yes it should.
-Realized capital gains are excluded from the definition.
-It includes income from  employment, self-employment, capital income and taxable transfer payments.
-Plus Eigenmietwert,
-	\item	Reported on national and cantonal level
-\end{itemize}
+%%
+% Population coverage
+% Analysen zu den Nullern aufzeigen> ist eine Hauptkritik an der Nutzung von Steuerdaten
+% Die Bedeutung von Normal- und Sonderfällen
+% Den Vergleich mit der HABE beschreiben, welche Einkommensgrösse nutzen wir aus der HABE um diese möglichst mit den Steuerdaten vergleichbar zu machen?
+Wir brauchen das Bruttoeinkommen nach Abzug der Sozialversicherungsbeiträge, dies lässt sich mit dem Reineinkommen vergleichen
 
-\subsection{Ways to tackle FTA-tax data specific problems}
+%%
+% Intertemporal comparison
+% Ziel eine möglichst lange Zeitreihe abzubilden. 
+% Technik der Lückenfüllung von Ben beschreiben> dies kurz anderen Techniken gegenüberstellen.
+% Konfidenzintervalle für die Perioden mit vielen fehlenden Werte?
+
+% eine / oder zwei Zeitreihen (eine lange und eine, die möglichst nahe an der richtigen Konzeptionalisierung ist: post direct taxed income, normal+sonderfälle) zusammen mit den bereits existierenden Reihen (HABE,LIS,SILC) ploten
 
 \textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
-Del et al.
-(2007) impute for non-fillers the 20 percentage of the annual average income.
+\citet{dell_income_2007} impute for non-fillers the 20 percentage of the annual average income.
 This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
-Furthermore Del et al.
-calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
+Furthermore the authors calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
 \\
@@
 %is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
+% Beim Imputieren m?sste man auch die Zahl der freiwilligen Nichtbeziehenden ber?cksichtigen.
 
-\textbf{deductions} Del et al.
-(2007:477):" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
-Gorgas and Schaltenegger (2011:5): ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
+\textbf{deductions} \citet[477]{dell_income_2007}:" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
+\citet[5]{schaltegger_evolution_2011 UNSICHER}: ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
 Können wir das auch?
@@
 In our case we can clearly reject the hypothesis that all interactions are zero ($p=0$).
-This leads to the conclusion  that gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted gini coefficients.
+This leads to the conclusion  that Gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted Gini coefficients.
 For example one might be interested in how inequality would had developed if the zero-share would have been constant over time.
 (Note RF: predict all data point using canton, time and the initial OR final zero-share to homogenize the time series)
-Furthermore we can quantify how large the bias is and we can do this seperately for tax periods or seperately for cantons.
+Furthermore we can quantify how large the bias is and we can do this separately for tax periods or separately for cantons.
 
@@
 Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
+
+%The discussion about problems with reporting income is fairly exhaustive. What about wealth?
+
+%It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
+\citet{credit_suisse_global_2011} makes use of the information in the ``Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
+\\
+
+%Additionally the Credit Suisse Reports states that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts \citep[8]{credit_suisse_global_2011}.
 
@@
 
+% M
+
 %%%-------------------------------------------------%%%

### 2014-06-13 paper/subdocuments/data_and_methods.Rnw
minor changes to data and methods and basic structure of results section

@@
 The central income reported through tax statistics is the taxable income.
-It includes all reported incomes (income from employment, income from property and received transfers\footnote{Realized capital gains are not taxed and Mean-tested benefits are not taxed and therefore not included in Tax data.
+It includes all reported incomes (income from employment, income from property and received transfers\footnote{ Mean-tested benefits are not taxed and therefore not included in Tax data.
 Income for low income groups are therefore underestimated.
-However, Piketty and Saez (2001) note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had nonly a trivial impact on top income shares.}) minus several deductions.
+However, Piketty and Saez (2001) note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions.
 It is therefore neither a pre-transfer income  nor a post-transfer income measure.
@@
 % Auf den ESTV-Tabellen sind jedoch noch mehr Abzüge abgebildet. Was soll/muss/können wir nun abziehen? Bin verwirrt...
+
+% Allenfalls darauf hinweisen, dass die Wahl der Einkommensdefinition von der Forschungsfrage abhängt.
+Inequality in market outcome (genderdifferences) requieres to look at market earings from employment, to assess the role of redistribution different steps of income has to be looked at.
+Following the idea of economic wellbeing usual disposable income is looked at
 
@@
 This might influence the assessment inequality, taking into account the change from traditional household and family structures over the last century.
-This is an inequality related issue, where relevant studies are missing.
 
@@
 % gibt es studien? Zur Bedeutung von ?quivalenzskalen und zur Bedeutung der Differenzierung von Haushalten und Individuuen?
+% inequality and family structure> review ny (McCall & Percheski (2010))
+% Was ist zu erwarten, wenn Äquivalenzskalen vernachlässigt werden? Soll man dies erwähnen oder einfach weglassen
 
@@
 This procedure ensures, that the inequality measure is not biased because of non-fillers, who doesn't appear in tax statistics.
-Leight (2002:594f) compares top income shares with other inequality measures and ask, if they are a useful measure of inequality in a society.
+Leight (2002:594f) compares top income shares with other inequality measures and asks, if they are a useful measure of inequality in a society.
 He tries to answer this question empiricaly by comparing measures of inequality based on top income shares with measures of household or familiy inequality.
@@
 
-An other critical issue is the problem of tax evasion, which definitely can bias the assessment of inequality.
+An other critical issue with tax data is the problem of tax evasion, which definitely can bias the assessment of inequality.
 Alvaredo and Saez (2006) for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
@@
 \citet{feld_tax_2006} examine the role of tax evasion in Switzerland by calculating the difference of the national accounts measures of primary income and the income reported to the tax authorities.
-Building on a time series of Pommerehne and Weck-Hannemann (1996) Feld and Frey (2002) and Frey and Feld (2002) extend the series.
 They can show, that the average level of income tax evasion from 1965 to 1995 varies between 13\% and 35\%.
@@
 %- flexibel applications of appropriate inequality measures
+% - appropriate income definition
 
@@
 % Population coverage
-% Analysen zu den Nullern aufzeigen> ist eine Hauptkritik an der Nutzung von Steuerdaten
+% Analysen zu den Nullern > ist eine Hauptkritik an der Nutzung von Steuerdaten
 % Die Bedeutung von Normal- und Sonderfällen

### 2014-06-13 paper/subdocuments/results.Rnw
minor changes to data and methods and basic structure of results section

@@
 
-\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
-Kantonsweise Grafiken
-Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
+% Brauchen wir glaubs nicht mehr
+%\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
+%Kantonsweise Grafiken
+%Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
 
-\subsection{Gini coefficients for taxable income, income before deductions and taxable income after tax}
+% Der Übersicht halber habe ich hier die Titel hier analog zum "Struktur-Issue" auf git gesetzt
+
+\subsection{Defining income}
+
+\emph{Gini coefficients for taxable income, income before deductions and taxable income after tax}
 
@@
 
-\subsection{Zero share}
+\emph{Gini coefficients considering mean-tested benefits}
 
-Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later)
+\subsection{Measuring inequality}
 
-\subsection{Income measures with and without zeros}
+% Zero-shares (alter Textbaustein)
+%Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later) 
+% Das allenfalls in der Methodensection machen?
+% Vergleich mit der Dell-Reihe zu Non-fillers
+% Überschneidungen gib es leider nicht 
+% Interessant wäre es auch, zu sehen ob es Brüche in den Zeitreihen der Zahl der Steuerpflichtigen und jener ohne Besteuerung gibt.
+
+\subsection{Population coverage}
+
+\emph{Income measures with and without zeros}
 
@@
 
-%Switzerland (income and wealth)
-%- graph with gini over time> identify periods of change
-%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
-%Cantonal level (income and wealth)
-%- different development on regional level 
-%Switzerland in international comparison
-%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality
+\emph{normal and special cases}
 
-%- Comparison with other countries (similar data)
+\emph{comparison of tax-data and survey data distribution}
 
-% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen.
-Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel.
-Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.
-
-%rf: genau. kantone unterstreichen die probleme mit dem 0er bias und die probleme mit "imputing the gap". mehr muss erstmal nicht sein.
+\subsection{Interremporal comparison}

### 2014-06-15 paper/paper.tex
normal and special cases

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 
@@
 
+% Allenfalls darauf hinweisen, dass die Wahl der Einkommensdefinition von der Forschungsfrage abhängt.
+Inequality in market outcome (genderdifferences) requieres to look at market earings from employment, to assess the role of redistribution different steps of income has to be looked at.
+Following the idea of economic wellbeing usual disposable income is looked at
+
 \emph{statistical units} \\
@@
 %%% Nach Lehrbuch müssten die methodischen Schritte hier beschrieben werden.
+% Man könnte zu den Zeitreihen jeweils die Varianz ausweisen-> dass würde helfen die Reihe zu beurteilen.
+Man sieht dann: aha, die Reihe ist im Vergleich zur anderen höher/tiefer.
+Ob die Reihen stärkeren Schwankungen unterliegen könnte einfach aus der Varianz gelesen werden
 
@@
 % Einkommen berechnen lassen
-% Es steht der Vorwurf im Raum, die unteren Einkommen seien falsch abgebildet, weil die Sozialhilfe nicht versteuert ist
+% Es steht der Vorwurf im Raum, die unteren Einkommen seien mit Steuerstatistiken schlecht abgebildet, weil bedarfsabhängige Leistungen nicht versteuert sind
+
+% Es müsste eine Reihe zum gesetzliche Grundbedarf vorliegen. Wüsste ich jetzt grad nicht von wo...
 
@@
 
+% Calculate the gini
+% Ein bisschen erklären, wie aus gruppierten Tabellen der Gini-Koeffizient berechnet wird.
+
+% Polaritätsindex
+% Alderson and Doran (2013:56) beschreiben zwei mögliche Indizies Median relative polarization index (MRP) eingeführt von (Morris, Bernhardt und Handcock 1994:217), der in den lower relative polarization index(LRP) und den upper relative polarization index (URP) zerlegt werden kann (ebd.209)
+% (1. Schritt) Probability density functions
+% (2.
+Schritt) Overall relative density function -> Achtung diese Dichtefunktion bildet sowohl die  Lokationsverschiebung als auch durch Verschiebung der Form ab.
+Eigentlich interessiert aber, der shape shift
+% Deshalb wird die Verteilung um die Lokationsverschiebung bereinigt und nur der shape-shift angeschaut.
+Ausgewiesen werden (MRP)-> hat polarisierung stattgefunden: LRP und URP-> ist polarisierung eher auf eine Zunahme armer (LRP) zurückzuführen oder auf eine Zunahme von Reichen?
+(URP)
+
 %%
 % Population coverage
-% Analysen zu den Nullern aufzeigen> ist eine Hauptkritik an der Nutzung von Steuerdaten
+% Analysen zu den Nullern > ist eine Hauptkritik an der Nutzung von Steuerdaten
 % Die Bedeutung von Normal- und Sonderfällen
@@
 
-\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
-Kantonsweise Grafiken
-Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
-
-\subsection{Gini coefficients for taxable income, income before deductions and taxable income after tax}
+% Brauchen wir glaubs nicht mehr
+%\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
+%Kantonsweise Grafiken
+%Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
+
+% Der Übersicht halber habe ich hier die Titel hier analog zum "Struktur-Issue" auf git gesetzt
+
+\subsection{Defining income}
+
+\emph{Gini coefficients for taxable income, income before deductions and taxable income after tax}
 
@@
 
-\subsection{Zero share}
-
-Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later)
-
-\subsection{Income measures with and without zeros}
+\emph{Gini coefficients considering mean-tested benefits}
+
+\subsection{Measuring inequality}
+
+% Zero-shares (alter Textbaustein)
+%Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later) 
+% Das allenfalls in der Methodensection machen?
+% Vergleich mit der Dell-Reihe zu Non-fillers
+% Überschneidungen gib es leider nicht 
+% Interessant wäre es auch, zu sehen ob es Brüche in den Zeitreihen der Zahl der Steuerpflichtigen und jener ohne Besteuerung gibt.
+
+\subsection{Population coverage}
+
+\emph{Normal versus special cases}
+
+The FTA stopped to publicly report data for special cases after tax period 1993/94. 
+For more recent periods however we can compare data from the 
+BRUELHARTPROJEKT (wie nennen wir das?) to the FTA normal cases, as these data are 
+identical but the BRÜLHART data include special cases. 
+We will have a look at both, the 1993/94 period as the last period where both 
+numbers were reported as well as 2009 where we compare data from the same source 
+but using differnt data sheets.
+
+%ABBILDUNGEN ZUM VERGLEICH NORMAL-/SONDERFÄLLE
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
+
+\includegraphics[width=\maxwidth]{figure/specialcases93941} \caption[]{\label{fig:specialcases93941}}
+\end{figure}
+
+\begin{figure}[]
+
+\includegraphics[width=\maxwidth]{figure/specialcases93942} \caption[]{\label{fig:specialcases93942}}
+\end{figure}
+
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
+
+\includegraphics[width=\maxwidth]{figure/specialcases091} \caption[]{\label{fig:specialcases091}}
+\end{figure}
+
+\begin{figure}[]
+
+\includegraphics[width=\maxwidth]{figure/specialcases092} \caption[]{\label{fig:specialcases092}}
+\end{figure}
+
+\end{knitrout}
+
+As we can see from figure~\ref{fig:specialcases93942} and \ref{fig:specialcases092}, special cases differ strongly 
+from normal cases within the low and top percentiles. Within the tax period 1993/94 the fifth percentile of 
+normal cases is 15.8\% higher than the fifth percentile of the combines data while the 95\% percentile of 
+normal cases is even 0.7\% lower if one leaves out special cases. This indicates special cases are very 
+differnt from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
+million tax units defined as being normal cases compared to 0.51 million special cases.}
+In 2009 the situation has the same pattern. The fifth percentile of normal cases has 11\% higher taxable 
+income compared to data where special cases are included. The five percent top incomes are 4.9\% higher 
+within all cases compared to normal cases only. The share of special cases however decreased 
+to 7.2\%.\footnote{2009: 3.42 million normal cases, 0.27 million special cases}
+
+\emph{Income measures with and without zeros}
 
@@
 
-%Switzerland (income and wealth)
-%- graph with gini over time> identify periods of change
-%- closer look at interesting periods with relative distribution methods-> is change because of down- or upgrading?
-%Cantonal level (income and wealth)
-%- different development on regional level 
-%Switzerland in international comparison
-%- Comparison with other data sources- LIS (inequality in Switzerland is decreasing, which is quite special, because the common pattern in western countries shows an increase in inequality
-
-%- Comparison with other countries (similar data)
-
-% Es scheinen mir aktuell etwas viele Ergebnisse zu sein und ich frage mich, ob wir Ergebnisse auf der Kantonsebene ausklammern sollen.
-Wenn wir drei Ergebniss haben (1) Unterscheidung von unterschiedlichen Perioden der Ungleichheitsentwicklung (2) Beurteilung dieser Perioden hinsichtlich dem Bereich der Veränderungen (Arme, Reiceh) und Entwicklung der Vermögensungleichheit, dann haben wir bereits viel.
-Falls wir spezifische Erkenntnisse nur aus dem Kantonsvergleich bekommen (Bias durch Nuller o.ä.), dann sollten wir trotzdem damit arbeiten.
-
-%rf: genau. kantone unterstreichen die probleme mit dem 0er bias und die probleme mit "imputing the gap". mehr muss erstmal nicht sein.
+\emph{normal and special cases}
+
+\emph{comparison of tax-data and survey data distribution}
+
+\subsection{Interremporal comparison}
 

### 2014-06-15 paper/subdocuments/results.Rnw
normal and special cases

@@
 
+\emph{Normal versus special cases}
+
+The FTA stopped to publicly report data for special cases after tax period 1993/94. 
+For more recent periods however we can compare data from the 
+BRUELHARTPROJEKT (wie nennen wir das?) to the FTA normal cases, as these data are 
+identical but the BRÜLHART data include special cases. 
+We will have a look at both, the 1993/94 period as the last period where both 
+numbers were reported as well as 2009 where we compare data from the same source 
+but using differnt data sheets.
+
+%ABBILDUNGEN ZUM VERGLEICH NORMAL-/SONDERFÄLLE
+<<specialcases9394, echo=FALSE,warning=FALSE,fig.cap=''>>=
+library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
+library(ggplot2)
+#library(reldist)
+normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
+normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
+pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")
+#special <- read.csv("data/ginis_und_perzentile_sonder.csv",sep="\t")
+normal$case <- "normal case"
+#special$case <- "special case"
+pooled$case <- "pooled, normal and special cases"
+d<-rbind(normal,pooled)
+d<-d %.% 
+  filter(kanton=="CH",steuerperiode==1993.5)
+
+start <- which(names(d)=="p1")
+end <- which(names(d)=="p99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
+normal <- data.frame(t(d[1,start:end]),percentile,case="normal")
+pooled <- data.frame(t(d[2,start:end]),percentile,case="pooled")
+names(normal)[1]=c("inc")
+names(pooled)[1]=c("inc")
+percentiles <- rbind(normal,pooled)  
+ggplot(percentiles, aes(x=percentile,y=inc,group=case,colour=case))+
+	geom_line()+theme_bw()+
+	scale_colour_grey()+
+	xlab("Percentile")+ylab("taxable income")+
+	ggtitle("Taxable incomes of normal and special cases 1993/94")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+
+#plot relative distribution
+rel <- data.frame(pooled$inc,normal$inc,pooled$percentile)
+rel$change <- rel$pooled.inc/rel$normal.inc
+ggplot(rel, aes(x=pooled.percentile,y=change))+
+	geom_line()+theme_bw()+
+	scale_colour_grey()+
+	geom_line(linetype="dashed",aes(y=1),size=2)+
+	xlab("Percentile")+ylab("taxable income")+
+	ggtitle("Taxable incomes of normal versus special cases 1993/94")
+
+#steuerpflichtige estv / stpf bruelhart
+#1- (2762419 / 3268548)
+@
+
+<<specialcases09, echo=FALSE,warning=FALSE,fig.cap=''>>=
+library(foreign)
+library(ggplot2)
+#library(reldist)
+normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2009)
+start <- which(names(normal)=="p1")
+end <- which(names(normal)=="p99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,99)
+normal <- data.frame(t(normal[1,start:end]),percentile)
+names(normal)[1]=c("inc")
+
+#bruelhart data
+pooled <- filter(read.dta("../../data/data_Schweiz.dta"), Veranlagungsperiode==2009, Einheit=="Total")
+start <- which(names(pooled)=="p1")
+p95 <- which(names(pooled)=="p95")
+end <- which(names(pooled)=="p99")
+pooled <- data.frame(t(pooled[1,c(start:p95,end)]),percentile)
+names(pooled)[1]=c("inc")
+pooled$inc<-pooled$inc/1000 #adjust scale
+normal$case <- "normal case"
+pooled$case <- "pooled, normal and special cases"
+percentiles <- rbind(normal,pooled)
+ggplot(percentiles, aes(x=percentile,y=inc,group=case,colour=case))+
+	geom_line()+theme_bw()+
+	scale_colour_grey()+
+	xlab("Percentile")+ylab("taxable income")+
+	ggtitle("Taxable incomes of normal and special cases 2009")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+
+#plot relative distribution
+rel <- data.frame(pooled$inc,normal$inc,pooled$percentile)
+rel$change <- rel$pooled.inc/rel$normal.inc
+ggplot(rel, aes(x=pooled.percentile,y=change))+
+	geom_line()+theme_bw()+
+	geom_line(linetype="dashed",aes(y=1),size=2)+
+	scale_colour_grey()+
+	xlab("Percentile")+ylab("taxable income")+
+	ggtitle("Taxable incomes of normal versus special cases 2009")
+
+#steuerpflichtige estv / stpf bruelhart
+#1- (3422210 / 3689063)
+@
+
+As we can see from figure~\ref{fig:specialcases93942} and \ref{fig:specialcases092}, special cases differ strongly 
+from normal cases within the low and top percentiles. Within the tax period 1993/94 the fifth percentile of 
+normal cases is 15.8\% higher than the fifth percentile of the combines data while the 95\% percentile of 
+normal cases is even 0.7\% lower if one leaves out special cases. This indicates special cases are very 
+differnt from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
+million tax units defined as being normal cases compared to 0.51 million special cases.}
+In 2009 the situation has the same pattern. The fifth percentile of normal cases has 11\% higher taxable 
+income compared to data where special cases are included. The five percent top incomes are 4.9\% higher 
+within all cases compared to normal cases only. The share of special cases however decreased 
+to 7.2\%.\footnote{2009: 3.42 million normal cases, 0.27 million special cases}
+
 \emph{Income measures with and without zeros}

### 2014-06-15 paper/paper.tex
deactivated introduction.Rnw while too many errors (cite)

@@
 %%%-------------------------------------------------%%%
-
-%%%-------------------------------------------------%%%
-%%% Sub document introduction %%%
-%%%-------------------------------------------------%%%
-
-\section{Introduction}
-
-% 1.Abschnitt: Bedeutung von Ungleichheit
-Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular \citep{wilkinson_income_2009}.
-While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders \citep{world_economic_forum_global_2013} and scholars alike.
-Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (\citealt{OECD, 2008 FEHLT}, \citealt{oecd_divided_2011}, \citealt{gornick_income_2013}, \citealt{salverda_changing_2014}).
-Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the ``hollowing of the the middle class'' \citep{Alderson and Doran 2013 FEHLT}.
-Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): ``our democracy is being put at peril.''
-\\
-
-% Allenfalls könnte man auch (Nolan und Whealan, 2014) erwähnen.
-"The social Impact of income inequality: Poverty, Deprivation, and Social Cohesion" Da werden die Thesen rund um die Folgen von Ungleichheit gut zusammengefasst.
-
-% 2.Abschnitt: Bedeutung der Datengrundlage und Einführung von Steuerdaten
-Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
-\citet[8]{Atkinson 2013:8 FEHLT} observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
-On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality undisputedly.
-These concerns have led to the search of alternative data sources, which can supplement the established survey data studies.
-Already \citet{kuznets_economic_1955} used tax data to examine the relationship between economic growth and personal distribution of income.
-Then it took several decades until \citet{piketty_les_2001, piketty_income_2003,piketty_income_2003-1} made the use of tax data fashionable again.
-Following his approach studies on several countries were conducted \citep{atkinson_top_2007,atkinson_top_2010}.
-Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database \citep{Alvaredo et al.
-2014 FEHLT}.
-\\
-
-% 3.Abschnitt: What do we know about Switzerland
-% Offical Data Collections on development of inequality
-As we focus our paper on the case of Switzerland, it is important to embed our work in the context of given publications concerning inequality in income.
-What is known about Switzerland so far?
-Looking for official data, three main sources has to be mentioned, which can be considered as de facto official data sources: EU-SILC, HBS and LIS-data.
-Figure XY shows the results stemming from this three sources while looking at Gini of equivalised disposable income.
-Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level.
-The main focus of EU-SILC is to collect data on a common ``framework'' to ensure comparability among EU-countries and countries living around or within the EU.
-As a Non-EU member Switzerland implemented the instrument not from the beginning (2004) but as from 2007.
-Therefore this times-series doesn't cover time points before 2007.
-As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concerning the distribution of income is the Household Budget Survey (HBS).
-The main focus of this survey lays in providing detailed data on household budgets.
-This allows researcher to look at different income concepts like income before and after public transfers.
-Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistent time series from 2000 to 2011.
-As it can be seen from graph XY the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period.
-A longer period is covered in the LIS-Data-set (1982-2004).
-Data-provider for the LIS Data is the Swiss Federal Statistical Office too.
-In contrast the the aforementioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
-All in all the LIS dataset contains the longest time series on inequality for Switzerland.
-Analyzing this data \citet{gornick_income_2013} found for Switzerland a quite substantially decreases in income inequality, contradictory to the development in most other western countries.
-This result is supported by \citet{grabka_evolution_2012} analyzing the Swiss Household Panel (2000-2009).\footnote{A further official database for income distribution is the OECD-Database.
-It includes measures from Income and Consumption survey as well.
-Additional data for 2008 is available from EU-Survey of Income and Living Conditions (EU-SILC).
-But this change in survey is considered as a strict break.
-Comparison before and since 2008 is not recommended (OECD 2012:315).
-For the sake of completeness the database construtcted by \citet{Deininger and Squire 2006 FEHLT} and the World Income Inequality Database (WIID) have to be mentioned.
-Both datasets do not contain figures  for recent years for Switzerland.
-A further important database on inequality is the GINI database which has been derived from the GINI Country Reports \citep{nolan_changing_2014}.
-But this dataset doesn't cover Switzerland.}  \\
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/Swiss_inequality_Plot}
-
-\end{knitrout}
-
-% 4.Abschnitt: What do we know about Switzerland
-% Studies with tax data
-Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
-\citet{dell_income_2007} used tax data from the Federal Tax Administration to assess the development of concentration of the highest incomes and wealth (top-shares).
-In contrast to most other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
-Using the same approach \citet{foellmi_volatile_2013} expand the Dell et al.
-time series to 2008 finding that the share of top income has risen, the top 0.01\% share even doubled in the last observed 20 years.
-A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
-\citet{Flückiger et al.
-(2007) FEHLT} and also \citet{jeitziner_regionale_2007, jeitziner_regionale_2009} report constant inequality from 1960-1996 respectively from 1995 to 2003.
-Covering a similar time period \citet{bauer_verteilung_1994} and \citet{Bolzani and Abul Naga 2002 FEHLT oder es ist 2001} found decreasing inequality.
-On the other hand \citet{Buchmann and Sacchi (1995) FEHLT oder incomplete, siehe buchmann_zur_????} and \citet{Ernst et al.
-2009 FEHLT} found an increase in the 1980s).} \\
-
-% 5. Abschnitt: Zwischenbilanz und Herausforderungen
-Divergence can be explained with several factors.
-First of all, different data sources were used.
-The official data providers trust on survey data, whereas the later mentioned publications use tax data.
-It is assumed that the coverage of top incomes is better in tax data than it is in survey data (non-respondent bias), which is a crucial issue concerning inequality.
-On the other hand the focus on top income neglects other changes in the distribution of income as it is not possible to see, whether newer concerns like the ``hollowing of the middle class'' occurred in Switzerland or not, which leads to the second point.
-Different measure of inequality hampers the comparability.
-Third, different income concepts and different units of analysis were used.
-As it is shown by \citet{modetta_einkommensungleichheit_2012} income distribution is strongly affected by governmental redistribution, reducing inequality substantially.
-With the focus on tax data the change in institutional settings is not covered.
-Also neglected is the household structure, whereas it is unclear how inequality is affected whether one looks at household income or at income of tax units.
-It can be assumed, that inequality corresponding to different concepts react differently on demographic change (change in household structure).
-
-% Zusammenfassung Stand der Forschung und festlegen unserer Ausrichtung
-
-Up to the day, Switzerland can be situated according to the actual level of income inequality in western societies as there is a huge effort to collect data which can be harmonized to comparable measures (see Luxembourg income study, EU-SILC).
-However, it is unclear how the bias through non-response affects the overall measure of inequality.
-For the US \citet{atkinson_top_2009} estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by \citet{Alvaredo (2010) FEHLT}.
-Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) is missing.
-Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.
-First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
-Second, we summarize and apply tax data specific techniques to construct suitable measures of inequality.
-We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
-Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through data source.
-
-% Könnte noch besser auf die geplanten Analysen abgestimmt sein (vgl. auf Github-Issues)
-
-% Was wollen wir wissen?
-
-% Wir müssen uns äussern zu: 
-% Analyseinheiten
-% Definition des Einkommens
-% Messung der Verteilung
-
-% Bei den Ergebnissen müsste es Grafiken haben zu:
-% 1) Unterschiedlichen Analyseneinheiten (Mit und ohne Nuller, Normal- und Sonderfälle)
-% 2) Unterschiedliche Einkommensgrössen (Reineinkommen, steuerbares Einkommen, pseudoverfügbares Einkommen (Abzüge von direkten Bundessteuern))
-% 3) Vergleich mit HABE
-
-% Textbaustein politische Dimension von Ungleichheit
-% What is known about Switzerland so far?
-Among the European Nations Switzerland can be situated in the middle-field when looking at inequality of disposable income (EU-SILC,Eurostat (2013)).
-Nevertheless in recent years Switzerland experienced  an ongoing political debate about the distribution of income in general and the widening of the gap between low and top earners.
-The debate went along with referendums trying to regulate the market outcomes.
-Whereas the "Rip-Off-Initiative" (Abzockerinitiative) found a surprisingly high majority of 68\%, the 1:12-Initiative, which aimed at the whole spectrum of the income distribution, was rejected.
-The voting about the minimum wage-initiative will be held on 9th February 2014.
-All these referendums questioned the current distribution of income and were accompanied by a broader discussion about the development of income inequality held in the media.
-Several official and semi-official publications addressed the question of rising inequality in Switzerland.
-\\
-
-% Alte Textbausteine zu Vermögen
-%While there are several studies about income inequality in Switzerland, the publications concerning the distribution of wealth are scarcer, albeit the distribution of wealth seems to be a relevant dimension shaping the economic well-being of individuals as well.
-The cross-national Data center in Luxembourg is expanding their efforts, constructing the first cross-national wealth database.
-But, up to the day information for Switzerland is not available.
-% ab hier wird es irgnediwe unklar
-%Shorrock et al.
-(2013) unify several datasources (including tax data) to assess the pattern of global wealth.
-Their databook includes figures for Switzerland.
-In contrast to inequality of income concerning inequality of wealth Switzerland takes a leading position \\
-% Albeit the awarness of an assessment of income and wealth simuntanously is rising, we focus in this paper on income inequality, which is the know as the core driver of economic inequlity
-% Dann lassen sich einige Passagen streichen.
-Man könnte auch auf verschiedene Publikationen verweisen, welche die Bedeutung von Vermögen untersuchen -> Buch von Piketty, Müller und Schoch und dann aber ziemlich schnell darauf verweisen, dass wir uns hier auf Einkommen konzentrieren.
-% Auch erste Dell Studie erwähnen -> es hat Zahlen zur Schweiz. Auch Probleme erwähnen (Fehlende Informationen zu Altersvorsorgevermögen -> hat aber niemand eine Lösung bisher.)
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# \\\#\# Attaching package: 'reshape'\\\#\# \\\#\# Die folgenden Objekte sind maskiert from 'package:plyr':\\\#\# \\\#\#\ \ \ \  rename, round\_any}}\end{kframe}
 \includegraphics[width=\maxwidth]{figure/different_ginis1}
@@
 
-\emph{Normal versus special cases}
+\subsubsection{Normal versus special cases}
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
-
-\includegraphics[width=\maxwidth]{figure/specialcases93941} \caption[]{\label{fig:specialcases93941}}
-\end{figure}
-
-\begin{figure}[]
-
-\includegraphics[width=\maxwidth]{figure/specialcases93942} \caption[]{\label{fig:specialcases93942}}
-\end{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/specialcases93941}
+
+\includegraphics[width=\maxwidth]{figure/specialcases93942}
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
-
-\includegraphics[width=\maxwidth]{figure/specialcases091} \caption[]{\label{fig:specialcases091}}
-\end{figure}
-
-\begin{figure}[]
-
-\includegraphics[width=\maxwidth]{figure/specialcases092} \caption[]{\label{fig:specialcases092}}
-\end{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/specialcases091}
+
+\includegraphics[width=\maxwidth]{figure/specialcases092}
 
@@
 normal cases is even 0.7\% lower if one leaves out special cases. This indicates special cases are very 
-differnt from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
+different from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
 million tax units defined as being normal cases compared to 0.51 million special cases.}
@@
 Including zeros leads to significantly higher gini coefficients.
-However we must keep in mind, that these might be artifically high values as we assume zero income for everyone in the zero group.
-We can conclude more from the graphic: the ratio between both measures seems to be quite constant although for aggregate Switzerland but there are minor deviations for multiple cantons as well as strong deviations for the cantons Geneva and Tessin.
+However we must keep in mind, that these might be artificially high values as we assume zero income for everyone in the zero group.
+We can conclude more from the graphic: the ratio between both measures seems to be quite constant although for aggregate Switzerland but there are minor deviations for multiple cantons as well as strong deviations for the cantons Geneva and Ticino.
 However the problems seem not to result from a shift in the zero-share over time but they are specific for the time-period when the tax system changed.
@@
 
-\subsection{Interremporal comparison}
+\subsection{Intertemporal comparison}
 

### 2014-06-15 paper/subdocuments/results.Rnw
normal/special cases results

@@
 %ABBILDUNGEN ZUM VERGLEICH NORMAL-/SONDERFÄLLE
-<<specialcases9394, echo=FALSE,warning=FALSE,fig.cap=''>>=
+<<specialcases9394, echo=FALSE,warning=FALSE>>=
 library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
@@
 
-<<specialcases09, echo=FALSE,warning=FALSE,fig.cap=''>>=
+<<specialcases09, echo=FALSE,warning=FALSE>>=
 library(foreign)
@@
 
-\subsection{Interremporal comparison}
+\subsection{Intertemporal comparison}

### 2014-06-16 paper/subdocuments/data_and_methods.Rnw
minor changes
normal and special cases
proofread subsection "standards for measuring".

@@
 Being called a war-tax in the beginning, the federal tax was renamed to crisis levy in 1934, defense-tax in 1939 and is finally known as direct federal tax since 1983.
-The time frame we were able to collect ranges from 1945 to 2010 including 44 tax periodes.
-\footnote{Between 1993 and 2003 there is no exact data available for Switzerland because of a system change from taxation assessed in arrears (Praenumerando-System) to taxation assessed on current year income (Postnumerando-System) which was implemented by cantons in different years.}  While the FTA provides data in electronic form since 1973 we collected earlier data by scanning hard copies.
+The time frame we were able to collect ranges from 1945 to 2010 including 44 tax periods for cantons and Switzerland.
+\footnote{Between 1993 and 2003 there is no exact data available for aggregate Switzerland (but for individual cantons) because of a system change from taxation assessed in arrears (Praenumerando-System) to taxation assessed on current year income (Postnumerando-System) which was implemented by cantons in different years.}  While the FTA provides data in electronic form since 1973 we collected earlier data by scanning hard copies.
 Data is available for Switzerland plus all cantons and basically covers every tax unit (individual or household) in Switzerland liable to pay federal taxes.
@@
 All other tax units and very few that are taxed based on the style of living because they don't work (Pauschalbesteuerte) are special cases.
+\\
+
+Data is provided by the FTA in an aggregate form for privacy reasons, i.e.
+they are classified into numerous income brackets.
+While we don't know individual incomes we still have sufficient information (number of tax units per income bracket plus sum of incomes within each bracket) to calculate percentiles, Gini coefficients and other desired measures.
+However, equivalization (weighting income by household members) is not possible from these aggregate data.
+In any parts of the article where we point to equivalized income, the data stem from BRÜLHART_WIE_NENNEN_WIR_DAS.
+Their measures are calculated from the same source (FTA data) but used the number of children and marriage status as an approximation of household size to calculate inequality measures that are pseudo equivalized.
 \\
@@
 However, recent activities emphasize the need of a broader conceptualization.
-A recent publication from the \citet{oecd_oecd_2013 UNSICHER} condense these ideas into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framework presented in this publication.
+A recent publication from the \citet{oecd_oecd_2013 UNSICHER} condense these ideas into the ICW framework (income, consumption and wealth), which is meant to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the expert group in developing the ICW Framework presented in this publication.
 Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_????
-UNSICHER}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
+UNSICHER}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}}.
+According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
 To gain policy relevant insight, it is recommended to look at the distribution of all three distributions simultaneously.
@@
 But it is also stated \citep[18]{oecd_oecd_2013 UNSICHER}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being.
-http://www.lisdatacenter.org/our-data/lws-database/} \\
+\url{http://www.lisdatacenter.org/our-data/lws-database/}} \\
 
 This last statement holds for Switzerland too, although the HBS study is strongly influenced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
-Albeit the awareness of an assessment of income, consumption and wealth simultaneously is rising, we  focus our analysis on income, which is undoubtedly a crucial indicator of economic wellbeing.
-It should be noted, that the Federal Tax Office publishes statistics on income and wealth.
-But it is not possible to analyze the jointly distribution on the individual or household level.
-Also measures of consumption are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow.
+Albeit the awareness of an assessment of income, consumption and wealth simultaneously is rising, we focus our analysis on income, which is undoubtedly a crucial indicator of economic well-being.
+It should be noted, that the Federal Tax Administration (FTA) publishes statistics on income and wealth but it is not possible to analyze the joint distribution on the individual or household level.
+Also measures of consumption are largely missing in tax data, albeit deductions can in some sense be understood as mandatory consumptions.
 
@@
 Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013 UNSICHER}, \citet[24]{united_nations_canberra_2011}).
-Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually leads to the disposable income.
-The income measure, which finaly shapes the possibility to consume.
+Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually lead to disposable income: the income measure, which finally shapes the possibility to consume.
 Within this framework common other income definitions are situated.
-These definitions are contrasted with income concepts, which can be derived out of the FTA Tax data.
+These definitions are contrasted with income concepts, which can be derived out of the FTA tax data.
 
@@
 The central income reported through tax statistics is the taxable income.
-It includes all reported incomes (income from employment, income from property and received transfers\footnote{ Mean-tested benefits are not taxed and therefore not included in Tax data.
+It includes all reported incomes (income from employment, income from property and received transfers\footnote{Means-tested benefits are not taxed and therefore not included in tax data.
 Income for low income groups are therefore underestimated.
-However, Piketty and Saez (2001) note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions.
+However, \citep{Piketty and Saez (2001) FEHLT} note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions.
 It is therefore neither a pre-transfer income nor a post-transfer income measure.
 It's rather something in between.
-As the FTA tax statistics include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions.
-This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction, Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income".
-As some deductions can be interpreted as compulsory expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption.
+As the FTA tax statistics include some but not all deductions\footnote{The difference between the real total income and the taxable income are deductions.
+These include: professional expenses, travel expenses, interest on debt, alimonies, training costs, two-earner deduction, party contributions, private pension provision ``Säule 3a'', buying into the pension plan and sideline deductions} it is possible to calculate a sort of "total income".
+As some deductions can be interpreted as compulsory expenses similar to taxes the step towards total income is a step away from the income, which can be used for consumption.
 Similar when calculating the disposable income out of the taxable income through accounting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
@@
 \\
+%RF (16.06.2014): den letzten abschnitt finde ich etwas schwer zu verstehen als jemand der nicht tief in der materie ist.
+vielleicht könnte man sowas machen: It is possible to calculate a kind of "total income" that is in between net and gross income.
+The principle applied is to only deduct expenditures that are unavoidable.
 
@@
 
-\emph{statistical units} \\
+\emph{Statistical units} \\
 
-The agreed standard on the statistical unit, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013 UNSICHER}.
-Indeed it are the individuals, who receive income, own wealth and experience economic wellbeing, but their possibility to do so, is strongly tied to the concept of household.
+The agreed standard on the statistical units, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013 UNSICHER}.
+Indeed it is the individuals, who receive income, own assets and experience economic well-being, but their possibility to do so, is strongly tied to the concept of household.
 This comprises all persons under the same housing arrangement.
-The basic underlying assumption for collection data on households level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children).
+The basic underlying assumption for collecting data on household level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children).
 Additionally, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing.
-The compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013 UNSICHER}, \citealt{Buhmann et al., 1998 FEHLT}).
+To compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013 UNSICHER}, \citealt{Buhmann et al., 1998 FEHLT}).
 \\
@@
 This doesn't imply, that those couples live together, as it is needed to satisfy the definition of a household.
-On the other hand, is it quit likely that more than one tax units live in the same household (unmarried/unregistered couples, see \citet[99]{muller_vermogenslage_2014}).
-It is therefore not directly possible to elicit households and household income out of tax data.
-This might influence the assessment inequality, taking into account the change from traditional household and family structures over the last century.
+On the other hand, is it quite likely that more than one tax unit live in the same household (unmarried/unregistered couples, see \citet[99]{muller_vermogenslage_2014}).
+It is therefore not directly possible to elicit households and household income from tax data.
+This might influence the assessment of inequality development, taking into account the change from traditional household and family structures over the last century.
 
@@
 The Gini coefficient is the most known measure and mainly used for international comparison.
-As it is derived from the lorenz-curve, the quantified amount of inequality can unpretentiously be described in a formal and visual way.
-Therefore the Gini coefficient is easely interpretable.
+As it is derived from the Lorenz-curve, the quantified amount of inequality can unpretentiously be described in a formal and visual way.
+Therefore the Gini coefficient is easily interpretable.
 Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}.
 (1) principle of population: the assessment of inequality is independent of the population size (2) Bresciani-Turroni: the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
-doubling of all income) (3) weak principle of transfers (pigou-dalton): transfers from richer households lead to a reduction of inequality.
-However, several drawbacks are reported in the literatur.
-The most important point is, that the underlying distributional form of the measured inequality is unknown and it is therefore not possibel to see if the measure is driven be a few rich or many poor individuals.
-This can also be problematic for comparision between countries or over time.
-In extreme cases two totally different distrubtions share the same gini-coefficient.\\
+doubling of all income) (3) weak principle of transfers (Pigou-Dalton): transfers from richer households lead to a reduction of inequality.
+However, several drawbacks are reported in the literature.
+The most important point is, that the underlying distributional form of the measured inequality is unknown and it is therefore not possible to see if the measure is driven by a few rich or many poor individuals.
+This can also be problematic for comparison between countries or over time.
+In extreme cases two totally different distributions share the same Gini-coefficient.\\
 
 The recent wave of tax-data studies do not report Gini-coefficents.
-Rather top income shares are informed on, which are calculated not only with taxdata, but togehter with external sources to produce the population and personal income control totals.
-This procedure ensures, that the inequality measure is not biased because of non-fillers, who doesn't appear in tax statistics.
-Leight (2002:594f) compares top income shares with other inequality measures and asks, if they are a useful measure of inequality in a society.
-He tries to answer this question empiricaly by comparing measures of inequality based on top income shares with measures of household or familiy inequality.
+Rather top income shares are informed on, which are calculated not only with tax data, but together with external sources to produce the population and personal income control totals (RF 16.06.2014: hä, was?
+gehts da um die Randverteilung der Bevölkerung?).
+This procedure ensures, that the inequality measure is not biased because of non-fillers, who do not appear in tax statistics.
+\citet{Leight (2002:594f) FEHLT} compares top income shares with other inequality measures and asks, whether they are a useful measure of inequality in a society.
+He tries to answer this question empirically by comparing measures of inequality based on top income shares with measures of household or family inequality.
 He finds a strong positive relationship, but concludes (P.600): top income shares are far from perfect as a measure of distribution of income across society.
@@
 Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent see above).
-A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between invidviduals, who are either both within the top group or both outsie the top group, then the share measure will remain unchanged.
+A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between individuals, who are either both within the top group or both outside the top group, then the share measure will remain unchanged.
 \\
@@
 The polarityindex is developed by \citep{Handock and Morris 1999} allows for this.
-Recenently this index was applied in the work of \citep{Alderson, Beckfield and Nielsen 2005 FEHLT} and \citep{Alderson and Doran 2013 FEHLT}.
+Recently this index was applied in the work of \citep{Alderson, Beckfield and Nielsen 2005 FEHLT} and \citep{Alderson and Doran 2013 FEHLT}.
 This approach is rooted in relative distribution methods.
 It includes a precise comparison of the shape of two distributions (groups, over time).
-The main advantage lies herin, that this approche allows to characterize the change, in the sense that it is possible to say, if a polarization occured (reduction of middle class), which equals an increase of inequality and if this change is driven by a change in the upper or lower part of the distribution.
+As the main advantage this approach allows to characterize the change in detail.
+It is possible to see which parts of the distribution changed, e.g whether a polarization occurred (reduction of middle class) -- which equals an increase of inequality -- or this change is driven by either a change in the upper or lower part of the distribution.
 \\
 
-The literature mentions several other metrics with desired properties, but we will not discuss them further (see for example \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}).
+The literature mentions several other metrics with desired properties we will not discuss here (see for example \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}).
 
@@
 1992 FEHLT}, but require a register for every unit, that is proportional to income.
-Currently there exists no such register for Switzerland \citep{ecoplan 2014:43 FEHLT}.
-Currently used micro datasets, which are used for official publications concerning inequality in Switzerland (SILC and HABE) are furthermore confronted with a constructed coverage problem, because these surveys rely on the phone register, which excludes the households not having a registered connection.
+Currently no such register exists for Switzerland \citep{ecoplan 2014:43 FEHLT}.
+Currently used micro datasets, which are used for official publications concerning inequality in Switzerland (SILC and HABE) are furthermore confronted with a constructed coverage problem, because these surveys rely on the phone register, which excludes households not having a registered connection.
 \\
@@
 The issue of incomplete coverage is less dramatical with tax data.
-Essentially every permanent resident in Switzerland, who has completed the age of 18 years (respectively 20 years prior to 1996) is taxed on a yearly base (or every two years before the change of the tax system).
+Essentially every permanent resident in Switzerland over 18 years of age (respectively 20 years of age prior to 1996) is taxed on a yearly base (or every two years before the change of the tax system).
 Essentially this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distribution.
-This includes a separation of normal cases, which embrace the majority of taxpayers, and the special cases, which cover Foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
+This includes a separation of normal cases, which embrace the majority of taxpayers, and the special cases, which cover (not only) foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
 Most important this includes high net wealth individuals taxed according to their expenditures.
@@
 This is possible for normal and special cases alike.
-From 1995/1996 until 2010 the number of non-taxed units are reported, but not for the years before.
+From 1995/1996 until 2010 the number of non-taxed units is reported, but not for the years before.
 Dell et al.
 (2007) try to estimate the fraction of non-taxed by comparing the reported numbers of tax units to census reports about the number of adult population.
-Following his etimations this fraction drops from 94\% in 1993/1994 to 63\% back in 1945/46.
+According to their estimations this fraction drops from 94\% in 1993/1994 to 63\% back in 1945/46.
 \\
 
-An other critical issue with tax data is the problem of tax evasion, which definitely can bias the assessment of inequality.
-Alvaredo and Saez (2006) for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
-Evasion can occured, when individuals try not to fill tax returns or by misreporting of incomes.
+Another critical issue with tax data is the problem of tax evasion, which definitely can bias the assessment of inequality.
+\citet{Alvaredo and Saez (2006) FEHLT} for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
+Evasion can occur, when individuals try not to fill tax returns or by misreporting of incomes.
 In Switzerland non-fillers show up in the tax-statistics either way, as long as they are registered.
-This person get an imputed income form an older tax return and infomration given by employers.
+This person gets an imputed income based on an older tax return and information given by employers.
 Only non registered non-fillers are not in the records.
 Therefore non-fillers are a minor problem.
-Not neglectable is the circumstance, that individuals misreport incomes.
+Not negligible is the circumstance, that individuals misreport incomes.
 \citet{feld_tax_2006} examine the role of tax evasion in Switzerland by calculating the difference of the national accounts measures of primary income and the income reported to the tax authorities.
@@
 % Beschreiben wie wir relative distribution technics einsetzen und den Polaritätsindex berechnen
+
+% Calculate the gini
+% Ein bisschen erklären, wie aus gruppierten Tabellen der Gini-Koeffizient berechnet wird.
+
+% Polaritätsindex
+% Alderson and Doran (2013:56) beschreiben zwei mögliche Indizies Median relative polarization index (MRP) eingeführt von (Morris, Bernhardt und Handcock 1994:217), der in den lower relative polarization index(LRP) und den upper relative polarization index (URP) zerlegt werden kann (ebd.209)
+% (1. Schritt) Probability density functions
+% (2.
+Schritt) Overall relative density function -> Achtung diese Dichtefunktion bildet sowohl die  Lokationsverschiebung als auch durch Verschiebung der Form ab.
+Eigentlich interessiert aber, der shape shift
+% Deshalb wird die Verteilung um die Lokationsverschiebung bereinigt und nur der shape-shift angeschaut.
+Ausgewiesen werden (MRP)-> hat polarisierung stattgefunden: LRP und URP-> ist polarisierung eher auf eine Zunahme armer (LRP) zurückzuführen oder auf eine Zunahme von Reichen?
+(URP)
 

### 2014-06-16 paper/subdocuments/introduction.Rnw
Literature update (introduction + bib-file)

@@
 While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders \citep{world_economic_forum_global_2013} and scholars alike.
-Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (\citealt{OECD, 2008 FEHLT}, \citealt{oecd_divided_2011}, \citealt{gornick_income_2013}, \citealt{salverda_changing_2014}).
-Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the ``hollowing of the the middle class'' \citep{Alderson and Doran 2013 FEHLT}.
-Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): ``our democracy is being put at peril.''
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (\citealt{oecd_growing_2008}, \citealt{oecd_divided_2011}, \citealt{gornick_income_2013}, \citealt{salverda_changing_2014}).
+Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the ``hollowing of the the middle class'' \citep{gornick_how_2013}.
+Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by \citet[117]{stiglitz_price_2012}: ``our democracy is being put at peril.''
 \\
@@
 Given the importance of the subject a constant reflection about reliability of empirical data seems appropriate.
-\citet[8]{Atkinson 2013:8 FEHLT} observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
+\citet{gornick_foreword_2013} observes advances in technology and methodology which improves the core sources of inequality research, the household surveys.
 On the other hand the labor intensive and expensive surveys around the world are subject to budget cuts and the instrument itself faces problems in form of low response rates, which affects the assessment of inequality undisputedly.
@@
 Following his approach studies on several countries were conducted \citep{atkinson_top_2007,atkinson_top_2010}.
-Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database \citep{Alvaredo et al.
-2014 FEHLT}.
+Today, all existing top income tax statistics based time series are collected and accessible through the world top incomes database \citep{alvaredo_world_2014}.
 \\
@@
 Comparison before and since 2008 is not recommended (OECD 2012:315).
-For the sake of completeness the database construtcted by \citet{Deininger and Squire 2006 FEHLT} and the World Income Inequality Database (WIID) have to be mentioned.
+For the sake of completeness the database construtcted by \citet{deininger_new_1996} and the World Income Inequality Database (WIID) have to be mentioned.
 Both datasets do not contain figures  for recent years for Switzerland.
@@
 A result which opposes the outcome of official data published by the Swiss Federal Statistical Office.\footnote{There are other studies on Switzerland covering different periods but not the recent years.
-\citet{Flückiger et al.
-(2007) FEHLT} and also \citet{jeitziner_regionale_2007, jeitziner_regionale_2009} report constant inequality from 1960-1996 respectively from 1995 to 2003.
-Covering a similar time period \citet{bauer_verteilung_1994} and \citet{Bolzani and Abul Naga 2002 FEHLT oder es ist 2001} found decreasing inequality.
-On the other hand \citet{Buchmann and Sacchi (1995) FEHLT oder incomplete, siehe buchmann_zur_????} and \citet{Ernst et al.
-2009 FEHLT} found an increase in the 1980s).} \\
+\citet{fluckiger_egalite_2002} and also \citet{jeitziner_regionale_2007, jeitziner_regionale_2009} report constant inequality from 1960-1996 respectively from 1995 to 2003.
+Covering a similar time period \citet{bauer_verteilung_1994} and \citet{bolzani_distribution_2001} found decreasing inequality.
+On the other hand \citet{buchmann_zur_1995} and \citet{ernst_inequality_2000} found an increase in the 1980s).} \\
 
@@
 
-% Zusammenfassung Stand der Forschung und festlegen unserer Ausrichtung
+% Zusammenfassung Stand der Forschung und Strukturierung des weiteren Artikels
 
@@
 However, it is unclear how the bias through non-response affects the overall measure of inequality.
-For the US \citet{atkinson_top_2009} estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by \citet{Alvaredo (2010) FEHLT}.
+For the US \citet{atkinson_top_2009} estimate that CPS survey data fail to capture about half of the overall increase in inequality measured by the Gini coefficient, a result confirmed by \citet{alvaredo_note_2011}.
 Likewise a long and consistent time-series allowing to identify and explore development patterns on every point of the distribution (not only the top-shares) is missing.
@@
 
-% Könnte noch besser auf die geplanten Analysen abgestimmt sein (vgl. auf Github-Issues)
+% Könnte noch besser auf die geplanten Analysen abgestimmt sein (vgl. auf Github-Issues oder definitives Paper)
 
-% Was wollen wir wissen?
-
-% Wir müssen uns äussern zu: 
-% Analyseinheiten
-% Definition des Einkommens
-% Messung der Verteilung
-
-% Bei den Ergebnissen müsste es Grafiken haben zu:
-% 1) Unterschiedlichen Analyseneinheiten (Mit und ohne Nuller, Normal- und Sonderfälle)
-% 2) Unterschiedliche Einkommensgrössen (Reineinkommen, steuerbares Einkommen, pseudoverfügbares Einkommen (Abzüge von direkten Bundessteuern))
-% 3) Vergleich mit HABE
+%%%%%%%%%%
+% Alte Textbausteine
 

### 2014-06-16 paper/subdocuments/abstract.Rnw
literature update (abstract)

@@
 \begin{abstract}
-There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (\citealt{cooperation_divided_2011 UNSICHER}, \citealt{gornick_income_2013}). 
+There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (\citealt{cooperation_divided_2011}, \citealt{gornick_income_2013}). 
 In Switzerland, however, the development is unclear, as there is evidence for trends in both directions.

### 2014-06-16 paper/subdocuments/data_and_methods.Rnw
update literature (data_and_methods.Rnw and bib.bib)

@@
 Income for low income groups are therefore underestimated.
-However, \citep{Piketty and Saez (2001) FEHLT} note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions.
+However, \citep{piketty_income_2003} note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions.
 It is therefore neither a pre-transfer income nor a post-transfer income measure.
@@
 Additionally, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing.
-To compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013 UNSICHER}, \citealt{Buhmann et al., 1998 FEHLT}).
+To compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013 UNSICHER}, \citealt{buhmann_equivalence_1988}).
 \\
@@
 Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}.
-(1) principle of population: the assessment of inequality is independent of the population size (2) Bresciani-Turroni: the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
-doubling of all income) (3) weak principle of transfers (Pigou-Dalton): transfers from richer households lead to a reduction of inequality.
+(1) ``\textbf{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textbf{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
+doubling of all income) (3) ``\emph{weak principle of transfers}'' or ``\emph{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality.
 However, several drawbacks are reported in the literature.
@@
 
+% namen mit principle ergänzen?
+
 The recent wave of tax-data studies do not report Gini-coefficents.
@@
 This procedure ensures, that the inequality measure is not biased because of non-fillers, who do not appear in tax statistics.
-\citet{Leight (2002:594f) FEHLT} compares top income shares with other inequality measures and asks, whether they are a useful measure of inequality in a society.
+\citet{leigh_how_2007} compares top income shares with other inequality measures and asks, whether they are a useful measure of inequality in a society.
 He tries to answer this question empirically by comparing measures of inequality based on top income shares with measures of household or family inequality.
-He finds a strong positive relationship, but concludes (P.600): top income shares are far from perfect as a measure of distribution of income across society.
-Top income shares hence inform not completely on how inequality evolves elsewhere in the distribution.
-Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent see above).
+He finds a strong positive relationship, but concludes (P.600): ``top income shares are far from perfect as a measure of distribution of income across society.'' Top income shares hence inform not completely on how inequality evolves elsewhere in the distribution.
+Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent as mentioned above).
 A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between individuals, who are either both within the top group or both outside the top group, then the share measure will remain unchanged.
@@
 
+% Leigh mit Seitenzahl zitieren. Auf deutsch würde ich noch ebd.
+
 Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading).
-The polarityindex is developed by \citep{Handock and Morris 1999} allows for this.
-Recently this index was applied in the work of \citep{Alderson, Beckfield and Nielsen 2005 FEHLT} and \citep{Alderson and Doran 2013 FEHLT}.
+The polarizationindex is developed by \citep{Handock and Morris 1999} allows for this.
+Recently this index was applied in the work of \citep{alderson_exactly_2005} and \citep{gornick_how_2013}.
 This approach is rooted in relative distribution methods.
@@
 As the main advantage this approach allows to characterize the change in detail.
-It is possible to see which parts of the distribution changed, e.g whether a polarization occurred (reduction of middle class) -- which equals an increase of inequality -- or this change is driven by either a change in the upper or lower part of the distribution.
+It is possible to see which parts of the distribution changed, e.g.
+whether a polarization occurred (reduction of middle class) -- which equals an increase of inequality -- or this change is driven by either a change in the upper or lower part of the distribution.
 \\
@@
 The magnitude of this bias in Switzerland, however, is unknown.
-Strategies to handle this kind of bias are discussed in the literature \citep{S?rndal et al.
-1992 FEHLT}, but require a register for every unit, that is proportional to income.
-Currently no such register exists for Switzerland \citep{ecoplan 2014:43 FEHLT}.
+Strategies to handle this kind of bias are discussed in the literature \citep{sarndal_model_2003}, but require a register for every unit, that is proportional to income.
+Currently no such register exists for Switzerland \citet[43]{muller_vermogenslage_2014}.
 Currently used micro datasets, which are used for official publications concerning inequality in Switzerland (SILC and HABE) are furthermore confronted with a constructed coverage problem, because these surveys rely on the phone register, which excludes households not having a registered connection.
@@
 Another critical issue with tax data is the problem of tax evasion, which definitely can bias the assessment of inequality.
-\citet{Alvaredo and Saez (2006) FEHLT} for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
+\citet{alvaredo_income_2009} for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
 Evasion can occur, when individuals try not to fill tax returns or by misreporting of incomes.

### 2014-06-16 paper/subdocuments/data_and_methods.Rnw
minor changes methods section

@@
 We will present these time-series in the result section and evaluate the use of the FTA Tax Data to assess the development of inequality in Switzerland
+
+% Nicht angesprochen sind in der Einleitung die Punkte 4 und 5 - Population coverage und intertemporal comparison. Der letzte Satz gefällt mir auch nicht so...
 
@@
 
-% Bei den Nullern erw?hnen, ab wann Angaben zur Zahl der Nuller vorliegen und ab wann nicht
-% Im Internet hat es Publikationen ab 1947/48
-% Es gibt auch Tabellen auf Gemeindeebene
-% Die Statitstischen Kennzahlen von Br?lhart erw?hnen?
-% Erw?hnen in welcher Form die Informationen publiziert sind. Tabulation by size of income. Not on individual or household level
+% Verschiedene Punkte werden zweimal erwähnt (hier und in den folgenden Abschnitten). Man muss mal alles am Stück lesen und schauen, ob es so passt.
 
-% http://www.estv.admin.ch/dokumentation/00075/00076/00701/index.html?lang=de
-% Kantone und Gemeinden: 2010 bis 1947/1948
-% Klassierte Tabellen für Abzüge, Einkommen (Steuerbares Einkommen und Reineinkommen) und Steuererträge
-% Normalfälle und Sonderfälle
-
-% Statistsiche Kennzahlen
-% Mit Belastung durch die direkte Bundessteuer: 1973/1974 bis 2010
-% Mit und ohne Belastung durch die direkte Bundessteuer: 1995/1996 bis 2010
-% steuerbares Einkommen/Reineinkommen
-% Jeweils Normal und Sonderfälle
-% Individual-Daten (nicht klassierte daten)
-% äquivalenzeinkommen
+% Verweis auf Brülhartdaten.
 
@@
 However, recent activities emphasize the need of a broader conceptualization.
-A recent publication from the \citet{oecd_oecd_2013 UNSICHER} condense these ideas into the ICW framework (income, consumption and wealth), which is meant to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the expert group in developing the ICW Framework presented in this publication.
-Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_????
-UNSICHER}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}}.
+A recent publication from the \citet{oecd_oecd_2013} condense these ideas into the ICW framework (income, consumption and wealth), which is meant to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the expert group in developing the ICW Framework presented in this publication.
+Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_2004}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}}.
 According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
@@
 Some households with low income, for example, may report adequate levels of consumption expenditure or wealth holdings, or vice-versa.
-But it is also stated \citep[18]{oecd_oecd_2013 UNSICHER}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being.
+But it is also stated \citep[18]{oecd_oecd_2013}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being.
 \url{http://www.lisdatacenter.org/our-data/lws-database/}} \\
@@
 Therefore the awareness of the analyzed concept is crucial.
-Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013 UNSICHER}, \citet[24]{united_nations_canberra_2011}).
+Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013}, \citet[24]{united_nations_canberra_2011}).
 Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually lead to disposable income: the income measure, which finally shapes the possibility to consume.
@@
 As the FTA tax statistics include some but not all deductions\footnote{The difference between the real total income and the taxable income are deductions.
-These include: professional expenses, travel expenses, interest on debt, alimonies, training costs, two-earner deduction, party contributions, private pension provision ``Säule 3a'', buying into the pension plan and sideline deductions} it is possible to calculate a sort of "total income".
+These include: professional expenses, travel expenses, interest on debt, alimonies, training costs, two-earner deduction, party contributions, private pension provision ``Säule 3a'', buying into the pension plan and sideline deductions} it is possible to calculate a sort of total income, which is called ``net inomce'' (Reineinkommen).
 As some deductions can be interpreted as compulsory expenses similar to taxes the step towards total income is a step away from the income, which can be used for consumption.
@@
 The principle applied is to only deduct expenditures that are unavoidable.
-
-% Die Deductions sind noch nicht gänzlich sauber abgebildet.
-Auf der Bilanz-Seite steht, zum steuerbaren Einkommen kommt man, indem vom Reineinkommen die Sozialabzüge(Kinder- und Fremdbetreung) sowie der Abzug für unterstützungsbedürftige Personen abzogen werden.
-% Auf den ESTV-Tabellen sind jedoch noch mehr Abzüge abgebildet. Was soll/muss/können wir nun abziehen? Bin verwirrt...
+%OH (16.06.2014): bin mir nicht sicher, ob's mit gross income einfacher ist (kommt in der Grafik nicht vor)
 
@@
 
-The agreed standard on the statistical units, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013 UNSICHER}.
+The agreed standard on the statistical units, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013}.
 Indeed it is the individuals, who receive income, own assets and experience economic well-being, but their possibility to do so, is strongly tied to the concept of household.
@@
 Additionally, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing.
-To compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013 UNSICHER}, \citealt{buhmann_equivalence_1988}).
+To compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013}, \citealt{buhmann_equivalence_1988}).
 \\
@@
 
-% Leigh mit Seitenzahl zitieren. Auf deutsch würde ich noch ebd.
+% Auf deutsch würde ich noch ebd.
 
@@
 \textbf{deductions} \citet[477]{dell_income_2007}:" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
-\citet[5]{schaltegger_evolution_2011 UNSICHER}: ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
+\citet[5]{schaltegger_evolution_2011}: ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
 Können wir das auch?

### 2014-06-17 paper/subdocuments/results.Rnw
applied relative distribution methods to compare FTA and HBS data

@@
 
+% Nicht eher nur CH-Plot?
+
 \emph{Gini coefficients considering mean-tested benefits}
@@
 pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")
-#special <- read.csv("data/ginis_und_perzentile_sonder.csv",sep="\t")
+#special <- read.csv("data/ginis_und_perzentile_sonder.csv",segep="\t")
 normal$case <- "normal case"
@@
 	scale_colour_grey()+
-	geom_line(linetype="dashed",aes(y=1),size=2)+
+	geom_line(linetype="dashed",aes(y=1),size=1)+
 	xlab("Percentile")+ylab("taxable income")+
@@
 
-\emph{normal and special cases}
-
 \emph{comparison of tax-data and survey data distribution}

### 2014-06-17 code/simulate_data.R

@@ -1,0 +1,170 @@
+## program:     simulate_data.R
+## task:        Simulate individual data points out of distribution percentiles and apply relative distribution techniques
+## project:     Inequality of income and wealth in switzerland
+## subproject:  Assessing inequality with tax data
+## date:        June2014
+
+
+##
+# Prepare FTA-data for 2010 (out of Rudis code) Normal cases
+normal <- read.csv("data/ginis_und_perzentile_normal.csv",sep="\t")
+normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
+normal<-normal %.% 
+  filter(kanton=="CH",steuerperiode==2010)
+start <- which(names(normal)=="p1")
+end <- which(names(normal)=="p95")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95)
+normal <- data.frame(t(d[1,start:end]),percentile,anz_pflichtige=89817)
+names(normal)[1]=c("inc")
+
+
+##
+# Simulate individual data points
+# Percentile-Plot (descriptive  purpose)
+ggplot(normal, aes(x=percentile,y=inc))+
+  geom_line()+theme_bw()
+# cumulative probabilites
+cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95)/100
+# probabilities
+prob<-c(cum.p[1],diff(cum.p), .01)
+# extreme values beyond x (to sample)
+#freq<-max(normal$anz_pflichtige)
+freq<-3000
+init<-0
+fin<-(abs(max(normal$inc,na.rm=TRUE))+5)
+# generate the sequence to take pairs from
+ival<-c(init,normal$inc,fin)
+len<-100 # sequence of each pair
+s<-sapply(2:length(ival),function(i){
+  seq(ival[i-1],ival[1],length.out=len)
+})
+s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
+normal.ind<-data.frame(s)
+names(normal.ind)[1]=c("inc")
+summary(normal.ind)
+hist(normal.ind$inc)
+normal.ind$inc<-normal.ind$inc*7 # Rescale income 
+hist(normal.ind$inc)
+
+##
+# Prepare HABE Data - 2010
+# ATTENTION: not weighted and with disposable income
[…]

### 2014-06-17 code/simulate_data.R
minor correction data preparation

@@ -5,6 +5,10 @@
 ## date:        June2014
 
 
+library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
+library(ggplot2)
+library(reldist)
+
 ##
 # Prepare FTA-data for 2010 (out of Rudis code) Normal cases
 normal <- read.csv("data/ginis_und_perzentile_normal.csv",sep="\t")
@@ -14,7 +18,7 @@ normal<-normal %.%
 start <- which(names(normal)=="p1")
 end <- which(names(normal)=="p95")
 percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95)
-normal <- data.frame(t(d[1,start:end]),percentile,anz_pflichtige=89817)
+normal <- data.frame(t(normal[1,start:end]),percentile,anz_pflichtige=89817)
 names(normal)[1]=c("inc")
 
 
@@ -67,7 +71,7 @@ habe10$inc_habe<-habe10$inc_habe*13/1000 # transform from monthly income to inco
 
 ##
 # Apply Relative Distribution Methods
-library(reldist)
+
 
 
 ##

### 2014-06-18 paper/subdocuments/data_and_methods.Rnw
entwurf fazit (summary)

@@
 
-One advantage of tax data over survey data is their observation period.
-While the FTA data we use range back to 1945, the earliest period of survey data is 2000 (HABE data).
-To be fair, survey data might cover the late periods (2011 and 2012) which are not covered by tax data.
+\emph{Summary}
+
+To define a standard of measuring economic resources and related inequality we discussed different dimensions of demands the data needs to meet.
+To sum up, ideally we want to measure \emph{disposable income} for \emph{all} swiss \emph{households} yearly for an extended period of time.
+Tax data has one weakness when it comes to the ideal income measure as we need to work with ``taxable income'' which is designed to serve taxation purposes.
+However, both concepts -- what needs to be taxed and what can be spent -- are related as both address a distinction between necessary and voluntary expenses.
+By subtracting taxes from taxable income we are confident to satisfyingly come close the theoretically ideal income measure of disposable income.
+The second most important drawback with tax data is that it does not adequately address households.
+There are few situations in which the tax unit equals the household, that is for individuals living alone, with married partner and/or deductible children and nobody else.
+This ``classic'' household setup however became less common in the last decades so we need to assume statements about household inequality based on taxed data became more and more biased.
+The swiss law seems to be slightly outdated in this context.
+Finally however, another advantage of tax data (over survey data) is indeed the observation period.
+While the FTA data we use range back to 1945, the earliest period of survey data is 2000 (HBS data).
 The striking feature of the long FTA time series is its consistency.
-Both population coverage and income measures are consistent for the complete observation period.
+Both population coverage and income measures are rather consistent for the complete observation period\footnote{Limitations might however exist due to tax evasion and major changes in the tax system, e.g.
+1945.}.
 This can not be said for survey data as these are based on samples and therefore require an ideal sampling design or reweighting to be representative and comparable over time.
 We will see how well sampling and weighting is done in the results section.
+
+%RF: damit haben wir WER, WANN, WAS, wollen wir noch ein WOMIT/WAS DARSTELLEN -> methodik -> reldist ?
 

### 2014-06-18 code/simulate_data.R
upate HBS comparison with a year with weights (2005)

@@ -9,6 +9,10 @@ library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
 library(ggplot2)
 library(reldist)
 
+#######################
+#     2010            #
+#######################
+
 ##
 # Prepare FTA-data for 2010 (out of Rudis code) Normal cases
 normal <- read.csv("data/ginis_und_perzentile_normal.csv",sep="\t")
@@ -154,7 +158,7 @@ par(mfrow=c(1,1))
 ##
 # statistical Summary 2
 
-# It seems, that the function needs weigths, so give it some weights...
+# It seems, that the function needs weigths, so we give it some weights...
 
 habe10$wgt<-1
 normal.ind$wgt<-1
@@ -170,5 +174,213 @@ format(rpluy(y=normal.ind$inc,yo=habe10$inc_habe,
              ywgt=normal.ind$wgt,yowgt=habe10$wgt,pvalue=TRUE,
              upper=TRUE),
        digits=3)
+
+
+#######################
+#     2005            #
+#######################
+
+
+####
+# We can do comparison with weights for other years - for example for 2005
+# Bruttoeinkommen stimmt in diesem Jahr (verfügbares Einkommen nicht)
+# Gewichte haben wir auch
+
+
+
+##
+# Prepare FTA-data for 2010 (out of Rudis code) Normal cases
+normal <- read.csv("data/ginis_und_perzentile_normal.csv",sep="\t")
+normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
+normal<-normal %.% 
+  filter(kanton=="CH",steuerperiode==2005)
+start <- which(names(normal)=="p1")
+end <- which(names(normal)=="p95")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95)
+normal <- data.frame(t(normal[1,start:end]),percentile,anz_pflichtige=89817)
+names(normal)[1]=c("inc")
+
+
+##
+# Simulate individual data points
+# Percentile-Plot (descriptive  purpose)
+ggplot(normal, aes(x=percentile,y=inc))+
+  geom_line()+theme_bw()
+# cumulative probabilites
+cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95)/100
+# probabilities
+prob<-c(cum.p[1],diff(cum.p), .01)
+# extreme values beyond x (to sample)
+#freq<-max(normal$anz_pflichtige)
+freq<-3000
+init<-0
[…]

### 2014-06-18 paper/subdocuments/data_and_methods.Rnw

@@
 \begin{tabular}{lcccc}
-                                   & FTA       & HABE      & EU-SILC   & LIS       \\
-Range                              & 1945-2010 & 2000-2011 & 2007-2012 & 1982-2004 \\
-Cross country comparison possible? & no        & no        & yes       & yes       \\
-Top incomes covered?               & yes       & no        & no        & no        \\
-Low incomes covered?               & no        & yes       & yes       & yes       \\
-Concepts of economic ressources?   & no        & yes       & ?         & ?         \\
-Equivalisation?                    & no        & yes       & ?         &          
+                                   & FTA       & Survey data\\
+Statistical unit                   & individuals/couples  & households\\                                   
+Concepts of economic ressources?   & limited   & yes (for recent periods)\\
+Equivalisation?                    & limited   & yes\\
+Top incomes covered?               & yes       & no\\
+Low incomes covered?               & no        & yes\\
+Range                              & 1945-2010 & 1982-2011\\
+Comparison over time possible? 	   & yes       & limited\\
+Cross country comparison possible? & limited   & limited
+
 \end{tabular}

### 2014-06-18 paper/subdocuments/introduction.Rnw

--- 2014-06-13 paper/subdocuments/introduction.Rnw
@@
 % 1.Abschnitt: Bedeutung von Ungleichheit
-Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular \citep{wilkinson_income_2009}.
+Economic resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular \citep{wilkinson_income_2009}.
 While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders \citep{world_economic_forum_global_2013} and scholars alike.
@@
 Already \citet{kuznets_economic_1955} used tax data to examine the relationship between economic growth and personal distribution of income.
-Then it took several decades until \citet{piketty_les_2001, piketty_income_2003,piketty_income_2003-1} made the use of tax data fashionable again.
+Then it took several decades until \citet{piketty_les_2001, piketty_income_2003,piketty_income_2003-}1 made the use of tax data fashionable again.
 Following his approach studies on several countries were conducted \citep{atkinson_top_2007,atkinson_top_2010}.

### 2014-06-18 paper/subdocuments/introduction.Rnw
recommit introduction

--- 2014-06-16 paper/subdocuments/introduction.Rnw
@@
 \\
+
+% Piketty droppen
 
@@
 Data-provider for the LIS Data is the Swiss Federal Statistical Office too.
-In contrast the the aforementioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000,2002,2004).
+In contrast the the aforementioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000, 2002, 2004).
 All in all the LIS dataset contains the longest time series on inequality for Switzerland.

### 2014-06-19 paper/paper.Rnw
normal and special cases
deactivated introduction.Rnw while too many errors (cite)
reactivated introduction.Rnw
compile-o-mat :)
different updates from yesterday

@@
 \usepackage{color} % used to mark parts that need to be edited
-\usepackage{natbib} %to use citealt citet citep
+\usepackage[comma,authoryear]{natbib} %to use citealt citet citep
+%\usepackage{biblatex} %to make use of the date field in .bib-files instead of year-field.
+%\usepackage{url}
 \usepackage{abstract} % Allows abstract customization
@@
 %%%-------------------------------------------------%%%
-%%% How did inequality change over time %%%
-%%%-------------------------------------------------%%%
-
-%<<subdoc_content_introduction, child='subdocuments/change.Rnw', eval=T>>=
-%@
-
-% Section from an older version. Currently disabled.
-
-%%%-------------------------------------------------%%%
 %%% Include data and methods %%%
@@
 
-<<subdoc_data_and_methods, child='subdocuments/data_and_methods.Rnw', eval=T>>=
+<<subdoc_data_and_methods, child='subdocuments/data.Rnw', eval=T>>=
+@
+
+<<subdoc_data_and_methods, child='subdocuments/methods.Rnw', eval=T>>=
 @

### 2014-06-19 paper/subdocuments/data.Rnw
different updates from yesterday

--- 2014-06-18 paper/subdocuments/data_and_methods.Rnw
@@
 Within this framework common other income definitions are situated.
-These definitions are contrasted with income concepts, which can be derived out of the FTA tax data.
 
@@
 
+% Es musse klar werden um was es geht.
+% These definitions are contrasted with income concepts, which can be derived out of the FTA tax data.
+
 The central income reported through tax statistics is the taxable income.
@@
 \\
+
+% market income mit primary income ersetzen!
+% Nicht vergessen: Abzüge können über die Zeit ändern. Was bedeutet das?
+
 %RF (16.06.2014): den letzten abschnitt finde ich etwas schwer zu verstehen als jemand der nicht tief in der materie ist.
@@
 
-\emph{Intertemporal comparison}
-
-\emph{Summary}
-
-To define a standard of measuring economic resources and related inequality we discussed different dimensions of demands the data needs to meet.
-To sum up, ideally we want to measure \emph{disposable income} for \emph{all} swiss \emph{households} yearly for an extended period of time.
-Tax data has one weakness when it comes to the ideal income measure as we need to work with ``taxable income'' which is designed to serve taxation purposes.
-However, both concepts -- what needs to be taxed and what can be spent -- are related as both address a distinction between necessary and voluntary expenses.
-By subtracting taxes from taxable income we are confident to satisfyingly come close the theoretically ideal income measure of disposable income.
-The second most important drawback with tax data is that it does not adequately address households.
-There are few situations in which the tax unit equals the household, that is for individuals living alone, with married partner and/or deductible children and nobody else.
-This ``classic'' household setup however became less common in the last decades so we need to assume statements about household inequality based on taxed data became more and more biased.
-The swiss law seems to be slightly outdated in this context.
-Finally however, another advantage of tax data (over survey data) is indeed the observation period.
-While the FTA data we use range back to 1945, the earliest period of survey data is 2000 (HBS data).
-The striking feature of the long FTA time series is its consistency.
-Both population coverage and income measures are rather consistent for the complete observation period\footnote{Limitations might however exist due to tax evasion and major changes in the tax system, e.g.
-1945.}.
-This can not be said for survey data as these are based on samples and therefore require an ideal sampling design or reweighting to be representative and comparable over time.
-We will see how well sampling and weighting is done in the results section.
-
-%RF: damit haben wir WER, WANN, WAS, wollen wir noch ein WOMIT/WAS DARSTELLEN -> methodik -> reldist ?
-
 \subsection{Comparison of tax data and other data sources - advantages and shortcomings}
@@
 Comparison over time possible? 	   & yes       & limited\\
-Cross country comparison possible? & limited   & limited
+Cross country comparison possible? & limited   & yes
 
@@
 \end{table}
+
+%RF: damit haben wir WER, WANN, WAS, wollen wir noch ein WOMIT/WAS DARSTELLEN -> methodik -> reldist ? jaa
+
+To define a standard of measuring economic resources and related inequality we discussed different dimensions of demands the data needs to meet.
+To sum up, ideally we want to measure \emph{disposable income} for \emph{all} swiss \emph{households} yearly for an extended period of time.
+Tax data has one weakness when it comes to the ideal income measure as we need to work with ``taxable income'' which is designed to serve taxation purposes.
+However, both concepts -- what needs to be taxed and what can be spent -- are related as both address a distinction between necessary and voluntary expenses.
+By subtracting taxes from taxable income we are confident to satisfyingly come close the theoretically ideal income measure of disposable income.
+The second most important drawback with tax data is that it does not adequately address households.
+There are few situations in which the tax unit equals the household, that is for individuals living alone, with married partner and/or deductible children and nobody else.
+This ``classic'' household setup however became less common in the last decades so we need to assume statements about household inequality based on taxed data became more and more biased.
+The swiss law seems to be slightly outdated in this context.
+Finally however, another advantage of tax data (over survey data) is indeed the observation period.
+While the FTA data we use range back to 1945, the earliest period of survey data is 2000 (HBS data).
+The striking feature of the long FTA time series is its consistency.
+Both population coverage and income measures are rather consistent for the complete observation period\footnote{Limitations might however exist due to tax evasion and major changes in the tax system, e.g.
+1945.}.
+This can not be said for survey data as these are based on samples and therefore require an ideal sampling design or reweighting to be representative and comparable over time.
+We will see how well sampling and weighting is done in the results section.
 
@@
 For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).
-
-\subsection{Outline of applied methods}
-
-% Diese Abschnitt muss ersetzt werden mit der von uns gewählten Vorgehensweise
-
-In the last section we described the advantages and drawback of taxdata discussing five aspects, which we regard as crucial concering the assessment of inequality.
-To get a feeling of the importance of these aspects, we exploit the FTA-tax data as far as possible and perform several insightful calculations adressing four of the five mentioned aspects.
-No further investigation is possible regarding aspect (1) concepts on measuring economic ressource.
-This is a point which cannot be further addresses with FTA tax data.
-But for the other four thematic areas, we can provide deeper insight.
-In general our main strategy is to apply different possible concepts within one conceptional area (income, units, measurement, population coverage) while holding other conceptual differences constant.
-With this strategy we want to show, the sensitiveness the assessement of inequality is especially if one looks at time trens.
-In this section we  describe which methods we use to produce the results in the next section.
-
-%%%
-% Was soll hier beschrieben werden und was direkt bei den Tabellen?
-%%% Gar nicht so einfach....
-
-%%% Nach Lehrbuch müssten die methodischen Schritte hier beschrieben werden.
-% Man könnte zu den Zeitreihen jeweils die Varianz ausweisen-> dass würde helfen die Reihe zu beurteilen.
-Man sieht dann: aha, die Reihe ist im Vergleich zur anderen höher/tiefer.
-Ob die Reihen stärkeren Schwankungen unterliegen könnte einfach aus der Varianz gelesen werden
-
-% Defining income
-% Beschreiben, dass sich aus den publizierten Einkommensgrössen und den Steuerbeträgen unterschiedliche 
-% Einkommen berechnen lassen
-% Es steht der Vorwurf im Raum, die unteren Einkommen seien mit Steuerstatistiken schlecht abgebildet, weil bedarfsabhängige Leistungen nicht versteuert sind
-
-% Es müsste eine Reihe zum gesetzliche Grundbedarf vorliegen. Wüsste ich jetzt grad nicht von wo...
-
-%%
-% statistical units
-% Beschreiben, dass zwar keine Haushaltseinkommen gebildet werden können und Skaleneffekte nicht abgebildet sind, dass eine Annäherung jedoch möglich ist
-% Definition der äquivalenzskala bei Brülhart (korrespondiert nicht mit OECD Definitionen, aus pragmatischen Gründen)
-
-%%
-% Measuring inequality
-% Beschreiben, wie sich aus gruppierten Steuerstatistiken der Gini-Koeffizient berechnen lässt
-% Beschreiben wie wir relative distribution technics einsetzen und den Polaritätsindex berechnen
-
-% Calculate the gini
-% Ein bisschen erklären, wie aus gruppierten Tabellen der Gini-Koeffizient berechnet wird.
-
-% Polaritätsindex
-% Alderson and Doran (2013:56) beschreiben zwei mögliche Indizies Median relative polarization index (MRP) eingeführt von (Morris, Bernhardt und Handcock 1994:217), der in den lower relative polarization index(LRP) und den upper relative polarization index (URP) zerlegt werden kann (ebd.209)
-% (1. Schritt) Probability density functions
-% (2.
-Schritt) Overall relative density function -> Achtung diese Dichtefunktion bildet sowohl die  Lokationsverschiebung als auch durch Verschiebung der Form ab.
-Eigentlich interessiert aber, der shape shift
-% Deshalb wird die Verteilung um die Lokationsverschiebung bereinigt und nur der shape-shift angeschaut.
-Ausgewiesen werden (MRP)-> hat polarisierung stattgefunden: LRP und URP-> ist polarisierung eher auf eine Zunahme armer (LRP) zurückzuführen oder auf eine Zunahme von Reichen?
-(URP)
-
-%%
-% Population coverage
-% Analysen zu den Nullern > ist eine Hauptkritik an der Nutzung von Steuerdaten
-% Die Bedeutung von Normal- und Sonderfällen
-% Den Vergleich mit der HABE beschreiben, welche Einkommensgrösse nutzen wir aus der HABE um diese möglichst mit den Steuerdaten vergleichbar zu machen?
-Wir brauchen das Bruttoeinkommen nach Abzug der Sozialversicherungsbeiträge, dies lässt sich mit dem Reineinkommen vergleichen
-
-%%
-% Intertemporal comparison
-% Ziel eine möglichst lange Zeitreihe abzubilden. 
-% Technik der Lückenfüllung von Ben beschreiben> dies kurz anderen Techniken gegenüberstellen.
-% Konfidenzintervalle für die Perioden mit vielen fehlenden Werte?
-
-% eine / oder zwei Zeitreihen (eine lange und eine, die möglichst nahe an der richtigen Konzeptionalisierung ist: post direct taxed income, normal+sonderfälle) zusammen mit den bereits existierenden Reihen (HABE,LIS,SILC) ploten
-
-\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
-\citet{dell_income_2007} impute for non-fillers the 20 percentage of the annual average income.
-This flattens the distribution on the left side, which is not a problem if you are interested in the top income shares, but it would surly affect overall measures of inequality.
-Furthermore the authors calculate the proportion of non-fillers by estimating the total of tax units out of the population records.
-\\
-
-\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
-This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
-This is why during the transitional period from 1995 to 2003 there is no uniform tax data published on the Swiss level but only data on the cantonal level \citep[8f]{foellmi_volatile_2013}.
-\\
-%es wird erwartet, dass der Wechsel Ungleichheitsmasse beeinflussen. Yearly fluctuations are dampened, when income is measured on a two-yearly basis.
-
-\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
-
-\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
-%is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
-% Beim Imputieren m?sste man auch die Zahl der freiwilligen Nichtbeziehenden ber?cksichtigen.
-
-\textbf{deductions} \citet[477]{dell_income_2007}:" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
-\citet[5]{schaltegger_evolution_2011}: ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
-Können wir das auch?
-Zumindest für gewisse Zeiträume?
-Das wäre noch gut.
-\\
-
-Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
-Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.
-On the other hand deductions can affect the distribution.
-There are recent studies about the correlation of progressivity and deductions in Switzerland, which examines if deductions have a ``perverse redistribution'' effect by redistributing income from the lower middle class to the upper middle class (see \citealt{peters_steuerabzuge:_2011} and \citet{Interpellation Barbara Gysel (2009) FEHLT.
-das interpellation barbara sieht falsch aus}.
-
-% Income share specific problems
-% Total income denominator Exogenous Approach -> net income reported in the national accounts.
-Endogenous Approach -> Dell et al.
-technic-> imputing 20\% of average personal income to non-fillers (which are mainly persons with low or no income).
-Honestly , is this appropriate?
-(Everyone did it)
-% Total of tax units in the country. exogenous approach -> construct number of total tax units artificially from other data sources
-% Endogenous approach -> reported in the tax tables
-
-%%%-------------------------------------------------%%%
-%%% Abschnitte aus ersten Version des Papers %%%
-%%%-------------------------------------------------%%%
-
-%\subsection{Hypotheses}
-%Based on the theories we test the following hypotheses:
-
-%\begin{itemize}
-%\item H1: Develpment of inequality is driven by sectoral change
-%\item H2: Development of inequality is driven by political change, i.e. economic crisis contribute to inequality because welfare states tend to be downsized
-%\end{itemize}
-
-%\subsection{Data and Variables}
-
-%We use data from the Swiss Federat Tax Administration (FTA) where our data about incomes ranges from the years 1941/42 to 2010.
-While the data results in a long and consistent time series to illustrate swiss inequality development, there are a few pitfalls we want to adress which might be of interest for other research on this topic (be it in Switzerland or other countries).
-
-%\subsubsection{Left censored data}
-
-%The FTA provides data about all tax units in Switzerland that are liable to pay federal taxes.A tax unit may be a single person or a household.
-The taxable population however is not identical to the population which should be used to calculate measures of inequality.
-Precisely, the data do not contain tax units with very little incomes so calculations based on these data treat the lowest percentiles equally to tax units with zero income.Figure X shows the threshold to be hit to enter the statistic.
-
-%\textcolor{red}{[FIGURE X ABOUT HERE] soll zeigen: Zeitreihe der Untergrenze von 1941-2010}
-
-%So first of all, there is a bias in the level of an inequality measure one could calculate with the   FTA data.
-Furthermore, also the changes over time might not be interpreted savely as over time the number of tax unit within this "hidden range" might vary or might even have a certain trend.
-We will adress this issue in detail in the methods chapter.
-
-%\subsubsection{Different measures, different populations}
-
-%The FTA data makes two kinds of distinctions.
-First, data was collected for so called "normal cases" and "special cases", i.e.
-a "normal case" is a taxable (for the complete tax period) person or household domiciled in a swiss canton without income from outside of Switzerland.
-A "special case" therefore is a diffuse reference category that contains tax units that are taxed at source, were not taxable for the complete tax period or generated additional income in another country.
-Second, the FTA reports two measures, that is taxable income and absolute income \textcolor{red}{(meine vorläufige Übersetzung von Reineinkommen)}.
-Absolute income is the sum of all incomes (earnings, interest income, rental incomes) minus expenses (e.g.
-from self-employment or credit cost).
-The taxable income is calculated as the difference of absolute income and deductions (e.g.
-children, insurance rates).
-The longest consistent time series exists for the taxable income of normal cases.
-So all statements we make with our data only apply to this subpopulation.
-
-%\subsubsection{Changes in taxation and measurement}
-%The swiss tax system is highly federal.
-That means, communities raise taxes which then go to the communities, the canton and the state.
-If we want to calculate overall swiss measures, we need to take into account, that cantons vary (between cantons and over time) with regard to the tax deductions that are possible and also the mechanism how taxes are collected.
-The latter adresses a comprehensive reorganization of the swiss tax system where between 1995 and 2003 cantons changed from taxing the past two years of income (postnumerando system) to taxing the present single year (praenumerando system).
-For details see \textcolor{red}{Martinez (xxxx) or some other author (xxxx)}.
-Aggregate measures of inequality therefore have to be estimated for the periods 1995 to 2003 which we adress shortly in the methods chapter.
-
-%\subsection{Methodology Used}
-%There are two steps of data analyses which need to be described to the reader.
-First, the estimation of the bias we introduce by estimating measures of inequality when tax units with too little income are not observed.
-Second, the steps undertaken to estimate aggregate swiss measures by imputing taxable income for those cantons and periods where the change of the tax system produced a gap (1995 to 2003, depending on the canton).
-
-%\subsubsection{Imputing the gap}
-%The imputation is not a focus of the paper so we basically follow the most simple approach of Martinez (xxxx).
-That is estimating the missing taxable income statistics via OLS using information from time trends and cantons.
-Our imputation model therefore includes canton inequality measures and periods dummies to explain aggregate swiss inequality.
-
-\subsubsection{Estimating the bias}
-For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation (lets call those ``zeros'' for convenience) .
-However starting 1995, the FTA provides exactly this information for each canton.
-This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010.
-Consequently we can obtain information whether the bias is stable over time (which makes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
-\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie ``je höher der Steuerfreibetrag, umso stärker der Bias''.
-Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias.
-Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
-
-One can take a first look at the descriptives plotting the share of zeros over time seperately for each canton.
-
-<<setup>>=
-library(foreign)
-df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
-df$zeros <- df$null_norm/df$cpop # i only use normal cases here
-@
-
-<<zero_descriptives>>=
-library(ggplot2)
-ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+theme_bw()
-ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")+theme_bw()
-@
-
-We can see multiple things here:
-
-\begin{enumerate}
-\item There is a small overall upward trend which we assume to be the Federal Administrations inflation adjustments to the tax threshold.
-\item Geneva and Tessin show wild changes but those might be explained by the tax gap ("Bemessungslücke") that people exploited when the cantons changed the tax system.
-It remains unclear however why we can't see similar patter within other cantons.
-\item There is some variance and we see different patterns over time and cantons.
-When estimating Gini coefficients or the like we must therefore assume that ignoring ``the zeros'' leads to a bias that is not stable over time.
-\end{enumerate}
-
-We might try two strategies to moderate the problem:
-
-\begin{enumerate}
-\item Add the zeros as a separate group
-\item Fit a model to predict the inequality measure (e.g. the Gini coefficient) using the share of zeros as a predictor
-\end{enumerate}
-
-By adding the zeros as a separate group we face several problems.
-With the exception of Geneva 1995/96 Pareto interpolation of percentiles p20 and above seems viable.
-Low percentiles however would need to be extrapolated so we might impose unrealistic assumptions.
-Even when estimating ``safe'' measures above p20 (better p50) or a Gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
-To be consistent with the measure of ``taxable income'' we could assume an income of zero for that group, calculate Gini coefficients and compare them with the original (uncorrected/plain) version.
-To get a rough number we could (by cantons) check the squared correlation between plain and corrected Gini.
-
-The second approach however is more robust as we do not impose additional assumptions but instead exercise some curve fitting.
-
-<<corrected_gini>>=
-df$J<-factor(df$steuerperiode)
-fit <- lm(G_steink~kanton+kanton:zeros+J+zeros:J,data=df)
-#summary(fit)
-SmoothCoefficientPlot <- function(models, modelnames = "", removeintercept = FALSE){
-  # models must be a list()
-
-  Alphas <- seq(1, 99, 2) / 100
-
-  Multiplier <- qnorm(1 - Alphas / 2)
-  zzTransparency <<- 1/(length(Multiplier)/4)
-  #CoefficientTables <- lapply(models, function(x){summary(x)$coef})
-  CoefficientTables <- summary(models)$coef
-  #TableRows <- unlist(lapply(CoefficientTables, nrow))
-  TableRows <- nrow(CoefficientTables)
-
-  if(modelnames[1] == ""){
-    ModelNameLabels <- rep(paste("Model", 1:length(TableRows)), TableRows)
-    } else {
-    ModelNameLabels <- rep(modelnames, TableRows)
-    }
-
-  MatrixofModels <- cbind(do.call(rbind, list(CoefficientTables)), ModelNameLabels)
-  if(removeintercept == TRUE){
-    MatrixofModels <- MatrixofModels[!rownames(MatrixofModels) == "(Intercept)", ]
-    }
-  MatrixofModels <- data.frame(cbind(rownames(MatrixofModels), MatrixofModels))
-
-  MatrixofModels <- data.frame(cbind(MatrixofModels, rep(Multiplier, each = nrow(MatrixofModels))))
-
-  colnames(MatrixofModels) <- c("IV", "Estimate", "StandardError", "TValue", "PValue", "ModelName", "Scalar")
-  MatrixofModels$IV <- factor(MatrixofModels$IV, levels = MatrixofModels$IV)
-  MatrixofModels[, -c(1, 6)] <- apply(MatrixofModels[, -c(1, 6)], 2, function(x){as.numeric(as.character(x))})
-  MatrixofModels$Emphasis <- by(1 - seq(0, 1, length = length(Multiplier) + 1)[-1], as.character(round(Multiplier, 5)), mean)[as.character(round(MatrixofModels$Scalar, 5))]
-
-  OutputPlot <- qplot(data = MatrixofModels, x = IV, y = Estimate,
-   ymin = Estimate - Scalar * StandardError, ymax = Estimate + Scalar * StandardError,
-   ylab = NULL, xlab = NULL, alpha = I(zzTransparency), colour = I(gray(0)), geom = "blank")
-  OutputPlot <- OutputPlot + geom_hline(yintercept = 0, lwd = I(7/12), colour = I(hsv(0/12, 7/12, 7/12)), alpha = I(5/12))
-  OutputPlot <- OutputPlot + geom_linerange(data = MatrixofModels, aes(size = as.integer(1/Emphasis)), alpha = I(zzTransparency), colour = I(gray(0)))
-  OutputPlot <- OutputPlot + scale_size_continuous(legend = FALSE)
-  OutputPlot <- OutputPlot + facet_grid(~ ModelName) + coord_flip() + geom_point(aes(x = IV, y = Estimate), colour = I(gray(0))) + theme_bw()
-  return(OutputPlot)
-  }
-
-library(grid) #for arrow() function
-SmoothCoefficientPlot(fit,modelnames=paste0("By cantons: Impact of the share of zeros on inequality (Gini) Rsq=", round(summary(fit)$r.squared,2)))+geom_text(show_guide=FALSE,x=77,y=2.2,size=4,label="effect variation\nover time")+geom_segment(aes(x = 85, y = 1.5, xend = 71, yend = 1.5))+geom_segment(aes(x = 71, y = 1.5, xend = 71, yend = 1.3))+geom_segment(aes(x = 85, y = 1.5, xend = 85, yend = 1.3))+geom_text(show_guide=FALSE,x=57,y=1.9,size=4,label="effect variation\nby cantons")+geom_segment(aes(x = 71, y = 1.2, xend = 45, yend = 1.2))+geom_segment(aes(x = 71, y = 1.2, xend = 71, yend = 1.0))+geom_segment(aes(x = 45, y = 1.2, xend = 45, yend = 1.0))+ggtitle("Bias variation by time and cantons")
-
-@
-
-The model outputs a test statisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical ``canton gini-level''.
-As the model has a decent fit we are not in great danger of omitted variable bias.
-Using a joint F-Test we can now test if all canton interactions are zero.
-
-<<ftest, echo=FALSE>>=
-library(survey, quietly=TRUE, warn.conflicts = FALSE)
-ftest <- regTermTest(fit,"kanton:zeros")
-@
-
-In our case we can clearly reject the hypothesis that all interactions are zero ($p=\Sexpr{round(ftest$p,3)}$).
-This leads to the conclusion  that Gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted Gini coefficients.
-For example one might be interested in how inequality would had developed if the zero-share would have been constant over time.
-(Note RF: predict all data point using canton, time and the initial OR final zero-share to homogenize the time series)
-Furthermore we can quantify how large the bias is and we can do this separately for tax periods or separately for cantons.
-
-For all cantons:
-
-<<magnitude_of_bias>>=
-reduced_fit <- lm(G_steink~kanton+factor(steuerperiode),data=df)
-#summary(reduced_fit)
-@
-
-We can see the the model fit reduces to explaining 61.5\% of the Gini variation versus 92.6\% when the information from the zero-shares was used.
-Although this so some extent attributable to the additional 26 parameters: this is huge.
-
-The model indicates some cases that deserve more attention: Schwyz (positive coefficient) and Geneva (negative coefficient) and the tax period 2000 as well as the most recent periods.
-
-A positive coefficient (e.g.
-Schwyz) can be read as follows: In periods with many zeros we measure higher Gini coefficients.
-We can derive, that the distribution of income is more skewed for high incomes than for low incomes.
-Simply speaking, the contrast between low and middle class is less pronounced than the contrast between middle and top class.
-One possible explanation would be that incomes stem from two different populations: 1) the people of Schwyz who possibly follow a log-normal or gamma distribution and 2) particularly rich people who moved to Schwyz to avoid taxes.
-
-A negative coefficient (e.g.
-Geneva) means the more zeros there are the smaller the Gini measure was compared to other tax periods within that canton (remember this is a fixed-effects model).
-This is the case we would usually expect: more zeros mask inequality that arises from the bottom.
-
-What can we conclude from that analysis?
-First one must notice that aggregate measure like the Gini (or others) do not always react in the same way when we cut off one part of the distribution, therefore the measures calculated from tax data is biased.
-On the other hand, the model coefficient of Switzerland as a whole is not significant suggesting that the cantonal biases cancel out each other.
-This seems plausible.
-Most of the ``tax optimization'' happens within Switzerland so the rich people who moved to Schwyz are now missing at another canton.
-
-We can even see more from the model coefficients.
-The period dummies indicate how the distribution of incomes has changed: compared to 1995, the subsequent periods have a negative coefficient, i.e.
-the bias we introduce by omitting the zeros increased, especially from the mid nineties to the mid 2000s.
-To simplify: cutting off zeros more and more seems to lead to an underestimation of our inequality measure, probably because the skewness in the left part of the distribution increased pointing to an increased pauperization that is masked by omitting the left tail of the income distribution.
-
-% Alter Textbaustein zu Vermögen
-%\emph{Defining wealth}
-%While the apropiateness of the conceptualization of income is widely discuded, there is no agreed definition of personal wealth yet and the appropriate methods of valuation are not always clear.
-Wealth can be defined as the marketable value of financial assets plus non-financial assets (housing and land) less debts (Credit Suisse 2011:5).
-
-%The discussion about problems with reporting income is fairly exhaustive. What about wealth?
-
-%It is well recognized that the traditional sources of wealth distribution data are unlikely to provide an accurate picture of wealth ownership in the top-tail of the distribution.
-\citet{credit_suisse_global_2011} makes use of the information in the ``Rich Lists'' published by Forbes Magazine to adjust the wealth distribution pattern in the highest wealth ranges.
-\\
-
-%Additionally the Credit Suisse Reports states that these data my be less subject to response bias, but my be more prone to valuation problems, especially in connection with pension assets and debts \citep[8]{credit_suisse_global_2011}.

### 2014-06-19 paper/subdocuments/methods.Rnw
different updates from yesterday

--- 2014-06-16 paper/subdocuments/data_and_methods.Rnw
@@
-%%%-------------------------------------------------%%%
-%%% Sub document for data and methods %%%
-%%%-------------------------------------------------%%%
-
-\section{Data, Measurement concepts and Methods}
-
-Studies on inequality have to address several thorny challenges.
-It starts with answering three crucial questions: First of all one has to define, which concepts should to be looked at.
-This refers to answering the question about inequality of what.
-Secondly one has to be clear about the unit of analysis.
-This refers to answering the question about inequality among whom.
-Thirdly, one has to choose an appropriate measure of inequality.
-All these questions are ideally answered considering theory and a given research question.
-Often it has to be answered in context of a given dataset.
-Therefore we start this section with a description of the FTA-Tax Data.
-Based on a review on the literature about the measurement concepts in an ideal world, we discuss the advantages and shortcomings of tax data compared to other data sources - namely survey data.
-We describe and explain methods and techniques needed to construct time-series of income inequality measures for Switzerland.
-We will present these time-series in the result section and evaluate the use of the FTA Tax Data to assess the development of inequality in Switzerland
-
-% Nicht angesprochen sind in der Einleitung die Punkte 4 und 5 - Population coverage und intertemporal comparison. Der letzte Satz gefällt mir auch nicht so...
-
-\subsection{Tax statistics in Switzerland}
-
-Our data comes from the Swiss Federal Tax Administration (FTA).
-Federal taxes are collected and documented by the FTA since 1915.
-Being called a war-tax in the beginning, the federal tax was renamed to crisis levy in 1934, defense-tax in 1939 and is finally known as direct federal tax since 1983.
-The time frame we were able to collect ranges from 1945 to 2010 including 44 tax periods for cantons and Switzerland.
-\footnote{Between 1993 and 2003 there is no exact data available for aggregate Switzerland (but for individual cantons) because of a system change from taxation assessed in arrears (Praenumerando-System) to taxation assessed on current year income (Postnumerando-System) which was implemented by cantons in different years.}  While the FTA provides data in electronic form since 1973 we collected earlier data by scanning hard copies.
-Data is available for Switzerland plus all cantons and basically covers every tax unit (individual or household) in Switzerland liable to pay federal taxes.
-This exempts all tax units with taxable income below a certain threshold (e.g.
-15.000 CHF in 2010).
-Furthermore the FTA differentiates between two groups of tax units, so called normal cases and special cases.
-A normal case is a tax unit residing in a swiss canton without foreign source income and being liable to taxation all year long.
-All other tax units and very few that are taxed based on the style of living because they don't work (Pauschalbesteuerte) are special cases.
-\\
-
-Data is provided by the FTA in an aggregate form for privacy reasons, i.e.
-they are classified into numerous income brackets.
-While we don't know individual incomes we still have sufficient information (number of tax units per income bracket plus sum of incomes within each bracket) to calculate percentiles, Gini coefficients and other desired measures.
-However, equivalization (weighting income by household members) is not possible from these aggregate data.
-In any parts of the article where we point to equivalized income, the data stem from BRÜLHART_WIE_NENNEN_WIR_DAS.
-Their measures are calculated from the same source (FTA data) but used the number of children and marriage status as an approximation of household size to calculate inequality measures that are pseudo equivalized.
-\\
-
-The FTA provides two income measures: taxable income and net income.
-Net income here is an administrative term and means taxable income plus social deductions (children and supported persons) but not including other deductions like donations or health-care costs.
-Both measures are designed for taxation purposes which might limit the suitability to measure inequality as we will discuss later.
-
-% Verschiedene Punkte werden zweimal erwähnt (hier und in den folgenden Abschnitten). Man muss mal alles am Stück lesen und schauen, ob es so passt.
-
-% Verweis auf Brülhartdaten.
-
-\subsection{Standards for measuring economic resources and inequality}
-
-\emph{Concepts on measuring economic resources}  \\
-
-Most studies on inequality focus on income inequality solely.
-However, recent activities emphasize the need of a broader conceptualization.
-A recent publication from the \citet{oecd_oecd_2013} condense these ideas into the ICW framework (income, consumption and wealth), which is meant to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the expert group in developing the ICW Framework presented in this publication.
-Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_2004}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}}.
-According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
-To gain policy relevant insight, it is recommended to look at the distribution of all three distributions simultaneously.
-Some households with low income, for example, may report adequate levels of consumption expenditure or wealth holdings, or vice-versa.
-But it is also stated \citep[18]{oecd_oecd_2013}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being.
-\url{http://www.lisdatacenter.org/our-data/lws-database/}} \\
-
-This last statement holds for Switzerland too, although the HBS study is strongly influenced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
-Albeit the awareness of an assessment of income, consumption and wealth simultaneously is rising, we focus our analysis on income, which is undoubtedly a crucial indicator of economic well-being.
-It should be noted, that the Federal Tax Administration (FTA) publishes statistics on income and wealth but it is not possible to analyze the joint distribution on the individual or household level.
-Also measures of consumption are largely missing in tax data, albeit deductions can in some sense be understood as mandatory consumptions.
-
-\emph{Defining income}
-
-The assessment of income inequality is influenced by the definition of the income itself.
-Market income or disposable income for example differ by substantial meaning and by the expected degree of inequality.
-Therefore the awareness of the analyzed concept is crucial.
-Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013}, \citet[24]{united_nations_canberra_2011}).
-Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually lead to disposable income: the income measure, which finally shapes the possibility to consume.
-Within this framework common other income definitions are situated.
-These definitions are contrasted with income concepts, which can be derived out of the FTA tax data.
-
-\begin{figure}
-\centering
-\includegraphics[width=\maxwidth]{figure/income_definition}
-\caption{Income definitions}
-\label{fig:incdef}
-\end{figure}
-
-% Quelle zu Abzügen:
-% http://www.bilanz.ch/invest/steuern/netto-oder-reineinkommen
-% Die rechte Seite mit den Steuergrössen könnte ausführlicher beschrieben werden
-% Beispielsweise könnte man zeigen was Nettoeinkommen, Reineinkommen und steuerbares Einkommen sind
-% Allerdings steht uns nur Reineinkommen und steuerbares Einkommen zur Verfügung
-
-% ACHTUNG: Wir m?ssen innerhalb des Papers einheitlich bleiben. In der Grafik steht jetzt "Total income" oben steht "Net income"
-
-% Man k?nnte es auch konzeptionell auf kantonale Steuerdaten erweitern. Besser w?re es, dies einfach im Ausblick zu erw?hnen
-
-The central income reported through tax statistics is the taxable income.
-It includes all reported incomes (income from employment, income from property and received transfers\footnote{Means-tested benefits are not taxed and therefore not included in tax data.
-Income for low income groups are therefore underestimated.
-However, \citep{piketty_income_2003} note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions.
-It is therefore neither a pre-transfer income nor a post-transfer income measure.
-It's rather something in between.
-As the FTA tax statistics include some but not all deductions\footnote{The difference between the real total income and the taxable income are deductions.
-These include: professional expenses, travel expenses, interest on debt, alimonies, training costs, two-earner deduction, party contributions, private pension provision ``Säule 3a'', buying into the pension plan and sideline deductions} it is possible to calculate a sort of total income, which is called ``net inomce'' (Reineinkommen).
-As some deductions can be interpreted as compulsory expenses similar to taxes the step towards total income is a step away from the income, which can be used for consumption.
-Similar when calculating the disposable income out of the taxable income through accounting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
-Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches, which represent the bulk of taxes in Switzerland, are missing.
-\\
-%RF (16.06.2014): den letzten abschnitt finde ich etwas schwer zu verstehen als jemand der nicht tief in der materie ist.
-vielleicht könnte man sowas machen: It is possible to calculate a kind of "total income" that is in between net and gross income.
-The principle applied is to only deduct expenditures that are unavoidable.
-%OH (16.06.2014): bin mir nicht sicher, ob's mit gross income einfacher ist (kommt in der Grafik nicht vor)
-
-% Allenfalls darauf hinweisen, dass die Wahl der Einkommensdefinition von der Forschungsfrage abhängt.
-Inequality in market outcome (genderdifferences) requieres to look at market earings from employment, to assess the role of redistribution different steps of income has to be looked at.
-Following the idea of economic wellbeing usual disposable income is looked at
-
-\emph{Statistical units} \\
-
-The agreed standard on the statistical units, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013}.
-Indeed it is the individuals, who receive income, own assets and experience economic well-being, but their possibility to do so, is strongly tied to the concept of household.
-This comprises all persons under the same housing arrangement.
-The basic underlying assumption for collecting data on household level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children).
-Additionally, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing.
-To compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013}, \citealt{buhmann_equivalence_1988}).
-\\
-
-In tax data, however, the units are represented according to administrative rules.
-Tax units therefore neither represent individuals in every case nor true households.
-Tax units rather represent individuals and couples, but couples, who are married or officially registered.
-This doesn't imply, that those couples live together, as it is needed to satisfy the definition of a household.
-On the other hand, is it quite likely that more than one tax unit live in the same household (unmarried/unregistered couples, see \citet[99]{muller_vermogenslage_2014}).
-It is therefore not directly possible to elicit households and household income from tax data.
-This might influence the assessment of inequality development, taking into account the change from traditional household and family structures over the last century.
-
-% Allenfalls die Br?lhart-Daten erw?hnen und die m?glichkeit der Haushaltsbildung ?ber Abz?ge. Falls wir auch eine Grafik zur Rolle der Haushaltszusammensetzung machen (ich denke schon)
-% gibt es studien? Zur Bedeutung von ?quivalenzskalen und zur Bedeutung der Differenzierung von Haushalten und Individuuen?
-% inequality and family structure> review ny (McCall & Percheski (2010))
-% Was ist zu erwarten, wenn Äquivalenzskalen vernachlässigt werden? Soll man dies erwähnen oder einfach weglassen
-
-\emph{Measuring inequality or concentration} \\
-To be able to make qualifying statements about a distribution or to compare different distributions, the concept of inequality turned out to be the most appropriate and thus the most commonly used dimension.
-The Gini coefficient is the most known measure and mainly used for international comparison.
-As it is derived from the Lorenz-curve, the quantified amount of inequality can unpretentiously be described in a formal and visual way.
-Therefore the Gini coefficient is easily interpretable.
-Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}.
-(1) ``\textbf{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textbf{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
-doubling of all income) (3) ``\emph{weak principle of transfers}'' or ``\emph{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality.
-However, several drawbacks are reported in the literature.
-The most important point is, that the underlying distributional form of the measured inequality is unknown and it is therefore not possible to see if the measure is driven by a few rich or many poor individuals.
-This can also be problematic for comparison between countries or over time.
-In extreme cases two totally different distributions share the same Gini-coefficient.\\
-
-% namen mit principle ergänzen?
-
-The recent wave of tax-data studies do not report Gini-coefficents.
-Rather top income shares are informed on, which are calculated not only with tax data, but together with external sources to produce the population and personal income control totals (RF 16.06.2014: hä, was?
-gehts da um die Randverteilung der Bevölkerung?).
-This procedure ensures, that the inequality measure is not biased because of non-fillers, who do not appear in tax statistics.
-\citet{leigh_how_2007} compares top income shares with other inequality measures and asks, whether they are a useful measure of inequality in a society.
-He tries to answer this question empirically by comparing measures of inequality based on top income shares with measures of household or family inequality.
-He finds a strong positive relationship, but concludes (P.600): ``top income shares are far from perfect as a measure of distribution of income across society.'' Top income shares hence inform not completely on how inequality evolves elsewhere in the distribution.
-Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent as mentioned above).
-A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between individuals, who are either both within the top group or both outside the top group, then the share measure will remain unchanged.
-\\
-
-% Auf deutsch würde ich noch ebd.
-
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading).
-The polarizationindex is developed by \citep{Handock and Morris 1999} allows for this.
-Recently this index was applied in the work of \citep{alderson_exactly_2005} and \citep{gornick_how_2013}.
-This approach is rooted in relative distribution methods.
-It includes a precise comparison of the shape of two distributions (groups, over time).
-As the main advantage this approach allows to characterize the change in detail.
-It is possible to see which parts of the distribution changed, e.g.
-whether a polarization occurred (reduction of middle class) -- which equals an increase of inequality -- or this change is driven by either a change in the upper or lower part of the distribution.
-\\
-
-The literature mentions several other metrics with desired properties we will not discuss here (see for example \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}).
-
-% Allenfalls auch auf Quintilvergleiche eingehen und Atkinsonmass
-
-\emph{Population Coverage}
-% beim letzten Punkt geht es um Steuerhinterziehung. Fällt das auch noch unter Population coverage?
-
-Often inequality is assessed on national level, which implies, that studies try to cover the whole population of the country of interest.
-This is a special thorny task for surveys working with samples, because nonresponse is a major source of bias \citep{bethlehem_handbook_2011}.
-\citep{korinek_survey_2006} show, that the position in the income distribution influence the probability to participate in a survey.
-Low income and high income households are more likely to refuse survey response, which leads to an overrepresentation of middle income households.
-This process can be referred to as the "middleclass bias" \citep{diekmann_empirische_2009}.
-Missing data in household surveys is therefore not missing at random, which has an impact on the measures of inequality.
-The magnitude of this bias in Switzerland, however, is unknown.
-Strategies to handle this kind of bias are discussed in the literature \citep{sarndal_model_2003}, but require a register for every unit, that is proportional to income.
-Currently no such register exists for Switzerland \citet[43]{muller_vermogenslage_2014}.
-Currently used micro datasets, which are used for official publications concerning inequality in Switzerland (SILC and HABE) are furthermore confronted with a constructed coverage problem, because these surveys rely on the phone register, which excludes households not having a registered connection.
-\\
-%(Problems of collective households, ecoplan 2014:45). Erw?hnen?
-
-The issue of incomplete coverage is less dramatical with tax data.
-Essentially every permanent resident in Switzerland over 18 years of age (respectively 20 years of age prior to 1996) is taxed on a yearly base (or every two years before the change of the tax system).
-Essentially this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distribution.
-This includes a separation of normal cases, which embrace the majority of taxpayers, and the special cases, which cover (not only) foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
-Most important this includes high net wealth individuals taxed according to their expenditures.
-Special attention has to be paid to tax units with none or very low incomes.
-Even though they have to hand in a tax return, their income does not show up in the statistics if their income after deductions falls below 15'000 CHF and they are therefore not taxed with direct federal taxes.
-This is possible for normal and special cases alike.
-From 1995/1996 until 2010 the number of non-taxed units is reported, but not for the years before.
-Dell et al.
-(2007) try to estimate the fraction of non-taxed by comparing the reported numbers of tax units to census reports about the number of adult population.
-According to their estimations this fraction drops from 94\% in 1993/1994 to 63\% back in 1945/46.
-\\
-
-Another critical issue with tax data is the problem of tax evasion, which definitely can bias the assessment of inequality.
-\citet{alvaredo_income_2009} for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
-Evasion can occur, when individuals try not to fill tax returns or by misreporting of incomes.
-In Switzerland non-fillers show up in the tax-statistics either way, as long as they are registered.
-This person gets an imputed income based on an older tax return and information given by employers.
-Only non registered non-fillers are not in the records.
-Therefore non-fillers are a minor problem.
-Not negligible is the circumstance, that individuals misreport incomes.
-\citet{feld_tax_2006} examine the role of tax evasion in Switzerland by calculating the difference of the national accounts measures of primary income and the income reported to the tax authorities.
-They can show, that the average level of income tax evasion from 1965 to 1995 varies between 13\% and 35\%.
-They suggest, that evasion is heavily driven by capital income tax evasion.
-\\
-
-\emph{Intertemporal comparison}
-
-One advantage of tax data over survey data is their observation period.
-While the FTA data we use range back to 1945, the earliest period of survey data is 2000 (HABE data).
-To be fair, survey data might cover the late periods (2011 and 2012) which are not covered by tax data.
-The striking feature of the long FTA time series is its consistency.
-Both population coverage and income measures are consistent for the complete observation period.
-This can not be said for survey data as these are based on samples and therefore require an ideal sampling design or reweighting to be representative and comparable over time.
-We will see how well sampling and weighting is done in the results section.
-
-\subsection{Comparison of tax data and other data sources - advantages and shortcomings}
-
-% Suitability of data sources can be judeged by different criterias. crossectional or longidutional perspective. longidutional perspective allows for dynamic
-
-%[ich stelle mir eine Tabelle mit verschiedenen Dimensionen vor, anhand derer eine Einordnung unterschiedlicher Datenquellen geschieht.]
-% Mögliche Vergleichsdimensionen
-%- Suitability of cross country comparison
-%- Coverage of the population of interesst
-%- Implementation of common concepts of economic ressources
-%- intertemporal comparision
-%- flexibel applications of appropriate inequality measures
-% - appropriate income definition
-
-\begin{table}[h]
-\begin{tabular}{lcccc}
-                                   & FTA       & HABE      & EU-SILC   & LIS       \\
-Range                              & 1945-2010 & 2000-2011 & 2007-2012 & 1982-2004 \\
-Cross country comparison possible? & no        & no        & yes       & yes       \\
-Top incomes covered?               & yes       & no        & no        & no        \\
-Low incomes covered?               & no        & yes       & yes       & yes       \\
-Concepts of economic ressources?   & no        & yes       & ?         & ?         \\
-Equivalisation?                    & no        & yes       & ?         &          
-\end{tabular}
-\end{table}
-
-% Braucht es alle drei Survey-Befragungen
-% ich würde nicht dichotomisieren, sondern ordinal bewerten (tief, mittel, hoch)
-
-% Die Punkte kurz beschreiben.
-% Zusammenfassung
-
-% Gedankenstütze
-%\textbf{Problems with household income surveys}
-%\begin{itemize}
-%\item Sample data (bias)
-%\item comparability between countries and over time (depends on income definition)
-%\item short time series
-%\end{itemize}
-
-%Burkhauser et al. (2009) compare inequality statistics form survey data and tax records to consolidate the findings about recent trends in the USA
-%Studies on problems with survey data See for other countries: \citealt{Siminski et al 2003 FEHLT} (Australia), \citealt{Brewer et al 2008 FEHLT}, UK, \citealt{Burkhauser et al. 2009 FEHLT}, US)
-
-%Warnings about the use of tax data
-%The use of tax data is often regarded by economists with considerable disbelief.
-These doubts are well justified for at least two reasons.
-The first is that tax data are collected as part of an administrative process, which is not tailored to the scientists' needs, so that the definition of income, income unit, etc., are not necessarily those that we would have chosen.
-This causes particular difficulties for comparisons across countries, but also for time-series analysis where there have been substantial changes in the tax system, such as the moves to and from the joint taxation of couples.
-Secondly, it is obvious that those paying tax have a financial incentive to present their affairs in a way that reduces tax liabilities.
-There is tax avoidance and tax evasion.
-The rich, in particular, have a strong incentive to understate their taxable incomes.
-Those with wealth take steps to ensure that the return comes in the form of asset appreciation, typically taxed at lower rates or not at all.
-Those with high salaries seek to ensure that part of their remuneration comes in forms, such as fringe benefits or stock-options which receive favorable tax treatment.
-Both groups may make use of tax havens that allow income to be moved beyond the reach of the national tax net.
-These shortcomings limit what can be said from tax data, but this does not mean that the data are worthless.
-Like all economic data, they measure with error the 'true' variable in which we are interested.
-%The data series presented here are fairly homogenous across countries, annual, long-run, and broken down by income source for several cases.
-Users should be aware also about their limitations.
-Firstly, the series measure only top income shares and hence are silent on how inequality evolves elsewhere in the distribution [why?].
-Secondly, the series are largely concerned with gross incomes before tax.
-Thirdly, the definition of income and the unit of observation (the individual vs.
-the family) vary across countries making comparability of levels across countries more difficult.
-Even within a country, there are breaks in comparability that arise because of changes in tax legislation affecting the definition of income, although most studies try to correct for such changes to create homogenous series.
-Finally and perhaps most important, the series might be biased because of tax avoidance and tax evasion.
-For the details, we refer users to the original papers (see also Atkinson, Piketty and Saez, 2011).
-
 \subsection{Outline of applied methods}

### 2014-06-19 code/simulate_data.R

@@ -42,36 +42,40 @@ init<-0
 fin<-(abs(max(normal$inc,na.rm=TRUE))+5)
 # generate the sequence to take pairs from
 ival<-c(init,normal$inc,fin)
-len<-100 # sequence of each pair
+len<-10000 # sequence of each pair
 s<-sapply(2:length(ival),function(i){
-  seq(ival[i-1],ival[1],length.out=len)
+  seq(ival[i-1],ival[i],length.out=len)
 })
 s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
 normal.ind<-data.frame(s)
 names(normal.ind)[1]=c("inc")
 summary(normal.ind)
 hist(normal.ind$inc)
-normal.ind$inc<-normal.ind$inc*7 # Rescale income 
-hist(normal.ind$inc)
+#normal.ind$inc<-normal.ind$inc*7 # Rescale income 
+#hist(normal.ind$inc)
+
+
+##
+# Interpolate CDF
 
 ##
 # Prepare HABE Data - 2010
 # ATTENTION: not weighted and with disposable income
+#habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+#habe10<-habe0911 %.% 
+ # filter(Jahr08==2010)
+#habe10<-data.frame(habe10$VerfuegbaresEinkommen08)
+#names(habe10)<-"inc_habe"
+#habe10$inc_habe[habe10$inc_habe<0]<-0 # One values is -20'000
+#habe10$inc_habe<-habe10$inc_habe*13/1000 # transform from monthly income to income per year
+
+# Alternative mit Bruttoeinkommen
 habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
 habe10<-habe0911 %.% 
   filter(Jahr08==2010)
-habe10<-data.frame(habe10$VerfuegbaresEinkommen08)
+habe10<-data.frame(habe10$Bruttoeinkommen08)
 names(habe10)<-"inc_habe"
-habe10$inc_habe[habe10$inc_habe<0]<-0 # One values is -20'000
[…]

### 2014-06-19 code/simulate_data.R
updated relative distribution comparison with narrowd comparison based on brülhart-data

@@ -8,6 +8,7 @@
 library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
 library(ggplot2)
 library(reldist)
+library(foreign)
 
 #######################
 #     2010            #
@@ -56,9 +57,6 @@ hist(normal.ind$inc)
 
 
 ##
-# Interpolate CDF
-
-##
 # Prepare HABE Data - 2010
 # ATTENTION: not weighted and with disposable income
 #habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
@@ -75,7 +73,7 @@ habe10<-habe0911 %.%
   filter(Jahr08==2010)
 habe10<-data.frame(habe10$Bruttoeinkommen08)
 names(habe10)<-"inc_habe"
-habe10$inc_habe<-habe10$inc_habe*13/1000 # transform from monthly income to income per year
+habe10$inc_habe<-habe10$inc_habe*12/1000 # transform from monthly income to income per year
 
 ##
 # Apply Relative Distribution Methods
@@ -95,11 +93,11 @@ plot(x = (density1$x), y = density1$y, type = "l",
      xlab = "Einkommen", ylab = "density",
      axes = FALSE,
      xlim = c(-70, 900),
-     ylim=c(0,1.626e-02))
+     ylim=c(0,1.559e-02))
 title(main="FTA vs HBS income distribution",cex=0.6)
 axis(side = 1)
 axis(side = 2)
-fig1legend <- list(x=c(400,400),y=c(1.626e-02,1.626e-02))
+fig1legend <- list(x=c(400,400),y=c(1.559e-02,1.559e-02))
 legend(fig1legend,lty=1:2,cex=0.5, bty="n",
        legend=c("HBS","FTA"))
 density2 <- density(normal.ind$inc)
@@ -121,7 +119,7 @@ title(main="Relative CDF",cex=0.6)
 fig2b <- reldist(y=normal.ind$inc,yo=habe10$inc_habe,
                  ci=FALSE,smooth=0.4,
                  yolabs=seq(-1,3,by=0.5),
-                 ylim=c(0,2.5),cex=0.8,
+                 ylim=c(0,5),cex=0.8,
                  ylab="Relative Density",
                  xlab="Proportion of HBS Distribution")
[…]

### 2014-06-19 code/simulate_data.R
update simulate_data.R

@@ -65,7 +65,7 @@ hist(normal.ind$inc)
 #habe10<-data.frame(habe10$VerfuegbaresEinkommen08)
 #names(habe10)<-"inc_habe"
 #habe10$inc_habe[habe10$inc_habe<0]<-0 # One values is -20'000
-#habe10$inc_habe<-habe10$inc_habe*13/1000 # transform from monthly income to income per year
+#habe10$inc_habe<-habe10$inc_habe*12/1000 # transform from monthly income to income per year
 
 # Alternative mit Bruttoeinkommen
 habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
@@ -389,21 +389,24 @@ title(main=paste("Overall realtive density",cex=0.6)
                   
 
 
-#####
-# Comparing HBS-data to FTA-data, we see that FTA-Data has a hugh middle-class bias
+###########################################################################################
+# Comparing HBS-data to FTA-data, we see that FTA-Data exceeds HBS-Data in the middle of the distribution
 # We think it's because we compare tax units to households
 # To find out about that, we try to focus the comparision on a) singles and b) married to see
+###########################################################################################
+
 
 # Get Brülhartdata for 2010
 estv2010 <- read.csv("data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
 
-# Select percentiles and all cases(total)
+
+###########################################
+# Select percentiles for all cases (total)
 
 estv2010.total<-estv2010[c(18:38),c(1,4)]
 
 ##
 # Simulate individual data points
-# Percentile-Plot (descriptive  purpose)
 # cumulative probabilites
 cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
 # probabilities
@@ -425,38 +428,320 @@ names(estv2010.total.ind)[1]=c("inc")
 summary(estv2010.total.ind)
 hist(estv2010.total.ind$inc)
[…]

### 2014-06-20 paper/subdocuments/results.Rnw
minor changes plot options

@@
 levels(dflong$Type)<-c("taxable income","income before deductions","taxable income after tax")
-ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line(size=1)+theme_bw()+xlab("tax period")
-ggplot(dflong, aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")
+ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,shape=Type))+
+  scale_y_continuous(limits=c(0.18,0.4))+
+  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
+  geom_point(aes(shape=Type),size=3)+
+  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Type))+
+  theme_bw()+
+  xlab("tax period")
+#ggplot(dflong, aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")
 @
@@
 \emph{Gini coefficients considering mean-tested benefits}
+
+% Wollen wir das ?berhaupt machen?
 
@@
 normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
-normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
+#normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
 pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")

### 2014-06-20 paper/subdocuments/results.Rnw
updated Brülhart data, plot net income vs equivalised net income

@@
 
-% Nicht eher nur CH-Plot?
+\subsection{Statistical units}
+
+<<setup_data>>=
+bd <- read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";")
+bd.long<-reshape(bd,
+varying = c("gini_reink","gini_reinka"),
+v.names = "Gini",
+timevar = "inc_def",
+times = c("Gini net income","Gini equivaliced net income"),
+direction ="long")
+@
+
+<<With and without equivalence scale>>=
+library(ggplot2)
+ggplot(db[db$Einheit=="Total",], aes(x=Veranlagungsperiode,y=Gini,shape=Type))+
+  scale_y_continuous(limits=c(0.18,0.4))+
+  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
+  geom_point(aes(shape=Type),size=3)+
+  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Type))+
+  theme_bw()+
+  xlab("tax period")
+
+@
+
+\subsection{Measuring inequality}
 

### 2014-06-20 paper/subdocuments/results.Rnw
integrated HABE comparison

--- 2014-06-20 paper/subdocuments/results.Rnw
@@
 library(reshape)
+number_ticks <- function(n) {function(limits) pretty(limits, n)}
 dflong <- melt(df[,c("G_steink","G_reink","G_taxed","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
@@
 
-% Nicht eher nur CH-Plot?
+\subsection{Statistical units}
+
+<<setup_data>>=
+library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
+# Ohne Nuller
+bd <- read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";")
+#Veranlagungsperiode -> Jahre
+bd$Jahr <- as.numeric(substr(bd$Veranlagungsperiode,1,4))
+bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] <- bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] - 2
+bd<-bd %.% 
+  filter(Einheit=="Total")
+bd<-subset(bd,Jahr<1995 | Jahr >2002)
+bd.long<-reshape(bd,
+varying = c("gini_reink","gini_reinka"),
+v.names = "Gini",
+timevar = "inc_def",
+times = c("Net income (only taxed)","Equivalised net income (only taxed)"),
+direction ="long")
+# mit Nuller
+bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
+#Veranlagungsperiode -> Jahre
+bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
+bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
+bd.0<-bd.0 %.% 
+  filter(Einheit=="Total")
+bd.0<-subset(bd.0,Jahr<1995 | Jahr >2002)
+bd.0.long<-reshape(bd.0,
+varying = c("gini_reink","gini_reinka"),
+v.names = "Gini",
+timevar = "inc_def",
+times = c("Net income","Equivalised net income"),
+direction ="long")
+bdlong<-rbind(bd.0.long,bd.long)
+@
+
+<<With and without equivalence scale>>=
+library(ggplot2)
+number_ticks <- function(n) {function(limits) pretty(limits, n)}
+ggplot(bdlong, aes(x=Jahr,y=Gini,shape=inc_def))+
+  scale_y_continuous(limits=c(0.18,0.6))+
+  scale_x_continuous(limits=c(1940,2012),breaks=number_ticks(10)) +
+  geom_point(aes(shape=inc_def),size=3)+
+  geom_line(data=bdlong,aes(linetype=inc_def))+
+  theme_bw()+
+  xlab("tax period")
+@
+
+\subsection{Measuring inequality}
 
@@
 
+<<setup_data>>=
+library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
+library(ggplot2)
+library(reldist)
+library(foreign)
+# Get Brülhartdata for 2010
+estv2010 <- read.csv("data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
+###########################################
+# Select percentiles for all cases (total)
+estv2010.total<-estv2010[c(18:38),c(1,4)]
+##
+# Simulate individual data points
+# cumulative probabilites
+cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
+# probabilities
+prob<-c(cum.p[1],diff(cum.p), .01)
+# extreme values beyond x (to sample)
+#freq<-max(normal$anz_pflichtige)
+freq<-3000
+init<-0
+fin<-(abs(max(estv2010.total$Total,na.rm=TRUE))+5)
+# generate the sequence to take pairs from
+ival<-c(init,estv2010.total$Total,fin)
+len<-10000 # sequence of each pair
+s<-sapply(2:length(ival),function(i){
+  seq(ival[i-1],ival[i],length.out=len)
+})
+s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
+estv2010.total.ind<-data.frame(s)
+names(estv2010.total.ind)[1]=c("inc")
+estv2010.total.ind$inc<-estv2010.total.ind$inc/1000
+
+# HABE - 2010 all cases
+
+habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+habe10<-habe0911 %.% 
+  filter(Jahr08==2010)
+habe10<-data.frame(habe10$VerfuegbaresEinkommen08)
+names(habe10)<-"inc_habe"
+habe10$inc_habe[habe10$inc_habe<0]<-0 # One values is -20'000
+habe10$inc_habe<-habe10$inc_habe*12/1000 # transform from monthly income to income per year
+
+# Probability density Function
+
+density1 <- density(habe10$inc_habe)
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "Einkommen", ylab = "density",
+     axes = FALSE,
+     xlim = c(-70, 4600),
+     ylim=c(0,1.103e-02))
+title(main="FTA vs HBS income distribution",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(2500,2500),y=c(1.103e-02,1.103e-02))
+legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+       legend=c("HBS","FTA"))
+density2 <- density(estv2010.total.ind$inc)
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+# Relative Distribution
+
+par(mfrow=c(1,3))
+g10 <- reldist(y=estv2010.total.ind$inc, yo=habe10$inc_habe,
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               ylim=c(0.5,5.0),
+               bar=TRUE, quiet=FALSE,
+               xlab="Proportion of HBS Distribution")
+title(main=paste("Overall realtive density"))
+abline(h=1,lty=2)
+g1A <- reldist(y=estv2010.total.ind$inc, yo=habe10$inc_habe,
+               show="effect",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of HBS Distribution")
+title(main=paste("Effect of Median shift"))
+abline(h=1,lty=2)
+gA0 <- reldist(y=estv2010.total.ind$inc, yo=habe10$inc_habe,
+               smooth=0.4, ci=FALSE,
+               show="residual",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of HBS Distribution")
+title(main=paste("Effect of different shape"))
+abline(h=1,lty=2)
+par(mfrow=c(1,1))
+
+########################
+# Separate comparison for unmarried
+
+# Select married - ESTV
+
+estv2010.married<-estv2010[c(18:38),c(1,2)]
+names(estv2010.married)[2]=c("married")
+##
+# Simulate individual data points
+# Percentile-Plot (descriptive  purpose)
+# cumulative probabilites
+cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
+# probabilities
+prob<-c(cum.p[1],diff(cum.p), .01)
+# extreme values beyond x (to sample)
+#freq<-max(normal$anz_pflichtige)
+freq<-3000
+init<-0
+fin<-(abs(max(estv2010.married$married,na.rm=TRUE))+5)
+# generate the sequence to take pairs from
+ival<-c(init,estv2010.married$married,fin)
+len<-10000 # sequence of each pair
+s<-sapply(2:length(ival),function(i){
+  seq(ival[i-1],ival[i],length.out=len)
+})
+s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
+estv2010.married.ind<-data.frame(s)
+names(estv2010.married.ind)[1]=c("inc")
+estv2010.married.ind$inc<-estv2010.married.ind$inc/1000
+
+# Select married - HABE
+
+habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+habe10<-habe0911 %.% 
+  filter(Jahr08==2010)
+habe10.married<-habe0911 %.% 
+  filter(Familientyp08>130,Jahr08==2010)
+habe10.married<-data.frame(habe10.married$VerfuegbaresEinkommen08)
+names(habe10.married)<-"inc_habe"
+habe10.married$inc_habe[habe10.married$inc_habe<0]<-0 # One values is -20'000
+habe10.married$inc_habe<-habe10.married$inc_habe*12/1000 # transform from monthly income to income per year
+
+# Probability Density Function
+
+density1 <- density(habe10.married$inc_habe)
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "Einkommen", ylab = "density",
+     axes = FALSE,
+     xlim = c(-257, 8000),
+     ylim=c(0,0.0101390))
+title(main="FTA vs HBS income distribution",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(2500,2500),y=c(0.0101390,0.0101390))
+legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+       legend=c("HBS","FTA"))
+density2 <- density(estv2010.married.ind$inc)
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+# Relative Distribution
+
+par(mfrow=c(1,3))
+g10 <- reldist(y=estv2010.married.ind$inc, yo=habe10.married$inc_habe,
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               ylim=c(0.5,5.0),
+               bar=TRUE, quiet=FALSE,
+               xlab="Proportion of HBS Distribution")
+title(main=paste("Overall realtive density"))
+abline(h=1,lty=2)
+g1A <- reldist(y=estv2010.married.ind$inc, yo=habe10.married$inc_habe,
+               show="effect",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of HBS Distribution")
+title(main=paste("Effect of Median shift"))
+abline(h=1,lty=2)
+gA0 <- reldist(y=estv2010.married.ind$inc, yo=habe10.married$inc_habe,
+               smooth=0.4, ci=FALSE,
+               show="residual",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of HBS Distribution")
+title(main=paste("Effect of different shape"))
+abline(h=1,lty=2)
+par(mfrow=c(1,1))
+
+#################
+# Separate comparison for unmaried
+
+##
+# Unmarried ESTV
+
+estv2010.unmarried<-estv2010[c(18:38),c(1,3)]
+names(estv2010.unmarried)[2]=c("unmarried")
+##
+# Simulate individual data points
+# Percentile-Plot (descriptive  purpose)
+# cumulative probabilites
+cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
+# probabilities
+prob<-c(cum.p[1],diff(cum.p), .01)
+# extreme values beyond x (to sample)
+#freq<-max(normal$anz_pflichtige)
+freq<-3000
+init<-0
+fin<-(abs(max(estv2010.unmarried$unmarried,na.rm=TRUE))+5)
+# generate the sequence to take pairs from
+ival<-c(init,estv2010.unmarried$unmarried,fin)
+len<-10000 # sequence of each pair
+s<-sapply(2:length(ival),function(i){
+  seq(ival[i-1],ival[i],length.out=len)
+})
+s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
+estv2010.unmarried.ind<-data.frame(s)
+names(estv2010.unmarried.ind)[1]=c("inc")
+estv2010.unmarried.ind$inc<-estv2010.unmarried.ind$inc/1000
+
+# Select unmarried - HABE
+
+habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+habe10<-habe0911 %.% 
+  filter(Jahr08==2010)
+habe10.unmarried<-habe0911 %.% 
+  filter(Familientyp08<140,Jahr08==2010)
+habe10.unmarried<-data.frame(habe10.unmarried$VerfuegbaresEinkommen08)
+names(habe10.unmarried)<-"inc_habe"
+habe10.unmarried$inc_habe[habe10.unmarried$inc_habe<0]<-0 # One values is -20'000
+habe10.unmarried$inc_habe<-habe10.unmarried$inc_habe*12/1000 # transform from monthly income to income per year
+
+# Probability Density Function
+
+density1 <- density(habe10.unmarried$inc_habe)
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "Einkommen", ylab = "density",
+     axes = FALSE,
+     xlim = c(-50, 2500),
+     ylim=c(0,1.764e-02))
+title(main="FTA vs HBS income distribution",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(1200,1200),y=c(1.764e-02,1.764e-02))
+legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+       legend=c("HBS","FTA"))
+density2 <- density(estv2010.unmarried.ind$inc)
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+# Relative Distribution
+
+par(mfrow=c(1,3))
+g10 <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               ylim=c(0.5,5.0),
+               bar=TRUE, quiet=FALSE,
+               xlab="Proportion of HBS Distribution")
+title(main=paste("Overall realtive density"))
+abline(h=1,lty=2)
+g1A <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
+               show="effect",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of HBS Distribution")
+title(main=paste("Effect of Median shift"))
+abline(h=1,lty=2)
+gA0 <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
+               smooth=0.4, ci=FALSE,
+               show="residual",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of HBS Distribution")
+title(main=paste("Effect of different shape"))
+abline(h=1,lty=2)
+par(mfrow=c(1,1))
+@
+
 \subsection{Intertemporal comparison}

### 2014-06-20 paper/subdocuments/results.Rnw
no weighting remark
included main analysis for subchapter "The measurement issue"

@@
 
-<<setup_data>>=
+<<setup_data_Brühlhart with and without taxed>>=
 library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
@@
 
+% Achtung: Das kann doch nicht stimmen, dass wir hier eine Reihe ab 1993 vorliegen haben.
+% Die ESTV-Zahlen beginnen ab 1995/1996... Ist das jetzt eine Periode, wo man noch Gesamtschweizerische Reihen rekonstruieren kann?
+
 <<With and without equivalence scale>>=
@@
 \subsection{Measuring inequality}
+
+% Es stellt sich die Frage für welchen Zeitraum der Vergleich gemacht wird.
+% Ich schlage vor, der Vergleich wird mit den Brülhartdaten gemacht und der Vergleich soll von 2003 bis 2010 reichen
+
+<<setup_data_Brülhart_2003>>=
+library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
+library(foreign)
+# Ohne Nuller
+bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
+#Veranlagungsperiode -> Jahre
+bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
+bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
+bd.2003<-bd.0 %.% 
+  filter(Jahr==2003,Einheit=="Total")
+start <- which(names(bd.2003)=="p1")
+end <- which(names(bd.2003)=="p99_99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+bd.2003<- data.frame(t(bd.2003[1,start:end]),percentile)
+names(bd.2003)[1]=c("inc")
+cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
+prob<-c(cum.p[1],diff(cum.p), .01)
+freq<-3000
+init<-0 
+bd.2003$inc<-as.numeric(as.character(bd.2003$inc))
+fin<-(abs(max(bd.2003$inc,na.rm=TRUE))+5)
+ival<-c(init,bd.2003$inc,fin)
+len<-10000
+s<-sapply(2:length(ival),function(i){
+  seq(ival[i-1],ival[i],length.out=len)
+})
+s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
+bd.2003.ind<-data.frame(s)
+names(bd.2003.ind)[1]=c("inc")
+bd.2003.ind$inc<-bd.2003.ind$inc/1000
+@
+
+<<setup_data_Brülhart_2010>>=
+library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
+library(foreign)
+# Ohne Nuller
+bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
+#Veranlagungsperiode -> Jahre
+bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
+bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
+bd.2010<-bd.0 %.% 
+  filter(Jahr==2010,Einheit=="Total")
+start <- which(names(bd.2010)=="p1")
+end <- which(names(bd.2010)=="p99_99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
+names(bd.2010)[1]=c("inc")
+cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
+prob<-c(cum.p[1],diff(cum.p), .01)
+freq<-3000
+init<-0 
+bd.2010$inc<-as.numeric(as.character(bd.2010$inc))
+fin<-(abs(max(bd.2010$inc,na.rm=TRUE))+5)
+ival<-c(init,bd.2010$inc,fin)
+len<-10000
+s<-sapply(2:length(ival),function(i){
+  seq(ival[i-1],ival[i],length.out=len)
+})
+s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
+bd.2010.ind<-data.frame(s)
+names(bd.2010.ind)[1]=c("inc")
+bd.2010.ind$inc<-bd.2010.ind$inc/1000
+@
+
+<<Comparing 2002 to 2010>>=
+library(reldist)
+
+# Probability Density Function
+
+density1 <- density(bd.2003.ind$inc)
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "taxable income", ylab = "density",
+     axes = FALSE,
+     xlim = c(-50, 5000),
+     ylim=c(0,0.0116031))
+title(main="2003 vs 2010",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(1200,1200),y=c(0.0116031,0.0116031))
+legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+       legend=c("2003","2010"))
+density2 <- density(bd.2010.ind$inc)
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+# Relative Distribution
+
+par(mfrow=c(1,3))
+g10 <- reldist(y=bd.2010.ind$inc, yo=bd.2003.ind$inc,
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               ylim=c(0.5,5.0),
+               bar=TRUE, quiet=FALSE,
+               xlab="Proportion of 2002 Distribution")
+title(main=paste("Overall realtive density"))
+abline(h=1,lty=2)
+g1A <- reldist(y=bd.2010.ind$inc, yo=bd.2003.ind$inc,
+               show="effect",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of 2002 Distribution")
+title(main=paste("Effect of Median shift"))
+abline(h=1,lty=2)
+gA0 <- reldist(y=bd.2010.ind$inc, yo=bd.2003.ind$inc,
+               smooth=0.4, ci=FALSE,
+               show="residual",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of 2002 Distribution")
+title(main=paste("Effect of different shape"))
+abline(h=1,lty=2)
+par(mfrow=c(1,1))
+@
+
+% Report Polarization index
 
@@
 
+% Auf neuste Zahlen anpassen
+
 As we can see from figure~\ref{fig:specialcases93942} and \ref{fig:specialcases092}, special cases differ strongly 
@@
 ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))
-
 @
@@
 # HABE - 2010 all cases
+
+#####################################################
+#         ACHTUNG: Bisher ohne Gewichte!            #
+#####################################################
 

### 2014-06-23 paper/subdocuments/results.Rnw
new figures for normal/special and minor fixes

@@
                xlab="Proportion of 2002 Distribution")
-title(main=paste("Overall realtive density"))
+title(main=paste("Overall relative density"))
 abline(h=1,lty=2)
@@
 The FTA stopped to publicly report data for special cases after tax period 1993/94. 
-For more recent periods however we can compare data from the 
-BRUELHARTPROJEKT (wie nennen wir das?) to the FTA normal cases, as these data are 
-identical but the BRÜLHART data include special cases. 
+For more recent periods however we can compare already calculated percentiles provided by the FTA which are based on both special and normal cases to normal cases alone.
 We will have a look at both, the 1993/94 period as the last period where both 
-numbers were reported as well as 2009 where we compare data from the same source 
-but using differnt data sheets.
+numbers were reported as well as 2010 where we compare data from the same source (FTA)
+but based on different reporting (aggregated data vs. percentiles).
 
@@
 normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
+normal <- select(normal, -G_reink,-G_taxed,-ppop0,-G_steink0)
 #normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
@@
 #1- (2762419 / 3268548)
+
+# imputing pseudo-individual data to use reldist
+# we cut at the 5th and 95th percentile because it is unclear what values to use to extrapolate
+ival<-na.omit(normal$inc[-1])
+perval<-normal$percentile[-1]
+normal_i<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+ival<-na.omit(pooled$inc)
+perval<-pooled$percentile[-1]
+pooled_i<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+# Probability Density Function
+
+density1 <- density(normal_i)
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "taxable income", ylab = "density"
+     #axes = FALSE,
+     #xlim = c(-50, 200),
+     #ylim=c(0,0.0116031)
+     )
+title(main="normal cases vs pooled, normal and special cases 1993/94",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(100,100),y=c(0.0116031,0.0116031))
+legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+       legend=c("normal cases","pooled, normal and special cases"))
+density2 <- density(pooled_i)
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+# Relative Distribution
+
+par(mfrow=c(1,3))
+reldist(y=pooled_i, yo=normal_i,
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               ylim=c(0.5,5.0),
+               bar=TRUE, quiet=FALSE,
+               xlab="Proportion of Normal-Case-Distribution")
+title(main=paste("Overall relative density 1993/94"))
+abline(h=1,lty=2)
+reldist(y=pooled_i, yo=normal_i,
+               show="effect",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of Normal-Case-Distribution")
+title(main=paste("Effect of Median shift"))
+abline(h=1,lty=2)
+reldist(y=pooled_i, yo=normal_i,
+               smooth=0.4, ci=FALSE,
+               show="residual",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of Normal-Case-Distribution")
+title(main=paste("Effect of different shape"))
+abline(h=1,lty=2)
+par(mfrow=c(1,1))
 @
 
-<<specialcases09, echo=FALSE,warning=FALSE>>=
+1993/94 including special cases changes the distribution slightly at both ends.
+Special cases have a slightly lower median income and therefore are more frequent in the lower percentiles and less frequent in the upper percentiles.
+In addition there is an effect of shape: the distribution of special cases is more skewed so that lower percentiles show lower incomes and higher percentiles show higher incomes compared to just normal cases.
+
+<<specialcases10, echo=FALSE,warning=FALSE>>=
 library(foreign)
@@
 #library(reldist)
-normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2009)
+normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2010)
 start <- which(names(normal)=="p1")
@@
 #bruelhart data
-pooled <- filter(read.dta("../../data/data_Schweiz.dta"), Veranlagungsperiode==2009, Einheit=="Total")
+pooled <- filter(read.csv("../../data/data_Schweiz.csv",sep=";"), Veranlagungsperiode==2010, Einheit=="Total")
 start <- which(names(pooled)=="p1")
@@
 	xlab("Percentile")+ylab("taxable income")+
-	ggtitle("Taxable incomes of normal and special cases 2009")+
+	ggtitle("Taxable incomes of normal and special cases 2010")+
   theme(legend.justification=c(0,1),legend.position=c(0, 1))
@@
 	xlab("Percentile")+ylab("taxable income")+
-	ggtitle("Taxable incomes of normal versus special cases 2009")
+	ggtitle("Taxable incomes of normal versus special cases 2010")
 
@@
 #1- (3422210 / 3689063)
+
+# imputing pseudo-individual data to use reldist
+# we cut at the 95th percentile because it is unclear what values to use to extrapolate
+ival<-na.omit(c(0,normal$inc))
+perval<-c(0,normal$percentile)
+normal_i<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+#nur bis 95
+ival<-na.omit(c(0,pooled$inc[-15]))
+perval<-c(0,pooled$percentile)
+pooled_i<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+# Probability Density Function
+
+density1 <- density(normal_i)
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "taxable income", ylab = "density"
+     #axes = FALSE,
+     #xlim = c(-50, 200),
+     #ylim=c(0,0.0116031)
+     )
+title(main="normal cases vs pooled, normal and special cases 2010",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(120,120),y=c(0.0116031,0.0116031))
+legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+       legend=c("normal cases","pooled, normal and special cases"))
+density2 <- density(pooled_i)
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+# Relative Distribution
+
+par(mfrow=c(1,3))
+reldist(y=pooled_i, yo=normal_i,
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               ylim=c(0.5,5.0),
+               bar=TRUE, quiet=FALSE,
+               xlab="Proportion of Normal-Case-Distribution")
+title(main=paste("Overall relative density 2010"))
+abline(h=1,lty=2)
+reldist(y=pooled_i, yo=normal_i,
+               show="effect",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               smooth=0.4, ci=FALSE,
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of Normal-Case-Distribution")
+title(main=paste("Effect of Median shift"))
+abline(h=1,lty=2)
+reldist(y=pooled_i, yo=normal_i,
+               smooth=0.4, ci=FALSE,
+               show="residual",
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,5.0), ylab="",
+               yolabs=seq(-1,3,by=0.5),
+               xlab="Proportion of Normal-Case-Distribution")
+title(main=paste("Effect of different shape"))
+abline(h=1,lty=2)
+par(mfrow=c(1,1))
 @
+
+2010 the picture is similar but more apparent: There is a small effect of median shift lowering incomes in the lowest percentiles. The effect of shape however contributes most.
 
@@
 
-As we can see from figure~\ref{fig:specialcases93942} and \ref{fig:specialcases092}, special cases differ strongly 
-from normal cases within the low and top percentiles. Within the tax period 1993/94 the fifth percentile of 
-normal cases is 15.8\% higher than the fifth percentile of the combines data while the 95\% percentile of 
-normal cases is even 0.7\% lower if one leaves out special cases. This indicates special cases are very 
-different from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
-million tax units defined as being normal cases compared to 0.51 million special cases.}
-In 2009 the situation has the same pattern. The fifth percentile of normal cases has 11\% higher taxable 
-income compared to data where special cases are included. The five percent top incomes are 4.9\% higher 
-within all cases compared to normal cases only. The share of special cases however decreased 
-to 7.2\%.\footnote{2009: 3.42 million normal cases, 0.27 million special cases}
+%As we can see from figure~\ref{fig:specialcases93942} and \ref{fig:specialcases092}, special cases differ strongly 
+%from normal cases within the low and top percentiles. Within the tax period 1993/94 the fifth percentile of 
+%normal cases is 15.8\% higher than the fifth percentile of the combines data while the 95\% percentile of 
+%normal cases is even 0.7\% lower if one leaves out special cases. This indicates special cases are very 
+%different from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
+%million tax units defined as being normal cases compared to 0.51 million special cases.}
+%In 2009 the situation has the same pattern. The fifth percentile of normal cases has 11\% higher taxable 
+%income compared to data where special cases are included. The five percent top incomes are 4.9\% higher 
+%within all cases compared to normal cases only. The share of special cases however decreased 
+%to 7.2\%.\footnote{2009: 3.42 million normal cases, 0.27 million special cases}
 
@@
                xlab="Proportion of HBS Distribution")
-title(main=paste("Overall realtive density"))
+title(main=paste("Overall relative density"))
 abline(h=1,lty=2)

### 2014-06-23 paper/subdocuments/results.Rnw
changed label results.Rnw

--- 2014-06-20 paper/subdocuments/results.Rnw
@@
 
-<<Comparing 2002 to 2010>>=
+<<Comparing 2003 to 2010>>=
 library(reldist)
@@
                bar=TRUE, quiet=FALSE,
-               xlab="Proportion of 2002 Distribution")
+               xlab="Proportion of 2003 Distribution")
 title(main=paste("Overall realtive density"))
@@
                yolabs=seq(-1,3,by=0.5),
-               xlab="Proportion of 2002 Distribution")
+               xlab="Proportion of 2003 Distribution")
 title(main=paste("Effect of Median shift"))
@@
                yolabs=seq(-1,3,by=0.5),
-               xlab="Proportion of 2002 Distribution")
+               xlab="Proportion of 2003 Distribution")
 title(main=paste("Effect of different shape"))

### 2014-06-23 paper/subdocuments/results.Rnw
2002->2003 fix

--- 2014-06-23 paper/subdocuments/results.Rnw
@@
                bar=TRUE, quiet=FALSE,
-               xlab="Proportion of 2002 Distribution")
+               xlab="Proportion of 2003 Distribution")
 title(main=paste("Overall relative density"))

### 2014-06-23 paper/subdocuments/results.Rnw
legenden+design, robuster code

@@
   theme_bw()+
-  xlab("tax period")
+  xlab("tax period")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))
 #ggplot(dflong, aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")
@@
   theme_bw()+
-  xlab("tax period")
+  xlab("tax period")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))
 @
@@
 <<Comparing 2003 to 2010>>=
-library(reldist)
+library(reldist, quietly=TRUE, warn.conflicts = FALSE)
 
@@
 fig1legend <- list(x=c(1200,1200),y=c(0.0116031,0.0116031))
-legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+legend(fig1legend,lty=1:2,cex=1, bty="n",
        legend=c("2003","2010"))
@@
 The FTA stopped to publicly report data for special cases after tax period 1993/94. 
-For more recent periods however we can compare already calculated percentiles provided by the FTA which are based on both special and normal cases to normal cases alone.
+For more recent periods however we can compare already calculated percentiles provided by the FTA\footnote{These calculations were done on commission of the FTA within the SNF project Sinergia Nr.
+130648 "The Swiss Confederation: A Natural Laboratory for Research on Fiscal and Political Decentralization" by Raphael Parchet and Stefanie Brilon in coordination with Prof.
+Dr.
+Marius Brülhart.} which are based on both special and normal cases to normal cases alone.
 We will have a look at both, the 1993/94 period as the last period where both 
@@
 %ABBILDUNGEN ZUM VERGLEICH NORMAL-/SONDERFÄLLE
-<<specialcases9394, echo=FALSE,warning=FALSE>>=
+<<specialcases9394, echo=FALSE,warning=FALSE,hide=TRUE>>=
 library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
@@
 axis(side = 2)
-fig1legend <- list(x=c(100,100),y=c(0.0116031,0.0116031))
-legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+fig1legend <- list(x=c(60,60),y=c(0.019,0.019))
+legend(fig1legend,lty=1:2, bty="n",
        legend=c("normal cases","pooled, normal and special cases"))
@@
 par(mfrow=c(1,3))
-reldist(y=pooled_i, yo=normal_i,
+a<-reldist(y=pooled_i, yo=normal_i,
                smooth=0.4, ci=FALSE,
@@
 abline(h=1,lty=2)
-reldist(y=pooled_i, yo=normal_i,
+b<-reldist(y=pooled_i, yo=normal_i,
                show="effect",
@@
 abline(h=1,lty=2)
-reldist(y=pooled_i, yo=normal_i,
+c<-reldist(y=pooled_i, yo=normal_i,
                smooth=0.4, ci=FALSE,
@@
 
-<<specialcases10, echo=FALSE,warning=FALSE>>=
+<<specialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=
 library(foreign)
@@
 axis(side = 2)
-fig1legend <- list(x=c(120,120),y=c(0.0116031,0.0116031))
-legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+fig1legend <- list(x=c(65,65),y=c(0.016,0.016))
+legend(fig1legend,lty=1:2, bty="n",
        legend=c("normal cases","pooled, normal and special cases"))
@@
 levels(dflong$Type)<-c("without zeros","including zeros")
-ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line(size=1)+theme_bw()+xlab("tax period")
-ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))
+ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line(size=1)+theme_bw()+xlab("tax period")+theme(legend.justification=c(0,1),legend.position=c(0, 1))
+ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))+theme(legend.justification=c(1,0),legend.position=c(1, 0))
 @
@@
 library(ggplot2)
-library(reldist)
+library(reldist, quietly=TRUE, warn.conflicts = FALSE)
 library(foreign)
 # Get Brülhartdata for 2010
-estv2010 <- read.csv("data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
+estv2010 <- read.csv("../../data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
 ###########################################
@@
 #####################################################
-
-habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
+try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
 habe10<-habe0911 %.% 
@@
 fig1legend <- list(x=c(2500,2500),y=c(1.103e-02,1.103e-02))
-legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+legend(fig1legend,lty=1:2,cex=1, bty="n",
        legend=c("HBS","FTA"))
@@
 # Select married - HABE
-
-habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
+try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
 habe10<-habe0911 %.% 
@@
 fig1legend <- list(x=c(2500,2500),y=c(0.0101390,0.0101390))
-legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+legend(fig1legend,lty=1:2,cex=1, bty="n",
        legend=c("HBS","FTA"))
@@
 # Select unmarried - HABE
-
-habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
+try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
+try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
 habe10<-habe0911 %.% 
@@
 fig1legend <- list(x=c(1200,1200),y=c(1.764e-02,1.764e-02))
-legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+legend(fig1legend,lty=1:2,cex=1, bty="n",
        legend=c("HBS","FTA"))

### 2014-06-24 paper/subdocuments/results.Rnw

--- 2014-06-23 paper/subdocuments/results.Rnw
@@
 density2 <- density(bd.2010.ind$inc)
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+# Variante mit Log
+
+density1 <- density(log(bd.2003.ind$inc))
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "log(taxable income)", ylab = "density",
+     axes = FALSE,
+     xlim = c(-8, 10),
+     ylim=c(0,0.5655167))
+title(main="2003 vs 2010",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(0,0),y=c(0.5655167,0.5655167))
+legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+       legend=c("2003","2010"))
+density2 <- density(log(bd.2010.ind$inc))
 lines(x = (density2$x), y = density2$y, type = "l",lty=2)

### 2014-06-24 paper/subdocuments/data.Rnw
nuller in summary erwähnt (nur in fussnote)

@@
 However, equivalization (weighting income by household members) is not possible from these aggregate data.
-In any parts of the article where we point to equivalized income, the data stem from BRÜLHART_WIE_NENNEN_WIR_DAS.
+In any parts of the article where we point to equivalized income, the data stem from ready-made calculations (percentiles, Gini coefficients) provided by the FTA\footnote{These calculations were done on commission of the FTA within the SNF project Sinergia Nr.
+130648 "The Swiss Confederation: A Natural Laboratory for Research on Fiscal and Political Decentralization" by Raphael Parchet and Stefanie Brilon in coordination with Prof.
+Dr.
+Marius Brülhart.}.
 Their measures are calculated from the same source (FTA data) but used the number of children and marriage status as an approximation of household size to calculate inequality measures that are pseudo equivalized.
@@
 Both population coverage and income measures are rather consistent for the complete observation period\footnote{Limitations might however exist due to tax evasion and major changes in the tax system, e.g.
-1945.}.
+1945.
+Furthermore, tax units with incomes too small to quality for federal taxation are not documented before 1995 and treated as zeros afterwards making it difficult to track income development in the lower percentiles.}.
 This can not be said for survey data as these are based on samples and therefore require an ideal sampling design or reweighting to be representative and comparable over time.

### 2014-06-24 code/simulate_data.R
minor changes

@@ -9,10 +9,14 @@ library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
 library(ggplot2)
 library(reldist)
 library(foreign)
+#########################
+# Compare FTA Data and HABE Data
+#############################
 
-#######################
+
+####
 #     2010            #
-#######################
+####
 
 ##
 # Prepare FTA-data for 2010 (out of Rudis code) Normal cases
@@ -178,9 +182,9 @@ format(rpluy(y=normal.ind$inc,yo=habe10$inc_habe,
        digits=3)
 
 
-#######################
+##
 #     2005            #
-#######################
+###
 
 
 ####
@@ -400,7 +404,7 @@ title(main=paste("Overall realtive density",cex=0.6)
 estv2010 <- read.csv("data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
 
 
-###########################################
+####
 # Select percentiles for all cases (total)
 
 estv2010.total<-estv2010[c(18:38),c(1,4)]
@@ -492,7 +496,7 @@ abline(h=1,lty=2)
 par(mfrow=c(1,1))
 
 
-########################
+####
 # Separate comparison for unmarried
 
 
@@ -590,9 +594,9 @@ abline(h=1,lty=2)
 par(mfrow=c(1,1))
 
 
-#################
+####
 # Separate comparison for unmaried
-
+####
 
 ##
 # Unmarried ESTV
@@ -690,8 +694,9 @@ par(mfrow=c(1,1))
 
 
 
-###
+####################################################
 # Sind die klassierten Steuerdaten so anders als die Brülhartdaten?
+#######################################################
 
 # Brülhart
 estv2010.total.ind
@@ -713,7 +718,7 @@ legend(fig1legend,lty=1:2,cex=0.5, bty="n",
 density2 <- density(estv2010.total.ind$inc)
 lines(x = (density2$x), y = density2$y, type = "l",lty=2)
 
-# Relative Distribution
+##############Relative Distribution
 
 par(mfrow=c(1,3))
 g10 <- reldist(y=estv2010.total.ind$inc, yo=normal.ind$inc,
@@ -743,5 +748,56 @@ gA0 <- reldist(y=estv2010.total.ind$inc, yo=normal.ind$inc,
[…]

### 2014-06-26 paper/subdocuments/results.Rnw
discard/stage spielchen

--- 2014-06-23 paper/subdocuments/results.Rnw
@@
 library(foreign)
-# Ohne Nuller
 bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
@@
 cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
-prob<-c(cum.p[1],diff(cum.p), .01)
+prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
 freq<-3000
@@
 
+# Variante mit Log
+
+density1 <- density(log(bd.2003.ind$inc))
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "log(taxable income)", ylab = "density",
+     axes = FALSE,
+     xlim = c(-8, 10),
+     ylim=c(0,0.5655167))
+title(main="2003 vs 2010",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(0,0),y=c(0.5655167,0.5655167))
+legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+       legend=c("2003","2010"))
+density2 <- density(log(bd.2010.ind$inc))
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
 # Relative Distribution
@@
                xlab="Proportion of 2003 Distribution")
-title(main=paste("Overall relative density"))
+title(main=paste("Overall realtive density"))
 abline(h=1,lty=2)
@@
 # probabilities
-prob<-c(cum.p[1],diff(cum.p), .01)
+prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
 # extreme values beyond x (to sample)
@@
      axes = FALSE,
-     xlim = c(-70, 4600),
-     ylim=c(0,1.103e-02))
+     xlim = c(-30, 1400),
+     ylim=c(0,1.097e-02))
 title(main="FTA vs HBS income distribution",cex=0.6)
@@
 axis(side = 2)
-fig1legend <- list(x=c(2500,2500),y=c(1.103e-02,1.103e-02))
+fig1legend <- list(x=c(800,800),y=c(1.097e-02,1.097e-02))
 legend(fig1legend,lty=1:2,cex=1, bty="n",
@@
      axes = FALSE,
-     xlim = c(-257, 8000),
-     ylim=c(0,0.0101390))
+     xlim = c(min(rbind(density1$x,density2$x)), max(rbind(density1$x,density2$x))),
+     ylim=c(0,max(rbind(density1$y,density2$y))))
 title(main="FTA vs HBS income distribution",cex=0.6)
@@
 axis(side = 2)
-fig1legend <- list(x=c(2500,2500),y=c(0.0101390,0.0101390))
+fig1legend <- list(x=c(median(rbind(density1$x,density2$x)),median(rbind(density1$x,density2$x))),y=c(max(rbind(density1$y,density2$y)),max(rbind(density1$y,density2$y))))
 legend(fig1legend,lty=1:2,cex=1, bty="n",
@@
      axes = FALSE,
-     xlim = c(-50, 2500),
-     ylim=c(0,1.764e-02))
+     xlim = c(min(rbind(density1$x,density2$x)), max(rbind(density1$x,density2$x))),
+     ylim=c(0,max(rbind(density1$y,density2$y))))
 title(main="FTA vs HBS income distribution",cex=0.6)
@@
 axis(side = 2)
-fig1legend <- list(x=c(1200,1200),y=c(1.764e-02,1.764e-02))
+fig1legend <- list(x=c(median(rbind(density1$x,density2$x)),median(rbind(density1$x,density2$x))),y=c(max(rbind(density1$y,density2$y)),max(rbind(density1$y,density2$y))))
 legend(fig1legend,lty=1:2,cex=1, bty="n",
@@
                xlab="Proportion of HBS Distribution")
-title(main=paste("Overall relative density"))
+title(main=paste("Overall realtive density"))
 abline(h=1,lty=2)

### 2014-06-26 paper/subdocuments/results.Rnw

@@
 fig1legend <- list(x=c(0,0),y=c(0.5655167,0.5655167))
-legend(fig1legend,lty=1:2,cex=0.5, bty="n",
+legend(fig1legend,lty=1:2,cex=1, bty="n",
        legend=c("2003","2010"))
@@
                yolabs=seq(-1,3,by=0.5),
-               ylim=c(0.5,5.0),
+               ylim=c(0.5,2),
                bar=TRUE, quiet=FALSE,
@@
                bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
+               ylim=c(0.5,2), ylab="",
                smooth=0.4, ci=FALSE,
@@
                bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
+               ylim=c(0.5,2), ylab="",
                yolabs=seq(-1,3,by=0.5),
@@
 par(mfrow=c(1,1))
+
+# Polarization index
+
+bd.2010.ind$wgt<-1
+bd.2003.ind$wgt<-1
+
+format(rpy(y=bd.2010.ind$inc,yo=bd.2003.ind$inc,
+           ywgt=bd.2010.ind$wgt,yowgt=bd.2003.ind$wgt,pvalue=FALSE),
+       digits=3) # Median Index, middle value is the estimate 
+format(rpluy(y=bd.2010.ind$inc,yo=bd.2003.ind$inc,
+             ywgt=bd.2010.ind$wgt,yowgt=bd.2003.ind$wgt,pvalue=FALSE),
+       digits=3) # Lower Index
+format(rpluy(y=bd.2010.ind$inc,yo=bd.2003.ind$inc,
+             ywgt=bd.2010.ind$wgt,yowgt=bd.2003.ind$wgt,pvalue=TRUE,
+             upper=TRUE),
+       digits=3) # Upper index
+
 @
@@
                yolabs=seq(-1,3,by=0.5),
-               ylim=c(0.5,5.0),
+               ylim=c(0.5,2.0),
                bar=TRUE, quiet=FALSE,
@@
                bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
+               ylim=c(0.5,2.0), ylab="",
                smooth=0.4, ci=FALSE,
@@
                bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
+               ylim=c(0.5,2.0), ylab="",
                yolabs=seq(-1,3,by=0.5),
@@
 levels(dflong$Type)<-c("without zeros","including zeros")
-ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line(size=1)+theme_bw()+xlab("tax period")+theme(legend.justification=c(0,1),legend.position=c(0, 1))
-ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))+theme(legend.justification=c(1,0),legend.position=c(1, 0))
+ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
+       aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
+  scale_y_continuous(limits=c(0.18,0.6))+
+  scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
+  geom_line(size=0.5)+
+  geom_point(aes(shape=Type),size=3)+
+  theme_bw()+
+  xlab("tax period")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+#ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))+theme(legend.justification=c(1,0),legend.position=c(1, 0))
 @
@@
 try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
+try(habe0911<-read.table("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
+g2009 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t")
+g2010 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t")
+g2011 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t")
+names(g2009)[3:4]<-c("SRH","Gewicht")
+names(g2010)[3:4]<-c("SRH","Gewicht")
+names(g2011)[3:4]<-c("SRH","Gewicht")
+jg091011 <- rbind(g2009,g2010,g2011)
+all.equal(match(jg091011$HaushaltID,habe0911$HaushaltID),1:nrow(habe0911))
+habe0911$jahresgewicht <- jg091011
+
 habe10<-habe0911 %.% 

### 2014-06-26 paper/subdocuments/results.Rnw
Update results.Rnw

@@
 prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
-freq<-3000
+freq<-100000
 init<-0 
@@
 names(bd.2010)[1]=c("inc")
-cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
-prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
-freq<-3000
-init<-0 
-bd.2010$inc<-as.numeric(as.character(bd.2010$inc))
-fin<-(abs(max(bd.2010$inc,na.rm=TRUE))+5)
-ival<-c(init,bd.2010$inc,fin)
-len<-10000
-s<-sapply(2:length(ival),function(i){
+bd.2010 <- rbind(c(0,0),bd.2010)
+ival<-bd.2010$inc
+perval<-bd.2010$percentile
+bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
   seq(ival[i-1],ival[i],length.out=len)
-})
-s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
-bd.2010.ind<-data.frame(s)
-names(bd.2010.ind)[1]=c("inc")
-bd.2010.ind$inc<-bd.2010.ind$inc/1000
+}))
+
+#cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
+#prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
+#freq<-1000
+#init<-0 
+#bd.2010$inc<-as.numeric(as.character(bd.2010$inc))
+#fin<-(abs(max(bd.2010$inc,na.rm=TRUE))+5)
+#ival<-c(init,bd.2010$inc,fin)
+#len<-10000
+#s<-sapply(2:length(ival),function(i){
+#  seq(ival[i-1],ival[i],length.out=len)
+#})
+#s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
+#bd.2010.ind<-data.frame(s)
+#names(bd.2010.ind)[1]=c("inc")
+#bd.2010.ind$inc<-bd.2010.ind$inc/1000
 @
@@
 
+# imputing pseudo-individual data to use reldist
+# we cut at the 5th and 95th percentile because it is unclear what values to use to extrapolate
+bd.2003 <- rbind(c(0,0),bd.2003)
+ival<-bd.2003$inc
+perval<-bd.2003$percentile
+bd.2003.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+bd.2010 <- rbind(c(0,0),bd.2010)
+ival<-bd.2010$inc
+perval<-bd.2010$percentile
+bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
 # Probability Density Function
 
-density1 <- density(bd.2003.ind$inc)
+density1 <- density(bd.2003.ind)
 plot(x = (density1$x), y = density1$y, type = "l",
      xlab = "taxable income", ylab = "density",
-     axes = FALSE,
-     xlim = c(-50, 5000),
-     ylim=c(0,0.0116031))
+     axes = FALSE)
 title(main="2003 vs 2010",cex=0.6)
@@
        legend=c("2003","2010"))
-density2 <- density(bd.2010.ind$inc)
+density2 <- density(bd.2010.ind)
 lines(x = (density2$x), y = density2$y, type = "l",lty=2)
@@
 
-density1 <- density(log(bd.2003.ind$inc))
+density1 <- density(log(bd.2003.ind))
 plot(x = (density1$x), y = density1$y, type = "l",
      xlab = "log(taxable income)", ylab = "density",
-     axes = FALSE,
-     xlim = c(-8, 10),
-     ylim=c(0,0.5655167))
-title(main="2003 vs 2010",cex=0.6)
+     axes = FALSE)
 axis(side = 1)
 axis(side = 2)
-fig1legend <- list(x=c(0,0),y=c(0.5655167,0.5655167))
+fig1legend <- list(x=c(5,5),y=c(0.5655167,0.5655167))
 legend(fig1legend,lty=1:2,cex=1, bty="n",
        legend=c("2003","2010"))
-density2 <- density(log(bd.2010.ind$inc))
+density2 <- density(log(bd.2010.ind))
 lines(x = (density2$x), y = density2$y, type = "l",lty=2)
@@
 abline(h=1,lty=2)
-gA0 <- reldist(y=bd.2010.ind$inc, yo=bd.2003.ind$inc,
+gA0 <- reldist(y=bd.2010.ind, yo=bd.2003.ind,
                smooth=0.4, ci=FALSE,
                show="residual",
+               entropy=NA,
                bar=TRUE, quiet=FALSE,
@@
                yolabs=seq(-1,3,by=0.5),
+               discrete=FALSE,
                xlab="Proportion of 2003 Distribution")
@@
 
-bd.2010.ind$wgt<-1
-bd.2003.ind$wgt<-1
+wgt<-rep(1,length(bd.2003.ind))
 
-format(rpy(y=bd.2010.ind$inc,yo=bd.2003.ind$inc,
-           ywgt=bd.2010.ind$wgt,yowgt=bd.2003.ind$wgt,pvalue=FALSE),
+format(rpy(y=bd.2010.ind,yo=bd.2003.ind,
+           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
        digits=3) # Median Index, middle value is the estimate 
-format(rpluy(y=bd.2010.ind$inc,yo=bd.2003.ind$inc,
-             ywgt=bd.2010.ind$wgt,yowgt=bd.2003.ind$wgt,pvalue=FALSE),
+format(rpluy(y=bd.2010.ind,yo=bd.2003.ind,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
        digits=3) # Lower Index
-format(rpluy(y=bd.2010.ind$inc,yo=bd.2003.ind$inc,
-             ywgt=bd.2010.ind$wgt,yowgt=bd.2003.ind$wgt,pvalue=TRUE,
+format(rpluy(y=bd.2010.ind,yo=bd.2003.ind,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
              upper=TRUE),
@@
                smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               ylim=c(0.5,5.0),
+               ylim=c(0.5,2.0),
                bar=TRUE, quiet=FALSE,
@@
                bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
+               ylim=c(0.5,2.0), ylab="",
                smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
                xlab="Proportion of Normal-Case-Distribution")
@@
                bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
-               yolabs=seq(-1,3,by=0.5),
+               ylim=c(0.5,2.0), ylab="",
                xlab="Proportion of Normal-Case-Distribution")
@@
 par(mfrow=c(1,1))
+
+wgt<-rep(1,length(normal_i))
+
+format(rpy(y=pooled_i,yo=normal_i,
+           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
+       digits=3) # Median Index, middle value is the estimate 
+format(rpluy(y=pooled_i,yo=normal_i,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
+       digits=3) # Lower Index
+format(rpluy(y=pooled_i,yo=normal_i,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
+             upper=TRUE),
+       digits=3) # Upper index
+
+gini(normal_i)-gini(pooled_i)
+
 @
@@
 
+<<with_without_zeros RelDist,warning=FALSE>>=
+bd.2010 <- filter(read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
+bd.0.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
+start <- which(names(bd.2010)=="p1")
+end <- which(names(bd.2010)=="p99_99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
+names(bd.2010)[1]=c("inc")
+bd.0.2010<- data.frame(t(bd.0.2010[1,start:end]),percentile)
+names(bd.0.2010)[1]=c("inc")
+
+bd.0.2010 <- rbind(c(0,0),bd.0.2010)
+ival<-bd.0.2010$inc
+perval<-bd.0.2010$percentile
+bd.0.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+bd.2010 <- rbind(c(0,0),bd.2010)
+ival<-bd.2010$inc
+perval<-bd.2010$percentile
+bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+# Prop-Density Plot
+density1 <- density(log(bd.2010.ind))
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "taxable income", ylab = "density"
+     #axes = FALSE,
+     #xlim = c(-50, 200),
+     #ylim=c(0,0.0116031)
+     )
+title(main="all tax units vs without non-taxed",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(6,6),y=c(0.6,0.6))
+legend(fig1legend,lty=1:2, bty="n",
+       legend=c("without non-taxed","all tax units"))
+density2 <- density(log(bd.0.2010.ind))
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+# Relative Distribution
+
+gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
+               smooth=0.4, ci=FALSE,
+               show="residual",
+               entropy=NA,
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,2), ylab="",
+               yolabs=seq(-1,3,by=0.5),
+               discrete=FALSE,
+               xlab="Proportion of 2010 Distribution (including non-taxed)")
+title(main=paste("Effect of different shape"))
+abline(h=1,lty=2)
+
+# Polarization index
+
+wgt<-rep(1,length(bd.2010.ind))
+
+format(rpy(y=bd.2010.ind,yo=bd.0.2010.ind,
+           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
+       digits=3) # Median Index, middle value is the estimate 
+format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
+       digits=3) # Lower Index
+format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
+             upper=TRUE),
+       digits=3) # Upper index
+
+@
+
 Including zeros leads to significantly higher gini coefficients.
@@
 library(foreign)
+
+bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
+bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
+bd.2010<-bd.0 %.% 
+  filter(Jahr==2010,Einheit=="Total")
+start <- which(names(bd.2010)=="p1")
+end <- which(names(bd.2010)=="p99_99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
+names(bd.2010)[1]=c("inc")
+bd.2010 <- rbind(c(0,0),bd.2010)
+ival<-bd.2010$inc
+perval<-bd.2010$percentile
+bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+bd.2010.ind<-bd.2010.ind/1000
+
+######
+# Old Version
 # Get Brülhartdata for 2010
 estv2010 <- read.csv("../../data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
-###########################################
+###
 # Select percentiles for all cases (total)
 estv2010.total<-estv2010[c(18:38),c(1,4)]
-##
 # Simulate individual data points
@@
      axes = FALSE,
-     xlim = c(-30, 1400),
-     ylim=c(0,1.097e-02))
+     #xlim = c(-30, 1400),
+     ylim=c(0,1.151e-02 ))
 title(main="FTA vs HBS income distribution",cex=0.6)
@@
 axis(side = 2)
-fig1legend <- list(x=c(800,800),y=c(1.097e-02,1.097e-02))
+fig1legend <- list(x=c(400,400),y=c(1.097e-02,1.097e-02))
 legend(fig1legend,lty=1:2,cex=1, bty="n",
        legend=c("HBS","FTA"))
-density2 <- density(estv2010.total.ind$inc)
+density2 <- density(bd.2010.ind)
 lines(x = (density2$x), y = density2$y, type = "l",lty=2)
@@
 
-par(mfrow=c(1,3))
-g10 <- reldist(y=estv2010.total.ind$inc, yo=habe10$inc_habe,
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               ylim=c(0.5,5.0),
-               bar=TRUE, quiet=FALSE,
-               xlab="Proportion of HBS Distribution")
-title(main=paste("Overall relative density"))
-abline(h=1,lty=2)
-g1A <- reldist(y=estv2010.total.ind$inc, yo=habe10$inc_habe,
-               show="effect",
-               bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               xlab="Proportion of HBS Distribution")
-title(main=paste("Effect of Median shift"))
-abline(h=1,lty=2)
-gA0 <- reldist(y=estv2010.total.ind$inc, yo=habe10$inc_habe,
+gA0 <- reldist(y=bd.2010.ind, yo=habe10$inc_habe,
                smooth=0.4, ci=FALSE,
@@
                bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
-               yolabs=seq(-1,3,by=0.5),
+               ylim=c(0.5,2.0), ylab="",
                xlab="Proportion of HBS Distribution")
@@
 abline(h=1,lty=2)
-par(mfrow=c(1,1))
 
@@
 
+bd.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Verheiratet / mariés")
+start <- which(names(bd.2010)=="p1")
+end <- which(names(bd.2010)=="p99_99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
+names(bd.2010)[1]=c("inc")
+bd.2010 <- rbind(c(0,0),bd.2010)
+ival<-bd.2010$inc
+perval<-bd.2010$percentile
+bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+bd.2010.ind<-bd.2010.ind/1000
+
+#######
+# Old Version
 estv2010.married<-estv2010[c(18:38),c(1,2)]
@@
 estv2010.married.ind$inc<-estv2010.married.ind$inc/1000
+#####
 
@@
      axes = FALSE,
-     xlim = c(min(rbind(density1$x,density2$x)), max(rbind(density1$x,density2$x))),
+     xlim = c(min(rbind(density1$x,density2$x)), 800),
      ylim=c(0,max(rbind(density1$y,density2$y))))
@@
        legend=c("HBS","FTA"))
-density2 <- density(estv2010.married.ind$inc)
+density2 <- density(bd.2010.ind)
 lines(x = (density2$x), y = density2$y, type = "l",lty=2)
@@
 
-par(mfrow=c(1,3))
-g10 <- reldist(y=estv2010.married.ind$inc, yo=habe10.married$inc_habe,
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               ylim=c(0.5,5.0),
-               bar=TRUE, quiet=FALSE,
-               xlab="Proportion of HBS Distribution")
-title(main=paste("Overall relative density"))
-abline(h=1,lty=2)
-g1A <- reldist(y=estv2010.married.ind$inc, yo=habe10.married$inc_habe,
-               show="effect",
-               bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               xlab="Proportion of HBS Distribution")
-title(main=paste("Effect of Median shift"))
-abline(h=1,lty=2)
-gA0 <- reldist(y=estv2010.married.ind$inc, yo=habe10.married$inc_habe,
+gA0 <- reldist(y=bd.2010.ind, yo=habe10.married$inc_habe,
                smooth=0.4, ci=FALSE,

### 2014-06-26 paper/subdocuments/results.Rnw
added the famous ben-plot

@@
 \subsection{Intertemporal comparison}
+
+<<benplot>>=
+### libs
+library(reshape, quietly=TRUE, warn.conflicts = FALSE)
+library(foreign)
+library(dplyr)
+library(reldist)
+library(ggplot2)
+
+### Fetch Ginis from different survey data
+number_ticks <- function(n) {function(limits) pretty(limits, n)}
+swissGini<-read.csv("../../data/swissGini.csv",header=TRUE,sep=";")
+
+### Ginis from Bens File (for imputed gap)
+ben <- filter(read.dta("../../data/estv-gini-kt.dta"),kanton=="CH2")
+
+### and for the rest of Ginis:
+estv <- read.dta("../../data/steuerdaten20140522_stata12.dta")
+
+estv_gini <- estv %.% 
+  filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %.% 
+  select(steuerperiode, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %.%
+  group_by(factor(steuerperiode)) %.%
+  mutate(mean_steink=steink/anz_pflichtige) %.%
+  summarise(gini(mean_steink,weights=anz_pflichtige))
+
+names(estv_gini) <- c("Year","FTA")
+estv_gini$Year<-as.numeric(as.character(estv_gini$Year))
+estv_gini$Year[estv_gini$Year<2003] <- estv_gini$Year[estv_gini$Year<2003]-2
+
+all_gini <- data.frame("Year"=seq(1916,2012,by=0.5))
+all_gini$FTA <- estv_gini$FTA[match(all_gini$Year,estv_gini$Year)]
+all_gini$HBS <- swissGini$HABE[match(all_gini$Year,swissGini$Year)]
+all_gini$EU.SILC <- swissGini$EU.SILC[match(all_gini$Year,swissGini$Year)]
+all_gini$LIS <- swissGini$LIS[match(all_gini$Year,swissGini$Year)]
+all_gini$FTA_imp <- ben$gini[match(all_gini$Year,ben$jahr)]
+
+all_gini_long<-melt(all_gini,id="Year")
+
+names(all_gini_long) <- c("Year","Data","Gini")
+ggplot(data=na.omit(all_gini_long),
+  aes(y=Gini,x=Year,shape=Data,linetype=Data))+
+  ylab("Gini from different sources") +
+  scale_y_continuous(limits=c(0.18,0.4)) +
+  scale_x_continuous(limits=c(1915,2012),breaks=number_ticks(10)) +
+  geom_point(size=3)+
+  geom_line()+
+  theme_bw()+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+
+@

### 2014-06-27 paper/subdocuments/results.Rnw
minor changes results.Rnw (including comparison for unmaried)

@@
 
+bd.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Unverheiratet / non-mariés")
+start <- which(names(bd.2010)=="p1")
+end <- which(names(bd.2010)=="p99_99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
+names(bd.2010)[1]=c("inc")
+bd.2010 <- rbind(c(0,0),bd.2010)
+ival<-bd.2010$inc
+perval<-bd.2010$percentile
+bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+bd.2010.ind<-bd.2010.ind/1000
+
 ##
@@
        legend=c("HBS","FTA"))
-density2 <- density(estv2010.unmarried.ind$inc)
+density2 <- density(bd.2010.ind)
 lines(x = (density2$x), y = density2$y, type = "l",lty=2)
@@
 abline(h=1,lty=2)
-gA0 <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
+gA0 <- reldist(y=bd.2010.ind, yo=habe10.unmarried$inc_habe,
                smooth=0.4, ci=FALSE,
@@
                bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
-               yolabs=seq(-1,3,by=0.5),
+               ylim=c(0.5,2.5), ylab="",
                xlab="Proportion of HBS Distribution")

### 2014-06-29 paper/subdocuments/methods.Rnw
revision: how zeros influence gini

@@
 
-\subsubsection{Estimating the bias}
-For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation (lets call those ``zeros'' for convenience) .
-However starting 1995, the FTA provides exactly this information for each canton.
-This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010.
-Consequently we can obtain information whether the bias is stable over time (which makes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
-\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie ``je höher der Steuerfreibetrag, umso stärker der Bias''.
-Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias.
-Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
+\subsubsection{Non-taxed}
+Some tax units are missing in the FTA data as they do not qualify for taxation below a certain income threshold.
+The income distribution is therefore cut at the left end.
+For most of the observed range it is not documented how many tax units fall into this unobserved category.
+Since 1995 however the FTA provides the number of people not being taxed (we will call these tax units ``zeros'').
+From 1995 to 2010 we can therefore estimate the bias we introduce when ignoring the zeros.
+In this section we will try to analyze whether the bias is stable over time (and cantons).
+If the bias was stable over time, we could still analyze trends and changes of inequality no matter how severe the bias is.
 
-One can take a first look at the descriptives plotting the share of zeros over time seperately for each canton.
+First it is useful to look at the descriptives plotting the share of zeros over time.
 
@@
 library(ggplot2)
-ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+theme_bw()
-ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")+theme_bw()
+ggplot(df[df$steuerperiode>=1995&df$kanton=="CH",], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+theme_bw()
+#ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")+theme_bw()
 @
 
-We can see multiple things here:
+The number of zeros increased slightly.
+Different interpretations seems possible.
+The increased share of zeros could be driven by the FTA's inflation adjustments or an increase in social deductions.
+Another plausible explanation is the increasing number of unmarried couples filling in two separate tax forms instead of one so we observe two ``small'' incomes instead of one ``large'' income, thus inflating the number of people falling below the taxation threshold.
 
-\begin{enumerate}
-\item There is a small overall upward trend which we assume to be the Federal Administrations inflation adjustments to the tax threshold.
-\item Geneva and Tessin show wild changes but those might be explained by the tax gap ("Bemessungslücke") that people exploited when the cantons changed the tax system.
-It remains unclear however why we can't see similar patter within other cantons.
-\item There is some variance and we see different patterns over time and cantons.
+%\begin{enumerate}
+%\item There is a small overall upward trend which we assume to be the Federal Administrations inflation adjustments to the tax threshold.
+%\item Geneva and Tessin show wild changes but those might be explained by the tax gap ("Bemessungslücke") that people exploited when the cantons changed the tax system.
+It %remains unclear however why we can't see similar patter within other cantons.
+%\item There is some variance and we see different patterns over time and cantons.
 When estimating Gini coefficients or the like we must therefore assume that ignoring ``the zeros'' leads to a bias that is not stable over time.
-\end{enumerate}
+%\end{enumerate}
 
-We might try two strategies to moderate the problem:
+As we want to report reliably inequality time series we pursue two strategies:
 
@@
 \item Add the zeros as a separate group
-\item Fit a model to predict the inequality measure (e.g. the Gini coefficient) using the share of zeros as a predictor
+\item Carry out a statistical test whether the share of zeros alters the Gini coefficient
 \end{enumerate}
 
-By adding the zeros as a separate group we face several problems.
-With the exception of Geneva 1995/96 Pareto interpolation of percentiles p20 and above seems viable.
-Low percentiles however would need to be extrapolated so we might impose unrealistic assumptions.
-Even when estimating ``safe'' measures above p20 (better p50) or a Gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
-To be consistent with the measure of ``taxable income'' we could assume an income of zero for that group, calculate Gini coefficients and compare them with the original (uncorrected/plain) version.
-To get a rough number we could (by cantons) check the squared correlation between plain and corrected Gini.
+By adding the zeros back as a separate group we need to make an assumption about the income distribution within that group.
+We know these cases had incomes between zero and the threshold (usually around 15.000 CHF for the period 1995-2010).
+In figure WITHWITHOUTZEROSTIMESERIES we show Gini coefficients calculated under the assumption of the zeros being zero.
+This is problematic for two reasons: first we know these units might earn more (up to the threshold) so Ginis calculated on this assumptions will be too high.
+Second we assume the zero group to be perfectly homogenous which on the contrary has a lowering effect on the Gini coefficient.
+Therefore we must at least assume a distribution for the zeros that has some variance and ranges from zero to the threshold.
+We decided to use the most simple approach fulfilling both criteria which is to interpolate a uniform distribution between zero and each tax periods threshold.
 
-The second approach however is more robust as we do not impose additional assumptions but instead exercise some curve fitting.
+The second approach investigates whether the FTA time series is suitable to analyze trends of inequality even if the inequality measures are biased.
+We can do this by regressing the inequality measure (here the Gini coefficient) on the share of zeros.
+Simply spoken we test if cutting a larger (left) tail of the distribution changes the shape of the distribution.
+We include Gini coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time).
 
@@
 
-The model outputs a test statisic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical ``canton gini-level''.
+The model outputs a test statistic for each canton that tells us whether the variation of the zero-rate over time leads to a significant deviation from the typical ``canton gini-level''.
 As the model has a decent fit we are not in great danger of omitted variable bias.
@@
 In our case we can clearly reject the hypothesis that all interactions are zero ($p=\Sexpr{round(ftest$p,3)}$).
-This leads to the conclusion  that Gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted Gini coefficients.
+This leads to the conclusion that Gini coefficients are biased by the variation of the zero-share.
+As a minimum a researcher using tax data should therefore control the share of unobserved people, while the best solution is to thoroughly the mechanisms behind a change in the zero share (e.g.
+this might be due to increased unemployment in one period and due to changes of the tax system in another period.).
+Fortunately the model coefficient of Switzerland as a whole is not significant suggesting that the cantonal biases cancel each other out.
+This seems plausible if the phenomenon is e.g.
+driven by tax competition because most of the ``tax optimization'' happens within Switzerland.
+% which is kind of obvious but at the same time we can use the model to report adjusted Gini coefficients.
 For example one might be interested in how inequality would had developed if the zero-share would have been constant over time.
 (Note RF: predict all data point using canton, time and the initial OR final zero-share to homogenize the time series)
-Furthermore we can quantify how large the bias is and we can do this separately for tax periods or separately for cantons.
+%Furthermore we can quantify how large the bias is and we can do this separately for tax periods or separately for cantons.
 
-For all cantons:
+%For all cantons:
 
-<<magnitude_of_bias>>=
-reduced_fit <- lm(G_steink~kanton+factor(steuerperiode),data=df)
-#summary(reduced_fit)
-@
+%<<magnitude_of_bias>>=
+%reduced_fit <- lm(G_steink~kanton+factor(steuerperiode),data=df)
+%#summary(reduced_fit)
+%@
 
-We can see the the model fit reduces to explaining 61.5\% of the Gini variation versus 92.6\% when the information from the zero-shares was used.
+%We can see the the model fit reduces to explaining 61.5\% of the Gini variation versus 92.6\% when the information from the zero-shares was used.
 Although this so some extent attributable to the additional 26 parameters: this is huge.
 
-The model indicates some cases that deserve more attention: Schwyz (positive coefficient) and Geneva (negative coefficient) and the tax period 2000 as well as the most recent periods.
+%The model indicates some cases that deserve more attention: Schwyz (positive coefficient) and Geneva (negative coefficient) and the tax period 2000 as well as the most recent periods.
 
-A positive coefficient (e.g.
-Schwyz) can be read as follows: In periods with many zeros we measure higher Gini coefficients.
-We can derive, that the distribution of income is more skewed for high incomes than for low incomes.
+Figure COEFPLOT can be read as follows: A positive coefficient (e.g.
+Schwyz) states that we measure higher Gini coefficients in periods with many zeros.
+We can derive, that the distribution of incomes is more skewed for high incomes than for low incomes.
 Simply speaking, the contrast between low and middle class is less pronounced than the contrast between middle and top class.
-One possible explanation would be that incomes stem from two different populations: 1) the people of Schwyz who possibly follow a log-normal or gamma distribution and 2) particularly rich people who moved to Schwyz to avoid taxes.
+One possible explanation would be that incomes stem from two different distributions (populations): 1) local people of Schwyz who follow a less skewed distribution and 2) particularly rich people who moved to Schwyz (to avoid taxes).
 
@@
 
-What can we conclude from that analysis?
+Another information we can read from the coefficient plot is the (zero-)adjusted inequality development represented by the period dummies.
+However these values indicate an unweighted change of the Gini coefficient which relates to cantons (not aggregate Switzerland).
+Fortunately the overall picture of inequality development looks very similar to the plain time series of Gini coefficients.
+%What can we conclude from that analysis?
 First one must notice that aggregate measure like the Gini (or others) do not always react in the same way when we cut off one part of the distribution, therefore the measures calculated from tax data is biased.
@@
 
-We can even see more from the model coefficients.
-The period dummies indicate how the distribution of incomes has changed: compared to 1995, the subsequent periods have a negative coefficient, i.e.
-the bias we introduce by omitting the zeros increased, especially from the mid nineties to the mid 2000s.
-To simplify: cutting off zeros more and more seems to lead to an underestimation of our inequality measure, probably because the skewness in the left part of the distribution increased pointing to an increased pauperization that is masked by omitting the left tail of the income distribution.
+And there is even more to see from the model coefficients: The period-zero-interactions in the model state how an increase of zeros (cutting off a larger piece of the left tail) affects the Gini coefficient in each period compared to 1995.
+We can see that 2004 to 2010 the (negative) effect is significantly larger than 1995.
+To simplify: cutting off zeros increasingly leads to an underestimation of inequality measures, probably because the skewness in the left part of the distribution increased.
+To simplify: in the last decade tax units are increasingly pushed below the observable threshold concealing part of the inequality development.
+One less dramatic interpretation of this result is the increasing number of unmarried couples filling in two separate tax forms.
 

### 2014-06-29 paper/options/global_r_options.Rnw
compile-o-mat

@@
   # Caching on by default
-  opts_chunk$set(cache=FALSE)
+  opts_chunk$set(cache=TRUE)
   # plot options 

### 2014-06-29 paper/paper.tex
compile-o-mat

--- 2014-06-15 paper/paper.tex
@@
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-
 \begin{document}
@@
 % 1.Abschnitt: Bedeutung von Ungleichheit
-Economic Resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular \citep{wilkinson_income_2009}.
+Economic resources can be seen as central indicator for life chances in general and a multitude of outcomes like physical and mental health, life expectancy and crime in particular \citep{wilkinson_income_2009}.
 While the study of social inequality can be considered as one of the core subjects of sociology in more recent years the concern about the widening gap was addressed by global leaders \citep{world_economic_forum_global_2013} and scholars alike.
-Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (\citealt{OECD, 2008 FEHLT}, \citealt{oecd_divided_2011}, \citealt{gornick_income_2013}, \citealt{salverda_changing_2014}).
-Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the ``hollowing of the the middle class'' \citep{Alderson and Doran 2013 FEHLT}.
-Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by Stiglitz (2012,117): ``our democracy is being put at peril.''
-\\
+Empirical evidence acknowledge the supposed trend that economic inequality increased in the majority of western countries over the last decades (\citealt{oecd_growing_2008}, \citealt{oecd_divided_2011}, \citealt{gornick_income_2013}, \citealt{salverda_changing_2014}).
+Although the rise was not uniform, a common pattern seems to be identifiable, which can be referred to as the ``hollowing of the the middle class'' \citep{gornick_how_2013}.
+Households are moving towards the top and the bottom of the distribution relative to the past, which is especially problematic as the middle class can be seen as the core of western democracies or as it is stated by \citet[117]{stiglitz_price_2012}: ``our democracy is being put at peril.''
+\\
+
+% Piketty droppen
 
@@
 
-% Könnte noch besser auf die geplanten Analysen abgestimmt sein (vgl. auf Github-Issues)
-
-% Was wollen wir wissen?
-
-% Wir müssen uns äussern zu: 
-% Analyseinheiten
-% Definition des Einkommens
-% Messung der Verteilung
-
-% Bei den Ergebnissen müsste es Grafiken haben zu:
-% 1) Unterschiedlichen Analyseneinheiten (Mit und ohne Nuller, Normal- und Sonderfälle)
-% 2) Unterschiedliche Einkommensgrössen (Reineinkommen, steuerbares Einkommen, pseudoverfügbares Einkommen (Abzüge von direkten Bundessteuern))
-% 3) Vergleich mit HABE
+% Könnte noch besser auf die geplanten Analysen abgestimmt sein (vgl. auf Github-Issues oder definitives Paper)
+
+%%%%%%%%%%
+% Alte Textbausteine
 
@@
 
+% Nicht angesprochen sind in der Einleitung die Punkte 4 und 5 - Population coverage und intertemporal comparison. Der letzte Satz gefällt mir auch nicht so...
+
 \subsection{Tax statistics in Switzerland}
@@
 
+Data is provided by the FTA in an aggregate form for privacy reasons, i.e.
+they are classified into numerous income brackets.
+While we don't know individual incomes we still have sufficient information (number of tax units per income bracket plus sum of incomes within each bracket) to calculate percentiles, Gini coefficients and other desired measures.
+However, equivalization (weighting income by household members) is not possible from these aggregate data.
+In any parts of the article where we point to equivalized income, the data stem from ready-made calculations (percentiles, Gini coefficients) provided by the FTA\footnote{These calculations were done on commission of the FTA within the SNF project Sinergia Nr.
+130648 "The Swiss Confederation: A Natural Laboratory for Research on Fiscal and Political Decentralization" by Raphael Parchet and Stefanie Brilon in coordination with Prof.
+Dr.
+Marius Brülhart.}.
+Their measures are calculated from the same source (FTA data) but used the number of children and marriage status as an approximation of household size to calculate inequality measures that are pseudo equivalized.
+\\
+
 The FTA provides two income measures: taxable income and net income.
@@
 
-% Bei den Nullern erw?hnen, ab wann Angaben zur Zahl der Nuller vorliegen und ab wann nicht
-% Im Internet hat es Publikationen ab 1947/48
-% Es gibt auch Tabellen auf Gemeindeebene
-% Die Statitstischen Kennzahlen von Br?lhart erw?hnen?
-% Erw?hnen in welcher Form die Informationen publiziert sind. Tabulation by size of income. Not on individual or household level
-
-% http://www.estv.admin.ch/dokumentation/00075/00076/00701/index.html?lang=de
-% Kantone und Gemeinden: 2010 bis 1947/1948
-% Klassierte Tabellen für Abzüge, Einkommen (Steuerbares Einkommen und Reineinkommen) und Steuererträge
-% Normalfälle und Sonderfälle
-
-% Statistsiche Kennzahlen
-% Mit Belastung durch die direkte Bundessteuer: 1973/1974 bis 2010
-% Mit und ohne Belastung durch die direkte Bundessteuer: 1995/1996 bis 2010
-% steuerbares Einkommen/Reineinkommen
-% Jeweils Normal und Sonderfälle
-% Individual-Daten (nicht klassierte daten)
-% äquivalenzeinkommen
+% Verschiedene Punkte werden zweimal erwähnt (hier und in den folgenden Abschnitten). Man muss mal alles am Stück lesen und schauen, ob es so passt.
+
+% Verweis auf Brülhartdaten.
 
@@
 However, recent activities emphasize the need of a broader conceptualization.
-A recent publication from the \citet{oecd_oecd_2013 UNSICHER} condense these ideas into the ICW framework (income, consumption and wealth), which is ment to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the Expert Group in developing the ICW Framework presented in this publication.
-Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_????
-UNSICHER}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}} According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
+A recent publication from the \citet{oecd_oecd_2013} condense these ideas into the ICW framework (income, consumption and wealth), which is meant to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the expert group in developing the ICW Framework presented in this publication.
+Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_2004}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}}.
+According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.
 To gain policy relevant insight, it is recommended to look at the distribution of all three distributions simultaneously.
 Some households with low income, for example, may report adequate levels of consumption expenditure or wealth holdings, or vice-versa.
-But it is also stated \citep[18]{oecd_oecd_2013 UNSICHER}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being.
-http://www.lisdatacenter.org/our-data/lws-database/} \\
+But it is also stated \citep[18]{oecd_oecd_2013}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being.
+\url{http://www.lisdatacenter.org/our-data/lws-database/}} \\
 
 This last statement holds for Switzerland too, although the HBS study is strongly influenced by the recommendations of the Canberra group handbook (United Nations, 2011), which concepts are part of the ICW framework.
-Albeit the awareness of an assessment of income, consumption and wealth simultaneously is rising, we  focus our analysis on income, which is undoubtedly a crucial indicator of economic wellbeing.
-It should be noted, that the Federal Tax Office publishes statistics on income and wealth.
-But it is not possible to analyze the jointly distribution on the individual or household level.
-Also measures of consumption are largely missing in tax data, albeit deductions can be understood as mandatory consumptions somehow.
+Albeit the awareness of an assessment of income, consumption and wealth simultaneously is rising, we focus our analysis on income, which is undoubtedly a crucial indicator of economic well-being.
+It should be noted, that the Federal Tax Administration (FTA) publishes statistics on income and wealth but it is not possible to analyze the joint distribution on the individual or household level.
+Also measures of consumption are largely missing in tax data, albeit deductions can in some sense be understood as mandatory consumptions.
 
@@
 Therefore the awareness of the analyzed concept is crucial.
-Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013 UNSICHER}, \citet[24]{united_nations_canberra_2011}).
-Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually leads to the disposable income.
-The income measure, which finaly shapes the possibility to consume.
+Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013}, \citet[24]{united_nations_canberra_2011}).
+Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually lead to disposable income: the income measure, which finally shapes the possibility to consume.
 Within this framework common other income definitions are situated.
-These definitions are contrasted with income concepts, which can be derived out of the FTA Tax data.
 
@@
 
+% Es musse klar werden um was es geht.
+% These definitions are contrasted with income concepts, which can be derived out of the FTA tax data.
+
 The central income reported through tax statistics is the taxable income.
-It includes all reported incomes (income from employment, income from property and received transfers\footnote{ Mean-tested benefits are not taxed and therefore not included in Tax data.
+It includes all reported incomes (income from employment, income from property and received transfers\footnote{Means-tested benefits are not taxed and therefore not included in tax data.
 Income for low income groups are therefore underestimated.
-However, Piketty and Saez (2001) note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions.
+However, \citep{piketty_income_2003} note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions.
 It is therefore neither a pre-transfer income nor a post-transfer income measure.
 It's rather something in between.
-As the FTA tax statistics include some but not all deductions\footnote{The difference between the real Total income and the taxable income are deductions.
-This includes: professional expenses, travel expenses, interest on debt, Alimonies, Training costs, two-earner deduction, Party Contributions, Payment into pillar 3a, purchases in the pension fund and sideline deductions} it is possible to calculate a sort of "total income".
-As some deductions can be interpreted as compulsory expenses similar to taxes the step towards Total income is a step away from the income, which can be used for consumption.
+As the FTA tax statistics include some but not all deductions\footnote{The difference between the real total income and the taxable income are deductions.
+These include: professional expenses, travel expenses, interest on debt, alimonies, training costs, two-earner deduction, party contributions, private pension provision ``Säule 3a'', buying into the pension plan and sideline deductions} it is possible to calculate a sort of total income, which is called ``net inomce'' (Reineinkommen).
+As some deductions can be interpreted as compulsory expenses similar to taxes the step towards total income is a step away from the income, which can be used for consumption.
 Similar when calculating the disposable income out of the taxable income through accounting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income).
@@
 
-% Die Deductions sind noch nicht gänzlich sauber abgebildet.
-Auf der Bilanz-Seite steht, zum steuerbaren Einkommen kommt man, indem vom Reineinkommen die Sozialabzüge(Kinder- und Fremdbetreung) sowie der Abzug für unterstützungsbedürftige Personen abzogen werden.
-% Auf den ESTV-Tabellen sind jedoch noch mehr Abzüge abgebildet. Was soll/muss/können wir nun abziehen? Bin verwirrt...
+% market income mit primary income ersetzen!
+% Nicht vergessen: Abzüge können über die Zeit ändern. Was bedeutet das?
+
+%RF (16.06.2014): den letzten abschnitt finde ich etwas schwer zu verstehen als jemand der nicht tief in der materie ist.
+vielleicht könnte man sowas machen: It is possible to calculate a kind of "total income" that is in between net and gross income.
+The principle applied is to only deduct expenditures that are unavoidable.
+%OH (16.06.2014): bin mir nicht sicher, ob's mit gross income einfacher ist (kommt in der Grafik nicht vor)
 
@@
 
-\emph{statistical units} \\
-
-The agreed standard on the statistical unit, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013 UNSICHER}.
-Indeed it are the individuals, who receive income, own wealth and experience economic wellbeing, but their possibility to do so, is strongly tied to the concept of household.
+\emph{Statistical units} \\
+
+The agreed standard on the statistical units, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013}.
+Indeed it is the individuals, who receive income, own assets and experience economic well-being, but their possibility to do so, is strongly tied to the concept of household.
 This comprises all persons under the same housing arrangement.
-The basic underlying assumption for collection data on households level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children).
+The basic underlying assumption for collecting data on household level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children).
 Additionally, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing.
-The compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013 UNSICHER}, \citealt{Buhmann et al., 1998 FEHLT}).
+To compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013}, \citealt{buhmann_equivalence_1988}).
 \\
@@
 The Gini coefficient is the most known measure and mainly used for international comparison.
-As it is derived from the lorenz-curve, the quantified amount of inequality can unpretentiously be described in a formal and visual way.
-Therefore the Gini coefficient is easely interpretable.
+As it is derived from the Lorenz-curve, the quantified amount of inequality can unpretentiously be described in a formal and visual way.
+Therefore the Gini coefficient is easily interpretable.
 Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}.
-(1) principle of population: the assessment of inequality is independent of the population size (2) Bresciani-Turroni: the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
-doubling of all income) (3) weak principle of transfers (pigou-dalton): transfers from richer households lead to a reduction of inequality.
-However, several drawbacks are reported in the literatur.
-The most important point is, that the underlying distributional form of the measured inequality is unknown and it is therefore not possibel to see if the measure is driven be a few rich or many poor individuals.
-This can also be problematic for comparision between countries or over time.
-In extreme cases two totally different distrubtions share the same gini-coefficient.\\
+(1) ``\textbf{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textbf{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
+doubling of all income) (3) ``\emph{weak principle of transfers}'' or ``\emph{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality.
+However, several drawbacks are reported in the literature.
+The most important point is, that the underlying distributional form of the measured inequality is unknown and it is therefore not possible to see if the measure is driven by a few rich or many poor individuals.
+This can also be problematic for comparison between countries or over time.
+In extreme cases two totally different distributions share the same Gini-coefficient.\\
+
+% namen mit principle ergänzen?
 
 The recent wave of tax-data studies do not report Gini-coefficents.
-Rather top income shares are informed on, which are calculated not only with taxdata, but togehter with external sources to produce the population and personal income control totals.
-This procedure ensures, that the inequality measure is not biased because of non-fillers, who doesn't appear in tax statistics.
-Leight (2002:594f) compares top income shares with other inequality measures and asks, if they are a useful measure of inequality in a society.
-He tries to answer this question empiricaly by comparing measures of inequality based on top income shares with measures of household or familiy inequality.
-He finds a strong positive relationship, but concludes (P.600): top income shares are far from perfect as a measure of distribution of income across society.
-Top income shares hence inform not completely on how inequality evolves elsewhere in the distribution.
-Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent see above).
-A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between invidviduals, who are either both within the top group or both outsie the top group, then the share measure will remain unchanged.
-\\
+Rather top income shares are informed on, which are calculated not only with tax data, but together with external sources to produce the population and personal income control totals (RF 16.06.2014: hä, was?
+gehts da um die Randverteilung der Bevölkerung?).
+This procedure ensures, that the inequality measure is not biased because of non-fillers, who do not appear in tax statistics.
+\citet{leigh_how_2007} compares top income shares with other inequality measures and asks, whether they are a useful measure of inequality in a society.
+He tries to answer this question empirically by comparing measures of inequality based on top income shares with measures of household or family inequality.
+He finds a strong positive relationship, but concludes (P.600): ``top income shares are far from perfect as a measure of distribution of income across society.'' Top income shares hence inform not completely on how inequality evolves elsewhere in the distribution.
+Furthermore, top income shares only weakly satisfy the Pigou-Dalton transfer principle (in contrast to the Gini-Coefficent as mentioned above).
+A transfer from rich to poor will indeed never increase the top income shares, but if the transfer is between individuals, who are either both within the top group or both outside the top group, then the share measure will remain unchanged.
+\\
+
+% Auf deutsch würde ich noch ebd.
 
 Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading).
-The polarityindex is developed by \citep{Handock and Morris 1999} allows for this.
-Recenently this index was applied in the work of \citep{Alderson, Beckfield and Nielsen 2005 FEHLT} and \citep{Alderson and Doran 2013 FEHLT}.
+The polarizationindex is developed by \citep{Handock and Morris 1999} allows for this.
+Recently this index was applied in the work of \citep{alderson_exactly_2005} and \citep{gornick_how_2013}.
 This approach is rooted in relative distribution methods.
 It includes a precise comparison of the shape of two distributions (groups, over time).
-The main advantage lies herin, that this approche allows to characterize the change, in the sense that it is possible to say, if a polarization occured (reduction of middle class), which equals an increase of inequality and if this change is driven by a change in the upper or lower part of the distribution.
-\\
-
-The literature mentions several other metrics with desired properties, but we will not discuss them further (see for example \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}).
+As the main advantage this approach allows to characterize the change in detail.
+It is possible to see which parts of the distribution changed, e.g.
+whether a polarization occurred (reduction of middle class) -- which equals an increase of inequality -- or this change is driven by either a change in the upper or lower part of the distribution.
+\\
+
+The literature mentions several other metrics with desired properties we will not discuss here (see for example \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}).
 
@@
 The magnitude of this bias in Switzerland, however, is unknown.
-Strategies to handle this kind of bias are discussed in the literature \citep{S?rndal et al.
-1992 FEHLT}, but require a register for every unit, that is proportional to income.
-Currently there exists no such register for Switzerland \citep{ecoplan 2014:43 FEHLT}.
-Currently used micro datasets, which are used for official publications concerning inequality in Switzerland (SILC and HABE) are furthermore confronted with a constructed coverage problem, because these surveys rely on the phone register, which excludes the households not having a registered connection.
+Strategies to handle this kind of bias are discussed in the literature \citep{sarndal_model_2003}, but require a register for every unit, that is proportional to income.
+Currently no such register exists for Switzerland \citet[43]{muller_vermogenslage_2014}.
+Currently used micro datasets, which are used for official publications concerning inequality in Switzerland (SILC and HABE) are furthermore confronted with a constructed coverage problem, because these surveys rely on the phone register, which excludes households not having a registered connection.
 \\
@@
 This is possible for normal and special cases alike.
-From 1995/1996 until 2010 the number of non-taxed units are reported, but not for the years before.
+From 1995/1996 until 2010 the number of non-taxed units is reported, but not for the years before.
 Dell et al.
 (2007) try to estimate the fraction of non-taxed by comparing the reported numbers of tax units to census reports about the number of adult population.
-Following his etimations this fraction drops from 94\% in 1993/1994 to 63\% back in 1945/46.
-\\
-
-An other critical issue with tax data is the problem of tax evasion, which definitely can bias the assessment of inequality.
-Alvaredo and Saez (2006) for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
-Evasion can occured, when individuals try not to fill tax returns or by misreporting of incomes.
+According to their estimations this fraction drops from 94\% in 1993/1994 to 63\% back in 1945/46.
+\\
+
+Another critical issue with tax data is the problem of tax evasion, which definitely can bias the assessment of inequality.
+\citet{alvaredo_income_2009} for example regard estimates of Spanish top incomes prior to 1981 as unreliable due to widespread tax evasion.
+Evasion can occur, when individuals try not to fill tax returns or by misreporting of incomes.
 In Switzerland non-fillers show up in the tax-statistics either way, as long as they are registered.
-This person get an imputed income form an older tax return and infomration given by employers.
+This person gets an imputed income based on an older tax return and information given by employers.
 Only non registered non-fillers are not in the records.
 Therefore non-fillers are a minor problem.
-Not neglectable is the circumstance, that individuals misreport incomes.
+Not negligible is the circumstance, that individuals misreport incomes.
 \citet{feld_tax_2006} examine the role of tax evasion in Switzerland by calculating the difference of the national accounts measures of primary income and the income reported to the tax authorities.
@@
 \\
-
-\emph{Intertemporal comparison}
-
-One advantage of tax data over survey data is their observation period.
-While the FTA data we use range back to 1945, the earliest period of survey data is 2000 (HABE data).
-To be fair, survey data might cover the late periods (2011 and 2012) which are not covered by tax data.
-The striking feature of the long FTA time series is its consistency.
-Both population coverage and income measures are consistent for the complete observation period.
-This can not be said for survey data as these are based on samples and therefore require an ideal sampling design or reweighting to be representative and comparable over time.
-We will see how well sampling and weighting is done in the results section.
 
@@
 \begin{tabular}{lcccc}
-                                   & FTA       & HABE      & EU-SILC   & LIS       \\
-Range                              & 1945-2010 & 2000-2011 & 2007-2012 & 1982-2004 \\
-Cross country comparison possible? & no        & no        & yes       & yes       \\
-Top incomes covered?               & yes       & no        & no        & no        \\
-Low incomes covered?               & no        & yes       & yes       & yes       \\
-Concepts of economic ressources?   & no        & yes       & ?         & ?         \\
-Equivalisation?                    & no        & yes       & ?         &          
+                                   & FTA       & Survey data\\
+Statistical unit                   & individuals/couples  & households\\                                   
+Concepts of economic ressources?   & limited   & yes (for recent periods)\\
+Equivalisation?                    & limited   & yes\\
+Top incomes covered?               & yes       & no\\
+Low incomes covered?               & no        & yes\\
+Range                              & 1945-2010 & 1982-2011\\
+Comparison over time possible? 	   & yes       & limited\\
+Cross country comparison possible? & limited   & yes
+
 \end{tabular}
 \end{table}
+
+%RF: damit haben wir WER, WANN, WAS, wollen wir noch ein WOMIT/WAS DARSTELLEN -> methodik -> reldist ? jaa
+
+To define a standard of measuring economic resources and related inequality we discussed different dimensions of demands the data needs to meet.
+To sum up, ideally we want to measure \emph{disposable income} for \emph{all} swiss \emph{households} yearly for an extended period of time.
+Tax data has one weakness when it comes to the ideal income measure as we need to work with ``taxable income'' which is designed to serve taxation purposes.
+However, both concepts -- what needs to be taxed and what can be spent -- are related as both address a distinction between necessary and voluntary expenses.
+By subtracting taxes from taxable income we are confident to satisfyingly come close the theoretically ideal income measure of disposable income.
+The second most important drawback with tax data is that it does not adequately address households.
+There are few situations in which the tax unit equals the household, that is for individuals living alone, with married partner and/or deductible children and nobody else.
+This ``classic'' household setup however became less common in the last decades so we need to assume statements about household inequality based on taxed data became more and more biased.
+The swiss law seems to be slightly outdated in this context.
+Finally however, another advantage of tax data (over survey data) is indeed the observation period.
+While the FTA data we use range back to 1945, the earliest period of survey data is 2000 (HBS data).
+The striking feature of the long FTA time series is its consistency.
+Both population coverage and income measures are rather consistent for the complete observation period\footnote{Limitations might however exist due to tax evasion and major changes in the tax system, e.g.
+1945.
+Furthermore, tax units with incomes too small to quality for federal taxation are not documented before 1995 and treated as zeros afterwards making it difficult to track income development in the lower percentiles.}.
+This can not be said for survey data as these are based on samples and therefore require an ideal sampling design or reweighting to be representative and comparable over time.
+We will see how well sampling and weighting is done in the results section.
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/different_ginis1}
-
-\includegraphics[width=\maxwidth]{figure/different_ginis2}
+\includegraphics[width=\maxwidth]{figure/different_ginis}
 
@@
 
+\subsection{Statistical units}
+
+% Achtung: Das kann doch nicht stimmen, dass wir hier eine Reihe ab 1993 vorliegen haben.
+% Die ESTV-Zahlen beginnen ab 1995/1996... Ist das jetzt eine Periode, wo man noch Gesamtschweizerische Reihen rekonstruieren kann?
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/With_and_without_equivalence_scale}
+
+\end{knitrout}
+
+\subsection{Measuring inequality}
+
+% Es stellt sich die Frage für welchen Zeitraum der Vergleich gemacht wird.
+% Ich schlage vor, der Vergleich wird mit den Brülhartdaten gemacht und der Vergleich soll von 2003 bis 2010 reichen
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/Comparing_2003_to_20101} 
+\begin{kframe}\begin{verbatim}
+## Smoothing using 0.4
+## Smoothing using 0.4
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/Comparing_2003_to_20102}
+
+\end{knitrout}
+
+% Report Polarization index
+
 \emph{Gini coefficients considering mean-tested benefits}
+
+% Wollen wir das ?berhaupt machen?
 
@@
 
-\emph{Normal versus special cases}
+\subsubsection{Normal versus special cases}
 
 The FTA stopped to publicly report data for special cases after tax period 1993/94. 
-For more recent periods however we can compare data from the 
-BRUELHARTPROJEKT (wie nennen wir das?) to the FTA normal cases, as these data are 
-identical but the BRÜLHART data include special cases. 
+For more recent periods however we can compare already calculated percentiles provided by the FTA\footnote{These calculations were done on commission of the FTA within the SNF project Sinergia Nr.
+130648 "The Swiss Confederation: A Natural Laboratory for Research on Fiscal and Political Decentralization" by Raphael Parchet and Stefanie Brilon in coordination with Prof.
+Dr.
+Marius Brülhart.} which are based on both special and normal cases to normal cases alone.
 We will have a look at both, the 1993/94 period as the last period where both 
-numbers were reported as well as 2009 where we compare data from the same source 
-but using differnt data sheets.
+numbers were reported as well as 2010 where we compare data from the same source (FTA)
+but based on different reporting (aggregated data vs. percentiles).
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
-
-\includegraphics[width=\maxwidth]{figure/specialcases93941} \caption[]{\label{fig:specialcases93941}}
-\end{figure}
-
-\begin{figure}[]
-
-\includegraphics[width=\maxwidth]{figure/specialcases93942} \caption[]{\label{fig:specialcases93942}}
-\end{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/specialcases93941}
+
+\includegraphics[width=\maxwidth]{figure/specialcases93942}
+
+\includegraphics[width=\maxwidth]{figure/specialcases93943} 
+\begin{kframe}\begin{verbatim}
+## Smoothing using 0.4
+## Smoothing using 0.4
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/specialcases93944}
 
@@
 
+1993/94 including special cases changes the distribution slightly at both ends.
+Special cases have a slightly lower median income and therefore are more frequent in the lower percentiles and less frequent in the upper percentiles.
+In addition there is an effect of shape: the distribution of special cases is more skewed so that lower percentiles show lower incomes and higher percentiles show higher incomes compared to just normal cases.
+
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
-
-\includegraphics[width=\maxwidth]{figure/specialcases091} \caption[]{\label{fig:specialcases091}}
-\end{figure}
-
-\begin{figure}[]
-
-\includegraphics[width=\maxwidth]{figure/specialcases092} \caption[]{\label{fig:specialcases092}}
-\end{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/specialcases101}
+
+\includegraphics[width=\maxwidth]{figure/specialcases102}
+
+\includegraphics[width=\maxwidth]{figure/specialcases103} 
+\begin{kframe}\begin{verbatim}
+## Smoothing using 0.4
+## Smoothing using 0.4
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/specialcases104}
 
@@
 
-As we can see from figure~\ref{fig:specialcases93942} and \ref{fig:specialcases092}, special cases differ strongly 
-from normal cases within the low and top percentiles. Within the tax period 1993/94 the fifth percentile of 
-normal cases is 15.8\% higher than the fifth percentile of the combines data while the 95\% percentile of 
-normal cases is even 0.7\% lower if one leaves out special cases. This indicates special cases are very 
-differnt from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
-million tax units defined as being normal cases compared to 0.51 million special cases.}
-In 2009 the situation has the same pattern. The fifth percentile of normal cases has 11\% higher taxable 
-income compared to data where special cases are included. The five percent top incomes are 4.9\% higher 
-within all cases compared to normal cases only. The share of special cases however decreased 
-to 7.2\%.\footnote{2009: 3.42 million normal cases, 0.27 million special cases}
+2010 the picture is similar but more apparent: There is a small effect of median shift lowering incomes in the lowest percentiles. The effect of shape however contributes most.
+
+% Auf neuste Zahlen anpassen
+
+%As we can see from figure~\ref{fig:specialcases93942} and \ref{fig:specialcases092}, special cases differ strongly 
+%from normal cases within the low and top percentiles. Within the tax period 1993/94 the fifth percentile of 
+%normal cases is 15.8\% higher than the fifth percentile of the combines data while the 95\% percentile of 
+%normal cases is even 0.7\% lower if one leaves out special cases. This indicates special cases are very 
+%different from normal cases as the share of special cases is only 15.5\%.\footnote{1993/94 there were 2.76 
+%million tax units defined as being normal cases compared to 0.51 million special cases.}
+%In 2009 the situation has the same pattern. The fifth percentile of normal cases has 11\% higher taxable 
+%income compared to data where special cases are included. The five percent top incomes are 4.9\% higher 
+%within all cases compared to normal cases only. The share of special cases however decreased 
+%to 7.2\%.\footnote{2009: 3.42 million normal cases, 0.27 million special cases}
 
@@
 Including zeros leads to significantly higher gini coefficients.
-However we must keep in mind, that these might be artifically high values as we assume zero income for everyone in the zero group.
-We can conclude more from the graphic: the ratio between both measures seems to be quite constant although for aggregate Switzerland but there are minor deviations for multiple cantons as well as strong deviations for the cantons Geneva and Tessin.
+However we must keep in mind, that these might be artificially high values as we assume zero income for everyone in the zero group.
+We can conclude more from the graphic: the ratio between both measures seems to be quite constant although for aggregate Switzerland but there are minor deviations for multiple cantons as well as strong deviations for the cantons Geneva and Ticino.
 However the problems seem not to result from a shift in the zero-share over time but they are specific for the time-period when the tax system changed.
 
-\emph{normal and special cases}
-
 \emph{comparison of tax-data and survey data distribution}
 
-\subsection{Interremporal comparison}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/setup_data1} 
+\begin{kframe}\begin{verbatim}
+## Smoothing using 0.4
+## Smoothing using 0.4
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/setup_data2}
+
+\includegraphics[width=\maxwidth]{figure/setup_data3} 
+\begin{kframe}\begin{verbatim}
+## Smoothing using 0.4
+## Smoothing using 0.4
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/setup_data4}
+
+\includegraphics[width=\maxwidth]{figure/setup_data5} 
+\begin{kframe}\begin{verbatim}
+## Smoothing using 0.4
+## Smoothing using 0.4
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/setup_data6}
+
+\end{knitrout}
+
+\subsection{Intertemporal comparison}
 

### 2014-06-29 paper/subdocuments/results.Rnw
cleaned Results.Rnw partly (until Sonder und Specialcases)

@@
 
-\emph{Gini coefficients for taxable income, income before deductions and taxable income after tax}
+\emph{Gini coefficients for taxable income, income before deductions and taxable income after fedaral tax}
 
@@
 dflong <- melt(df[,c("G_steink","G_reink","G_taxed","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
-names(dflong)[3]<-"Type"
+names(dflong)[3]<-"Income"
 names(dflong)[4]<-"Gini"
-levels(dflong$Type)<-c("taxable income","income before deductions","taxable income after tax")
-ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,shape=Type))+
-  scale_y_continuous(limits=c(0.18,0.4))+
+levels(dflong$Income)<-c("taxable income","income before deductions","taxable income after tax")
+ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,shape=Income))+
+  scale_y_continuous(limits=c(0.3,0.55))+
   scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
-  geom_point(aes(shape=Type),size=3)+
-  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Type))+
+  geom_point(aes(shape=Income),size=3)+
+  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Income))+
   theme_bw()+
@@
 
-\subsection{Statistical units}
+\emph{Equivalence scale}
 
@@
   filter(Einheit=="Total")
-bd.0<-subset(bd.0,Jahr<1995 | Jahr >2002)
+bd.0<-subset(bd.0,Jahr<1993 | Jahr >2002)
 bd.0.long<-reshape(bd.0,
@@
 bdlong<-rbind(bd.0.long,bd.long)
+names(bdlong)[61]<-"Income"
 @
 
-% Achtung: Das kann doch nicht stimmen, dass wir hier eine Reihe ab 1993 vorliegen haben.
-% Die ESTV-Zahlen beginnen ab 1995/1996... Ist das jetzt eine Periode, wo man noch Gesamtschweizerische Reihen rekonstruieren kann?
-
-<<With and without equivalence scale>>=
+<<equivalencescale>>=
 library(ggplot2)
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
-ggplot(bdlong, aes(x=Jahr,y=Gini,shape=inc_def))+
-  scale_y_continuous(limits=c(0.18,0.6))+
-  scale_x_continuous(limits=c(1940,2012),breaks=number_ticks(10)) +
-  geom_point(aes(shape=inc_def),size=3)+
-  geom_line(data=bdlong,aes(linetype=inc_def))+
+ggplot(bdlong, aes(x=Jahr,y=Gini,shape=Income))+
+  scale_y_continuous(limits=c(0.3,0.55))+
+  scale_x_continuous(limits=c(1970,2012),breaks=number_ticks(10)) +
+  geom_point(aes(shape=Income),size=3)+
+  geom_line(data=bdlong,aes(linetype=Income))+
   theme_bw()+
@@
 
-% Es stellt sich die Frage für welchen Zeitraum der Vergleich gemacht wird.
-% Ich schlage vor, der Vergleich wird mit den Brülhartdaten gemacht und der Vergleich soll von 2003 bis 2010 reichen
+% Vergleich über die Zeit da, wo zum ersten Mal Gesamtschweizerische Daten vorliegen bis zur neusten Reihe: 2003 vs 2010
 
@@
 cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
-prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
-freq<-100000
-init<-0 
-bd.2003$inc<-as.numeric(as.character(bd.2003$inc))
-fin<-(abs(max(bd.2003$inc,na.rm=TRUE))+5)
-ival<-c(init,bd.2003$inc,fin)
-len<-10000
-s<-sapply(2:length(ival),function(i){
+bd.2003 <- rbind(c(0,0),bd.2003)
+ival<-bd.2003$inc
+perval<-bd.2003$percentile
+bd.2003.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
   seq(ival[i-1],ival[i],length.out=len)
-})
-s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
-bd.2003.ind<-data.frame(s)
-names(bd.2003.ind)[1]=c("inc")
-bd.2003.ind$inc<-bd.2003.ind$inc/1000
+}))
+
+# Alt
+#prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
+#init<-0 
+##freq<-100000
+#bd.2003$inc<-as.numeric(as.character(bd.2003$inc))
+#fin<-(abs(max(bd.2003$inc,na.rm=TRUE))+5)
+#ival<-c(init,bd.2003$inc,fin)
+#len<-10000
+#s<-sapply(2:length(ival),function(i){
+#  seq(ival[i-1],ival[i],length.out=len)
+#})
+#s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
+#names(bd.2003.ind)[1]=c("inc")
+#3bd.2003.ind<-data.frame(s)
+#bd.2003.ind$inc<-bd.2003.ind$inc/1000
 @
@@
 }))
-
-#cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
-#prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
-#freq<-1000
-#init<-0 
-#bd.2010$inc<-as.numeric(as.character(bd.2010$inc))
-#fin<-(abs(max(bd.2010$inc,na.rm=TRUE))+5)
-#ival<-c(init,bd.2010$inc,fin)
-#len<-10000
-#s<-sapply(2:length(ival),function(i){
-#  seq(ival[i-1],ival[i],length.out=len)
-#})
-#s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
-#bd.2010.ind<-data.frame(s)
-#names(bd.2010.ind)[1]=c("inc")
-#bd.2010.ind$inc<-bd.2010.ind$inc/1000
+bd.2010.ind<-bd.2010.ind/1000
 @
 
-<<Comparing 2003 to 2010>>=
+<<reldist20032010>>=
 library(reldist, quietly=TRUE, warn.conflicts = FALSE)
-
-# imputing pseudo-individual data to use reldist
-# we cut at the 5th and 95th percentile because it is unclear what values to use to extrapolate
-bd.2003 <- rbind(c(0,0),bd.2003)
-ival<-bd.2003$inc
-perval<-bd.2003$percentile
-bd.2003.ind<-unlist(sapply(2:length(ival),function(i){
-  len=(perval[i]-perval[i-1])*100
-  seq(ival[i-1],ival[i],length.out=len)
-}))
-
-bd.2010 <- rbind(c(0,0),bd.2010)
-ival<-bd.2010$inc
-perval<-bd.2010$percentile
-bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
-  len=(perval[i]-perval[i-1])*100
-  seq(ival[i-1],ival[i],length.out=len)
-}))
 
@@
 
-density1 <- density(bd.2003.ind)
-plot(x = (density1$x), y = density1$y, type = "l",
-     xlab = "taxable income", ylab = "density",
-     axes = FALSE)
-title(main="2003 vs 2010",cex=0.6)
-axis(side = 1)
-axis(side = 2)
-fig1legend <- list(x=c(1200,1200),y=c(0.0116031,0.0116031))
-legend(fig1legend,lty=1:2,cex=1, bty="n",
-       legend=c("2003","2010"))
-density2 <- density(bd.2010.ind)
-lines(x = (density2$x), y = density2$y, type = "l",lty=2)
-
-# Variante mit Log
-
+par(mfrow=c(1,2))
 density1 <- density(log(bd.2003.ind))
@@
 
-par(mfrow=c(1,3))
-g10 <- reldist(y=bd.2010.ind$inc, yo=bd.2003.ind$inc,
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               ylim=c(0.5,2),
-               bar=TRUE, quiet=FALSE,
-               xlab="Proportion of 2003 Distribution")
-title(main=paste("Overall realtive density"))
-abline(h=1,lty=2)
-g1A <- reldist(y=bd.2010.ind$inc, yo=bd.2003.ind$inc,
-               show="effect",
-               bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,2), ylab="",
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               xlab="Proportion of 2003 Distribution")
-title(main=paste("Effect of Median shift"))
-abline(h=1,lty=2)
+# Overall rel dist und median shif effect werden deaktiviert
+# g10 <- reldist(y=bd.2010.ind$inc, yo=bd.2003.ind$inc,
+#               smooth=0.4, ci=FALSE,
+#              ylim=c(0.5,2),
+#               yolabs=seq(-1,3,by=0.5),
+#               bar=TRUE, quiet=FALSE,
+#               xlab="Proportion of 2003 Distribution")
+#title(main=paste("Overall realtive density"))
+#abline(h=1,lty=2)
+#g1A <- reldist(y=bd.2010.ind$inc, yo=bd.2003.ind$inc,
+#              show="effect",
+#               bar=TRUE, quiet=FALSE,
+#               ylim=c(0.5,2), ylab="",
+#               smooth=0.4, ci=FALSE,
+#               yolabs=seq(-1,3,by=0.5),
+#               xlab="Proportion of 2003 Distribution")
+#title(main=paste("Effect of Median shift"))
+#abline(h=1,lty=2)
 gA0 <- reldist(y=bd.2010.ind, yo=bd.2003.ind,
@@
                show="residual",
-               entropy=NA,
                bar=TRUE, quiet=FALSE,
                ylim=c(0.5,2), ylab="",
-               yolabs=seq(-1,3,by=0.5),
                discrete=FALSE,
@@
 par(mfrow=c(1,1))
+@
 
-# Polarization index
-
+<<polindex>>=
 wgt<-rep(1,length(bd.2003.ind))
-
 format(rpy(y=bd.2010.ind,yo=bd.2003.ind,
@@
 @
-
-% Report Polarization index
-
-\emph{Gini coefficients considering mean-tested benefits}
-
-% Wollen wir das ?berhaupt machen?
-
-\subsection{Measuring inequality}
 
@@
 taxable income with and without zeros
+
+% Beim Vergleich der Datensätze unterschieden nach nur besteuerten und unter berücksichtigung der nicht besteuerten, stellt sich die Frage wie die unteren Einkommen imputiert werden.
+Wir haben uns überlegt, dass dies vor allem singles (übrige) betrifft, weil Familien in den unteren Einkommenstabellen bereits sehr selten vorkommen.
+Deswegen reicht es nur die Untergrenze für singels zu berücksichtigen.
 

### 2014-06-30 paper/subdocuments/data.Rnw
updated summary table

@@
                                    & FTA       & Survey data\\
-Statistical unit                   & individuals/couples  & households\\                                   
-Concepts of economic ressources?   & limited   & yes (for recent periods)\\
-Equivalisation?                    & limited   & yes\\
-Top incomes covered?               & yes       & no\\
+Statistical unit                   & Tax units (individuals/married couples)  & Households\\                                   
+Concepts of economic resources  & Data-driven   & Theory-driven\\
+Population coverage & total population & sample \\ 
+Top incomes covered                 & Yes   & No\\
 Low incomes covered?               & no        & yes\\
-Range                              & 1945-2010 & 1982-2011\\
-Comparison over time possible? 	   & yes       & limited\\
-Cross country comparison possible? & limited   & yes
+Comparison over time possible 	   & Long       & Short \\
+Cross country comparison possible? & limited   & possible
 

### 2014-06-30 paper/subdocuments/results.Rnw
update Cleaning of results.Rnw (not finished)

@@
 \section{Results}
-
-% Brauchen wir glaubs nicht mehr
-%\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
-%Kantonsweise Grafiken
-%Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
-
-% Der Übersicht halber habe ich hier die Titel hier analog zum "Struktur-Issue" auf git gesetzt
 
@@
 }))
+bd.2003.ind<-bd.2003.ind/1000
 
 # Alt
+#####
 #prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
@@
 #bd.2003.ind$inc<-bd.2003.ind$inc/1000
+#####
 @
@@
 axis(side = 2)
-fig1legend <- list(x=c(5,5),y=c(0.5655167,0.5655167))
-legend(fig1legend,lty=1:2,cex=1, bty="n",
+fig1legend <- list(x=c(0,0),y=c(0.5655167,0.5655167))
+legend(fig1legend,lty=1:2,cex=0.7, bty="n",
        legend=c("2003","2010"))
@@
 
-<<polindex>>=
+<<polindex20032010>>=
 wgt<-rep(1,length(bd.2003.ind))
@@
        digits=3) # Upper index
-
 @
 
-% Zero-shares (alter Textbaustein)
-%Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later) 
-% Das allenfalls in der Methodensection machen?
-% Vergleich mit der Dell-Reihe zu Non-fillers
-% Überschneidungen gib es leider nicht 
-% Interessant wäre es auch, zu sehen ob es Brüche in den Zeitreihen der Zahl der Steuerpflichtigen und jener ohne Besteuerung gibt.
+% Der Vergleich liesse sich auch mit den Perzentilen des Reineinkommens auf der Grundlage der aggregierten ESTV-Daten machen.Dies wäre ein Vorteil, falls die Nuller in den Individual-Daten der ESTV als Nuller geführt sind und nicht mit dem tatsächlichen steuerbaren Einkommen
+
+% FEHLT: Tabelle mit Median, lower und upper polaritätsindex plus Gini bzw. Differenz der Ginis
 
@@
 
-%ABBILDUNGEN ZUM VERGLEICH NORMAL-/SONDERFÄLLE
-<<specialcases9394, echo=FALSE,warning=FALSE,hide=TRUE>>=
+% Variante 1993/1994
+<<setup_data_ESTVspecialcases9394, echo=FALSE,warning=FALSE,hide=TRUE>>=
 library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
@@
 percentiles <- rbind(normal,pooled)
-ggplot(percentiles, aes(x=percentile,y=inc,group=case,colour=case))+
-	geom_line()+theme_bw()+
-	scale_colour_grey()+
-	xlab("Percentile")+ylab("taxable income")+
-	ggtitle("Taxable incomes of normal and special cases 1993/94")+
-  theme(legend.justification=c(0,1),legend.position=c(0, 1))
-
-#plot relative distribution
-rel <- data.frame(pooled$inc,normal$inc,pooled$percentile)
-rel$change <- rel$pooled.inc/rel$normal.inc
-ggplot(rel, aes(x=pooled.percentile,y=change))+
-	geom_line()+theme_bw()+
-	scale_colour_grey()+
-	geom_line(linetype="dashed",aes(y=1),size=1)+
-	xlab("Percentile")+ylab("taxable income")+
-	ggtitle("Taxable incomes of normal versus special cases 1993/94")
-
-#steuerpflichtige estv / stpf bruelhart
-#1- (2762419 / 3268548)
 
 # imputing pseudo-individual data to use reldist
-# we cut at the 5th and 95th percentile because it is unclear what values to use to extrapolate
 ival<-na.omit(normal$inc[-1])
@@
 
+@
+
+<<specialcases9394, echo=FALSE,warning=FALSE,hide=TRUE>>=
+#plot relative distribution
+
 # Probability Density Function
 
+par(mfrow=c(1,2))
 density1 <- density(normal_i)
@@
 axis(side = 2)
-fig1legend <- list(x=c(60,60),y=c(0.019,0.019))
+fig1legend <- list(x=c(70,70),y=c(0.016,0.016))
 legend(fig1legend,lty=1:2, bty="n",
-       legend=c("normal cases","pooled, normal and special cases"))
+       legend=c("normal cases","pooled, normal and special cases"),cex=0.5)
 density2 <- density(pooled_i)
@@
 
-# Relative Distribution
+# Overall Reldist and Effect of Median shift
+########
 
-par(mfrow=c(1,3))
-a<-reldist(y=pooled_i, yo=normal_i,
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               ylim=c(0.5,2.0),
-               bar=TRUE, quiet=FALSE,
-               xlab="Proportion of Normal-Case-Distribution")
-title(main=paste("Overall relative density 1993/94"))
-abline(h=1,lty=2)
-b<-reldist(y=pooled_i, yo=normal_i,
-               show="effect",
-               bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,2.0), ylab="",
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               xlab="Proportion of Normal-Case-Distribution")
-title(main=paste("Effect of Median shift"))
-abline(h=1,lty=2)
+#par(mfrow=c(1,3))
+#a<-reldist(y=pooled_i, yo=normal_i,
+               #smooth=0.4, ci=FALSE,
+               #yolabs=seq(-1,3,by=0.5),
+              #ylim=c(0.5,2.0),
+               #bar=TRUE, quiet=FALSE,
+               #xlab="Proportion of Normal-Case-Distribution")
+#title(main=paste("Overall relative density 1993/94"))
+#abline(h=1,lty=2)
+#b<-reldist(y=pooled_i, yo=normal_i,
+               #show="effect",
+               #bar=TRUE, quiet=FALSE,
+               #ylim=c(0.5,2.0), ylab="",
+               #smooth=0.4, ci=FALSE,
+               #yolabs=seq(-1,3,by=0.5),
+               #xlab="Proportion of Normal-Case-Distribution")
+#title(main=paste("Effect of Median shift"))
+#abline(h=1,lty=2)
+######
+
+# Effect of differen shape
 c<-reldist(y=pooled_i, yo=normal_i,
@@
                ylim=c(0.5,2.0), ylab="",
-               yolabs=seq(-1,3,by=0.5),
                xlab="Proportion of Normal-Case-Distribution")
@@
 par(mfrow=c(1,1))
+
+# Older Versions Rel-Dist-Plots (1993/1994)
+########
+#ggplot(percentiles, aes(x=percentile,y=inc,group=case,colour=case))+
+#  geom_line()+theme_bw()+
+#  scale_colour_grey()+
+#	xlab("Percentile")+ylab("taxable income")+
+#	ggtitle("Taxable incomes of normal and special cases 1993/94")+
+#  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+# rel <- data.frame(pooled$inc,normal$inc,pooled$percentile)
+# rel$change <- rel$pooled.inc/rel$normal.inc
+# ggplot(rel, aes(x=pooled.percentile,y=change))+
+#	geom_line()+theme_bw()+
+#	scale_colour_grey()+
+#	geom_line(linetype="dashed",aes(y=1),size=1)+
+#	xlab("Percentile")+ylab("taxable income")+
+#	ggtitle("Taxable incomes of normal versus special cases 1993/94")
+#####
+
+@
+
+<<polindexspecialcases9394>>=
+wgt<-rep(1,length(normal_i))
+format(rpy(y=pooled_i,yo=normal_i,
+           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
+       digits=3) # Median Index, middle value is the estimate 
+format(rpluy(y=pooled_i,yo=normal_i,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
+       digits=3) # Lower Index
+format(rpluy(y=pooled_i,yo=normal_i,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
+             upper=TRUE),
+       digits=3) # Upper index
+
+gini(pooled_i)-gini(normal_i)
 @
@@
 
-<<specialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=
+% Variante 2010
+<<setup_data_specialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=
 library(foreign)
@@
 percentiles <- rbind(normal,pooled)
-ggplot(percentiles, aes(x=percentile,y=inc,group=case,colour=case))+
-	geom_line()+theme_bw()+
-	scale_colour_grey()+
-	xlab("Percentile")+ylab("taxable income")+
-	ggtitle("Taxable incomes of normal and special cases 2010")+
-  theme(legend.justification=c(0,1),legend.position=c(0, 1))
-
-#plot relative distribution
-rel <- data.frame(pooled$inc,normal$inc,pooled$percentile)
-rel$change <- rel$pooled.inc/rel$normal.inc
-ggplot(rel, aes(x=pooled.percentile,y=change))+
-	geom_line()+theme_bw()+
-	geom_line(linetype="dashed",aes(y=1),size=2)+
-	scale_colour_grey()+
-	xlab("Percentile")+ylab("taxable income")+
-	ggtitle("Taxable incomes of normal versus special cases 2010")
-
-#steuerpflichtige estv / stpf bruelhart
-#1- (3422210 / 3689063)
 
@@
 # we cut at the 95th percentile because it is unclear what values to use to extrapolate
+# (Cutting through omitting na)
 ival<-na.omit(c(0,normal$inc))
@@
 
+##Older Versions of Reldist-Plot
+######
+#ggplot(percentiles, aes(x=percentile,y=inc,group=case,colour=case))+
+#  geom_line()+theme_bw()+
+#	scale_colour_grey()+
+#	xlab("Percentile")+ylab("taxable income")+
+#	ggtitle("Taxable incomes of normal and special cases 2010")+
+#  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+#####
+@
+
+<<specialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=
+
+#Older Version of Relative Distribution plot
+######## 
+#rel <- data.frame(pooled$inc,normal$inc,pooled$percentile)
+#rel$change <- rel$pooled.inc/rel$normal.inc
+#ggplot(rel, aes(x=pooled.percentile,y=change))+
+	#geom_line()+theme_bw()+
+	#geom_line(linetype="dashed",aes(y=1),size=2)+
+	#scale_colour_grey()+
+	#xlab("Percentile")+ylab("taxable income")+
+	#ggtitle("Taxable incomes of normal versus special cases 2010")
+
+#steuerpflichtige estv / stpf bruelhart
+#1- (3422210 / 3689063)
+######
+
 # Probability Density Function
 
+par(mfrow=c(1,2))
 density1 <- density(normal_i)
@@
      )
-title(main="normal cases vs pooled, normal and special cases 2010",cex=0.6)
+title(main="normal cases vs pooled, normal and special cases 2010",cex.main=0.9)
 axis(side = 1)
@@
 legend(fig1legend,lty=1:2, bty="n",
-       legend=c("normal cases","pooled, normal and special cases"))
+       legend=c("normal cases","pooled, normal and special cases"),cex=0.5)
 density2 <- density(pooled_i)
@@
 
-# Relative Distribution
+# Relative Distribution - Overall and Effect of Median shift
+#########
 
-par(mfrow=c(1,3))
-a<-reldist(y=pooled_i, yo=normal_i,
-               smooth=0.4, ci=FALSE,
-               ylim=c(0.5,2.0),
-               bar=TRUE, quiet=FALSE,
-               xlab="Proportion of Normal-Case-Distribution")
-title(main=paste("Overall relative density 2010"))
-abline(h=1,lty=2)
-b<-reldist(y=pooled_i, yo=normal_i,
-               show="effect",
-               bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,2.0), ylab="",
-               smooth=0.4, ci=FALSE,
-               xlab="Proportion of Normal-Case-Distribution")
-title(main=paste("Effect of Median shift"))
-abline(h=1,lty=2)
+#a<-reldist(y=pooled_i, yo=normal_i,
+#               smooth=0.4, ci=FALSE,
+#               ylim=c(0.5,2.0),
+#               bar=TRUE, quiet=FALSE,
+#               xlab="Proportion of Normal-Case-Distribution")
+# title(main=paste("Overall relative density 2010"))
+# abline(h=1,lty=2)
+# b<-reldist(y=pooled_i, yo=normal_i,
+#               show="effect",
+#               bar=TRUE, quiet=FALSE,
+#               ylim=c(0.5,2.0), ylab="",
+#               smooth=0.4, ci=FALSE,
+#               xlab="Proportion of Normal-Case-Distribution")
+# title(main=paste("Effect of Median shift"))
+# abline(h=1,lty=2)
+######
+
+# Relative Distribution - Effect of different shapes
 c<-reldist(y=pooled_i, yo=normal_i,
@@
 par(mfrow=c(1,1))
+@
+
+<<polindexspecialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=
 
@@
 gini(normal_i)-gini(pooled_i)
-
 @
@@
 
-\emph{Income measures with and without zeros}
-
-taxable income with and without zeros
+\emph{taxable income with and without non-taxed}
 

### 2014-06-30 paper/subdocuments/results.Rnw
cleaned results.Rnw and update of HBS vs FTA comparison

@@
        digits=3) # Upper index
-@
-
-% Der Vergleich liesse sich auch mit den Perzentilen des Reineinkommens auf der Grundlage der aggregierten ESTV-Daten machen.Dies wäre ein Vorteil, falls die Nuller in den Individual-Daten der ESTV als Nuller geführt sind und nicht mit dem tatsächlichen steuerbaren Einkommen
+
+gini(bd.2010.ind)-gini(bd.2003.ind)
+
+@
+
+% Der Vergleich liesse sich auch mit den Perzentilen des Reineinkommens auf der Grundlage der aggregierten ESTV-Daten machen.
+Dies wäre ein Vorteil, falls die Nuller in den Individual-Daten der ESTV als Nuller geführt sind und nicht mit dem tatsächlichen steuerbaren Einkommen
 
@@
 Deswegen reicht es nur die Untergrenze für singels zu berücksichtigen.
-
-<<with_without_zeros,warning=FALSE>>=
-dflong <- melt(df[,c("G_steink","G_steink0","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
+Diese liegt 2010 bei 16'900 verändert sich aber über die Zeit
+
+<<with_without_zeros_Gini,warning=FALSE>>=
+df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
+library(reshape)
+dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
 names(dflong)[3]<-"Type"
 names(dflong)[4]<-"Gini"
-levels(dflong$Type)<-c("without zeros","including zeros")
+levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
 ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
@@
   theme(legend.justification=c(0,1),legend.position=c(0, 1))
+
+# Kantonale Unterschiede
+#####
 #ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))+theme(legend.justification=c(1,0),legend.position=c(1, 0))
-@
-
-<<with_without_zeros RelDist,warning=FALSE>>=
+#####
+@
+
+% Wenn wir annehmen, dass die Brülhart-Daten, sowohl die unteren als auch die oberen Perzentile angemessen abbildend, dann reicht der Vergleich.
+Etwas merkwürdig ist es jetzt, wenn wir die Gini-Reihe mit den ESTV Daten machen und dann den Vergleich mit den Brülhartdaten
+
+<<with_without_zeros_RelDist_mitBrülhart,warning=FALSE>>=
 bd.2010 <- filter(read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
@@
 
+% Wenn etwas bei den unteren Perzentilen nicht stimmt (nicht Besteuerte als Null gewertet), dann müsste der Vergleich mit den ESTV-Nullern gemacht werden.
+% Es ist mir aktuell unklar bei welchem Datensazt jetzt die Perzentile mit imputierten Nullern vorliegt.
+
+<<with_without_zeros_RelDist_mitESTV,warning=FALSE>>=
+
+## Nicht fertig angepasst!!!
+estvSteink<-(read.csv("../../data/ginis_und_perzentile_normal.csv", header=TRUE,sep="\t"))
+estvSteink<-filter(estvSteink,kanton=="CH")
+start <- which(names(estvSteink)=="p1")
+end <- which(names(estvSteink)=="p9999")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+estvReink<- data.frame(t(estvReink[1,start:end]),percentile)
+
+names(bd.2010)[1]=c("inc")
+bd.0.2010<- data.frame(t(bd.0.2010[1,start:end]),percentile)
+names(bd.0.2010)[1]=c("inc")
+
+bd.0.2010 <- rbind(c(0,0),bd.0.2010)
+ival<-bd.0.2010$inc
+perval<-bd.0.2010$percentile
+bd.0.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+bd.2010 <- rbind(c(0,0),bd.2010)
+ival<-bd.2010$inc
+perval<-bd.2010$percentile
+bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+
+# Prop-Density Plot
+density1 <- density(log(bd.2010.ind))
+plot(x = (density1$x), y = density1$y, type = "l",
+     xlab = "taxable income", ylab = "density"
+     #axes = FALSE,
+     #xlim = c(-50, 200),
+     #ylim=c(0,0.0116031)
+     )
+title(main="all tax units vs without non-taxed",cex=0.6)
+axis(side = 1)
+axis(side = 2)
+fig1legend <- list(x=c(6,6),y=c(0.6,0.6))
+legend(fig1legend,lty=1:2, bty="n",
+       legend=c("without non-taxed","all tax units"))
+density2 <- density(log(bd.0.2010.ind))
+lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+# Relative Distribution
+
+gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
+               smooth=0.4, ci=FALSE,
+               show="residual",
+               entropy=NA,
+               bar=TRUE, quiet=FALSE,
+               ylim=c(0.5,2), ylab="",
+               yolabs=seq(-1,3,by=0.5),
+               discrete=FALSE,
+               xlab="Proportion of 2010 Distribution (including non-taxed)")
+title(main=paste("Effect of different shape"))
+abline(h=1,lty=2)
+
+# Polarization index
+
+wgt<-rep(1,length(bd.2010.ind))
+
+format(rpy(y=bd.2010.ind,yo=bd.0.2010.ind,
+           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
+       digits=3) # Median Index, middle value is the estimate 
+format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
+       digits=3) # Lower Index
+format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
+             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
+             upper=TRUE),
+       digits=3) # Upper index
+
+@
+
 Including zeros leads to significantly higher gini coefficients.
@@
 
-<<setup_data>>=
+<<setup_data Brülhart_mitNuller_und_HABE_2010,warning=FALSE>>=
 library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
-library(ggplot2)
-library(reldist, quietly=TRUE, warn.conflicts = FALSE)
 library(foreign)
-
-bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
-bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
-bd.2010<-bd.0 %.% 
-  filter(Jahr==2010,Einheit=="Total")
+bd.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
 start <- which(names(bd.2010)=="p1")
@@
 
-bd.2010.ind<-bd.2010.ind/1000
-
-######
-# Old Version
-# Get Brülhartdata for 2010
-estv2010 <- read.csv("../../data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
-###
-# Select percentiles for all cases (total)
-estv2010.total<-estv2010[c(18:38),c(1,4)]
-# Simulate individual data points
-# cumulative probabilites
-cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
-# probabilities
-prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
-# extreme values beyond x (to sample)
-#freq<-max(normal$anz_pflichtige)
-freq<-3000
-init<-0
-fin<-(abs(max(estv2010.total$Total,na.rm=TRUE))+5)
-# generate the sequence to take pairs from
-ival<-c(init,estv2010.total$Total,fin)
-len<-10000 # sequence of each pair
-s<-sapply(2:length(ival),function(i){
-  seq(ival[i-1],ival[i],length.out=len)
-})
-s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
-estv2010.total.ind<-data.frame(s)
-names(estv2010.total.ind)[1]=c("inc")
-estv2010.total.ind$inc<-estv2010.total.ind$inc/1000
+bd.2010.ind<-bd.2010.ind/1000 # Reskalierung auf 1000er
 
@@
 #####################################################
-#         ACHTUNG: Bisher ohne Gewichte!            #
-#####################################################
+library(reshape)
+library(dplyr)
 try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
@@
 try(habe0911<-read.table("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
-g2009 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t")
-g2010 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t")
-g2011 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t")
+try(g2009 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
+try(g2010 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
+try(g2011 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
+try(g2009 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
+try(g2010 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
+try(g2011 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
 names(g2009)[3:4]<-c("SRH","Gewicht")
@@
 all.equal(match(jg091011$HaushaltID,habe0911$HaushaltID),1:nrow(habe0911))
-habe0911$jahresgewicht <- jg091011
-
-habe10<-habe0911 %.% 
-  filter(Jahr08==2010)
-habe10<-data.frame(habe10$VerfuegbaresEinkommen08)
-names(habe10)<-"inc_habe"
+habe0911$jahresgewicht <- jg091011$Gewicht
+
+habe10<-subset(habe0911,Jahr08==2010)
+habe10$psteink<-habe10$Bruttoeinkommen08+habe10$A31m+habe10$A35m
+
+# Wir brauchen (a) verfügbares Einkommen (b) Bruttoeinkommen und c(Bruttoeinkommen nach Soziversicherungsbeiträgen(A31m) und Monetäre Trasnferausgaben an andere HaushalteA35m)
+habe10<-data.frame(habe10$VerfuegbaresEinkommen08*12/1000,habe10$Bruttoeinkommen08*12/1000,habe10$psteink*12/1000,habe10$jahresgewicht)
+names(habe10)<-c("inc_habe","brutto","psteink","gew")
 habe10$inc_habe[habe10$inc_habe<0]<-0 # One values is -20'000
-habe10$inc_habe<-habe10$inc_habe*12/1000 # transform from monthly income to income per year
-
-# Probability density Function
-
-density1 <- density(habe10$inc_habe)
+@
+
+<<BrülhartvsHABE,warning=FALSE>>=
+library(ggplot2)
+library(reldist, quietly=TRUE, warn.conflicts = FALSE)
+
+# Probability density Function (unterschiedliche Einkommen)
+######
+
+#density1 <- density(bd.2010.ind)
+#plot(x = (density1$x), y = density1$y, type = "l",
+#     xlab = "Einkommen", ylab = "density",
+#     axes = FALSE,
+#     xlim = c(-30, 800))
+#     #ylim=c(0,1.151e-02 ))
+#title(main="FTA vs HBS income distribution",cex=0.6)
+#axis(side = 1)
+#axis(side = 2)
+#fig1legend <- list(x=c(400,400),y=c(0.004,0.004))
+#legend(fig1legend,lty=1:4,cex=1, bty="n",
+#       legend=c("FTA","HBS VerfE","HBS Brutto","HBS Brutto(kor)"))
+#density2 <- density(habe10$inc_habe,weights=habe10$gew/sum(habe10$gew))
+#lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+#density3 <- density(habe10$brutto,weights=habe10$gew/sum(habe10$gew))
+#lines(x = (density3$x), y = density3$y, type = "l",lty=3)
+#density4 <- density(habe10$psteink,weights=habe10$gew/sum(habe10$gew))
+#lines(x = (density4$x), y = density4$y, type = "l",lty=4)
+######
+
+par(mfrow=c(1,2))
+# Probability density Function (FTA (Brülhart vs HBS Brutto (korrigiert)))
+density1 <- density(bd.2010.ind)
 plot(x = (density1$x), y = density1$y, type = "l",
@@
      axes = FALSE,
-     #xlim = c(-30, 1400),
-     ylim=c(0,1.151e-02 ))
+     xlim = c(-30, 800))
+     #ylim=c(0,1.151e-02 ))
 title(main="FTA vs HBS income distribution",cex=0.6)
@@
 axis(side = 2)
-fig1legend <- list(x=c(400,400),y=c(1.097e-02,1.097e-02))
+fig1legend <- list(x=c(400,400),y=c(0.004,0.004))
 legend(fig1legend,lty=1:2,cex=1, bty="n",
-       legend=c("HBS","FTA"))
-density2 <- density(bd.2010.ind)
+       legend=c("FTA","HBS"))
+density2 <- density(habe10$psteink,weights=habe10$gew/sum(habe10$gew))
 lines(x = (density2$x), y = density2$y, type = "l",lty=2)
@@
 
-gA0 <- reldist(y=bd.2010.ind, yo=habe10$inc_habe,
+gA0 <- reldist(y=habe10$psteink, yo=bd.2010.ind,
+               ywgt=habe10$gew,
                smooth=0.4, ci=FALSE,
@@
                ylim=c(0.5,2.0), ylab="",
-               xlab="Proportion of HBS Distribution")
+               xlab="Proportion of FTA Distribution")
 title(main=paste("Effect of different shape"))
 abline(h=1,lty=2)
+par(mfrow=c(1,1))
 
@@
 
-#######
-# Old Version
-estv2010.married<-estv2010[c(18:38),c(1,2)]
-names(estv2010.married)[2]=c("married")
-##
-# Simulate individual data points
-# Percentile-Plot (descriptive  purpose)
-# cumulative probabilites
-cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
-# probabilities
-prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
-# extreme values beyond x (to sample)
-#freq<-max(normal$anz_pflichtige)
-freq<-3000
-init<-0
-fin<-(abs(max(estv2010.married$married,na.rm=TRUE))+5)
-# generate the sequence to take pairs from
-ival<-c(init,estv2010.married$married,fin)
-len<-10000 # sequence of each pair
-s<-sapply(2:length(ival),function(i){
-  seq(ival[i-1],ival[i],length.out=len)
-})
-s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
-estv2010.married.ind<-data.frame(s)
-names(estv2010.married.ind)[1]=c("inc")
-estv2010.married.ind$inc<-estv2010.married.ind$inc/1000
-#####
-
 # Select married - HABE
-try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
-try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
-habe10<-habe0911 %.% 
-  filter(Jahr08==2010)
+
 habe10.married<-habe0911 %.% 
   filter(Familientyp08>130,Jahr08==2010)
-habe10.married<-data.frame(habe10.married$VerfuegbaresEinkommen08)
-names(habe10.married)<-"inc_habe"
+habe10.married$psteink<-habe10.married$Bruttoeinkommen08+habe10.married$A31m+habe10.married$A35m
+# Wir brauchen (a) verfügbares Einkommen (b) Bruttoeinkommen und c(Bruttoeinkommen nach Soziversicherungsbeiträgen(A31m) und Monetäre Trasnferausgaben an andere HaushalteA35m)
+habe10.married<-data.frame(habe10.married$VerfuegbaresEinkommen08*12/1000,habe10.married$Bruttoeinkommen08*12/1000,habe10.married$psteink*12/1000,habe10.married$jahresgewicht)
+names(habe10.married)<-c("inc_habe","brutto","psteink","gew")
 habe10.married$inc_habe[habe10.married$inc_habe<0]<-0 # One values is -20'000
-habe10.married$inc_habe<-habe10.married$inc_habe*12/1000 # transform from monthly income to income per year
 
@@
 
-density1 <- density(habe10.married$inc_habe)
+par(mfrow=c(1,2))
+density1 <- density(bd.2010.ind)
 plot(x = (density1$x), y = density1$y, type = "l",
@@
        legend=c("HBS","FTA"))
-density2 <- density(bd.2010.ind)
+density2 <- density(habe10.married$psteink,weights=habe10.married$gew/sum(habe10.married$gew))
 lines(x = (density2$x), y = density2$y, type = "l",lty=2)
@@
 
-gA0 <- reldist(y=bd.2010.ind, yo=habe10.married$inc_habe,
+gA0 <- reldist(y=habe10.married$inc_habe, yo=bd.2010.ind,
+               ywgt=habe10.married$gew,
                smooth=0.4, ci=FALSE,
@@
 # Separate comparison for unmaried
+
+### Not corrected!!!
 

### 2014-06-30 paper/subdocuments/results.Rnw
noch mehr aufgeräumt

@@
 <<setup_gini_data>>=
-library(foreign)
+library(foreign,quietly=TRUE, warn.conflicts=FALSE)
 library(ggplot2)
@@
 
-<<reldist20032010>>=
+<<reldist20032010, echo=FALSE>>=
 library(reldist, quietly=TRUE, warn.conflicts = FALSE)
@@
 
-<<polindex20032010>>=
-wgt<-rep(1,length(bd.2003.ind))
-format(rpy(y=bd.2010.ind,yo=bd.2003.ind,
-           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Median Index, middle value is the estimate 
-format(rpluy(y=bd.2010.ind,yo=bd.2003.ind,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Lower Index
-format(rpluy(y=bd.2010.ind,yo=bd.2003.ind,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
-             upper=TRUE),
-       digits=3) # Upper index
-
-gini(bd.2010.ind)-gini(bd.2003.ind)
-
+<<polindex20032010, results='asis'>>=
+# Setup functin to produce latex tables with polarization indices and difference in ginis
+polTable <- function(y, yo, ywgt=1, yowgt=1,ndig=2,title="Inequality Indices") {
+stopifnot(is.numeric(y), is.numeric(yo)) # all TRUE
+#y vector with new distribution
+#yo vector with reference distribution
+#ywgt yowgt: weights to use (optional)
+#ndig: digits to display, default 2
+#title: table title (optional)
+
+# load neccessary libraries
+library(reldist,quietly=TRUE,warn.conflicts=FALSE) #for Latex tables
+library(Hmisc,quietly=TRUE,warn.conflicts=FALSE)
+
+# Create vector of weights 1 if nothing else is defined
+if(length(ywgt)==1 & ywgt==1) ywgt <- rep(1, length(y))
+if(length(yowgt)==1 & yowgt==1) yowgt <- rep(1, length(yo))
+
+med <- rpy(y=y,yo=yo, ywgt=ywgt,yowgt=yowgt,pvalue=FALSE)[2] # Median Index, middle value is the estimate so we take [2]
+lower <- rpluy(y=y,yo=yo, ywgt=ywgt,yowgt=yowgt,pvalue=FALSE)[2] # Lower Index
+upper <- rpluy(y=y,yo=yo, ywgt=ywgt,yowgt=yowgt,pvalue=FALSE, upper=TRUE)[2] # Upper index
+dgini <- gini(y)-gini(yo)
+
+latex(format(c("Median Index"=med, "Lower Index"=lower, "Upper Index"=upper, "$\\Delta$ Gini"=dgini),digits=ndig),file="",title=title)
+}
+
+polTable(bd.2010.ind,bd.2003.ind)
 @
@@
 % Variante 1993/1994
-<<setup_data_ESTVspecialcases9394, echo=FALSE,warning=FALSE,hide=TRUE>>=
+<<setup_data_ESTVspecialcases9394>>=
 library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
-library(ggplot2)
+library(ggplot2, quietly=TRUE, warn.conflicts = FALSE)
 #library(reldist)
 normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
-normal <- select(normal, -G_reink,-G_taxed,-ppop0,-G_steink0)
+normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
 #normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
 pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")
+pooled <- select(pooled, kanton, steuerperiode, p1:p99)
 #special <- read.csv("data/ginis_und_perzentile_sonder.csv",segep="\t")
@@
 
-<<specialcases9394, echo=FALSE,warning=FALSE,hide=TRUE>>=
+<<specialcases9394>>=
 #plot relative distribution
@@
 
-<<polindexspecialcases9394>>=
-wgt<-rep(1,length(normal_i))
-format(rpy(y=pooled_i,yo=normal_i,
-           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Median Index, middle value is the estimate 
-format(rpluy(y=pooled_i,yo=normal_i,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Lower Index
-format(rpluy(y=pooled_i,yo=normal_i,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
-             upper=TRUE),
-       digits=3) # Upper index
-
-gini(pooled_i)-gini(normal_i)
+<<polindexspecialcases9394, results='asis'>>=
+polTable(pooled_i,normal_i)
 @
@@
 % Variante 2010
-<<setup_data_specialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=
+<<setup_data_specialcases10>>=
 library(foreign)
@@
 normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2010)
+normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
 start <- which(names(normal)=="p1")
@@
 
-<<specialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=
+<<specialcases10>>=
 
@@
 
-<<polindexspecialcases10, echo=FALSE,warning=FALSE,hide=TRUE>>=
-
-wgt<-rep(1,length(normal_i))
-
-format(rpy(y=pooled_i,yo=normal_i,
-           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Median Index, middle value is the estimate 
-format(rpluy(y=pooled_i,yo=normal_i,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Lower Index
-format(rpluy(y=pooled_i,yo=normal_i,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
-             upper=TRUE),
-       digits=3) # Upper index
-
-gini(normal_i)-gini(pooled_i)
+<<polindexspecialcases10, results='asis'>>=
+polTable(pooled_i,normal_i)
 @
@@
 
-<<with_without_zeros_RelDist_mitBrülhart,warning=FALSE>>=
+<<with_without_zeros_RelDist_mitBruelhart,warning=FALSE>>=
 bd.2010 <- filter(read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
@@
 abline(h=1,lty=2)
-
-# Polarization index
-
-wgt<-rep(1,length(bd.2010.ind))
-
-format(rpy(y=bd.2010.ind,yo=bd.0.2010.ind,
-           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Median Index, middle value is the estimate 
-format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Lower Index
-format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
-             upper=TRUE),
-       digits=3) # Upper index
-
+@
+
+<<polindexzeros, results='asis'>>=
+polTable(bd.2010.ind,bd.0.2010.ind)
 @
@@
 abline(h=1,lty=2)
-
-# Polarization index
-
-wgt<-rep(1,length(bd.2010.ind))
-
-format(rpy(y=bd.2010.ind,yo=bd.0.2010.ind,
-           ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Median Index, middle value is the estimate 
-format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE),
-       digits=3) # Lower Index
-format(rpluy(y=bd.2010.ind,yo=bd.0.2010.ind,
-             ywgt=wgt,yowgt=wgt,pvalue=FALSE,
-             upper=TRUE),
-       digits=3) # Upper index
-
+@
+
+<<polindexzeros2, results='asis'>>=
+polTable(bd.2010.ind,bd.0.2010.ind)
 @
@@
 #####################################################
-library(reshape)
-library(dplyr)
+library(reshape, quietly=TRUE, warn.conflicts=FALSE)
+library(dplyr, quietly=TRUE, warn.conflicts=FALSE)
 try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
@@
 try(g2011 <- read.csv("C:/Users/rudi/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
+try(g2009 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
+try(g2010 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2010_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
+try(g2011 <- read.csv("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE2011_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
 try(g2009 <- read.csv("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE2009_Gewichte_140625UOe.txt",sep="\t"),silent=TRUE)
@@
 jg091011 <- rbind(g2009,g2010,g2011)
-all.equal(match(jg091011$HaushaltID,habe0911$HaushaltID),1:nrow(habe0911))
+#all.equal(match(jg091011$HaushaltID,habe0911$HaushaltID),1:nrow(habe0911)) ## proves it is ok to just attach weights in the default order
 habe0911$jahresgewicht <- jg091011$Gewicht
@@
 
-<<BrülhartvsHABE,warning=FALSE>>=
+<<BruelhartvsHABE,warning=FALSE>>=
 library(ggplot2)
@@
 # Unmarried ESTV
-
+estv2010 <- read.csv("../../data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
 estv2010.unmarried<-estv2010[c(18:38),c(1,3)]
@@
 # Select unmarried - HABE
-try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
-try(habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
 habe10<-habe0911 %.% 
@@
 
+#1918 fixen
+estv <- within(estv, {
+    steink[steuerperiode==1918]<-(eink_ll[steuerperiode==1918]+eink_ul[steuerperiode==1918])/2*anz_pflichtige[steuerperiode==1918]
+})
+
 estv_gini <- estv %.% 

### 2014-06-30 paper/subdocuments/methods.Rnw
switched coefplot to bmp for performance and corrected latex error

@@
 Simply spoken we test if cutting a larger (left) tail of the distribution changes the shape of the distribution.
-We include Gini coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time).
+We include Gini coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time)}.
 
-<<corrected_gini>>=
+<<corrected_gini, dev='bmp'>>=
 df$J<-factor(df$steuerperiode)
@@
 
-library(grid) #for arrow() function
+library(grid,quietly=TRUE, warn.conflicts=FALSE) #for arrow() function
 SmoothCoefficientPlot(fit,modelnames=paste0("By cantons: Impact of the share of zeros on inequality (Gini) Rsq=", round(summary(fit)$r.squared,2)))+geom_text(show_guide=FALSE,x=77,y=2.2,size=4,label="effect variation\nover time")+geom_segment(aes(x = 85, y = 1.5, xend = 71, yend = 1.5))+geom_segment(aes(x = 71, y = 1.5, xend = 71, yend = 1.3))+geom_segment(aes(x = 85, y = 1.5, xend = 85, yend = 1.3))+geom_text(show_guide=FALSE,x=57,y=1.9,size=4,label="effect variation\nby cantons")+geom_segment(aes(x = 71, y = 1.2, xend = 45, yend = 1.2))+geom_segment(aes(x = 71, y = 1.2, xend = 71, yend = 1.0))+geom_segment(aes(x = 45, y = 1.2, xend = 45, yend = 1.0))+ggtitle("Bias variation by time and cantons")

### 2014-06-30 paper/subdocuments/data.Rnw
fixed 1 latex error

@@
 A recent publication from the \citet{oecd_oecd_2013} condense these ideas into the ICW framework (income, consumption and wealth), which is meant to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the expert group in developing the ICW Framework presented in this publication.
-Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_2004}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}}.
+Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_2004} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}}.
 According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being.

### 2014-06-30 paper/subdocuments/introduction.Rnw
fixed 1 latex error

@@
 Therefore this times-series doesn't cover time points before 2007.
-As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm last accessed 21.Mai 2014.} The second important source concerning the distribution of income is the Household Budget Survey (HBS).
+As graph XY shows, following the results from SILC income inequality decreased from 2007 to 2013.\footnote{Data shown in the graph was downloaded from the Eurostat Metadata-portal \url{http://epp.eurostat.ec.europa.eu/cache/ITY_SDDS/EN/ilc_esms.htm} last accessed 21.Mai 2014.} The second important source concerning the distribution of income is the Household Budget Survey (HBS).
 The main focus of this survey lays in providing detailed data on household budgets.
@@
 @
+
+% Change variable-name for the legend
 

### 2014-06-30 paper/paper.tex
compile-o-mat

@@
 
+% Change variable-name for the legend
+
 % 4.Abschnitt: What do we know about Switzerland
@@
 
-\subsubsection{Estimating the bias}
-For most of the observed range (1941 to 2010) we do not have any information how many tax units fall into the category of having income that is not zero but is too little to qualify for federal taxation (lets call those ``zeros'' for convenience) .
-However starting 1995, the FTA provides exactly this information for each canton.
-This enables us to estimate the bias we introduce for each canton and each period between 1995 and 2010.
-Consequently we can obtain information whether the bias is stable over time (which makes it possible to safely interpret the changes of inequality over time) and whether the bias is different for each canton.
-\textcolor{red}{Unterschiede zwischen Kantonen wären gut um zu argumentieren, dass andere Länder auch davon betroffen sind, in etwa sowas wie ``je höher der Steuerfreibetrag, umso stärker der Bias''.
-Länder de erst sehr spät besteuern (und über nicht Besteuerte dann auch nicht Buch führen) haben einen krassen Bias.
-Wir könnten dann empfehlungen geben, ab welchem Perzentil man save interpretieren kann oder so.}
-
-One can take a first look at the descriptives plotting the share of zeros over time seperately for each canton.
+\subsubsection{Non-taxed}
+Some tax units are missing in the FTA data as they do not qualify for taxation below a certain income threshold.
+The income distribution is therefore cut at the left end.
+For most of the observed range it is not documented how many tax units fall into this unobserved category.
+Since 1995 however the FTA provides the number of people not being taxed (we will call these tax units ``zeros'').
+From 1995 to 2010 we can therefore estimate the bias we introduce when ignoring the zeros.
+In this section we will try to analyze whether the bias is stable over time (and cantons).
+If the bias was stable over time, we could still analyze trends and changes of inequality no matter how severe the bias is.
+
+First it is useful to look at the descriptives plotting the share of zeros over time.
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/zero_descriptives1}
-
-\includegraphics[width=\maxwidth]{figure/zero_descriptives2}
+\includegraphics[width=\maxwidth]{figure/zero_descriptives}
 
@@
 
-We can see multiple things here:
-
-\begin{enumerate}
-\item There is a small overall upward trend which we assume to be the Federal Administrations inflation adjustments to the tax threshold.
-\item Geneva and Tessin show wild changes but those might be explained by the tax gap ("Bemessungslücke") that people exploited when the cantons changed the tax system.
-It remains unclear however why we can't see similar patter within other cantons.
-\item There is some variance and we see different patterns over time and cantons.
+The number of zeros increased slightly.
+Different interpretations seems possible.
+The increased share of zeros could be driven by the FTA's inflation adjustments or an increase in social deductions.
+Another plausible explanation is the increasing number of unmarried couples filling in two separate tax forms instead of one so we observe two ``small'' incomes instead of one ``large'' income, thus inflating the number of people falling below the taxation threshold.
+
+%\begin{enumerate}
+%\item There is a small overall upward trend which we assume to be the Federal Administrations inflation adjustments to the tax threshold.
+%\item Geneva and Tessin show wild changes but those might be explained by the tax gap ("Bemessungslücke") that people exploited when the cantons changed the tax system.
+It %remains unclear however why we can't see similar patter within other cantons.
+%\item There is some variance and we see different patterns over time and cantons.
 When estimating Gini coefficients or the like we must therefore assume that ignoring ``the zeros'' leads to a bias that is not stable over time.
-\end{enumerate}
-
-We might try two strategies to moderate the problem:
+%\end{enumerate}
+
+As we want to report reliably inequality time series we pursue two strategies:
 
@@
 \item Add the zeros as a separate group
-\item Fit a model to predict the inequality measure (e.g. the Gini coefficient) using the share of zeros as a predictor
+\item Carry out a statistical test whether the share of zeros alters the Gini coefficient
 \end{enumerate}
 
-By adding the zeros as a separate group we face several problems.
-With the exception of Geneva 1995/96 Pareto interpolation of percentiles p20 and above seems viable.
-Low percentiles however would need to be extrapolated so we might impose unrealistic assumptions.
-Even when estimating ``safe'' measures above p20 (better p50) or a Gini coefficient we need to make an assumption about the income structure of that group (a distribution, a mean income or zero-income).
-To be consistent with the measure of ``taxable income'' we could assume an income of zero for that group, calculate Gini coefficients and compare them with the original (uncorrected/plain) version.
-To get a rough number we could (by cantons) check the squared correlation between plain and corrected Gini.
-
-The second approach however is more robust as we do not impose additional assumptions but instead exercise some curve fitting.
+By adding the zeros back as a separate group we need to make an assumption about the income distribution within that group.
+We know these cases had incomes between zero and the threshold (usually around 15.000 CHF for the period 1995-2010).
+In figure WITHWITHOUTZEROSTIMESERIES we show Gini coefficients calculated under the assumption of the zeros being zero.
+This is problematic for two reasons: first we know these units might earn more (up to the threshold) so Ginis calculated on this assumptions will be too high.
+Second we assume the zero group to be perfectly homogenous which on the contrary has a lowering effect on the Gini coefficient.
+Therefore we must at least assume a distribution for the zeros that has some variance and ranges from zero to the threshold.
+We decided to use the most simple approach fulfilling both criteria which is to interpolate a uniform distribution between zero and each tax periods threshold.
+
+The second approach investigates whether the FTA time series is suitable to analyze trends of inequality even if the inequality measures are biased.
+We can do this by regressing the inequality measure (here the Gini coefficient) on the share of zeros.
+Simply spoken we test if cutting a larger (left) tail of the distribution changes the shape of the distribution.
+We include Gini coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time)}.
 
@@
 In our case we can clearly reject the hypothesis that all interactions are zero ($p=0$).
-This leads to the conclusion  that Gini coefficients are biased by the variation of the zero-share which is kind of obvious but at the same time we can use the model to report adjusted Gini coefficients.
+This leads to the conclusion that Gini coefficients are biased by the variation of the zero-share.
+As a minimum a researcher using tax data should therefore control the share of unobserved people, while the best solution is to thoroughly the mechanisms behind a change in the zero share (e.g.
+this might be due to increased unemployment in one period and due to changes of the tax system in another period.).
+Fortunately the model coefficient of Switzerland as a whole is not significant suggesting that the cantonal biases cancel each other out.
+This seems plausible if the phenomenon is e.g.
+driven by tax competition because most of the ``tax optimization'' happens within Switzerland.
+% which is kind of obvious but at the same time we can use the model to report adjusted Gini coefficients.
 For example one might be interested in how inequality would had developed if the zero-share would have been constant over time.
 (Note RF: predict all data point using canton, time and the initial OR final zero-share to homogenize the time series)
-Furthermore we can quantify how large the bias is and we can do this separately for tax periods or separately for cantons.
-
-For all cantons:
-
-We can see the the model fit reduces to explaining 61.5\% of the Gini variation versus 92.6\% when the information from the zero-shares was used.
+%Furthermore we can quantify how large the bias is and we can do this separately for tax periods or separately for cantons.
+
+%For all cantons:
+
+%<<magnitude_of_bias>>=
+%reduced_fit <- lm(G_steink~kanton+factor(steuerperiode),data=df)
+%#summary(reduced_fit)
+%@
+
+%We can see the the model fit reduces to explaining 61.5\% of the Gini variation versus 92.6\% when the information from the zero-shares was used.
 Although this so some extent attributable to the additional 26 parameters: this is huge.
 
-The model indicates some cases that deserve more attention: Schwyz (positive coefficient) and Geneva (negative coefficient) and the tax period 2000 as well as the most recent periods.
-
-A positive coefficient (e.g.
-Schwyz) can be read as follows: In periods with many zeros we measure higher Gini coefficients.
-We can derive, that the distribution of income is more skewed for high incomes than for low incomes.
+%The model indicates some cases that deserve more attention: Schwyz (positive coefficient) and Geneva (negative coefficient) and the tax period 2000 as well as the most recent periods.
+
+Figure COEFPLOT can be read as follows: A positive coefficient (e.g.
+Schwyz) states that we measure higher Gini coefficients in periods with many zeros.
+We can derive, that the distribution of incomes is more skewed for high incomes than for low incomes.
 Simply speaking, the contrast between low and middle class is less pronounced than the contrast between middle and top class.
-One possible explanation would be that incomes stem from two different populations: 1) the people of Schwyz who possibly follow a log-normal or gamma distribution and 2) particularly rich people who moved to Schwyz to avoid taxes.
+One possible explanation would be that incomes stem from two different distributions (populations): 1) local people of Schwyz who follow a less skewed distribution and 2) particularly rich people who moved to Schwyz (to avoid taxes).
 
@@
 
-% Brauchen wir glaubs nicht mehr
-%\textcolor{red}{Gesamtschweizer Grafik (mit imputierten Daten) einmal eine Linie mit Bias, einmal ohne
-%Kantonsweise Grafiken
-%Kuznets / U-Turn, Test der Hypothesen mit den final Daten. Der link zwischen rein deskriptiv  und Theorie plus Modell ist aktuell noch ein krasser Drahtseilakt...}
-
-% Der Übersicht halber habe ich hier die Titel hier analog zum "Struktur-Issue" auf git gesetzt
-
 \subsection{Defining income}
 
-\emph{Gini coefficients for taxable income, income before deductions and taxable income after tax}
+\emph{Gini coefficients for taxable income, income before deductions and taxable income after fedaral tax}
 
@@
 
-\subsection{Statistical units}
-
-% Achtung: Das kann doch nicht stimmen, dass wir hier eine Reihe ab 1993 vorliegen haben.
-% Die ESTV-Zahlen beginnen ab 1995/1996... Ist das jetzt eine Periode, wo man noch Gesamtschweizerische Reihen rekonstruieren kann?
+\emph{Equivalence scale}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/With_and_without_equivalence_scale}
+\includegraphics[width=\maxwidth]{figure/equivalencescale}
 
@@
 
-% Es stellt sich die Frage für welchen Zeitraum der Vergleich gemacht wird.
-% Ich schlage vor, der Vergleich wird mit den Brülhartdaten gemacht und der Vergleich soll von 2003 bis 2010 reichen
+% Vergleich über die Zeit da, wo zum ersten Mal Gesamtschweizerische Daten vorliegen bis zur neusten Reihe: 2003 vs 2010
 
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/Comparing_2003_to_20101} 
-\begin{kframe}\begin{verbatim}
-## Smoothing using 0.4
-## Smoothing using 0.4
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
 ## Smoothing using 0.4
@@
 \end{kframe}
-\includegraphics[width=\maxwidth]{figure/Comparing_2003_to_20102}
+\includegraphics[width=\maxwidth]{figure/reldist20032010}
 
@@
 
-% Report Polarization index
-
-\emph{Gini coefficients considering mean-tested benefits}
-
-% Wollen wir das ?berhaupt machen?
-
-\subsection{Measuring inequality}
-
-% Zero-shares (alter Textbaustein)
-%Estimation using population statistics (1945/46-1993/94) and official tax administration statistics (1995/96 and later) 
-% Das allenfalls in der Methodensection machen?
-% Vergleich mit der Dell-Reihe zu Non-fillers
-% Überschneidungen gib es leider nicht 
-% Interessant wäre es auch, zu sehen ob es Brüche in den Zeitreihen der Zahl der Steuerpflichtigen und jener ohne Besteuerung gibt.
+\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: splines}}\end{kframe}% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title)
+%
+\begin{table}[!tbp]
+\begin{center}
+\begin{tabular}{ll}
+\hline\hline
+\multicolumn{1}{l}{Inequality Indices}&\multicolumn{1}{c}{}\tabularnewline
+\hline
+Median Index&0.058\tabularnewline
+Lower Index&0.072\tabularnewline
+Upper Index&0.045\tabularnewline
+$\Delta$ Gini&0.025\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+% Der Vergleich liesse sich auch mit den Perzentilen des Reineinkommens auf der Grundlage der aggregierten ESTV-Daten machen.
+Dies wäre ein Vorteil, falls die Nuller in den Individual-Daten der ESTV als Nuller geführt sind und nicht mit dem tatsächlichen steuerbaren Einkommen
+
+% FEHLT: Tabelle mit Median, lower und upper polaritätsindex plus Gini bzw. Differenz der Ginis
 
@@
 
-%ABBILDUNGEN ZUM VERGLEICH NORMAL-/SONDERFÄLLE
+% Variante 1993/1994
+
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/specialcases93941}
-
-\includegraphics[width=\maxwidth]{figure/specialcases93942}
-
-\includegraphics[width=\maxwidth]{figure/specialcases93943} 
-\begin{kframe}\begin{verbatim}
-## Smoothing using 0.4
-## Smoothing using 0.4
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
 ## Smoothing using 0.4
@@
 \end{kframe}
-\includegraphics[width=\maxwidth]{figure/specialcases93944}
+\includegraphics[width=\maxwidth]{figure/specialcases9394}
 
 \end{knitrout}
+
+% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title) 
+%
+\begin{table}[!tbp]
+\begin{center}
+\begin{tabular}{ll}
+\hline\hline
+\multicolumn{1}{l}{Inequality Indices}&\multicolumn{1}{c}{}\tabularnewline
+\hline
+Median Index&0.020\tabularnewline
+Lower Index&0.029\tabularnewline
+Upper Index&0.010\tabularnewline
+$\Delta$ Gini&0.013\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
+% Variante 2010
+
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/specialcases101}
-
-\includegraphics[width=\maxwidth]{figure/specialcases102}
-
-\includegraphics[width=\maxwidth]{figure/specialcases103} 
-\begin{kframe}\begin{verbatim}
-## Smoothing using 0.4
-## Smoothing using 0.4
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
 ## Smoothing using 0.4
@@
 \end{kframe}
-\includegraphics[width=\maxwidth]{figure/specialcases104}
+\includegraphics[width=\maxwidth]{figure/specialcases10}
 
 \end{knitrout}
+
+% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title) 
+%
+\begin{table}[!tbp]
+\begin{center}
+\begin{tabular}{ll}
+\hline\hline
+\multicolumn{1}{l}{Inequality Indices}&\multicolumn{1}{c}{}\tabularnewline
+\hline
+Median Index&0.170\tabularnewline
+Lower Index&0.206\tabularnewline
+Upper Index&0.134\tabularnewline
+$\Delta$ Gini&0.044\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-\emph{Income measures with and without zeros}
-
-taxable income with and without zeros
+\emph{taxable income with and without non-taxed}
+
+% Beim Vergleich der Datensätze unterschieden nach nur besteuerten und unter berücksichtigung der nicht besteuerten, stellt sich die Frage wie die unteren Einkommen imputiert werden.
+Wir haben uns überlegt, dass dies vor allem singles (übrige) betrifft, weil Familien in den unteren Einkommenstabellen bereits sehr selten vorkommen.
+Deswegen reicht es nur die Untergrenze für singels zu berücksichtigen.
+Diese liegt 2010 bei 16'900 verändert sich aber über die Zeit
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/with_without_zeros1}
-
-\includegraphics[width=\maxwidth]{figure/with_without_zeros2}
+\includegraphics[width=\maxwidth]{figure/with_without_zeros_Gini}
 
 \end{knitrout}
+
+% Wenn wir annehmen, dass die Brülhart-Daten, sowohl die unteren als auch die oberen Perzentile angemessen abbildend, dann reicht der Vergleich.
+Etwas merkwürdig ist es jetzt, wenn wir die Gini-Reihe mit den ESTV Daten machen und dann den Vergleich mit den Brülhartdaten
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/with_without_zeros_RelDist_mitBruelhart1} 
+\begin{kframe}\begin{verbatim}
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/with_without_zeros_RelDist_mitBruelhart2}
+
+\end{knitrout}
+
+% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title) 
+%
+\begin{table}[!tbp]
+\begin{center}
+\begin{tabular}{ll}
+\hline\hline
+\multicolumn{1}{l}{Inequality Indices}&\multicolumn{1}{c}{}\tabularnewline
+\hline
+Median Index&-0.1084\tabularnewline
+Lower Index&-0.2155\tabularnewline
+Upper Index&-0.0013\tabularnewline
+$\Delta$ Gini&-0.0952\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+% Wenn etwas bei den unteren Perzentilen nicht stimmt (nicht Besteuerte als Null gewertet), dann müsste der Vergleich mit den ESTV-Nullern gemacht werden.
+% Es ist mir aktuell unklar bei welchem Datensazt jetzt die Perzentile mit imputierten Nullern vorliegt.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt 'estvReink' nicht gefunden}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: undefined columns selected}}\end{kframe}
+\includegraphics[width=\maxwidth]{figure/with_without_zeros_RelDist_mitESTV1} 
+\begin{kframe}\begin{verbatim}
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/with_without_zeros_RelDist_mitESTV2}
+
+\end{knitrout}
+
+% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title) 
+%
+\begin{table}[!tbp]
+\begin{center}
+\begin{tabular}{ll}
+\hline\hline
+\multicolumn{1}{l}{Inequality Indices}&\multicolumn{1}{c}{}\tabularnewline
+\hline
+Median Index&-0.1084\tabularnewline
+Lower Index&-0.2155\tabularnewline
+Upper Index&-0.0013\tabularnewline
+$\Delta$ Gini&-0.0952\tabularnewline
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/setup_data1} 
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/BruelhartvsHABE1} 
+\begin{kframe}\begin{verbatim}
+## Smoothing using 0.4
+\end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/BruelhartvsHABE2}
+
+\includegraphics[width=\maxwidth]{figure/BruelhartvsHABE3} 
 \begin{kframe}\begin{verbatim}
@@
 \end{kframe}
-\includegraphics[width=\maxwidth]{figure/setup_data2}
-
-\includegraphics[width=\maxwidth]{figure/setup_data3} 
-\begin{kframe}\begin{verbatim}
-## Smoothing using 0.4
-## Smoothing using 0.4
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/setup_data4}
-
-\includegraphics[width=\maxwidth]{figure/setup_data5} 
-\begin{kframe}\begin{verbatim}
-## Smoothing using 0.4
-## Smoothing using 0.4
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/setup_data6}
+\includegraphics[width=\maxwidth]{figure/BruelhartvsHABE4}
 
@@
 
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/benplot}
+
+\end{knitrout}
+
 %%%-------------------------------------------------%%%
@@
 %Methodological Main findings 
-%(1) Tax data is useful because its possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
-Its superior to survey data, because the latter is affected by non-responded bias.
+%(1) Tax data is useful because it's possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
+It's superior to survey data, because the latter is affected by non-responded bias.
 Tax data includes data about the whole population [was finden wir wegen den Nullern raus] .
 or at least it should (tax evasion).(3) The FTA-Tax data has different shortcomings, which cannot be handled like a measure of income and wealth which (a) cannot be corrected for household size and (b) is neither a pre-transfer nor a post-transfer income measure.
-Therefor its comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
+Therefor it's comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
 This FTA-Tax data problem can be handled with non-aggregated tax-data .
@@
 
-% M
+% Mögliche Einstiegsfloskel
+% This article highlights the main disparities between taxdata based assessment of inequality opossed to given standard concepts of inequality as it is mainly applied with household-surveys.
+We reflected all important concerns based on the case of Switzerland.
 

### 2014-07-01 paper/subdocuments/results.Rnw
fixed issue with invisible tables (remove table env)

@@
 
-<<polindex20032010, results='asis'>>=
+<<poltable>>=
 # Setup functin to produce latex tables with polarization indices and difference in ginis
@@
 library(Hmisc,quietly=TRUE,warn.conflicts=FALSE)
+library(xtable,quietly=TRUE,warn.conflicts=FALSE)
 
@@
 
-latex(format(c("Median Index"=med, "Lower Index"=lower, "Upper Index"=upper, "$\\Delta$ Gini"=dgini),digits=ndig),file="",title=title)
+latex(format(c("Median Index"=med, "Lower Index"=lower, "Upper Index"=upper, "$\\Delta$ Gini"=dgini),digits=ndig),file="",title=title,table.env=FALSE)
 }
 
+<<polindex20032010, results='asis'>>=
 polTable(bd.2010.ind,bd.2003.ind)

### 2014-07-01 paper/paper.tex
compile-o-mat now with visible tables

@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# reldist: Relative Distribution Methods\\\#\# Version 1.6-2 created on 2013-03-03.\\\#\# copyright (c) 2003, Mark S.
+Handcock, University of California-Los Angeles\\\#\#\ \ For citation information, type citation("{}reldist"{}).\\\#\#\ \ Type help(package="{}reldist"{}) to get started.\\\#\# \\\#\# \\\#\# Attaching package: 'nlme'\\\#\# \\\#\# Das folgende Objekt ist maskiert from 'package:dplyr':\\\#\# \\\#\#\ \ \ \  collapse\\\#\# \\\#\# This is mgcv 1.7-26.
+For overview type 'help("{}mgcv-package"{})'.}}\begin{verbatim}
 ## Smoothing using 0.4
@@
 
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: splines}}\end{kframe}% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title)
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: splines}}\end{kframe}% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title, table.env = FALSE)
 %
-\begin{table}[!tbp]
 \begin{center}
@@
 \end{center}
-\end{table}
 
@@
 
-% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title) 
+% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title, table.env = FALSE) 
 %
-\begin{table}[!tbp]
 \begin{center}
@@
 \hline
-Median Index&0.170\tabularnewline
-Lower Index&0.206\tabularnewline
-Upper Index&0.134\tabularnewline
-$\Delta$ Gini&0.044\tabularnewline
+Median Index&0.031\tabularnewline
+Lower Index&0.039\tabularnewline
+Upper Index&0.022\tabularnewline
+$\Delta$ Gini&0.020\tabularnewline
 \hline

### 2014-07-02 paper/subdocuments/results.Rnw
cleaned results.Rnw Comparison of FTA and HBS (unmarried)

@@
 
-\subsubsection{Normal versus special cases}
+\emph{Normal versus special cases}
 
@@
 
-##
-# Unmarried ESTV
-estv2010 <- read.csv("../../data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
-estv2010.unmarried<-estv2010[c(18:38),c(1,3)]
-names(estv2010.unmarried)[2]=c("unmarried")
-##
-# Simulate individual data points
-# Percentile-Plot (descriptive  purpose)
-# cumulative probabilites
-cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
-# probabilities
-prob<-c(cum.p[1],diff(cum.p), 1-0.9999)
-# extreme values beyond x (to sample)
-#freq<-max(normal$anz_pflichtige)
-freq<-3000
-init<-0
-fin<-(abs(max(estv2010.unmarried$unmarried,na.rm=TRUE))+5)
-# generate the sequence to take pairs from
-ival<-c(init,estv2010.unmarried$unmarried,fin)
-len<-10000 # sequence of each pair
-s<-sapply(2:length(ival),function(i){
-  seq(ival[i-1],ival[i],length.out=len)
-})
-s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
-estv2010.unmarried.ind<-data.frame(s)
-names(estv2010.unmarried.ind)[1]=c("inc")
-estv2010.unmarried.ind$inc<-estv2010.unmarried.ind$inc/1000
-
 # Select unmarried - HABE
@@
 
+par(mfrow=c(1,2))
 density1 <- density(habe10.unmarried$inc_habe)
@@
 # Relative Distribution
-
-par(mfrow=c(1,3))
-g10 <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               ylim=c(0.5,5.0),
-               bar=TRUE, quiet=FALSE,
-               xlab="Proportion of HBS Distribution")
-title(main=paste("Overall realtive density"))
-abline(h=1,lty=2)
-g1A <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
-               show="effect",
-               bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,5.0), ylab="",
-               smooth=0.4, ci=FALSE,
-               yolabs=seq(-1,3,by=0.5),
-               xlab="Proportion of HBS Distribution")
-title(main=paste("Effect of Median shift"))
-abline(h=1,lty=2)
+# Overall and Median effect
+#####
+#g10 <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
+ #              smooth=0.4, ci=FALSE,
+#               yolabs=seq(-1,3,by=0.5),
+   #            ylim=c(0.5,5.0),
+    #           bar=TRUE, quiet=FALSE,
+     #          xlab="Proportion of HBS Distribution")
+#title(main=paste("Overall realtive density"))
+#abline(h=1,lty=2)
+#g1A <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
+ #              show="effect",
+  #             bar=TRUE, quiet=FALSE,
+  #             ylim=c(0.5,5.0), ylab="",
+  #             smooth=0.4, ci=FALSE,
+  #             yolabs=seq(-1,3,by=0.5),
+  #             xlab="Proportion of HBS Distribution")
+#title(main=paste("Effect of Median shift"))
+#abline(h=1,lty=2)
+
 gA0 <- reldist(y=bd.2010.ind, yo=habe10.unmarried$inc_habe,

### 2014-07-02 paper/subdocuments/methods.Rnw
completed relative distribution part in methods.Rnw

@@
 
-In the last section we described the advantages and drawback of taxdata discussing five aspects, which we regard as crucial concering the assessment of inequality.
-To get a feeling of the importance of these aspects, we exploit the FTA-tax data as far as possible and perform several insightful calculations adressing four of the five mentioned aspects.
-No further investigation is possible regarding aspect (1) concepts on measuring economic ressource.
-This is a point which cannot be further addresses with FTA tax data.
-But for the other four thematic areas, we can provide deeper insight.
-In general our main strategy is to apply different possible concepts within one conceptional area (income, units, measurement, population coverage) while holding other conceptual differences constant.
-With this strategy we want to show, the sensitiveness the assessement of inequality is especially if one looks at time trens.
-In this section we  describe which methods we use to produce the results in the next section.
+The RD-framework is based on the concept of a ``relative distribution'', a transformation of the data from two distributions into a single distribution that contains all of the information necessary for scale-invariant comparison.
+This allows the make distribution differences ``visual'' in an elegant way and it is also a base for summary statistics, which are more sensitive to detailed theoretical hypotheses in contrast to other measures like the gini-coefficient or top income-shares, which inform either about the whole population or only about the top part of the distribution.\\
+
+% Probability Density function as a base for the RD
+The goal of RD is to study the differences between two distributions.
+A common example could be the income distribution for man and women.
+Subject of the comparison can also be two distributions describing the same population, but stemming from two different data sources like survey or tax data or even comparison of the same source/population but for different time points, like the income distribution out of tax data for Switzerland today compared to an earlier time point.
+To describe what we gone compare, we start with defining the two distributions.
+On represents the reference population (\Y_{0}) and the other the comparison population \Y.
+Y represents our measure of interest (income).
+A first visual approach is to compare the two probability density functions (PDF).
+The PDF is a function (fy) which describes the distribution of probability over the outcome set and is defined for all possible values of y.
+Also this function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%.
+Out of the comparison of the PDF, it is possible to see, which values of x are more and less probable.
+This allows allready to spot distributional differences over the whole scale of Y visualy.
+\\
+
+The PDF can be characterized by its cumulative distribution function (CDF).
+The CDF can be formulated as F(y), which represents the cumulative probability at point x, which is the probability that a randomly chosen value is less than or equal to y.
+The relative distribution of \Y to \Y_{0} is then defined as
+
+\begin{equation}
+R=F_{0}(Y)
+\end{equation}
+
+R is obtained from \Y by transforming it by the CDF for \Y_{0}, \F_{0}. R measures therefore the relative rank of Y compared to Yo.
+
+% Relative Distribution
+
+\begin{equation}
+g(r)=\frac{f(F_{0}^{-1}(r))}{^{f_{0}(F_{0}^{-1}(r))}}
+\end{equation}
+
+We can calculate the Probability Density Function (g(r)) of RD, where \r represents the proportion of values and F_{0}^{-1}(r) is the inverse cumulative distribution function, also called the quantile function.
+g(r) can be interpreted as a density Ratio, which is defined as the ratio of these two quantities of every percentile of the reference distribution [0,1].
+With a complete overlap of both distributions the Probability Density Function of the RD is 1 at every point of the PDF.
+On the other hand, values higher than 1 represent higher probabilities in the comparison distribution than in the references distribution at this specific point and values lower than 1 respectively represent lower probabilities.
+It is a proper PDF in the sense that it integrates to 1 over the unit interval.\\
+
+% Median and shape differences
+What we get through the above transformation of two distributions is the overall relative probability density.
+But Differences between distributions can be divided into two basic components: changes in location and changes in shape.
+If the comparative distribution is a simple location-shifted version of the reference distribution, then the difference between the two distributions can be parsimoniously summarized by this shift.
+Differences that remain after a location adjustment are differences in ``shape'' (scale, skew and other distributional characteristics).
+When both types of shifts are operating, or when factors other than scale are changing in the shape component, we need a way to separate out the various effects.
+If we want to identify the effect of a location shift and separate it from other changes in the distribution, it is necessary to specify what scale this shift operates on.
+It is possible to adjust distribution by any measure of central tendency.
+We choose here a median location adjustment because the median is a natural, robust and scale invariant unit of measurement.
+Because our interest lies in analyzing distributional differences concerning the degree of inequality, we will focus in the results section on shape differences and look therefore at the relative distribution after the distributions are adjusted for location differences.
+
+% RD based Summary Meausres
+Distributional polarization is of particular interest in the study of inequality.
+However, common inequality indicator (for example Gini or Theil’s index) are not designed to distinguish between growth in the upper and lower tails.
+Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail), downgrading (increases in lower tail).
+The polarization index plays the same role as the difference in Gini-indices, but it can be decomposed to compare the growth in the upper and lower tails.
+Because it is based on the relative distribution it provides a simple link between what is observed in the graphical display and what is measured by the numerical summary.
+\\
+
+The median relative polarization index (MRP) is then the mean absolute deviation around the median of the location-adjusted relative distribution, scaled to produce an index that varies between -1 and 1.
+Given the scaling, a value of zero represents no differences in distributional shape; positive values represent more polarization (increases in the tails of the distribution); and negative values represent less polarization (convergence towards the center of the distribution).
+The measure catches only differences in distributional shape (not location).
+And it has several interesting features.
+MRP can be interpreted in terms of a proportional shift of mass in the distribution from more central to less central values.
+A value of 0.1, for example, is equivalent to a 10\% population shift from the center of the distribution to the upper and lower quartiles and the MRP is decomposable along the scale of y.
+This makes it possible to compare the contribution of each section of the distribution to the overall polarization.
+A natural decomposition is the contributions made by components above and below the median (of g(r)).
 
@@
 
-\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
+%\textbf{Incomplete coverage of the population (left censored data.)} What can be done about the not-taxed?
 \citet{dell_income_2007} impute for non-fillers the 20 percentage of the annual average income.
@@
 
-\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
+%\textbf{changes in taxation system  (switch from annual to biannual taxation)} In the mid-1990s a fundamental change in the Swiss tax system took place by switching form the two-years based praenumerando taxation to the one-year based postnumerando taxation.
 This change was enacted with a transitional period of several years, during which each canton could choose when to adopt the new system.
@@
 
-\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
+%\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
 
-\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
+%\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
 %is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
@@
 
-\textbf{deductions} \citet[477]{dell_income_2007}:" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
-\citet[5]{schaltegger_evolution_2011}: ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
+%\textbf{deductions} \citet[477]{dell_income_2007}:" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
+%\citet[5]{schaltegger_evolution_2011}: ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''.
 Können wir das auch?
@@
 
-Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
+%Studies on income try to focus on the disposable income, which subtracts certain expenditures from the primary income.
 Deductions reflect somehow compulsory expenditures and thus taxable income can be seen as a sort of pseudo disposable income.

### 2014-07-02 paper/subdocuments/introduction.Rnw
introduced figure environments

@@
 
-<<Swiss inequality Plot>>=
+<<Swiss inequality Plot, fig.cap='',fig.pos='H'>>=
 library(ggplot2)

### 2014-07-02 paper/subdocuments/results.Rnw
introduced figure environments

@@
 
-\emph{Gini coefficients for taxable income, income before deductions and taxable income after fedaral tax}
+\emph{Gini coefficients for taxable income, income before deductions and taxable income after federal tax}
 
@@
 
-<<different_ginis>>=
+<<different_ginis, fig.cap='',fig.pos='H'>>=
 df$G_reink[df$G_reink==1]<-NA
@@
 
-<<equivalencescale>>=
+<<equivalencescale, fig.cap='',fig.pos='H'>>=
 library(ggplot2)
@@
 
-<<reldist20032010, echo=FALSE>>=
+<<reldist20032010, echo=FALSE, fig.cap='',fig.pos='H'>>=
 library(reldist, quietly=TRUE, warn.conflicts = FALSE)
@@
 
-The FTA stopped to publicly report data for special cases after tax period 1993/94. 
-For more recent periods however we can compare already calculated percentiles provided by the FTA\footnote{These calculations were done on commission of the FTA within the SNF project Sinergia Nr.
+The FTA distinguishes normal from special cases as described in the data section.
+To test whether it matters which cases the researcher looks at we want to compare the distributions of normal and special cases.
+Unfortunately, the FTA stopped to publicly report data for special cases after tax period 1993/94.
+Therefore we will compare the two distributions for a rather old dataset.
+However the FTA does report aggregate statistics (e.g.
+percentiles) based on a pool of all cases (normal and special) for more recent periods\footnote{These calculations were done on commission of the FTA within the SNF project Sinergia Nr.
 130648 "The Swiss Confederation: A Natural Laboratory for Research on Fiscal and Political Decentralization" by Raphael Parchet and Stefanie Brilon in coordination with Prof.
 Dr.
-Marius Brülhart.} which are based on both special and normal cases to normal cases alone.
-We will have a look at both, the 1993/94 period as the last period where both 
-numbers were reported as well as 2010 where we compare data from the same source (FTA)
-but based on different reporting (aggregated data vs. percentiles).
+Marius Brülhart.} which allows us to do a corresponding analysis for 2010 as well.
 
@@
 
-<<specialcases9394>>=
+<<specialcases9394, fig.cap='',fig.pos='H'>>=
 #plot relative distribution
@@
 
-1993/94 including special cases changes the distribution slightly at both ends.
-Special cases have a slightly lower median income and therefore are more frequent in the lower percentiles and less frequent in the upper percentiles.
-In addition there is an effect of shape: the distribution of special cases is more skewed so that lower percentiles show lower incomes and higher percentiles show higher incomes compared to just normal cases.
+1993/94 a pooled data set of normal and special cases has a slightly higher density at both ends compared to data based in normal cases only.
+Special cases appear to have a slightly lower median income and their distribution is more skewed.
+Therefore special cases are more polarized than normal cases, i.e.
+striving away from the median (positive Median Index of 0.02).
+This tendency is more pronounces in the lower than upper part  of the distribution (Lower Index of 0.029 compared to Upper Index if 0.01).
+As the special cases consist of a broad mix of individuals it remains unclear which factor explains the differences of both distributions.
+Possible explanations can be immigrants partly concentrating in lower income percentiles, low income artists who belong to the special cases or a more technical selection effect: tax units not liable for taxation throughout the whole year are special cases; those cases might have lower incomes, e.g.
+if they moved and stopped working.
+To get a more complete picture we can look how the two distributions relate to each other in 2010 (see FIGURE SPECIAL2010).
 
@@
 
-<<specialcases10>>=
+<<specialcases10, fig.cap='',fig.pos='H'>>=
 
@@
 
-2010 the picture is similar but more apparent: There is a small effect of median shift lowering incomes in the lowest percentiles. The effect of shape however contributes most.
+2010 the picture is similar but more apparent: Special cases appear more frequent around the lower percentiles of the pooled distribution (Lower Index of 0.039), however 2010 there is a more noteworthy effect in the upper part of the distribution (Upper Index of 0.022).
+According to figure \ref{fig:specialcases10} we can attribute this effect to the top percentiles.
+This gives credibility to the thesis that rich immigrants whose number increased between 1994 and 2010 drive the effect.
 
@@
 
-<<with_without_zeros_Gini,warning=FALSE>>=
+<<with_without_zeros_Gini,warning=FALSE, fig.cap='',fig.pos='H'>>=
 df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
@@
 
-<<with_without_zeros_RelDist_mitBruelhart,warning=FALSE>>=
+<<with_without_zeros_RelDist_mitBruelhart,warning=FALSE, fig.cap='',fig.pos='H'>>=
 bd.2010 <- filter(read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
@@
 
-<<with_without_zeros_RelDist_mitESTV,warning=FALSE>>=
+<<with_without_zeros_RelDist_mitESTV,warning=FALSE, fig.cap='',fig.pos='H'>>=
 
@@
 
-<<BruelhartvsHABE,warning=FALSE>>=
+<<BruelhartvsHABE,warning=FALSE, fig.cap='',fig.pos='H'>>=
 library(ggplot2)
@@
 
-<<benplot>>=
+<<benplot, fig.cap='',fig.pos='H'>>=
  ### libs

### 2014-07-02 paper/subdocuments/methods.Rnw
fixed formulae

@@
 The RD-framework is based on the concept of a ``relative distribution'', a transformation of the data from two distributions into a single distribution that contains all of the information necessary for scale-invariant comparison.
-This allows the make distribution differences ``visual'' in an elegant way and it is also a base for summary statistics, which are more sensitive to detailed theoretical hypotheses in contrast to other measures like the gini-coefficient or top income-shares, which inform either about the whole population or only about the top part of the distribution.\\
+This allows to make distribution differences ``visual'' in an elegant way and it is also a base for summary statistics, which are more sensitive to detailed theoretical hypotheses in contrast to other measures like the gini-coefficient or top income-shares, which inform either about the whole population or only about the top part of the distribution.\\
 
@@
 Subject of the comparison can also be two distributions describing the same population, but stemming from two different data sources like survey or tax data or even comparison of the same source/population but for different time points, like the income distribution out of tax data for Switzerland today compared to an earlier time point.
-To describe what we gone compare, we start with defining the two distributions.
-On represents the reference population (\Y_{0}) and the other the comparison population \Y.
-Y represents our measure of interest (income).
+To describe what we are going to compare, we start with defining the two distributions.
+One represents the reference population $Y_{0}$ and the other the comparison population $Y$.
+$Y$ represents our measure of interest (income).
 A first visual approach is to compare the two probability density functions (PDF).
-The PDF is a function (fy) which describes the distribution of probability over the outcome set and is defined for all possible values of y.
+The PDF is a function $f(y)$ which describes the distribution of probability over the outcome set and is defined for all possible values of $y$.
 Also this function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%.
-Out of the comparison of the PDF, it is possible to see, which values of x are more and less probable.
-This allows allready to spot distributional differences over the whole scale of Y visualy.
+Out of the comparison of the PDF, it is possible to see, which values of $x$ are more and less probable.
+This allows already to spot distributional differences over the whole scale of $Y$ visually.
 \\
@@
 The PDF can be characterized by its cumulative distribution function (CDF).
-The CDF can be formulated as F(y), which represents the cumulative probability at point x, which is the probability that a randomly chosen value is less than or equal to y.
-The relative distribution of \Y to \Y_{0} is then defined as
+The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $x$, which is the probability that a randomly chosen value is less than or equal to $y$.
+The relative distribution of $Y$ to $Y_{0}$ is then defined as
 
@@
 
-R is obtained from \Y by transforming it by the CDF for \Y_{0}, \F_{0}. R measures therefore the relative rank of Y compared to Yo.
+R is obtained from $Y$ by transforming it by the CDF for $Y_{0}$, $\F_{0}$. $R$ therefore measures the relative rank of $Y$ compared to $Y_{0}$.
 
@@
 
-We can calculate the Probability Density Function (g(r)) of RD, where \r represents the proportion of values and F_{0}^{-1}(r) is the inverse cumulative distribution function, also called the quantile function.
-g(r) can be interpreted as a density Ratio, which is defined as the ratio of these two quantities of every percentile of the reference distribution [0,1].
-With a complete overlap of both distributions the Probability Density Function of the RD is 1 at every point of the PDF.
+We can calculate the Probability Density Function $g(r)$ of RD, where $r$ represents the proportion of values and $F_{0}^{-1}(r)$ is the inverse cumulative distribution function, also called the quantile function.
+$g(r)$ can be interpreted as a density ratio, which is defined as the ratio of these two quantities evaluated at every percentile of the reference distribution [0,1].
+With a complete overlap of both distributions the probability density function of the RD is 1 at every point of the PDF.
 On the other hand, values higher than 1 represent higher probabilities in the comparison distribution than in the references distribution at this specific point and values lower than 1 respectively represent lower probabilities.
@@
 What we get through the above transformation of two distributions is the overall relative probability density.
-But Differences between distributions can be divided into two basic components: changes in location and changes in shape.
+But differences between distributions can be divided into two basic components: changes in location and changes in shape.
 If the comparative distribution is a simple location-shifted version of the reference distribution, then the difference between the two distributions can be parsimoniously summarized by this shift.
-Differences that remain after a location adjustment are differences in ``shape'' (scale, skew and other distributional characteristics).
+Differences that remain after a location adjustment are differences in ``shape'' (scale, skewness and other distributional characteristics).
 When both types of shifts are operating, or when factors other than scale are changing in the shape component, we need a way to separate out the various effects.
 If we want to identify the effect of a location shift and separate it from other changes in the distribution, it is necessary to specify what scale this shift operates on.
-It is possible to adjust distribution by any measure of central tendency.
-We choose here a median location adjustment because the median is a natural, robust and scale invariant unit of measurement.
+It is possible to adjust distributions by any measure of central tendency.
+Here we choose a median location adjustment because the median is a natural, robust and scale invariant unit of measurement.
 Because our interest lies in analyzing distributional differences concerning the degree of inequality, we will focus in the results section on shape differences and look therefore at the relative distribution after the distributions are adjusted for location differences.
@@
 Distributional polarization is of particular interest in the study of inequality.
-However, common inequality indicator (for example Gini or Theil’s index) are not designed to distinguish between growth in the upper and lower tails.
-Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail), downgrading (increases in lower tail).
+However, common inequality indicators (for example Gini or Theil’s index) are not designed to distinguish between growth in the upper and lower tails.
+Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail) or downgrading (increases in lower tail).
 The polarization index plays the same role as the difference in Gini-indices, but it can be decomposed to compare the growth in the upper and lower tails.
@@
 MRP can be interpreted in terms of a proportional shift of mass in the distribution from more central to less central values.
-A value of 0.1, for example, is equivalent to a 10\% population shift from the center of the distribution to the upper and lower quartiles and the MRP is decomposable along the scale of y.
+A value of 0.1, for example, is equivalent to a 10\% population shift from the center of the distribution to the upper and lower quartiles and the MRP is decomposable along the scale of $y$.
 This makes it possible to compare the contribution of each section of the distribution to the overall polarization.
-A natural decomposition is the contributions made by components above and below the median (of g(r)).
+A natural decomposition is the contributions made by components above and below the median (of $g(r)$).
 
@@
 
-<<zero_descriptives>>=
+<<zero_descriptives, fig.cap='',fig.pos='H'>>=
 library(ggplot2)
-ggplot(df[df$steuerperiode>=1995&df$kanton=="CH",], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+theme_bw()
+ggplot(df[df$steuerperiode>=1995&df$kanton=="CH",], aes(x=steuerperiode,y=zeros))+geom_line()+theme_bw()+xlab("tax period")+ylab("share of non-taxed")
 #ggplot(df[df$steuerperiode>=1995,], aes(x=steuerperiode,y=zeros))+geom_line()+facet_wrap(~kanton)+geom_line(aes(y=G_steink),color="blue")+theme_bw()
@@
 
-The number of zeros increased slightly.
+The number of zeros increased slightly according to figure \ref{fig:zero_descriptives}.
 Different interpretations seems possible.
@@
 We know these cases had incomes between zero and the threshold (usually around 15.000 CHF for the period 1995-2010).
-In figure WITHWITHOUTZEROSTIMESERIES we show Gini coefficients calculated under the assumption of the zeros being zero.
+In figure \ref{fig:with_without_zeros_Gini} we show Gini coefficients calculated under the assumption of the zeros being zero.
 This is problematic for two reasons: first we know these units might earn more (up to the threshold) so Ginis calculated on this assumptions will be too high.
@@
 
-<<corrected_gini, dev='bmp'>>=
+<<corrected_gini, dev='bmp', fig.cap='',fig.pos='H'>>=
 df$J<-factor(df$steuerperiode)
@@
 
-Figure COEFPLOT can be read as follows: A positive coefficient (e.g.
+Figure \ref{fig:corrected_gini} can be read as follows: A positive coefficient (e.g.
 Schwyz) states that we measure higher Gini coefficients in periods with many zeros.

### 2014-07-02 paper/subdocuments/methods.Rnw
added subsubtitel methods.Rnw

@@
 % Diese Abschnitt muss ersetzt werden mit der von uns gewählten Vorgehensweise
+
+\subsubsection{Relative distribution (RD) methods }
 
@@
 Also this function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%.
-Out of the comparison of the PDF, it is possible to see, which values of $x$ are more and less probable.
+Out of the comparison of the PDF, it is possible to see, which values of $y$ are more and less probable.
 This allows already to spot distributional differences over the whole scale of $Y$ visually.
@@
 The PDF can be characterized by its cumulative distribution function (CDF).
-The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $x$, which is the probability that a randomly chosen value is less than or equal to $y$.
+The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $y$, which is the probability that a randomly chosen value is less than or equal to $y$.
 The relative distribution of $Y$ to $Y_{0}$ is then defined as
@@
 
-We can calculate the Probability Density Function $g(r)$ of RD, where $r$ represents the proportion of values and $F_{0}^{-1}(r)$ is the inverse cumulative distribution function, also called the quantile function.
+We can calculate the Probability Density Function $g(r)$ of R, where $r$ represents the proportion of values and $F_{0}^{-1}(r)$ is the inverse cumulative distribution function, also called the quantile function.
 $g(r)$ can be interpreted as a density ratio, which is defined as the ratio of these two quantities evaluated at every percentile of the reference distribution [0,1].
-With a complete overlap of both distributions the probability density function of the RD is 1 at every point of the PDF.
+With a complete overlap of both distributions the probability density function of the R is 1 at every point of the PDF.
 On the other hand, values higher than 1 represent higher probabilities in the comparison distribution than in the references distribution at this specific point and values lower than 1 respectively represent lower probabilities.
@@
 
-% RD based Summary Meausres
+% R based Summary Meausres
 Distributional polarization is of particular interest in the study of inequality.

### 2014-07-03 paper/subdocuments/data.Rnw
fixed issue with invisible tables (remove table env)

@@
 
-\begin{table}[h]
-\begin{tabular}{lcccc}
+\multicolinterrupt{
+\begin{table}[H]
+\caption{\label{datacomp}} 
+\begin{center}
+\begin{tabular}{lcc}
+\hline\hline
                                    & FTA       & Survey data\\
+\hline
 Statistical unit                   & Tax units (individuals/married couples)  & Households\\                                   
@@
 Comparison over time possible 	   & Long       & Short \\
-Cross country comparison possible? & limited   & possible
-
+Cross country comparison possible? & limited   & possible\\
+\hline
 \end{tabular}
+\end{center}
 \end{table}
-
+}
 %RF: damit haben wir WER, WANN, WAS, wollen wir noch ein WOMIT/WAS DARSTELLEN -> methodik -> reldist ? jaa

### 2014-07-03 paper/subdocuments/results.Rnw

@@
 
-\subsection{Defining income}
-
-\emph{Gini coefficients for taxable income, income before deductions and taxable income after federal tax}
-
-<<setup_gini_data>>=
+<<libraries, include=FALSE>>=
 library(foreign,quietly=TRUE, warn.conflicts=FALSE)
 library(ggplot2)
+library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
+library(reshape)
+# before loading reldist we load nlme and mgcv. else reldist prints library loading messages
+library(nlme,quietly=TRUE, warn.conflicts=FALSE)
+library(mgcv,quietly=TRUE, warn.conflicts=FALSE)
+library(reldist,quietly=TRUE,warn.conflicts=FALSE) #for reldist plots and inequality measures
+library(Hmisc,quietly=TRUE,warn.conflicts=FALSE) #for polTable function / Latex tables
+@
+
+\subsection{Defining income}
+
+\emph{Gini coefficients for taxable income, income before deductions and taxable income after federal tax}
+
+<<setup_gini_data>>=
 df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
@@
 <<setup_data_Brühlhart with and without taxed>>=
-library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
 # Ohne Nuller
@@
 <<equivalencescale, fig.cap='',fig.pos='H'>>=
-library(ggplot2)
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
@@
 <<setup_data_Brülhart_2003>>=
-library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
-library(foreign)
 bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
@@
 <<setup_data_Brülhart_2010>>=
-library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
-library(foreign)
 bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
@@
 <<reldist20032010, echo=FALSE, fig.cap='',fig.pos='H'>>=
-library(reldist, quietly=TRUE, warn.conflicts = FALSE)
 
@@
 # Setup functin to produce latex tables with polarization indices and difference in ginis
-polTable <- function(y, yo, ywgt=1, yowgt=1,ndig=2,title="Inequality Indices") {
-stopifnot(is.numeric(y), is.numeric(yo)) # all TRUE
+polTable <- function(y, yo, ywgt=1, yowgt=1,ndig=2,title="Inequality Indices",caption='',tablabel) {
+stopifnot(is.numeric(y), is.numeric(yo),exists("tablabel")) # all TRUE
 #y vector with new distribution
@@
 
-# load neccessary libraries
-library(reldist,quietly=TRUE,warn.conflicts=FALSE) #for Latex tables
-library(Hmisc,quietly=TRUE,warn.conflicts=FALSE)
-library(xtable,quietly=TRUE,warn.conflicts=FALSE)
-
 # Create vector of weights 1 if nothing else is defined
@@
 
-latex(format(c("Median Index"=med, "Lower Index"=lower, "Upper Index"=upper, "$\\Delta$ Gini"=dgini),digits=ndig),file="",title=title,table.env=FALSE)
+latex(format(c("Median Index"=med, "Lower Index"=lower, "Upper Index"=upper, "$\\Delta$ Gini"=dgini),digits=ndig),file="",title=title,where='H',label=tablabel,caption=caption)
 }
@@
 <<polindex20032010, results='asis'>>=
-polTable(bd.2010.ind,bd.2003.ind)
+polTable(bd.2010.ind,bd.2003.ind, tablabel="pol20032010")
 @
@@
 <<setup_data_ESTVspecialcases9394>>=
-library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
-library(ggplot2, quietly=TRUE, warn.conflicts = FALSE)
-#library(reldist)
+#library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
+#library(ggplot2, quietly=TRUE, warn.conflicts = FALSE)
+
 normal <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
@@
 <<polindexspecialcases9394, results='asis'>>=
-polTable(pooled_i,normal_i)
-@
-
-1993/94 a pooled data set of normal and special cases has a slightly higher density at both ends compared to data based in normal cases only.
+polTable(pooled_i,normal_i,tablabel="pol9394spec")
+@
+
+1993/94 a pooled data set of normal and special cases has a slightly higher density at both ends compared to data based in normal cases only (see figure \ref{fig:specialcases9394}).
 Special cases appear to have a slightly lower median income and their distribution is more skewed.
-Therefore special cases are more polarized than normal cases, i.e.
+Therefore special cases are more polarized than normal cases (see table \ref{pol9394spec}), i.e.
 striving away from the median (positive Median Index of 0.02).
-This tendency is more pronounces in the lower than upper part  of the distribution (Lower Index of 0.029 compared to Upper Index if 0.01).
+This tendency is more pronounced in the lower than upper part  of the distribution (Lower Index of 0.029 compared to Upper Index if 0.01).
 As the special cases consist of a broad mix of individuals it remains unclear which factor explains the differences of both distributions.
@@
 if they moved and stopped working.
-To get a more complete picture we can look how the two distributions relate to each other in 2010 (see FIGURE SPECIAL2010).
+To get a more complete picture we can look how the two distributions relate to each other in 2010 (see figure \ref{fig:specialcases10}).
 
@@
 <<setup_data_specialcases10>>=
-library(foreign)
-library(ggplot2)
+#library(foreign)
+#library(ggplot2)
 #library(reldist)
@@
 <<polindexspecialcases10, results='asis'>>=
-polTable(pooled_i,normal_i)
+polTable(pooled_i,normal_i,tablabel="pol2010spec")
 @
@@
 df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
-library(reshape)
+#library(reshape)
 dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
@@
 <<polindexzeros, results='asis'>>=
-polTable(bd.2010.ind,bd.0.2010.ind)
+polTable(bd.2010.ind,bd.0.2010.ind,tablabel="polzeros")
 @
@@
 <<polindexzeros2, results='asis'>>=
-polTable(bd.2010.ind,bd.0.2010.ind)
+polTable(bd.2010.ind,bd.0.2010.ind,tablabel="polzeros2")
 @
@@
 <<setup_data Brülhart_mitNuller_und_HABE_2010,warning=FALSE>>=
-library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
-library(foreign)
+#library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
+#library(foreign)
 bd.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
@@
 #####################################################
-library(reshape, quietly=TRUE, warn.conflicts=FALSE)
-library(dplyr, quietly=TRUE, warn.conflicts=FALSE)
+#library(reshape, quietly=TRUE, warn.conflicts=FALSE)
+#library(dplyr, quietly=TRUE, warn.conflicts=FALSE)
 try(habe0911<-read.table("C:/Users/Hackstutz/Dropbox/Ungleichheit/HABE/HABE091011/HABE091011_Standard_130717UOe.txt", header=TRUE),silent=TRUE)
@@
 <<BruelhartvsHABE,warning=FALSE, fig.cap='',fig.pos='H'>>=
-library(ggplot2)
-library(reldist, quietly=TRUE, warn.conflicts = FALSE)
+#library(ggplot2)
+#library(reldist, quietly=TRUE, warn.conflicts = FALSE)
 
@@
 plot(x = (density1$x), y = density1$y, type = "l",
-     xlab = "Einkommen", ylab = "density",
+     xlab = "income", ylab = "density",
      axes = FALSE,
@@
 
+From the discussion in the data section we would expect differences between the income distributions from survey and tax data.
+Within the FTA data we observe zeros for incomes below the threshold to be liable for federal tax while survey data might cover this range.
+On the other hand we expect underreporting from both lowest percentiles and highest percentiles (middle class bias) within the survey data.
+Figure \ref{fig:BruelhartvsHABE1} (left) however reveals a more critical issue related to tax data, that is the median location of income compared to survey data.
+Although we try to measure similar concepts of income, survey data shows a median (93.000 CHF) more than twice as big as tax data (44.600 CHF).
+The issue here is clearly the assumptions of household composition.
+While survey data is likely to capture the correct household composition, tax data can only approximate households by using marital status.
+Splitting the data in married and unmarried tax units underpins this argument.
+From figure \ref{fig:BruelhartvsHABE2} we can see that married tax units (FTA data) and household with married couples (survey data) are better (but still not perfectly) comparable.
+Figure \ref{fig:BruelhartvsHABE2} (right) shows the expected shape difference between the two distributions:  survey data has a bias towards the (median-adjusted) 80\% to 90\% percentile (of the FTA data distribution).
+Top percentiles are badly covered by survey data, suggesting that inequality measures based on survey data might underestimate inequality development that arises from changes in the incomes of the rich.
+Though, survey data can be of interest if one is interested in the lower 20\% of the income distribution.
+
+The bad household approximation within tax data needs to be addressed in further research. One possible way would be to link tax data to household ID's from the residential register.
+
 \subsection{Intertemporal comparison}
@@
  ### libs
-library(reshape, quietly=TRUE, warn.conflicts = FALSE)
-library(foreign)
-library(dplyr)
-library(reldist)
-library(ggplot2)
+#library(reshape, quietly=TRUE, warn.conflicts = FALSE)
 
@@
 names(all_gini_long) <- c("Year","Data","Gini")
+
+all_gini_long$annotation<-""
+all_gini_long$annotation[all_gini_long$Year==1939.5] <- "Tolle\nSteuererhöhung"
+all_gini_long$annotation[all_gini_long$Year==1955.5] <- "Tolle\nSteuersenkung"
+
+annotations <- data.frame("Year"=c(1939.5,1955.5),"label"=c("Tolle\nSteuererhöhung","Tolle\nSteuersenkung"))
+
 ggplot(data=na.omit(all_gini_long),
@@
   geom_line()+
+  geom_text(aes(label=annotation,y=0.29),color="#555555",size=3)+
+  geom_segment(aes(x = 1939.5, y = 0.295, xend = 1939.5, yend = Gini[Year==1939.5]),color="#aaaaaa") +
+  geom_segment(aes(x = 1955.5, y = 0.295, xend = 1955.5, yend = Gini[Year==1955.5]),color="#aaaaaa") +
   theme_bw()+

### 2014-07-03 paper/subdocuments/methods.Rnw
titel RD-section
added introduction to methods section

@@
 
-\subsubsection{Relative distribution (RD) methods }
+In the last section we described the advantages and drawback of taxdata discussing five aspects, which we regard as crucial concering the assessment of inequality.
+To get a feeling of the importance of these aspects, we exploit the FTA-tax data as far as possible and perform several insightful calculations adressing four of the five mentioned aspects.
+No further investigation is possible regarding aspect (1) concepts on measuring economic ressource.
+But for the other four thematic areas, we can provide deeper insight.
+In general our main strategy is to apply different possible concepts within one conceptional area (income, units, measurement, population coverage) while holding other conceptual differences constant.
+With this strategy we want to show, the sensitiveness the assessement of inequality is especially if one looks at time trends.\\
+
+To fulfill the above described task, we use two techniques.
+To assess the development over time, we calculate in generall gini-coefficents for all possible time points, allowing us the make time trends visible and to compare the result out of the FTA-tax data to other publications.
+For selected periodes we expand the analysis with relative distribution methods, which allows an indepth analysis of distributional differences and therefore catches the shortcomins of gini-coefficents.
+\\
+
+To calculate Gini-coefficents we follow common definitions, like it is describded for example in \citep{jann_einfuhrung_2005}.
+Because the relative distribution methods are not that widely know, we provide a short introduction of the important concepts in this section.
+We expand the methodsection with an extended analysis about the importance of non-taxed, because the issue of non-taxed is a especially thorny one.
+
+\subsubsection{Relative distribution (RD) framework}
 

### 2014-07-03 paper/subdocuments/methods.Rnw
minor changes methods.Rnw

--- 2014-07-02 paper/subdocuments/methods.Rnw
@@
-\subsection{Outline of applied methods}
+\subsection{Outline of research strategy and applied methods}
 
@@
 
-\subsubsection{Relative distribution (RD) methods }
+In the last section we described the advantages and drawback of tax-data discussing five aspects, which we regard as crucial concerning the assessment of inequality.
+To get a feeling of the importance of these aspects, we exploit the FTA-tax data as far as possible and perform several insightful calculations addressing four of the five mentioned aspects.
+No further investigation is possible regarding aspect (1) concepts on measuring economic resource.
+But for the other four thematic areas, we can provide deeper insight.
+In general our main strategy is to apply different possible concepts within one conceptional area (income, units, measurement, population coverage) while holding other conceptual differences constant.
+With this strategy we want to show, the sensitiveness the assessment of inequality is especially if one looks at time trends.\\
+
+To fulfill the above described task, we use two techniques.
+To assess the development over time, we calculate in general gini-coefficients for all possible time points, allowing us the make time trends visible and to compare the result out of the FTA-tax data to other publications.
+For selected periods we expand the analysis with relative distribution methods, which allows an in-depth analysis of distributional differences and therefore catches the shortcomings of gini-coefficients.
+\\
+
+To calculate Gini-coefficients we follow common definitions, like it is described for example in \citep{jann_einfuhrung_2005}.
+Because the relative distribution methods are not that widely know, we provide a short introduction of the important concepts in this section.
+We additional expand the method section with an extended analysis about the importance of non-taxed, because the issue of non-taxed is a especially thorny one.
+
+\subsubsection{Relative distribution (RD) framework}
 
@@
 Subject of the comparison can also be two distributions describing the same population, but stemming from two different data sources like survey or tax data or even comparison of the same source/population but for different time points, like the income distribution out of tax data for Switzerland today compared to an earlier time point.
-To describe what we are going to compare, we start with defining the two distributions.
+To describe how the two distributions are gone be transformed into a relative distribution, we start with defining the two distributions.
 One represents the reference population $Y_{0}$ and the other the comparison population $Y$.
-$Y$ represents our measure of interest (income).
+$x$ represents our measure of interest (income).
 A first visual approach is to compare the two probability density functions (PDF).
-The PDF is a function $f(y)$ which describes the distribution of probability over the outcome set and is defined for all possible values of $y$.
+The PDF is a function $f(x)$ which describes the distribution of probability over the outcome set and is defined for all possible values of $x$.
 Also this function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%.
-Out of the comparison of the PDF, it is possible to see, which values of $y$ are more and less probable.
+Out of the comparison of the PDF, it is possible to see, which values of $x$ are more and less probable.
 This allows already to spot distributional differences over the whole scale of $Y$ visually.
@@
 The PDF can be characterized by its cumulative distribution function (CDF).
-The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $y$, which is the probability that a randomly chosen value is less than or equal to $y$.
+The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $x$, which is the probability that a randomly chosen value is less than or equal to $x$.
 The relative distribution of $Y$ to $Y_{0}$ is then defined as

### 2014-07-03 paper/subdocuments/methods.Rnw
minor changes methods.Rnw

@@
 The PDF is a function $f(x)$ which describes the distribution of probability over the outcome set and is defined for all possible values of $x$.
-Also this function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%.
+This function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%.
 Out of the comparison of the PDF, it is possible to see, which values of $x$ are more and less probable.
-This allows already to spot distributional differences over the whole scale of $Y$ visually.
+This already allows to spot distributional differences over the whole scale of $x$ visually.
 \\
@@
 The PDF can be characterized by its cumulative distribution function (CDF).
-The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $x$, which is the probability that a randomly chosen value is less than or equal to $x$.
+The CDF can be formulated as $F(x)$, which represents the probability that a randomly chosen value is less than or equal to $x$.
 The relative distribution of $Y$ to $Y_{0}$ is then defined as
@@
 
-R is obtained from $Y$ by transforming it by the CDF for $Y_{0}$, $\F_{0}$. $R$ therefore measures the relative rank of $Y$ compared to $Y_{0}$.
+R is obtained from $Y$ by transforming it by the CDF for $Y_{0}$, $F_{0}$. $R$ therefore measures the relative rank of $Y$ compared to $Y_{0}$.
 
@@
 % Median and shape differences
-What we get through the above transformation of two distributions is the overall relative probability density.
+What we got through the above transformation of two distributions is the overall relative probability density.
 But differences between distributions can be divided into two basic components: changes in location and changes in shape.
@@
 Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail) or downgrading (increases in lower tail).
-The polarization index plays the same role as the difference in Gini-indices, but it can be decomposed to compare the growth in the upper and lower tails.
+The polarization index developed by \citep{Handock and Morris 1999} adresses this issue, because it is decomposable to distinguish differences in the upper and lower tails.
 Because it is based on the relative distribution it provides a simple link between what is observed in the graphical display and what is measured by the numerical summary.
@@
 
-The median relative polarization index (MRP) is then the mean absolute deviation around the median of the location-adjusted relative distribution, scaled to produce an index that varies between -1 and 1.
+The median relative polarization index (MRP) is defined as the mean absolute deviation around the median of the location-adjusted relative distribution, scaled to produce an index that varies between -1 and 1.
 Given the scaling, a value of zero represents no differences in distributional shape; positive values represent more polarization (increases in the tails of the distribution); and negative values represent less polarization (convergence towards the center of the distribution).

### 2014-07-03 paper/paper.Rnw
introduced figure environments
added command to interrupt/span multicol
viele mini fixes aus gemeinsamer session

@@
 \usepackage{multicol} % Used for the two-column layout of the document
+\usepackage{float} %we need this to use figure environments within multicol.
+knitr figure chunks need to include fig.pos='H'.
+'H' stems from package float and deactivates floating (which multicol does not like)
 \usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
@@
 
+%%%%%% new command to break multicol environment from time to time
+\newcounter{tempcolnum}
+\makeatletter
+\newcommand{\multicolinterrupt}[1]{% Stuff to span both rows
+\setcounter{tempcolnum}{\col@number}
+\end{multicols}
+#1%
+\begin{multicols}{\value{tempcolnum}}
+}
+\makeatother
+%%%%%%%
+
 \usepackage{titlesec} % Allows customization of titles
@@
 \fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
+\newdateformat{mydate}{\monthname[\THEMONTH] \THEYEAR}
+\fancyhead[C]{Assessing income inequality with tax data $\bullet$ \mydate $\bullet$ Working paper} % Custom header text
 \fancyfoot[RO,LE]{\thepage} % Custom footer text
@@
 
-\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}} % Article title
+\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing income inequality with tax data - Switzerland from 1917 to 2010}} % Article title
 

### 2014-07-03 paper/subdocuments/abstract.Rnw
viele mini fixes aus gemeinsamer session

@@
 \begin{abstract}
-There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (\citealt{cooperation_divided_2011}, \citealt{gornick_income_2013}). 
+There is empirical evidence that economic inequality increased in the majority of western countries over the last decades (\citealt{oecd_divided_2011}, \citealt{gornick_income_2013}). 
 In Switzerland, however, the development is unclear, as there is evidence for trends in both directions.
@@
 We present common and new strategies to handle tax data specific methodological difficulties and compare results out of aggregated federal tax statistics to results from the Household Budget Survey (HBS).
-We can show to which extend survey data underestimate inequality in income. 
-Following the results out of the tax-data Switzerland experienced in slight rise in inequality in recent years, similar to other western countries, but only because of rise in upper percentiles of the income distribution.
+We find theoretical inconsistencies in both tax and survey data which both result in an underestimation of inequality.
+Following the results from tax-data Switzerland experienced in slight rise in inequality in recent years, similar to other western countries due to a polarization of the income distribution.
 

### 2014-07-03 paper/subdocuments/data.Rnw
viele mini fixes aus gemeinsamer session

@@
 
-\section{Data, Measurement concepts and Methods}
+\section{Data and measurement concepts}
 
@@
 Based on a review on the literature about the measurement concepts in an ideal world, we discuss the advantages and shortcomings of tax data compared to other data sources - namely survey data.
-We describe and explain methods and techniques needed to construct time-series of income inequality measures for Switzerland.
-We will present these time-series in the result section and evaluate the use of the FTA Tax Data to assess the development of inequality in Switzerland
 
@@
 This exempts all tax units with taxable income below a certain threshold (e.g.
-15.000 CHF in 2010).
+CHF 16.900 for singles in 2010).
 Furthermore the FTA differentiates between two groups of tax units, so called normal cases and special cases.
@@
 \emph{Concepts on measuring economic resources}  \\
-
 Most studies on inequality focus on income inequality solely.
@@
 It should be noted, that the Federal Tax Administration (FTA) publishes statistics on income and wealth but it is not possible to analyze the joint distribution on the individual or household level.
-Also measures of consumption are largely missing in tax data, albeit deductions can in some sense be understood as mandatory consumptions.
+Also measures of consumption are largely missing in tax data, albeit deductions can in some sense be understood as mandatory consumptions.\\
 
@@
 
-\begin{figure}
+\multicolinterrupt{
+\begin{figure}[H]
 \centering
@@
 \end{figure}
+}
 
@@
 \emph{Statistical units} \\
-
 The agreed standard on the statistical units, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013}.
-Indeed it is the individuals, who receive income, own assets and experience economic well-being, but their possibility to do so, is strongly tied to the concept of household.
-This comprises all persons under the same housing arrangement.
+Indeed it are individuals, who receive income, own assets and experience economic well-being, but their possibility to do so, is strongly tied to the concept of household.
+A household comprises all persons under the same housing arrangement.
 The basic underlying assumption for collecting data on household level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children).
@@
 This doesn't imply, that those couples live together, as it is needed to satisfy the definition of a household.
-On the other hand, is it quite likely that more than one tax unit live in the same household (unmarried/unregistered couples, see \citet[99]{muller_vermogenslage_2014}).
+On the other hand, is it quite likely that more than one tax unit lives in the same household (unmarried/unregistered couples, see \citet[99]{muller_vermogenslage_2014}).
 It is therefore not directly possible to elicit households and household income from tax data.
 This might influence the assessment of inequality development, taking into account the change from traditional household and family structures over the last century.
+\\
 
@@
 (1) ``\textbf{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textbf{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
-doubling of all income) (3) ``\emph{weak principle of transfers}'' or ``\emph{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality.
+doubling of all income) (3) ``\textit{weak principle of transfers}'' or ``\textit{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality.
 However, several drawbacks are reported in the literature.
@@
 The recent wave of tax-data studies do not report Gini-coefficents.
-Rather top income shares are informed on, which are calculated not only with tax data, but together with external sources to produce the population and personal income control totals (RF 16.06.2014: hä, was?
-gehts da um die Randverteilung der Bevölkerung?).
+Rather top income shares are informed on, which are calculated not only with tax data, but corrected using marginal distributions of income and population using external sources (census data and national accounts).
 This procedure ensures, that the inequality measure is not biased because of non-fillers, who do not appear in tax statistics.
@@
 
-Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow better analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading).
-The polarizationindex is developed by \citep{Handock and Morris 1999} allows for this.
+Newer branches of inequality studies emphasize the need for broader measures of inequality, which allow more detailed analyses about the change of inequality and namely statements about the area of change (downgrading/upgrading).
+The polarization index developed by \citep{handcock_relative_1999} is capable of doing this.
 Recently this index was applied in the work of \citep{alderson_exactly_2005} and \citep{gornick_how_2013}.
@@
 
-The literature mentions several other metrics with desired properties we will not discuss here (see for example \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}).
+The literature mentions several other metrics with desired properties we will not discuss here (see for example \citet{cowell_chapter_2000} and \citet{hao_assessing_2010}).\\
 
@@
 
-The issue of incomplete coverage is less dramatical with tax data.
+The issue of incomplete coverage is less dramatic with tax data.
 Essentially every permanent resident in Switzerland over 18 years of age (respectively 20 years of age prior to 1996) is taxed on a yearly base (or every two years before the change of the tax system).
-Essentially this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distribution.
+Basically this leads to a full representation of the adult population of Switzerland and a complete coverage of the income distribution.
 This includes a separation of normal cases, which embrace the majority of taxpayers, and the special cases, which cover (not only) foreign nationals living in Switzerland but with a yearly or any other temporary resident permit only.
@@
 Special attention has to be paid to tax units with none or very low incomes.
-Even though they have to hand in a tax return, their income does not show up in the statistics if their income after deductions falls below 15'000 CHF and they are therefore not taxed with direct federal taxes.
+Even though they have to hand in a tax return, their income does not show up in the statistics if their income after deductions falls below the threshold (e.g.
+CHF 16.900 for singles in 2010) and they are therefore not taxed with direct federal taxes.
 This is possible for normal and special cases alike.
@@
 
-\subsection{Comparison of tax data and other data sources - advantages and shortcomings}
+\emph{Comparison of tax data and other data sources - advantages and shortcomings}
+
+Based on this sections previous discussion table~\ref{datacomp} summarizes important differences between tax and survey data related to analyses of income inequality.
 
@@
 \begin{table}[H]
-\caption{\label{datacomp}} 
+\caption{Comparison of tax and survey data\label{datacomp}} 
 \begin{center}
@@
 \hline\hline
-                                   & FTA       & Survey data\\
+                                   & Tax data       & Survey data\\
 \hline
-Statistical unit                   & Tax units (individuals/married couples)  & Households\\                                   
-Concepts of economic resources  & Data-driven   & Theory-driven\\
+Statistical unit                   & tax units (individuals/married couples)  & households\\                                   
+Concepts of economic resources  & data-driven   & theory-driven\\
 Population coverage & total population & sample \\ 
-Top incomes covered                 & Yes   & No\\
+Top incomes covered                 & yes   & no\\
 Low incomes covered?               & no        & yes\\
-Comparison over time possible 	   & Long       & Short \\
+Comparison over time possible 	   & long       & short \\
 Cross country comparison possible? & limited   & possible\\
@@
 To define a standard of measuring economic resources and related inequality we discussed different dimensions of demands the data needs to meet.
-To sum up, ideally we want to measure \emph{disposable income} for \emph{all} swiss \emph{households} yearly for an extended period of time.
+To sum up, ideally we want to measure \textit{disposable income} for \textit{all} swiss \textit{households} yearly for an extended period of time.
 Tax data has one weakness when it comes to the ideal income measure as we need to work with ``taxable income'' which is designed to serve taxation purposes.
 However, both concepts -- what needs to be taxed and what can be spent -- are related as both address a distinction between necessary and voluntary expenses.
-By subtracting taxes from taxable income we are confident to satisfyingly come close the theoretically ideal income measure of disposable income.
+By subtracting federal taxes from taxable income we are getting closer to the theoretically ideal income measure of disposable income.
+However we can not subtract the more important cantonal and communal taxes.
 The second most important drawback with tax data is that it does not adequately address households.
@@
 This ``classic'' household setup however became less common in the last decades so we need to assume statements about household inequality based on taxed data became more and more biased.
-The swiss law seems to be slightly outdated in this context.
+%The swiss law seems to be slightly outdated in this context.
 Finally however, another advantage of tax data (over survey data) is indeed the observation period.

### 2014-07-03 paper/subdocuments/introduction.Rnw
viele mini fixes aus gemeinsamer session

--- 2014-06-30 paper/subdocuments/introduction.Rnw
@@
 Looking for official data, three main sources has to be mentioned, which can be considered as de facto official data sources: EU-SILC, HBS and LIS-data.
-Figure XY shows the results stemming from this three sources while looking at Gini of equivalised disposable income.
+Figure \ref{fig:SwissinequalityPlot} shows the results stemming from this three sources while looking at Gini of equivalised disposable income.
 Up to the day, EU-SILC or Statistics on Income and Living Conditions is the main source used for policy monitoring at EU-level.
@@
 Since 2000 the survey has been conducted on a continuous basis, which allows to look at a consistent time series from 2000 to 2011.
-As it can be seen from graph XY the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period.
+As it can be seen from figure \ref{fig:SwissinequalityPlot} the trend is rather stable.\footnote{figures shown in the graph were calculated out of the original datasets, which were kindly provided by Swiss Federal Statistical Office.} Both time-series (SILC and HBS) cover a relatively short time period.
 A longer period is covered in the LIS-Data-set (1982-2004).
 Data-provider for the LIS Data is the Swiss Federal Statistical Office too.
-In contrast the the aforementioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000, 2002, 2004).
+In contrast to the aforementioned surveys the LIS-data is harmonized out of three surveys: Swiss Income and Wealth Survey (1982), Swiss Poverty Survey (1992) and the Income and Consumption survey (2000, 2002, 2004).
 All in all the LIS dataset contains the longest time series on inequality for Switzerland.
@@
 
-<<Swiss inequality Plot>>=
+<<SwissinequalityPlot, fig.cap='Trends in income inequality',fig.pos='H'>>=
 library(ggplot2)
@@
 swissGini_long<-melt(swissGini,id="Year")
+swissGini_long$Source<-swissGini_long$variable
 ggplot(data=swissGini_long,
-  aes(y=value,x=Year),shape=variable)+
+  aes(y=value,x=Year),shape=Source)+
   ylab("Gini of equivalised disposable income") +
@@
   scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
-  geom_point(aes(shape=variable),size=3)+
-  geom_line(data=swissGini_long[!is.na(swissGini_long$value),],aes(linetype=variable))+
+  geom_point(aes(shape=Source),size=3)+
+  geom_line(data=swissGini_long[!is.na(swissGini_long$value),],aes(linetype=Source))+
   theme_bw()
@@
 Whereas the aforementioned publications focused on disposable household income from survey data, the revival of tax-data-inequality studies lead to fruitful insights for Switzerland as well.
-\citet{dell_income_2007} used tax data from the Federal Tax Administration to assess the development of concentration of the highest incomes and wealth (top-shares).
+\citet{dell_income_2007} used tax data from the Federal Tax Administration to assess the concentration of the highest incomes and wealth (top-shares) over time.
 In contrast to most other examined countries, Switzerland did not experience a reduction in income and wealth concentration from the pre-First World War period to the decades following the second World War (up to 1996).
@@
 Third, different income concepts and different units of analysis were used.
-As it is shown by \citet{modetta_einkommensungleichheit_2012} income distribution is strongly affected by governmental redistribution, reducing inequality substantially.
+As it is shown by \citet{modetta_einkommensungleichheit_2012} the income distribution is strongly affected by governmental redistribution, reducing inequality substantially.
 With the focus on tax data the change in institutional settings is not covered.
@@
 Building on recent developments in the field of inequality research, we assess the suitability of the publicly accessible tax data to report inequality and its changes over time.
-First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being.
-Second, we summarize and apply tax data specific techniques to construct suitable measures of inequality.
-We expand the given set by an in depth discussion of a newly applied step to handle the incomplete coverage in tax data statistics.
-Third, we compare our results to results from a relevant Inequality-Survey in Switzerland (HBS) to assess the bias through data source.
+First of all this includes a discussion of the accessible measures in context of a reflection about the state of the art conceptualization of economic resources as an indicator for economic well-being (see section \textit{II.
+Data and measurement concepts}).
+Second, we examine the role of methodological decisions along with an in-depth distributional analysis (see sections \textit{III.
+Methods} and \textit{IV.
+Results}).
+Against the background of all methodological issues we discuss the long-term inequality development in Switzerland -- as well as more recent periods in detail -- using all relevant data.
 

### 2014-07-03 paper/subdocuments/methods.Rnw
viele mini fixes aus gemeinsamer session

@@
+\section{Methods}
+
 \subsection{Outline of research strategy and applied methods}
@@
 To fulfill the above described task, we use two techniques.
-To assess the development over time, we calculate in general gini-coefficients for all possible time points, allowing us the make time trends visible and to compare the result out of the FTA-tax data to other publications.
-For selected periods we expand the analysis with relative distribution methods, which allows an in-depth analysis of distributional differences and therefore catches the shortcomings of gini-coefficients.
+To assess the development over time, we calculate in general Gini-coefficients for all possible time points, allowing us the make time trends visible and to compare the result out of the FTA-tax data to other publications.
+For selected periods we expand the analysis with relative distribution methods, which allows an in-depth analysis of distributional differences and therefore catches the shortcomings of Gini-coefficients.
 \\
@@
 To calculate Gini-coefficients we follow common definitions, like it is described for example in \citep{jann_einfuhrung_2005}.
-Because the relative distribution methods are not that widely know, we provide a short introduction of the important concepts in this section.
+Because the relative distribution methods are not that widely known, we provide a short introduction of the important concepts in this section.
 We additional expand the method section with an extended analysis about the importance of non-taxed, because the issue of non-taxed is a especially thorny one.
@@
 The RD-framework is based on the concept of a ``relative distribution'', a transformation of the data from two distributions into a single distribution that contains all of the information necessary for scale-invariant comparison.
-This allows to make distribution differences ``visual'' in an elegant way and it is also a base for summary statistics, which are more sensitive to detailed theoretical hypotheses in contrast to other measures like the gini-coefficient or top income-shares, which inform either about the whole population or only about the top part of the distribution.\\
+This allows to make distribution differences ``visual'' in an elegant way and it is also a base for summary statistics, which are more sensitive to detailed theoretical hypotheses in contrast to other measures like the Gini-coefficient or top income-shares, which inform either about the whole population or only about the top part of the distribution.\\
 
@@
 
-R is obtained from $Y$ by transforming it by the CDF for $Y_{0}$, $F_{0}$. $R$ therefore measures the relative rank of $Y$ compared to $Y_{0}$.
+$R$ is obtained from $Y$ by transforming it by the CDF for $Y_{0}$, $F_{0}$. $R$ therefore measures the relative rank of $Y$ compared to $Y_{0}$.
 
@@
 $g(r)$ can be interpreted as a density ratio, which is defined as the ratio of these two quantities evaluated at every percentile of the reference distribution [0,1].
-With a complete overlap of both distributions the probability density function of the R is 1 at every point of the PDF.
+With a complete overlap of both distributions the probability density function of the $R$ is 1 at every point of the PDF.
 On the other hand, values higher than 1 represent higher probabilities in the comparison distribution than in the references distribution at this specific point and values lower than 1 respectively represent lower probabilities.
@@
 Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail) or downgrading (increases in lower tail).
-The polarization index developed by \citep{Handock and Morris 1999} adresses this issue, because it is decomposable to distinguish differences in the upper and lower tails.
+The polarization index developed by \citep{Handock and Morris 1999} addresses this issue, because it is decomposable to distinguish differences in the upper and lower tails.
 Because it is based on the relative distribution it provides a simple link between what is observed in the graphical display and what is measured by the numerical summary.
@@
 This makes it possible to compare the contribution of each section of the distribution to the overall polarization.
-A natural decomposition is the contributions made by components above and below the median (of $g(r)$).
+A natural decomposition is the contributions made by components above (upper polarization index, URP) and below (lower polarization index, LRP) the median (of $g(r)$).
 
@@
 
-<<zero_descriptives, fig.cap='',fig.pos='H'>>=
+<<zero_descriptives, fig.cap='Share of tax units below federal tax liability threshold',fig.pos='H'>>=
 library(ggplot2)
@@
 We know these cases had incomes between zero and the threshold (usually around 15.000 CHF for the period 1995-2010).
-In figure \ref{fig:with_without_zeros_Gini} we show Gini coefficients calculated under the assumption of the zeros being zero.
+In figure \ref{fig:with_without_zeros_Gini} we show Gini-coefficients calculated under the assumption of the zeros being zero.
 This is problematic for two reasons: first we know these units might earn more (up to the threshold) so Ginis calculated on this assumptions will be too high.
-Second we assume the zero group to be perfectly homogenous which on the contrary has a lowering effect on the Gini coefficient.
+Second we assume the zero group to be perfectly homogeneous which on the contrary has a lowering effect on the Gini coefficient.
 Therefore we must at least assume a distribution for the zeros that has some variance and ranges from zero to the threshold.
@@
 The second approach investigates whether the FTA time series is suitable to analyze trends of inequality even if the inequality measures are biased.
-We can do this by regressing the inequality measure (here the Gini coefficient) on the share of zeros.
+We can do this by regressing the inequality measure (here the Gini-coefficient) on the share of zeros.
 Simply spoken we test if cutting a larger (left) tail of the distribution changes the shape of the distribution.
-We include Gini coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time)}.
+We include Gini-coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time)}.
 
-<<corrected_gini, dev='bmp', fig.cap='',fig.pos='H'>>=
+<<corrected_gini, fig.cap='Bias variation by time and cantons',fig.pos='H'>>=
 df$J<-factor(df$steuerperiode)

### 2014-07-03 paper/subdocuments/results.Rnw
viele mini fixes aus gemeinsamer session

@@
 
-\emph{Equivalence scale}
+\subsection{Statistical units -- Equivalence scale}
 
@@
 
+par(mfrow=c(1,2))
 # Prop-Density Plot
@@
 abline(h=1,lty=2)
+
+par(mfrow=c(1,1))
 @
@@
 % Es ist mir aktuell unklar bei welchem Datensazt jetzt die Perzentile mit imputierten Nullern vorliegt.
-
-<<with_without_zeros_RelDist_mitESTV,warning=FALSE, fig.cap='',fig.pos='H'>>=
-
-## Nicht fertig angepasst!!!
-estvSteink<-(read.csv("../../data/ginis_und_perzentile_normal.csv", header=TRUE,sep="\t"))
-estvSteink<-filter(estvSteink,kanton=="CH")
-start <- which(names(estvSteink)=="p1")
-end <- which(names(estvSteink)=="p9999")
-percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
-estvReink<- data.frame(t(estvReink[1,start:end]),percentile)
-
-names(bd.2010)[1]=c("inc")
-bd.0.2010<- data.frame(t(bd.0.2010[1,start:end]),percentile)
-names(bd.0.2010)[1]=c("inc")
-
-bd.0.2010 <- rbind(c(0,0),bd.0.2010)
-ival<-bd.0.2010$inc
-perval<-bd.0.2010$percentile
-bd.0.2010.ind<-unlist(sapply(2:length(ival),function(i){
-  len=(perval[i]-perval[i-1])*100
-  seq(ival[i-1],ival[i],length.out=len)
-}))
-
-bd.2010 <- rbind(c(0,0),bd.2010)
-ival<-bd.2010$inc
-perval<-bd.2010$percentile
-bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
-  len=(perval[i]-perval[i-1])*100
-  seq(ival[i-1],ival[i],length.out=len)
-}))
-
-# Prop-Density Plot
-density1 <- density(log(bd.2010.ind))
-plot(x = (density1$x), y = density1$y, type = "l",
-     xlab = "taxable income", ylab = "density"
-     #axes = FALSE,
-     #xlim = c(-50, 200),
-     #ylim=c(0,0.0116031)
-     )
-title(main="all tax units vs without non-taxed",cex=0.6)
-axis(side = 1)
-axis(side = 2)
-fig1legend <- list(x=c(6,6),y=c(0.6,0.6))
-legend(fig1legend,lty=1:2, bty="n",
-       legend=c("without non-taxed","all tax units"))
-density2 <- density(log(bd.0.2010.ind))
-lines(x = (density2$x), y = density2$y, type = "l",lty=2)
-
-# Relative Distribution
-
-gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
-               smooth=0.4, ci=FALSE,
-               show="residual",
-               bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,2), ylab="",
-               discrete=FALSE,
-               xlab="Proportion of 2010 Distribution (including non-taxed)")
-title(main=paste("Effect of different shape"))
-abline(h=1,lty=2)
-@
-
-<<polindexzeros2, results='asis'>>=
-polTable(bd.2010.ind,bd.0.2010.ind,tablabel="polzeros2")
-@
 

### 2014-07-03 paper/subdocuments/methods.Rnw
titel-levels method.Rnw

@@
 \section{Methods}
-
-\subsection{Outline of research strategy and applied methods}
 
@@
 
-\subsubsection{Relative distribution (RD) framework}
+\subsection{Relative distribution (RD) framework}
 
@@
 
-\subsubsection{Non-taxed}
+\subsection{Non-taxed}
 Some tax units are missing in the FTA data as they do not qualify for taxation below a certain income threshold.

### 2014-07-04 paper/subdocuments/results.Rnw
super crude benplot text + timeline additions

@@
 estv <- within(estv, {
-    steink[steuerperiode==1918]<-(eink_ll[steuerperiode==1918]+eink_ul[steuerperiode==1918])/2*anz_pflichtige[steuerperiode==1918]
+    steink[bemessungsjahre==1917]<-(eink_ll[bemessungsjahre==1917]+eink_ul[bemessungsjahre==1917])/2*anz_pflichtige[bemessungsjahre==1917]
 })
@@
   filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %.% 
-  select(steuerperiode, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %.%
-  group_by(factor(steuerperiode)) %.%
+  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %.%
+  group_by(factor(bemessungsjahre)) %.%
   mutate(mean_steink=steink/anz_pflichtige) %.%
@@
 estv_gini$Year<-as.numeric(as.character(estv_gini$Year))
-estv_gini$Year[estv_gini$Year<2003] <- estv_gini$Year[estv_gini$Year<2003]-2
 
@@
 
-all_gini_long$annotation<-""
-all_gini_long$annotation[all_gini_long$Year==1939.5] <- "Tolle\nSteuererhöhung"
-all_gini_long$annotation[all_gini_long$Year==1955.5] <- "Tolle\nSteuersenkung"
+all_gini_long$annotation=""
+all_gini_long$ypos=-1
 
-annotations <- data.frame("Year"=c(1939.5,1955.5),"label"=c("Tolle\nSteuererhöhung","Tolle\nSteuersenkung"))
+all_gini_long<-within(all_gini_long, {
+annotation[Year==1921] <- "different data source: cantonal registers"
+annotation[Year==1940] <- "increased taxation"
+annotation[Year==1945.5] <- "increased effort to fight tax evasion"
+annotation[Year==1967.5] <- "wide-ranging tax amnesty"
+ypos[Year==1921] <- 0.27
+ypos[Year==1940] <- 0.305
+ypos[Year==1945.5] <- 0.29
+ypos[Year==1967.5] <- 0.285
+
+})
 
@@
   geom_line()+
-  geom_text(aes(label=annotation,y=0.29),color="#555555",size=3)+
-  geom_segment(aes(x = 1939.5, y = 0.295, xend = 1939.5, yend = Gini[Year==1939.5]),color="#aaaaaa") +
-  geom_segment(aes(x = 1955.5, y = 0.295, xend = 1955.5, yend = Gini[Year==1955.5]),color="#aaaaaa") +
+  geom_text(aes(label=annotation,y=ypos),color="#555555",size=3)+
+  geom_segment(aes(x = 1921, y = 0.27, xend = 1921, yend = Gini[Year==1921]),color="#aaaaaa") +
+  geom_segment(aes(x = 1940, y = 0.305, xend = 1940, yend = Gini[Year==1940]),color="#aaaaaa") +
+  geom_segment(aes(x = 1945.5, y = 0.29, xend = 1945.5, yend = Gini[Year==1945.5]),color="#aaaaaa") +
+  geom_segment(aes(x = 1967.5, y = 0.285, xend = 1967.5, yend = Gini[Year==1967.5]),color="#aaaaaa") +
   theme_bw()+
@@
 @
+
+Figure\ref{fig:benplot} displays the most relevant Gini-coefficients that can be calculated for Switzerland.
+Although we can not adjust the Gini-coefficients to be perfectly valid, we can discuss the picture against the background of our analyses.
+For the periods before the second world war it is difficult to draw secure conclusions because these data points are based on unreliable data.
+1933 only 13.7\% of the population filled in a tax form \citep{dell_income_2005}.
+During the war we see a tendency towards lower income inequality.
+This might be attributable to a changed data base as the amount of non-fillers decreased during the war and shortly after.
+The period after the world war is characterized by strong economic growth as well as an increase in inequality.
+Our interpretation is that high income percentiles disproportionately profited from the economic upturn.
+After the oil crises there were alternating phases of social welfare expansion and economic upturns.
+
+The more interesting part of the picture are the years around and past the millennium.
+Between 1997/98 and 2003 we see a gap in the FTA series caused by the change in the tax system.
+Our approach to impute the gap is to interpolate the income sums of all FTA-reported income brackets.
+This is fruitful because cantons switched the tax system in different years so we gain at least some information about the trend within the gap.
+The spike 2001 can be explained by tax tricks: within the period the tax system changed, individuals were able to save taxes by shifting parts of their incomes into this period.
+
+After the financial crisis most of the data sources (except EU SILC) state decreasing inequality.
+Our interpretation is that top percentiles lost relevant parts of their incomes from capital investments.

### 2014-07-04 paper/subdocuments/discussion.Rnw
fixed encoding errors, switched to utf8
discussion first draft

@@
 
+In this paper we attempted to address four methodological issues related to inequality research using tax data.
+Can we measure what we want to measure?
+Do data include the cases we are interested in?
+The paper revealed tax data afflicted with numerous problems when it comes to analyzing inequality.
+On the pro side tax data is not a sample but covers the full population liable to federal tax.
+The documentation of tax data for almost 100 years makes it the only option if the researcher is interested in long term analyses of swiss income inequality, take it or leave it.
+On the contra side the researcher needs to keep in mind several limitations.
+According to our analyses we can give advice which issues are more or less serious.
+The biggest issue is that tax data does not capture households.
+In comparison to survey data we have seen that tax data extremely overestimates the share of singles/single households if we assume marital status to be a proxy for household composition.
+Luckily there is a way to fix this shortcoming for further research.
+For analyses of household income inequality, the researcher should gather household information, e.g.
+by linking tax data to the population register.
+The second most important drawback of tax data is the non-reporting of the non-taxed (low income) tax units.
+The best solution here would be to gather data directly from the FTA where these information are available, although not published.
+As a minimum the researcher should keep an eye on the share of population ``disappearing'' below the threshold.
+The larger the share, the more likely it is to overlook increases in inequality if tax data is used without these cases.
+Minor issues are whether the researcher looks at all cases (including special cases) the FTA may provide or only normal cases.
+However this issue became more relevant in recent years as high income special cases became more frequent.
+Our recommendation is to use both normal and special cases wherever possible but it is not an important issue.
+It is also not very important for the researcher whether the researcher looks at equivalized or non-equivalized incomes as long as the exact household structure is unknown.
+A more fruitful approach is to do separate analyses for married and unmarried tax units and treat the latter with healthy portion of caution.
+
 % Hier könnte allenfalls der Bogen zur Theorie gespannt werden
@@
 %Methodological Main findings 
-%(1) Tax data is useful because it’s possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
-It’s superior to survey data, because the latter is affected by non-responded bias.
-Tax data includes data about the whole population … [was finden wir wegen den Nullern raus] .
+%(1) Tax data is useful because it's possible to construct consistent measures over time for income and wealth, which is useful to asses changes over time and conducted studies about impact of structural changes.(2) Concerning accuracy of inequality tax data has advantages and shortcomings.
+It's superior to survey data, because the latter is affected by non-responded bias.
+Tax data includes data about the whole population [was finden wir wegen den Nullern raus] .
 or at least it should (tax evasion).(3) The FTA-Tax data has different shortcomings, which cannot be handled like a measure of income and wealth which (a) cannot be corrected for household size and (b) is neither a pre-transfer nor a post-transfer income measure.
-Therefor it’s comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
+Therefor it's comparability with other surveys is harsh as long as the main sources of data on inequality are based on measures, which cannot be replicated with taxable income.
 This FTA-Tax data problem can be handled with non-aggregated tax-data .
@@
 %(3) While the overall Swiss inequality remained pretty stable over decades, inequality between cantons underwent a bizarre development.
+
+% Mögliche Einstiegsfloskel
+% This article highlights the main disparities between taxdata based assessment of inequality opossed to given standard concepts of inequality as it is mainly applied with household-surveys.
+We reflected all important concerns based on the case of Switzerland.

### 2014-07-04 paper/subdocuments/results.Rnw
added introduction and description of defining income to results.Rnw

@@
 
+In this section we will present results following the strategy introducded in the last section.
+As describded this includes changeing possible concepts within one field (incomde defintion, statistical unit, measurement and population coverage) while holding other conceptional differences constant.
+To get the best possible comparison we rely on different data sources (FTA-tax tax data and HBS-survey data) and different time periodes.
+The presentation of results therefore includes a despriction of the uses data.
+
 \subsection{Defining income}
@@
 \emph{Gini coefficients for taxable income, income before deductions and taxable income after federal tax}
+
+We start with the effect of income definition on the assessment of income inequality.
+As described in the previous section income definitions are either closer to the market outcome or the the income, which shapes the possibility to consume (disposable income).
+With FTA-Tax data in-between income measures can be defined.
+This includes
+
+\begin{itemize}
+  \item \textit{taxable income}: all incomes declared for taxes (earnings, income from property and current transfers received) minus several deductions
+  \item \textit{net income}: Is a sort of total income, which excludes some (but not all) deductions
+  \item \textit{taxable income after federal tax}: By subtracting federal taxes from the taxable income, we get closer to the disposable income
+\end{itemize}
+
+Following this three definitions we calculate Gini-coefficients out of the FTA-tax statistics.
+Because information about net income per income bracket is only reported from 1983/1984 until 2010, we can show differences over this time period.
 
@@
 names(dflong)[4]<-"Gini"
-levels(dflong$Income)<-c("taxable income","income before deductions","taxable income after tax")
+levels(dflong$Income)<-c("taxable income","net income","taxable income after federal tax")
 ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,shape=Income))+
@@
 @
+
+As \ref{fig:different_ginis shows, the development for the three defined measures of income is quit parallel except for the 1980s.
+In this time periode the Gini-coefficent for net income veers.
+This has probably to do with a change in regulations of deductions and shows that interpretation over time has to be very carefull, becauses changes in taxation or regulation systems can affect the outcome.
+In generall inequality assesst with taxable income is higher than inequality assesst with net income or taxable income after federal taxes.
+While it is not surprising that federal taxes reduces inequality slightly because of the progressivity of the taxes and it is somehow astonishing that adding deductions (from net income to taxable income) increases inequality slightly.
 

### 2014-07-04 paper/subdocuments/discussion.Rnw
first sentence discussion.Rnw

@@
 
-In this paper we attempted to address four methodological issues related to inequality research using tax data.
+This article highlights the main disparities between taxdata based assessment of inequality opossed to given standard concepts of inequality as it is mainly applied with household-surveys.
+We reflected all important concerns based on the case of Switzerland.
 Can we measure what we want to measure?

### 2014-07-04 paper/subdocuments/results.Rnw
completed part measuring inequality

@@
 
-As \ref{fig:different_ginis shows, the development for the three defined measures of income is quit parallel except for the 1980s.
+As \ref{fig:different_ginis} shows, the development for the three defined measures of income is quit parallel except for the 1980s.
 In this time periode the Gini-coefficent for net income veers.
@@
 In generall inequality assesst with taxable income is higher than inequality assesst with net income or taxable income after federal taxes.
-While it is not surprising that federal taxes reduces inequality slightly because of the progressivity of the taxes and it is somehow astonishing that adding deductions (from net income to taxable income) increases inequality slightly.
+While it is not surprising that federal taxes reduces inequality slightly because of the progressivity of the taxes and it is somehow astonishing that adding deductions (from net income to taxable income) increases inequality.
 
 \subsection{Statistical units -- Equivalence scale}
+
+The main issue concerning the unit of analysis is not easy solved with tax-data, because the concepts of tax-units and households are not perfectly congruent (we will take up this issue, when comparing survey data and the FTA tax data later on.) But we can examine, how the assessment of inequality is affected by the implementation of equivalence scale.
+We do this by looking and Gini-time series for net income with and without implementation of a equivalence scale.
+The scale is constructed by using information out of tax data.
+The income of single household/tax unit are divided by 1 (no change), for married the equivalence-factor is 1.5.
+For every child or other by the tax-unit supported person a value of 0.3 is added.
+This measures are provided directly by the FTA and are not calculated by us.
+Because excluding the group of not-taxed leads to a longer time-series we provide four time-series in total (two possibilities to compare the effect of a equivalence scale)
 
@@
 
+The implementation of a equivalence scale does not have a major impact on the assessment of inequality.
+(see \ref{fig:equivalencescale}).
+Over the whole observed time period the two lines, which can be compared, move more or less parallel.
+Because tax units only approximately depict households, it has to be said, that the implemented equivalence scale automatically has it's drawbacks.
+This hinders a pure assessment of the effect of a equivalence scale.
+
 \subsection{Measuring inequality}
 
-% Vergleich über die Zeit da, wo zum ersten Mal Gesamtschweizerische Daten vorliegen bis zur neusten Reihe: 2003 vs 2010
+Here we examine how interpretation can change, when we expand the analysis by using relative distribution methods and not only look at Gini-Coefficients.
+We therefore use the published percentiles about the distribution of taxable income  on the FTA webpage .
+We prefer these measures over the calculates measures out of the published income brackets statistics, because they represent the distribution at both tails more accurate since they are based directly on the information about every single tax units\footnote{When calculating percentiles out of the income bracket statistic we loose relevant information at the edges.
+First, we don't have information about taxable income of tax-units falling below the income threshold for federal taxation.
+We only know how many persons fall in this category.However, the percentiles reported on the FTA webpage are based on the true taxable income (also for units below the threshold), which allows a more precise estimation of the lower percentiles.
+Secondly, it is especially hard to estimate the highest top income percentiles out of the aggregated tax statistics, leaving us with information only until the 95\%-percentiles, while the reported percentiles reach the 99.99\%-percentiles}.
+We use the reported Measures at the cost of time.
+The longest time-period we can compare out of these data reach from 2003 to 2010.
+Gini changed during this time from 0.47 to 0.50, which equals a moderate increase of inequality.
+The in-depth distributional analysis allows us to see, how this change translated into different shapes of the distributions.
+\\
+
+As described in the method section, the analysis includes a comparison of the density distribution of the log taxable income.
+We define the 2003 distribution as the reference group and compare it to the distribution of 2010.
+We then compare the relative density of the median adjusted distributions, to see where the two distributions differ.
+Finally, we calculate the median relative polarization index (MRP), to quantify the degree of polarization.
+We also show the decomposition into the upper and lower polarization index (URP and LRP).
 
@@
 
-% FEHLT: Tabelle mit Median, lower und upper polaritätsindex plus Gini bzw. Differenz der Ginis
+While looking relative density of shape differences (see right part of \ref{fig:reldist20032010}) it gets clearly visible, how the two compared distributions differ.
+The pattern suggest that from 2003 to 2010 a moderate polarization occured, which is represented in a lower relative density in the middle seven deciles (d.30 to d.90), while the density ratio is higher in the top decile and the two lowest deciles.
+Comparing 2003 to 2010 tax unites moved to the tails.
+This pattern is quantifed with the inequality indices reported in table 2.
+Looking at the lower index and the upper index shows, that the polarization is slightly stronger driven by the downgrading process.
 

### 2014-07-04 paper/subdocuments/results.Rnw
updated taxable income with and without results.Rnw

@@
 First, we don't have information about taxable income of tax-units falling below the income threshold for federal taxation.
-We only know how many persons fall in this category.However, the percentiles reported on the FTA webpage are based on the true taxable income (also for units below the threshold), which allows a more precise estimation of the lower percentiles.
+We only know how many persons fall in this category.
+However, the percentiles reported on the FTA webpage are based on the true taxable income (also for units below the threshold), which allows a more precise estimation of the lower percentiles.
 Secondly, it is especially hard to estimate the highest top income percentiles out of the aggregated tax statistics, leaving us with information only until the 95\%-percentiles, while the reported percentiles reach the 99.99\%-percentiles}.
@@
 While looking relative density of shape differences (see right part of \ref{fig:reldist20032010}) it gets clearly visible, how the two compared distributions differ.
-The pattern suggest that from 2003 to 2010 a moderate polarization occured, which is represented in a lower relative density in the middle seven deciles (d.30 to d.90), while the density ratio is higher in the top decile and the two lowest deciles.
+The pattern suggest that from 2003 to 2010 a moderate polarization occurred, which is represented in a lower relative density in the middle seven deciles (d.30 to d.90), while the density ratio is higher in the top decile and the two lowest deciles.
 Comparing 2003 to 2010 tax unites moved to the tails.
-This pattern is quantifed with the inequality indices reported in table 2.
-Looking at the lower index and the upper index shows, that the polarization is slightly stronger driven by the downgrading process.
+This pattern is quantified with the inequality indices reported in table 2.
+Comparing the lower and the upper index shows, that the polarization is slightly stronger driven by the downgrading process.
 
@@
 
+FTA-Tax statistics sometimes exlcude tax units, which reach below the treshold to be taxed (let's call them zeros).
+We compare here three Gini-time-series \ref{fig:with_without_zeros_Gini}.
+Excluding zeros leads to a dramatical drop of the gini-coefficent, which is not realy surprising.
+On the other hand inequality is overestimated when assuming non-taxed tax units have zero taxable income.
+Rather we must assume taxable income for zeros lay between zero and the taxation treshold.
+We adress this by presenting a third time-series, where we assume non-taxed have a taxable income equal halve the threshold for single tax units (around CHF 8000).
+This results slightly lower Gini-coefficents.
+
 % Beim Vergleich der Datensätze unterschieden nach nur besteuerten und unter berücksichtigung der nicht besteuerten, stellt sich die Frage wie die unteren Einkommen imputiert werden.
@@
 
-<<with_without_zeros_Gini,warning=FALSE, fig.cap='',fig.pos='H'>>=
-df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
-#library(reshape)
-dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
-names(dflong)[3]<-"Type"
-names(dflong)[4]<-"Gini"
-levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
-ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
-       aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
-  scale_y_continuous(limits=c(0.18,0.6))+
-  scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
-  geom_line(size=0.5)+
-  geom_point(aes(shape=Type),size=3)+
-  theme_bw()+
-  xlab("tax period")+
-  theme(legend.justification=c(0,1),legend.position=c(0, 1))
-
-# Kantonale Unterschiede
-#####
-#ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))+theme(legend.justification=c(1,0),legend.position=c(1, 0))
-#####
-@
+%<<with_without_zeros_Gini,warning=FALSE, fig.cap='',fig.pos='H'>>=
+%df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
+%#library(reshape)
+%dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
+%names(dflong)[3]<-"Type"
+%names(dflong)[4]<-"Gini"
+%levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
+%ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
+     %  aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
+ % scale_y_continuous(limits=c(0.18,0.6))+
+ % scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
+ % geom_line(size=0.5)+
+ % geom_point(aes(shape=Type),size=3)+
+ % theme_bw()+
+ % xlab("tax period")+
+ % theme(legend.justification=c(0,1),legend.position=c(0, 1))
+
+%# Kantonale Unterschiede
+%#####
+%#ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))+theme(legend.justification=c(1,0),legend.position=c(1, 0))
+%#####
+%@
 
@@
 
-<<with_without_zeros_RelDist_mitBruelhart,warning=FALSE, fig.cap='',fig.pos='H'>>=
-bd.2010 <- filter(read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
-bd.0.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
-start <- which(names(bd.2010)=="p1")
-end <- which(names(bd.2010)=="p99_99")
-percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
-bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
-names(bd.2010)[1]=c("inc")
-bd.0.2010<- data.frame(t(bd.0.2010[1,start:end]),percentile)
-names(bd.0.2010)[1]=c("inc")
-
-bd.0.2010 <- rbind(c(0,0),bd.0.2010)
-ival<-bd.0.2010$inc
-perval<-bd.0.2010$percentile
-bd.0.2010.ind<-unlist(sapply(2:length(ival),function(i){
-  len=(perval[i]-perval[i-1])*100
-  seq(ival[i-1],ival[i],length.out=len)
-}))
-
-bd.2010 <- rbind(c(0,0),bd.2010)
-ival<-bd.2010$inc
-perval<-bd.2010$percentile
-bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
-  len=(perval[i]-perval[i-1])*100
-  seq(ival[i-1],ival[i],length.out=len)
-}))
-
-par(mfrow=c(1,2))
-# Prop-Density Plot
-density1 <- density(log(bd.2010.ind))
-plot(x = (density1$x), y = density1$y, type = "l",
-     xlab = "taxable income", ylab = "density"
-     #axes = FALSE,
-     #xlim = c(-50, 200),
-     #ylim=c(0,0.0116031)
+%<<with_without_zeros_RelDist_mitBruelhart,warning=FALSE, fig.cap='',fig.pos='H'>>=
+%bd.2010 <- filter(read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
+%bd.0.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
+%start <- which(names(bd.2010)=="p1")
+%end <- which(names(bd.2010)=="p99_99")
+%percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+%bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
+%names(bd.2010)[1]=c("inc")
+%bd.0.2010<- data.frame(t(bd.0.2010[1,start:end]),percentile)
+%names(bd.0.2010)[1]=c("inc")
+
+%bd.0.2010 <- rbind(c(0,0),bd.0.2010)
+%ival<-bd.0.2010$inc
+%perval<-bd.0.2010$percentile
+%bd.0.2010.ind<-unlist(sapply(2:length(ival),function(i){
+ % len=(perval[i]-perval[i-1])*100
+%  seq(ival[i-1],ival[i],length.out=len)
+%}))
+
+%bd.2010 <- rbind(c(0,0),bd.2010)
+%ival<-bd.2010$inc
+%perval<-bd.2010$percentile
+%bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+%  len=(perval[i]-perval[i-1])*100
+%  seq(ival[i-1],ival[i],length.out=len)
+%}))
+
+%par(mfrow=c(1,2))
+%# Prop-Density Plot
+%density1 <- density(log(bd.2010.ind))
+%plot(x = (density1$x), y = density1$y, type = "l",
+%     xlab = "taxable income", ylab = "density"
+%     #axes = FALSE,
+%     #xlim = c(-50, 200),
+%     #ylim=c(0,0.0116031)
      )
-title(main="all tax units vs without non-taxed",cex=0.6)
-axis(side = 1)
-axis(side = 2)
-fig1legend <- list(x=c(6,6),y=c(0.6,0.6))
-legend(fig1legend,lty=1:2, bty="n",
-       legend=c("without non-taxed","all tax units"))
-density2 <- density(log(bd.0.2010.ind))
-lines(x = (density2$x), y = density2$y, type = "l",lty=2)
-
-# Relative Distribution
-
-gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
-               smooth=0.4, ci=FALSE,
-               show="residual",
-               bar=TRUE, quiet=FALSE,
-               ylim=c(0.5,2), ylab="",
-               discrete=FALSE,
-               xlab="Proportion of 2010 Distribution (including non-taxed)")
-title(main=paste("Effect of different shape"))
-abline(h=1,lty=2)
-
-par(mfrow=c(1,1))
-@
-
-<<polindexzeros, results='asis'>>=
-polTable(bd.2010.ind,bd.0.2010.ind,tablabel="polzeros")
-@
+%title(main="all tax units vs without non-taxed",cex=0.6)
+%axis(side = 1)
+%axis(side = 2)
+%fig1legend <- list(x=c(6,6),y=c(0.6,0.6))
+%legend(fig1legend,lty=1:2, bty="n",
+%       legend=c("without non-taxed","all tax units"))
+%density2 <- density(log(bd.0.2010.ind))
+%lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+%# Relative Distribution
+
+%gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
+%               smooth=0.4, ci=FALSE,
+%               show="residual",
+%               bar=TRUE, quiet=FALSE,
+%               ylim=c(0.5,2), ylab="",
+%               discrete=FALSE,
+%               xlab="Proportion of 2010 Distribution (including non-taxed)")
+%title(main=paste("Effect of different shape"))
+%abline(h=1,lty=2)
+
+%par(mfrow=c(1,1))
+@
+
+%<<polindexzeros, results='asis'>>=
+%polTable(bd.2010.ind,bd.0.2010.ind,tablabel="polzeros")
+%@
 
@@
 
-Including zeros leads to significantly higher gini coefficients.
+%Including zeros leads to significantly higher gini coefficients.
 However we must keep in mind, that these might be artificially high values as we assume zero income for everyone in the zero group.

### 2014-07-04 paper/subdocuments/results.Rnw
captions added results.Rnw

@@
 
-<<different_ginis, fig.cap='',fig.pos='H'>>=
+<<different_ginis, fig.cap='Gini coefficients for different income measures',fig.pos='H'>>=
 df$G_reink[df$G_reink==1]<-NA
@@
 
-<<equivalencescale, fig.cap='',fig.pos='H'>>=
+<<equivalencescale, fig.cap='Gini coefficients with and without equivalization',fig.pos='H'>>=
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
@@
 
-<<reldist20032010, echo=FALSE, fig.cap='',fig.pos='H'>>=
+<<reldist20032010, echo=FALSE, fig.cap='Distribution change over time',fig.pos='H'>>=
 
@@
 
-<<specialcases9394, fig.cap='',fig.pos='H'>>=
+<<specialcases9394, fig.cap='Comparison of normal and special cases 1993/94',fig.pos='H'>>=
 #plot relative distribution
@@
 
-<<specialcases10, fig.cap='',fig.pos='H'>>=
+<<specialcases10, fig.cap='Comparison of normal and special cases 2010',fig.pos='H'>>=
 
@@
 
-%<<with_without_zeros_Gini,warning=FALSE, fig.cap='',fig.pos='H'>>=
-%df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
-%#library(reshape)
-%dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
-%names(dflong)[3]<-"Type"
-%names(dflong)[4]<-"Gini"
-%levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
-%ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
-     %  aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
- % scale_y_continuous(limits=c(0.18,0.6))+
- % scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
- % geom_line(size=0.5)+
- % geom_point(aes(shape=Type),size=3)+
- % theme_bw()+
- % xlab("tax period")+
- % theme(legend.justification=c(0,1),legend.position=c(0, 1))
-
-%# Kantonale Unterschiede
-%#####
-%#ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))+theme(legend.justification=c(1,0),legend.position=c(1, 0))
-%#####
-%@
+<<with_without_zeros_Gini,warning=FALSE, fig.cap='Gini coefficients with, without and imputed non-taxed units',fig.pos='H'>>=
+df <- read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t")
+#library(reshape)
+dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
+names(dflong)[3]<-"Type"names(dflong)[4]<-"Gini"
+levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
+ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
+       aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
+  scale_y_continuous(limits=c(0.18,0.6))+
+  scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
+  geom_line(size=0.5)+
+  geom_point(aes(shape=Type),size=3)+
+  theme_bw()+
+  xlab("tax period")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+
+# Kantonale Unterschiede
+#####
+#ggplot(dflong[dflong$steuerperiode>=1995.5,], aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")+theme(axis.text.x = element_text(angle = 60, hjust = 1))+theme(legend.justification=c(1,0),legend.position=c(1, 0))
+#####
+@
 
@@
 %par(mfrow=c(1,1))
-@
+%@
 
@@
 
-<<BruelhartvsHABE,warning=FALSE, fig.cap='',fig.pos='H'>>=
+<<BruelhartvsHABE,warning=FALSE, fig.cap='Comparison of tax and survey data',fig.pos='H'>>=
 #library(ggplot2)
@@
 bd.2010.ind<-bd.2010.ind/1000
-
+@
+
+<<BruelhartvsHABEmarried,warning=FALSE, fig.cap='Comparison of tax and survey data for married',fig.pos='H'>>=
 # Select married - HABE
@@
 par(mfrow=c(1,1))
-
+@
+
+<<BruelhartvsHABEunmarried,warning=FALSE, fig.cap='Comparison of tax and survey data for unmarried',fig.pos='H'>>=
 #################
 # Separate comparison for unmaried
-
-### Not corrected!!!
 
@@
 On the other hand we expect underreporting from both lowest percentiles and highest percentiles (middle class bias) within the survey data.
-Figure \ref{fig:BruelhartvsHABE1} (left) however reveals a more critical issue related to tax data, that is the median location of income compared to survey data.
+Figure \ref{fig:BruelhartvsHABE} (left) however reveals a more critical issue related to tax data, that is the median location of income compared to survey data.
 Although we try to measure similar concepts of income, survey data shows a median (93.000 CHF) more than twice as big as tax data (44.600 CHF).
@@
 Splitting the data in married and unmarried tax units underpins this argument.
-From figure \ref{fig:BruelhartvsHABE2} we can see that married tax units (FTA data) and household with married couples (survey data) are better (but still not perfectly) comparable.
-Figure \ref{fig:BruelhartvsHABE2} (right) shows the expected shape difference between the two distributions:  survey data has a bias towards the (median-adjusted) 80\% to 90\% percentile (of the FTA data distribution).
+From figure \ref{fig:BruelhartvsHABEmarried} we can see that married tax units (FTA data) and household with married couples (survey data) are better (but still not perfectly) comparable.
+Figure \ref{fig:BruelhartvsHABEmarried} (right) shows the expected shape difference between the two distributions:  survey data has a bias towards the (median-adjusted) 80\% to 90\% percentile (of the FTA data distribution).
 Top percentiles are badly covered by survey data, suggesting that inequality measures based on survey data might underestimate inequality development that arises from changes in the incomes of the rich.
@@
 
-<<benplot, fig.cap='',fig.pos='H'>>=
+<<benplot, fig.cap='Gini coefficients over time',fig.pos='H'>>=
  ### libs

### 2014-07-04 paper/subdocuments/data.Rnw
textbf -> textit

@@
 Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}.
-(1) ``\textbf{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textbf{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
+(1) ``\textit{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textit{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
 doubling of all income) (3) ``\textit{weak principle of transfers}'' or ``\textit{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality.

### 2014-07-04 paper/subdocuments/results.Rnw
textbf -> textit

@@
 dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
-names(dflong)[3]<-"Type"names(dflong)[4]<-"Gini"
+names(dflong)[3]<-"Type"
+names(dflong)[4]<-"Gini"
 levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
@@
 
-estv_gini <- estv %.% 
-  filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %.% 
-  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %.%
-  group_by(factor(bemessungsjahre)) %.%
-  mutate(mean_steink=steink/anz_pflichtige) %.%
-  summarise(gini(mean_steink,weights=anz_pflichtige))
+#estv_gini <- estv %.% 
+#  filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %.% 
+#  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %.%
+#  group_by(factor(bemessungsjahre)) %.%
+#  mutate(mean_steink=steink/anz_pflichtige) %.%
+#  summarise(gini(mean_steink,weights=anz_pflichtige))
+
+load("estv_gini.RData")
 

### 2014-07-10 paper/subdocuments/data.Rnw
minor spelling corrections

--- 2014-07-03 paper/subdocuments/data.Rnw
@@
 This is a special thorny task for surveys working with samples, because nonresponse is a major source of bias \citep{bethlehem_handbook_2011}.
-\citep{korinek_survey_2006} show, that the position in the income distribution influence the probability to participate in a survey.
+\citet{korinek_survey_2006} show, that the position in the income distribution influence the probability to participate in a survey.
 Low income and high income households are more likely to refuse survey response, which leads to an overrepresentation of middle income households.
@@
 From 1995/1996 until 2010 the number of non-taxed units is reported, but not for the years before.
-Dell et al.
-(2007) try to estimate the fraction of non-taxed by comparing the reported numbers of tax units to census reports about the number of adult population.
+citet{dell_income_2005} try to estimate the fraction of non-taxed by comparing the reported numbers of tax units to census reports about the number of adult population.
 According to their estimations this fraction drops from 94\% in 1993/1994 to 63\% back in 1945/46.

### 2014-07-10 paper/subdocuments/methods.Rnw
minor spelling corrections

@@
 To fulfill the above described task, we use two techniques.
-To assess the development over time, we calculate in general Gini-coefficients for all possible time points, allowing us the make time trends visible and to compare the result out of the FTA-tax data to other publications.
+To assess the development over time, we calculate in general Gini-coefficients for all possible time points, allowing us to make time trends visible and to compare the result out of the FTA-tax data to other publications.
 For selected periods we expand the analysis with relative distribution methods, which allows an in-depth analysis of distributional differences and therefore catches the shortcomings of Gini-coefficients.
@@
 Subject of the comparison can also be two distributions describing the same population, but stemming from two different data sources like survey or tax data or even comparison of the same source/population but for different time points, like the income distribution out of tax data for Switzerland today compared to an earlier time point.
-To describe how the two distributions are gone be transformed into a relative distribution, we start with defining the two distributions.
+To describe how the two distributions are going to be transformed into a relative distribution, we start with defining the two distributions.
 One represents the reference population $Y_{0}$ and the other the comparison population $Y$.
@@
 Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail) or downgrading (increases in lower tail).
-The polarization index developed by \citep{Handock and Morris 1999} addresses this issue, because it is decomposable to distinguish differences in the upper and lower tails.
+The polarization index developed by \citep{handcock_relative_1999} addresses this issue, because it is decomposable to distinguish differences in the upper and lower tails.
 Because it is based on the relative distribution it provides a simple link between what is observed in the graphical display and what is measured by the numerical summary.

### 2014-09-17 paper/subdocuments/results.Rnw
export reldist data to stata creating stata plot (reldist) minor edits to the text in results.Rnw

@@
 In this section we will present results following the strategy introducded in the last section.
-As describded this includes changeing possible concepts within one field (incomde defintion, statistical unit, measurement and population coverage) while holding other conceptional differences constant.
+As describded this includes changeing possible concepts within one field (income definition, statistical unit, measurement and population coverage) while holding other conceptional differences constant.
 To get the best possible comparison we rely on different data sources (FTA-tax tax data and HBS-survey data) and different time periodes.
-The presentation of results therefore includes a despriction of the uses data.
+The presentation of results therefore includes a description of the used data.
 
@@
 We start with the effect of income definition on the assessment of income inequality.
-As described in the previous section income definitions are either closer to the market outcome or the the income, which shapes the possibility to consume (disposable income).
+As described in the previous section income definitions are either closer to the market outcome or the income, which shapes the possibility to consume (disposable income).
 With FTA-Tax data in-between income measures can be defined.
@@
 
-As \ref{fig:different_ginis} shows, the development for the three defined measures of income is quit parallel except for the 1980s.
+As \ref{fig:different_ginis} shows, the development for the three defined measures of income is quite parallel except for the 1980s.
 In this time periode the Gini-coefficent for net income veers.
-This has probably to do with a change in regulations of deductions and shows that interpretation over time has to be very carefull, becauses changes in taxation or regulation systems can affect the outcome.
-In generall inequality assesst with taxable income is higher than inequality assesst with net income or taxable income after federal taxes.
-While it is not surprising that federal taxes reduces inequality slightly because of the progressivity of the taxes and it is somehow astonishing that adding deductions (from net income to taxable income) increases inequality.
+This has probably to do with a change in regulations of deductions and shows that interpretation over time has to be very careful, becauses changes in taxation or regulation systems can affect the outcome.
+In general inequality assessed with taxable income is higher than inequality assessed with net income or taxable income after federal taxes.
+While it is not surprising that federal taxes reduces inequality slightly because of the progressivity of the taxes it is somehow astonishing that adding deductions (from net income to taxable income) increases inequality.
+This implies deductions to be rather flat not progressive.
 
@@
 
-The main issue concerning the unit of analysis is not easy solved with tax-data, because the concepts of tax-units and households are not perfectly congruent (we will take up this issue, when comparing survey data and the FTA tax data later on.) But we can examine, how the assessment of inequality is affected by the implementation of equivalence scale.
-We do this by looking and Gini-time series for net income with and without implementation of a equivalence scale.
+The main issue concerning the unit of analysis is not easy to solve with tax-data, because the concepts of tax-units and households are not perfectly congruent (we will take up this issue, when comparing survey data and the FTA tax data later on.) However we can examine, how the assessment of inequality is affected by the implementation of equivalence scale.
+We do this by looking at Gini-time series for net income with and without implementation of a equivalence scale.
 The scale is constructed by using information out of tax data.
-The income of single household/tax unit are divided by 1 (no change), for married the equivalence-factor is 1.5.
-For every child or other by the tax-unit supported person a value of 0.3 is added.
-This measures are provided directly by the FTA and are not calculated by us.
-Because excluding the group of not-taxed leads to a longer time-series we provide four time-series in total (two possibilities to compare the effect of a equivalence scale)
+The incomes of single households are divided by 1 (no change), for married tax units the equivalence-factor is 1.5.
+For every child and person supported by the tax-unit a value of 0.3 is added to the denominator.
+These measures are provided directly by the FTA and are not calculated by us.
+Because excluding the group of not-taxed leads to a longer time-series we provide four time-series in total (two possibilities to compare the effect of a equivalence scale).
 
@@
 We use the reported Measures at the cost of time.
-The longest time-period we can compare out of these data reach from 2003 to 2010.
+The longest time-period we can compare out of these data reaches from 2003 to 2010.
 Gini changed during this time from 0.47 to 0.50, which equals a moderate increase of inequality.
@@
 par(mfrow=c(1,1))
+
+# Export data to use with Stata
+bd2010_2003 <- rbind(data.frame(cbind("bd" = bd.2010.ind,"group"=2010)),data.frame("bd" = bd.2003.ind,"group"=2003))
+write.dta(bd2010_2003, file="../../data/stata_data/bd2010_2003.dta")
 @
@@
 
+# Export data to use with Stata
+ns1993_94 <- rbind(data.frame(cbind("income" = pooled_i,"group"=0)),data.frame("income" = normal_i,"group"=1))
+ns1993_94$group <- factor(ns1993_94$group,labels=c("pooled","normal"))
+write.dta(ns1993_94, file="../../data/stata_data/ns1993_94.dta")
+
 # Older Versions Rel-Dist-Plots (1993/1994)
@@
 par(mfrow=c(1,1))
+
+# Export data to use with Stata
+ns2010 <- rbind(data.frame(cbind("income" = pooled_i,"group"=0)),data.frame("income" = normal_i,"group"=1))
+ns2010$group <- factor(ns2010$group,labels=c("pooled","normal"))
+write.dta(ns2010, file="../../data/stata_data/ns2010.dta")
+
 @
@@
 
+# Export data to use with Stata
+habe_bd_2010 <- rbind(data.frame(cbind("income" = habe10$psteink,"group"=0,"weight"=habe10$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
+habe_bd_2010$group <- factor(habe_bd_2010$group,labels=c("habe","bd"))
+write.dta(habe_bd_2010, file="../../data/stata_data/habe_bd_2010.dta")
+
 ########################
@@
 par(mfrow=c(1,1))
+
+# Export data to use with Stata
+habe_bd_2010_married <- rbind(data.frame(cbind("income" = habe10.married$inc_habe,"group"=0,"weight"=habe10.married$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
+habe_bd_2010_married$group <- factor(habe_bd_2010_married$group,labels=c("habe","bd"))
+write.dta(habe_bd_2010_married, file="../../data/stata_data/habe_bd_2010_married.dta")
+
 @
@@
 par(mfrow=c(1,1))
+
+# Export data to use with Stata
+habe_bd_2010_unmarried <- rbind(data.frame(cbind("income" = habe10.unmarried$inc_habe,"group"=0)),data.frame("income" = bd.2010.ind,"group"=1))
+habe_bd_2010_unmarried$group <- factor(habe_bd_2010_unmarried$group,labels=c("habe","bd"))
+write.dta(habe_bd_2010_unmarried, file="../../data/stata_data/habe_bd_2010_unmarried.dta")
+
 @

### 2014-09-23 paper/subdocuments/results.Rnw

@@
 
-# Export data to use with Stata
-habe_bd_2010 <- rbind(data.frame(cbind("income" = habe10$psteink,"group"=0,"weight"=habe10$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
+# Export data to use with Stata (psteink durch VerfuegbaresEinkommen08 ersetzt)
+habe_bd_2010 <- rbind(data.frame(cbind("income" = habe10$VerfuegbaresEinkommen08,"group"=0,"weight"=habe10$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
 habe_bd_2010$group <- factor(habe_bd_2010$group,labels=c("habe","bd"))
@@
 habe10.unmarried<-habe0911 %.% 
-  filter(Familientyp08<140,Jahr08==2010)
-habe10.unmarried<-data.frame(habe10.unmarried$VerfuegbaresEinkommen08)
-names(habe10.unmarried)<-"inc_habe"
+  filter(Familientyp08<140,Jahr08==2010) %.%
+  select(VerfuegbaresEinkommen08,jahresgewicht)
+#habe10.unmarried<-data.frame(habe10.unmarried$VerfuegbaresEinkommen08)
+
+names(habe10.unmarried)<-c("inc_habe","gew")
 habe10.unmarried$inc_habe[habe10.unmarried$inc_habe<0]<-0 # One values is -20'000
@@
 
-# Export data to use with Stata
-habe_bd_2010_unmarried <- rbind(data.frame(cbind("income" = habe10.unmarried$inc_habe,"group"=0)),data.frame("income" = bd.2010.ind,"group"=1))
+# Export data to use with Stata (gewicht hinzugefügt)
+habe_bd_2010_unmarried <- rbind(data.frame(cbind("income" = habe10.unmarried$inc_habe,"group"=0,"weight"=habe10.unmarried$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
 habe_bd_2010_unmarried$group <- factor(habe_bd_2010_unmarried$group,labels=c("habe","bd"))

### 2014-09-23 paper/subdocuments/results.Rnw
noch was vergessen

@@
 
-# Export data to use with Stata (psteink durch VerfuegbaresEinkommen08 ersetzt)
-habe_bd_2010 <- rbind(data.frame(cbind("income" = habe10$VerfuegbaresEinkommen08,"group"=0,"weight"=habe10$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
+# Export data to use with Stata (psteink durch VerfuegbaresEinkommen08 (inc_habe) ersetzt)
+habe_bd_2010 <- rbind(data.frame(cbind("income" = habe10$inc_habe,"group"=0,"weight"=habe10$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
 habe_bd_2010$group <- factor(habe_bd_2010$group,labels=c("habe","bd"))

### 2014-11-25 paper/paper.tex
major sync,word files, figures, bern tax data analyses

--- 2014-06-30 paper/paper.tex
@@
 \usepackage{graphicx}
+\usepackage{datetime} %for automatic date in the header
 
@@
 \fancyfoot{} % Blank out the default footer
-\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
+\newdateformat{mydate}{\monthname[\THEMONTH] \THEYEAR}
+\fancyhead[C]{Assessing income inequality with tax data $\bullet$ \mydate\today~$\bullet$ Working paper} % Custom header text
 \fancyfoot[RO,LE]{\thepage} % Custom footer text
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/Swiss_inequality_Plot}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/SwissinequalityPlot} \caption[Trends in income inequality]{Trends in income inequality\label{fig:SwissinequalityPlot}}
+\end{figure}
 
@@
 Furthermore it has several desired statistical properties \citet{engelhardt_modelle_2000}.
-(1) ``\textbf{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textbf{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
-doubling of all income) (3) ``\emph{weak principle of transfers}'' or ``\emph{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality.
+(1) ``\textit{principle of population}'': the assessment of inequality is independent of the population size (2) ``\textit{Requirement of Bresciani-Turroni}'': the measure is sensitive for changes of income shares, but not for absolute changes (e.g.
+doubling of all income) (3) ``\textit{weak principle of transfers}'' or ``\textit{requirement of Pigou-Dalton}'': transfers from richer households lead to a reduction of inequality.
 However, several drawbacks are reported in the literature.
@@
 
-\begin{table}[h]
-\begin{tabular}{lcccc}
-                                   & FTA       & Survey data\\
-Statistical unit                   & Tax units (individuals/married couples)  & Households\\                                   
-Concepts of economic resources  & Data-driven   & Theory-driven\\
+\multicolinterrupt{
+\begin{table}[H]
+\caption{Comparison of tax and survey data\label{datacomp}} 
+\begin{center}
+\begin{tabular}{lcc}
+\hline\hline
+                                   & Tax data       & Survey data\\
+\hline
+Statistical unit                   & tax units (individuals/married couples)  & households\\                                   
+Concepts of economic resources  & data-driven   & theory-driven\\
 Population coverage & total population & sample \\ 
-Top incomes covered                 & Yes   & No\\
+Top incomes covered                 & yes   & no\\
 Low incomes covered?               & no        & yes\\
-Comparison over time possible 	   & Long       & Short \\
-Cross country comparison possible? & limited   & possible
-
+Comparison over time possible 	   & long       & short \\
+Cross country comparison possible? & limited   & possible\\
+\hline
 \end{tabular}
+\end{center}
 \end{table}
-
+}
 %RF: damit haben wir WER, WANN, WAS, wollen wir noch ein WOMIT/WAS DARSTELLEN -> methodik -> reldist ? jaa
@@
 
-\subsection{Outline of applied methods}
+\section{Methods}
 
@@
 
-In the last section we described the advantages and drawback of taxdata discussing five aspects, which we regard as crucial concering the assessment of inequality.
-To get a feeling of the importance of these aspects, we exploit the FTA-tax data as far as possible and perform several insightful calculations adressing four of the five mentioned aspects.
-No further investigation is possible regarding aspect (1) concepts on measuring economic ressource.
-This is a point which cannot be further addresses with FTA tax data.
+In the last section we described the advantages and drawback of tax-data discussing five aspects, which we regard as crucial concerning the assessment of inequality.
+To get a feeling of the importance of these aspects, we exploit the FTA-tax data as far as possible and perform several insightful calculations addressing four of the five mentioned aspects.
+No further investigation is possible regarding aspect (1) concepts on measuring economic resource.
 But for the other four thematic areas, we can provide deeper insight.
 In general our main strategy is to apply different possible concepts within one conceptional area (income, units, measurement, population coverage) while holding other conceptual differences constant.
-With this strategy we want to show, the sensitiveness the assessement of inequality is especially if one looks at time trens.
-In this section we  describe which methods we use to produce the results in the next section.
+With this strategy we want to show, the sensitiveness the assessment of inequality is especially if one looks at time trends.\\
+
+To fulfill the above described task, we use two techniques.
+To assess the development over time, we calculate in general Gini-coefficients for all possible time points, allowing us the make time trends visible and to compare the result out of the FTA-tax data to other publications.
+For selected periods we expand the analysis with relative distribution methods, which allows an in-depth analysis of distributional differences and therefore catches the shortcomings of Gini-coefficients.
+\\
+
+To calculate Gini-coefficients we follow common definitions, like it is described for example in \citep{jann_einfuhrung_2005}.
+Because the relative distribution methods are not that widely known, we provide a short introduction of the important concepts in this section.
+We additional expand the method section with an extended analysis about the importance of non-taxed, because the issue of non-taxed is a especially thorny one.
+
+\subsection{Relative distribution (RD) framework}
+
+The RD-framework is based on the concept of a ``relative distribution'', a transformation of the data from two distributions into a single distribution that contains all of the information necessary for scale-invariant comparison.
+This allows to make distribution differences ``visual'' in an elegant way and it is also a base for summary statistics, which are more sensitive to detailed theoretical hypotheses in contrast to other measures like the Gini-coefficient or top income-shares, which inform either about the whole population or only about the top part of the distribution.\\
+
+% Probability Density function as a base for the RD
+The goal of RD is to study the differences between two distributions.
+A common example could be the income distribution for man and women.
+Subject of the comparison can also be two distributions describing the same population, but stemming from two different data sources like survey or tax data or even comparison of the same source/population but for different time points, like the income distribution out of tax data for Switzerland today compared to an earlier time point.
+To describe how the two distributions are gone be transformed into a relative distribution, we start with defining the two distributions.
+One represents the reference population $Y_{0}$ and the other the comparison population $Y$.
+$x$ represents our measure of interest (income).
+A first visual approach is to compare the two probability density functions (PDF).
+The PDF is a function $f(x)$ which describes the distribution of probability over the outcome set and is defined for all possible values of $x$.
+This function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%.
+Out of the comparison of the PDF, it is possible to see, which values of $x$ are more and less probable.
+This already allows to spot distributional differences over the whole scale of $x$ visually.
+\\
+
+The PDF can be characterized by its cumulative distribution function (CDF).
+The CDF can be formulated as $F(x)$, which represents the probability that a randomly chosen value is less than or equal to $x$.
+The relative distribution of $Y$ to $Y_{0}$ is then defined as
+
+\begin{equation}
+R=F_{0}(Y)
+\end{equation}
+
+$R$ is obtained from $Y$ by transforming it by the CDF for $Y_{0}$, $F_{0}$. $R$ therefore measures the relative rank of $Y$ compared to $Y_{0}$.
+
+% Relative Distribution
+
+\begin{equation}
+g(r)=\frac{f(F_{0}^{-1}(r))}{^{f_{0}(F_{0}^{-1}(r))}}
+\end{equation}
+
+We can calculate the Probability Density Function $g(r)$ of R, where $r$ represents the proportion of values and $F_{0}^{-1}(r)$ is the inverse cumulative distribution function, also called the quantile function.
+$g(r)$ can be interpreted as a density ratio, which is defined as the ratio of these two quantities evaluated at every percentile of the reference distribution [0,1].
+With a complete overlap of both distributions the probability density function of the $R$ is 1 at every point of the PDF.
+On the other hand, values higher than 1 represent higher probabilities in the comparison distribution than in the references distribution at this specific point and values lower than 1 respectively represent lower probabilities.
+It is a proper PDF in the sense that it integrates to 1 over the unit interval.\\
+
+% Median and shape differences
+What we got through the above transformation of two distributions is the overall relative probability density.
+But differences between distributions can be divided into two basic components: changes in location and changes in shape.
+If the comparative distribution is a simple location-shifted version of the reference distribution, then the difference between the two distributions can be parsimoniously summarized by this shift.
+Differences that remain after a location adjustment are differences in ``shape'' (scale, skewness and other distributional characteristics).
+When both types of shifts are operating, or when factors other than scale are changing in the shape component, we need a way to separate out the various effects.
+If we want to identify the effect of a location shift and separate it from other changes in the distribution, it is necessary to specify what scale this shift operates on.
+It is possible to adjust distributions by any measure of central tendency.
+Here we choose a median location adjustment because the median is a natural, robust and scale invariant unit of measurement.
+Because our interest lies in analyzing distributional differences concerning the degree of inequality, we will focus in the results section on shape differences and look therefore at the relative distribution after the distributions are adjusted for location differences.
+
+% R based Summary Meausres
+Distributional polarization is of particular interest in the study of inequality.
+However, common inequality indicators (for example Gini or Theils index) are not designed to distinguish between growth in the upper and lower tails.
+Even if the measures register increasing inequality over time, one cannot distinguish a polarization of the distribution (increases in both tails) from upgrading (increases in the upper tail) or downgrading (increases in lower tail).
+The polarization index developed by \citep{Handock and Morris 1999} addresses this issue, because it is decomposable to distinguish differences in the upper and lower tails.
+Because it is based on the relative distribution it provides a simple link between what is observed in the graphical display and what is measured by the numerical summary.
+\\
+
+The median relative polarization index (MRP) is defined as the mean absolute deviation around the median of the location-adjusted relative distribution, scaled to produce an index that varies between -1 and 1.
+Given the scaling, a value of zero represents no differences in distributional shape; positive values represent more polarization (increases in the tails of the distribution); and negative values represent less polarization (convergence towards the center of the distribution).
+The measure catches only differences in distributional shape (not location).
+And it has several interesting features.
+MRP can be interpreted in terms of a proportional shift of mass in the distribution from more central to less central values.
+A value of 0.1, for example, is equivalent to a 10\% population shift from the center of the distribution to the upper and lower quartiles and the MRP is decomposable along the scale of $y$.
+This makes it possible to compare the contribution of each section of the distribution to the overall polarization.
+A natural decomposition is the contributions made by components above (upper polarization index, URP) and below (lower polarization index, LRP) the median (of $g(r)$).
 
@@
 
-\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
-
-\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
+%\textbf{Estimating percentiles from bracket income tabulation} Pareto interpolation \\
+
+%\textbf{Missing of mean-tested benefits as part of the income} -> imputation with recommendation for minimum level for basic needs defined by the SKOS.\\
 %is never mentioned as a problem, but it seems to me a better way to approach the non-taxed issue, than dell way (20 % of average income)
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/zero_descriptives}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/zero_descriptives} \caption[Share of tax units below federal tax liability threshold]{Share of tax units below federal tax liability threshold\label{fig:zero_descriptives}}
+\end{figure}
 
@@
 We know these cases had incomes between zero and the threshold (usually around 15.000 CHF for the period 1995-2010).
-In figure WITHWITHOUTZEROSTIMESERIES we show Gini coefficients calculated under the assumption of the zeros being zero.
+In figure \ref{fig:with_without_zeros_Gini} we show Gini-coefficients calculated under the assumption of the zeros being zero.
 This is problematic for two reasons: first we know these units might earn more (up to the threshold) so Ginis calculated on this assumptions will be too high.
-Second we assume the zero group to be perfectly homogenous which on the contrary has a lowering effect on the Gini coefficient.
+Second we assume the zero group to be perfectly homogeneous which on the contrary has a lowering effect on the Gini coefficient.
 Therefore we must at least assume a distribution for the zeros that has some variance and ranges from zero to the threshold.
@@
 The second approach investigates whether the FTA time series is suitable to analyze trends of inequality even if the inequality measures are biased.
-We can do this by regressing the inequality measure (here the Gini coefficient) on the share of zeros.
+We can do this by regressing the inequality measure (here the Gini-coefficient) on the share of zeros.
 Simply spoken we test if cutting a larger (left) tail of the distribution changes the shape of the distribution.
-We include Gini coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time)}.
+We include Gini-coefficient for all cantons in our analyses to improve the power of our statistical test\footnote{Furthermore without cantonal data (but only data for aggregate Switzerland) it is impossible to distinguish whether inequality changed due to the number of unobserved people or any other variable that changed over time)}.
 
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/corrected_gini}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/corrected_gini} \caption[Bias variation by time and cantons]{Bias variation by time and cantons\label{fig:corrected_gini}}
+\end{figure}
 
@@
 
+In this section we will present results following the strategy introducded in the last section.
+As describded this includes changeing possible concepts within one field (incomde defintion, statistical unit, measurement and population coverage) while holding other conceptional differences constant.
+To get the best possible comparison we rely on different data sources (FTA-tax tax data and HBS-survey data) and different time periodes.
+The presentation of results therefore includes a despriction of the uses data.
+
 \subsection{Defining income}
 
-\emph{Gini coefficients for taxable income, income before deductions and taxable income after fedaral tax}
+\emph{Gini coefficients for taxable income, income before deductions and taxable income after federal tax}
+
+We start with the effect of income definition on the assessment of income inequality.
+As described in the previous section income definitions are either closer to the market outcome or the the income, which shapes the possibility to consume (disposable income).
+With FTA-Tax data in-between income measures can be defined.
+This includes
+
+\begin{itemize}
+  \item \textit{taxable income}: all incomes declared for taxes (earnings, income from property and current transfers received) minus several deductions
+  \item \textit{net income}: Is a sort of total income, which excludes some (but not all) deductions
+  \item \textit{taxable income after federal tax}: By subtracting federal taxes from the taxable income, we get closer to the disposable income
+\end{itemize}
+
+Following this three definitions we calculate Gini-coefficients out of the FTA-tax statistics.
+Because information about net income per income bracket is only reported from 1983/1984 until 2010, we can show differences over this time period.
 
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/different_ginis}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/different_ginis} \caption[Gini coefficients for different income measures]{Gini coefficients for different income measures\label{fig:different_ginis}}
+\end{figure}
 
@@
 
-\emph{Equivalence scale}
+As \ref{fig:different_ginis} shows, the development for the three defined measures of income is quit parallel except for the 1980s.
+In this time periode the Gini-coefficent for net income veers.
+This has probably to do with a change in regulations of deductions and shows that interpretation over time has to be very carefull, becauses changes in taxation or regulation systems can affect the outcome.
+In generall inequality assesst with taxable income is higher than inequality assesst with net income or taxable income after federal taxes.
+While it is not surprising that federal taxes reduces inequality slightly because of the progressivity of the taxes and it is somehow astonishing that adding deductions (from net income to taxable income) increases inequality.
+
+\subsection{Statistical units -- Equivalence scale}
+
+The main issue concerning the unit of analysis is not easy solved with tax-data, because the concepts of tax-units and households are not perfectly congruent (we will take up this issue, when comparing survey data and the FTA tax data later on.) But we can examine, how the assessment of inequality is affected by the implementation of equivalence scale.
+We do this by looking and Gini-time series for net income with and without implementation of a equivalence scale.
+The scale is constructed by using information out of tax data.
+The income of single household/tax unit are divided by 1 (no change), for married the equivalence-factor is 1.5.
+For every child or other by the tax-unit supported person a value of 0.3 is added.
+This measures are provided directly by the FTA and are not calculated by us.
+Because excluding the group of not-taxed leads to a longer time-series we provide four time-series in total (two possibilities to compare the effect of a equivalence scale)
 
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/equivalencescale}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/equivalencescale} \caption[Gini coefficients with and without equivalization]{Gini coefficients with and without equivalization\label{fig:equivalencescale}}
+\end{figure}
 
@@
 
+The implementation of a equivalence scale does not have a major impact on the assessment of inequality.
+(see \ref{fig:equivalencescale}).
+Over the whole observed time period the two lines, which can be compared, move more or less parallel.
+Because tax units only approximately depict households, it has to be said, that the implemented equivalence scale automatically has it's drawbacks.
+This hinders a pure assessment of the effect of a equivalence scale.
+
 \subsection{Measuring inequality}
 
-% Vergleich über die Zeit da, wo zum ersten Mal Gesamtschweizerische Daten vorliegen bis zur neusten Reihe: 2003 vs 2010
+Here we examine how interpretation can change, when we expand the analysis by using relative distribution methods and not only look at Gini-Coefficients.
+We therefore use the published percentiles about the distribution of taxable income  on the FTA webpage .
+We prefer these measures over the calculates measures out of the published income brackets statistics, because they represent the distribution at both tails more accurate since they are based directly on the information about every single tax units\footnote{When calculating percentiles out of the income bracket statistic we loose relevant information at the edges.
+First, we don't have information about taxable income of tax-units falling below the income threshold for federal taxation.
+We only know how many persons fall in this category.
+However, the percentiles reported on the FTA webpage are based on the true taxable income (also for units below the threshold), which allows a more precise estimation of the lower percentiles.
+Secondly, it is especially hard to estimate the highest top income percentiles out of the aggregated tax statistics, leaving us with information only until the 95\%-percentiles, while the reported percentiles reach the 99.99\%-percentiles}.
+We use the reported Measures at the cost of time.
+The longest time-period we can compare out of these data reach from 2003 to 2010.
+Gini changed during this time from 0.47 to 0.50, which equals a moderate increase of inequality.
+The in-depth distributional analysis allows us to see, how this change translated into different shapes of the distributions.
+\\
+
+As described in the method section, the analysis includes a comparison of the density distribution of the log taxable income.
+We define the 2003 distribution as the reference group and compare it to the distribution of 2010.
+We then compare the relative density of the median adjusted distributions, to see where the two distributions differ.
+Finally, we calculate the median relative polarization index (MRP), to quantify the degree of polarization.
+We also show the decomposition into the upper and lower polarization index (URP and LRP).
 
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/reldist20032010}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/reldist20032010} \caption[Distribution change over time]{Distribution change over time\label{fig:reldist20032010}}
+\end{figure}
 
@@
 
-\begin{kframe}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: splines}}\end{kframe}% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title)
-%
-\begin{table}[!tbp]
+%latex.default(format(c(`Median Index` = med, `Lower Index` = lower,     `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),     file = "", title = title, where = "H", label = tablabel,     caption = caption)%
+\begin{table}[H]
+\caption{\label{pol20032010}} 
 \begin{center}
@@
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \end{table}
@@
 
-% FEHLT: Tabelle mit Median, lower und upper polaritätsindex plus Gini bzw. Differenz der Ginis
+While looking relative density of shape differences (see right part of \ref{fig:reldist20032010}) it gets clearly visible, how the two compared distributions differ.
+The pattern suggest that from 2003 to 2010 a moderate polarization occurred, which is represented in a lower relative density in the middle seven deciles (d.30 to d.90), while the density ratio is higher in the top decile and the two lowest deciles.
+Comparing 2003 to 2010 tax unites moved to the tails.
+This pattern is quantified with the inequality indices reported in table 2.
+Comparing the lower and the upper index shows, that the polarization is slightly stronger driven by the downgrading process.
 
@@
 
-\subsubsection{Normal versus special cases}
-
-The FTA stopped to publicly report data for special cases after tax period 1993/94. 
-For more recent periods however we can compare already calculated percentiles provided by the FTA\footnote{These calculations were done on commission of the FTA within the SNF project Sinergia Nr.
+\emph{Normal versus special cases}
+
+The FTA distinguishes normal from special cases as described in the data section.
+To test whether it matters which cases the researcher looks at we want to compare the distributions of normal and special cases.
+Unfortunately, the FTA stopped to publicly report data for special cases after tax period 1993/94.
+Therefore we will compare the two distributions for a rather old dataset.
+However the FTA does report aggregate statistics (e.g.
+percentiles) based on a pool of all cases (normal and special) for more recent periods\footnote{These calculations were done on commission of the FTA within the SNF project Sinergia Nr.
 130648 "The Swiss Confederation: A Natural Laboratory for Research on Fiscal and Political Decentralization" by Raphael Parchet and Stefanie Brilon in coordination with Prof.
 Dr.
-Marius Brülhart.} which are based on both special and normal cases to normal cases alone.
-We will have a look at both, the 1993/94 period as the last period where both 
-numbers were reported as well as 2010 where we compare data from the same source (FTA)
-but based on different reporting (aggregated data vs. percentiles).
+Marius Brülhart.} which allows us to do a corresponding analysis for 2010 as well.
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/specialcases9394}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/specialcases9394} \caption[Comparison of normal and special cases 1993/94]{Comparison of normal and special cases 1993/94\label{fig:specialcases9394}}
+\end{figure}
 
@@
 
-% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title) 
-%
-\begin{table}[!tbp]
+%latex.default(format(c(`Median Index` = med, `Lower Index` = lower,     `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),     file = "", title = title, where = "H", label = tablabel,     caption = caption)%
+\begin{table}[H]
+\caption{\label{pol9394spec}} 
 \begin{center}
@@
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \end{table}
 
-1993/94 including special cases changes the distribution slightly at both ends.
-Special cases have a slightly lower median income and therefore are more frequent in the lower percentiles and less frequent in the upper percentiles.
-In addition there is an effect of shape: the distribution of special cases is more skewed so that lower percentiles show lower incomes and higher percentiles show higher incomes compared to just normal cases.
+1993/94 a pooled data set of normal and special cases has a slightly higher density at both ends compared to data based in normal cases only (see figure \ref{fig:specialcases9394}).
+Special cases appear to have a slightly lower median income and their distribution is more skewed.
+Therefore special cases are more polarized than normal cases (see table \ref{pol9394spec}), i.e.
+striving away from the median (positive Median Index of 0.02).
+This tendency is more pronounced in the lower than upper part  of the distribution (Lower Index of 0.029 compared to Upper Index if 0.01).
+As the special cases consist of a broad mix of individuals it remains unclear which factor explains the differences of both distributions.
+Possible explanations can be immigrants partly concentrating in lower income percentiles, low income artists who belong to the special cases or a more technical selection effect: tax units not liable for taxation throughout the whole year are special cases; those cases might have lower incomes, e.g.
+if they moved and stopped working.
+To get a more complete picture we can look how the two distributions relate to each other in 2010 (see figure \ref{fig:specialcases10}).
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/specialcases10}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/specialcases10} \caption[Comparison of normal and special cases 2010]{Comparison of normal and special cases 2010\label{fig:specialcases10}}
+\end{figure}
 
@@
 
-% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title) 
-%
-\begin{table}[!tbp]
+%latex.default(format(c(`Median Index` = med, `Lower Index` = lower,     `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),     file = "", title = title, where = "H", label = tablabel,     caption = caption)%
+\begin{table}[H]
+\caption{\label{pol2010spec}} 
 \begin{center}
@@
 \hline
-Median Index&0.170\tabularnewline
-Lower Index&0.206\tabularnewline
-Upper Index&0.134\tabularnewline
-$\Delta$ Gini&0.044\tabularnewline
+Median Index&0.031\tabularnewline
+Lower Index&0.039\tabularnewline
+Upper Index&0.022\tabularnewline
+$\Delta$ Gini&0.020\tabularnewline
 \hline
-\end{tabular}
-\end{center}
+\end{tabular}\end{center}
+
 \end{table}
 
-2010 the picture is similar but more apparent: There is a small effect of median shift lowering incomes in the lowest percentiles. The effect of shape however contributes most.
+2010 the picture is similar but more apparent: Special cases appear more frequent around the lower percentiles of the pooled distribution (Lower Index of 0.039), however 2010 there is a more noteworthy effect in the upper part of the distribution (Upper Index of 0.022).
+According to figure \ref{fig:specialcases10} we can attribute this effect to the top percentiles.
+This gives credibility to the thesis that rich immigrants whose number increased between 1994 and 2010 drive the effect.
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/with_without_zeros_Gini}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/with_without_zeros_Gini} \caption[Gini coefficients with, without and imputed non-taxed units]{Gini coefficients with, without and imputed non-taxed units\label{fig:with_without_zeros_Gini}}
+\end{figure}
 
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/with_without_zeros_RelDist_mitBruelhart1} 
-\begin{kframe}\begin{verbatim}
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/with_without_zeros_RelDist_mitBruelhart2}
-
-\end{knitrout}
-
-% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title) 
-%
-\begin{table}[!tbp]
-\begin{center}
-\begin{tabular}{ll}
-\hline\hline
-\multicolumn{1}{l}{Inequality Indices}&\multicolumn{1}{c}{}\tabularnewline
-\hline
-Median Index&-0.1084\tabularnewline
-Lower Index&-0.2155\tabularnewline
-Upper Index&-0.0013\tabularnewline
-$\Delta$ Gini&-0.0952\tabularnewline
-\hline
-\end{tabular}
-\end{center}
-\end{table}
+%<<with_without_zeros_RelDist_mitBruelhart,warning=FALSE, fig.cap='',fig.pos='H'>>=
+%bd.2010 <- filter(read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
+%bd.0.2010 <- filter(read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";"),Veranlagungsperiode==2010,Einheit=="Total")
+%start <- which(names(bd.2010)=="p1")
+%end <- which(names(bd.2010)=="p99_99")
+%percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+%bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
+%names(bd.2010)[1]=c("inc")
+%bd.0.2010<- data.frame(t(bd.0.2010[1,start:end]),percentile)
+%names(bd.0.2010)[1]=c("inc")
+
+%bd.0.2010 <- rbind(c(0,0),bd.0.2010)
+%ival<-bd.0.2010$inc
+%perval<-bd.0.2010$percentile
+%bd.0.2010.ind<-unlist(sapply(2:length(ival),function(i){
+ % len=(perval[i]-perval[i-1])*100
+%  seq(ival[i-1],ival[i],length.out=len)
+%}))
+
+%bd.2010 <- rbind(c(0,0),bd.2010)
+%ival<-bd.2010$inc
+%perval<-bd.2010$percentile
+%bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+%  len=(perval[i]-perval[i-1])*100
+%  seq(ival[i-1],ival[i],length.out=len)
+%}))
+
+%par(mfrow=c(1,2))
+%# Prop-Density Plot
+%density1 <- density(log(bd.2010.ind))
+%plot(x = (density1$x), y = density1$y, type = "l",
+%     xlab = "taxable income", ylab = "density"
+%     #axes = FALSE,
+%     #xlim = c(-50, 200),
+%     #ylim=c(0,0.0116031)
+     )
+%title(main="all tax units vs without non-taxed",cex=0.6)
+%axis(side = 1)
+%axis(side = 2)
+%fig1legend <- list(x=c(6,6),y=c(0.6,0.6))
+%legend(fig1legend,lty=1:2, bty="n",
+%       legend=c("without non-taxed","all tax units"))
+%density2 <- density(log(bd.0.2010.ind))
+%lines(x = (density2$x), y = density2$y, type = "l",lty=2)
+
+%# Relative Distribution
+
+%gA0 <- reldist(y=bd.2010.ind, yo=bd.0.2010.ind,
+%               smooth=0.4, ci=FALSE,
+%               show="residual",
+%               bar=TRUE, quiet=FALSE,
+%               ylim=c(0.5,2), ylab="",
+%               discrete=FALSE,
+%               xlab="Proportion of 2010 Distribution (including non-taxed)")
+%title(main=paste("Effect of different shape"))
+%abline(h=1,lty=2)
+
+%par(mfrow=c(1,1))
+%@
+
+%<<polindexzeros, results='asis'>>=
+%polTable(bd.2010.ind,bd.0.2010.ind,tablabel="polzeros")
+%@
 
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: Objekt 'estvReink' nicht gefunden}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: undefined columns selected}}\end{kframe}
-\includegraphics[width=\maxwidth]{figure/with_without_zeros_RelDist_mitESTV1} 
-\begin{kframe}\begin{verbatim}
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/with_without_zeros_RelDist_mitESTV2}
-
-\end{knitrout}
-
-% latex.default(format(c(`Median Index` = med, `Lower Index` = lower,      `Upper Index` = upper, `$\\Delta$ Gini` = dgini), digits = ndig),      file = "", title = title) 
-%
-\begin{table}[!tbp]
-\begin{center}
-\begin{tabular}{ll}
-\hline\hline
-\multicolumn{1}{l}{Inequality Indices}&\multicolumn{1}{c}{}\tabularnewline
-\hline
-Median Index&-0.1084\tabularnewline
-Lower Index&-0.2155\tabularnewline
-Upper Index&-0.0013\tabularnewline
-$\Delta$ Gini&-0.0952\tabularnewline
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-Including zeros leads to significantly higher gini coefficients.
+%Including zeros leads to significantly higher gini coefficients.
 However we must keep in mind, that these might be artificially high values as we assume zero income for everyone in the zero group.
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{verbatim}
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/BruelhartvsHABE1} 
-\begin{kframe}\begin{verbatim}
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/BruelhartvsHABE2}
-
-\includegraphics[width=\maxwidth]{figure/BruelhartvsHABE3} 
-\begin{kframe}\begin{verbatim}
-## Smoothing using 0.4
-## Smoothing using 0.4
-## Smoothing using 0.4
-\end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/BruelhartvsHABE4}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/BruelhartvsHABE} \caption[Comparison of tax and survey data]{Comparison of tax and survey data\label{fig:BruelhartvsHABE}}
+\end{figure}
 
@@
 
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/BruelhartvsHABEmarried} \caption[Comparison of tax and survey data for married]{Comparison of tax and survey data for married\label{fig:BruelhartvsHABEmarried}}
+\end{figure}
+
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/BruelhartvsHABEunmarried} \caption[Comparison of tax and survey data for unmarried]{Comparison of tax and survey data for unmarried\label{fig:BruelhartvsHABEunmarried}}
+\end{figure}
+
+\end{knitrout}
+
+From the discussion in the data section we would expect differences between the income distributions from survey and tax data.
+Within the FTA data we observe zeros for incomes below the threshold to be liable for federal tax while survey data might cover this range.
+On the other hand we expect underreporting from both lowest percentiles and highest percentiles (middle class bias) within the survey data.
+Figure \ref{fig:BruelhartvsHABE} (left) however reveals a more critical issue related to tax data, that is the median location of income compared to survey data.
+Although we try to measure similar concepts of income, survey data shows a median (93.000 CHF) more than twice as big as tax data (44.600 CHF).
+The issue here is clearly the assumptions of household composition.
+While survey data is likely to capture the correct household composition, tax data can only approximate households by using marital status.
+Splitting the data in married and unmarried tax units underpins this argument.
+From figure \ref{fig:BruelhartvsHABEmarried} we can see that married tax units (FTA data) and household with married couples (survey data) are better (but still not perfectly) comparable.
+Figure \ref{fig:BruelhartvsHABEmarried} (right) shows the expected shape difference between the two distributions:  survey data has a bias towards the (median-adjusted) 80\% to 90\% percentile (of the FTA data distribution).
+Top percentiles are badly covered by survey data, suggesting that inequality measures based on survey data might underestimate inequality development that arises from changes in the incomes of the rich.
+Though, survey data can be of interest if one is interested in the lower 20\% of the income distribution.
+
+The bad household approximation within tax data needs to be addressed in further research. One possible way would be to link tax data to household ID's from the residential register.
+
 \subsection{Intertemporal comparison}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\maxwidth]{figure/benplot}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[H]
+
+\includegraphics[width=\maxwidth]{figure/benplot} \caption[Gini coefficients over time]{Gini coefficients over time\label{fig:benplot}}
+\end{figure}
 
@@
 
+Figure\ref{fig:benplot} displays the most relevant Gini-coefficients that can be calculated for Switzerland.
+Although we can not adjust the Gini-coefficients to be perfectly valid, we can discuss the picture against the background of our analyses.
+For the periods before the second world war it is difficult to draw secure conclusions because these data points are based on unreliable data.
+1933 only 13.7\% of the population filled in a tax form \citep{dell_income_2005}.
+During the war we see a tendency towards lower income inequality.
+This might be attributable to a changed data base as the amount of non-fillers decreased during the war and shortly after.
+The period after the world war is characterized by strong economic growth as well as an increase in inequality.
+Our interpretation is that high income percentiles disproportionately profited from the economic upturn.
+After the oil crises there were alternating phases of social welfare expansion and economic upturns.
+
+The more interesting part of the picture are the years around and past the millennium.
+Between 1997/98 and 2003 we see a gap in the FTA series caused by the change in the tax system.
+Our approach to impute the gap is to interpolate the income sums of all FTA-reported income brackets.
+This is fruitful because cantons switched the tax system in different years so we gain at least some information about the trend within the gap.
+The spike 2001 can be explained by tax tricks: within the period the tax system changed, individuals were able to save taxes by shifting parts of their incomes into this period.
+
+After the financial crisis most of the data sources (except EU SILC) state decreasing inequality.
+Our interpretation is that top percentiles lost relevant parts of their incomes from capital investments.
+
 %%%-------------------------------------------------%%%
@@
 \section{Discussion}
+
+This article highlights the main disparities between taxdata based assessment of inequality opossed to given standard concepts of inequality as it is mainly applied with household-surveys.
+We reflected all important concerns based on the case of Switzerland.
+Can we measure what we want to measure?
+Do data include the cases we are interested in?
+The paper revealed tax data afflicted with numerous problems when it comes to analyzing inequality.
+On the pro side tax data is not a sample but covers the full population liable to federal tax.
+The documentation of tax data for almost 100 years makes it the only option if the researcher is interested in long term analyses of swiss income inequality, take it or leave it.
+On the contra side the researcher needs to keep in mind several limitations.
+According to our analyses we can give advice which issues are more or less serious.
+The biggest issue is that tax data does not capture households.
+In comparison to survey data we have seen that tax data extremely overestimates the share of singles/single households if we assume marital status to be a proxy for household composition.
+Luckily there is a way to fix this shortcoming for further research.
+For analyses of household income inequality, the researcher should gather household information, e.g.
+by linking tax data to the population register.
+The second most important drawback of tax data is the non-reporting of the non-taxed (low income) tax units.
+The best solution here would be to gather data directly from the FTA where these information are available, although not published.
+As a minimum the researcher should keep an eye on the share of population ``disappearing'' below the threshold.
+The larger the share, the more likely it is to overlook increases in inequality if tax data is used without these cases.
+Minor issues are whether the researcher looks at all cases (including special cases) the FTA may provide or only normal cases.
+However this issue became more relevant in recent years as high income special cases became more frequent.
+Our recommendation is to use both normal and special cases wherever possible but it is not an important issue.
+It is also not very important for the researcher whether the researcher looks at equivalized or non-equivalized incomes as long as the exact household structure is unknown.
+A more fruitful approach is to do separate analyses for married and unmarried tax units and treat the latter with healthy portion of caution.
 

### 2014-11-25 code/ben_plot.r

@@ -1,0 +1,25 @@
+library(foreign)
+library(dplyr)
+library(reldist)
+ben <- read.dta("C:/Users/Hackstutz/Dropbox/Git/swiss_inequality_development/data/estv-gini-kt.dta")
+estv <- read.dta("C:/Users/Hackstutz/Dropbox/Git/swiss_inequality_development/data/steuerdaten20140522_stata12.dta")
+filter(ben, kanton=="CH")
+
+#estv <- arrange(estv, steuerperiode, kanton, eink_ll)
+#1945 fixen (funktioniert nicht)
+#index45 <- which(estv$steuerperiode==1945.5)
+#estv$steink[index45] <- estv$steink[index45+481]-estv$steink_diff[index45]
+#estv$anz_pflichtige[index45] <- estv$anz_pflichtige[index45+481]-estv$anz_pflichtige_diff[index45]
+
+estv_gini <- estv %.% 
+	filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %.% 
+	select(steuerperiode, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %.%
+	group_by(factor(steuerperiode)) %.%
+	mutate(mean_steink=steink/anz_pflichtige) %.%
+	summarise(gini(mean_steink,weights=anz_pflichtige))
+
+	#%.%
+	#group_by(factor(steuerperiode)) %.%
+	#summarise(gini(steink,weights=anz_pflichtige))
+
+unique(ave(estv_ch$steink, estv_ch$steuerperiode,FUN=gini,weights=estv_ch$anz_pflichtige))

### 2014-11-25 code/poltable_draft.r

@@ -1,0 +1,25 @@
+polTable <- function(y, yo, ywgt=1, yowgt=1,ndig=2,title="Inequality Indices") {
+stopifnot(is.numeric(y), is.numeric(yo)) # all TRUE
+#y vector with new distribution
+#yo vector with reference distribution
+#ywgt yowgt: weights to use (optional)
+#ndig: digits to display, default 2
+#title: table title (optional)
+
+# load neccessary libraries
+library(reldist,quietly=TRUE,warn.conflicts=FALSE) #for Latex tables
+library(Hmisc,quietly=TRUE,warn.conflicts=FALSE)
+
+# Create vector of weights 1 if nothing else is defined
+if(length(ywgt)==1 & ywgt==1) ywgt <- rep(1, length(y))
+if(length(yowgt)==1 & yowgt==1) yowgt <- rep(1, length(yo))
+
+med <- rpy(y=y,yo=yo, ywgt=ywgt,yowgt=yowgt,pvalue=FALSE)[2] # Median Index, middle value is the estimate so we take [2]
+lower <- rpluy(y=y,yo=yo, ywgt=ywgt,yowgt=yowgt,pvalue=FALSE)[2] # Lower Index
+upper <- rpluy(y=y,yo=yo, ywgt=ywgt,yowgt=yowgt,pvalue=FALSE, upper=TRUE)[2] # Upper index
+dgini <- gini(y)-gini(yo)
+
+latex(format(c("Median Index"=med, "Lower Index"=lower, "Upper Index"=upper, "$\\Delta$ Gini"=dgini),digits=ndig),file="",title=title)
+}
+
+polTable(bd.2010.ind,bd.2003.ind)

### 2014-11-25 paper/subdocuments/results.Rnw
update reldist graph

@@
 bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] <- bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] - 2
-bd<-bd %.% 
+bd<-bd %>% 
   filter(Einheit=="Total")
@@
 bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
-bd.0<-bd.0 %.% 
+bd.0<-bd.0 %>% 
   filter(Einheit=="Total")
@@
 bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
-bd.2003<-bd.0 %.% 
+bd.2003<-bd.0 %>% 
   filter(Jahr==2003,Einheit=="Total")
@@
 bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
-bd.2010<-bd.0 %.% 
+bd.2010<-bd.0 %>% 
   filter(Jahr==2010,Einheit=="Total")
@@
 normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
-#normal <- normal %.% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
+#normal <- normal %>% select(-G_reink, -G_taxed, -ppop0, -G_steink0)
 pooled <- read.csv("../../data/ginis_und_perzentile_beides.csv",sep="\t")
@@
 d<-rbind(normal,pooled)
-d<-d %.% 
+d<-d %>% 
   filter(kanton=="CH",steuerperiode==1993.5)
@@
 
-habe10.married<-habe0911 %.% 
+habe10.married<-habe0911 %>% 
   filter(Familientyp08>130,Jahr08==2010)
@@
 # Select unmarried - HABE
-habe10<-habe0911 %.% 
+habe10<-habe0911 %>% 
   filter(Jahr08==2010)
-habe10.unmarried<-habe0911 %.% 
-  filter(Familientyp08<140,Jahr08==2010) %.%
+habe10.unmarried<-habe0911 %>% 
+  filter(Familientyp08<140,Jahr08==2010) %>%
   select(VerfuegbaresEinkommen08,jahresgewicht)
@@
 
+#bern kantonal (liegt lokal wegen datenschutz)
+bern_kt <- read.dta("H:/Assessing Inequalty/taxDataBE2012.dta")
+bern_kt2012 <- bern_kt %>% filter(stj==2012) %>% group_by(hh_id) %>% select(tot_erwerbeink,tot_vermeink,primeink,tot_transeink,TOTEINK,hh_id)
+#summarise(bern_kt2012, n=n()) %>% arrange(desc(n))
+#habe0911 %>%mutate(anzahl_personen=Anzahl1524Personen08+Anzahl2534Personen08+Anzahl3544Personen08+Anzahl4554Personen08+Anzahl5564Personen08+Anzahl6574Personen08+Anzahl7599Personen08) %>% group_by(anzahl_personen) %>% summarise(n=n())%>%arrange(desc(n))
+# gr??ter haushalt 7 personen gross (>=15jahre)
+# dort machen wir einen schnitt bei den kantonalen daten um wohnheime etc auszuschlie?en (wie es HABE auch gemacht hat)
+bern_kt2012_plain <- bern_kt2012  %>% filter(n()<=7) #5% werden gefiltert
+bern_kt2012 <- bern_kt2012_plain %>% mutate(TOTEINK_hh=sum(TOTEINK)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2. bis x.
+
+# f?r habe (bern) abz?glich sozialversicherungsbeitr?ge
+habe11_bern <- habe0911 %>% filter(Kanton08==2) %>% 
+  mutate(income=(Bruttoeinkommen08+A31m)*12/1000,weight=jahresgewicht,group=0) %>% 
+  select(income, weight,group)
+
+# export nach stata (ungewichtet)
+bern_habe <- rbind(habe11_bern,data.frame("income" = bern_kt2012$TOTEINK_hh/1000,"group"=1,"weight"=1))
+bern_habe$group <- factor(bern_habe$group,labels=c("habe","tax"))
+bern_habe <- bern_habe %>% mutate(income=ifelse(income<0,0,income))
+write.dta(bern_habe, file="../../data/stata_data/bern_habe.dta")
+
+# vergleich innerhalb von bern tax data: haushalte vs. plain
+# oben wurde plain und nicht plain definiert, das fügen wir jetzt zusammen
+# export nach stata (ungewichtet)
+bern_plain_hh <- rbind(data.frame("income" = bern_kt2012_plain$TOTEINK/1000,"group"=0,"weight"=1),data.frame("income" = bern_kt2012$TOTEINK_hh/1000,"group"=1,"weight"=1))
+bern_plain_hh$group <- factor(bern_plain_hh$group,labels=c("tax units","households"))
+bern_plain_hh <- bern_plain_hh %>% mutate(income=ifelse(income<0,0,income))
+write.dta(bern_plain_hh, file="../../data/stata_data/bern_plain_hh.dta")
+
 @
@@
 ### and for the rest of Ginis:
-estv <- read.dta("../../data/steuerdaten20140522_stata12.dta")
+estv <- read.dta("../../data/steuerdaten20141118_stata12.dta")
 
@@
 
-#estv_gini <- estv %.% 
-#  filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %.% 
-#  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %.%
-#  group_by(factor(bemessungsjahre)) %.%
-#  mutate(mean_steink=steink/anz_pflichtige) %.%
+#estv_gini <- estv %>% 
+#  filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %>% 
+#  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %>%
+#  group_by(factor(bemessungsjahre)) %>%
+#  mutate(mean_steink=steink/anz_pflichtige) %>%
 #  summarise(gini(mean_steink,weights=anz_pflichtige))

### 2014-11-28 paper/subdocuments/results.Rnw
update both graphics (6er Stata and 4er R figures)

@@
 <<different_ginis, fig.cap='Gini coefficients for different income measures',fig.pos='H'>>=
-df$G_reink[df$G_reink==1]<-NA
 df$G_taxed[df$G_taxed==1]<-NA
@@
 levels(dflong$Income)<-c("taxable income","net income","taxable income after federal tax")
-ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,shape=Income))+
-  scale_y_continuous(limits=c(0.3,0.55))+
-  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
+fig_incomes <- ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,shape=Income))+
+  scale_y_continuous(limits=c(0.28,0.55))+
+  scale_x_continuous(limits=c(1941,2011),breaks=number_ticks(10)) +
   geom_point(aes(shape=Income),size=3)+
@@
   theme_bw()+
-  xlab("tax period")+
-  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+  xlab("Year")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
+  ggtitle("Using net, taxable or taxed income")
+fig_incomes
 #ggplot(dflong, aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")
@@
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
-ggplot(bdlong, aes(x=Jahr,y=Gini,shape=Income))+
-  scale_y_continuous(limits=c(0.3,0.55))+
-  scale_x_continuous(limits=c(1970,2012),breaks=number_ticks(10)) +
+fig_equiv <- ggplot(bdlong, aes(x=Jahr,y=Gini,shape=Income))+
+  scale_y_continuous(limits=c(0.28,0.55))+
+  #scale_x_continuous(limits=c(1970,2012),breaks=number_ticks(10)) +
+  scale_x_continuous(limits=c(1970,2011),breaks=number_ticks(10)) +
   geom_point(aes(shape=Income),size=3)+
@@
   theme_bw()+
-  xlab("tax period")+
-  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+  xlab("Year")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
+  ggtitle("Using (pseudo) equivalence scales")
+fig_equiv
 @
@@
 levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
-ggplot(dflong[dflong$kanton=="CH" & dflong$steuerperiode>=1995.5,], 
+fig_zeros <- ggplot(dflong[dflong$kanton=="CH",], 
        aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
-  scale_y_continuous(limits=c(0.18,0.6))+
-  scale_x_continuous(limits=c(1994,2012),breaks=number_ticks(10)) +
+  scale_y_continuous(limits=c(0.28,0.55))+
+  scale_x_continuous(limits=c(2003,2012),breaks=number_ticks(10)) +
   geom_line(size=0.5)+
@@
   xlab("tax period")+
-  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
+  ggtitle("How to treat non-taxed")
+fig_zeros
 
@@
 bern_kt2012_plain <- bern_kt2012  %>% filter(n()<=7) #5% werden gefiltert
-bern_kt2012 <- bern_kt2012_plain %>% mutate(TOTEINK_hh=sum(TOTEINK)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2. bis x.
+#bern_kt2012 <- bern_kt2012_plain %>% mutate(TOTEINK_hh=sum(TOTEINK)) %>% filter(!duplicated(hh_id))
+bern_kt2012 <- bern_kt2012_plain %>% mutate(primeink_hh=sum(primeink)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2. bis x.
 
@@
 # export nach stata (ungewichtet)
-bern_habe <- rbind(habe11_bern,data.frame("income" = bern_kt2012$TOTEINK_hh/1000,"group"=1,"weight"=1))
+bern_habe <- rbind(habe11_bern,data.frame("income" = bern_kt2012$primeink_hh/1000,"group"=1,"weight"=1))
 bern_habe$group <- factor(bern_habe$group,labels=c("habe","tax"))
@@
 # export nach stata (ungewichtet)
-bern_plain_hh <- rbind(data.frame("income" = bern_kt2012_plain$TOTEINK/1000,"group"=0,"weight"=1),data.frame("income" = bern_kt2012$TOTEINK_hh/1000,"group"=1,"weight"=1))
+bern_plain_hh <- rbind(data.frame("income" = bern_kt2012_plain$primeink/1000,"group"=0,"weight"=1),data.frame("income" = bern_kt2012$primeink_hh/1000,"group"=1,"weight"=1))
 bern_plain_hh$group <- factor(bern_plain_hh$group,labels=c("tax units","households"))
@@
 ### and for the rest of Ginis:
-estv <- read.dta("../../data/steuerdaten20141118_stata12.dta")
+estv <- read.dta("../../data/steuerdaten20141126_stata12.dta")
 
@@
 
-#estv_gini <- estv %>% 
-#  filter(kanton=="CH",eink_art=="Nach steuerbarem Einkommen",fall=="Normalfälle",anz_pflichtige>0,steuerperiode!=1945.5) %>% 
-#  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %>%
-#  group_by(factor(bemessungsjahre)) %>%
-#  mutate(mean_steink=steink/anz_pflichtige) %>%
-#  summarise(gini(mean_steink,weights=anz_pflichtige))
-
-load("estv_gini.RData")
+estv_gini <- estv %>% 
+  filter(kanton=="CH",
+         eink_art=="Nach steuerbarem Einkommen",
+         fall=="Normalfälle",
+         anz_pflichtige>0) %>%
+        # steuerperiode!=1945.5) %>% 
+  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %>%
+  group_by(factor(bemessungsjahre)) %>%
+  mutate(mean_steink=steink/anz_pflichtige) %>%
+  summarise(gini(mean_steink,weights=anz_pflichtige))
+
+#load("estv_gini.RData")
 
@@
 
+# kleiner einschub: verschiedene masse
+library(IC2)
+getAtkinson <- function(x,w){
+  calcAtkinson(x,w=w)$ineq$index
+}
+getTheil <- function(x,w){
+  calcGEI(x,w=w,alpha=1)$ineq$index
+}
+
+estv_masse <- estv %>% 
+  filter(kanton=="CH",
+         eink_art=="Nach steuerbarem Einkommen",
+         fall=="Normalfälle",
+         anz_pflichtige>0) %>%
+        # steuerperiode!=1945.5) %>% 
+  select(bemessungsjahre, steink, kanton, eink_ll, eink_ul,anz_pflichtige) %>%
+  group_by(factor(bemessungsjahre)) %>%
+  mutate(mean_steink=steink/anz_pflichtige) %>%
+  summarise(Gini=gini(mean_steink,weights=anz_pflichtige),Atkinson=getAtkinson(mean_steink,w=anz_pflichtige),Theil=getTheil(mean_steink,w=anz_pflichtige))
+estv_masse$id<-1
+names(estv_masse)<-c("Year","Gini","Atkinson","Theil","id")
+try(detach("package:reshape", unload=TRUE))
+library(reshape2)
+estv_masse_long <- melt(estv_masse,id=c("id","Year"))
+
+estv_masse_long$measure <- factor(estv_masse_long$variable)
+estv_masse_long$Year <- as.double(as.character(estv_masse_long$Year))
+fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log(value),shape=measure))+
+  #scale_y_continuous(limits=c(0,0.55))+
+  scale_x_continuous(limits=c(1918,2011),breaks=number_ticks(10)) +
+  geom_point(aes(shape=measure),size=3)+
+  geom_line(aes(linetype=measure))+
+  theme_bw()+
+  xlab("Year")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
+  ggtitle("Using different inequality measures")
+fig_measures
+
+# output als 4er grafik
+library(gridExtra)
+png("../figure/figure1234.png",width=1200,height=675)
+grid.arrange(fig_measures,fig_equiv,fig_incomes,fig_zeros)
+dev.off()
 @

### 2014-12-03 paper/subdocuments/results.Rnw
4er und 6er grafik angepasst d.Gini berechnet

@@
 fig_incomes
+# was ist die maximale Bandbreite zwischen den Punkten?
+max(filter(df,kanton=="CH") %>% mutate(d_steink_reink=abs(G_steink-G_reink),d_steink_taxed=abs(G_steink-G_taxed),d_reink_taxed=abs(G_reink-G_taxed)) %>% select(starts_with("d_")),na.rm=TRUE)
+# 0.026
+
 #ggplot(dflong, aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")
@@
 fig_equiv
+# Gini Range
+max(bd %>% mutate(d_steink=abs(gini_steink-gini_steinka),d_reink=abs(gini_reink-gini_reinka)) %>% select(starts_with("d_")),na.rm=TRUE)
+# 0.011
 @
@@
   scale_y_continuous(limits=c(0.28,0.55))+
-  scale_x_continuous(limits=c(2003,2012),breaks=number_ticks(10)) +
+  scale_x_continuous(limits=c(2003,2011),breaks=number_ticks(10)) +
   geom_line(size=0.5)+
@@
 fig_zeros
+
+# Gini range
+max(filter(df,kanton=="CH") %>% mutate(d_with_without=abs(G_steink-G_steink0)) %>% select(starts_with("d_")),na.rm=TRUE)
+# 0.18
 
@@
 bern_kt <- read.dta("H:/Assessing Inequalty/taxDataBE2012.dta")
-bern_kt2012 <- bern_kt %>% filter(stj==2012) %>% group_by(hh_id) %>% select(tot_erwerbeink,tot_vermeink,primeink,tot_transeink,TOTEINK,hh_id)
+bern_kt2012 <- bern_kt %>% filter(stj==2012) %>% group_by(hh_id) %>% select(tot_erwerbeink,tot_vermeink,primeink,tot_transeink,TOTEINK,Satzeink_BUND,hh_id)
 #summarise(bern_kt2012, n=n()) %>% arrange(desc(n))
@@
 #bern_kt2012 <- bern_kt2012_plain %>% mutate(TOTEINK_hh=sum(TOTEINK)) %>% filter(!duplicated(hh_id))
-bern_kt2012 <- bern_kt2012_plain %>% mutate(primeink_hh=sum(primeink)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2. bis x.
+bern_kt2012 <- bern_kt2012_plain %>% mutate(primeink_hh=sum(primeink),Satzeink_BUND_hh=sum(Satzeink_BUND)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2. bis x.
 
@@
 write.dta(bern_habe, file="../../data/stata_data/bern_habe.dta")
+# Ginis berechnen
+gini_habe <- gini(bern_habe$income[bern_habe$group=="habe"],weights=bern_habe$weight[bern_habe$group=="habe"]); gini_habe
+gini_tax <- gini(bern_habe$income[bern_habe$group=="tax"]); gini_tax
+round(gini_habe-gini_tax,3)
 
@@
 # export nach stata (ungewichtet)
-bern_plain_hh <- rbind(data.frame("income" = bern_kt2012_plain$primeink/1000,"group"=0,"weight"=1),data.frame("income" = bern_kt2012$primeink_hh/1000,"group"=1,"weight"=1))
+bern_plain_hh <- rbind(data.frame("income" = bern_kt2012_plain$Satzeink_BUND/1000,"group"=0,"weight"=1),data.frame("income" = bern_kt2012$Satzeink_BUND_hh/1000,"group"=1,"weight"=1))
 bern_plain_hh$group <- factor(bern_plain_hh$group,labels=c("tax units","households"))
@@
 write.dta(bern_plain_hh, file="../../data/stata_data/bern_plain_hh.dta")
+# Ginis berechnen
+gini_taxunits <- gini(bern_plain_hh$income[bern_plain_hh$group=="tax units"]); gini_taxunits
+gini_hh <- gini(bern_plain_hh$income[bern_plain_hh$group=="households"]); gini_hh
+round(gini_taxunits-gini_hh,3)
 
@@
 all_gini$LIS <- swissGini$LIS[match(all_gini$Year,swissGini$Year)]
-all_gini$FTA_imp <- ben$gini[match(all_gini$Year,ben$jahr)]
+#all_gini$FTA_imp <- ben$gini[match(all_gini$Year,ben$jahr)]
 
@@
 
-ggplot(data=na.omit(all_gini_long),
+benplot <- ggplot(data=na.omit(all_gini_long),
   aes(y=Gini,x=Year,shape=Data,linetype=Data))+
   ylab("Gini from different sources") +
-  scale_y_continuous(limits=c(0.18,0.4)) +
-  scale_x_continuous(limits=c(1915,2012),breaks=number_ticks(10)) +
+  scale_y_continuous(limits=c(0.18,0.45)) +
+  scale_x_continuous(limits=c(1945,2012),breaks=number_ticks(10)) +
   geom_point(size=3)+
   geom_line()+
-  geom_text(aes(label=annotation,y=ypos),color="#555555",size=3)+
-  geom_segment(aes(x = 1921, y = 0.27, xend = 1921, yend = Gini[Year==1921]),color="#aaaaaa") +
-  geom_segment(aes(x = 1940, y = 0.305, xend = 1940, yend = Gini[Year==1940]),color="#aaaaaa") +
-  geom_segment(aes(x = 1945.5, y = 0.29, xend = 1945.5, yend = Gini[Year==1945.5]),color="#aaaaaa") +
-  geom_segment(aes(x = 1967.5, y = 0.285, xend = 1967.5, yend = Gini[Year==1967.5]),color="#aaaaaa") +
+  #geom_text(aes(label=annotation,y=ypos),color="#555555",size=3)+
+  #geom_segment(aes(x = 1921, y = 0.27, xend = 1921, yend = Gini[Year==1921]),color="#aaaaaa") +
+  #geom_segment(aes(x = 1940, y = 0.305, xend = 1940, yend = Gini[Year==1940]),color="#aaaaaa") +
+  #geom_segment(aes(x = 1945.5, y = 0.29, xend = 1945.5, yend = Gini[Year==1945.5]),color="#aaaaaa") +
+  #geom_segment(aes(x = 1967.5, y = 0.285, xend = 1967.5, yend = Gini[Year==1967.5]),color="#aaaaaa") +
   theme_bw()+
   theme(legend.justification=c(0,1),legend.position=c(0, 1))
+
+png("../figure/ben_plot.png",width=800,height=450)
+benplot
+dev.off()
 
@@
 estv_masse_long$Year <- as.double(as.character(estv_masse_long$Year))
-fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log(value),shape=measure))+
+estv_masse_long <- estv_masse_long %>% group_by(measure) %>% mutate(log_value=log(value)-log(value)[1])
+fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log_value,shape=measure))+
   #scale_y_continuous(limits=c(0,0.55))+
-  scale_x_continuous(limits=c(1918,2011),breaks=number_ticks(10)) +
+  scale_x_continuous(limits=c(1917,2011),breaks=number_ticks(10)) +
   geom_point(aes(shape=measure),size=3)+
@@
 png("../figure/figure1234.png",width=1200,height=675)
-grid.arrange(fig_measures,fig_equiv,fig_incomes,fig_zeros)
+grid.arrange(fig_incomes,fig_equiv,fig_measures,fig_zeros)
 dev.off()

### 2014-12-03 paper/subdocuments/results.Rnw
gini berechnet für special cases

@@
 write.dta(ns1993_94, file="../../data/stata_data/ns1993_94.dta")
+# Ginis berechnen
+gini_allcases <- gini(ns1993_94$income[ns1993_94$group=="pooled"]); gini_allcases
+gini_normal <- gini(ns1993_94$income[ns1993_94$group=="normal"]); gini_normal
+round(gini_allcases-gini_normal,3)
+# 0.013
 
@@
 write.dta(ns2010, file="../../data/stata_data/ns2010.dta")
+
+#Gini Range
+gini_allcases <- gini(ns2010$income[ns1993_94$group=="pooled"]); gini_allcases
+gini_normal <- gini(ns2010$income[ns1993_94$group=="normal"]); gini_normal
+round(gini_allcases-gini_normal,3)
+# 0.025
 

### 2014-12-03 paper/subdocuments/results.Rnw
vertippt: fixed

@@
 #Gini Range
-gini_allcases <- gini(ns2010$income[ns1993_94$group=="pooled"]); gini_allcases
-gini_normal <- gini(ns2010$income[ns1993_94$group=="normal"]); gini_normal
+gini_allcases <- gini(ns2010$income[ns2010$group=="pooled"]); gini_allcases
+gini_normal <- gini(ns2010$income[ns2010$group=="normal"]); gini_normal
 round(gini_allcases-gini_normal,3)
-# 0.025
+# 0.02
 

### 2014-12-05 paper/subdocuments/results.Rnw
Corrected income definition for HABE

@@
 habe11_bern <- habe0911 %>% filter(Kanton08==2) %>% 
-  mutate(income=(Bruttoeinkommen08+A31m)*12/1000,weight=jahresgewicht,group=0) %>% 
+  mutate(income=(Primaereinkommen08)*12/1000,weight=jahresgewicht,group=0) %>% 
   select(income, weight,group)

### 2014-12-07 paper/subdocuments/results.Rnw
grosses update. 4er und 6er plot. update von 2010 auf 2011

@@
 bd<-bd %>% 
-  filter(Einheit=="Total")
+  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_reink,gini_reinka)
 bd<-subset(bd,Jahr<1995 | Jahr >2002)
+# 2011 zufügen
+bd.2011 <- read.csv("../../data/NP_2011.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011)
+levels(bd.0$Veranlagungsperiode)<- c(levels(bd.0$Veranlagungsperiode),"2011")
+bd <- rbind(bd,bd.2011[,names(bd)])
 bd.long<-reshape(bd,
@@
 bd.0<-bd.0 %>% 
-  filter(Einheit=="Total")
+  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_reink,gini_reinka)
 bd.0<-subset(bd.0,Jahr<1993 | Jahr >2002)
+# 2011 zufügen
+bd.0.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011)
+levels(bd.0$Veranlagungsperiode)<- c(levels(bd.0$Veranlagungsperiode),"2011")
+bd.0 <- rbind(bd.0,bd.0.2011[,names(bd.0)])
 bd.0.long<-reshape(bd.0,
@@
 bdlong<-rbind(bd.0.long,bd.long)
-names(bdlong)[61]<-"Income"
+names(bdlong)[3]<-"Income"
 @
@@
 # Gini Range
-max(bd %>% mutate(d_steink=abs(gini_steink-gini_steinka),d_reink=abs(gini_reink-gini_reinka)) %>% select(starts_with("d_")),na.rm=TRUE)
-# 0.011
+max(bd %>% mutate(d_reink=abs(gini_reink-gini_reinka)) %>% select(starts_with("d_")),na.rm=TRUE)
+max(bd.0 %>% mutate(d_reink=abs(gini_reink-gini_reinka)) %>% select(starts_with("d_")),na.rm=TRUE)
+# 0.012
 @
@@
 bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
-bd.2010<-bd.0 %>% 
-  filter(Jahr==2010,Einheit=="Total")
-start <- which(names(bd.2010)=="p1")
-end <- which(names(bd.2010)=="p99_99")
+bd.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(Einheit=="Total")
+#bd.2010<-bd.0 %>% 
+#  filter(Jahr==2010,Einheit=="Total")
+#start <- which(names(bd.2010)=="p1")
+#end <- which(names(bd.2010)=="p99_99")
+#percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+#bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
+#names(bd.2010)[1]=c("inc")
+#bd.2010 <- rbind(c(0,0),bd.2010)
+#ival<-bd.2010$inc
+#perval<-bd.2010$percentile
+#bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+#  len=(perval[i]-perval[i-1])*100
+#  seq(ival[i-1],ival[i],length.out=len)
+#}))
+#bd.2010.ind<-bd.2010.ind/1000
+
+start <- which(names(bd.2011)=="p1")
+end <- which(names(bd.2011)=="p99_99")
 percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
-bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
-names(bd.2010)[1]=c("inc")
-bd.2010 <- rbind(c(0,0),bd.2010)
-ival<-bd.2010$inc
-perval<-bd.2010$percentile
-bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+bd.2011<- data.frame(t(bd.2011[1,start:end]),percentile)
+names(bd.2011)[1]=c("inc")
+bd.2011 <- rbind(c(0,0),bd.2011)
+ival<-bd.2011$inc
+perval<-bd.2011$percentile
+bd.2011.ind<-unlist(sapply(2:length(ival),function(i){
   len=(perval[i]-perval[i-1])*100
@@
 }))
-bd.2010.ind<-bd.2010.ind/1000
+bd.2011.ind<-bd.2011.ind/1000
 @
@@
 # Export data to use with Stata
-bd2010_2003 <- rbind(data.frame(cbind("bd" = bd.2010.ind,"group"=2010)),data.frame("bd" = bd.2003.ind,"group"=2003))
-write.dta(bd2010_2003, file="../../data/stata_data/bd2010_2003.dta")
+bd2011_2003 <- rbind(data.frame(cbind("bd" = bd.2011.ind,"group"=2011)),data.frame("bd" = bd.2003.ind,"group"=2003))
+write.dta(bd2011_2003, file="../../data/stata_data/bd2011_2003.dta")
 @
@@
 
-% Variante 2010
-<<setup_data_specialcases10>>=
+% Variante 2011
+<<setup_data_specialcases11>>=
 #library(foreign)
@@
 #library(reldist)
-normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2010)
+normal <- filter(read.csv("../../data/ginis_und_perzentile_normal.csv",sep="\t"),kanton=="CH",steuerperiode==2011)
 normal <- select(normal, p1:p99, kanton, steuerperiode, -p96,-p97,-p98)
@@
 #bruelhart data
-pooled <- filter(read.csv("../../data/data_Schweiz.csv",sep=";"), Veranlagungsperiode==2010, Einheit=="Total")
+#pooled <- filter(read.csv("../../data/data_Schweiz.csv",sep=";"), Veranlagungsperiode==2010, Einheit=="Total")
+pooled <- read.csv("../../data/NP_2011.csv",dec=",") %>% filter(Einheit=="Total")
 start <- which(names(pooled)=="p1")
@@
 
-<<specialcases10, fig.cap='Comparison of normal and special cases 2010',fig.pos='H'>>=
+<<specialcases11, fig.cap='Comparison of normal and special cases 2011',fig.pos='H'>>=
 
@@
 # Export data to use with Stata
-ns2010 <- rbind(data.frame(cbind("income" = pooled_i,"group"=0)),data.frame("income" = normal_i,"group"=1))
-ns2010$group <- factor(ns2010$group,labels=c("pooled","normal"))
-write.dta(ns2010, file="../../data/stata_data/ns2010.dta")
+ns2011 <- rbind(data.frame(cbind("income" = pooled_i,"group"=0)),data.frame("income" = normal_i,"group"=1))
+ns2011$group <- factor(ns2011$group,labels=c("pooled","normal"))
+write.dta(ns2011, file="../../data/stata_data/ns2011.dta")
 
 #Gini Range
-gini_allcases <- gini(ns2010$income[ns2010$group=="pooled"]); gini_allcases
-gini_normal <- gini(ns2010$income[ns2010$group=="normal"]); gini_normal
-round(gini_allcases-gini_normal,3)
+#gini_allcases <- gini(ns2011$income[ns2011$group=="pooled"]); gini_allcases
+#gini_normal <- gini(ns2011$income[ns2011$group=="normal"]); gini_normal
+#round(gini_allcases-gini_normal,3)
 # 0.02
-
+# Es ist quatsch hier die interpolierten Reihen zu nehmen
+# Gini 2011 Pooled (BH-Daten): 0.408
+# Gini 2011 Normal (unsere ESTV Daten): 0.366
 @
@@
 # Export data to use with Stata
-habe_bd_2010_married <- rbind(data.frame(cbind("income" = habe10.married$inc_habe,"group"=0,"weight"=habe10.married$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
-habe_bd_2010_married$group <- factor(habe_bd_2010_married$group,labels=c("habe","bd"))
-write.dta(habe_bd_2010_married, file="../../data/stata_data/habe_bd_2010_married.dta")
+#habe_bd_2010_married <- rbind(data.frame(cbind("income" = habe10.married$inc_habe,"group"=0,"weight"=habe10.married$gew)),data.frame("income" = bd.2010.ind,"group"=1,"weight"=1))
+#habe_bd_2010_married$group <- factor(habe_bd_2010_married$group,labels=c("habe","bd"))
+#write.dta(habe_bd_2010_married, file="../../data/stata_data/habe_bd_2010_married.dta")
+
+# 2011
+habe11.married<-habe0911 %>% 
+  filter(Familientyp08>130,Jahr08==2011) %>% 
+  mutate(psteink=Bruttoeinkommen08+A31m+A35m) 
+# Wir brauchen (a) verfügbares Einkommen (b) Bruttoeinkommen und c(Bruttoeinkommen nach Soziversicherungsbeiträgen(A31m) und Monetäre Trasnferausgaben an andere HaushalteA35m)
+habe11.married<-data.frame(habe11.married$VerfuegbaresEinkommen08*12/1000,habe11.married$Bruttoeinkommen08*12/1000,habe11.married$psteink*12/1000,habe11.married$jahresgewicht)
+names(habe11.married)<-c("inc_habe","brutto","psteink","gew")
+habe11.married$inc_habe[habe11.married$inc_habe<0]<-0 # One values is -500
+
+bd.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(grepl("Verheiratet",Einheit))
+start <- which(names(bd.2011)=="p1")
+end <- which(names(bd.2011)=="p99_99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+bd.2011<- data.frame(t(bd.2011[1,start:end]),percentile)
+names(bd.2011)[1]=c("inc")
+bd.2011 <- rbind(c(0,0),bd.2011)
+ival<-bd.2011$inc
+perval<-bd.2011$percentile
+bd.2011.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+bd.2011.ind<-bd.2011.ind/1000
+
+habe_bd_2011_married <- rbind(data.frame(cbind("income" = habe11.married$inc_habe,"group"=0,"weight"=habe11.married$gew)),data.frame("income" = bd.2011.ind,"group"=1,"weight"=1))
+habe_bd_2011_married$group <- factor(habe_bd_2011_married$group,labels=c("habe","bd"))
+write.dta(habe_bd_2011_married, file="../../data/stata_data/habe_bd_2011_married.dta")
 
@@
   #scale_y_continuous(limits=c(0,0.55))+
-  scale_x_continuous(limits=c(1917,2011),breaks=number_ticks(10)) +
+  scale_x_continuous(limits=c(1941,2011),breaks=number_ticks(10)) +
   geom_point(aes(shape=measure),size=3)+

### 2014-12-10 paper/subdocuments/results.Rnw
minor fixes 6er plot

@@
 bern_kt <- read.dta("H:/Assessing Inequalty/taxDataBE2012.dta")
-bern_kt2012 <- bern_kt %>% filter(stj==2012) %>% group_by(hh_id) %>% select(tot_erwerbeink,tot_vermeink,primeink,tot_transeink,TOTEINK,Satzeink_BUND,hh_id)
+set0 <- function(x) ifelse(x<0,0,x)
+bern_kt <- bern_kt %>% mutate(primeink2 = set0(hem)+set0(nem)+set0(esm)+set0(sgm)+set0(hef)+set0(nef)+set0(esf)+set0(sgf)+set0(erfolg_s)+set0(erfolg_l)+set0(kk_ertrag)+set0(bg_ertrag)+ set0(wsert)+set0(eink_lieg)+set0(eg_ertrag)+set0(stm)+set0(stf)+set0(lov)+set0(lmv)+set0(lng))
+bern_kt2012 <- bern_kt %>% filter(stj==2012) %>% group_by(hh_id) %>% select(tot_erwerbeink,tot_vermeink,primeink,primeink2,tot_transeink,TOTEINK,Satzeink_BUND,hh_id)
 #summarise(bern_kt2012, n=n()) %>% arrange(desc(n))
@@
 #bern_kt2012 <- bern_kt2012_plain %>% mutate(TOTEINK_hh=sum(TOTEINK)) %>% filter(!duplicated(hh_id))
-bern_kt2012 <- bern_kt2012_plain %>% mutate(primeink_hh=sum(primeink),Satzeink_BUND_hh=sum(Satzeink_BUND)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2. bis x.
+bern_kt2012 <- bern_kt2012_plain %>% mutate(primeink_hh=sum(primeink),primeink2_hh=sum(primeink2),Satzeink_BUND_hh=sum(Satzeink_BUND)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2.
+bis x.
 
@@
 # export nach stata (ungewichtet)
-bern_habe <- rbind(habe11_bern,data.frame("income" = bern_kt2012$primeink_hh/1000,"group"=1,"weight"=1))
+bern_habe <- rbind(habe11_bern,data.frame("income" = bern_kt2012$primeink2_hh/1000,"group"=1,"weight"=1))
 bern_habe$group <- factor(bern_habe$group,labels=c("habe","tax"))
@@
 round(gini_habe-gini_tax,3)
+# 0.457 0.519 diff=-0.061
 

### 2014-12-10 paper/subdocuments/results.Rnw
changed fig_equiv to steink not reink

@@
 bd<-bd %>% 
-  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_reink,gini_reinka)
+  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
 bd<-subset(bd,Jahr<1995 | Jahr >2002)
 # 2011 zufügen
-bd.2011 <- read.csv("../../data/NP_2011.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011)
-levels(bd.0$Veranlagungsperiode)<- c(levels(bd.0$Veranlagungsperiode),"2011")
+bd.2011 <- read.csv("../../data/NP_2011.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011)%>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
+levels(bd$Veranlagungsperiode)<- c(levels(bd$Veranlagungsperiode),"2011")
 bd <- rbind(bd,bd.2011[,names(bd)])
 bd.long<-reshape(bd,
-varying = c("gini_reink","gini_reinka"),
+varying = c("gini_steink","gini_steinka"),
 v.names = "Gini",
 timevar = "inc_def",
-times = c("Net income (only taxed)","Equivalised net income (only taxed)"),
+times = c("taxable income (only taxed)","Equivalised taxable income (only taxed)"),
 direction ="long")
@@
 bd.0<-bd.0 %>% 
-  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_reink,gini_reinka)
+  filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
 bd.0<-subset(bd.0,Jahr<1993 | Jahr >2002)
 # 2011 zufügen
-bd.0.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011)
+bd.0.2011 <- read.csv("../../data/NP_2011_mitnull.csv",dec=",") %>% filter(Einheit=="Total") %>% mutate(Veranlagungsperiode=2011,Jahr=2011) %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
 levels(bd.0$Veranlagungsperiode)<- c(levels(bd.0$Veranlagungsperiode),"2011")
@@
 bd.0.long<-reshape(bd.0,
-varying = c("gini_reink","gini_reinka"),
+varying = c("gini_steink","gini_steinka"),
 v.names = "Gini",
 timevar = "inc_def",
-times = c("Net income","Equivalised net income"),
+times = c("taxable income","Equivalised taxable income"),
 direction ="long")
@@
 # Gini Range
-max(bd %>% mutate(d_reink=abs(gini_reink-gini_reinka)) %>% select(starts_with("d_")),na.rm=TRUE)
-max(bd.0 %>% mutate(d_reink=abs(gini_reink-gini_reinka)) %>% select(starts_with("d_")),na.rm=TRUE)
-# 0.012
+max(bd %>% mutate(d_reink=abs(gini_steink-gini_steinka)) %>% select(starts_with("d_")),na.rm=TRUE)
+max(bd.0 %>% mutate(d_reink=abs(gini_steink-gini_steinka)) %>% select(starts_with("d_")),na.rm=TRUE)
+# 0.011
 @

### 2014-12-10 paper/subdocuments/results.Rnw
added code to include Bern Data from BFH location

@@
 #bern kantonal (liegt lokal wegen datenschutz)
-bern_kt <- read.dta("H:/Assessing Inequalty/taxDataBE2012.dta")
+try(bern_kt <- read.dta("H:/Assessing Inequalty/taxDataBE2012.dta"))
+try(bern_kt<-read.dta("U:/Persönlich/Forschung/Uni Bern/taxDataBE2012.dta"))
 set0 <- function(x) ifelse(x<0,0,x)

### 2014-12-16 paper/subdocuments/results.Rnw
graphen überarbeitet bemessungsjahre statt steuerperiode start jetzt 1945

@@
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
-dflong <- melt(df[,c("G_steink","G_reink","G_taxed","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
+dflong <- melt(df[,c("G_steink","G_reink","G_taxed","kanton","bemessungsjahre")], id=c("kanton","bemessungsjahre"))
 names(dflong)[3]<-"Income"
@@
 levels(dflong$Income)<-c("taxable income","net income","taxable income after federal tax")
-fig_incomes <- ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,shape=Income))+
+fig_incomes <- ggplot(dflong[dflong$kanton=="CH",], aes(x=bemessungsjahre,y=Gini,shape=Income))+
   scale_y_continuous(limits=c(0.28,0.55))+
-  scale_x_continuous(limits=c(1941,2011),breaks=number_ticks(10)) +
+  scale_x_continuous(limits=c(1945,2011),breaks=number_ticks(10)) +
   geom_point(aes(shape=Income),size=3)+
@@
   theme(legend.justification=c(0,1),legend.position=c(0, 1))+
-  ggtitle("Using net, taxable or taxed income")
+  ggtitle("Income definitions within tax data")
 fig_incomes
@@
 timevar = "inc_def",
-times = c("taxable income (only taxed)","Equivalised taxable income (only taxed)"),
+times = c("taxable income (only taxed)","equivalised taxable income (only taxed)"),
 direction ="long")
@@
 timevar = "inc_def",
-times = c("taxable income","Equivalised taxable income"),
+times = c("taxable income","equivalised taxable income"),
 direction ="long")
@@
 bd.2003.ind<-bd.2003.ind/1000
+
+bd.2010<-bd.0 %>% 
+  filter(Jahr==2010,Einheit=="Total")
+start <- which(names(bd.2010)=="p1")
+end <- which(names(bd.2010)=="p99_99")
+percentile <- c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)
+bd.2010<- data.frame(t(bd.2010[1,start:end]),percentile)
+names(bd.2010)[1]=c("inc")
+cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
+bd.2010 <- rbind(c(0,0),bd.2010)
+ival<-bd.2010$inc
+perval<-bd.2010$percentile
+bd.2010.ind<-unlist(sapply(2:length(ival),function(i){
+  len=(perval[i]-perval[i-1])*100
+  seq(ival[i-1],ival[i],length.out=len)
+}))
+bd.2010.ind<-bd.2010.ind/1000
 
@@
 write.dta(bd2011_2003, file="../../data/stata_data/bd2011_2003.dta")
+
+bd2011_2010 <- rbind(data.frame(cbind("bd" = bd.2011.ind,"group"=2011)),data.frame("bd" = bd.2010.ind,"group"=2010))
+write.dta(bd2011_2010, file="../../data/stata_data/bd2011_2010.dta")
 @
@@
 #library(reshape)
-dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
+dflong <- melt(df[,c("G_steink","G_steink0","G_steink_halb","kanton","bemessungsjahre")], id=c("kanton","bemessungsjahre"))
 names(dflong)[3]<-"Type"
 names(dflong)[4]<-"Gini"
-levels(dflong$Type)<-c("without zeros","including zeros","imputed values for zeros")
-fig_zeros <- ggplot(dflong[dflong$kanton=="CH",], 
-       aes(x=steuerperiode,y=Gini,shape=Type,linetype=Type))+
+levels(dflong$Type)<-c("without non-taxed","including non-taxed","imputed values for non-taxed")
+fig_zeros <- ggplot(dflong[dflong$kanton=="CH"&dflong$bemessungsjahre>=2003,], 
+       aes(x=bemessungsjahre,y=Gini,shape=Type,linetype=Type))+
   scale_y_continuous(limits=c(0.28,0.55))+
-  scale_x_continuous(limits=c(2003,2011),breaks=number_ticks(10)) +
+  scale_x_continuous(limits=c(1970,2011),breaks=number_ticks(10)) +
   geom_line(size=0.5)+
@@
   theme_bw()+
-  xlab("tax period")+
+  xlab("Year")+
   theme(legend.justification=c(0,1),legend.position=c(0, 1))+
-  ggtitle("How to treat non-taxed")
+  ggtitle("Influence of non-taxed")
 fig_zeros
@@
   #scale_y_continuous(limits=c(0,0.55))+
-  scale_x_continuous(limits=c(1941,2011),breaks=number_ticks(10)) +
+  scale_x_continuous(limits=c(1945,2011),breaks=number_ticks(10)) +
   geom_point(aes(shape=measure),size=3)+

### 2014-12-16 paper/subdocuments/results.Rnw
modified visual parameters for figure1234

@@
 levels(dflong$Type)<-c("without non-taxed","including non-taxed","imputed values for non-taxed")
-fig_zeros <- ggplot(dflong[dflong$kanton=="CH"&dflong$bemessungsjahre>=2003,], 
+dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<2003&dflong$Type=="including non-taxed"]<-NA
+dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<2003&dflong$Type=="imputed values for non-taxed"]<-NA
+fig_zeros <- ggplot(dflong[dflong$kanton=="CH",], 
        aes(x=bemessungsjahre,y=Gini,shape=Type,linetype=Type))+
@@
 fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log_value,shape=measure))+
-  #scale_y_continuous(limits=c(0,0.55))+
+  scale_y_continuous(limits=c(-0.7,0.45))+
   scale_x_continuous(limits=c(1945,2011),breaks=number_ticks(10)) +

### 2014-12-16 paper/subdocuments/results.Rnw
indexed fig_measures to 1945/1946

@@
 estv_masse$id<-1
+estv_masse<-estv_masse[as.numeric(estv_masse$Year)>4,]
+
 names(estv_masse)<-c("Year","Gini","Atkinson","Theil","id")

### 2014-12-17 paper/subdocuments/results.Rnw
6er plot neu mit bern-daten-update

@@
 #bern kantonal (liegt lokal wegen datenschutz)
-try(bern_kt <- read.dta("H:/Assessing Inequalty/taxDataBE2012.dta"))
-try(bern_kt<-read.dta("U:/Persönlich/Forschung/Uni Bern/taxDataBE2012.dta"))
-set0 <- function(x) ifelse(x<0,0,x)
+#try(bern_kt <- read.dta("H:/Assessing Inequalty/taxDataBE2012.dta"))
+#try(bern_kt<-read.dta("U:/Persönlich/Forschung/Uni Bern/taxDataBE2012.dta"))
+try(bern_kt <- read.csv("G:/stick/bern//BFH-Grunddaten 2012 Output V1.1.csv",sep=";"))
+#> nrow(bern_kt)
+#[1] 622368
+nrow(filter(bern_kt,VON=="2012-01-01",BIS=="2012-12-31"))
+#[1] 605975 wir filtern ca. 3% unterjährige
+bern_kt <- filter(bern_kt,VON=="2012-01-01",BIS=="2012-12-31")
+names(bern_kt)<-tolower(names(bern_kt))
+set0 <- function(x) ifelse(x<0 | is.na(x),0,x)
 bern_kt <- bern_kt %>% mutate(primeink2 = set0(hem)+set0(nem)+set0(esm)+set0(sgm)+set0(hef)+set0(nef)+set0(esf)+set0(sgf)+set0(erfolg_s)+set0(erfolg_l)+set0(kk_ertrag)+set0(bg_ertrag)+ set0(wsert)+set0(eink_lieg)+set0(eg_ertrag)+set0(stm)+set0(stf)+set0(lov)+set0(lmv)+set0(lng))
-bern_kt2012 <- bern_kt %>% filter(stj==2012) %>% group_by(hh_id) %>% select(tot_erwerbeink,tot_vermeink,primeink,primeink2,tot_transeink,TOTEINK,Satzeink_BUND,hh_id)
+
+#ALT aber evtl noch mal nüzlich nach der aufbereitung: bern_kt2012 <- bern_kt %>% filter(stj==2012) %>% group_by(hh_id) %>% select(tot_erwerbeink,tot_vermeink,primeink,primeink2,tot_transeink,TOTEINK,Satzeink_BUND,hh_id)
+
 #summarise(bern_kt2012, n=n()) %>% arrange(desc(n))
@@
 # dort machen wir einen schnitt bei den kantonalen daten um wohnheime etc auszuschlie?en (wie es HABE auch gemacht hat)
-bern_kt2012_plain <- bern_kt2012  %>% filter(n()<=7) #5% werden gefiltert
+#alt: bern_kt2012_plain <- bern_kt2012  %>% filter(n()<=7) #5% werden gefiltert
+bern_kt2012_plain <- bern_kt  %>% group_by(hh_id) %>% filter(n()<=7) #5% werden gefiltert #neu
 #bern_kt2012 <- bern_kt2012_plain %>% mutate(TOTEINK_hh=sum(TOTEINK)) %>% filter(!duplicated(hh_id))
-bern_kt2012 <- bern_kt2012_plain %>% mutate(primeink_hh=sum(primeink),primeink2_hh=sum(primeink2),Satzeink_BUND_hh=sum(Satzeink_BUND)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2.
-bis x.
+bern_kt2012 <- bern_kt2012_plain %>% mutate(primeink2_hh=sum(primeink2),eink_satz_dbs_hh=sum(eink_satz_dbs)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2. bis x.
 
@@
 round(gini_habe-gini_tax,3)
-# 0.457 0.519 diff=-0.061
+# gini_tax-gini_habe diff=-0.069
 
@@
 # export nach stata (ungewichtet)
-bern_plain_hh <- rbind(data.frame("income" = bern_kt2012_plain$Satzeink_BUND/1000,"group"=0,"weight"=1),data.frame("income" = bern_kt2012$Satzeink_BUND_hh/1000,"group"=1,"weight"=1))
+bern_plain_hh <- rbind(data.frame("income" = bern_kt2012_plain$eink_satz_dbs/1000,"group"=0,"weight"=1),data.frame("income" = bern_kt2012$eink_satz_dbs_hh/1000,"group"=1,"weight"=1))
 bern_plain_hh$group <- factor(bern_plain_hh$group,labels=c("tax units","households"))
@@
 round(gini_taxunits-gini_hh,3)
-
+#diff: 0.059
 @

### 2014-12-21 paper/subdocuments/results.Rnw
überarbeiten der conclusion

@@
 max(filter(df,kanton=="CH") %>% mutate(d_with_without=abs(G_steink-G_steink0)) %>% select(starts_with("d_")),na.rm=TRUE)
+max(filter(df,kanton=="CH") %>% mutate(d_with_without=abs(G_steink-G_steink_halb)) %>% select(starts_with("d_")),na.rm=TRUE)
 # 0.18

### 2014-12-23 paper/subdocuments/results.Rnw
1995 wieder eingefügt in fig2 und fig4

@@
   filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
-bd<-subset(bd,Jahr<1995 | Jahr >2002)
+bd<-subset(bd,(Jahr<=1995 | Jahr >2002)&Veranlagungsperiode!=1995)
 # 2011 zufügen
@@
   filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
-bd.0<-subset(bd.0,Jahr<1993 | Jahr >2002)
+bd.0<-subset(bd.0,(Jahr<=1995 | Jahr >2002)&Veranlagungsperiode!=1995)
 # 2011 zufügen
@@
 levels(dflong$Type)<-c("without non-taxed","including non-taxed","imputed values for non-taxed")
-dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<2003&dflong$Type=="including non-taxed"]<-NA
-dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<2003&dflong$Type=="imputed values for non-taxed"]<-NA
+dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<1995&dflong$Type=="including non-taxed"]<-NA
+dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<1995&dflong$Type=="imputed values for non-taxed"]<-NA
 fig_zeros <- ggplot(dflong[dflong$kanton=="CH",], 
@@
 # export nach stata (ungewichtet)
-bern_plain_hh <- rbind(data.frame("income" = bern_kt2012_plain$eink_satz_dbs/1000,"group"=0,"weight"=1),data.frame("income" = bern_kt2012$eink_satz_dbs_hh/1000,"group"=1,"weight"=1))
+bern_kt2012_plain7 <- bern_kt  %>% group_by(hh_id)
+bern_kt2012_7 <- bern_kt2012_plain7 %>% mutate(primeink2_hh=sum(primeink2),eink_satz_dbs_hh=sum(eink_satz_dbs)) %>% filter(!duplicated(hh_id)) #nimmt die erste unique beobachtung und droppt 2. bis x.
+
+bern_plain_hh <- rbind(data.frame("income" = bern_kt2012_plain7$eink_satz_dbs/1000,"group"=0,"weight"=1),data.frame("income" = bern_kt2012_7$eink_satz_dbs_hh/1000,"group"=1,"weight"=1))
 bern_plain_hh$group <- factor(bern_plain_hh$group,labels=c("tax units","households"))
@@
 round(gini_taxunits-gini_hh,3)
-#diff: 0.059
+#diff: 0.058
 @

### 2014-12-23 paper/subdocuments/results.Rnw
93er punkt raus und names() zeile vertauscht

@@
   filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
-bd.0<-subset(bd.0,(Jahr<=1995 | Jahr >2002)&Veranlagungsperiode!=1995)
+bd.0<-subset(bd.0,(Jahr==1995|Jahr>=2003)&Veranlagungsperiode!=1995)
 # 2011 zufügen
@@
 estv_masse$id<-1
+names(estv_masse)<-c("Year","Gini","Atkinson","Theil","id")
 estv_masse<-estv_masse[as.numeric(estv_masse$Year)>4,]
-
-names(estv_masse)<-c("Year","Gini","Atkinson","Theil","id")
 try(detach("package:reshape", unload=TRUE))

### 2015-01-09 paper/subdocuments/results.Rnw
1993 rein, 1995 raus.

@@
 library(reshape)
+df <- df %>% filter(kanton=="CH"&(bemessungsjahre<=1993.5|bemessungsjahre>=2003))
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
@@
   filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
-bd<-subset(bd,(Jahr<=1995 | Jahr >2002)&Veranlagungsperiode!=1995)
+bd<-subset(bd,(Jahr<=1993 | Jahr >2002)&Veranlagungsperiode!=1995)
 # 2011 zufügen
@@
   filter(Einheit=="Total") %>% select(Veranlagungsperiode,Jahr,gini_steink,gini_steinka)
-bd.0<-subset(bd.0,(Jahr==1995|Jahr>=2003)&Veranlagungsperiode!=1995)
+bd.0<-subset(bd.0,(Jahr<=1993|Jahr>=2003)&Veranlagungsperiode!=1995)
 # 2011 zufügen
@@
 levels(dflong$Type)<-c("without non-taxed","including non-taxed","imputed values for non-taxed")
-dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<1995&dflong$Type=="including non-taxed"]<-NA
-dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre<1995&dflong$Type=="imputed values for non-taxed"]<-NA
-fig_zeros <- ggplot(dflong[dflong$kanton=="CH",], 
+#dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre!=1993.5&dflong$bemessungsjahre<2003&dflong$Type=="including non-taxed"]<-NA
+#dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre!=1993.5&dflong$bemessungsjahre<2003&dflong$Type=="imputed values for non-taxed"]<-NA
+fig_zeros <- ggplot(filter(dflong,kanton=="CH"&
+  (((bemessungsjahre<=1993.5|bemessungsjahre>=2003)&Type=="without non-taxed")|
+  ((bemessungsjahre==1993.5|bemessungsjahre>=2003)&Type!="without non-taxed"))), 
        aes(x=bemessungsjahre,y=Gini,shape=Type,linetype=Type))+
@@
          eink_art=="Nach steuerbarem Einkommen",
+         bemessungsjahre<=1993.5|bemessungsjahre>=2003,
          fall=="Normalfälle",
@@
 
-estv_masse_long$measure <- factor(estv_masse_long$variable)
+estv_masse_long$Measure <- factor(estv_masse_long$variable)
 estv_masse_long$Year <- as.double(as.character(estv_masse_long$Year))
-estv_masse_long <- estv_masse_long %>% group_by(measure) %>% mutate(log_value=log(value)-log(value)[1])
-fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log_value,shape=measure))+
+estv_masse_long <- estv_masse_long %>% group_by(Measure) %>% mutate(log_value=log(value)-log(value)[1])
+fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log_value,shape=Measure))+
   scale_y_continuous(limits=c(-0.7,0.45))+
   scale_x_continuous(limits=c(1945,2011),breaks=number_ticks(10)) +
-  geom_point(aes(shape=measure),size=3)+
-  geom_line(aes(linetype=measure))+
+  geom_point(aes(shape=Measure),size=3)+
+  geom_line(aes(linetype=Measure))+
   theme_bw()+

### 2015-01-09 paper/subdocuments/results.Rnw
mini ding vergessen

@@
          fall=="Normalfälle",
-         anz_pflichtige>0) %>%
+         anz_pflichtige>0,
+         bemessungsjahre<=1993.5|bemessungsjahre>=2003) %>%
         # steuerperiode!=1945.5) %>% 

### 2015-01-12 paper/subdocuments/results.Rnw
updated section special cases plus minor corrections

@@
 #try(bern_kt <- read.dta("H:/Assessing Inequalty/taxDataBE2012.dta"))
-#try(bern_kt<-read.dta("U:/Persönlich/Forschung/Uni Bern/taxDataBE2012.dta"))
+try(bern_kt<-read.csv("U:/Persönlich/Forschung/Uni Bern/BFH-Grunddaten 2012 Output V1.1.csv",sep=";"))
 try(bern_kt <- read.csv("G:/stick/bern//BFH-Grunddaten 2012 Output V1.1.csv",sep=";"))

### 2015-01-13 paper/subdocuments/results.Rnw
minor corrections results.Rnw

@@
 
+# Compare ginis for married among tax units and for HBS
+round(gini(habe11.married$psteink,w=habe11.married$gew)-gini(bd.2011.ind),3)
+# gini_tax-gini_habe diff=-0.066
+
 habe_bd_2011_married <- rbind(data.frame(cbind("income" = habe11.married$inc_habe,"group"=0,"weight"=habe11.married$gew)),data.frame("income" = bd.2011.ind,"group"=1,"weight"=1))
@@
 round(gini_habe-gini_tax,3)
-# gini_tax-gini_habe diff=-0.069
+# gini_tax-gini_habe diff=-0.066
 

### 2015-04-14 paper/subdocuments/results.Rnw
grafiken wieder einzeln, legendensortierung, neue grafik bern_ginis und benplot74

@@
 
+<<graphic_options>>=
+## Graphic Options ----------------------------------------
+pointsize = 3
+linesize = 0.7
+theme_set(theme_bw(base_size = 16)+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+		)
+@
+
 <<different_ginis, fig.cap='Gini coefficients for different income measures',fig.pos='H'>>=
@@
 levels(dflong$Income)<-c("taxable income","net income","taxable income after federal tax")
+dflong$Income <- factor(dflong$Income,levels=c("taxable income","taxable income after federal tax","net income"))
 fig_incomes <- ggplot(dflong[dflong$kanton=="CH",], aes(x=bemessungsjahre,y=Gini,shape=Income))+
@@
   scale_x_continuous(limits=c(1945,2011),breaks=number_ticks(10)) +
-  geom_point(aes(shape=Income),size=3)+
-  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Income))+
+  geom_point(aes(shape=Income),size=pointsize)+
+  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Income),size=linesize)+
   theme_bw()+
@@
 names(bdlong)[3]<-"Income"
+bdlong$Income <- factor(bdlong$Income,levels=c("equivalised taxable income","taxable income","equivalised taxable income (only taxed)","taxable income (only taxed)"))
 @
@@
   scale_x_continuous(limits=c(1970,2011),breaks=number_ticks(10)) +
-  geom_point(aes(shape=Income),size=3)+
-  geom_line(data=bdlong,aes(linetype=Income))+
+  geom_point(aes(shape=Income),size=pointsize)+
+  geom_line(data=bdlong,aes(linetype=Income),size=linesize)+
   theme_bw()+
@@
 levels(dflong$Type)<-c("without non-taxed","including non-taxed","imputed values for non-taxed")
+dflong$Type <- factor(dflong$Type, levels=c("including non-taxed","imputed values for non-taxed","without non-taxed"))
 #dflong$Gini[dflong$kanton=="CH"&dflong$bemessungsjahre!=1993.5&dflong$bemessungsjahre<2003&dflong$Type=="including non-taxed"]<-NA
@@
   scale_x_continuous(limits=c(1970,2011),breaks=number_ticks(10)) +
-  geom_line(size=0.5)+
-  geom_point(aes(shape=Type),size=3)+
+  geom_line(size=linesize)+
+  geom_point(aes(shape=Type),size=pointsize)+
   theme_bw()+
@@
   filter(Familientyp08>130,Jahr08==2010)
+
+# Wir brauchen (a) verfügbares Einkommen (b) Bruttoeinkommen und c(Bruttoeinkommen nach Soziversicherungsbeiträgen(A31m) und Monetäre Trasnferausgaben an andere HaushalteA35m)
 habe10.married$psteink<-habe10.married$Bruttoeinkommen08+habe10.married$A31m+habe10.married$A35m
-# Wir brauchen (a) verfügbares Einkommen (b) Bruttoeinkommen und c(Bruttoeinkommen nach Soziversicherungsbeiträgen(A31m) und Monetäre Trasnferausgaben an andere HaushalteA35m)
+
+# Dann noch eine extra Version wo die Sozialhilfe abgezogen wird
+habe10.married$psteink_ohne_soz<-habe10.married$Bruttoeinkommen08+habe10.married$A31m+habe10.married$A35m- habe10.married$E23
+
+# data-select für "ohne soz"
+habe10.married_ohne_soz<-data.frame(habe10.married$VerfuegbaresEinkommen08*12/1000,habe10.married$Bruttoeinkommen08*12/1000,habe10.married$psteink_ohne_soz*12/1000,habe10.married$jahresgewicht)
+names(habe10.married_ohne_soz)<-c("inc_habe","brutto","psteink","gew")
+habe10.married_ohne_soz$inc_habe[habe10.married_ohne_soz$inc_habe<0]<-0 # One values is -20'000
+
+# data-select für "mit soz"
 habe10.married<-data.frame(habe10.married$VerfuegbaresEinkommen08*12/1000,habe10.married$Bruttoeinkommen08*12/1000,habe10.married$psteink*12/1000,habe10.married$jahresgewicht)
@@
 # 2011
+
+# Wir brauchen (a) verfügbares Einkommen (b) Bruttoeinkommen und c(Bruttoeinkommen nach Soziversicherungsbeiträgen(A31m) und Monetäre Trasnferausgaben an andere HaushalteA35m)
 habe11.married<-habe0911 %>% 
@@
   mutate(psteink=Bruttoeinkommen08+A31m+A35m)
-# Wir brauchen (a) verfügbares Einkommen (b) Bruttoeinkommen und c(Bruttoeinkommen nach Soziversicherungsbeiträgen(A31m) und Monetäre Trasnferausgaben an andere HaushalteA35m)
+
+# und nochmal ohne Sozialhilfe
+habe11.married_ohne_soz<-habe0911 %>% 
+  filter(Familientyp08>130,Jahr08==2011) %>% 
+  mutate(psteink=Bruttoeinkommen08+A31m+A35m-E23)
+
 habe11.married<-data.frame(habe11.married$VerfuegbaresEinkommen08*12/1000,habe11.married$Bruttoeinkommen08*12/1000,habe11.married$psteink*12/1000,habe11.married$jahresgewicht)
@@
 habe11.married$inc_habe[habe11.married$inc_habe<0]<-0 # One values is -500
+
+# und nochmal für ohne_soz
+habe11.married_ohne_soz<-data.frame(habe11.married_ohne_soz$VerfuegbaresEinkommen08*12/1000,habe11.married_ohne_soz$Bruttoeinkommen08*12/1000,habe11.married_ohne_soz$psteink*12/1000,habe11.married_ohne_soz$jahresgewicht)
+names(habe11.married_ohne_soz)<-c("inc_habe","brutto","psteink","gew")
+habe11.married_ohne_soz$inc_habe[habe11.married_ohne_soz$inc_habe<0]<-0 # One values is -500
 
@@
 write.dta(habe_bd_2011_married, file="../../data/stata_data/habe_bd_2011_married.dta")
+
+# und nochmal ohne Sozialhilfe (Vorsicht: hier nehmen wir psteink, vorher verf. einkommen (inc_habe))
+habe_bd_2011_married_ohne_soz <- rbind(data.frame(cbind("income" = habe11.married_ohne_soz$psteink,"group"=0,"weight"=habe11.married_ohne_soz$gew)),data.frame("income" = bd.2011.ind,"group"=1,"weight"=1))
+habe_bd_2011_married_ohne_soz$group <- factor(habe_bd_2011_married_ohne_soz$group,labels=c("habe","bd"))
+write.dta(habe_bd_2011_married_ohne_soz, file="../../data/stata_data/habe_bd_2011_married_ohne_soz.dta")
 
@@
 
+<<bern_kantonale_analysen>>=
+# core variables from bern 2002-2012
+bern <- read.dta("G:/stick/bern/mini_taxdata_BE_all.dta")
+bern_gini <- bern %>%
+  mutate(stbeink_no_dbs = stbeink_BUND - eink_st_dbs,
+         stbeink_no_tax = stbeink_BUND - steuern_eink) %>% # nur steuern auf einkommen (nicht vermögen) werden abgezogen
+  group_by(stj) %>%
+  summarise(G_TOTEINK = gini(TOTEINK),
+            G_stbeink_BUND = gini(stbeink_BUND),
+            G_stbeink_no_dbs = gini(stbeink_no_dbs),
+            G_stbeink_no_tax = gini(stbeink_no_tax))
+
+library(reshape2)
+
+# wide to long
+bern_gini_long <- melt(bern_gini,id.vars="stj",variable.name="Income",value.name="Gini")
+
+# legend order
+levels(bern_gini_long$Income) <- c("total income", "taxable income", "taxable income after federal income tax", "taxable income after all income taxes")
+bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("taxable income", "taxable income after federal income tax", "taxable income after all income taxes","total income"))
+
+fig_bern_gini <- ggplot(bern_gini_long, aes(x=stj,y=Gini)) +
+  #scale_y_continuous(limits=c(0.28,0.55))+
+  scale_x_continuous(breaks=number_ticks(5)) +
+  geom_point(aes(shape=Income),size=pointsize)+
+  geom_line(aes(linetype=Income),size=linesize)+
+  theme_bw()+
+  xlab("Year")+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
+  ggtitle("Gini coefficients of different income measures")
+fig_bern_gini
+
+pdf("../figure/fig_bern_gini.pdf",width=10.5,height=6)
+fig_bern_gini
+dev.off()
+@
+
 From the discussion in the data section we would expect differences between the income distributions from survey and tax data.
@@
 
+all_gini_long$Data <- factor(all_gini_long$Data,levels=c("FTA","LIS","EU.SILC","HBS"))
 benplot <- ggplot(data=na.omit(all_gini_long),
@@
   scale_x_continuous(limits=c(1945,2012),breaks=number_ticks(10)) +
-  geom_point(size=3)+
-  geom_line()+
+  geom_point(size=pointsize)+
+  geom_line(size=linesize)+
   #geom_text(aes(label=annotation,y=ypos),color="#555555",size=3)+
@@
 
+benplot74 <- ggplot(data=na.omit(all_gini_long),
+  aes(y=Gini,x=Year,shape=Data,linetype=Data))+
+  ylab("Gini from different sources") +
+  scale_y_continuous(limits=c(0.18,0.45)) +
+  scale_x_continuous(limits=c(1973,2012),breaks=number_ticks(10)) +
+  geom_point(size=pointsize)+
+  geom_line(size=linesize)+
+  theme_bw()+
+  theme(legend.justification=c(0,1),legend.position=c(0, 1))
+
 png("../figure/ben_plot.png",width=800,height=450)
 benplot
+dev.off()
+
+pdf("../figure/ben_plot.pdf",width=10.5,height=6)
+benplot
+dev.off()
+
+pdf("../figure/ben_plot74.pdf",width=10.5,height=6)
+benplot74
 dev.off()
@@
 estv_masse_long <- estv_masse_long %>% group_by(Measure) %>% mutate(log_value=log(value)-log(value)[1])
+
 fig_measures <- ggplot(estv_masse_long, aes(x=Year,y=log_value,shape=Measure))+
@@
 dev.off()
+
+#output als einzelne pdfs
+pdf("../figure/fig_incomes.pdf",width=10.5,height=6)
+fig_incomes
+dev.off()
+pdf("../figure/fig_equiv.pdf",width=10.5,height=6)
+fig_equiv
+dev.off()
+pdf("../figure/fig_zeros.pdf",width=10.5,height=6)
+fig_zeros
+dev.off()
+pdf("../figure/fig_measures.pdf",width=10.5,height=6)
+fig_measures
+dev.off()
 @

### 2015-04-20 paper/subdocuments/results.Rnw
update latest version of paper updated sissGini Data with HBS data several updates of results.Rnw

@@
 #bern kantonal (liegt lokal wegen datenschutz)
-#try(bern_kt <- read.dta("H:/Assessing Inequalty/taxDataBE2012.dta"))
-try(bern_kt<-read.csv("U:/Persönlich/Forschung/Uni Bern/BFH-Grunddaten 2012 Output V1.1.csv",sep=";"))
+try(bern_kt<-read.dta("U:/Persönlich/Forschung/Uni Bern/taxdata_BE_12.dta"))
 try(bern_kt <- read.csv("G:/stick/bern//BFH-Grunddaten 2012 Output V1.1.csv",sep=";"))
@@
 bern_kt <- filter(bern_kt,VON=="2012-01-01",BIS=="2012-12-31")
+bern_kt <- filter(bern_kt,Satzeink_BUND<=TOTEINK,erm==0)
+bern_kt <- filter(bern_kt,erm==0)
 names(bern_kt)<-tolower(names(bern_kt))
@@
 # core variables from bern 2002-2012
-bern <- read.dta("G:/stick/bern/mini_taxdata_BE_all.dta")
+try(bern <- read.dta("G:/stick/bern/mini_taxdata_BE_all.dta"))
+try(bern<-read.dta("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/Kantonale Steuerdaten/Bern/taxdata_BE_reduced.dta"))
+P:\WGS\FBS\ISS\Projekte laufend\SNF Ungleichheit\Datengrundlagen\Kantonale Steuerdaten\Bern
+bern <- filter(bern_kt,Satzeink_BUND<=TOTEINK,erm==0)
+bern <- filter(bern_kt,erm==0)
+bern$verfeink<-bern$TOTEINK+bern$privattrans_bez-bern$steuertotal
+
+#bern_gini <- bern %>%
+#  mutate(stbeink_no_dbs = stbeink_BUND - eink_st_dbs,
+#         stbeink_no_tax = stbeink_BUND - steuern_eink) %>% # nur steuern auf einkommen (nicht vermögen) werden abgezogen
+#  group_by(stj) %>%
+#  summarise(G_TOTEINK = gini(TOTEINK),
+#            G_stbeink_BUND = gini(stbeink_BUND),
+#            G_stbeink_no_dbs = gini(stbeink_no_dbs),
+#            G_stbeink_no_tax = gini(stbeink_no_tax))
+
 bern_gini <- bern %>%
-  mutate(stbeink_no_dbs = stbeink_BUND - eink_st_dbs,
-         stbeink_no_tax = stbeink_BUND - steuern_eink) %>% # nur steuern auf einkommen (nicht vermögen) werden abgezogen
   group_by(stj) %>%
     summarise(G_TOTEINK = gini(TOTEINK),
-            G_stbeink_BUND = gini(stbeink_BUND),
-            G_stbeink_no_dbs = gini(stbeink_no_dbs),
-            G_stbeink_no_tax = gini(stbeink_no_tax))
+            G_satzeink_BUND = gini(Satzeink_KANTON),
+            G_TOTEINK = gini(TOTEINK),
+            G_verfeink = gini(verfeink))
 
@@
 # legend order
-levels(bern_gini_long$Income) <- c("total income", "taxable income", "taxable income after federal income tax", "taxable income after all income taxes")
-bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("taxable income", "taxable income after federal income tax", "taxable income after all income taxes","total income"))
+levels(bern_gini_long$Income) <- c("total income", "taxable income", "Disposable income")
+bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("total income", "taxable income", "Disposable income")
 
@@
   theme(legend.justification=c(0,1),legend.position=c(0, 1))+
-  ggtitle("Gini coefficients of different income measures")
+  ggtitle("Gini coefficients of different income concepts")
 fig_bern_gini
@@
 number_ticks <- function(n) {function(limits) pretty(limits, n)}
-swissGini<-read.csv("../../data/swissGini.csv",header=TRUE,sep=";")
+try(swissGini<-read.csv("../../data/swissGini.csv",header=TRUE,sep=";"))
+try(swissGini<-read.csv("C:/Users/hlo1/swiss_inequality_development/data/swissGini.csv",header=TRUE,sep=";"))
 
@@
 ### and for the rest of Ginis:
-estv <- read.dta("../../data/steuerdaten20141126_stata12.dta")
+try(estv <- read.dta("../../data/steuerdaten20141126_stata12.dta"))
+try(estv <- read.dta("C:/Users/hlo1/swiss_inequality_development/data/steuerdaten20141126_stata12.dta"))
 
@@
 all_gini_long$Data <- factor(all_gini_long$Data,levels=c("FTA","LIS","EU.SILC","HBS"))
+
 benplot <- ggplot(data=na.omit(all_gini_long),
   aes(y=Gini,x=Year,shape=Data,linetype=Data))+
-  ylab("Gini from different sources") +
+  ylab("Gini") +
   scale_y_continuous(limits=c(0.18,0.45)) +
   scale_x_continuous(limits=c(1945,2012),breaks=number_ticks(10)) +
-  geom_point(size=pointsize)+
-  geom_line(size=linesize)+
+  geom_point(size=4)+
+  geom_line(size=1)+
   #geom_text(aes(label=annotation,y=ypos),color="#555555",size=3)+

### 2015-06-05 paper/subdocuments/results.Rnw
ginis und grafiken überarbeitet mit erm!=0 und satzeink<=toteink.

@@
 try(bern_kt<-read.dta("U:/Persönlich/Forschung/Uni Bern/taxdata_BE_12.dta"))
-try(bern_kt <- read.csv("G:/stick/bern//BFH-Grunddaten 2012 Output V1.1.csv",sep=";"))
+try(bern_kt <- read.dta("G:/stick/bern/taxdata_BE_12.dta"))
 #> nrow(bern_kt)
 #[1] 622368
-nrow(filter(bern_kt,VON=="2012-01-01",BIS=="2012-12-31"))
+nrow(filter(bern_kt,von=="2012-01-01",bis=="2012-12-31"))
 #[1] 605975 wir filtern ca. 3% unterjährige
-bern_kt <- filter(bern_kt,VON=="2012-01-01",BIS=="2012-12-31")
+bern_kt <- filter(bern_kt,von=="2012-01-01",bis=="2012-12-31")
 bern_kt <- filter(bern_kt,Satzeink_BUND<=TOTEINK,erm==0)
 bern_kt <- filter(bern_kt,erm==0)
-names(bern_kt)<-tolower(names(bern_kt))
+#names(bern_kt)<-tolower(names(bern_kt))
 set0 <- function(x) ifelse(x<0 | is.na(x),0,x)
@@
 P:\WGS\FBS\ISS\Projekte laufend\SNF Ungleichheit\Datengrundlagen\Kantonale Steuerdaten\Bern
-bern <- filter(bern_kt,Satzeink_BUND<=TOTEINK,erm==0)
-bern <- filter(bern_kt,erm==0)
+bern <- filter(bern,Satzeink_BUND<=TOTEINK,erm==0)
+bern <- filter(bern,erm==0)
 bern$verfeink<-bern$TOTEINK+bern$privattrans_bez-bern$steuertotal
@@
     summarise(G_TOTEINK = gini(TOTEINK),
-            G_satzeink_BUND = gini(Satzeink_KANTON),
+            G_satzeink_KANTON = gini(Satzeink_KANTON),
             G_TOTEINK = gini(TOTEINK),
@@
 
+detach("package:reshape", unload=TRUE) #das ist nötig falls man allen code vorher verwendet hat, da reshape dann geladen ist und reshape und reshape2 gleichnamige funktionen haben die sie intern aufrufen.
 library(reshape2)
@@
 # legend order
-levels(bern_gini_long$Income) <- c("total income", "taxable income", "Disposable income")
-bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("total income", "taxable income", "Disposable income")
+levels(bern_gini_long$Income) <- c("total income", "taxable income", "disposable income")
+bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("total income", "taxable income", "disposable income"))
+bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("taxable income","total income","disposable income"))
 
@@
   xlab("Year")+
-  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
+  theme(legend.justification=c(0,1),legend.position=c(0, 0.75))+
   ggtitle("Gini coefficients of different income concepts")

### 2015-06-16 paper/subdocuments/results.Rnw
tiny changes in the results part...

--- 2015-04-20 paper/subdocuments/results.Rnw
@@
 bern_kt <- filter(bern_kt,VON=="2012-01-01",BIS=="2012-12-31")
-bern_kt <- filter(bern_kt,Satzeink_BUND<=TOTEINK,erm==0)
+bern_kt <- filter(bern_kt,Satzeink_BUND<=TOTEINK)
 bern_kt <- filter(bern_kt,erm==0)
@@
 try(bern<-read.dta("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/Kantonale Steuerdaten/Bern/taxdata_BE_reduced.dta"))
-P:\WGS\FBS\ISS\Projekte laufend\SNF Ungleichheit\Datengrundlagen\Kantonale Steuerdaten\Bern
-bern <- filter(bern_kt,Satzeink_BUND<=TOTEINK,erm==0)
-bern <- filter(bern_kt,erm==0)
+
 bern$verfeink<-bern$TOTEINK+bern$privattrans_bez-bern$steuertotal
@@
 bern_gini_long <- melt(bern_gini,id.vars="stj",variable.name="Income",value.name="Gini")
-
+names(bern_gini_long)<-c("stj","Income","Gini")
 # legend order
 levels(bern_gini_long$Income) <- c("total income", "taxable income", "Disposable income")
-bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("total income", "taxable income", "Disposable income")
+bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("total income", "taxable income", "Disposable income"))
 
 fig_bern_gini <- ggplot(bern_gini_long, aes(x=stj,y=Gini)) +
-  #scale_y_continuous(limits=c(0.28,0.55))+
+  scale_y_continuous(limits=c(0.28,0.55))+
   scale_x_continuous(breaks=number_ticks(5)) +
-  geom_point(aes(shape=Income),size=pointsize)+
-  geom_line(aes(linetype=Income),size=linesize)+
+  geom_point(aes(shape=Income),size=5)+
+  geom_line(aes(linetype=Income),size=1)+
   theme_bw()+
   xlab("Year")+
-  theme(legend.justification=c(0,1),legend.position=c(0, 1))+
+  theme(legend.justification=c(0,1),legend.position=c(0, 0.3))+
   ggtitle("Gini coefficients of different income concepts")

### 2015-06-17 paper/subdocuments/results.Rnw
merge und bern grafik

--- 2015-06-05 paper/subdocuments/results.Rnw
@@
 #[1] 605975 wir filtern ca. 3% unterjährige
-bern_kt <- filter(bern_kt,von=="2012-01-01",bis=="2012-12-31")
-bern_kt <- filter(bern_kt,Satzeink_BUND<=TOTEINK,erm==0)
-bern_kt <- filter(bern_kt,erm==0)
+
+try(bern_kt <- filter(bern_kt,VON=="2012-01-01",BIS=="2012-12-31",Satzeink_BUND<=TOTEINK,erm==0))
+try(bern_kt <- filter(bern_kt,von=="2012-01-01",bis=="2012-12-31",Satzeink_BUND<=TOTEINK,erm==0))
+
 #names(bern_kt)<-tolower(names(bern_kt))
@@
 try(bern<-read.dta("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/Kantonale Steuerdaten/Bern/taxdata_BE_reduced.dta"))
-P:\WGS\FBS\ISS\Projekte laufend\SNF Ungleichheit\Datengrundlagen\Kantonale Steuerdaten\Bern
+
 bern <- filter(bern,Satzeink_BUND<=TOTEINK,erm==0)
-bern <- filter(bern,erm==0)
 bern$verfeink<-bern$TOTEINK+bern$privattrans_bez-bern$steuertotal
@@
 bern_gini_long <- melt(bern_gini,id.vars="stj",variable.name="Income",value.name="Gini")
-
+names(bern_gini_long)<-c("stj","Income","Gini")
 # legend order
@@
 fig_bern_gini <- ggplot(bern_gini_long, aes(x=stj,y=Gini)) +
-  #scale_y_continuous(limits=c(0.28,0.55))+
+  scale_y_continuous(limits=c(0.28,0.6))+
   scale_x_continuous(breaks=number_ticks(5)) +
-  geom_point(aes(shape=Income),size=pointsize)+
-  geom_line(aes(linetype=Income),size=linesize)+
+  geom_point(aes(shape=Income),size=4)+
+  geom_line(aes(linetype=Income),size=1)+
   theme_bw()+
   xlab("Year")+
-  theme(legend.justification=c(0,1),legend.position=c(0, 0.75))+
+  theme(legend.justification=c(0,1),legend.position=c(0, 0.3))+
   ggtitle("Gini coefficients of different income concepts")

### 2015-06-17 paper/subdocuments/results.Rnw
update Paper and results.Rnw

@@
 #bern kantonal (liegt lokal wegen datenschutz)
-try(bern_kt<-read.dta("U:/Persönlich/Forschung/Uni Bern/taxdata_BE_12.dta"))
+try(bern_kt<-read.dta("U:/Persönlich/Forschung/Uni Bern/Steuerdaten_BE_2002_2012.dta"))
 try(bern_kt <- read.dta("G:/stick/bern/taxdata_BE_12.dta"))
@@
 
-bern <- filter(bern,Satzeink_BUND<=TOTEINK,erm==0)
+bern <- filter(bern,Satzeink_KANTON<=TOTEINK,erm==0)
 bern$verfeink<-bern$TOTEINK+bern$privattrans_bez-bern$steuertotal
@@
 fig_bern_gini <- ggplot(bern_gini_long, aes(x=stj,y=Gini)) +
-  scale_y_continuous(limits=c(0.28,0.6))+
+  scale_y_continuous(limits=c(0.28,0.55))+
   scale_x_continuous(breaks=number_ticks(5)) +
-  geom_point(aes(shape=Income),size=4)+
+  geom_point(aes(shape=Income),size=3)+
   geom_line(aes(linetype=Income),size=1)+

### 2015-06-17 paper/subdocuments/results.Rnw

@@
 bern <- filter(bern,Satzeink_KANTON<=TOTEINK,erm==0)
+bern <- filter(bern,Satzeink_BUND<=TOTEINK,erm==0)
 bern$verfeink<-bern$TOTEINK+bern$privattrans_bez-bern$steuertotal
@@
 fig_bern_gini <- ggplot(bern_gini_long, aes(x=stj,y=Gini)) +
-  scale_y_continuous(limits=c(0.28,0.55))+
+  scale_y_continuous(limits=c(0.28,0.6))+
   scale_x_continuous(breaks=number_ticks(5)) +

### 2015-11-18 paper/subdocuments/results.Rnw
cleaned results.Rnw
figures

@@
 names(dflong)[4]<-"Gini"
-levels(dflong$Income)<-c("taxable income","net income","taxable income after federal tax")
-dflong$Income <- factor(dflong$Income,levels=c("taxable income","taxable income after federal tax","net income"))
+levels(dflong$Income)<-c("taxable income (total income - all deductions)","net income (total income - non-social deductions)","taxable income after federal tax")
+dflong$Income <- factor(dflong$Income,levels=c("taxable income (total income - all deductions)","taxable income after federal tax","net income (total income - non-social deductions)"))
 fig_incomes <- ggplot(dflong[dflong$kanton=="CH",], aes(x=bemessungsjahre,y=Gini,shape=Income))+
@@
   ((bemessungsjahre==1993.5|bemessungsjahre>=2003)&Type!="without non-taxed"))), 
-       aes(x=bemessungsjahre,y=Gini,shape=Type,linetype=Type))+
+       aes(x=bemessungsjahre,y=Gini,linetype=Type))+
   scale_y_continuous(limits=c(0.28,0.55))+
   scale_x_continuous(limits=c(1970,2011),breaks=number_ticks(10)) +
+  #scale_linetype_manual("Taxable income",values=c("including non-taxed"=1,"imputed values for non-taxed"=2,"without non-taxed"=3))+
+  #scale_shape_manual("Taxable income",values=c("including non-taxed"=1,"imputed values for non-taxed"=2,"without non-taxed"=3))+
   geom_line(size=linesize)+
-  geom_point(aes(shape=Type),size=pointsize)+
+  geom_point(aes(shape=Type), fill = "red",size=pointsize)+
   theme_bw()+
@@
   theme(legend.justification=c(0,1),legend.position=c(0, 1))+
-  ggtitle("Influence of non-taxed")
+  ggtitle("Influence of non-taxed")+
+  labs(shape="Taxable income")+
+  labs(linetype="Taxable income")
 fig_zeros
@@
 bern <- filter(bern,Satzeink_KANTON<=TOTEINK,erm==0)
-bern <- filter(bern,Satzeink_BUND<=TOTEINK,erm==0)
 bern$verfeink<-bern$TOTEINK+bern$privattrans_bez-bern$steuertotal
@@
 # legend order
-levels(bern_gini_long$Income) <- c("total income", "taxable income", "disposable income")
-bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("total income", "taxable income", "disposable income"))
-bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("taxable income","total income","disposable income"))
+levels(bern_gini_long$Income) <- c("total income (earnings, self-employment, capital income)", "taxable income (total income - all deductions)", "disposable income (total income - taxes: federal, cantonal, communal and churches)")
+bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("total income (earnings, self-employment, capital income)", "taxable income (total income - all deductions)", "disposable income (total income - taxes: federal, cantonal, communal and churches)"))
+bern_gini_long$Income <- factor(bern_gini_long$Income,levels=c("taxable income (total income - all deductions)","total income (earnings, self-employment, capital income)","disposable income (total income - taxes: federal, cantonal, communal and churches)"))
 
